### 2017-11-08 latex/ROADEF/article_roadef_2018.tex
added latex formulation draft with roadef format

% Exemple d'utilisation de la classe roadef2018 pour le congrès ROADEF 2018 (http://roadef2018.labsticc.fr)

\documentclass{roadef}

\begin{document}

% Le titre du papier
\title{Maintenance planning on french military aircraft operations}

% Le titre court
\def\shorttitle{Titre court}

% Les auteurs et leur numéro d'affiliation
\author{Franco Peschiera\inst{1}, Alain Haït\inst{1}, Olga Battaïa\inst{1} }

% Les affiliations (par ordre croissant des numéros d'affiliation) séparées par \and
\institute{
ISAE... \\
\email{\{franco.peschiera\}@isae-supaero.fr}
}

% Création de la page de titre
\maketitle
\thispagestyle{empty}

% Les mots-clés
\keywords{optimisation, planification, militaire, maintenance}

\section{Introduction}

The Flight and Maintenance Planning is a very well known and studied problem in the aviation industry. It

...

Military Flight and Maintenance Planning...

\section{MFMP}

    \subsection{Tasks}
    \label{def:task}

    There is a set of (XXX) tasks to do in a given planning horizon divided into (XXX) periods of the same length. Each task (XXX) has a given starting (XXX) an ending (XXX) time (the duration is measured in a certain number of periods) and needs to use a certain quantity of resources (XXX). Not all (XXX) resources can be used to accomplish any task: each task has specific specifications that a resource need to satisfy in order to be a candidate. In other words, there is a parameter (XXX) that indicates task-resource compatibility. A resource can be a candidate for more than one task.

    The assignment of a resource to a task is not made necessarily for the whole duration of the task. After a minimum amount of time (XXX), a resource can be liberated and exchanged for another one. The total number of resources being used at any given time in a specific task can never be less than the required XXX.

    \subsection{Resources}

    The resources require a recurrent maintenance to be in the correct state to accomplish such tasks correctly. These maintenances need to take place before a given amount of time (ElapsedTime XXX) has passed since the previous maintenance or before the resource has been used for a given amount of time (UsageTime XXX) since the previous maintenance, whichever happens first. A resource without the proper maintenance cannot be used to satisfy tasks and is consider worthless.

    Additionally, after an absolute amount of time and/or usage, the resource becomes obsolete. There is no way to reverse this process.

    Each resources starts the planning period with a specific status given by:

    \begin{itemize}
        \item UsageTime.
        \item ElapsedTime.
        \item UsageTimeAbs.
        \item ElapseTimeAbs.
        \item LastMaintenanceType.
    \end{itemize}

    Finally, resources are organized into families or groups. Each resource inside in a family or group shares the same types of maintenances and, usually, the same kind of tasks, among other information.

    \paragraph{Resource's states}
    \label{def:res-state}

    The following are the possible states for a resource in any given time:

    \begin{itemize}
        \item Assigned to a task (see \ref{def:task}).
        \item Under maintenance (see \ref{def:maint}).
        \item Under storage (see \ref{def:sto}).
        \item Waiting maintenance.
        \item Obsolete.
        \item Available (none of the above).
    \end{itemize}

    \subsection{Maintenances}
    \label{def:maint}

    These maintenances take an amount of time equal to (XXX) periods and cannot be interrupted. The state of the resource after the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (ElapsedTime XXX and UsageTime XXX).

    In other words: after a given maintenance XXX, a resource resets its ElapsedTime and UsageTime back to 0, so it can continue being assigned to tasks.

    The number of resources under maintenance in any given period is limited to a maximum capacity (XXX) depending on each period.

    Not all maintenances are the same for any given resource. Each resources follows a sequence of maintenances. In other words, the second maintenance of a given resource is different in nature from the first maintenance for that same resource. Examples of differences on consecutive maintenances are the duration it takes or whether it restores storage capacity (see \ref{def:sto}).

    \subsection{Storage}
    \label{def:sto}

    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (ElapsedTime). In order to avoid this problem, the resource can be put into a "storage" state.

    A resource in this states has to be kept for a minimum time (XXX). While in this state it cannot receive maintenance or be assigned any task.

    \subsection{Objectives}

    The are multiple objectives that need to be taken into account:

    \begin{enumerate}
        \item Minimize the maximum number of resources under maintenance in any given period.
        \item Maximize the minimum number of resources available (see \ref{def:res-state}) in any given period.
    \end{enumerate}

    Every resource has the capacity to be stored. This capacity (measured in number of periods XXX) is expended every time the resource is stored. In order for a resource to recover storage capacity, it needs to received a specific maintenance.

    % \begin{figure}[!ht]
    %     \begin{center}
    %         \includegraphics[height=2cm,clip=true]{roadef_logo.eps}
    %         \caption[Fig]{Logo de la ROADEF}
    %         \label{logoRoadef}
    %     \end{center}
    % \end{figure}

    % \begin{equation}
    % E=MC^2\label{emc}
    % \end{equation}

    % \begin{theoreme}
    %   Un exemple de théorème. Les environnements suivants sont également disponibles : remarque, propriété, corollaire, définition, notation, proposition, exemple, preuve.
    % \end{theoreme}

    % \subsection{Tableau}

    % Le titre du tableau devra être positionné sous le tableau (par exemple Tableau \ref{tableau}).

    % \begin{table}[!ht]
    %     \begin{center}
    %         \begin{tabular}{lrr}
    %             \hline
    %             & \multicolumn{1}{c}{Colonne 1} & \multicolumn{1}{c}{Colonne 2}\\
    %             \hline
    %             Ligne 1 & L1C1 & L1C2\\
    %             Ligne 2 & L2C1 & L2C2\\
    %             \hline
    %         \end{tabular}
    %         \caption{Exemple de tableau}
    %     \label{tableau}
    %     \end{center}
    % \end{table}

    % \subsection{Liste}

    % Voici une liste :

    % \begin{itemize}
    % \item remarque
    % \item propriété
    % \end{itemize}

\section{Model}

    \begin{equation}
    E=MC^2\label{emc}
    \end{equation}

\section{Conclusions et perspectives}

% La bibliographie

\bibliographystyle{plain}

% Version "on-line" de la bibliographie, mais il est
% également possible d'utiliser un fichier ".bib" et d'utiliser BibTeX

\begin{thebibliography}{2}
\bibitem{toth02}
Paolo Toth and Daniele Vigo.
\newblock \emph{The Vehicle Routing Problem}.
\newblock Monographs on Discrete Mathematics and Applications. Society for Industrial and Applied Mathematics, 2002.
\bibitem{kirkpatrick83}
Scott Kirkpatrick, C~Daniel Gelatt, and Mario~P Vecchi.
\newblock Optimization by simmulated annealing.
\newblock \emph{science}, 220\penalty0 (4598):\penalty0 671--680, 1983.

\end{thebibliography}

\end{document}

### 2017-11-09 latex/ROADEF/article_roadef_2018.tex
I update the problem description with more details and letters

@@
 % Les mots-clés
-\keywords{optimisation, planification, militaire, maintenance}
+\keywords{optimization, planning, military, maintenance}
 
@@
 
-\section{MFMP}
+TODO.
+
+\section{The problem}
+
+    % Missing: 
+    % * horizon constraints (initial, ending)
+    % * Prefer new
+    % * Total hours
+    % *
 
@@
 
-    There is a set of (XXX) tasks to do in a given planning horizon divided into (XXX) periods of the same length.
-Each task (XXX) has a given starting (XXX) an ending (XXX) time (the duration is measured in a certain number of periods) and needs to use a certain quantity of resources (XXX).
-Not all (XXX) resources can be used to accomplish any task: each task has specific specifications that a resource need to satisfy in order to be a candidate.
-In other words, there is a parameter (XXX) that indicates task-resource compatibility.
+    There is a set of $\vert V \vert$ tasks to do in a given planning horizon divided into $t \in T$ periods of the same length.
+Each task $v \in V$ has a given starting ($st_v \in T$) an ending ($et_v \in T$) time.
+The duration is measured in a certain number of periods.
+Each task requires (seizes) a certain quantity of resources ($Q_v$).
+
+    Not all $a \in A$ resources can be used to accomplish any task: each task has specific specifications that a resource need to satisfy in order to be a candidate.
+In other words, there is a binary parameter $p01va_{va}$ that indicates task-resource compatibility.
 A resource can be a candidate for more than one task.
 
-    The assignment of a resource to a task is not made necessarily for the whole duration of the task.
-After a minimum amount of time (XXX), a resource can be liberated and exchanged for another one.
-The total number of resources being used at any given time in a specific task can never be less than the required XXX.
+    The assignment of a resource to a task is not decided for the whole duration of the task.
+After a minimum amount of time $minTime_v$, a resource can be liberated and exchanged for another one.
+The total number of resources being used at any given time in a specific task can never be less than the required $REQ_v$.
 
@@
 
-    The resources require a recurrent maintenance to be in the correct state to accomplish such tasks correctly.
-These maintenances need to take place before a given amount of time (ElapsedTime XXX) has passed since the previous maintenance or before the resource has been used for a given amount of time (UsageTime XXX) since the previous maintenance, whichever happens first.
-A resource without the proper maintenance cannot be used to satisfy tasks and is consider worthless.
+    Resources are what makes possible the completion of tasks.
+These specific resources suffer from wear and tear and so require a recurrent preventive maintenance in order to be in a correct state.
+These maintenances need to take place before a given amount of time (we will call it elapsed time $et$) has passed since the previous maintenance or before the resource has been used for a given amount of time (we will call it usage time $ut$) since the previous maintenance, whichever happens first.
+A resource without the proper maintenance cannot be used to satisfy tasks and is considered worthless.
 
-    Additionally, after an absolute amount of time and/or usage, the resource becomes obsolete. There is no way to reverse this process.
+    Additionally, after an absolute amount of time and/or usage ($aet$ or $aut$), the resource becomes obsolete. There is no way to reverse this process.
 
@@
     \begin{itemize}
-        \item UsageTime.
-        \item ElapsedTime.
-        \item UsageTimeAbs.
-        \item ElapseTimeAbs.
-        \item LastMaintenanceType.
+        \item remaining UsageTime.
+        \item remaining ElapsedTime.
+        \item remaining UsageTimeAbs.
+        \item remaining ElapseTimeAbs.
+        \item LastMaintenanceType (see \ref{def:maint}).
     \end{itemize}
@@
 
-    These maintenances take an amount of time equal to (XXX) periods and cannot be interrupted.
-The state of the resource after the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (ElapsedTime XXX and UsageTime XXX).
+    These maintenances take an amount of time equal to $mTime_m$ periods and cannot be interrupted.
+The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (ElapsedTime and UsageTime).
 
-    In other words: after a given maintenance XXX, a resource resets its ElapsedTime and UsageTime back to 0, so it can continue being assigned to tasks.
+    In other words, after receiving maintenance $m \in M$, a resource restores its remaining ElapsedTime and remaining UsageTime back to their max values $et$ and $ut$ respectively.
 
-    The number of resources under maintenance in any given period is limited to a maximum capacity (XXX) depending on each period.
+    The number of resources under maintenance in any given period is limited to a maximum capacity ($mCap_t$) depending on each period.
 
@@
 In other words, the second maintenance of a given resource is different in nature from the first maintenance for that same resource.
-Examples of differences on consecutive maintenances are the duration it takes or whether it restores storage capacity (see \ref{def:sto}).
+Examples of differences on consecutive maintenances for a given resource are the duration ($mTime_m$) it takes or whether it restores storage capacity (see \ref{def:sto}).
 
@@
 
-    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (ElapsedTime).
+    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (the elapsed time $et$).
 In order to avoid this problem, the resource can be put into a "storage" state.
 
-    A resource in this states has to be kept for a minimum time (XXX). While in this state it cannot receive maintenance or be assigned any task.
+    A resource in this states has to be kept for a minimum time of $sMin$ periods. While in this state it cannot receive maintenance or be assigned any task.
+
+    Every resource has the capacity to be stored, although it is limited.
+This capacity (measured in a number of periods $sCap$) is expended every time the resource is stored.
+In order for a resource to reset its remaining storage capacity, it needs to received a specific maintenance (see \ref{dec:maint}).
 
@@
 
-    The are multiple objectives that need to be taken into account:
+    The are multiple objectives that need to be taken into account.
+
+    Given that the creation of new tasks and the duration of maintenance times are considered stochastic, the basic goal is to maximize the robustness of the planning.
+
+    Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the maintenance tasks over the planning horizon.
 
     \begin{enumerate}
+        \item Maximize the minimum number of available resources (see \ref{def:res-state}) in any given period.
         \item Minimize the maximum number of resources under maintenance in any given period.
-        \item Maximize the minimum number of resources available (see \ref{def:res-state}) in any given period.
     \end{enumerate}
 
-    Every resource has the capacity to be stored.
-This capacity (measured in number of periods XXX) is expended every time the resource is stored.
-In order for a resource to recover storage capacity, it needs to received a specific maintenance.
+    \subsection{Decisions}
+
+    The decisions involve, basically, the management of these resources.
+
+    We can summarize the main decisions into the following:
+
+    \begin{itemize}
+        \item deciding the state of each resource $a$ in each period $t$.
+        \item if a resource $a$ is assigned to some task in period $t$: decide which task $v$.
+    \end{itemize}
 
@@
 
-\section{Model}
+\section{The model}
+
+    TODO.
 
@@
 \section{Conclusions et perspectives}
+
+    TODO.
 
@@
 
-\begin{thebibliography}{2}
-\bibitem{toth02}
-Paolo Toth and Daniele Vigo.
-\newblock \emph{The Vehicle Routing Problem}.
-\newblock Monographs on Discrete Mathematics and Applications. Society for Industrial and Applied Mathematics, 2002.
-\bibitem{kirkpatrick83}
-Scott Kirkpatrick, C~Daniel Gelatt, and Mario~P Vecchi.
-\newblock Optimization by simmulated annealing.
-\newblock \emph{science}, 220\penalty0 (4598):\penalty0 671--680, 1983.
+% \begin{thebibliography}{2}
+% \bibitem{toth02}
+% Paolo Toth and Daniele Vigo.
+% \newblock \emph{The Vehicle Routing Problem}.
+% \newblock Monographs on Discrete Mathematics and Applications. Society for Industrial and Applied Mathematics, 2002.
+% \bibitem{kirkpatrick83}
+% Scott Kirkpatrick, C~Daniel Gelatt, and Mario~P Vecchi.
+% \newblock Optimization by simmulated annealing.
+% \newblock \emph{science}, 220\penalty0 (4598):\penalty0 671--680, 1983.
 

### 2017-11-10 latex/ROADEF/article_roadef_2018.tex
Added constraints, variables and parameters. First draft version.

@@
 \documentclass{roadef}
+\usepackage{amsmath}
 
@@
 
-% Les auteurs et leur numéro d'affiliation
+% Les AUTeurs et leur numéro d'affiliation
 \author{Franco Peschiera\inst{1}, Alain Haït\inst{1}, Olga Battaïa\inst{1} }
@@
     There is a set of $\vert V \vert$ tasks to do in a given planning horizon divided into $t \in T$ periods of the same length.
-Each task $v \in V$ has a given starting ($st_v \in T$) an ending ($et_v \in T$) time.
+Each task $v \in V$ has a given starting ($startTIME_v \in T$) an ending ($endTIME_v \in T$) time.
 The duration is measured in a certain number of periods.
@@
     Not all $a \in A$ resources can be used to accomplish any task: each task has specific specifications that a resource need to satisfy in order to be a candidate.
-In other words, there is a binary parameter $p01va_{va}$ that indicates task-resource compatibility.
+In other words, there is a binary parameter $P01VA_{va}$ that indicates task-resource compatibility.
 A resource can be a candidate for more than one task.
 
-    The assignment of a resource to a task is not decided for the whole duration of the task.
-After a minimum amount of time $minTime_v$, a resource can be liberated and exchanged for another one.
+    The taskment of a resource to a task is not decided for the whole duration of the task.
+After a minimum amount of time $minTIME_v$, a resource can be liberated and exchanged for another one.
 The total number of resources being used at any given time in a specific task can never be less than the required $REQ_v$.
@@
 These specific resources suffer from wear and tear and so require a recurrent preventive maintenance in order to be in a correct state.
-These maintenances need to take place before a given amount of time (we will call it elapsed time $et$) has passed since the previous maintenance or before the resource has been used for a given amount of time (we will call it usage time $ut$) since the previous maintenance, whichever happens first.
+These maintenances need to take place before a given amount of time (we will call it elapsed time $elapsedTIME$) has passed since the previous maintenance or before the resource has been used for a given amount of time (we will call it usage time $usageTIME$) since the previous maintenance, whichever happens first.
 A resource without the proper maintenance cannot be used to satisfy tasks and is considered worthless.
 
-    Additionally, after an absolute amount of time and/or usage ($aet$ or $aut$), the resource becomes obsolete. There is no way to reverse this process.
+    Additionally, after an absolute amount of time and/or usage ($AET$ or $AUT$), the resource becomes obsolete. There is no way to reverse this process.
 
@@
     \begin{itemize}
-        \item Assigned to a task (see \ref{def:task}).
+        \item tasked to a task (see \ref{def:task}).
         \item Under maintenance (see \ref{def:maint}).
@@
 
-    These maintenances take an amount of time equal to $mTime_m$ periods and cannot be interrupted.
+    These maintenances take an amount of time equal to $mTIME_m$ periods and cannot be interrupted.
 The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (ElapsedTime and UsageTime).
 
-    In other words, after receiving maintenance $m \in M$, a resource restores its remaining ElapsedTime and remaining UsageTime back to their max values $et$ and $ut$ respectively.
+    In other words, after receiving maintenance $m \in M$, a resource restores its remaining ElapsedTime and remaining UsageTime back to their max values $elapsedTIME$ and $usageTIME$ respectively.
 
@@
 In other words, the second maintenance of a given resource is different in nature from the first maintenance for that same resource.
-Examples of differences on consecutive maintenances for a given resource are the duration ($mTime_m$) it takes or whether it restores storage capacity (see \ref{def:sto}).
+Examples of differences on consecutive maintenances for a given resource are the duration ($mTIME_m$) it takes or whether it restores storage capacity (see \ref{def:sto}).
 
@@
 
-    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (the elapsed time $et$).
+    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (the elapsed time $elapsedTIME$).
 In order to avoid this problem, the resource can be put into a "storage" state.
 
-    A resource in this states has to be kept for a minimum time of $sMin$ periods. While in this state it cannot receive maintenance or be assigned any task.
+    A resource in this states has to be kept for a minimum time of $sMin$ periods. While in this state it cannot receive maintenance or be tasked any task.
 
@@
 This capacity (measured in a number of periods $sCap$) is expended every time the resource is stored.
-In order for a resource to reset its remaining storage capacity, it needs to received a specific maintenance (see \ref{dec:maint}).
+In order for a resource to reset its remaining storage capacity, it needs to received a specific maintenance (see \ref{def:maint}).
 
@@
         \item deciding the state of each resource $a$ in each period $t$.
-        \item if a resource $a$ is assigned to some task in period $t$: decide which task $v$.
+        \item if a resource $a$ is tasked to some task in period $t$: decide which task $v$.
     \end{itemize}
@@
 
-    TODO.
+    Below are the parameters used in the model.
 
-    \begin{equation}
-    E=MC^2\label{emc}
-    \end{equation}
+    \begin{tabular}{p{40mm}lp{80mm}}
+    $CAP_t$               & : & number of resources that can be in maintenance in period $t$. \\
+    $CAND_v$              & : & resources that satisfy task's $v$ needs. \\
+    $REQ\_H_{vt}$             & : & time consumed of any resource when assigned to task $v$ in period $t$. \\
+    $REQ\_R_{vt}$            & : & number or resources to assign to task $v$ in period $t$. \\            
+    $MAX\_USAGE$         & : & maximum amount of used time that a resource can have between two maintenances. \\    
+    $MAX\_ELAPSED$       & : & maximum amount of elapsed time that a resource can have between two maintenances. \\    
+    $M\_DURATION$        & : & maintenance duration, in periods. \\        
+    $REM\_USED\_INIT_a$     & : & remaining used time for resource $a$ at the beginning of horizon. \\        
+    $REM\_ELAPSED\_INIT_a$  & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
+    \end{tabular}
+    \bigskip
+
+    Below are the variables used in the model.
+
+    \section{Decision variables}
+
+    \begin{tabular}{p{30mm}lp{90mm}}
+        $ret_{at}$       & : & remaining elapsed time for resource $a$ after period $t$ \\  
+        $rut_{at}$       & : & remaining usage time for resource $a$ after period $t$ \\  
+        $task_{avt}$    & : & 1 if resource $a \in CAND_v$ is assigned task $v$ in period $t$ \\  
+        $state_{ast}$   & : & 1 if resource $a$ is assigned state $s$ in period $t$ \\  
+        $used_{at}$     & : & time consumed by resource $a$ in period $t$ \\
+        $start_{at}$     & : & 1 if resource $a$ starts maintenance in period $t$ \\
+        $available$     & : & minimum number of available resources in any period \\
+        $maintenance$     & : & maxium number of resources in maintenance in any period \\
+    \end{tabular}
+    \bigskip
+
+    \begin{align}
+        %%%%%%%%%%%%%%%%%%%
+        % MAIN VARS
+        %%%%%%%%%%%%%%%%%%%
+        % definition of consumed resource hours:
+        & used_{at} = \sum_{v \in \mathcal{V}} task_{avt} \times REQ\_H_{vt}
+                \hspace{10mm} a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
+        % maximum capacity:
+        & \sum_{a \in \mathcal{A}} state_{ast} \leq CAP_{t}
+                \hspace{10mm} t \in \mathcal{T}, s = 'M' \label{eq:capacity}\\
+        % minimum resources
+        & \sum_{a \in \mathcal{A}} task_{avt} \geq REQ\_R_{vt}
+                \hspace{10mm} t \in \mathcal{T}, v \in \mathcal{V} \label{eq:resources}\\
+        % max one task per period
+        & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:resources}\\
+        %%%%%%%%%%%%%%%%%%%
+        % BALANCES
+        %%%%%%%%%%%%%%%%%%%
+        % remaining used time
+        & rut_{at} \leq (rut_{at-1} - used_{at}) + (MAX\_USAGE \times start_{at})
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance1}\\
+        % remaining elapsed time. TODO: improve this constraint like the maintenance state duration
+        & ret_{at} \leq (ret_{at-1} - 1) + (MAX\_ELAPSED \times start_{at})
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}\\
+        %%%%%%%%%%%%%%%%%%%
+        % STATES
+        %%%%%%%%%%%%%%%%%%%
+        % if start maintenance, at least X periods
+        & state_{ast_2} \geq start_{at_1} \\
+                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION, a \in \mathcal{A}, s = 'M' \label{eq:state1}\\
+        % if assigned to task; in workng state
+        & state_{ast} \geq task_{avt}
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'V' \label{eq:state2}\\
+        % only one state. If we take out the 'do nothing' state, this could be \leq
+        & \sum_{s \in \mathcal{S}} state_{ast} = 1
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state3}\\
+        %%%%%%%%%%%%%%%%%%%
+        % OBJECTIVE
+        %%%%%%%%%%%%%%%%%%%
+        % min available
+        & available \leq \sum_{a} state_{ast}
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'N' \label{eq:available}\\
+        % max maintenance
+        & maintenance \geq \sum_{a} state_{ast}
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'N' \label{eq:maintenance}\\
+        %%%%%%%%%%%%%%%%%%%
+        % HORIZON
+        %%%%%%%%%%%%%%%%%%%
+        % rut initialize
+        & rut_{at} = REM\_USED\_INIT_a
+                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:initialize1}\\
+        % ret initialize
+        & ret_{at} = REM\_ELAPSED\_INIT_a
+                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:initialize2}\\
+    \end{align}
 
@@
 
-\end{thebibliography}
+% \end{thebibliography}
 

### 2017-11-13 latex/ROADEF/article_roadef_2018.tex
added explanations to constraints and corrected inconsistencies in naming

@@
 \usepackage{amsmath}
+
+% \usepackage{fontspec}
 
@@
 
-    There is a set of $\vert V \vert$ tasks to do in a given planning horizon divided into $t \in T$ periods of the same length.
-Each task $v \in V$ has a given starting ($startTIME_v \in T$) an ending ($endTIME_v \in T$) time.
-The duration is measured in a certain number of periods.
-Each task requires (seizes) a certain quantity of resources ($Q_v$).
+    Each $v \in \mathcal{V}$ task needs to be accomplished in a planning horizon divided into $t \in \mathcal{T}$ periods of the same length.
+Each task has a given starting ($startTIME_v \in T$) an ending ($endTIME_v \in T$) time.
+The starting and ending times are given and cannot be changed.
 
-    Not all $a \in A$ resources can be used to accomplish any task: each task has specific specifications that a resource need to satisfy in order to be a candidate.
-In other words, there is a binary parameter $P01VA_{va}$ that indicates task-resource compatibility.
+    Given that the starting and ending times belong to $\mathcal{T}$, the duration of a task is measured in a certain number of periods.
+Each task requires (seizes) a certain quantity of resources ($REQ_{vt}$).
+For resources, see \ref{def:resource}.
+
+    Not all $a \in \mathcal{A}$ resources can be used to accomplish any task: each task has specific requirements that a resource needs to satisfy in order to be a candidate.
+In other words, for each task $t$, there is a set $CAND_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to that task.
 A resource can be a candidate for more than one task.
 
-    The taskment of a resource to a task is not decided for the whole duration of the task.
-After a minimum amount of time $minTIME_v$, a resource can be liberated and exchanged for another one.
-The total number of resources being used at any given time in a specific task can never be less than the required $REQ_v$.
+    % The assignment of a resource to a task is not decided for the whole duration of the task.
+After a minimum amount of time ($minTIME_v$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished.
+The total number of resources being used at any given time in a specific task can never be less than the required $REQ_{vt}$.
 
     \subsection{Resources}
+    \label{def:resource}
 
     Resources are what makes possible the completion of tasks.
-These specific resources suffer from wear and tear and so require a recurrent preventive maintenance in order to be in a correct state.
-These maintenances need to take place before a given amount of time (we will call it elapsed time $elapsedTIME$) has passed since the previous maintenance or before the resource has been used for a given amount of time (we will call it usage time $usageTIME$) since the previous maintenance, whichever happens first.
-A resource without the proper maintenance cannot be used to satisfy tasks and is considered worthless.
+These specific resources suffer from wear and tear and so require a recurrent preventive maintenance (see \ref{def:maint}) in order to be in a correct state.
+These maintenances need to take place before a given amount of time (we will call it $MAX\_ELPASED$) has passed since the previous maintenance or before the resource has been used for a given amount of time (we will call it $MAX\_USAGE$) since the previous maintenance, whichever happens first.
+While a resource $a$ has not received the proper maintenance, it cannot be used to satisfy tasks and is considered useless.
 
-    Additionally, after an absolute amount of time and/or usage ($AET$ or $AUT$), the resource becomes obsolete. There is no way to reverse this process.
+    We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance.
+
+    We will call "remaining usage time" (or $rut$) to the amount of time that the resource needs to be used in order for it to need a maintenance.
+
+    % Additionally, after an absolute amount of time and/or usage ($AET$ or $AUT$), the resource becomes obsolete. There is no way to reverse this process.
 
@@
     \begin{itemize}
-        \item remaining UsageTime.
-        \item remaining ElapsedTime.
-        \item remaining UsageTimeAbs.
-        \item remaining ElapseTimeAbs.
-        \item LastMaintenanceType (see \ref{def:maint}).
+        \item remaining usage time.
+        \item remaining elapsed time.
+        % \item remaining absolute elapsed time.
+        % \item remaining absolute usage time.
+        % \item LastMaintenanceType (see \ref{def:maint}).
     \end{itemize}
@@
 
-    The following are the possible states for a resource in any given time:
+    The following are the possible $s \in \mathcal{S}$ states for a resource in any given time:
 
     \begin{itemize}
-        \item tasked to a task (see \ref{def:task}).
+        \item Assigned to a task (see \ref{def:task}).
         \item Under maintenance (see \ref{def:maint}).
         \item Under storage (see \ref{def:sto}).
-        \item Waiting maintenance.
-        \item Obsolete.
+        \item Waiting maintenances.
+        % \item Obsolete.
         \item Available (none of the above).
@@
 
-    These maintenances take an amount of time equal to $mTIME_m$ periods and cannot be interrupted.
-The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (ElapsedTime and UsageTime).
+    Maintenances are the process by which resources that have reached a limit in their usage (but not their absolute limit) can return to a state where they can continue to be used in tasks.
 
-    In other words, after receiving maintenance $m \in M$, a resource restores its remaining ElapsedTime and remaining UsageTime back to their max values $elapsedTIME$ and $usageTIME$ respectively.
+    Each maintenance belongs to a specific type of maintenance $m \in \mathcal{M}$. These types of maintenance differentiate between each other by having different characteristics.
 
-    The number of resources under maintenance in any given period is limited to a maximum capacity ($mCap_t$) depending on each period.
+    These maintenances take an amount of time equal to $M\_DURATION_m$ periods and cannot be interrupted.
+The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (remaining elapsed time and remaining usage time).
 
-    Not all maintenances are the same for any given resource.
+    In other words, after receiving maintenance of type $m$, a resource restores its remaining elapsed time and remaining usage time back to their max values $MAX\_ELAPSED$ and $MAX\_USAGE$ respectively.
+
+    The number of resources under maintenance in any given period is limited to a maximum capacity ($CAP_t$) depending on each period.
+
+    % Not all maintenances are the same for any given resource.
 Each resources follows a sequence of maintenances.
@@
 
-    % \begin{theoreme}
-    %   Un exemple de théorème. Les environnements suivants sont également disponibles : remarque, propriété, corollaire, définition, notation, proposition, exemple, preuve.
-    % \end{theoreme}
-
-    % \subsection{Tableau}
-
-    % Le titre du tableau devra être positionné sous le tableau (par exemple Tableau \ref{tableau}).
-
-    % \begin{table}[!ht]
-    %     \begin{center}
-    %         \begin{tabular}{lrr}
-    %             \hline
-    %             & \multicolumn{1}{c}{Colonne 1} & \multicolumn{1}{c}{Colonne 2}\\
-    %             \hline
-    %             Ligne 1 & L1C1 & L1C2\\
-    %             Ligne 2 & L2C1 & L2C2\\
-    %             \hline
-    %         \end{tabular}
-    %         \caption{Exemple de tableau}
-    %     \label{tableau}
-    %     \end{center}
-    % \end{table}
-
-    % \subsection{Liste}
-
-    % Voici une liste :
-
-    % \begin{itemize}
-    % \item remarque
-    % \item propriété
-    % \end{itemize}
-
 \section{The model}
@@
 
-    \begin{tabular}{p{40mm}lp{80mm}}
+    \begin{tabular}{p{30mm}lp{90mm}}
     $CAP_t$                 & : & number of resources that can be in maintenance in period $t$. \\
@@
     $M\_DURATION$           & : & maintenance duration, in periods. \\        
-    $REM\_USED\_INIT_a$     & : & remaining used time for resource $a$ at the beginning of horizon. \\        
-    $REM\_ELAPSED\_INIT_a$  & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
+    $RUT\_INI_a$            & : & remaining used time for resource $a$ at the beginning of horizon. \\        
+    $RET\_INI_a$            & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
+    $STATE\_INI_{at}$       & : & some $s \in \mathcal{S}$ in case the resource $a$ has a fixed state. If not; this is not filled. \\
     \end{tabular}
@@
     \begin{tabular}{p{30mm}lp{90mm}}
-        $ret_{at}$       & : & remaining elapsed time for resource $a$ after period $t$ \\  
-        $rut_{at}$       & : & remaining usage time for resource $a$ after period $t$ \\  
+        $ret_{at}$      & : & remaining elapsed time for resource $a$ after period $t$ (including t=0) \\  
+        $rut_{at}$      & : & remaining usage time for resource $a$ after period $t$ (including t=0) \\  
         $task_{avt}$    & : & 1 if resource $a \in CAND_v$ is assigned task $v$ in period $t$ \\  
@@
         %%%%%%%%%%%%%%%%%%%
-        % definition of consumed resource hours:
-        & used_{at} = \sum_{v \in \mathcal{V}} task_{avt} \times REQ\_H_{vt}
-                \hspace{10mm} a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
         % maximum capacity:
@@
         & \sum_{a \in \mathcal{A}} task_{avt} \geq REQ\_R_{vt}
-                \hspace{10mm} t \in \mathcal{T}, v \in \mathcal{V} \label{eq:resources}\\
+                \hspace{10mm} t \in \mathcal{T}, v \in \mathcal{V} \label{eq:taskres}\\
         % max one task per period
         & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:resources}\\
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
+    \end{align}
+
+    These constraints are the the main constraints of the model.
+
+    \ref{eq:taskres} and \ref{eq:capacity} limit the number of resources assigned to task and the number of resource maintenances that can be done in any given month.
+    \ref{eq:maxtasks} limits the number of assigned tasks for a resource to one.
+
+    \begin{align}
         %%%%%%%%%%%%%%%%%%%
@@
         %%%%%%%%%%%%%%%%%%%
+        % definition of consumed resource hours:
+        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times REQ\_H_{vt}
+                \hspace{10mm} a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
         % remaining used time
@@
                 \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}\\
+    \end{align}
+
+    These constraints calculate the balances of hours for each resource.
+
+    \ref{eq:usage} defines the consumption of hours for each resource at each period.
+This is then used in \ref{eq:balance1} to get the remaining usage time at each period for each resource.
+Similarly, the \ref{eq:balance2} calculates the remaining elapsed time at each period and resource.
+
+    \begin{align}
         %%%%%%%%%%%%%%%%%%%
@@
         % if start maintenance, at least X periods
-        & state_{ast_2} \geq start_{at_1} \\
-                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION, a \in \mathcal{A}, s = 'M' \label{eq:state1}\\
-        % if assigned to task; in workng state
+        & state_{ast_2} \geq start_{at_1} \notag\\
+                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = 'M' \label{eq:start1}\\
+        % if maintenance doesn't start, we cannot have maintenance:
+        & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} \notag\\
+                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = 'M' \label{eq:start2}\\
+        % if assigned to task: in working state
         & state_{ast} \geq task_{avt}
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'V' \label{eq:state2}\\
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'V' \label{eq:state1}\\
         % only one state. If we take out the 'do nothing' state, this could be \leq
         & \sum_{s \in \mathcal{S}} state_{ast} = 1
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state3}\\
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}\\
+    \end{align}
+
+    These constraints involve the states assignments to resources.
+
+    \ref{eq:start1} and \ref{eq:start2} tie the start of each maintenance with the maintenance state so resources need to have started a maintenance to receive it and have to have maintenance for a minimum period after having starting it.
+
+    \ref{eq:state1} makes working states with being assigned to a task. \ref{eq:state2} makes resources have only one state at each period.
+
+    \begin{align}
         %%%%%%%%%%%%%%%%%%%
@@
                 \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'N' \label{eq:maintenance}\\
+    \end{align}
+
+    These constraints define the possible objectives.
+
+    \ref{eq:available} is defined as the minimum number of available resources in any period.
+
+    \ref{eq:maintenance} is defined as the maximum number of resources under maintenance in any period.
+
+    \begin{align}
         %%%%%%%%%%%%%%%%%%%
@@
         %%%%%%%%%%%%%%%%%%%
+        % TODO: initialize maintenance states too!
         % rut initialize
-        & rut_{at} = REM\_USED\_INIT_a
-                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:initialize1}\\
+        & rut_{at} = RUT\_INI_a
+                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
         % ret initialize
-        & ret_{at} = REM\_ELAPSED\_INIT_a
-                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:initialize2}\\
+        & ret_{at} = RET\_INI_a
+                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
+        % fixed states during planning period
+        & state_{ast} = STATE\_INI_{at}
+                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}\, | if\, STATE\_INI_{at} \neq \{\} \label{eq:fixedstates}\\
     \end{align}
+
+    These constraints feed the relevant historic data for resources.
+
+    \ref{eq:remaining1} and \ref{eq:remaining2} fill the last period's remaining used and elapsed time respectively.
+
+    \ref{eq:fixedstates} fixes the states to their planned state.
+This is specially useful for the beginning of the planning period, where we could have maintenance states already decided for some resources.
 

### 2017-11-13 latex/ROADEF/article_roadef_2018.tex
added R scripts to generate calendar, input data with calendar in Excel and small aesthetic modifications in latex

@@
-% Exemple d'utilisation de la classe roadef2018 pour le congrès ROADEF 2018 (http://roadef2018.labsticc.fr)
-
 \documentclass{roadef}
@@
 % \usepackage{fontspec}
+% This command is to use simple quotes inside math expressions:
+\newcommand{\mq}[1] {`#1\textrm'}
 
@@
 
-% Les AUTeurs et leur numéro d'affiliation
-\author{Franco Peschiera\inst{1}, Alain Haït\inst{1}, Olga Battaïa\inst{1} }
+\author{Franco Peschiera, Alain Haït, Olga Battaïa }
 
-% Les affiliations (par ordre croissant des numéros d'affiliation) séparées par \and
 \institute{
@@
 
-% Création de la page de titre
 \maketitle
@@
 
-% Les mots-clés
 \keywords{optimization, planning, military, maintenance}
@@
         $available$     & : & minimum number of available resources in any period \\
-        $maintenance$   & : & maxium number of resources in maintenance in any period \\
+        $maintenance$   & : & maximum number of resources in maintenance in any period \\
     \end{tabular}
@@
         & \sum_{a \in \mathcal{A}} state_{ast} \leq CAP_{t}
-                \hspace{10mm} t \in \mathcal{T}, s = 'M' \label{eq:capacity}\\
+                & t \in \mathcal{T}, s = 'M' \label{eq:capacity}\\
         % minimum resources
         & \sum_{a \in \mathcal{A}} task_{avt} \geq REQ\_R_{vt}
-                \hspace{10mm} t \in \mathcal{T}, v \in \mathcal{V} \label{eq:taskres}\\
+                & t \in \mathcal{T}, v \in \mathcal{V} \label{eq:taskres}\\
         % max one task per period
         & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
+                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
     \end{align}
@@
         & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times REQ\_H_{vt}
-                \hspace{10mm} a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
+                & a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
         % remaining used time
         & rut_{at} \leq (rut_{at-1} - used_{at}) + (MAX\_USAGE \times start_{at})
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance1}\\
+                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance1}\\
         % remaining elapsed time. TODO: improve this constraint like the maintenance state duration
         & ret_{at} \leq (ret_{at-1} - 1) + (MAX\_ELAPSED \times start_{at})
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}\\
+                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}\\
     \end{align}
@@
         & state_{ast_2} \geq start_{at_1} \notag\\
-                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = 'M' \label{eq:start1}\\
+                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start1}\\
         % if maintenance doesn't start, we cannot have maintenance:
         & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} \notag\\
-                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = 'M' \label{eq:start2}\\
+                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start2}\\
         % if assigned to task: in working state
-        & state_{ast} \geq task_{avt}
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'V' \label{eq:state1}\\
+        & state_{ast} \geq task_{avt} \notag\\
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{V} \label{eq:state1}\\
         % only one state. If we take out the 'do nothing' state, this could be \leq
-        & \sum_{s \in \mathcal{S}} state_{ast} = 1
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}\\
+        & \sum_{s \in \mathcal{S}} state_{ast} = 1 \notag\\
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}\\
     \end{align}
@@
         & available \leq \sum_{a} state_{ast}
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'N' \label{eq:available}\\
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:available}\\
         % max maintenance
         & maintenance \geq \sum_{a} state_{ast}
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = 'N' \label{eq:maintenance}\\
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:maintenance}\\
     \end{align}
@@
         & rut_{at} = RUT\_INI_a
-                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
+                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
         % ret initialize
         & ret_{at} = RET\_INI_a
-                \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
+                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
         % fixed states during planning period
-        & state_{ast} = STATE\_INI_{at}
-                \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}\, | if\, STATE\_INI_{at} \neq \{\} \label{eq:fixedstates}\\
+        & state_{ast} = (STATE\_INI_{at} = s)
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}\, | if\, STATE\_INI_{at} \neq \mq{} \label{eq:fixedstates}\\
     \end{align}
@@
 
-% Version "on-line" de la bibliographie, mais il est
-% également possible d'utiliser un fichier ".bib" et d'utiliser BibTeX
-
 % \begin{thebibliography}{2}

### 2017-11-13 R/scripts/import_results.R

@@ -1,0 +1,67 @@
+library(tidyverse)
+library(readxl)
+library(magrittr)
+library(stringr)
+library(zoo)
+library(data.table)
+library(lubridate)
+library(timevis)
+library(Hmisc)
+library(RColorBrewer)
+
+path <- '/home/pchtsp/Documents/projects/OPTIMA/data/raw/Planifs M2000.xlsm'
+
+
+sheets <- readxl::excel_sheets(path)
+table <- read_xlsx(path, "M2D")
+months_names <- str_split("Ja  Fe  Ma  Av  Mi  Jn  Jt  Au  Se  Oc  No  De", "\\s+") %>% unlist
+equiv <- data.table(value=months_names, num=c(1:length(months_names)) %>% str_pad(2, "left", "0"))
+
+table_names <- 
+    table %>% 
+    slice(1) %>% 
+    gather() %>% 
+    full_join(equiv) %>% 
+    # mutate(year= years) %>% 
+    mutate(w_year = str_detect(key, "(^\\d{4}\\s)|^2"),
+           year = if_else(w_year, key, "") %>% str_extract("^\\d{4}") %>% na.locf(na.rm = FALSE),
+           colname= if_else(is.na(year) | is.na(num), make.names(key), paste(year, num, sep="-"))) 
+
+states <- 
+    table %>% 
+    set_names(table_names$colname) %>% 
+    select(UNIT= X__3, starts_with("20")) %>% 
+    slice(2:78) %>% 
+    gather(key= "month", value="state", starts_with("20")) %>% 
+    filter(str_sub(UNIT, 1, 1)=="D") %>% 
+    arrange(UNIT, month) %>% 
+    group_by(UNIT) %>% 
+    mutate(prev_state= state %>% lag,
+           first_row= row_number()==1,
+           change = state != prev_state) %>% 
+    filter(first_row | change) %>% 
+    mutate(post_month = month %>% lead %>% paste0("-01") %>% ymd() %>% add(0) %>% format("%Y-%m")) %>% 
+    ungroup() %>% 
+    mutate(id= c(1:nrow(.))) %>% 
+    select(id, start= month, end= post_month, content= state, group= UNIT ) %>% 
+    mutate(bucket= content %>% as.integer %>% cut2(g= 5),
+           maint = content %>% str_detect('^V\\d+'),
+           color= RColorBrewer::brewer.pal(5, "YlOrRd")[bucket],
+           color= if_else(maint, "black", color),
+           color= if_else(color %>% is.na, '', color),
[…]

### 2017-11-15 latex/ROADEF/article_roadef_2018.tex
added modifications to article and to R scripts that makes graph

@@
 
-\author{Franco Peschiera, Alain Haït, Olga Battaïa }
+\author{Franco Peschiera\inst{1}, Alain Haït\inst{1}, Olga Battaïa\inst{1}, Nicolas Dupin\inst{2}}
 
+% Les affiliations (par ordre croissant des numéros d'affiliation) séparées par \and
 \institute{
-ISAE... \\
-\email{\{franco.peschiera\}@isae-supaero.fr}
+Institut Supérieur de l'Aéronautique et de l'Espace \\
+\email{\{franco.peschiera,alain.haït,olga.battaia\}@isae-supaero.fr}
+\and
+Direction Générale de l'Armement \\
+\email{nicolas.dupin@intradef.gouv.fr}
 }
@@
 
+    Le problème des planifications de maintenances des avions commerciales est un sujet qui a été étudié beaucoup dans la littérature.
+Dans ce problème on affecte des avions à vols au même temps qu'on les affecte à des maintenances.
+Il est aussi similaire à la planification des maintenances de matériel roulant.
+
+    Par contre, le problème des planifications de maintenances des avions militaires n'a pas été étudié que pour quelques auteurs.
+Les principales différences se trouve aux objectives, à la taille de l'horizon de planification et au fait que les avions retourne toujours à la même base après les opérations.
+
+    L'objectif dans le Military Flight and Maintenance Planning n'est pas ce de minimiser le coût total de la planification si non ce de lisser les plus possible les opérations de maintenances au long de l'horizon au même temps que de lisser le plus possible la disponibilité des avions \cite{Kozanidis2008}, \cite{Cho2011} et \cite{Verhoeff2015}.
+
+    En plus, chaque armée est différente une de l'autre et, donc, les contraints et buts qui conditionnent les opérations de maintenances sont aussi différentes.
+
+    On a formulé des nouveaux contraints, comme la possibilité de stocker temporairement des avions ou la nécessité de affecter des avions dans un temps minimal au missions.
+Ces contraints font que le modèle devient plus complexe en ajoutant une dépendance plus fort entre les mois consécutifs.
+
 \section{The problem}
@@
 
-    Each $v \in \mathcal{V}$ task needs to be accomplished in a planning horizon divided into $t \in \mathcal{T}$ periods of the same length.
+    Each $v \in \mathcal{V}$ task needs to be done in a planning horizon divided into $t \in \mathcal{T}$ periods of the same length.
 Each task has a given starting ($startTIME_v \in T$) an ending ($endTIME_v \in T$) time.
@@
 
-    Given that the starting and ending times belong to $\mathcal{T}$, the duration of a task is measured in a certain number of periods.
-Each task requires (seizes) a certain quantity of resources ($REQ_{vt}$).
-For resources, see \ref{def:resource}.
+    Each task $v$ requires (seizes) a certain quantity of resources $a \in \mathcal{A}$ each period $t$ ($REQ_{vt}$). For resources, see \ref{def:resource}.
 
-    Not all $a \in \mathcal{A}$ resources can be used to accomplish any task: each task has specific requirements that a resource needs to satisfy in order to be a candidate.
-In other words, for each task $t$, there is a set $CAND_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to that task.
+    Not all $a$ resources can be used to accomplish any task: each task has specific requirements that a resource needs to satisfy in order to be a candidate.
+In other words, for each task $v$, there is a set $cd_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to that task.
 A resource can be a candidate for more than one task.
@@
 These specific resources suffer from wear and tear and so require a recurrent preventive maintenance (see \ref{def:maint}) in order to be in a correct state.
-These maintenances need to take place before a given amount of time (we will call it $MAX\_ELPASED$) has passed since the previous maintenance or before the resource has been used for a given amount of time (we will call it $MAX\_USAGE$) since the previous maintenance, whichever happens first.
+These maintenances need to take place before a given amount of time (that we will call $MAX\_ELPASED$) has passed since the previous maintenance or before the resource has been used for a given amount of time (that we will call $mu$) since the previous maintenance, whichever happens first.
 While a resource $a$ has not received the proper maintenance, it cannot be used to satisfy tasks and is considered useless.
 
-    We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance.
+    We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance at any given period.
 
-    We will call "remaining usage time" (or $rut$) to the amount of time that the resource needs to be used in order for it to need a maintenance.
+    We will call "remaining usage time" (or $rut$) to the amount of time that the resource needs to be used in order for it to need a maintenance at any given period.
 
@@
 
-    The following are the possible $s \in \mathcal{S}$ states for a resource in any given time:
+    The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time:
 
@@
 
-    These maintenances take an amount of time equal to $M\_DURATION_m$ periods and cannot be interrupted.
+    These maintenances take an amount of time equal to $md_m$ periods and cannot be interrupted.
 The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (remaining elapsed time and remaining usage time).
 
-    In other words, after receiving maintenance of type $m$, a resource restores its remaining elapsed time and remaining usage time back to their max values $MAX\_ELAPSED$ and $MAX\_USAGE$ respectively.
+    In other words, after receiving a maintenance, a resource restores its remaining elapsed time and remaining usage time back to their max values $me$ and $mu$ respectively.
 
-    The number of resources under maintenance in any given period is limited to a maximum capacity ($CAP_t$) depending on each period.
+    The number of resources under maintenance in any given period is limited to a maximum capacity ($c_t$) which depends on each period.
 
@@
 
-    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (the elapsed time $elapsedTIME$).
+    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (govern by the elapsed time $elapsedTIME$).
 In order to avoid this problem, the resource can be put into a "storage" state.
@@
 
-    % \begin{figure}[!ht]
-    %     \begin{center}
-    %         \includegraphics[height=2cm,clip=true]{roadef_logo.eps}
-    %         \caption[Fig]{Logo de la ROADEF}
-    %         \label{logoRoadef}
-    %     \end{center}
-    % \end{figure}
-
-    % \begin{equation}
-    % E=MC^2\label{emc}
-    % \end{equation}
-
 \section{The model}
@@
 
-    \begin{tabular}{p{30mm}lp{90mm}}
-    $CAP_t$                 & : & number of resources that can be in maintenance in period $t$. \\
-    $CAND_v$                & : & resources that satisfy task's $v$ needs. \\
-    $REQ\_H_{vt}$           & : & time consumed of any resource when assigned to task $v$ in period $t$. \\
-    $REQ\_R_{vt}$           & : & number or resources to assign to task $v$ in period $t$. \\            
-    $MAX\_USAGE$            & : & maximum amount of used time that a resource can have between two maintenances. \\    
-    $MAX\_ELAPSED$          & : & maximum amount of elapsed time that a resource can have between two maintenances. \\    
-    $M\_DURATION$           & : & maintenance duration, in periods. \\        
-    $RUT\_INI_a$            & : & remaining used time for resource $a$ at the beginning of horizon. \\        
-    $RET\_INI_a$            & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
-    $STATE\_INI_{at}$       & : & some $s \in \mathcal{S}$ in case the resource $a$ has a fixed state. If not; this is not filled. \\
+    \begin{tabular}{p{10mm}lp{110mm}}
+    $c_t$                 & : & number of resources that can be in maintenance in period $t$. \\
+    $cd_v$                & : & resources that satisfy task's $v$ needs. \\
+    $rh_{vt}$           & : & time consumed of any resource when assigned to task $v$ in period $t$. \\
+    $rr_{vt}$           & : & number or resources to assign to task $v$ in period $t$. \\            
+    $mu$            & : & maximum used time for each resource between maintenances. \\    
+    $me$          & : & maximum elapsed time for each resource between maintenances. \\    
+    $md$           & : & maintenance duration, in periods. \\        
+    $rut\_i_a$            & : & remaining used time for resource $a$ at the beginning of horizon. \\        
+    $ret\_i_a$            & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
+    $state\_i_{at}$       & : & some $s \in \mathcal{S}$ in case the resource $a$ has a fixed state. \\
     \end{tabular}
@@
 
-    \section{Decision variables}
-
-    \begin{tabular}{p{30mm}lp{90mm}}
-        $ret_{at}$      & : & remaining elapsed time for resource $a$ after period $t$ (including t=0) \\  
-        $rut_{at}$      & : & remaining usage time for resource $a$ after period $t$ (including t=0) \\  
-        $task_{avt}$    & : & 1 if resource $a \in CAND_v$ is assigned task $v$ in period $t$ \\  
+    \begin{tabular}{p{20mm}lp{100mm}}
+        $ret_{at}$      & : & remaining elapsed time for resource $a$ after period $t$ \\  
+        $rut_{at}$      & : & remaining usage time for resource $a$ after period $t$ \\  
+        $task_{avt}$    & : & 1 if resource $a \in cd_v$ is assigned task $v$ in period $t$ \\  
         $state_{ast}$   & : & 1 if resource $a$ is assigned state $s$ in period $t$ \\  
@@
         % maximum capacity:
-        & \sum_{a \in \mathcal{A}} state_{ast} \leq CAP_{t}
+        & \sum_{a \in \mathcal{A}} state_{ast} \leq c_{t}
                 & t \in \mathcal{T}, s = 'M' \label{eq:capacity}\\
         % minimum resources
-        & \sum_{a \in \mathcal{A}} task_{avt} \geq REQ\_R_{vt}
+        & \sum_{a \in \mathcal{A}} task_{avt} \geq rr_{vt}
                 & t \in \mathcal{T}, v \in \mathcal{V} \label{eq:taskres}\\
@@
         & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
-                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
+                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}
     \end{align}
@@
         % definition of consumed resource hours:
-        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times REQ\_H_{vt}
+        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times rh_{vt}
                 & a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
         % remaining used time
-        & rut_{at} \leq (rut_{at-1} - used_{at}) + (MAX\_USAGE \times start_{at})
+        & rut_{at} \leq (rut_{at-1} - used_{at}) + (mu \times start_{at})
                 & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance1}\\
         % remaining elapsed time. TODO: improve this constraint like the maintenance state duration
-        & ret_{at} \leq (ret_{at-1} - 1) + (MAX\_ELAPSED \times start_{at})
-                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}\\
+        & ret_{at} \leq (ret_{at-1} - 1) + (me \times start_{at})
+                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}
     \end{align}
@@
         % if start maintenance, at least X periods
-        & state_{ast_2} \geq start_{at_1} \notag\\
-                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start1}\\
+        & state_{ast_2} \geq start_{at_1} 
+                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + md - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start1}\\
         % if maintenance doesn't start, we cannot have maintenance:
-        & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} \notag\\
-                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + M\_DURATION - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start2}\\
+        & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} 
+                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + md - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start2}\\
         % if assigned to task: in working state
-        & state_{ast} \geq task_{avt} \notag\\
+        & state_{ast} \geq task_{avt} 
                 & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{V} \label{eq:state1}\\
         % only one state. If we take out the 'do nothing' state, this could be \leq
-        & \sum_{s \in \mathcal{S}} state_{ast} = 1 \notag\\
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}\\
+        & \sum_{s \in \mathcal{S}} state_{ast} = 1 
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}
     \end{align}
@@
         & maintenance \geq \sum_{a} state_{ast}
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:maintenance}\\
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:maintenance}
     \end{align}
@@
         % rut initialize
-        & rut_{at} = RUT\_INI_a
+        & rut_{at} = rut\_ini_a
                 & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
         % ret initialize
-        & ret_{at} = RET\_INI_a
+        & ret_{at} = ret\_ini_a
                 & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
         % fixed states during planning period
-        & state_{ast} = (STATE\_INI_{at} = s)
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}\, | if\, STATE\_INI_{at} \neq \mq{} \label{eq:fixedstates}\\
+        & state_{ast} = 1
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}; s \in \mathcal{S}\, | if\, state\_ini_{at} = s \label{eq:fixedstates}
     \end{align}
@@
 \bibliographystyle{plain}
-
+\bibliography{MFMP}
 % \begin{thebibliography}{2}

### 2017-11-16 latex/ROADEF/article_roadef_2018.tex
modifications to latex article, added bibliography

@@
 % Le titre du papier
-\title{Maintenance planning on french military aircraft operations}
+\title{Maintenance planning on French military aircraft operations}
 
@@
 
-\author{Franco Peschiera\inst{1}, Alain Haït\inst{1}, Olga Battaïa\inst{1}, Nicolas Dupin\inst{2}}
+\author{Franco Peschiera\inst{1}, Alain Haït\inst{1}, Olga Battaïa\inst{1}, Nicolas Dupin\inst{1}}
 
@@
 Institut Supérieur de l'Aéronautique et de l'Espace \\
-\email{\{franco.peschiera,alain.haït,olga.battaia\}@isae-supaero.fr}
-\and
-Direction Générale de l'Armement \\
-\email{nicolas.dupin@intradef.gouv.fr}
+\email{\{franco.peschiera,alain.hait,olga.battaia,nicolas.dupin\}@isae-supaero.fr}
 }
@@
 
-    The Flight and Maintenance Planning is a very well known and studied problem in the aviation industry. It
+%  TODO!: re-arrange all this better.
 
-    ...
+    The Military Flight and Maintenance Planning is a problem consisting on the assignment of missions and maintenances tasks to military aircraft.
+It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically but they are assumed to include a round-trip to the base.
+It also includes different objectives and constraints.
 
-    Military Flight and Maintenance Planning...
+    Although the former has been studied in literature in \cite{Cho2011}, \cite{Kozanidis2008}, \cite{Verhoeff2015} and \cite{Chastellux2017}, it has not received as much attention as the latter.
 
-    TODO.
+    % First of all, the objective function's motivation is on reducing the load on the maintenance facilities and guaranteeing the availability of the fleet instead of reducing the cost.
+Also, the maintenance operations take considerably longer and the rules that govern the maintenances are different.
 
-    Le problème des planifications de maintenances des avions commerciales est un sujet qui a été étudié beaucoup dans la littérature.
-Dans ce problème on affecte des avions à vols au même temps qu'on les affecte à des maintenances.
-Il est aussi similaire à la planification des maintenances de matériel roulant.
+    % The temporal scope of the problem is also bigger, making the normal planning horizon bigger and each period bigger too: months instead of days.
+On the other hand, the physical scope, while present in the commercial planning, is not present in the military one: aircraft are assumed to return to the origin every time.
 
-    Par contre, le problème des planifications de maintenances des avions militaires n'a pas été étudié que pour quelques auteurs.
-Les principales différences se trouve aux objectives, à la taille de l'horizon de planification et au fait que les avions retourne toujours à la même base après les opérations.
-
-    L'objectif dans le Military Flight and Maintenance Planning n'est pas ce de minimiser le coût total de la planification si non ce de lisser les plus possible les opérations de maintenances au long de l'horizon au même temps que de lisser le plus possible la disponibilité des avions \cite{Kozanidis2008}, \cite{Cho2011} et \cite{Verhoeff2015}.
-
-    En plus, chaque armée est différente une de l'autre et, donc, les contraints et buts qui conditionnent les opérations de maintenances sont aussi différentes.
-
-    On a formulé des nouveaux contraints, comme la possibilité de stocker temporairement des avions ou la nécessité de affecter des avions dans un temps minimal au missions.
-Ces contraints font que le modèle devient plus complexe en ajoutant une dépendance plus fort entre les mois consécutifs.
+    % Finally, there are some particular constraints that are specific for this specific problem that have not been found in other works.
+For example: the possibility to temporally stock the aircraft, the treatment of missions as a complex tasks with requirements and not as an aggregate periodic demand among others.
 
@@
 
-    % Missing: 
-    % * horizon constraints (initial, ending)
-    % * Prefer new
-    % * Total hours
-    % *
+    In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks".
 
-    \subsection{Tasks}
-    \label{def:task}
+    % \subsection{Tasks}
+    % \label{def:task}
 
     Each $v \in \mathcal{V}$ task needs to be done in a planning horizon divided into $t \in \mathcal{T}$ periods of the same length.
-Each task has a given starting ($startTIME_v \in T$) an ending ($endTIME_v \in T$) time.
-The starting and ending times are given and cannot be changed.
+Each task has a given starting an ending time that cannot be changed.
+They also require a certain quantity of assigned resources $a \in \mathcal{A}$ each period $t$, defined by $rr_{vt}$ and, if assigned, it uses them for a quantity of time in each period, defined by $rh_{vt}$.
 
-    Each task $v$ requires (seizes) a certain quantity of resources $a \in \mathcal{A}$ each period $t$ ($REQ_{vt}$). For resources, see \ref{def:resource}.
-
-    Not all $a$ resources can be used to accomplish any task: each task has specific requirements that a resource needs to satisfy in order to be a candidate.
-In other words, for each task $v$, there is a set $cd_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to that task.
-A resource can be a candidate for more than one task.
+    There is a set $cd_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to each task.
 
@@
 
-    \subsection{Resources}
-    \label{def:resource}
+    % \subsection{Resources}
+    % \label{def:resource}
 
-    Resources are what makes possible the completion of tasks.
-These specific resources suffer from wear and tear and so require a recurrent preventive maintenance (see \ref{def:maint}) in order to be in a correct state.
-These maintenances need to take place before a given amount of time (that we will call $MAX\_ELPASED$) has passed since the previous maintenance or before the resource has been used for a given amount of time (that we will call $mu$) since the previous maintenance, whichever happens first.
-While a resource $a$ has not received the proper maintenance, it cannot be used to satisfy tasks and is considered useless.
+    Resources suffer from wear and tear and so require a recurrent preventive maintenance in order to be in a correct state.
+These maintenances need to take place before the first of two things occurs: a given amount of elapsed time $me$ has passed since the previous maintenance or a given amount of usage time $mu$ has passed since the previous maintenance.
 
-    We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance at any given period.
+    % We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance at any given period.
 
-    We will call "remaining usage time" (or $rut$) to the amount of time that the resource needs to be used in order for it to need a maintenance at any given period.
+    % We will call "remaining usage time" (or $rut$) to the amount of time that the resource needs to be used in order for it to need a maintenance at any given period.
 
@@
 
-    Each resources starts the planning period with a specific status given by:
+    % Each resources starts the planning period with a specific status given by:
 
-    \begin{itemize}
-        \item remaining usage time.
-        \item remaining elapsed time.
-        % \item remaining absolute elapsed time.
-        % \item remaining absolute usage time.
-        % \item LastMaintenanceType (see \ref{def:maint}).
-    \end{itemize}
+    % \begin{itemize}
+    %     \item remaining usage time.
+    %     \item remaining elapsed time.
+    %     % \item remaining absolute elapsed time.
+    %     % \item remaining absolute usage time.
+    %     % \item LastMaintenanceType (see \ref{def:maint}).
+    % \end{itemize}
 
-    Finally, resources are organized into families or groups.
+    % Finally, resources are organized into families or groups.
 Each resource inside in a family or group shares the same types of maintenances and, usually, the same kind of tasks, among other information.
 
-    \paragraph{Resource's states}
-    \label{def:res-state}
+    % \paragraph{Resource's states}
+    % \label{def:res-state}
 
-    The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time:
+    The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time: Assigned to a task ('V'), Receiving maintenance ('M'), Available ('N').
 
-    \begin{itemize}
-        \item Assigned to a task (see \ref{def:task}).
-        \item Under maintenance (see \ref{def:maint}).
-        \item Under storage (see \ref{def:sto}).
-        \item Waiting maintenances.
-        % \item Obsolete.
-        \item Available (none of the above).
-    \end{itemize}
+    % \begin{itemize}
+    %     \item Assigned to a task (see \ref{def:task}).
+    %     \item Under maintenance (see \ref{def:maint}).
+    %     % \item Under storage (see \ref{def:sto}).
+    %     \item Waiting maintenances.
+    %     % \item Obsolete.
+    %     \item Available (none of the above).
+    % \end{itemize}
 
-    \subsection{Maintenances}
-    \label{def:maint}
+    % \subsection{Maintenances}
+    % \label{def:maint}
 
-    Maintenances are the process by which resources that have reached a limit in their usage (but not their absolute limit) can return to a state where they can continue to be used in tasks.
+    % Maintenances are the process by which resources that have reached a limit in their usage (but not their absolute limit) can return to a state where they can continue to be used in tasks.
 
-    Each maintenance belongs to a specific type of maintenance $m \in \mathcal{M}$. These types of maintenance differentiate between each other by having different characteristics.
+    % Each maintenance belongs to a specific type of maintenance $m \in \mathcal{M}$. These types of maintenance differentiate between each other by having different characteristics.
 
     These maintenances take an amount of time equal to $md_m$ periods and cannot be interrupted.
-The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance (remaining elapsed time and remaining usage time).
+The number of resources under maintenance in any given period is limited to a maximum capacity which depends on each period and is represented by $c_t$.
+    % The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance.
 
-    In other words, after receiving a maintenance, a resource restores its remaining elapsed time and remaining usage time back to their max values $me$ and $mu$ respectively.
-
-    The number of resources under maintenance in any given period is limited to a maximum capacity ($c_t$) which depends on each period.
+    % In other words, after receiving a maintenance, a resource restores its remaining elapsed time and remaining usage time back to their max values $me$ and $mu$ respectively.
 
@@
 
-    \subsection{Storage}
-    \label{def:sto}
+    % \subsection{Storage}
+    % \label{def:sto}
 
-    We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (govern by the elapsed time $elapsedTIME$).
+    % We have already explained that even if a resource is not being used, it still needs to have a maintenance after a given amount of time (govern by the elapsed time $elapsedTIME$).
 In order to avoid this problem, the resource can be put into a "storage" state.
 
-    A resource in this states has to be kept for a minimum time of $sMin$ periods. While in this state it cannot receive maintenance or be tasked any task.
+    % A resource in this states has to be kept for a minimum time of $sMin$ periods. While in this state it cannot receive maintenance or be tasked any task.
 
-    Every resource has the capacity to be stored, although it is limited.
+    % Every resource has the capacity to be stored, although it is limited.
 This capacity (measured in a number of periods $sCap$) is expended every time the resource is stored.
@@
 
-    \subsection{Objectives}
+    % \subsection{Objectives}
 
-    The are multiple objectives that need to be taken into account.
+    % The are multiple objectives that need to be taken into account.
 
-    Given that the creation of new tasks and the duration of maintenance times are considered stochastic, the basic goal is to maximize the robustness of the planning.
+    % Given that the creation of new tasks and the duration of maintenance times are considered stochastic, the basic goal is to maximize the robustness of the planning.
 
-    Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the maintenance tasks over the planning horizon.
+    % Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the maintenance tasks over the planning horizon.
 
-    \begin{enumerate}
-        \item Maximize the minimum number of available resources (see \ref{def:res-state}) in any given period.
-        \item Minimize the maximum number of resources under maintenance in any given period.
-    \end{enumerate}
+    % \begin{enumerate}
+    %     \item Maximize the minimum number of available resources (see \ref{def:res-state}) in any given period.
+    %     \item Minimize the maximum number of resources under maintenance in any given period.
+    % \end{enumerate}
+
+\section{The model}
 
@@
 
-    The decisions involve, basically, the management of these resources.
+    The decisions involve, basically, the management of the previously defined resources.
+We can summarize the main decisions the following way: if resource $a$ is in state $s$ in each period $t$ and if resource $a$ is assigned to task $v$ in period $t$.
 
-    We can summarize the main decisions into the following:
-
-    \begin{itemize}
-        \item deciding the state of each resource $a$ in each period $t$.
-        \item if a resource $a$ is tasked to some task in period $t$: decide which task $v$.
-    \end{itemize}
-
-\section{The model}
-
-    Below are the parameters used in the model.
+    \subsection{Parameters}
 
@@
     $md$             & : & maintenance duration, in periods. \\        
-    $rut\_i_a$            & : & remaining used time for resource $a$ at the beginning of horizon. \\        
-    $ret\_i_a$            & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
-    $state\_i_{at}$       & : & some $s \in \mathcal{S}$ in case the resource $a$ has a fixed state. \\
+    % $rut\_i_a$       & : & remaining used time for resource $a$ at the beginning of horizon. \\        
+    % $ret\_i_a$       & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
+    % $state\_i_{at}$  & : & some $s \in \mathcal{S}$ in case the resource $a$ has a fixed state. \\
     \end{tabular}
@@
 
-    Below are the variables used in the model.
+    \subsection{Variables}
 
-    \begin{tabular}{p{20mm}lp{100mm}}
-        $ret_{at}$      & : & remaining elapsed time for resource $a$ after period $t$ \\  
-        $rut_{at}$      & : & remaining usage time for resource $a$ after period $t$ \\  
-        $task_{avt}$    & : & 1 if resource $a \in cd_v$ is assigned task $v$ in period $t$ \\  
-        $state_{ast}$   & : & 1 if resource $a$ is assigned state $s$ in period $t$ \\  
-        $used_{at}$     & : & time consumed by resource $a$ in period $t$ \\
-        $start_{at}$    & : & 1 if resource $a$ starts maintenance in period $t$ \\
-        $available$     & : & minimum number of available resources in any period \\
-        $maintenance$   & : & maximum number of resources in maintenance in any period \\
+    \begin{tabular}{p{15mm}lp{105mm}}
+        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assigned task $v$ in period $t$, 0 otherwise. \\  
+        $state_{ast}$       & : & 1 if resource $a$ is assigned state $s$ in period $t$, 0 otherwise. \\  
+        $start_{at}$        & : & 1 if resource $a$ starts maintenance in period $t$, 0 otherwise. \\
+        $ret_{at}$          & : & remaining elapsed time (integer) for resource $a$ after period $t$. \\  
+        $rut_{at}$          & : & remaining usage time (integer) for resource $a$ after period $t$. \\  
+        $used_{at}$         & : & time (integer) used by resource $a$ in period $t$. \\
+        $min\_avail$        & : & min number (integer) of available resources in any period. \\
+        $max\_maint$        & : & max number (integer) of resources in maintenance in any period. \\
     \end{tabular}
     \bigskip
+
+    \subsection{Constraints}
+
+    The objective is to both minimize the maximum number of maintenances and maximize the minimal number of available aircraft:
+
+    \begin{align}
+        & \text{Min}\, (max\_maint) \\
+        & \text{Max}\, (min\_avail)
+    \end{align}
 
@@
 
-    These constraints are the the main constraints of the model.
+    % These constraints are the the main constraints of the model.
 
-    \ref{eq:taskres} and \ref{eq:capacity} limit the number of resources assigned to task and the number of resource maintenances that can be done in any given month.
-    \ref{eq:maxtasks} limits the number of assigned tasks for a resource to one.
+    The limit in the number of resources assigned to task and the number of resource maintenances that can be done in any given month are modeles via \ref{eq:taskres} and \ref{eq:capacity} respectively.
+    Constraint \ref{eq:maxtasks} limits the number of assigned tasks for a resource to one.
 
@@
 
-    These constraints calculate the balances of hours for each resource.
+    % These constraints calculate the balances of hours for each resource.
 
-    \ref{eq:usage} defines the consumption of hours for each resource at each period.
+    The definition of the consumption of hours for each resource at each period is forced by \ref{eq:usage}.
 This is then used in \ref{eq:balance1} to get the remaining usage time at each period for each resource.
@@
 
-    These constraints involve the states assignments to resources.
+    % These constraints involve the states assignments to resources.
 
-    \ref{eq:start1} and \ref{eq:start2} tie the start of each maintenance with the maintenance state so resources need to have started a maintenance to receive it and have to have maintenance for a minimum period after having starting it.
+    The start of each maintenance with the maintenance state is done by \ref{eq:start1} and \ref{eq:start2}.
+This way resources need to have started a maintenance to receive it and have to have maintenance for a minimum period after having starting it.
 
-    \ref{eq:state1} makes working states with being assigned to a task. \ref{eq:state2} makes resources have only one state at each period.
+    Constraint \ref{eq:state1} ties the states with the assignment to a task. Resources have only one state at each period and this is guaranteed by \ref{eq:state2}.
 
@@
         % min available
-        & available \leq \sum_{a} state_{ast}
+        & min\_avail \leq \sum_{a} state_{ast}
                 & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:available}\\
         % max maintenance
-        & maintenance \geq \sum_{a} state_{ast}
+        & max\_maint \geq \sum_{a} state_{ast}
                 & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:maintenance}
@@
 
-    These constraints define the possible objectives.
+    % These constraints define the possible objectives.
 
-    \ref{eq:available} is defined as the minimum number of available resources in any period.
+    The minimum number of available resources in any period is controlled by \ref{eq:available}.
 
-    \ref{eq:maintenance} is defined as the maximum number of resources under maintenance in any period.
+    The definition of the maximum number of resources under maintenance in any period is done by \ref{eq:maintenance}.
 
-    \begin{align}
-        %%%%%%%%%%%%%%%%%%%
-        % HORIZON
-        %%%%%%%%%%%%%%%%%%%
-        % TODO: initialize maintenance states too!
-        % rut initialize
-        & rut_{at} = rut\_ini_a
-                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
-        % ret initialize
-        & ret_{at} = ret\_ini_a
-                & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
-        % fixed states during planning period
-        & state_{ast} = 1
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}; s \in \mathcal{S}\, | if\, state\_ini_{at} = s \label{eq:fixedstates}
-    \end{align}
+    % \begin{align}
+    %     %%%%%%%%%%%%%%%%%%%
+    %     % HORIZON
+    %     %%%%%%%%%%%%%%%%%%%
+    %     % TODO: initialize maintenance states too!
+    %     % rut initialize
+    %     & rut_{at} = rut\_ini_a
+    %             & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
+    %     % ret initialize
+    %     & ret_{at} = ret\_ini_a
+    %             & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
+    %     % fixed states during planning period
+    %     & state_{ast} = 1
+    %             & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}; s \in \mathcal{S}\, | if\, state\_ini_{at} = s \label{eq:fixedstates}
+    % \end{align}
 
-    These constraints feed the relevant historic data for resources.
+    % % These constraints feed the relevant historic data for resources.
 
-    \ref{eq:remaining1} and \ref{eq:remaining2} fill the last period's remaining used and elapsed time respectively.
+    % The last period's remaining used and elapsed time are filled by \ref{eq:remaining1} and \ref{eq:remaining2} respectively.
 
-    \ref{eq:fixedstates} fixes the states to their planned state.
+    % In order to fix the states to their planned state we need \ref{eq:fixedstates}.
 This is specially useful for the beginning of the planning period, where we could have maintenance states already decided for some resources.
 
-\section{Conclusions et perspectives}
+% \section{Conclusions et perspectives}
 
-    TODO.
+%     TODO.
 

### 2017-11-21 latex/ROADEF/article_roadef_2018.tex
changes to latex document

@@
 
+    The problem consists in the planning of the maintenance operations of military aircraft while complying with scheduled missions.
 In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks".
@@
 
-    The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time: Assigned to a task ('V'), Receiving maintenance ('M'), Available ('N').
+    The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time: Assigned to a task ('V'), Receiving maintenance ('M'), Available ('A'), Needing maintenance ('N').
 
@@
 
-    These maintenances take an amount of time equal to $md_m$ periods and cannot be interrupted.
+    These maintenances take an amount of time equal to $md$ periods and cannot be interrupted.
 The number of resources under maintenance in any given period is limited to a maximum capacity which depends on each period and is represented by $c_t$.
@@
 
-    \subsection{Decisions}
+    % \subsection{Decisions}
 
-    The decisions involve, basically, the management of the previously defined resources.
-We can summarize the main decisions the following way: if resource $a$ is in state $s$ in each period $t$ and if resource $a$ is assigned to task $v$ in period $t$.
+    % \subsection{Parameters}
 
-    \subsection{Parameters}
-
-    \begin{tabular}{p{10mm}lp{110mm}}
-    $c_t$            & : & number of resources that can be in maintenance in period $t$. \\
-    $cd_v$           & : & resources that satisfy task's $v$ needs. \\
-    $rh_{vt}$        & : & time consumed of any resource when assigned to task $v$ in period $t$. \\
-    $rr_{vt}$        & : & number or resources to assign to task $v$ in period $t$. \\            
-    $mu$             & : & maximum used time for each resource between maintenances. \\    
-    $me$             & : & maximum elapsed time for each resource between maintenances. \\    
-    $md$             & : & maintenance duration, in periods. \\        
+    % \begin{tabular}{p{10mm}lp{110mm}}
+    % $c_t$            & : & number of resources that can be in maintenance in period $t$. \\
+    % $cd_v$           & : & resources that satisfy task's $v$ needs. \\
+    % $rh_{vt}$        & : & time consumed of any resource when assigned to task $v$ in period $t$. \\
+    % $rr_{vt}$        & : & number or resources to assign to task $v$ in period $t$. \\            
+    % $mu$             & : & maximum used time for each resource between maintenances. \\    
+    % $me$             & : & maximum elapsed time for each resource between maintenances. \\    
+    % $md$             & : & maintenance duration, in periods. \\        
     % $rut\_i_a$       & : & remaining used time for resource $a$ at the beginning of horizon. \\        
@@
     % $state\_i_{at}$  & : & some $s \in \mathcal{S}$ in case the resource $a$ has a fixed state. \\
-    \end{tabular}
-    \bigskip
+    % \end{tabular}
+    % \bigskip
 
@@
 
+    The decisions involve, in essence, the management of the previously defined resources.
+We can summarize the main decisions the following way: if resource $a$ is in state $s$ in each period $t$ and if resource $a$ is assigned to task $v$ in period $t$.
+
     \begin{tabular}{p{15mm}lp{105mm}}
-        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assigned task $v$ in period $t$, 0 otherwise. \\  
+        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assignigned task $v$ in period $t$, 0 otherwise. \\  
         $state_{ast}$       & : & 1 if resource $a$ is assigned state $s$ in period $t$, 0 otherwise. \\  
@@
         $ret_{at}$          & : & remaining elapsed time (integer) for resource $a$ after period $t$. \\  
-        $rut_{at}$          & : & remaining usage time (integer) for resource $a$ after period $t$. \\  
+        $rut_{at}$          & : & remaining usage time (integer) for remainingsource $a$ after period $t$. \\  
         $used_{at}$         & : & time (integer) used by resource $a$ in period $t$. \\
@@
     \end{tabular}
-    \bigskip
+    % \bigskip
 
@@
         & \sum_{a \in \mathcal{A}} state_{ast} \leq c_{t}
-                & t \in \mathcal{T}, s = 'M' \label{eq:capacity}\\
+                & t \in \mathcal{T}, s = \mq{M} \label{eq:capacity}\\
         % minimum resources
@@
         & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
-                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}
+                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
+        % if assigned to task: in working state
+        & state_{ast} \geq task_{avt} 
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, v \in \mathcal{V}, s = \mq{V} \label{eq:state1}\\
+        % only one state. If we take out the 'do nothing' state, this could be \leq
+        & \sum_{s \in \mathcal{S}} state_{ast} = 1 
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}
     \end{align}
@@
 
-    The limit in the number of resources assigned to task and the number of resource maintenances that can be done in any given month are modeles via \ref{eq:taskres} and \ref{eq:capacity} respectively.
+    The limit in the number of resources assigned to task and the number of resources under maintenance that can be done in any given month are modeled via \ref{eq:taskres} and \ref{eq:capacity} respectively.
     Constraint \ref{eq:maxtasks} limits the number of assigned tasks for a resource to one.
+
+    Constraint \ref{eq:state1} ties the states with the assignment to a task. Resources have only one state at each period and this is guaranteed by \ref{eq:state2}.
 
@@
         & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} 
-                & \hspace{10mm} (t_1) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + md - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start2}\\
-        % if assigned to task: in working state
-        & state_{ast} \geq task_{avt} 
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{V} \label{eq:state1}\\
-        % only one state. If we take out the 'do nothing' state, this could be \leq
-        & \sum_{s \in \mathcal{S}} state_{ast} = 1 
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}
+                & \hspace{10mm} t_2 \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + md - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start2}
     \end{align}
@@
 
-    Constraint \ref{eq:state1} ties the states with the assignment to a task. Resources have only one state at each period and this is guaranteed by \ref{eq:state2}.
-
     \begin{align}
@@
         & min\_avail \leq \sum_{a} state_{ast}
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:available}\\
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{A} \label{eq:available}\\
         % max maintenance
         & max\_maint \geq \sum_{a} state_{ast}
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{N} \label{eq:maintenance}
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{M} \label{eq:maintenance}
     \end{align}
@@
 
-% \section{Conclusions et perspectives}
+\section{Conclusions et perspectives}
 
-%     TODO.
+    The present modelization is a new way to formulate the Military Flight Planning Model.
+Unlike previous work, it assigns aircraft to missions, instead of just using an aggregate demand.
+This is useful for later improving the model by including minimum durations for mission assignment and other more complex constraints.
 

### 2017-11-22 latex/ROADEF/article_roadef_2018.tex
modifications to resume document

@@
 
-% Les affiliations (par ordre croissant des numéros d'affiliation) séparées par \and
+% Les affiliations (par ordre croissant des num?os d'affiliation) s?ar?s par \and
 \institute{
@@
 \section{Introduction}
-
-%  TODO!: re-arrange all this better.
 
@@
 Each task has a given starting an ending time that cannot be changed.
-They also require a certain quantity of assigned resources $a \in \mathcal{A}$ each period $t$, defined by $rr_{vt}$ and, if assigned, it uses them for a quantity of time in each period, defined by $rh_{vt}$.
+They also require a certain quantity of assigned resources $a \in \mathcal{A}$ each period $t$, defined by $rr_{v}$ and, if assigned, it uses them for a quantity of time in each period, defined by $rh_{v}$.
 
@@
         % minimum resources
-        & \sum_{a \in \mathcal{A}} task_{avt} \geq rr_{vt}
-                & t \in \mathcal{T}, v \in \mathcal{V} \label{eq:taskres}\\
+        & \sum_{a \in \mathcal{A}} task_{avt} \geq rr_{t}
+                & t \in \mathcal{T}, v \in \mathcal{V},  \label{eq:taskres}\\
         % max one task per period
@@
                 & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, v \in \mathcal{V}, s = \mq{V} \label{eq:state1}\\
+        % if no working state, no assigned to any task
+        & state_{ast} + task_{avt} \leq 1
+                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, v \in \mathcal{V}, s \neq \mq{V} \label{eq:state3}\\
         % only one state. If we take out the 'do nothing' state, this could be \leq
@@
 
-    Constraint \ref{eq:state1} ties the states with the assignment to a task. Resources have only one state at each period and this is guaranteed by \ref{eq:state2}.
+    Constraint \ref{eq:state1} and \ref{eq:state3} tie states with the assignment to a task. Resources have only one state at each period and this is guaranteed by \ref{eq:state2}.
 
@@
         % definition of consumed resource hours:
-        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times rh_{vt}
+        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times rh_{v}
                 & a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
@@
 \bibliographystyle{plain}
+\selectlanguage{french}
 \bibliography{MFMP}

### 2017-11-24 latex/ROADEF/article_roadef_2018.tex
Finally been able to produce R graphs from python based on the solutions of the pulp model.

@@
 
-% Les affiliations (par ordre croissant des num?os d'affiliation) s?ar?s par \and
 \institute{
@@
 
-    \subsection{Variables}
+    % \subsection{Variables}
 
@@
         $ret_{at}$          & : & remaining elapsed time (integer) for resource $a$ after period $t$. \\  
-        $rut_{at}$          & : & remaining usage time (integer) for remainingsource $a$ after period $t$. \\  
+        $rut_{at}$          & : & remaining usage time (integer) for resource $a$ after period $t$. \\  
         $used_{at}$         & : & time (integer) used by resource $a$ in period $t$. \\
@@
     \end{tabular}
-    % \bigskip
+    \bigskip
 
-    \subsection{Constraints}
+    % \subsection{Constraints}
 
-    The objective is to both minimize the maximum number of maintenances and maximize the minimal number of available aircraft:
+    The objective is to both minimize the maximum number of maintenances $\text{Min}\, (max\_maint)$ and maximize the minimal number of available aircraft \text{Max}\, (min\_avail).
 
-    \begin{align}
-        & \text{Min}\, (max\_maint) \\
-        & \text{Max}\, (min\_avail)
-    \end{align}
+    % \begin{align}
+    %     & \text{Min}\, (max\_maint) \\
+    %     & \text{Max}\, (min\_avail)
+    % \end{align}
 
@@
 
-\section{Conclusions et perspectives}
+\section{Conclusions}
 
     The present modelization is a new way to formulate the Military Flight Planning Model.
-Unlike previous work, it assigns aircraft to missions, instead of just using an aggregate demand.
-This is useful for later improving the model by including minimum durations for mission assignment and other more complex constraints.
+Unlike previous work, it assigns aircraft to missions instead of just using an aggregate hourly demand per period.
+This is useful in order to improve the model by including minimum durations for mission assignment and other more complex constraints.
 

### 2017-11-24 R/scripts/import_results.R

@@ -8,6 +8,7 @@ library(lubridate)
 library(timevis)
 library(Hmisc)
 library(RColorBrewer)
+library(htmlwidgets)
 
 collapse_states <- function(table, column_name){
     
@@ -65,7 +66,8 @@ if (FALSE){
         select(id, start= month, end= post_month, content = bucket, group= UNIT, style) %>% 
         slice(1:1000)
     
-    groups <- data.table(id= states %>% distinct(group) %>% unlist, content= states %>% distinct(group) %>% unlist)
+    groups_c <- states %>% distinct(group) %>% unlist
+    groups <- data.table(id= groups_c, content= groups_c)
     
     config <- list(
         stack = FALSE,
@@ -76,5 +78,7 @@ if (FALSE){
         margin = 0
     )
     
-    timevis(states, groups= groups, options= config)
+    result <- timevis(states, groups= groups, options= config)
 }
+
+saveWidget(result, file = "/home/pchtsp/Downloads/test.html")

### 2017-11-24 python/scripts/graphs.py

@@ -5,6 +5,8 @@ import rpy2.robjects as ro
 from rpy2.robjects.packages import importr
 # import pandas.rpy.common as com
 from rpy2.robjects import pandas2ri
+import palettable.colorbrewer as cbrewer
+
 # import rpy2.robjects.lib.timevis as timevis
 # we start with some pickle
 
@@ -33,31 +35,63 @@ table_n = \
     join(table['task']).\
     join(table['used'])
 
-table_n.state = np.where(table_n.state != 'V',
+table_n['content'] = np.where(table_n.state != 'V',
                          table_n.state,
                          table_n['task'] + ' (' + table_n['used'].map(str) + ')')
 
-# table_nn = table_n[table_n.state != 'A'].state.reset_index()
-table_nn = table_n.state.reset_index()
-    # .\
-    # rename(columns={'UNIT':'id', 'state': 'content'})
+table_nn = table_n[['content', 'task', 'state']].reset_index()
+last_period = table_nn.month.values.max()
 
 table_nn.sort_values(indeces, inplace=True)
-table_nn['prev'] = table_nn.groupby('UNIT').state.shift(1)
+table_nn['prev'] = table_nn.groupby('UNIT').content.shift(1)
 table_nn['nrow'] = table_nn.groupby('UNIT').cumcount() + 1
 
-table_nn[np.logical_or(table_nn.prev==table_nn.state, table_nn['nrow']==1)]
+table_nn = table_nn[np.logical_or(table_nn.prev!=table_nn.content, table_nn['nrow']==1)]
 table_nn['end'] = table_nn.groupby('UNIT').month.shift(-1)
 
-ro.r('''
-       library(timevis)
-''')
+table_nn.end = np.where(pd.isna(table_nn.end), last_period, table_nn.end)
+
+colors = {'A': "white", 'M': '#FFFFB2', 'V': "#BD0026"}
+
+table_nn['color'] = table_nn.state.replace(colors)
+table_nn['style'] = table_nn.color.map("background-color:{0};border-color:{0}".format)
+
+table_nn = table_nn.reset_index().reset_index().\
+    rename(columns={'level_0': 'id', 'UNIT': 'group', 'month': 'start'})\
+    [['id', 'start', 'end', 'content', 'group', 'style']]
+
+groups = pd.DataFrame({'id': table_nn.group.unique(), 'content': table_nn.group.unique()})
+
[…]

### 2017-11-27 latex/ROADEF/article_roadef_2018.tex
Edited latex summary to almost 2 pages

@@
 
-    Although the former has been studied in literature in \cite{Cho2011}, \cite{Kozanidis2008}, \cite{Verhoeff2015} and \cite{Chastellux2017}, it has not received as much attention as the latter.
+    Although the former has been studied in literature in \cite{Cho2011}, \cite{Chastellux2017}, \cite{Kozanidis2008}, it has not received as much attention as the latter.
 
@@
 
-    The problem consists in the planning of the maintenance operations of military aircraft while complying with scheduled missions.
+    The problem consists in the planning of the maintenance operations for military aircraft while complying with already scheduled missions.
 In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks".
@@
 
-    Each $v \in \mathcal{V}$ task needs to be done in a planning horizon divided into $t \in \mathcal{T}$ periods of the same length.
-Each task has a given starting an ending time that cannot be changed.
-They also require a certain quantity of assigned resources $a \in \mathcal{A}$ each period $t$, defined by $rr_{v}$ and, if assigned, it uses them for a quantity of time in each period, defined by $rh_{v}$.
-
+    A series of $v \in \mathcal{V}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
+Each task require a certain number $rr_{v}$ of resources $a \in \mathcal{A}$ and, while assigned, it wears them for a quantity of time defined by $rh_{v}$ in each period.
 There is a set $cd_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to each task.
@@
 These maintenances need to take place before the first of two things occurs: a given amount of elapsed time $me$ has passed since the previous maintenance or a given amount of usage time $mu$ has passed since the previous maintenance.
+The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time: Assigned to a task ('V'), Receiving maintenance ('M'), Available ('A'), Needing maintenance ('N').
+These maintenances take an amount of time equal to $md$ periods and cannot be interrupted.
+The number of resources under maintenance in any given period is limited by $c_t$.
 
@@
 
-    The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time: Assigned to a task ('V'), Receiving maintenance ('M'), Available ('A'), Needing maintenance ('N').
-
     % \begin{itemize}
@@
 
-    These maintenances take an amount of time equal to $md$ periods and cannot be interrupted.
-The number of resources under maintenance in any given period is limited to a maximum capacity which depends on each period and is represented by $c_t$.
     % The state of the resource after exiting the maintenance period is of "as good as new" in terms of the two indicators that decide the maintenance.
@@
     \begin{tabular}{p{15mm}lp{105mm}}
-        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assignigned task $v$ in period $t$, 0 otherwise. \\  
+        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assigned task $v$ in period $t$, 0 otherwise. \\  
         $state_{ast}$       & : & 1 if resource $a$ is assigned state $s$ in period $t$, 0 otherwise. \\  
@@
 
-    The objective is to both minimize the maximum number of maintenances $\text{Min}\, (max\_maint)$ and maximize the minimal number of available aircraft \text{Max}\, (min\_avail).
+    The objective is to both minimize the maximum number of maintenances $\text{Min}\, (max\_maint)$ and maximize the minimal number of available aircraft \text{Max}\, (min\_avail) subject to the constraints listed below.
 
@@
         %%%%%%%%%%%%%%%%%%%
-        % MAIN VARS
+        % MAIN CONSTRAINTS
         %%%%%%%%%%%%%%%%%%%
@@
         % minimum resources
-        & \sum_{a \in \mathcal{A}} task_{avt} \geq rr_{t}
+        & \sum_{a \in cd_v} task_{avt} \geq rr_{t}
                 & t \in \mathcal{T}, v \in \mathcal{V},  \label{eq:taskres}\\
@@
         & state_{ast} \geq task_{avt} 
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, v \in \mathcal{V}, s = \mq{V} \label{eq:state1}\\
+                & \hspace{10mm} t \in \mathcal{T}, a \in cd_v, v \in \mathcal{V}, s = \mq{V} \label{eq:state1}\\
         % if no working state, no assigned to any task
         & state_{ast} + task_{avt} \leq 1
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, v \in \mathcal{V}, s \neq \mq{V} \label{eq:state3}\\
+                & \hspace{10mm} t \in \mathcal{T}, a \in cd_v, v \in \mathcal{V}, s \neq \mq{V} \label{eq:state3}\\
         % only one state. If we take out the 'do nothing' state, this could be \leq
@@
     % These constraints are the the main constraints of the model.
-
-    The limit in the number of resources assigned to task and the number of resources under maintenance that can be done in any given month are modeled via \ref{eq:taskres} and \ref{eq:capacity} respectively.
-    Constraint \ref{eq:maxtasks} limits the number of assigned tasks for a resource to one.
-
-    Constraint \ref{eq:state1} and \ref{eq:state3} tie states with the assignment to a task. Resources have only one state at each period and this is guaranteed by \ref{eq:state2}.
+    % The number of resources assigned to a single task and under maintenance 
+    Tasks' resource requirements are defined by \ref{eq:taskres}, maintenance capacity is controlled by \ref{eq:capacity}.
+Constraints \ref{eq:state1} and \ref{eq:state3} tie states with the assignment to a task.
+Maximum tasks and states assigned to each resource is guaranteed by \ref{eq:maxtasks} and \ref{eq:state2}.
 
@@
     The definition of the consumption of hours for each resource at each period is forced by \ref{eq:usage}.
-This is then used in \ref{eq:balance1} to get the remaining usage time at each period for each resource.
+The monthly remaining usage and elapsed time are calculated by \ref{eq:balance1} and \ref{eq:balance2} respectively.
+     % This is then applied in \ref{eq:balance1} to get the remaining usage time at each period for each resource.
 Similarly, the \ref{eq:balance2} calculates the remaining elapsed time at each period and resource.
@@
 
-    The start of each maintenance with the maintenance state is done by \ref{eq:start1} and \ref{eq:start2}.
-This way resources need to have started a maintenance to receive it and have to have maintenance for a minimum period after having starting it.
+    The coupling between the start of each maintenance and the maintenance state is done by \ref{eq:start1} and \ref{eq:start2}.
+    % This way resources need to have started a maintenance to receive it and have to have maintenance for a minimum period after having starting it.
 
@@
     % These constraints define the possible objectives.
+    The definitions of the two previously defined objectives is done by \ref{eq:available} and \ref{eq:maintenance}.
 
-    The minimum number of available resources in any period is controlled by \ref{eq:available}.
-
+    % The minimum number of available resources in any period is controlled by \ref{eq:available}.
 The definition of the maximum number of resources under maintenance in any period is done by \ref{eq:maintenance}.
@@
 Unlike previous work, it assigns aircraft to missions instead of just using an aggregate hourly demand per period.
-This is useful in order to improve the model by including minimum durations for mission assignment and other more complex constraints.
+This is useful in order to improve the model by including minimum durations for mission assignment.
 

### 2017-11-27 R/scripts/import_results.R

@@ -81,4 +81,4 @@ if (FALSE){
     result <- timevis(states, groups= groups, options= config)
 }
 
-saveWidget(result, file = "/home/pchtsp/Downloads/test.html")
+saveWidget(result, file = "/home/pchtsp/Downloads/test.html", selfcontained = FALSE)

### 2017-11-27 latex/ROADEF/article_roadef_2018.tex
experiments with model and changes after Olga's revision

@@
 \section{Introduction}
-
-    The Military Flight and Maintenance Planning is a problem consisting on the assignment of missions and maintenances tasks to military aircraft.
+    The Military Flight and Maintenance Planning is a problem consisting on the assignment of missions and maintenance tasks to military aircraft.
 It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically but they are assumed to include a round-trip to the base.
 It also includes different objectives and constraints.
-
-    Although the former has been studied in literature in \cite{Cho2011}, \cite{Chastellux2017}, \cite{Kozanidis2008}, it has not received as much attention as the latter.
+Although the former has been studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008}, it has not received as much attention as the latter.
 
@@
 
-\section{The problem}
+%\section{The problem}
 
-    The problem consists in the planning of the maintenance operations for military aircraft while complying with already scheduled missions.
+    The problem consists in the planning of the maintenance operations for military aircraft and the assignment of these aircraft to already scheduled missions.
 In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks".
-
-    % \subsection{Tasks}
-    % \label{def:task}
-
 A series of $v \in \mathcal{V}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
-Each task require a certain number $rr_{v}$ of resources $a \in \mathcal{A}$ and, while assigned, it wears them for a quantity of time defined by $rh_{v}$ in each period.
-There is a set $cd_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to each task.
+Each task requires a certain number $rr_{v}$ of resources $a \in \mathcal{A}$ which it employs for a time duration defined by $rh_{v}$ in each period.
+Set $cd_{v} \subset \mathcal{A}$ lists the resources that can be assigned to each task.
 
@@
 
-    Resources suffer from wear and tear and so require a recurrent preventive maintenance in order to be in a correct state.
-These maintenances need to take place before the first of two things occurs: a given amount of elapsed time $me$ has passed since the previous maintenance or a given amount of usage time $mu$ has passed since the previous maintenance.
-The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time: Assigned to a task ('V'), Receiving maintenance ('M'), Available ('A'), Needing maintenance ('N').
-These maintenances take an amount of time equal to $md$ periods and cannot be interrupted.
-The number of resources under maintenance in any given period is limited by $c_t$.
-
+    Resources require recurrent preventive maintenance operations.
+The frequency of maintenance operations is defined by 2 factors: the amount of time $me$ elapsed since the last maintenance operation and the amount of time $mu$ for which the resource has been used since the last maintenance operation.
+The following states $s \in \mathcal{S}$ are possible for a resource: assigned to a task ('V'), receiving maintenance ('M'), available ('A'), needing maintenance ('N').
+A maintenance operation takes an amount of time equal to $md$ periods and cannot be interrupted.
+The number of resources under maintenance in any period is restricted by its capacity $c_t$.
+In the following we present a new mathematical model for this problem.
+Unlike previous work, it assigns aircraft to missions, instead of just using an aggregate demand.
     % We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance at any given period.
@@
 
-\section{The model}
+\section{Mathematical model}
 
@@
 
-    The decisions involve, in essence, the management of the previously defined resources.
+    The decision variables manage the previously defined resources.
 We can summarize the main decisions the following way: if resource $a$ is in state $s$ in each period $t$ and if resource $a$ is assigned to task $v$ in period $t$.
@@
     \begin{tabular}{p{15mm}lp{105mm}}
-        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assigned task $v$ in period $t$, 0 otherwise. \\  
-        $state_{ast}$       & : & 1 if resource $a$ is assigned state $s$ in period $t$, 0 otherwise. \\  
+        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assigned to task $v$ in period $t$, 0 otherwise. \\  
+        $state_{ast}$       & : & 1 if resource $a$ is in state $s$ in period $t$, 0 otherwise. \\  
         $start_{at}$        & : & 1 if resource $a$ starts maintenance in period $t$, 0 otherwise. \\
@@
 
-    The objective is to both minimize the maximum number of maintenances $\text{Min}\, (max\_maint)$ and maximize the minimal number of available aircraft \text{Max}\, (min\_avail) subject to the constraints listed below.
+    The objective is to both minimize the maximum number of maintenances and maximize the minimal number of available aircraft:
 
-    % \begin{align}
-    %     & \text{Min}\, (max\_maint) \\
-    %     & \text{Max}\, (min\_avail)
-    % \end{align}
+    \begin{align}
+        & \text{Min}\, (max\_maint) \\
+        & \text{Max}\, (min\_avail)
+    \end{align}
+
+    Subject to:
 
@@
         % minimum resources
-        & \sum_{a \in cd_v} task_{avt} \geq rr_{t}
+        & \sum_{a \in cd_v} task_{avt} \geq rr_{v}
                 & t \in \mathcal{T}, v \in \mathcal{V},  \label{eq:taskres}\\
@@
 
-\section{Conclusions}
+%\section{Conclusions}
 
-    The present modelization is a new way to formulate the Military Flight Planning Model.
+ %   The present modelization is a new way to formulate the Military Flight Planning Model.
 Unlike previous work, it assigns aircraft to missions instead of just using an aggregate hourly demand per period.

### 2017-11-27 python/package/model.py

@@ -152,12 +152,17 @@ ub = {
 
 vt = [(v, t) for v in tasks for t in periods if start_time[v] <= t <= end_time[v]]
 avt = [(a, v, t) for a in resources for (v, t) in vt if a in candidates[v]]
-at = [(a, t) for a in resources for t in periods]  # if periods_pos[t] % 2 == 0?
+at = [(a, t) for a in resources for t in periods]
 ast = [(a, s, t) for (a, t) in at for s in states]
+
+# this is tricky: we  limit the possibilities of starting a maintenance:
+at_start = [(a, t) for a, t in at if periods_pos[t] % 2 == 0]
+at_start_not = [tup for tup in at if tup not in at_start]
 
 at0 = [(a, t) for a in resources for t in periods_0]
 att = [(a, t1, t2) for (a, t1) in at for t2 in periods if
-       periods_pos[t1] <= periods_pos[t2] <= periods_pos[t1] + duration - 1]
+       periods_pos[t1] <= periods_pos[t2] <= periods_pos[t1] + duration - 1 and \
+       (a, t1) in at_start]
 
 # a_t = {t: a for t in periods for a in resources if (a, t) in at}
 a_t = tup_to_dict(at, result_col=0, is_list=True)
@@ -169,7 +174,7 @@ t1_at2 = tup_to_dict(att, result_col=1, is_list=True)
 
 # binary:
 task = pl.LpVariable.dicts("task", avt, 0, 1, pl.LpInteger)
-start = pl.LpVariable.dicts("start", at, 0, 1, pl.LpInteger)
+start = pl.LpVariable.dicts("start", at_start, 0, 1, pl.LpInteger)
 state = pl.LpVariable.dicts("state", ast, 0, 1, pl.LpInteger)
 
 # numeric:
@@ -214,13 +219,23 @@ for (a, v, t) in avt:
 # remaining used time calculations:
 # remaining elapsed time calculations:
 # *maybe* reformulate this
-for (a, t) in v_at:
-    if t != periods[0]:
+# We only increase the remainders if in that month we could start a maintenance
+for (a, t) in at_start:
+    if t != first_period:
         model += rut[(a, t)] <= rut[(a, previous[t])] - used[(a, t)] + max_used_time * start[(a, t)]
         model += ret[(a, t)] <= ret[(a, previous[t])] - 1 + max_elapsed_time * start[(a, t)]
     else:
[…]

### 2017-12-18 R/dga_diaporama/dga.Rmd
add presentation files for DGA meeting.

---
title: "Plannification de maintenances des aéronefs militaires"
author: "Franco Peschiera"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
bibliography: MFMP.bib
output: 
  slidy_presentation: 
    incremental: yes
    css: "./../../css/baobab.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Qui sui je

* Péruvien (et français).
* Ingénieur Industriel par l'Universidad Politecnica de Madrid.
* Passionné par les mathématiques appliqués et l'informatique.
* Fondé et travaillé dans une entreprise de technologie à Madrid pendant les 7 derniers années.

# Où et avec qui je travaille

* A Toulouse dans l'ISAE-SUPAERO.
* Plus spécifiquement dans le Département d'Ingénierie des Systèmes Complexes (DISC).
* Mes directeurs de thèse sont: Alain Haït, Olga Battaïa, Nicolas Dupin.

# Mon sujet de thèse: MFMP

> Military Flight and Maintenance Planning.

* Fait partie de la recherche opérationnelle: mathématiques dédiés à résoudre des gros et complexes problèmes de l'industrie.
* Un problème d'affectation des missions et planification des maintenances aux aéronefs militaires.
* Similaire au problème FMP: Flight and Maintenance Planning .
* Résolu par techniques de modélisation comme PLNE, méta-heuristiques ou heuristiques.

# Example

```{r echo=FALSE}
htmltools::includeHTML("calendrier_20171116.html")
```

# Etat de l'art

* En @Cho2011, des affectations quotidiens des heures aux avions militaires.
* En @Verhoeff2015 et @Kozanidis2008, des affectations mensuels des maintenances et opérations sont faits avec contraintes des disponibilité, niveau de service et soutenabilité.
* En @Chastellux2017, un premier modèle d'affectation de missions aux aéronefs en lieu des heures est présenté.

# A quoi ça sert

* En faisant meilleures planifications, on fait une meilleur utilisation de ressources.
* Lisser les maintenances pendant tout le horizon future.
* garantir la disponibilité des aéronefs pour possibles nouvelles missions.

# Bibliographie

### 2017-12-18 R/dga_diaporama/dga.Rmd
Added graphing functions to solution. Added changes to dga presentation.

@@
 
-# Qui sui je
+# Qui suis-je
 
@@
 ```{r echo=FALSE}
-htmltools::includeHTML("calendrier_20171116.html")
+htmltools::includeHTML("./../../img/calendrier_20171116.html")
 ```
@@
 
-* En faisant meilleures planifications, on fait une meilleur utilisation de ressources.
-* Lisser les maintenances pendant tout le horizon future.
+* En faisant meilleures planifications, on fait une meilleur utilisation des ressources.
+* Lisser les maintenances pendant tout le horizon future fait possible 
 * garantir la disponibilité des aéronefs pour possibles nouvelles missions.

### 2017-12-18 python/package/aux.py

@@ -163,17 +163,9 @@ def get_property_from_dic(dic, property):
 def get_timestamp():
     return datetime.datetime.now().strftime("%Y%m%d%H%M")
 
-# TODO: this function
-def clean_exp(path, clean=True):
-    """
-    loads and cleans all experiments that are incomplete
-    :param path: path to experiments
-    :param clean: if set to false it returns the experiments to clean instead of delete them
-    :return: deleted experiments
-    """
-    return 0
 
 if __name__ == "__main__":
+
     content = tup_to_dicts({1: {2: {4: 5}}}, (1, 2, 3), 1)
     result = {}
     result = dicts_to_tup(result, [], {1: {2: {4: 5}}})

### 2017-12-18 python/package/solution.py

@@ -1,6 +1,10 @@
 # /usr/bin/python3
 
 import package.aux as aux
+import pandas as pd
+import rpy2.robjects as ro
+from rpy2.robjects.packages import importr
+from rpy2.robjects import pandas2ri
 
 
 class Solution(object):
@@ -38,3 +42,48 @@ class Solution(object):
         statesMissions = aux.dict_to_tup(self.get_state()) + aux.dict_to_tup(self.get_tasks())
         result = aux.tup_to_start_finish(statesMissions)
         return result
+
+    def print_solution(self, path):
+        cols = ['group', 'start', 'content', 'end']
+        table = pd.DataFrame(self.get_schedule(),
+                             columns=cols
+                             )
+        table.end = table.end.apply(lambda x: aux.get_next_month(x))
+        table['status'] = table.content.str.replace(r"\d+", "")
+        colors = {'A': "white", 'M': '#FFFFB2', 'O': "#BD0026"}
+        table['style'] = \
+            table.status.replace(colors). \
+                map("background-color:{0};border-color:{0}".format)
+        table = table.sort_values("group").reset_index().rename(columns={'index': "id"})
+        cols2 = cols + ['style', 'id']
+        table[cols2] = table[cols2].astype(str)
+        # table = table.astype(
+        #     {'content': 'U5', 'id': 'int64', 'group': 'U5', 'start': 'U7', 'end': 'U7', 'status': 'U5', 'style': 'U40'})
+        groups = pd.DataFrame({'id': table.group.unique(), 'content': table.group.unique()})
+
+        timevis = importr('timevis')
+        htmlwidgets = importr('htmlwidgets')
+        rdf = pandas2ri.py2ri(table)
+        rdfgroups = pandas2ri.py2ri(groups)
+
+        options = ro.ListVector({
+            "stack": False,
+            "editable": True,
+            "align": "center",
+            "orientation": "top",
+            # "snap": None,
+            "margin": 0
+        })
+
+        graph = timevis.timevis(rdf, groups=rdfgroups, options=options, width="100%")
+        htmlwidgets.saveWidget(graph, file=path, selfcontained=False)
+        return
+        # print(graph)
+
+
+if __name__ == "__main__":
+    path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201712121208/"
+    path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201712142345/"
[…]

### 2017-12-18 python/package/tests.py

@@ -5,7 +5,8 @@ import pandas as pd
 import package.data_input as di
 import package.solution as sol
 import package.instance as inst
-
+import os
+import shutil
 
 """
 
@@ -33,9 +34,12 @@ class Experiment(object):
         self.solution = solution
 
     @classmethod
-    def from_dir(cls, path, format='json'):
-        instance = di.load_data(path + "data_in." + format)
-        solution = di.load_data(path + "data_out." + format)
+    def from_dir(cls, path, format='json', prefix="data_"):
+        files = [os.path.join(path, prefix + f + "." + format) for f in ['in', 'out']]
+        if not np.all([os.path.exists(f) for f in files]):
+            return None
+        instance = di.load_data(files[0])
+        solution = di.load_data(files[1])
         return cls(inst.Instance(instance), sol.Solution(solution))
 
     def check_solution(self):
@@ -139,6 +143,30 @@ class Experiment(object):
         return maintenance_duration_incorrect
 
 
+def clean_experiments(path, clean=True):
+    """
+    loads and cleans all experiments that are incomplete
+    :param path: path to experiments
+    :param clean: if set to false it only does not delete them
+    :return: deleted experiments
+    """
+    exps_paths = [os.path.join(path, f) for f in os.listdir(path) if os.path.isdir(os.path.join(path, f))]
+    to_delete = []
+    for e in exps_paths:
+        exp = Experiment.from_dir(e, format="json")
+        if exp is None:
+            exp = Experiment.from_dir(e, format="pickle")
+        to_delete.append(exp is None)
+    exps_to_delete = np.array(exps_paths)[to_delete]
+    if clean:
+        for ed in exps_to_delete:
+            shutil.rmtree(ed)
+    return exps_to_delete
+    # np.array(exps_paths).__len__()
+    #
+    # return 0
+
+
 if __name__ == "__main__":
     path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201712121208/"
     path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201712142345/"
@@ -148,20 +176,24 @@ if __name__ == "__main__":
     exp = Experiment.from_dir(path)
     results = exp.check_solution()
 
-    [k for (k, v) in results.items() if len(v) > 0]
[…]

### 2017-12-18 python/scripts/graphs.py

@@ -3,16 +3,16 @@ import pandas as pd
 import numpy as np
 import rpy2.robjects as ro
 from rpy2.robjects.packages import importr
-# import pandas.rpy.common as com
 from rpy2.robjects import pandas2ri
+# import pandas.rpy.common as com
 import palettable.colorbrewer as cbrewer
+import package.tests as tests
 
 # import rpy2.robjects.lib.timevis as timevis
 # we start with some pickle
 
-directory_path = '/home/pchtsp/Documents/projects/OPTIMA/python/experiments/201711231500/'
-
-solution = aux.load_data(directory_path + "data_out.pickle")
+# directory_path = '/home/pchtsp/Documents/projects/OPTIMA/python/experiments/201711231500/'
+directory_path = '/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201712181704'
 
 # to make a graph, we need to have something like the following:
 
@@ -88,7 +88,7 @@ graph = timevis.timevis(rdf, groups= rdfgroups, options = options, width="100%")
 
 print(graph)
 
-htmlwidgets.saveWidget(graph, file= "/home/pchtsp/Downloads/calendar_20171124.html", selfcontained = False)
+htmlwidgets.saveWidget(graph, file= "/home/pchtsp/Downloads/calendar_20171124.html", selfcontained=False)
 
 # pandas2ri.ri2py()
 

### 2017-12-19 R/dga_diaporama/dga.Rmd
New version of dga presentation

--- 2017-12-18 R/dga_diaporama/dga.Rmd
@@
 bibliography: MFMP.bib
+nocite: "@*"
 output: 
@@
 
-# Qui sui je
+# Qui suis-je
 
-* Péruvien (et français).
+* Péruvien et français.
 * Ingénieur Industriel par l'Universidad Politecnica de Madrid.
 * Passionné par les mathématiques appliqués et l'informatique.
-* Fondé et travaillé dans une entreprise de technologie à Madrid pendant les 7 derniers années.
+* Fondé et travaillé dans une entreprise de conseil en optimisation et recherche opérationnelle à Madrid pendant les 7 derniers années.
 
@@
 
-* A Toulouse dans l'ISAE-SUPAERO.
+* A Toulouse à l'ISAE-SUPAERO.
 * Plus spécifiquement dans le Département d'Ingénierie des Systèmes Complexes (DISC).
-* Mes directeurs de thèse sont: Alain Haït, Olga Battaïa, Nicolas Dupin.
+* Mes directeurs de thèse sont: Alain Haït et Olga Battaïa.
+* Le sujet est issu d'une collaboration avec Nicolas Dupin de la DGA.
 
@@
 
-* Fait partie de la recherche opérationnelle: mathématiques dédiés à résoudre des gros et complexes problèmes de l'industrie.
+* Fait partie de la recherche opérationnelle: mathématiques dédiés à résoudre des problèmes complexes de l'industrie.
 * Un problème d'affectation des missions et planification des maintenances aux aéronefs militaires.
-* Similaire au problème FMP: Flight and Maintenance Planning .
-* Résolu par techniques de modélisation comme PLNE, méta-heuristiques ou heuristiques.
+* Résolu par techniques comme PLNE, méta-heuristiques ou heuristiques.
 
-# Example
+# Exemple
 
 ```{r echo=FALSE}
-htmltools::includeHTML("calendrier_20171116.html")
+htmltools::includeHTML("./../../img/calendar.html")
 ```
 
-# Etat de l'art
+# Exemple (2)
 
-* En @Cho2011, des affectations quotidiens des heures aux avions militaires.
-* En @Verhoeff2015 et @Kozanidis2008, des affectations mensuels des maintenances et opérations sont faits avec contraintes des disponibilité, niveau de service et soutenabilité.
-* En @Chastellux2017, un premier modèle d'affectation de missions aux aéronefs en lieu des heures est présenté.
+<style>
+  .col2 {
+    columns: 2 200px;         /* number of columns and width in pixels*/
+    -webkit-columns: 2 200px; /* chrome, safari */
+    -moz-columns: 2 200px;    /* firefox */
+  }
+</style>
+
+<div class="col2">
+```{r echo=FALSE}
+htmltools::includeHTML("./../../img/maintenances.html")
+htmltools::includeHTML("./../../img/unavailable.html")
+```
+</div>
 
@@
 
-* En faisant meilleures planifications, on fait une meilleur utilisation de ressources.
-* Lisser les maintenances pendant tout le horizon future.
-* garantir la disponibilité des aéronefs pour possibles nouvelles missions.
+* Faire meilleures planifications pour une meilleure utilisation des ressources.
+* Lisser les maintenances sur l'horizon de planification pour maximiser l'efficacité de l'équipe de maintenance.
+* Garantir la disponibilité des aéronefs pour de possibles nouvelles missions.
+
+# Pour quoi un doctorat?
+
+* Ces types de problèmes sont difficiles à résoudre.
+* Pour trouver des bons solutions aux problèmes il faut souvent chercher la structure du problème, sa complexité et ses similitudes / differences avec des autres problèmes.
+
+<!-- # Etat de l'art -->
+
+<!-- * En @Cho2011, des affectations quotidiens des heures aux avions militaires. -->
+<!-- * En @Verhoeff2015 et @Kozanidis2008, des affectations mensuels des maintenances et opérations sont faits avec contraintes des disponibilité, niveau de service et soutenabilité. -->
+<!-- * En @Chastellux2017, un premier modèle d'affectation de missions aux aéronefs en lieu des heures est présenté. -->
 

### 2017-12-19 latex/ROADEF2018/article_roadef_2018_alt.tex
changed name of ROADEF dir

--- 2017-11-27 latex/ROADEF/article_roadef_2018.tex
@@
 % Le titre court
-\def\shorttitle{Titre court}
+% \def\shorttitle{Titre court}
 
@@
 \institute{
-Institut Supérieur de l'Aéronautique et de l'Espace \\
+ISAE-SUPAERO, Université de Toulouse, France \\
 \email{\{franco.peschiera,alain.hait,olga.battaia,nicolas.dupin\}@isae-supaero.fr}
@@
 \section{Introduction}
-    The Military Flight and Maintenance Planning is a problem consisting on the assignment of missions and maintenance tasks to military aircraft.
-It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically but they are assumed to include a round-trip to the base.
+    The Military Flight and Maintenance Planning Problem considered here aims to assign missions and maintenance tasks to military aircraft.
+It is a variant of the Flight and Maintenance Planning problem where flights are not modelled geographically since a round-trip to the base is assumed for each flight.
 It also includes different objectives and constraints.
-Although the former has been studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008}, it has not received as much attention as the latter.
+Although the former has been studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015}, it has not received as much attention as the latter.
+In the following, the problem with already scheduled missions is considered.
 
@@
 
-    The problem consists in the planning of the maintenance operations for military aircraft and the assignment of these aircraft to already scheduled missions.
+    The problem consists in assigning military aircraft to a given set of missions while scheduling the maintenance operations over a time horizon.
     In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks".
-A series of $v \in \mathcal{V}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
-Each task requires a certain number $rr_{v}$ of resources $a \in \mathcal{A}$ which it employs for a time duration defined by $rh_{v}$ in each period.
-Set $cd_{v} \subset \mathcal{A}$ lists the resources that can be assigned to each task.
+A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
+Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
+Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.
+Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
+Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
+The number of resources in use for each period $d_t$ is also known.% (it is obtained by summing all $r_j$ for tasks $j \in J_t$).
 
     % The assignment of a resource to a task is not decided for the whole duration of the task.
-After a minimum amount of time ($minTIME_v$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished.
+After a minimum amount of time ($minTIME_j$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished.
 The total number of resources being used at any given time in a specific task can never be less than the required $REQ_{vt}$.
 
-    % \subsection{Resources}
+    % \subsection{iesources}
     % \label{def:resource}
 
-    Resources require recurrent preventive maintenance operations.
-The frequency of maintenance operations is defined by 2 factors: the amount of time $me$ elapsed since the last maintenance operation and the amount of time $mu$ for which the resource has been used since the last maintenance operation.
-The following states $s \in \mathcal{S}$ are possible for a resource: assigned to a task ('V'), receiving maintenance ('M'), available ('A'), needing maintenance ('N').
-A maintenance operation takes an amount of time equal to $md$ periods and cannot be interrupted.
-The number of resources under maintenance in any period is restricted by its capacity $c_t$.
-In the following we present a new mathematical model for this problem.
-Unlike previous work, it assigns aircraft to missions, instead of just using an aggregate demand.
+    Resources require recurrent preventive maintenance operations since the realization of tasks diminish the remaining usage time of resources.
+A resource cannot be used for a task $j$ if its remaining usage time is less that usage $h_j$ required by task $j$.
+A maintenance operation assigns to the resource exactly $H$ units of remaining usage time.
+The remaining usage time not used before the maintenance operation is lost.
+%Independently of its usage, each resource has to be maintained each $t^u$ periods of time since the end of the last maintenance operation.
+		A maintenance operation takes exactly $m$ periods and cannot be interrupted: during this time the resource cannot be assigned to any task.
+Some resources can be in maintenance at the beginning of the planning horizon, $n_t$ is used for the number of aircraft in such planned maintenance per period $t$ and defined only for the first $m-1$ time periods.
+In the next section, we present a new mathematical model for this problem.
     % We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance at any given period.
@@
 
-    % \paragraph{Resource's states}
+    % \paragraph{iesource's states}
     % \label{def:res-state}
@@
 
-    % \subsection{Decisions}
+    The decision variables manage the previously defined resources.
+%We can summarize the main decisions the following way: if resource $i$ starts maintenance in period $t$ and if resource $i$ is assigned to task $j$ in period $t$.
+We will call $\mathcal{T}_0$ to periods that include the period before the first, $0$, and $d_t$ the number of resources assigned to some task in period $t$.
+This parameter only should take some value in the first $md-1$ periods of the planning horizon.
 
-    % \subsection{Parameters}
-
-    % \begin{tabular}{p{10mm}lp{110mm}}
-    % $c_t$            & : & number of resources that can be in maintenance in period $t$. \\
-    % $cd_v$           & : & resources that satisfy task's $v$ needs. \\
-    % $rh_{vt}$        & : & time consumed of any resource when assigned to task $v$ in period $t$. \\
-    % $rr_{vt}$        & : & number or resources to assign to task $v$ in period $t$. \\            
-    % $mu$             & : & maximum used time for each resource between maintenances. \\    
-    % $me$             & : & maximum elapsed time for each resource between maintenances. \\    
-    % $md$             & : & maintenance duration, in periods. \\        
-    % $rut\_i_a$       & : & remaining used time for resource $a$ at the beginning of horizon. \\        
-    % $ret\_i_a$       & : & remaining elapsed time for resource $a$ at the beginning of horizon. \\
-    % $state\_i_{at}$  & : & some $s \in \mathcal{S}$ in case the resource $a$ has a fixed state. \\
-    % \end{tabular}
-    % \bigskip
-
-    % \subsection{Variables}
-
-    The decision variables manage the previously defined resources.
-We can summarize the main decisions the following way: if resource $a$ is in state $s$ in each period $t$ and if resource $a$ is assigned to task $v$ in period $t$.
-
-    \begin{tabular}{p{15mm}lp{105mm}}
-        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assigned to task $v$ in period $t$, 0 otherwise. \\  
-        $state_{ast}$       & : & 1 if resource $a$ is in state $s$ in period $t$, 0 otherwise. \\  
-        $start_{at}$        & : & 1 if resource $a$ starts maintenance in period $t$, 0 otherwise. \\
-        $ret_{at}$          & : & remaining elapsed time (integer) for resource $a$ after period $t$. \\  
-        $rut_{at}$          & : & remaining usage time (integer) for resource $a$ after period $t$. \\  
-        $used_{at}$         & : & time (integer) used by resource $a$ in period $t$. \\
-        $min\_avail$        & : & min number (integer) of available resources in any period. \\
-        $max\_maint$        & : & max number (integer) of resources in maintenance in any period. \\
+    \begin{tabular}{p{5mm}lp{165mm}}
+        $A_{jti}$           & =&1 if task $j \in J$ in period $t\in T_j$ is realized with resource $i \in a_j$, 0 otherwise. \\  
+        % $state_{ist}$       & = & 1 if resource $i$ is in state $s$ in period $t$, 0 otherwise. \\  
+        $M_{it}$            & =&1 if resource $i \in I$ starts maintenance in period $t \in \mathcal{T}$, 0 otherwise. \\
+        $rut_{it}$          & =&remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
+        % $ret_{it}$          & = & remaining elapsed time (integer) for resource $i$ after period $t \in \mathcal{T}0$. \\  
+        % $used_{it}$         & = & time (integer) used by resource $i$ in period $t$. \\
+        $U_{max}$        & =&maximal number (integer) of unavailable resources in any period. \\
+        $M_{max}$        & =&maximal number (integer) of resources in maintenance in any period. \\
     \end{tabular}
-    \bigskip
-
+Note that  $A_{jti}$ and $M_{it}$ are initially set up to 0 for all resources already in maintenance at the beginning of the planning period for the remaining time periods of maintenance.
+The remaining usage time for each resource at the beginning of the planning period is used to initialize $rut_{i0}$.
     % \subsection{Constraints}
 
-    The objective is to both minimize the maximum number of maintenances and maximize the minimal number of available aircraft:
-
+    The objective is to simultaneously minimize the maximum number of maintenances and the maximum number of unavailable aircraft. 
+ % The objective is to minimize the maximum number of unavailable aircraft at any time of the planning horizon:
     \begin{align}
-        & \text{Min}\, (max\_maint) \\
-        & \text{Max}\, (min\_avail)
+        & \text{Min}\; w \times M_{max} + U_{max}
     \end{align}
-
-    Subject to:
-
+    where relative weight $w$ is chosen by the decision maker. The following constraints are used.       
     \begin{align}
@@
         %%%%%%%%%%%%%%%%%%%
-        % maximum capacity:
-        & \sum_{a \in \mathcal{A}} state_{ast} \leq c_{t}
-                & t \in \mathcal{T}, s = \mq{M} \label{eq:capacity}\\
+        % maximum capacity1:
+        & \sum_{t' = 1}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \leq M_{max}
+                & t =1, 2, ..., m-1 \label{eq:capacity1}\\
+               % maximum capacity2:                
+               & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} \leq M_{max}
+        & t =m, ..., |\mathcal{T}|  \label{eq:capacity2}\\
+               %avail
+               & \sum_{t' = 1}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} + n_{t} + d_t\leq U_{max} 
+                & t =1, 2, ..., m-1 \label{eq:avalaibility1}\\
+               % maximum capacity2:                
+               & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} + d_t\leq U_{max} 
+        & t =m, ..., |\mathcal{T}|  \label{eq:avalaibility2}\\
+        % % maximum capacity:
+        % & \sum_{t' = \max{\{1, t - m+1\}}}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \leq M_{max}
+        %         & t \in \mathcal{T} \label{eq:capacity}\\
+        % % maximum avaliability:
+        % & \sum_{t' = \max{\{1, t - m+1\}}}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} + d_{t} + n_{t} \leq U_{max}
+        %         & \hspace{10mm} t \in \mathcal{T}  \label{eq:avalaibility}\\
+        % or maintenance or task:
         % minimum resources
-        & \sum_{a \in cd_v} task_{avt} \geq rr_{v}
-                & t \in \mathcal{T}, v \in \mathcal{V},  \label{eq:taskres}\\
-        % max one task per period
-        & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
-                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
-        % if assigned to task: in working state
-        & state_{ast} \geq task_{avt} 
-                & \hspace{10mm} t \in \mathcal{T}, a \in cd_v, v \in \mathcal{V}, s = \mq{V} \label{eq:state1}\\
-        % if no working state, no assigned to any task
-        & state_{ast} + task_{avt} \leq 1
-                & \hspace{10mm} t \in \mathcal{T}, a \in cd_v, v \in \mathcal{V}, s \neq \mq{V} \label{eq:state3}\\
-        % only one state. If we take out the 'do nothing' state, this could be \leq
-        & \sum_{s \in \mathcal{S}} state_{ast} = 1 
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}
+        & \sum_{i \in a_j} A_{jti} = r_{j}
+                & j \in \mathcal{J}, t \in T_j  \label{eq:taskres}\\
+			  & \sum_{t' = \max{\{1, t - m+1\}}}^{t} M_{it'} + \sum_{j \in J_t \cap O_i} A_{jti} \leq 1
+                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
     \end{align}
-
+    % $$H*M_{it} - Heures de vol entre t_1 et t \leq H*(M_{it_1}+1) i \in I, t \geq t-1 \in T$$
+    % $$t_1M_{it_1} \geq tM_{it} - D_{max} \foreach i \in I, t>t-1, t_1\geq D_{max}-D_i$$
     % These constraints are the the main constraints of the model.
     % The number of resources assigned to a single task and under maintenance 
-    Tasks' resource requirements are defined by \ref{eq:taskres}, maintenance capacity is controlled by \ref{eq:capacity}.
-Constraints \ref{eq:state1} and \ref{eq:state3} tie states with the assignment to a task.
-Maximum tasks and states assigned to each resource is guaranteed by \ref{eq:maxtasks} and \ref{eq:state2}.
-
+    % Maintenance capacity is controlled by \ref{eq:capacity}.
+The number of unavailable resources is defined by \ref{eq:avalaibility}.
+Tasks' resource requirements are defined by \ref{eq:taskres}.
+Constraints \ref{eq:state} guarantee that a resource can be used only for one task or maintenance operation at the same period.
+    Maintenance capacity is controlled by (\ref{eq:capacity1})-(\ref{eq:capacity2}).
+The number of unavailable resources is defined by (\ref{eq:avalaibility1})-(\ref{eq:avalaibility2}).
+Tasks' resource requirements are defined by (\ref{eq:taskres}).
+Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task or maintenance operation at the same period.
     \begin{align}
@@
         % definition of consumed resource hours:
-        & used_{at} \geq \sum_{v \in \mathcal{V}} task_{avt} \times rh_{v}
-                & a \in \mathcal{A}, t \in \mathcal{T} \label{eq:usage}\\
+        % & used_{it} \geq \sum_{j \in \mathcal{J}} A_{ijt} \times h_{j}
+        %         & i \in \mathcal{I}, t \in \mathcal{T} \label{eq:usage}\\
+        % definition of consumed resource hours (bigger):
+        % & used_{it} \geq A_{ijt} \times h_{j}
+        %         & i \in \mathcal{I}, t \in \mathcal{T}, j \in \mathcal{J} \label{eq:usage2}\\
         % remaining used time
-        & rut_{at} \leq (rut_{at-1} - used_{at}) + (mu \times start_{at})
-                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance1}\\
+         & rut_{it} \leq rut_{it-1} - \sum_{j \in J_t \cap O_i} A_{jti} \times h_{j} + H \times M_{it}
+                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:balance1}\\
         % remaining elapsed time. TODO: improve this constraint like the maintenance state duration
-        & ret_{at} \leq (ret_{at-1} - 1) + (me \times start_{at})
-                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:balance2}
+        % & ret_{it} \leq (ret_{it-1} - 1) + (me \times M_{it})
+        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:balance2}\\
+        %& rut_{i0} = rut^{Init}_i
+          %      & i \in \mathcal{I} \label{eq:initial}\\
+        % & ret_{it} = retInit_r
+        %         & t = 0, i \in \mathcal{I} \label{eq:initial2}\\
+        % & ret_{it} \leq me
+        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
+        & rut_{it} \geq H \times M_{it}
+                & i \in \mathcal{I}, t \in \mathcal{T}\label{eq:balance2}\\	
+				& rut_{it} \in [0,H]
+                & i \in \mathcal{I}, t \in \mathcal{T} \label{eq:mu}
     \end{align}
-
         % These constraints calculate the balances of hours for each resource.
-
-    The definition of the consumption of hours for each resource at each period is forced by \ref{eq:usage}.
-The monthly remaining usage and elapsed time are calculated by \ref{eq:balance1} and \ref{eq:balance2} respectively.
+    The remaining usage time is defined by (\ref{eq:balance1}) and its limits by (\ref{eq:balance2})-(\ref{eq:mu}). 
      % This is then applied in \ref{eq:balance1} to get the remaining usage time at each period for each resource.
@@
 
-    \begin{align}
+    % \begin{align}
         %%%%%%%%%%%%%%%%%%%
@@
         % if start maintenance, at least X periods
-        & state_{ast_2} \geq start_{at_1} 
-                & \hspace{10mm} (t_1, t_2) \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + md - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start1}\\
-        % if maintenance doesn't start, we cannot have maintenance:
-        & \sum_{t_1 \in \mathcal{T}} start_{at_1} \geq state_{ast_2} 
-                & \hspace{10mm} t_2 \in \mathcal{T}, t_1 \leq t_2 \leq t_1 + md - 1, a \in \mathcal{A}, s = \mq{M} \label{eq:start2}
-    \end{align}
+        % & state_{ist} \geq M_{it_1} 
+        %         & \hspace{10mm} (t_1, t) \in \mathcal{T}, t_1 \leq t \leq t_1 + md - 1, i \in \mathcal{I}, s = \mq{M} \label{eq:start1}\\
+        % % if maintenance doesn't start, we cannot have maintenance:
+        % & \sum_{t_1 \in \mathcal{T}} M_{it_1} \geq state_{ist} 
+        %         & \hspace{10mm} t \in \mathcal{T}, t_1 \leq t \leq t_1 + md - 1, i \in \mathcal{I}, s = \mq{M} \label{eq:start2}\\
+        % availqbility definition:
+        % & Umax \leq \sum_{i} state_{ist}
+        %         & \hspace{10mm} t \in \mathcal{T}, i \in \mathcal{I}, s = \mq{A} \label{eq:available}
+    % \end{align}
 
@@
 
-    The coupling between the start of each maintenance and the maintenance state is done by \ref{eq:start1} and \ref{eq:start2}.
+    % The coupling between the start of each maintenance and the maintenance state is done by \ref{eq:start1} and \ref{eq:start2}.
+The definition of the previously mentioned objective is done by \ref{eq:available}.
     % This way resources need to have started a maintenance to receive it and have to have maintenance for a minimum period after having starting it.
 
-    \begin{align}
+    % \begin{align}
         %%%%%%%%%%%%%%%%%%%
@@
         % min available
-        & min\_avail \leq \sum_{a} state_{ast}
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{A} \label{eq:available}\\
         % max maintenance
-        & max\_maint \geq \sum_{a} state_{ast}
-                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}, s = \mq{M} \label{eq:maintenance}
-    \end{align}
+        % & Mmax \geq \sum_{i} state_{ist}
+        %         & \hspace{10mm} t \in \mathcal{T}, i \in \mathcal{I}, s = \mq{M} \label{eq:maintenance}
+    % \end{align}
 
     % These constraints define the possible objectives.
-    The definitions of the two previously defined objectives is done by \ref{eq:available} and \ref{eq:maintenance}.
 
@@
     %     % rut initialize
-    %     & rut_{at} = rut\_ini_a
-    %             & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining1}\\
+    %     & rut_{it} = rut\_ini_a
+    %             & \hspace{10mm} t = 0, i \in \mathcal{I} \label{eq:remaining1}\\
     %     % ret initialize
-    %     & ret_{at} = ret\_ini_a
-    %             & \hspace{10mm} t = 0, a \in \mathcal{A} \label{eq:remaining2}\\
+    %     & ret_{it} = ret\_ini_a
+    %             & \hspace{10mm} t = 0, i \in \mathcal{I} \label{eq:remaining2}\\
     %     % fixed states during planning period
-    %     & state_{ast} = 1
-    %             & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A}; s \in \mathcal{S}\, | if\, state\_ini_{at} = s \label{eq:fixedstates}
+    %     & state_{ist} = 1
+    %             & \hspace{10mm} t \in \mathcal{T}, i \in \mathcal{I}; s \in \mathcal{S}\, | if\, state\_ini_{it} = s \label{eq:fixedstates}
     % \end{align}
@@
 
-%\section{Conclusions}
-
- %   The present modelization is a new way to formulate the Military Flight Planning Model.
-Unlike previous work, it assigns aircraft to missions instead of just using an aggregate hourly demand per period.
-This is useful in order to improve the model by including minimum durations for mission assignment.
+\section{Conclusions and future work} 
+    The present work proposes a new model for the Military Flight and Maintenance Problem.
+This formulation sets an interesting base for further model extensions in order to integrate more complex real constraints.
 
@@
 \bibliographystyle{plain}
-\selectlanguage{french}
+% \selectlanguage{english}
 \bibliography{MFMP}
-% \begin{thebibliography}{2}
-% \bibitem{toth02}
-% Paolo Toth and Daniele Vigo.
-% \newblock \emph{The Vehicle Routing Problem}.
-% \newblock Monographs on Discrete Mathematics and Applications. Society for Industrial and Applied Mathematics, 2002.
-% \bibitem{kirkpatrick83}
-% Scott Kirkpatrick, C~Daniel Gelatt, and Mario~P Vecchi.
-% \newblock Optimization by simmulated annealing.
-% \newblock \emph{science}, 220\penalty0 (4598):\penalty0 671--680, 1983.
-
-% \end{thebibliography}
 

### 2017-12-20 R/dga_diaporama/dga.Rmd
edited DGA presentation and gantt files.

@@
 * Péruvien et français.
-* Ingénieur Industriel par l'Universidad Politecnica de Madrid.
+* Ingénieur Industriel de l'Universidad Politecnica de Madrid.
 * Passionné par les mathématiques appliqués et l'informatique.
@@
 
-* Fait partie de la recherche opérationnelle: mathématiques dédiés à résoudre des problèmes complexes de l'industrie.
-* Un problème d'affectation des missions et planification des maintenances aux aéronefs militaires.
+* Fait partie de la recherche opérationnelle: mathématiques dédiées à résoudre des problèmes complexes de l'industrie.
+* Un problème d'affectation des missions et planification des maintenances des aéronefs militaires.
 * Résolu par techniques comme PLNE, méta-heuristiques ou heuristiques.
@@
 
-* Faire meilleures planifications pour une meilleure utilisation des ressources.
+* Faire de meilleures planifications pour une meilleure utilisation des ressources.
 * Lisser les maintenances sur l'horizon de planification pour maximiser l'efficacité de l'équipe de maintenance.
@@
 * Ces types de problèmes sont difficiles à résoudre.
-* Pour trouver des bons solutions aux problèmes il faut souvent chercher la structure du problème, sa complexité et ses similitudes / differences avec des autres problèmes.
+* Domaine scientifique: l'optimisation combinatoire.
+* Pour trouver des bonnes solutions aux problèmes il faut souvent chercher la structure du problème, sa complexité et ses similitudes / differences avec d'autres problèmes.
 

### 2018-01-09 R/dga_diaporama/dga.Rmd
changed MOSIM article and dga presentation

@@
 author: "Franco Peschiera"
-date: '`r format(Sys.time(), "%Y-%m-%d")`'
+# date: "`r format(Sys.time(), "%Y-%m-%d")`"
+output:
+  beamer_presentation:
+    incremental: yes
+  slidy_presentation:
+    css: ./../../css/baobab.css
+    incremental: yes
+nocite: '@*'
 bibliography: MFMP.bib
-nocite: "@*"
-output: 
-  slidy_presentation: 
-    incremental: yes
-    css: "./../../css/baobab.css"
 ---
@@
 
+![Maintenances](./../../img/calendar.png)
 ```{r echo=FALSE}
-htmltools::includeHTML("./../../img/calendar.html")
+#htmltools::includeHTML("./../../img/calendar.html")
 ```
@@
 
-<style>
-  .col2 {
-    columns: 2 200px;         /* number of columns and width in pixels*/
-    -webkit-columns: 2 200px; /* chrome, safari */
-    -moz-columns: 2 200px;    /* firefox */
-  }
-</style>
+<!-- <style> -->
+<!--   .col2 { -->
+<!--     columns: 2 200px;         /* number of columns and width in pixels*/ -->
+<!--     -webkit-columns: 2 200px; /* chrome, safari */ -->
+<!--     -moz-columns: 2 200px;    /* firefox */ -->
+<!--   } -->
+<!-- </style> -->
 
-<div class="col2">
+<!-- <div class="col2"> -->
+![Maintenances](./../../img/maintenances.png)
+
+# Exemple (3)
+![Affectations](./../../img/unavailable.png)
 ```{r echo=FALSE}
-htmltools::includeHTML("./../../img/maintenances.html")
-htmltools::includeHTML("./../../img/unavailable.html")
+# htmltools::includeHTML("./../../img/maintenances.html")
+# htmltools::includeHTML("./../../img/unavailable.html")
 ```
-</div>
+<!-- </div> -->
 

### 2018-01-09 latex/ROADEF2018/article_roadef_2018_alt.tex
changed MOSIM article and dga presentation

@@
     The Military Flight and Maintenance Planning Problem considered here aims to assign missions and maintenance tasks to military aircraft.
-It is a variant of the Flight and Maintenance Planning problem where flights are not modelled geographically since a round-trip to the base is assumed for each flight.
+It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically since a round-trip to the base is assumed for each flight.
 It also includes different objectives and constraints.
@@
 % \selectlanguage{english}
-\bibliography{MFMP}
+\bibliography{./../biblio/MFMP}
 

### 2018-01-11 latex/MOSIM2018/article.tex
Added latex files with bibliography and MOSIM article

\documentclass[a4paper,twocolumn,fleqn]{article}
\textwidth 17cm \textheight 247mm
\topmargin -4mm
\hoffset -9mm \voffset -14mm
%\setlength{\topmargin}{-0.8truecm}
\setlength{\oddsidemargin}{0.3cm}
\setlength{\evensidemargin}{0.3cm}
\setlength{\columnsep}{8mm}
\setlength{\parindent}{0mm}
\setlength{\parskip}{-2.0ex}
\setlength{\mathindent}{0mm}
\flushbottom

\usepackage{epsfig}
%\usepackage{timesnew}
\usepackage{harvard}
\usepackage{amsmath}
\usepackage{booktabs}

\usepackage[latin1]{inputenc}
\usepackage{aeguill}
\usepackage[frenchb,english]{babel}
\usepackage{caption}

\setlength{\parskip}{2ex} \pagestyle{myheadings}
\markright{\hspace*{3.8cm} \textit{MOSIM18 - June 27-29, 2018 - Toulouse - France}}

\renewcommand{\thepage}{}
\renewcommand{\refname}{REFERENCES}

\makeatletter
\renewcommand\section{\@startsection{section}{1}{\z@}%
                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
                       {2\p@ \@plus 0\p@ \@minus 0\p@}%
                       {\normalsize\textbf}}

\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
                       {2\p@ \@plus 0\p@ \@minus 0\p@}%
                       {\normalsize\textbf}}

\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
                       {1\p@ \@plus 0\p@ \@minus 0\p@}%
                       {\normalsize\itshape\bfseries}}
\makeatother

\begin{document}

\title{ \vspace*{-25mm}
\centering
\fbox{\normalsize
\begin{minipage}{17cm}
\centering
\em \small{$12^{th}$ International Conference on MOdeling, Optimization and SIMlation - MOSIM18 - June 27-29 2018  \\ Toulouse - France "The rise of connected systems in industry and services"}
\end{minipage} }\\
\vspace*{6mm}
{\Large \textbf{Maintenance planning on French military aircraft operations}}}

\author{
\begin{tabular}{c}
\bf \normalsize {Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin} \\
\\
     \normalsize ISAE-SUPAERO \\
     \normalsize Université de Toulouse \\
     \normalsize 57045 Metz cedex 01 - France\\
     \normalsize {\{franco.peschiera,alain.hait,olga.battaia,nicolas.dupin\}@isae-supaero.fr}
     \\
\end{tabular}
}

\date{\begin{minipage}{17cm}
\normalsize
{\bf  ABSTRACT:}
\rm
{\em The Military Flight and Maintenance Planning problem is presented. In this problem, preventive maintenance operations are scheduled for military aircraft along with the assignment of regular missions. The scheduling of maintenances and mission assignments is a tedious task currently done manually by specialized personnel. A MIP mathematical model is formulated in order to produce better quality solutions and reduce the time it takes to find them. The quality of the solutions is measured via a multi-objective function that smooths both maintenances and aircraft unavailability among time periods. A real world dataset provided by the French Airforce is used to test the implementation.}\\~\\
{\bf KEYWORDS:}
\rm
{\em optimization, planning, military, maintenance}
\end{minipage}
}
\maketitle

% Here I put the list of things we will leave for another article:
% 1. stochasticity
% 2. symmetry breaking inside model.
% 3. storage?

\section{Introduction}

  The Military Flight and Maintenance Planning problem considered assigns missions and schedules maintenances to aircraft. In this problem, aircraft are needed to comply with a series of planned missions while at the same time needing to receive frequent preventive maintenances in order to be ready for new missions.

  It is a variant of the Flight and Maintenance Planning problem, where maintenances and flights are scheduled to commercial aircraft. The main differences are the fact that flights are not modeled geographically: all aircraft are assumed to exit the main depot and return to the same depot at the end of the flight.

  Other differences include the following.

  First of all, the objective function's motivation is on reducing the load on the maintenance facilities and guaranteeing the availability of the fleet instead of reducing the cost. Also, the maintenance operations take considerably longer and the rules that govern the maintenances are different.

  The temporal scope of the problem is also bigger, making the normal planning horizon bigger and each period bigger too: months instead of days. On the other hand, the physical scope, while present in the commercial planning, is not present in the military one: aircraft are assumed to return to the origin every time.

\section{Related work}

  % TODO: add FMP biblio?

  % The Military Flight and Maintenance Planning Problem considered here aims to assign missions and maintenance tasks to military aircraft. It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically since a round-trip to the base is assumed for each flight.

  % It also includes different objectives and constraints.

  % Although the former has been studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015}, it has not received as much attention as the latter. In the following, the problem with already scheduled missions is considered.

  In \cite{Kozanidis2008}, a model was made for the Hellenic Air Force. It assigned flight hours to aircraft and scheduled flexible maintenance periods based on a "remaining usage time" rule. The objectives were to maximize this "remaining usage time" in the aircraft as well as to smooth the availability of aircraft during the planning period. Instances of 24 aircraft and 6 monthly periods were tested with Mixed Integer Programming (MIP) formulations and heuristics.

  In \cite{Cho2011}, the objective was to smooth the maintenances as much as possible. Randomly generated instances of 15 aircraft and 520 bi-daily periods were tested with MIP formulations. The final state of aircraft was taken into account in order to distribute as much as possible the "remaining usage time" of aircraft.

  The concept of mission is not introduced as part of these formulations, only a demand of flight hours is given for all aircraft in total. The aircraft themselves have been assumed as homogeneous and interchangeable. Due to this aggregation of demand, there are no minimal time assignments for aircraft to missions. Missions are important in order to guarantee that aircraft have the required capabilities needed for the missions it is assigned to.

  % Additionally, the possibility of storage of aircraft is not present in the previous work.

  Lastly, in \cite{Chastellux2017} a mission-centered modeled is presented. Here, cycles, composed of one maintenance followed by mission assignments were used. Instances of 6 - 70 aircraft and 60 monthly periods were tested, founding close to optimal solutions for small to medium instances using MIP formulations and heuristics based on these MIP formulations.

  Here, an alternative model is presented that simplifies the formulation and should permit the inclusion of the rest of constraints without sacrificing optimality.

\section{The problem}

  The problem consists in assigning resources to predefined tasks and scheduling periodic preventive maintenances for these same resources in order to keep them ready for doing tasks in the future. In order to generalize the formulation of the model, in the description presented below, the following naming convention has been adopted: aircraft are called resources, missions are called tasks.

  \subsection{Tasks}
    \label{def:task}

    There is a fixed set of $j \in \mathcal{J}$ tasks to be accomplished over an horizon of time divided into $t \in \mathcal{T}$ discrete periods. In order to be accomplished, these tasks require the assignment of a specific number of resources $r_{j}$ each period of time the task is active. Tasks have both a fixed start and end time that define the set of periods of time it is active.

    During each assigned period, tasks use the assigned resources for a time $h_j$.

    The assignment of a resource to a task is not decided for the whole duration of the task. After a minimum amount of time ($mt$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished. The total number of resources being used at any given time in a specific task can never be less than the required $r_{j}$.

    Finally, tasks have a set of requisites that the resources that can be assigned to it need to comply with.

    % TODO: add graph showing some tasks.

  \subsection{Resources}
    \label{def:res}

    There is a set $i \in \mathcal{I}$ of available resources that are assigned to tasks in order to accomplish them. Each resource can only be assigned to a single task in any given period. These resources suffer from wear and tear and require regular maintenances during their lifetime. The need for these maintenances is calculated based on two indicators every resource has.

    We will call "remaining elapsed time" (or $ret_{it}$) to the amount of time (measured in time periods) that needs to pass in order for the resource $i$ to need a maintenance at any given period $t$. In a similar way, we will call "remaining usage time" (or $rut_{it}$) to the amount of time that the resource $i$ needs to be used in order for it to need a maintenance at any given period $t$.

    Additionally, after an absolute amount of time and/or usage ($aet_i$ or $aut_i$), the resource becomes obsolete. There is no way to reverse this process.

    At any given period, including at the start of the planning period, each resources has a specific status given by:

    \begin{itemize}
        \item remaining usage time.
        \item remaining elapsed time.
        \item remaining absolute elapsed time.
        \item remaining absolute usage time.
        % \item remaining storage time (see \ref{def:sto}).
        \item type of last maintenance (see \ref{def:maint}).
    \end{itemize}

  \subsection{Maintenances}
    \label{def:maint}

    Maintenances are the process by which resources that have reached a limit in their usage can return to a state where they can continue to be used in tasks.

    % Each maintenance belongs to a specific type of maintenance $m \in \mathcal{M}$. These types of maintenance differentiate between each other by having potentially different characteristics such as: different durations, the provision of new functionalities for the resource or the restoration of the storage capacity (see \ref{def:sto}). The maintenance type is not a decision to be made since there is a specific sequence of maintenance types a resources needs to follow according to the resource type.

    Maintenances have a fix duration of $m$ periods.

    After receiving a maintenance, a resource restores its remaining elapsed time and remaining usage time back to their max values $E$ and $H$ respectively.

    Finally, resources are organized into families or groups. Each resource inside a family or group shares the same types of maintenances and, usually, is able to do the same type of tasks.

% \subsection{Storages}
%   \label{def:sto}

%   Following the rule of remaining elapsed time, even if a resource is not being used, it still needs to have a maintenance after a given amount of time has passed. In order to avoid this problem, the resource can be put into a storage state.

%   A resource in this states has to be kept for a minimum time of $sm$ periods. While in this state it cannot receive maintenance or be assigned any task.

%   Every resource has the capacity to be stored and this capacity is measured in a number of periods $sc$. In order for a resource to restore its remaining storage capacity, it needs to receive a specific maintenance (see \ref{def:maint}). Similar to the remaining elapsed time, after these maintenances, the resource recovers its storage capacity up to a certain level $S$.

  \subsection{Possible states}

    As a summary, the following are the possible logical states that a resource can be in:

      \begin{itemize}
          \item Assigned to a task (see \ref{def:task}).
          \item Under maintenance (see \ref{def:maint}).
          % \item Under storage (see \ref{def:sto}).
            \item Obsolete (see \ref{def:res}).
          \item Available.
      \end{itemize}

    Figure \ref{fig:solution} shows an example of an assignment of missions (red) and maintenances (blue) to aircraft (codes A1 - A9).

      \begin{figure}
        \centering
          \includegraphics[width=0.5\textwidth]{./../../img/calendar.png}
        \caption{Example solution}
        \label{fig:solution}
      \end{figure}

  \subsection{Horizon}
    \label{def:hor}

    In planning tasks and maintenances, it is important to take into account the initial state of each resource. This initial state can be a maintenance or an assigned task. If a resource is under maintenance, it needs to continue in this state for its remaining maintenance time. Tasks' assignments should be taken into account in a similar manner.

    On the other hand, the remaining used and elapsed times need to be assigned to each resource at the beginning of the planning period.

    Finally, the end state of each resource needs to be addressed too. This implies guaranteeing that the remaining (elapsed, used) time of resources is sustainable for future tasks and the final states of resources are not too skewed.

  \subsection{Objectives}

    The are multiple objectives that need to be taken into account.

    Given that the creation of new tasks and the duration of maintenance are considered stochastic, the basic goal is to maximize the robustness of the planning.

    Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of resources under maintenance over the planning horizon.

    This is summarized by the following two sentences:

    \begin{enumerate}
        \item Minimize the maximum number of unavailable resources in any given period.
        \item Minimize the maximum number of resources under maintenance in any given period.
    \end{enumerate}

    As it will be shown in the mathematical formulation, these objectives are quite related one with the other: the more resources are in maintenance in a given period, the more unavailable resources will be had.

\section{Optimization model}

  \subsection{Parameters}

    \begin{tabular}{p{8mm}p{67mm}}
        $h_j$             & amount of usage time for resources used in task $j$. \\
        $r_j$             & amount of resources that task $j$ needs when active. \\
        $mt$              & minimum number of periods a resource can be assigned to a task. \\
        $aut_t$           & maximal absolute usage time for a resource. \\
        $aet_i$           & maximal absolute elapsed time for a resource. \\
        $m$               & maintenance duration in number of periods. \\
        $E$               & remaining elapsed time after a maintenance. \\
        $H$               & remaining usage time after a maintenance. \\
        % $S$               & remaining storage time after (certain) maintenances. \\
        $w$               & weight of smoothness for maintenances. \\
        $n_t$             & number of resources in already-planned maintenance per period.\\
        $d_t$             & number of resources that need to be assigned in total in period $t$. \\
        $rut^{Init}_{i}$  & remaining usage time per resource at the start of the planning period. \\
        $ret^{Init}_{i}$  & remaining elapsed time per resource at the start of the planning period. \\
    \end{tabular}

  \subsection{Sets}

    \begin{tabular}{p{5mm}p{70mm}}
        $T_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ will be active. \\
        $J_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
        $a_{j}$   &  resources $i \in \mathcal{I}$ that can be assigned to each task $j$. \\
        $O_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
        $T^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - m+1\}},  ..., {t}$\}. \\
    \end{tabular}

  \subsection{Variables}

     The decision variables manage the previously defined resources.

    \begin{tabular}{p{8mm}p{67mm}}
        $A_{jti}$   &  1 if task $j \in J$ in period $t\in T_j$ is realized with resource $i \in a_j$, 0 otherwise. \\  
        $M_{it}$    &  1 if resource $i \in I$ starts maintenance in period $t \in \mathcal{T}$, 0 otherwise. \\
        $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
        $ret_{it}$  &  remaining elapsed time (integer) for resource $i$ after period $t \in \mathcal{T}0$. \\  
        $U_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
        $M_{max}$   &  maximal number (integer) of resources in maintenance in any period. \\
    \end{tabular}

    Note that  $A_{jti}$ and $M_{it}$ are initially set up to 0 for all resources already in maintenance at the beginning of the planning period for the remaining time periods of maintenance. The remaining usage time for each resource at the beginning of the planning period is used to initialize $rut_{i0}$.

  \subsection{Constraints}

    The objective is to simultaneously minimize the maximum number of maintenances and the maximum number of unavailable aircraft.

    \begin{align}
        & \text{Min}\; w1 \times M_{max} + w2 \times U_{max}
    \end{align}
    where relative weight $w$ is chosen by the decision maker. 
    The following constraints are used in the model:       
    \begin{align}
        % maximum capacity1:
        & \sum_{t' \in T^{s}_t} \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \leq M_{max}
          & t \in \mathcal{T} \label{eq:capacity1}\\
               % maximum capacity2:                
        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} \leq M_{max}
        % & t =m, ..., |\mathcal{T}|  \label{eq:capacity2}\\
               %avail
       & \sum_{t' \in T^{s}_t} \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \notag \\
           &\hspace*{10mm}+ d_t\leq U_{max} 
        &t \in \mathcal{T} \label{eq:avalaibility1}\\
               % maximum capacity2:                
        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} + d_t\leq U_{max} 
        % & t =m, ..., |\mathcal{T}|  \label{eq:avalaibility2}\\
        & \sum_{i \in a_j} A_{jti} = r_{j}
                & j \in \mathcal{J}, t \in T_j  \label{eq:taskres}\\
        & \sum_{t' \in T^{s}_t} M_{it'} + \sum_{j \in J_t \cap O_i} A_{jti} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

    Maintenance capacity is controlled by (\ref{eq:capacity1}). The number of unavailable resources is defined by (\ref{eq:avalaibility1})). Tasks' resource requirements are defined by (\ref{eq:taskres}). Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task or maintenance operation at the same period.  
    \begin{align}
        % remaining used time
         & rut_{it} \leq rut_{it-1} + H M_{it} \notag \\ 
           &\hspace*{4mm}- \sum_{j \in J_t \cap O_i} A_{jti} h_{j} \notag \\
                && t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
        & rut_{i0} = rut^{Init}_i
               & i \in \mathcal{I} \label{eq:rut_initial}\\
        & rut_{it} \geq H M_{it}
                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
        & rut_{it} \in [0,H]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\
        & ret_{it} \leq ret_{it-1} - 1 \notag \\
          &\hspace*{4mm}+ E M_{it}
                & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
        & ret_{i0} = ret^{Init}_r
                & i \in \mathcal{I} \label{eq:ret_initial}\\
        & ret_{it} \geq E M_{it}
                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
        & ret_{it} \in [0,me]
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}
    \end{align}
        % These constraints calculate the balances of hours for each resource.
    The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}). 
    Similarly, the remaining elapsed time is defined by (\ref{eq:ret_upper})-(\ref{eq:ret_initial}) and its limits by (\ref{eq:ret_lower})-(\ref{eq:me}).

\section{Datasets}

  The dataset that was used was provided by the French Direction Générale de l'Armement (DGA).
  Based on the information provided, a number of datasets were created to run different tests.

  \subsection{Tasks}

    The main characteristics of the dataset are the following. 11 tasks, 127 resources.
    Most tasks demand between 1 and 8 resources per month. The biggest and second biggest tasks demand 50 and 20 resources respectively each month.
    More details in table \ref{tab:tasks}.
    Candidates are calculated based on each task's requirements matching the candidate's capabilities.

    It is possible to see that task 08 is impossible to satisfy given that there are less candidates than resources needed. This task has been taken out from all of the instances.

    \begin{table}
    \begin{tabular}{lrrr}
    \toprule
    {} &  usage (h) &  \# resource &  \# candidates \\
    \midrule
    O1  &         80 &           8 &            50 \\
    O10 &         23 &          50 &            74 \\
    O11 &         23 &           6 &             8 \\
    O12 &         23 &           8 &            16 \\
    O2  &         60 &           2 &            50 \\
    O3  &         50 &           2 &            16 \\
    O4  &         50 &           2 &            50 \\
    O5  &         23 &           4 &            29 \\
    O6  &         23 &          20 &            29 \\
    O7  &         23 &           1 &            16 \\
    O8  &         23 &           4 &             3 \\
    O9  &         23 &           1 &            29 \\
    \bottomrule
    \end{tabular}
    \caption{Tasks' comparative table}
    \vspace{-0.5cm}
    \label{tab:tasks}
    \end{table}

  \subsection{Maintenances}

    All maintenances take 6 months (periods). Each maintenance provides with 1000 hours of usage time to the resource and a resource can be without maintenance for a maximum of 60 periods. The nominative capacity for maintenance is 18 resources at any given time.

    \subsection{Resources}

      The distribution of initial usage times is shown in the figure \ref{fig:histogramUsage}. It can be seen that the aircraft do not necessarily follow an horizontal line.

      \begin{figure}
        \centering
          \includegraphics[width=0.5\textwidth]{./../../img/initial_used.png}
        \caption{Distribution of Initial Usage Times among resources}
        \label{fig:histogramUsage}
      \end{figure}

\section{Heuristics}

\section{Results}

  All tests were run on a 64 bit Ubuntu 16.04 workstation with 16GB of RAM and Intel i7-6700HQ CPU \@ 2.60GHz x 8 processor.
  Several experiments were done with the MIP formulation. Table \ref{tab:instance} shows the characteristics of each instance. 
  Most of the columns describe the size of the instance: number of periods, number of tasks and number of assignments.
  The assign column indicates the number of assignments that are needed to be done to all the task-periods.

  Both weights on the objective function were set at 1. For a sensibility analysis on these two parameters, see section \ref{sec:multi}.

  \begin{table}[h]
  \begin{center}

\begin{tabular}{lrrrrrr}
\toprule
   id &  $|\mathcal{T}|$ &  assign &  $|\mathcal{J}|$ &  vars &  cons &  nonzeros \\
\midrule
 2321 &               11 &    1080 &               11 &  7270 &  6329 &     41266 \\
 1839 &               41 &    2069 &               10 & 26171 & 25062 &    169532 \\
 2127 &               31 &    1610 &               10 & 19789 & 18750 &    126376 \\
\bottomrule
\end{tabular}

  \caption{Instances' description}
  \vspace{-0.5cm}
  \label{tab:instance}
  \end{center}
  \end{table}

    Table \ref{tab:results} show the details of the solving process for each instance.

  \begin{table}[h]
  \begin{center}

\begin{tabular}{lrrrr}
\toprule
   id &  gap (\%) &    OF &  time (s) &  bound \\
\midrule
 2321 &      0.7 & 139.0 &       500 &  138.0 \\
 1839 &     90.5 & 855.0 &      3600 &   81.7 \\
 2127 &      1.2 &  83.0 &      3600 &   82.0 \\
\bottomrule
\end{tabular}

  \caption{Solutions' description}
  \vspace{-0.5cm}
  \label{tab:results}
  \end{center}
  \end{table}

  \section{Multiobjective analysis}
  \label{sec:multi}

    In order to understand the sensibility of the weights $w1$ and $w2$ on the solution, a multiobjective analysis was done. In it, both weights were changed between 0 and 1 in intervals of 0.1 so that they always summed 1. Table \ref{tab:multiobj} shows the obtained results for each instance. The base instance that was used was the one of 31 periods and 10 tasks.

    The instances were not solved until optimum value but it can be seen that the remaining gap is very low (1.5\% maximum).

    In figure \ref{fig:pareto} it's possible to see that solutions W01, W05, W06 and W09 dominate the other solutions, having the latter three the same objective function.

  \begin{table}[h]
  \begin{center}

\begin{tabular}{lrrrrr}
\toprule
  id &  maint &   w2 &   w1 &  unavailable &  gap (\%) \\
\midrule
 W00 &     21 &  1.0 &  0.0 &           67 &     1.49 \\
 W01 &     16 &  0.9 &  0.1 &           67 &     0.00 \\
 W10 &     15 &  0.0 &  1.0 &           69 &     0.00 \\
 W02 &     19 &  0.8 &  0.2 &           67 &     0.35 \\
 W03 &     17 &  0.7 &  0.3 &           68 &     0.58 \\
 W04 &     15 &  0.6 &  0.4 &           69 &     0.43 \\
 W05 &     15 &  0.5 &  0.5 &           68 &     1.20 \\
 W06 &     15 &  0.4 &  0.6 &           68 &     1.10 \\
 W07 &     16 &  0.3 &  0.7 &           70 &     0.97 \\
 W08 &     16 &  0.2 &  0.8 &           70 &     0.78 \\
 W09 &     15 &  0.1 &  0.9 &           68 &     0.49 \\
\bottomrule
\end{tabular}

  \caption{Multiobjective comparison}
  \vspace{-0.5cm}
  \label{tab:multiobj}
  \end{center}
  \end{table}

      \begin{figure}
        \centering
          \includegraphics[width=0.5\textwidth]{./../../img/multiobjective.png}
        \caption{Pareto diagram of weight combinations}
        \label{fig:pareto}
      \end{figure}

\section{Conclusions}

  The results show that a direct exact method to solve the whole dataset is not ideal.

\bibliographystyle{dcu}
\bibliography{./../biblio/MFMP}

\end{document}

### 2018-01-12 latex/MOSIM2018/article.tex
added latex tables in separate files. Updated article with a lot of changes

@@
 \vspace*{6mm}
-{\Large \textbf{Maintenance planning on French military aircraft operations}}}
+{\Large \textbf{Bi-objective MIP formulation for the optimization of maintenance planning on French military aircraft operations}}}
 
@@
      \normalsize Université de Toulouse \\
-     \normalsize 57045 Metz cedex 01 - France\\
+     \normalsize 31055 Toulouse cedex 4 - France\\
      \normalsize {\{franco.peschiera,alain.hait,olga.battaia,nicolas.dupin\}@isae-supaero.fr}
@@
 A MIP mathematical model is formulated in order to produce better quality solutions and reduce the time it takes to find them.
-The quality of the solutions is measured via a multi-objective function that smooths both maintenances and aircraft unavailability among time periods.
+The quality of the solutions is measured via a bi-objective function that smooths both maintenances and aircraft unavailability among time periods.
 A real world dataset provided by the French Airforce is used to test the implementation.}\\~\\
@@
 \rm
-{\em optimization, planning, military, maintenance}
+{\em optimization, planning, aircraft maintenance, MIP formulation, bi-objective optimization}
 \end{minipage}
@@
 
+  In section \ref{sec:soa} we present previous work in this domain.
+Section \ref{sec:problem} describes with detail the problem, its different elements, conditions and objectives, while the particular dataset used in this study is described in \ref{sec:data}.
+Section \ref{sec:model} presents a complete MIP formulation to solve the described problem.
+
+  Section \ref{sec:results} shows the results obtained while solving several instances of the problem and finally, section \ref{sec:conclusions} mentions the conclusions obtained from the experimentation and the possible direction of further work in this domain.
+
 \section{Related work}
+  \label{sec:soa}
 
@@
 
-  Here, an alternative model is presented that simplifies the formulation and should permit the inclusion of the rest of constraints without sacrificing optimality.
+  % Here, an alternative model is presented that simplifies the formulation and should permit the inclusion of the rest of constraints without sacrificing optimality.
 
-\section{The problem}
-
+\section{Problem statement}
+  \label{sec:problem}
   The problem consists in assigning resources to predefined tasks and scheduling periodic preventive maintenances for these same resources in order to keep them ready for doing tasks in the future.
@@
 
-    % TODO: add graph showing some tasks.
-
   \subsection{Resources}
@@
 
-    Additionally, after an absolute amount of time and/or usage ($aet_i$ or $aut_i$), the resource becomes obsolete. There is no way to reverse this process.
+    % Additionally, after an absolute amount of time and/or usage ($aet_i$ or $aut_i$), the resource becomes obsolete. There is no way to reverse this process.
 
@@
         \item remaining elapsed time.
-        \item remaining absolute elapsed time.
-        \item remaining absolute usage time.
+        % \item remaining absolute elapsed time.
+        % \item remaining absolute usage time.
         % \item remaining storage time (see \ref{def:sto}).
@@
           % \item Under storage (see \ref{def:sto}).
-            \item Obsolete (see \ref{def:res}).
+            % \item Obsolete (see \ref{def:res}).
           \item Available.
@@
 
-    Figure \ref{fig:solution} shows an example of an assignment of missions (red) and maintenances (blue) to aircraft (codes A1 - A9).
+    Figure \ref{fig:solution} shows an extraction of an assignment of missions (red) and maintenances (blue) to aircraft (codes A1 - A9).
 
@@
           \includegraphics[width=0.5\textwidth]{./../../img/calendar.png}
-        \caption{Example solution}
+        \caption{Visual representation of part of the solution to an instance of the problem showing maintenances and assignments}
         \label{fig:solution}
@@
 
-    The are multiple objectives that need to be taken into account.
+    The are multiple objectives that need to be taken into account. In this study two will be addressed.
 
-    Given that the creation of new tasks and the duration of maintenance are considered stochastic, the basic goal is to maximize the robustness of the planning.
+    Given that the creation of new tasks and the duration of maintenance are considered stochastic, one basic goal is to maximize the robustness of the planning by having the greatest amount of available resources at every period of the planning horizon.
 
@@
 \section{Optimization model}
-
+  \label{sec:model}
   \subsection{Parameters}
@@
         $mt$              & minimum number of periods a resource can be assigned to a task. \\
-        $aut_t$           & maximal absolute usage time for a resource. \\
-        $aet_i$           & maximal absolute elapsed time for a resource. \\
+        % $aut_i$           & maximal absolute usage time for resource $i$. \\
+        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
         $m$               & maintenance duration in number of periods. \\
@@
         % $S$               & remaining storage time after (certain) maintenances. \\
-        $w$               & weight of smoothness for maintenances. \\
-        $n_t$             & number of resources in already-planned maintenance per period.\\
+        $w1$               & weight of smoothness for maintenances. \\
+        $w2$               & weight of smoothness for unavailabilities. \\
+        $n_t$             & number of resources in already-planned maintenance in period $t$.\\
         $d_t$             & number of resources that need to be assigned in total in period $t$. \\
-        $rut^{Init}_{i}$  & remaining usage time per resource at the start of the planning period. \\
-        $ret^{Init}_{i}$  & remaining elapsed time per resource at the start of the planning period. \\
+        $rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning period. \\
+        $ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning period. \\
+        $ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning period. \\
+        $rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning period. \\
     \end{tabular}
@@
     \begin{tabular}{p{5mm}p{70mm}}
-        $T_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ will be active. \\
-        $J_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
-        $a_{j}$   &  resources $i \in \mathcal{I}$ that can be assigned to each task $j$. \\
-        $O_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
-        $T^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - m+1\}},  ..., {t}$\}. \\
+        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ will be active. \\
+        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
+        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to each task $j$. \\
+        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
+        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - m+1\}},  ..., {t}$\}. \\
     \end{tabular}
@@
     \begin{tabular}{p{8mm}p{67mm}}
-        $A_{jti}$   &  1 if task $j \in J$ in period $t\in T_j$ is realized with resource $i \in a_j$, 0 otherwise. \\  
+        $A_{jti}$   &  1 if task $j \in J$ in period $t\in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
         $M_{it}$    &  1 if resource $i \in I$ starts maintenance in period $t \in \mathcal{T}$, 0 otherwise. \\
@@
     \end{align}
-    where relative weight $w$ is chosen by the decision maker. 
+    where weights $w1$ and $w2$ are chosen by the decision maker. 
     The following constraints are used in the model:       
@@
         % maximum capacity1:
-        & \sum_{t' \in T^{s}_t} \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \leq M_{max}
+        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \leq M_{max}
           & t \in \mathcal{T} \label{eq:capacity1}\\
@@
                %avail
-       & \sum_{t' \in T^{s}_t} \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \notag \\
+       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \notag \\
            &\hspace*{10mm}+ d_t\leq U_{max} 
@@
         % & t =m, ..., |\mathcal{T}|  \label{eq:avalaibility2}\\
-        & \sum_{i \in a_j} A_{jti} = r_{j}
-                & j \in \mathcal{J}, t \in T_j  \label{eq:taskres}\\
-        & \sum_{t' \in T^{s}_t} M_{it'} + \sum_{j \in J_t \cap O_i} A_{jti} \leq 1 
+        & \sum_{i \in \mathcal{I}_j} A_{jti} = r_{j}
+                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
+        & \sum_{t' \in \mathcal{T}^{s}_t} M_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} A_{jti} \leq 1 
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
@@
     Maintenance capacity is controlled by (\ref{eq:capacity1}).
-The number of unavailable resources is defined by (\ref{eq:avalaibility1})).
+The number of unavailable resources is defined by (\ref{eq:avalaibility1}).
 Tasks' resource requirements are defined by (\ref{eq:taskres}).
@@
          & rut_{it} \leq rut_{it-1} + H M_{it} \notag \\ 
-           &\hspace*{4mm}- \sum_{j \in J_t \cap O_i} A_{jti} h_{j} \notag \\
+           &\hspace*{4mm}- \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} A_{jti} h_{j} \notag \\
                 && t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
@@
         & ret_{it} \in [0,me]
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}
+                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
+        & \sum_{i \in \mathcal{I}} ret_{it} \geq ret^{Init}_{sum}
+              & t = |\mathcal{T}| \label{eq:min_ret}\\
+        & \sum_{i \in \mathcal{I}} rut_{it} \geq rut^{Init}_{sum}
+              & t = |\mathcal{T}| \label{eq:min_rut}
     \end{align}
@@
     Similarly, the remaining elapsed time is defined by (\ref{eq:ret_upper})-(\ref{eq:ret_initial}) and its limits by (\ref{eq:ret_lower})-(\ref{eq:me}). 
+    Finally, \ref{eq:min_ret} and \ref{eq:min_rut} guarantee that resources have, globally, the same number of remaining used and elapsed times that at the beginning.
 
 \section{Datasets}
-
-  The dataset that was used was provided by the French Direction Générale de l'Armement (DGA).
-  Based on the information provided, a number of datasets were created to run different tests.
+  \label{sec:data}
+  The dataset that was used was provided by the French Airforce and already used by the French Defense Procurement Agency in \cite{Chastellux2017}.
+  Based on the information provided, a number of datasets were created with subsets of the main dataset in order to run different tests.
 
@@
     \begin{table}
-    \begin{tabular}{lrrr}
-    \toprule
-    {} &  usage (h) &  \# resource &  \# candidates \\
-    \midrule
-    O1  &         80 &           8 &            50 \\
-    O10 &         23 &          50 &            74 \\
-    O11 &         23 &           6 &             8 \\
-    O12 &         23 &           8 &            16 \\
-    O2  &         60 &           2 &            50 \\
-    O3  &         50 &           2 &            16 \\
-    O4  &         50 &           2 &            50 \\
-    O5  &         23 &           4 &            29 \\
-    O6  &         23 &          20 &            29 \\
-    O7  &         23 &           1 &            16 \\
-    O8  &         23 &           4 &             3 \\
-    O9  &         23 &           1 &            29 \\
-    \bottomrule
-    \end{tabular}
-    \caption{Tasks' comparative table}
+
+    \input{tables/tasks.tex}
+
+    \caption{Tasks' main characteristics including number of candidate resources and monthly needs}
     \vspace{-0.5cm}
@@
           \includegraphics[width=0.5\textwidth]{./../../img/initial_used.png}
-        \caption{Distribution of Initial Usage Times among resources}
+        \caption{Distribution of Usage Times among resources at the beginning of the planning horizon}
         \label{fig:histogramUsage}
@@
 
-\section{Heuristics}
+% \section{Heuristics}
 
 \section{Results}
+  \label{sec:results}
 
@@
 
+  Table \ref{tab:summary} gives an overview on the results obtained for each of the executed instances.
+
   Both weights on the objective function were set at 1. For a sensibility analysis on these two parameters, see section \ref{sec:multi}.
@@
 
-\begin{tabular}{lrrrrrr}
-\toprule
-   id &  $|\mathcal{T}|$ &  assign &  $|\mathcal{J}|$ &  vars &  cons &  nonzeros \\
-\midrule
- 2321 &               11 &    1080 &               11 &  7270 &  6329 &     41266 \\
- 1839 &               41 &    2069 &               10 & 26171 & 25062 &    169532 \\
- 2127 &               31 &    1610 &               10 & 19789 & 18750 &    126376 \\
-\bottomrule
-\end{tabular}
+  \input{tables/instance.tex}
 
-  \caption{Instances' description}
+  \caption{Instances used in the experiments with indicators of size}
   \vspace{-0.5cm}
@@
 
+    \begin{table}[h]
+  \begin{center}
+
+  \input{tables/summary.tex}
+
+  \caption{Solution information for each tested instance}
+  \vspace{-0.5cm}
+  \label{tab:summary}
+  \end{center}
+  \end{table}
+
     Table \ref{tab:results} show the details of the solving process for each instance.
+As can be seen, cuts do not improve significantly the relaxation (bound cuts is very close to the root solution).
+In fact, the original lineal relaxation (root) is quite similar to the final lower bound (bound).
+This suggests further improvement in more compact modelization could help reduce the gap faster.
+
+    The biggest amount of cuts that were used were 'Implied bound' and 'Mixed integer rounding', depending on the instance.
 
@@
 
-\begin{tabular}{lrrrr}
-\toprule
-   id &  gap (\%) &    OF &  time (s) &  bound \\
-\midrule
- 2321 &      0.7 & 139.0 &       500 &  138.0 \\
- 1839 &     90.5 & 855.0 &      3600 &   81.7 \\
- 2127 &      1.2 &  83.0 &      3600 &   82.0 \\
-\bottomrule
-\end{tabular}
+  \input{tables/results.tex}
 
-  \caption{Solutions' description}
+  \caption{Solution details on the progress of the lower bound}
   \vspace{-0.5cm}
@@
 
-  \section{Multiobjective analysis}
+  \subsection{Bi-objective analysis}
   \label{sec:multi}
 
-    In order to understand the sensibility of the weights $w1$ and $w2$ on the solution, a multiobjective analysis was done.
+    In order to understand the sensibility of the weights $w1$ and $w2$ on the solution, a bi-objective analysis was done.
 In it, both weights were changed between 0 and 1 in intervals of 0.1 so that they always summed 1.
 Table \ref{tab:multiobj} shows the obtained results for each instance.
-The base instance that was used was the one of 31 periods and 10 tasks.
+The base instance that was used for this comparison was the one of 31 periods and 10 tasks corresponding to id 2127.
 
-    The instances were not solved until optimum value but it can be seen that the remaining gap is very low (1.5\% maximum).
+    The instances were stopped after 3600 seconds and most did not solve until optimality. Nevertheless, it can be seen that the remaining gap is very low (1.5\% maximum).
 
-    In figure \ref{fig:pareto} it's possible to see that solutions W01, W05, W06 and W09 dominate the other solutions, having the latter three the same objective function.
+    In figure \ref{fig:pareto} it is possible to see that solutions W01, W05, W06 and W09 dominate the rest, having the latter three the same objective function.
+
+    One last test was made by constraining the solution to be the intermediate point (15, 67) shown in the graph. After 3600 seconds, an integer solution was not found.
 
@@
 
-\begin{tabular}{lrrrrr}
-\toprule
-  id &  maint &   w2 &   w1 &  unavailable &  gap (\%) \\
-\midrule
- W00 &     21 &  1.0 &  0.0 &           67 &     1.49 \\
- W01 &     16 &  0.9 &  0.1 &           67 &     0.00 \\
- W10 &     15 &  0.0 &  1.0 &           69 &     0.00 \\
- W02 &     19 &  0.8 &  0.2 &           67 &     0.35 \\
- W03 &     17 &  0.7 &  0.3 &           68 &     0.58 \\
- W04 &     15 &  0.6 &  0.4 &           69 &     0.43 \\
- W05 &     15 &  0.5 &  0.5 &           68 &     1.20 \\
- W06 &     15 &  0.4 &  0.6 &           68 &     1.10 \\
- W07 &     16 &  0.3 &  0.7 &           70 &     0.97 \\
- W08 &     16 &  0.2 &  0.8 &           70 &     0.78 \\
- W09 &     15 &  0.1 &  0.9 &           68 &     0.49 \\
-\bottomrule
-\end{tabular}
+  \input{tables/multiobj.tex}
 
-  \caption{Multiobjective comparison}
+  \caption{Comparison of instances where variations on the weight of its two objectives has been done}
   \vspace{-0.5cm}
@@
       \includegraphics[width=0.5\textwidth]{./../../img/multiobjective.png}
-        \caption{Pareto diagram of weight combinations}
+    \caption{Pareto diagram comparing the quality of both objectives for each instance}
     \label{fig:pareto}
@@
 
-\section{Conclusions}
+\section{Conclusions and further work}
+  \label{sec:conclusions}
 
-  The results show that a direct exact method to solve the whole dataset is not ideal.
+  The results show that a direct exact method to solve the whole dataset may not be ideal.
+Although medium size instances have been solved, the complete dataset is not solved in a reasonable time for more horizons bigger than 10 periods.
+Still, the analysis on the progress of the lower bound shows that further work can be done to improve the model in order for it to have a better relaxation.
+One possibility is replacing the constraints that count the remaining usage time and elapsed time for more tighter constraints.
+Another possibility is slightly changing the objective function so it can measure more gradually the smoothness of the objective and help get better lower bounds.
+
+  Regarding the multi-objective nature of the problem, the bi-objective analysis appears to indicate the existence of a small number of solutions that are Pareto optimal.
+If this is the case, an iterative approach that explores these solutions could be done.
+Further tests with different, bigger, instances could reveal if this is the case.
+Improvement in the mono-objective modeling could also lead to producing non-dominated solutions.
+In addition to this, new objectives could be included into the analysis, such as smoothing the remaining usage time among periods instead of doing it only in the last period.
+
+  Another possible reason for the difficult in solving big instances could be the quantity of symmetries among candidates for each task: there exist a very big number of possible candidates to be potentially assigned for each task with respect with the actual needed resources of the task.
+This could be handled by assigning priorities to some combinations of resource-task or by limiting the number of candidates that can be assign to each task by some heuristic that clusters resources.
 

### 2018-01-14 python/package/data_input.py
changed article; tables and executions

@@ -293,18 +293,20 @@ def get_log_info_cplex(path):
     result = re.findall(regex, content)
     if result:
         presolve['time'] = result[0]
-
     regex = r'^\*?\s+{0}\s+{0}\s+{0}?\s+{0}?\s+{0}?\s+(Cuts: )?{0}?\s+{0}\s+{0}?\%?$'.format(numberSearch)
     result = re.findall(regex, content, flags=re.MULTILINE)
     after_cuts = -1
     first_relax = -1
+    progress = pd.DataFrame()
     if result:
         first_relax = float(result[0][2])
         for r in result:
             if r[0] == '0' and r[1] == '2':
                 after_cuts = float(r[2])
                 break
-
+        relax, obj, it = zip(*[[r[2], r[4], r[7]] for r in result if
+                             r[2] != '' if r[4] != '' if r[6] != ''])
+        progress = pd.DataFrame({'relax': relax, 'obj': obj, 'it': it})
     return {
         'bound_out': bound,
         'objective_out': objective,
@@ -317,7 +319,9 @@ def get_log_info_cplex(path):
         'cutsTime': cutsTime,
         'presolve': presolve,
         'first_relaxed': first_relax,
-        'after_cuts': after_cuts
+        'after_cuts': after_cuts,
+        'progress': progress
+
     }
 
 

### 2018-01-14 python/scripts/exec_parallel.py

@@ -44,12 +44,12 @@ model_data = di.combine_data_states(model_data, historic_data)
 
 # this is for testing purposes:
 num_start_period = 0
-num_max_periods = 30
+num_max_periods = 20
 model_data['parameters']['start'] = \
     aux.shift_month(model_data['parameters']['start'], num_start_period)
 model_data['parameters']['end'] = \
     aux.shift_month(model_data['parameters']['start'], num_max_periods)
-forbidden_tasks = ['O10', 'O8']
+forbidden_tasks = ['O10', 'O8', '06']
 # forbidden_tasks = ['O8']  # this task has less candidates than what it asks.
 model_data['tasks'] = \
     {k: v for k, v in model_data['tasks'].items() if k not in forbidden_tasks}
@@ -58,13 +58,13 @@ model_data['tasks'] = \
 instance = inst.Instance(model_data)
 
 def_options = {
-    'timeLimit': 7200
+    'timeLimit': 3600
     , 'gap': 0
     , 'solver': "CPLEX"
     , 'path':
         '/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/weights2/'
     , "model": "no_states"
-    , "comments": "periods 0 to 30 without tasks: O10, O8"
+    # , "comments": "periods 0 to 30 without tasks: O10, O8"
 }
 
 # solving part:
@@ -73,10 +73,25 @@ weight_options = [round(0.1*value, 1) for value in range(11)]
 
 if __name__ == "__main__":
 
-    time = datetime.datetime.now()
-    working(instance, def_options, weight_options)
-    duration = datetime.datetime.now() - time
-    print("The time it took = {} seconds".format(duration.seconds))
+    # time = datetime.datetime.now()
+    # working(instance, def_options, weight_options)
+    # duration = datetime.datetime.now() - time
+    # print("The time it took = {} seconds".format(duration.seconds))
 
+    results = {}
 
+    for pos, w in enumerate(weight_options):
+        name = str(int(w * 10))
+        instance.data["parameters"]['maint_weight'] = w
+        instance.data["parameters"]['unavail_weight'] = 1 - w
+        options = dict(def_options)
+        options["weights"] = {'maint_weight': w, 'unavail_weight': 1 - w}
+        options["path"] += name + '/'
+        di.export_data(options['path'], instance.data, name="data_in", file_type='json')
+        di.export_data(options['path'], options, name="options", file_type='json')
+
+        # solving part:
[…]

### 2018-01-14 python/scripts/graphs2.py

@@ -130,7 +130,7 @@ if __name__ == "__main__":
     # Multiobjective
     ################################################
     # paths = os.listdir()
-    path_comp = path_abs + "weights/"
+    path_comp = path_abs + "weights2/"
     exps = exp.list_experiments(path_comp)
     # pp.pprint(exps)
 

### 2018-01-14 latex/MOSIM2018/article.tex
changed article; tables and executions
added instances, graphs and corrections. Changes Pareto analysis. Added progress graph.

@@
 
-\section{Related work}
-  \label{sec:soa}
-
-  % TODO: add FMP biblio?
-
-  % The Military Flight and Maintenance Planning Problem considered here aims to assign missions and maintenance tasks to military aircraft.
-It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically since a round-trip to the base is assumed for each flight.
-
-  % It also includes different objectives and constraints.
-
-  % Although the former has been studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015}, it has not received as much attention as the latter.
-In the following, the problem with already scheduled missions is considered.
-
-  In \cite{Kozanidis2008}, a model was made for the Hellenic Air Force.
-It assigned flight hours to aircraft and scheduled flexible maintenance periods based on a "remaining usage time" rule.
-The objectives were to maximize this "remaining usage time" in the aircraft as well as to smooth the availability of aircraft during the planning period.
-Instances of 24 aircraft and 6 monthly periods were tested with Mixed Integer Programming (MIP) formulations and heuristics.
-
-  In \cite{Cho2011}, the objective was to smooth the maintenances as much as possible.
-Randomly generated instances of 15 aircraft and 520 bi-daily periods were tested with MIP formulations.
-The final state of aircraft was taken into account in order to distribute as much as possible the "remaining usage time" of aircraft.
-
-  The concept of mission is not introduced as part of these formulations, only a demand of flight hours is given for all aircraft in total.
-The aircraft themselves have been assumed as homogeneous and interchangeable.
-Due to this aggregation of demand, there are no minimal time assignments for aircraft to missions.
-Missions are important in order to guarantee that aircraft have the required capabilities needed for the missions it is assigned to.
-
-  % Additionally, the possibility of storage of aircraft is not present in the previous work.
-
-  Lastly, in \cite{Chastellux2017} a mission-centered modeled is presented.
-Here, cycles, composed of one maintenance followed by mission assignments were used.
-Instances of 6 - 70 aircraft and 60 monthly periods were tested, founding close to optimal solutions for small to medium instances using MIP formulations and heuristics based on these MIP formulations.
-
-  % Here, an alternative model is presented that simplifies the formulation and should permit the inclusion of the rest of constraints without sacrificing optimality.
-
 \section{Problem statement}
@@
           \includegraphics[width=0.5\textwidth]{./../../img/calendar.png}
-        \caption{Visual representation of part of the solution to an instance of the problem showing maintenances and assignments}
+        \caption{Visual representation of part of the solution to an instance of the problem showing maintenances and assignments.
+Resources are listed in the vertical axis and the horizontal axis defines the time.
+Red periods correspond to task assignments while blue ones to maintenance assignments.}
         \label{fig:solution}
@@
 
+\section{Related work}
+  \label{sec:soa}
+
+  % TODO: add FMP biblio?
+
+  % The Military Flight and Maintenance Planning Problem considered here aims to assign missions and maintenance tasks to military aircraft.
+It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically since a round-trip to the base is assumed for each flight.
+
+  % It also includes different objectives and constraints.
+
+  % Although the former has been studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015}, it has not received as much attention as the latter.
+In the following, the problem with already scheduled missions is considered.
+
+  In \cite{Kozanidis2008}, a model was made for the Hellenic Air Force.
+It assigned flight hours to aircraft and scheduled flexible maintenance periods based on a "remaining usage time" rule.
+The objectives were to maximize this "remaining usage time" in the aircraft as well as to smooth the availability of aircraft during the planning period.
+Instances of 24 aircraft and 6 monthly periods were tested with Mixed Integer Programming (MIP) formulations and heuristics.
+
+  In \cite{Cho2011}, the objective was to smooth the maintenances as much as possible.
+Randomly generated instances of 15 aircraft and 520 bi-daily periods were tested with MIP formulations.
+The final state of aircraft was taken into account in order to distribute as much as possible the "remaining usage time" of aircraft.
+
+  The concept of mission is not introduced as part of these formulations, only a demand of flight hours is given for all aircraft in total.
+The aircraft themselves have been assumed as homogeneous and interchangeable.
+Due to this aggregation of demand, there are no minimal time assignments for aircraft to missions.
+Missions are important in order to guarantee that aircraft have the required capabilities needed for the missions they are assigned to.
+
+  % Additionally, the possibility of storage of aircraft is not present in the previous work.
+
+  Lastly, in \cite{Chastellux2017} a mission-centered modeled is presented.
+Here, cycles, composed of one maintenance followed by mission assignments were used.
+Instances of 6 - 70 aircraft and 60 monthly periods were tested, founding close to optimal solutions for small to medium instances using MIP formulations and heuristics based on these MIP formulations.
+
+  % Here, an alternative model is presented that simplifies the formulation and should permit the inclusion of the rest of constraints without sacrificing optimality.
+
 \section{Optimization model}
@@
         % $S$               & remaining storage time after (certain) maintenances. \\
-        $w1$               & weight of smoothness for maintenances. \\
-        $w2$               & weight of smoothness for unavailabilities. \\
+        $w_1$               & weight of smoothness for maintenances. \\
+        $w_2$               & weight of smoothness for unavailabilities. \\
         $n_t$             & number of resources in already-planned maintenance in period $t$.\\
@@
     \begin{align}
-        & \text{Min}\; w1 \times M_{max} + w2 \times U_{max}
+        & \text{Min}\; w_1 \times M_{max} + w_2 \times U_{max}
     \end{align}
-    where weights $w1$ and $w2$ are chosen by the decision maker. 
+    where weights $w_1$ and $w_2$ are chosen by the decision maker. 
     The following constraints are used in the model:       
@@
 
-    \caption{Tasks' main characteristics including number of candidate resources and monthly needs}
+    \caption{Tasks' main characteristics including number of candidate resources and monthly needs.
+It reads in the following way: Task 'O1' consumes 80 flight hours of each of 8 resources each month.
+These resources need to be chosen from a pool of 50 possible candidates.}
     \vspace{-0.5cm}
@@
 
-  \caption{Instances used in the experiments with indicators of size}
+  \caption{Instances used in the experiments with indicators of size.
+"id" is the instance; $|\mathcal{T}|$ the number of periods; "assign" the number of assignments of resources to tasks; $|\mathcal{J}|$ the number of tasks; "vars" the number of variables; "cons" the number of constraints; and "nonzeros" the number of non zero values in the matrix}
   \vspace{-0.5cm}
@@
 
-  \caption{Solution information for each tested instance}
+  \caption{Solution information for each tested instance.
+"id" is the instance identifier; "objective" the best integer solution found; "gap" the percentage gap between the integer solution and the lineal relaxation; "time (s)" the time limit that was chosen; and "bound" the lower bound from the lineal relaxation}
   \vspace{-0.5cm}
@@
     The biggest amount of cuts that were used were 'Implied bound' and 'Mixed integer rounding', depending on the instance.
+Being these two generic cuts, it appears the solver has not detected any particular structure as part of this problem.
 
@@
 
-  \caption{Solution details on the progress of the lower bound}
+  \caption{Solution details on the progress of the lower bound.
+"id" shows the name of the instance; "root" the relaxation before cuts; "b.
+cuts" the relaxation after the cuts; "bound" the relaxation at the time limit; "cuts (\#)" the number of cuts; and "cuts (s)" the time the cuts took in seconds}
   \vspace{-0.5cm}
@@
 
+  \begin{table}[h]
+  \begin{center}
+
+  \input{tables/results2.tex}
+
+  \caption{Solution details on the progress of the integer solution.
+"id" shows the name of the instance; "first" is the first solution integer found; "sol.
+cuts" the solution after the cuts; "last" is the last solution found at the time limit; "cuts (\#)" the number of cuts; and "cuts (s)" the time the cuts took in seconds}
+  \vspace{-0.5cm}
+  \label{tab:results2}
+  \end{center}
+  \end{table}
+
+  Figure \ref{fig:progress} shows an example of progress for instance with id=1817.
+
+  \begin{figure}
+    \centering
+      \includegraphics[width=0.5\textwidth]{./../../img/progress.png}
+    \caption{Solving progress of the integer solution (in blue) and the lower bound (red) along the solver iterations (x axis). Instance id=1817.}
+    \label{fig:progress}
+  \end{figure}
+
   \subsection{Bi-objective analysis}
@@
 
-    In order to understand the sensibility of the weights $w1$ and $w2$ on the solution, a bi-objective analysis was done.
+    In order to understand the sensibility of the weights $w_1$ and $w_2$ on the solution, a bi-objective analysis was done.
 In it, both weights were changed between 0 and 1 in intervals of 0.1 so that they always summed 1.
 Table \ref{tab:multiobj} shows the obtained results for each instance.
-The base instance that was used for this comparison was the one of 31 periods and 10 tasks corresponding to id 2127.
+The base instance that was used for this comparison was the one of 21 periods and 9 tasks corresponding to id 1331.
 
-    The instances were stopped after 3600 seconds and most did not solve until optimality. Nevertheless, it can be seen that the remaining gap is very low (1.5\% maximum).
+    The instances that were chosen solved up to optimality.
 
-    In figure \ref{fig:pareto} it is possible to see that solutions W01, W05, W06 and W09 dominate the rest, having the latter three the same objective function.
-
-    One last test was made by constraining the solution to be the intermediate point (15, 67) shown in the graph. After 3600 seconds, an integer solution was not found.
+    In figure \ref{fig:pareto} it is possible to see that there are at least three Pareto optimal points.
+Further analysis must be done with different instances to obtain insights on the relation between the two objectives.
 
@@
 
-  \caption{Comparison of instances where variations on the weight of its two objectives has been done}
+  \caption{Comparison of instances where variations on the weight of its two objectives has been done. Instance id=1331.}
   \vspace{-0.5cm}
@@
       \includegraphics[width=0.5\textwidth]{./../../img/multiobjective.png}
-    \caption{Pareto diagram comparing the quality of both objectives for each instance}
+    \caption{Pareto diagram comparing the quality of both objectives for each instance. Instance id=1331.}
     \label{fig:pareto}
@@
 If this is the case, an iterative approach that explores these solutions could be done.
-Further tests with different, bigger, instances could reveal if this is the case.
+To confirm the true number of solutions, an analysis using epsilon constrains should follow.
+Further tests with different, bigger, instances could also reveal more information regarding the relation between the two objectives.
 Improvement in the mono-objective modeling could also lead to producing non-dominated solutions.
@@
   Another possible reason for the difficult in solving big instances could be the quantity of symmetries among candidates for each task: there exist a very big number of possible candidates to be potentially assigned for each task with respect with the actual needed resources of the task.
-This could be handled by assigning priorities to some combinations of resource-task or by limiting the number of candidates that can be assign to each task by some heuristic that clusters resources.
+This could be handled by assigning priorities to some combinations of resource-task or by limiting the number of candidates that can be assigned to each task by some heuristic that clusters resources with tasks.
+
+  As new constraints are added, such as minimum durations for assignments and the possibility to "store" resources for a predefined time, the present instances will grow bigger still.
+This, of course, will require new better techniques to decompose the problem into smaller problems using heuristics or take a decomposition approach such as Column Generation.
 

### 2018-01-14 python/package/data_input.py

@@ -242,7 +242,8 @@ def get_log_info_gurobi(path):
 
 
 def get_log_info_cplex(path):
-    # path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201801110004/results.log"
+    # path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201801141331/results.log"
+    # path = '/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201801131817/results.log'
     with open(path, 'r') as f:
         content = f.read()
     numberSearch = r'([\de\.\+]+)'
@@ -297,16 +298,21 @@ def get_log_info_cplex(path):
     result = re.findall(regex, content, flags=re.MULTILINE)
     after_cuts = -1
     first_relax = -1
+    first_solution = -1
+    sol_after_cuts = -1
     progress = pd.DataFrame()
     if result:
         first_relax = float(result[0][2])
         for r in result:
             if r[0] == '0' and r[1] == '2':
                 after_cuts = float(r[2])
+                if r[4]:
+                    sol_after_cuts= float(r[4])
                 break
-        relax, obj, it = zip(*[[r[2], r[4], r[7]] for r in result if
-                             r[2] != '' if r[4] != '' if r[6] != ''])
-        progress = pd.DataFrame({'relax': relax, 'obj': obj, 'it': it})
+        for r in result[1:]:
+            if r[4] != "":
+                first_solution = float(r[4])
+                break
     return {
         'bound_out': bound,
         'objective_out': objective,
@@ -320,9 +326,25 @@ def get_log_info_cplex(path):
         'presolve': presolve,
         'first_relaxed': first_relax,
         'after_cuts': after_cuts,
-        'progress': progress
-
+        'progress': progress,
+        'first_solution': first_solution,
+        'sol_after_cuts': sol_after_cuts
     }
+
+
+def get_log_info_cplex_progress(path):
+    # path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201801131817/results.log"
+    with open(path, 'r') as f:
+        content = f.read()
+    numberSearch = r'([\de\.\+]+)'
+    regex = r'^\*?\s+{0}\s+{0}\s+{0}?\s+{0}?\s+{0}?\s+(Cuts: )?{0}?\s+{0}\s+{0}?\%?$'.format(numberSearch)
[…]

### 2018-01-14 python/scripts/excec.py

@@ -20,7 +20,7 @@ if __name__ == "__main__":
         aux.shift_month(model_data['parameters']['start'], num_start_period)
     model_data['parameters']['end'] = \
         aux.shift_month(model_data['parameters']['start'], num_max_periods)
-    forbidden_tasks = ['O10', 'O8', '06']
+    forbidden_tasks = ['O10', 'O8', 'O6']
     # forbidden_tasks = ['O8']  # this task has less candidates than what it asks.
     model_data['tasks'] = \
         {k: v for k, v in model_data['tasks'].items() if k not in forbidden_tasks}

### 2018-01-14 python/scripts/graphs2.py

@@ -55,7 +55,7 @@ if __name__ == "__main__":
 
     plot = ggplot(aes(x='initial_used'), data=data) + geom_histogram() + \
            theme(axis_text=element_text(size=15)) + \
-           xlab("Initial Remaining Usage Time (hours)")
+           xlab(element_text("Initial Remaining Usage Time (hours)", size=15))
     plot.save(path_img + 'initial_used.png')
 
     ################################################
@@ -95,9 +95,14 @@ if __name__ == "__main__":
                    'gap_out': 'gap (%)', 'bound_out': 'bound'
                    }
     cols_rename2 = {
-        'index': 'id', 'after_cuts': 'bound cuts',
+        'index': 'id', 'after_cuts': 'b. cuts',
         'first_relaxed': 'root', '# cuts': 'cuts (#)',
         'cutsTime': 'cuts (s)', 'bound_out': 'bound'}
+
+    cols_rename3 = {
+        'index': 'id', 'sol_after_cuts': 'sol. cuts',
+        'first_solution': 'first', 'objective_out': 'last'
+    }
 
     table = pd.DataFrame.from_dict(exps, orient="index")
     table = table[np.all((table.model == 'no_states',
@@ -112,17 +117,24 @@ if __name__ == "__main__":
                    ][0])
     table['index'] = table['index'].str.slice(8)
 
-    table1 = table.rename(columns=cols_rename2)[
-        ['id', 'root', 'bound cuts',
+    table2 = table.rename(columns=cols_rename2)[
+        ['id', 'root', 'b. cuts',
          'bound', 'cuts (#)', 'cuts (s)']]
-    latex = table1.to_latex(bold_rows=True, index=False, float_format='%.1f')
+    latex = table2.to_latex(bold_rows=True, index=False, float_format='%.1f')
     with open(path_latex + 'MOSIM2018/tables/results.tex', 'w') as f:
         f.write(latex)
 
-    table2 = table.rename(columns=cols_rename)[
+    table1 = table.rename(columns=cols_rename)[
         ['id', 'objective', 'gap (%)', 'time (s)', 'bound']]
-    latex = table2.to_latex(bold_rows=True, index=False, float_format='%.1f')
+    latex = table1.to_latex(bold_rows=True, index=False, float_format='%.1f')
     with open(path_latex + 'MOSIM2018/tables/summary.tex', 'w') as f:
+        f.write(latex)
+
+    table3 = table.rename(columns=cols_rename3)[
[…]

### 2018-01-14 latex/MOSIM2018/article.tex
modifications to result tables, added commentary to MOSIM article and changes to heuristic

@@
 
-  The Military Flight and Maintenance Planning problem considered assigns missions and schedules maintenances to aircraft.
-In this problem, aircraft are needed to comply with a series of planned missions while at the same time needing to receive frequent preventive maintenances in order to be ready for new missions.
+  This paper presents a specific version of the Flight and Maintenance Problem, which we call Military Flight and Maintenance Problem.
+We provide a description of the problem, a mathematical formulation to obtain quality solutions for a bi-objective optimization function and experimental results to show both the quality of the solutions and the limitations of the technique.
 
-  It is a variant of the Flight and Maintenance Planning problem, where maintenances and flights are scheduled to commercial aircraft.
-The main differences are the fact that flights are not modeled geographically: all aircraft are assumed to exit the main depot and return to the same depot at the end of the flight.
+  The article is structured as follows:
 
-  Other differences include the following.
+  Section \ref{sec:problem} describes with detail the problem, its different elements, conditions and objectives.
+We also make a review of previous work done in the subject in section \ref{sec:soa} where we compare the addressed problems with our own and explain the differences.
 
-  First of all, the objective function's motivation is on reducing the load on the maintenance facilities and guaranteeing the availability of the fleet instead of reducing the cost.
-Also, the maintenance operations take considerably longer and the rules that govern the maintenances are different.
+  Later, in section \ref{sec:data}, the specific dataset at hand is described.
+Section \ref{sec:model} presents a complete MIP formulation to solve the described problem and section \ref{sec:results} shows the commented results obtained while solving several instances of the problem based on the provided dataset.
 
-  The temporal scope of the problem is also bigger, making the normal planning horizon bigger and each period bigger too: months instead of days.
-On the other hand, the physical scope, while present in the commercial planning, is not present in the military one: aircraft are assumed to return to the origin every time.
-
-  In section \ref{sec:soa} we present previous work in this domain.
-Section \ref{sec:problem} describes with detail the problem, its different elements, conditions and objectives, while the particular dataset used in this study is described in \ref{sec:data}.
-Section \ref{sec:model} presents a complete MIP formulation to solve the described problem.
-
-  Section \ref{sec:results} shows the results obtained while solving several instances of the problem and finally, section \ref{sec:conclusions} mentions the conclusions obtained from the experimentation and the possible direction of further work in this domain.
+  Finally, section \ref{sec:conclusions} mentions the conclusions obtained from the experimentation and the possible direction of further work.
 
@@
 
+  The Military Flight and Maintenance Planning problem considered assigns missions and schedules maintenances to aircraft.
+In this problem, aircraft are needed to comply with a series of planned missions while at the same time needing to receive frequent preventive maintenances in order to be ready for new missions.
+
+  It is a variant of the Flight and Maintenance Planning problem, where maintenances and flights are scheduled to commercial aircraft.
+The main differences are the fact that flights are not modeled geographically: all aircraft are assumed to exit the main depot and return to the same depot at the end of the flight.
+
+  Other differences include the following.
+
+  First of all, the objective function's motivation is on reducing the load on the maintenance facilities and guaranteeing the availability of the fleet instead of reducing the cost.
+Also, the maintenance operations take considerably longer and the rules that govern the maintenances are different.
+
+  The temporal scope of the problem is also bigger, making the normal planning horizon bigger and each period bigger too: months instead of days.
+On the other hand, the physical scope, while present in the commercial planning, is not present in the military one: aircraft are assumed to return to the origin every time.
+
   In \cite{Kozanidis2008}, a model was made for the Hellenic Air Force.
@@
 
-  % Here, an alternative model is presented that simplifies the formulation and should permit the inclusion of the rest of constraints without sacrificing optimality.
+  Here, an alternative model is presented that simplifies the formulation and should permit the inclusion of the rest of constraints without sacrificing optimality.
 
@@
   \label{sec:data}
-  The dataset that was used was provided by the French Airforce and already used by the French Defense Procurement Agency in \cite{Chastellux2017}.
-  Based on the information provided, a number of datasets were created with subsets of the main dataset in order to run different tests.
+  The dataset being used was provided by the French Airforce and already used by the French Defense Procurement Agency in \cite{Chastellux2017}.
+  Based on the information provided, a number of instances were created as subsets of the main dataset in order to run different experiments.
 
@@
 
-    The main characteristics of the dataset are the following. 11 tasks, 127 resources.
+    There are a total of 12 tasks.
     Most tasks demand between 1 and 8 resources per month. The biggest and second biggest tasks demand 50 and 20 resources respectively each month.
-    More details in table \ref{tab:tasks}.
+    More details of task differences are found in table \ref{tab:tasks}.
     Candidates are calculated based on each task's requirements matching the candidate's capabilities.
@@
     All maintenances take 6 months (periods).
-Each maintenance provides with 1000 hours of usage time to the resource and a resource can be without maintenance for a maximum of 60 periods.
+Each maintenance provides with 1000 hours of usage time to the resource and a resource can be without maintenance for a maximum of 60 consecutive periods.
 The nominative capacity for maintenance is 18 resources at any given time.
@@
 
+    There are 127 resources in total. Each one with its functionalities and past assignments.
     The distribution of initial usage times is shown in the figure \ref{fig:histogramUsage}. It can be seen that the aircraft do not necessarily follow an horizontal line.
@@
   All tests were run on a 64 bit Ubuntu 16.04 workstation with 16GB of RAM and Intel i7-6700HQ CPU \@ 2.60GHz x 8 processor.
-  Several experiments were done with the MIP formulation. Table \ref{tab:instance} shows the characteristics of each instance. 
-  Most of the columns describe the size of the instance: number of periods, number of tasks and number of assignments.
-  The assign column indicates the number of assignments that are needed to be done to all the task-periods.
+  Several experiments were done with the MIP formulation.
+Table \ref{tab:instance} shows the characteristics of each experiment.
+The main differences between instances is the number of tasks and the number of periods.
 
   Table \ref{tab:summary} gives an overview on the results obtained for each of the executed instances.
+As can be seen, smaller instances are solved up to optimality while medium size ones are stopped after one hour.
 
@@
 "id" shows the name of the instance; "root" the relaxation before cuts; "b.
-cuts" the relaxation after the cuts; "bound" the relaxation at the time limit; "cuts (\#)" the number of cuts; and "cuts (s)" the time the cuts took in seconds}
+cuts" the relaxation after the cuts; "bound" the relaxation at the time limit; "cuts (\#)" the number of cuts; and "cuts (s)" the time the cuts took in seconds.}
   \vspace{-0.5cm}
@@
   \end{table}
+
+  The table \ref{tab:results2} show the progress of the integer solution for each instance.
+For more detail of this progress, figure \ref{fig:progress} shows an example of progress for instance with id=1817.
+It can be seen that initial solutions are quite big in comparison with the optimal.
+
+  For two, rather small, instances the solver was able to find an optimal solution before starting to branch. In these two instances, the missing values have been replaced by "-1".
+
+  These results show how the branching is not providing with much improvement in increasing the lower bound or finding better integer solutions.
+This hints into a possible reason for the difficult in solving big instances: the quantity of symmetries in the model, represented as candidates for each task: there exist a very big number of possible candidates to be potentially assigned for each task with respect with the actual needed resources for the task.
+This reduces the capacity of the branching the solver does by making it explore many nodes that are really almost equivalent.
 
@@
 "id" shows the name of the instance; "first" is the first solution integer found; "sol.
-cuts" the solution after the cuts; "last" is the last solution found at the time limit; "cuts (\#)" the number of cuts; and "cuts (s)" the time the cuts took in seconds}
+cuts" the solution after the cuts; "last" is the last solution found at the time limit.}
   \vspace{-0.5cm}
@@
   \end{table}
-
-  Figure \ref{fig:progress} shows an example of progress for instance with id=1817.
 
@@
 
+    The bi-objective analysis appears to indicate the existence of a small number of solutions that are Pareto optimal.
+If this is the case, an iterative approach that explores these solutions could be done.
+Improvement in the mono-objective modeling could also lead to producing non-dominated solutions.
+
   \begin{table}[h]
@@
 
-  \caption{Comparison of instances where variations on the weight of its two objectives has been done. Instance id=1331.}
+  \caption{Comparison of instances where variations on the weight of its two objectives has been done. Instance id=1331. All experiments were solved to optimality.}
   \vspace{-0.5cm}
@@
 Although medium size instances have been solved, the complete dataset is not solved in a reasonable time for more horizons bigger than 10 periods.
-Still, the analysis on the progress of the lower bound shows that further work can be done to improve the model in order for it to have a better relaxation.
-One possibility is replacing the constraints that count the remaining usage time and elapsed time for more tighter constraints.
-Another possibility is slightly changing the objective function so it can measure more gradually the smoothness of the objective and help get better lower bounds.
+Still, the analysis on the progress of the lower bound shows that further work can be done to improve the model in order for it to have a tighter, better relaxation.
+One possibility is to replace the constraints that count the remaining usage time and elapsed time for tighter constraints.
+Another possibility is slightly changing the objective function so it can measure more gradually the smoothness of the objective and help get better bounds.
 
-  Regarding the multi-objective nature of the problem, the bi-objective analysis appears to indicate the existence of a small number of solutions that are Pareto optimal.
-If this is the case, an iterative approach that explores these solutions could be done.
-To confirm the true number of solutions, an analysis using epsilon constrains should follow.
+  Regarding the multi-objective nature of the problem, to confirm the true number of solutions, an analysis using epsilon constrains should follow.
 Further tests with different, bigger, instances could also reveal more information regarding the relation between the two objectives.
-Improvement in the mono-objective modeling could also lead to producing non-dominated solutions.
 In addition to this, new objectives could be included into the analysis, such as smoothing the remaining usage time among periods instead of doing it only in the last period.
 
-  Another possible reason for the difficult in solving big instances could be the quantity of symmetries among candidates for each task: there exist a very big number of possible candidates to be potentially assigned for each task with respect with the actual needed resources of the task.
-This could be handled by assigning priorities to some combinations of resource-task or by limiting the number of candidates that can be assigned to each task by some heuristic that clusters resources with tasks.
+  Regarding the branching and symmetries: this could be handled by assigning priorities to some combinations of resource-task or by limiting the number of candidates that can be assigned to each task by some heuristic that clusters resources with tasks.
+Finally, arbitrarily preassigning (fixing) the final state (in terms of remaining usage time) to each resource could also help break those symmetries.
 
-  As new constraints are added, such as minimum durations for assignments and the possibility to "store" resources for a predefined time, the present instances will grow bigger still.
-This, of course, will require new better techniques to decompose the problem into smaller problems using heuristics or take a decomposition approach such as Column Generation.
+  As new constraints are added, such as minimum durations for assignments and the possibility to "store" resources for a predefined time, it is predicted that the present instances will grow bigger.
+This, of course, will require new and better techniques to decompose the problem into smaller problems using heuristics or to take a decomposition approach such as Column Generation.
+
+  Finally, this work presents a first step useful into the solving of the Military Flight and Maintenance Planning.
+The model presented provides already good initial solutions and also the potential of solving even bigger, more complex problems.
 

### 2018-01-14 python/package/data_input.py

@@ -244,6 +244,7 @@ def get_log_info_gurobi(path):
 def get_log_info_cplex(path):
     # path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201801141331/results.log"
     # path = '/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201801131817/results.log'
+    # path = '/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201801102259/results.log'
     with open(path, 'r') as f:
         content = f.read()
     numberSearch = r'([\de\.\+]+)'
@@ -294,6 +295,11 @@ def get_log_info_cplex(path):
     result = re.findall(regex, content)
     if result:
         presolve['time'] = result[0]
+    regex = r'Solution time =\s+{0} sec\.\s+Iterations = {0}\s+Nodes = {0}'.format(numberSearch)
+    result = re.findall(regex, content)
+    time_out = -1
+    if result:
+        time_out = float(result[0][0])
     regex = r'^\*?\s+{0}\s+{0}\s+{0}?\s+{0}?\s+{0}?\s+(Cuts: )?{0}?\s+{0}\s+{0}?\%?$'.format(numberSearch)
     result = re.findall(regex, content, flags=re.MULTILINE)
     after_cuts = -1
@@ -317,6 +323,7 @@ def get_log_info_cplex(path):
         'bound_out': bound,
         'objective_out': objective,
         'gap_out': gap,
+        'time_out': time_out,
         'cons': int(size.group(1)),
         'vars': int(size.group(2)),
         'nonzeros': int(size.group(3)),

### 2018-01-14 python/package/heuristics.py

@@ -14,15 +14,23 @@ class Greedy(test.Experiment):
         self.solution.data['aux'] = {'ret': {}, 'rut': {}}
         for resource in self.instance.get_resources():
             for time in ['rut', 'ret']:
+                label = 'initial_{}'.format(self.label_rt(time))
+                initial = self.instance.data['resources'][resource][label]
                 label = 'max_{}_time'.format(self.label_rt(time))
+                max_time = self.instance.get_param(label)
+                period = aux.get_prev_month(self.instance.get_param('start'))
+                self.set_remainingtime(resource, period, time, min(max_time, initial))
                 self.update_time_usage(resource,
                                        self.instance.get_periods(),
-                                       previous_value=self.instance.get_param(label),
                                        time=time)
 
         self.options = options
 
     def fill_mission(self, task):
+
+        # label = 'max_' + self.label_rt(time) + '_time'
+        max_rut = self.instance.get_param('max_used_time')
+
         dtype = 'U7'
         # get task candidates: we copy the list.
         candidates = self.instance.data['tasks'][task]['candidates'][:]
@@ -33,6 +41,8 @@ class Greedy(test.Experiment):
             )[task]
 
         while len(rem_resources) > 0 and len(candidates) > 0:
+            # delete periods that have 0 rem_resources to assign
+            rem_resources = {p: v for p, v in rem_resources.items() if v > 0}
             # candidate = 'A124'
             candidate = candidates[0]
             # get free periods for candidate
@@ -54,38 +64,44 @@ class Greedy(test.Experiment):
                         maint_period = period
                         break
                     rem_resources[period] -= 1
-                self.update_time_usage(candidate, period, self.get_next_maintenance(candidate, period), 'rut')
+                    self.update_time_usage(candidate, periods=aux.get_months(period, end), time='rut')
+                    # self.update_time_usage(candidate, period, self.get_next_maintenance(candidate, period), 'rut')
 
-            if maint_period != '':
-                # find soonest period to start maintenance:
-                maint_start = self.get_soonest_maint(candidate, maint_period)
-                if maint_start is None:
-                    # this means that we cannot assign tasks BUT
-                    # we cannot assign maintenance either :/
-                    # we should then take out the candidate:
-                    candidates.pop(0)
-                    continue
-                maint_end = aux.shift_month(maint_start, 5)
[…]

### 2018-01-14 python/scripts/excec.py

@@ -20,8 +20,9 @@ if __name__ == "__main__":
         aux.shift_month(model_data['parameters']['start'], num_start_period)
     model_data['parameters']['end'] = \
         aux.shift_month(model_data['parameters']['start'], num_max_periods)
-    forbidden_tasks = ['O10', 'O8', 'O6']
-    # forbidden_tasks = ['O8']  # this task has less candidates than what it asks.
+    # forbidden_tasks = ['O10', 'O8', 'O6']
+    # forbidden_tasks = ['O10', 'O8']
+    forbidden_tasks = ['O8']  # this task has less candidates than what it asks.
     model_data['tasks'] = \
         {k: v for k, v in model_data['tasks'].items() if k not in forbidden_tasks}
     # this was for testing purposes
@@ -29,7 +30,7 @@ if __name__ == "__main__":
     instance = inst.Instance(model_data)
 
     options = {
-        'timeLimit': 3600
+        'timeLimit': 7200
         , 'gap': 0
         , 'solver': "CPLEX"
         , 'path':
@@ -45,4 +46,3 @@ if __name__ == "__main__":
     solution = md.model_no_states(instance, options)
     if solution is not None:
         di.export_data(options['path'], solution.data, name="data_out", file_type='json')
-

### 2018-01-14 python/scripts/exec_heur.py

@@ -1,0 +1,33 @@
+import package.solution as sol
+import package.heuristics as heur
+import package.data_input as di
+import package.instance as inst
+
+
+if __name__ == "__main__":
+
+    path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201801141334/"
+    model_data = di.load_data(path + "data_in.json")
+    # this was for testing purposes
+
+    instance = inst.Instance(model_data)
+    solution = sol.Solution({'state': {}, 'task': {}})
+    heur = heur.Greedy(instance, solution)
+    # heur.instance.get_tasks()
+    # heur.expand_resource_period(heur.solution.data['task'], 'A100', '2017-01')
+    tasks = list(heur.instance.get_tasks().keys())
+    # tasks = ['O5']
+    for task in tasks:
+        heur.fill_mission(task)
+    rut_old = dict(heur.solution.data['aux']['rut'])
+    # heur.solution.print_solution("/home/pchtsp/Downloads/calendar_temp1.html")
+    checks = heur.check_solution()
+    rut_new = heur.solution.data['aux']['rut']
+    print([k for k, v in checks.items() if len(v) > 0])
+    heur.get_objective_function()
+    # checks['usage']
+    # {k: v for k, v in checks.items() if len(v) > 0}.keys()
+    # checks.keys()
+    # path = "/home/pchtsp/Documents/projects/OPTIMA_documents/results/experiments/201712190002/"
+    # self = Greedy.from_dir(path)
+    #

### 2018-01-14 python/scripts/graphs2.py

@@ -59,39 +59,34 @@ if __name__ == "__main__":
     plot.save(path_img + 'initial_used.png')
 
     ################################################
-    # Instances table
+    # Instances and Results table
     ################################################
 
     paths = os.listdir(path_abs)
     exps = exp.list_experiments(path_abs)
     # pp.pprint(exps)
     # input: num missions, num periods, variables, constraints, assignments, timeLimit
-    input_cols = ['periods', 'assignments', 'tasks', 'vars', 'cons', 'nonzeros']
+    input_cols = ['index', 'tasks', 'periods', 'assignments', 'vars', 'cons', 'nonzeros']
     cols_rename = {'periods': '$|\mathcal{T}|$', 'tasks': '$|\mathcal{J}|$', 'assignments': 'assign',
                    'timeLimit': 'time (s)', 'index': 'id'}
     table = pd.DataFrame.from_dict(exps, orient="index")
     table = table[np.all((table.model == 'no_states',
                   table.gap == 0,
                   table.timeLimit >= 500),
-                 axis=0)] \
-        [input_cols].reset_index().rename(columns=cols_rename)
-    table.id = table.id.str.slice(8)
-    latex = table.to_latex(escape=False, bold_rows=True, index=False, float_format='%.0f')
+                 axis=0)].reset_index().sort_values(['tasks', 'periods'])
+
+    table['index'] = table['index'].str.slice(8)
+
+    table_in = table[input_cols].rename(columns=cols_rename)
+    latex = table_in.to_latex(escape=False, bold_rows=True, index=False, float_format='%.0f')
     with open(path_latex + 'MOSIM2018/tables/instance.tex', 'w') as f:
         f.write(latex)
 
-    ################################################
-    # Results table
-    ################################################
-
-    paths = os.listdir(path_abs)
-    exps = exp.list_experiments(path_abs)
-    # pp.pprint(exps)
     # output: number of cuts, initial relaxation, relaxation after cuts, end relaxation, cuts' time.
     # variable number before and after cuts
     # input_cols = ['timeLimit']
     cols_rename = {
-        'timeLimit': 'time (s)', 'index': 'id', 'objective_out': 'objective',
+        'time_out': 'time (s)', 'index': 'id', 'objective_out': 'objective',
[…]

### 2018-01-15 latex/ROADEF2018/article_roadef_2018_alt.tex
corrected ROADEF article and cleaned old ROADEF article files

@@
 In the following, the problem with already scheduled missions is considered.
+With respect to \cite{Chastellux2017}, a new model will be presented.
 
@@
 
-    Resources require recurrent preventive maintenance operations since the realization of tasks diminish the remaining usage time of resources.
-A resource cannot be used for a task $j$ if its remaining usage time is less that usage $h_j$ required by task $j$.
+    Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
+A resource cannot be used for a task $j$ if its remaining usage time is less than usage $h_j$ required by task $j$.
 A maintenance operation assigns to the resource exactly $H$ units of remaining usage time.

### 2018-01-15 latex/MOSIM2018/article.tex
changes in MOSIM article: renaming of parameters and state of the art

@@
     There is a fixed set of $j \in \mathcal{J}$ tasks to be accomplished over an horizon of time divided into $t \in \mathcal{T}$ discrete periods.
-In order to be accomplished, these tasks require the assignment of a specific number of resources $r_{j}$ each period of time the task is active.
+In order to be accomplished, these tasks require the assignment of a specific number of resources $R_j$ each period of time the task is active.
 Tasks have both a fixed start and end time that define the set of periods of time it is active.
 
-    During each assigned period, tasks use the assigned resources for a time $h_j$.
+    During each assigned period, tasks use the assigned resources for a time $H_j$.
 
     The assignment of a resource to a task is not decided for the whole duration of the task.
-After a minimum amount of time ($mt$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished.
-The total number of resources being used at any given time in a specific task can never be less than the required $r_{j}$.
+After a minimum amount of time ($MT$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished.
+The total number of resources being used at any given time in a specific task can never be less than the required $R_j$.
 
@@
 
-    Maintenances have a fix duration of $m$ periods.
+    Maintenances have a fix duration of $M$ periods.
 
@@
 
+  In \cite{Verhoeff2015}, a model for the Royal Netherlands Air Force was developed.
+Here, three criteria for improving the planning robustness are presented and applied in a MIP model: availability, serviceability and sustainability.
+
+  With respect to heuristic methods for solving this problem, both \cite{Kozanidis2008} and \cite{Cho2011} already presented alternatives to the MIP formulation for solving bigger instances with reduced scope.
+In \cite{steiner2006heuristic}, a constructive, stepped heuristic approach is described to obtain balanced maintenances for the Swiss Air Force.
+Both calendar-based maintenances actions (CBMA) and usage-based maintenance actions (UBMA) are used.
+In \cite{gavranis2015exact}, an exact algorithm was presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
+Finally, in \cite{kozanidis2014heuristics}, two heuristics to solve big instances of the problem by taking consecutive decisions period by period with the help of mathematical models at each step.
+
+  Regarding operational-level planning, a more detailed implementation is found in \cite{safaei2011workforce}, where the specific conditions during the maintenances operations are taken into account to schedule the maintenance jobs, such as the workforce skills and availability.
+
   The concept of mission is not introduced as part of these formulations, only a demand of flight hours is given for all aircraft in total.
@@
 
-  Lastly, in \cite{Chastellux2017} a mission-centered modeled is presented.
+  Lastly, in \cite{Chastellux2016} a mission-centered modeled is presented.
 Here, cycles, composed of one maintenance followed by mission assignments were used.
@@
     \begin{tabular}{p{8mm}p{67mm}}
-        $h_j$             & amount of usage time for resources used in task $j$. \\
-        $r_j$             & amount of resources that task $j$ needs when active. \\
-        $mt$              & minimum number of periods a resource can be assigned to a task. \\
+        $H_j$             & amount of usage time for resources used in task $j$. \\
+        $R_j$             & amount of resources that task $j$ needs when active. \\
+        $MT$              & minimum number of periods a resource can be assigned to a task. \\
         % $aut_i$           & maximal absolute usage time for resource $i$. \\
         % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
-        $m$               & maintenance duration in number of periods. \\
+        $M$               & maintenance duration in number of periods. \\
         $E$               & remaining elapsed time after a maintenance. \\
@@
         % $S$               & remaining storage time after (certain) maintenances. \\
-        $w_1$               & weight of smoothness for maintenances. \\
-        $w_2$               & weight of smoothness for unavailabilities. \\
-        $n_t$             & number of resources in already-planned maintenance in period $t$.\\
-        $d_t$             & number of resources that need to be assigned in total in period $t$. \\
-        $rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning period. \\
-        $ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning period. \\
-        $ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning period. \\
-        $rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning period. \\
+        $W_1$               & weight of smoothness for maintenances. \\
+        $W_2$               & weight of smoothness for unavailabilities. \\
+        $N_t$             & number of resources in already-planned maintenance in period $t$.\\
+        $D_t$             & number of resources that need to be assigned in total in period $t$. \\
+        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning period. \\
+        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning period. \\
+        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning period. \\
+        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning period. \\
     \end{tabular}
@@
     \begin{tabular}{p{8mm}p{67mm}}
-        $A_{jti}$   &  1 if task $j \in J$ in period $t\in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
-        $M_{it}$    &  1 if resource $i \in I$ starts maintenance in period $t \in \mathcal{T}$, 0 otherwise. \\
+        $a_{jti}$   &  1 if task $j \in J$ in period $t\in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
+        $m_{it}$    &  1 if resource $i \in I$ starts maintenance in period $t \in \mathcal{T}$, 0 otherwise. \\
         $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
         $ret_{it}$  &  remaining elapsed time (integer) for resource $i$ after period $t \in \mathcal{T}0$. \\  
-        $U_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
-        $M_{max}$   &  maximal number (integer) of resources in maintenance in any period. \\
+        $u_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
+        $m_{max}$   &  maximal number (integer) of resources in maintenance in any period. \\
     \end{tabular}
 
-    Note that  $A_{jti}$ and $M_{it}$ are initially set up to 0 for all resources already in maintenance at the beginning of the planning period for the remaining time periods of maintenance.
+    Note that  $a_{jti}$ and $m_{it}$ are initially set up to 0 for all resources already in maintenance at the beginning of the planning period for the remaining time periods of maintenance.
 The remaining usage time for each resource at the beginning of the planning period is used to initialize $rut_{i0}$.
@@
     \begin{align}
-        & \text{Min}\; w_1 \times M_{max} + w_2 \times U_{max}
+        & \text{Min}\; W_1 m_{max} + W_2 u_{max}
     \end{align}
-    where weights $w_1$ and $w_2$ are chosen by the decision maker. 
+    where weights $W_1$ and $W_2$ are chosen by the decision maker. 
     The following constraints are used in the model:       
@@
         % maximum capacity1:
-        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \leq M_{max}
+        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
           & t \in \mathcal{T} \label{eq:capacity1}\\
                % maximum capacity2:                
-        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} \leq M_{max}
+        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} m_{it'} \leq m_{max}
         % & t =m, ..., |\mathcal{T}|  \label{eq:capacity2}\\
                %avail
-       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} M_{it'} + n_{t} \notag \\
-           &\hspace*{10mm}+ d_t\leq U_{max} 
+       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \notag \\
+           &\hspace*{10mm}+ D_t\leq u_{max} 
         &t \in \mathcal{T} \label{eq:avalaibility1}\\
                % maximum capacity2:                
-        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} M_{it'} + d_t\leq U_{max} 
+        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} m_{it'} + D_t\leq u_{max} 
         % & t =m, ..., |\mathcal{T}|  \label{eq:avalaibility2}\\
-        & \sum_{i \in \mathcal{I}_j} A_{jti} = r_{j}
+        & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
                 & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
-        & \sum_{t' \in \mathcal{T}^{s}_t} M_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} A_{jti} \leq 1 
+        & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
@@
         % remaining used time
-         & rut_{it} \leq rut_{it-1} + H M_{it} \notag \\ 
-           &\hspace*{4mm}- \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} A_{jti} h_{j} \notag \\
+         & rut_{it} \leq rut_{it-1} + H m_{it} \notag \\ 
+           &\hspace*{4mm}- \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j \notag \\
                 && t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
-        & rut_{i0} = rut^{Init}_i
+        & rut_{i0} = Rut^{Init}_i
                & i \in \mathcal{I} \label{eq:rut_initial}\\
-        & rut_{it} \geq H M_{it}
+        & rut_{it} \geq H m_{it}
                 & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
@@
         & ret_{it} \leq ret_{it-1} - 1 \notag \\
-          &\hspace*{4mm}+ E M_{it}
+          &\hspace*{4mm}+ E m_{it}
                 & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
-        & ret_{i0} = ret^{Init}_r
+        & ret_{i0} = Ret^{Init}_i
                 & i \in \mathcal{I} \label{eq:ret_initial}\\
-        & ret_{it} \geq E M_{it}
+        & ret_{it} \geq E m_{it}
                 & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
@@
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
-        & \sum_{i \in \mathcal{I}} ret_{it} \geq ret^{Init}_{sum}
+        & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
               & t = |\mathcal{T}| \label{eq:min_ret}\\
-        & \sum_{i \in \mathcal{I}} rut_{it} \geq rut^{Init}_{sum}
+        & \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
               & t = |\mathcal{T}| \label{eq:min_rut}
@@
 
-  \begin{table}[h]
+  \begin{table}
   \begin{center}
@@
 
-    \begin{table}[h]
+    \begin{table}
   \begin{center}
@@
 
-    In order to understand the sensibility of the weights $w_1$ and $w_2$ on the solution, a bi-objective analysis was done.
+    In order to understand the sensibility of the weights $W_1$ and $W_2$ on the solution, a bi-objective analysis was done.
 In it, both weights were changed between 0 and 1 in intervals of 0.1 so that they always summed 1.

### 2018-01-19 latex/MOSIM2018/article.tex
changed tables of MOSIM and some changes in article

@@
   Based on the information provided, a number of instances were created as subsets of the main dataset in order to run different experiments.
+The criteria for the creation of the instances was the following:
+
+    \begin{enumerate}
+        \item The base instances was the biggest instance where an integer solution was found in less than 3600 seconds: all 11 tasks and 31 periods.
+        \item Take out the task with the biggest number of required resources.
+        \item Increase the number of periods in multiples of 10 until no integer solutions can be found in 3600 seconds.
+        \item Repeat step 2.
+    \end{enumerate}
 
@@
 
-    There are a total of 12 tasks.
+    There are a total of 11 tasks.
     Most tasks demand between 1 and 8 resources per month. The biggest and second biggest tasks demand 50 and 20 resources respectively each month.
-    More details of task differences are found in table \ref{tab:tasks}.
     Candidates are calculated based on each task's requirements matching the candidate's capabilities.
 
-    It is possible to see that task 08 is impossible to satisfy given that there are less candidates than resources needed. This task has been taken out from all of the instances.
-
-    \begin{table}
-
-    \input{tables/tasks.tex}
-
-    \caption{Tasks' main characteristics including number of candidate resources and monthly needs.
-It reads in the following way: Task 'O1' consumes 80 flight hours of each of 8 resources each month.
-These resources need to be chosen from a pool of 50 possible candidates.}
-    \vspace{-0.5cm}
-    \label{tab:tasks}
-    \end{table}
+    Some tasks have a limited duration, meaning they start and end during the planning horizon. Others are active during the whole planning horizon.
 
@@
   The table \ref{tab:results2} show the progress of the integer solution for each instance.
-For more detail of this progress, figure \ref{fig:progress} shows an example of progress for instance with id=1817.
+For more detail of this progress, figure \ref{fig:progress} shows an example of progress for instance with id=I\_7.
 It can be seen that initial solutions are quite big in comparison with the optimal.
@@
       \includegraphics[width=0.5\textwidth]{./../../img/progress.png}
-    \caption{Solving progress of the integer solution (in blue) and the lower bound (red) along the solver iterations (x axis). Instance id=1817.}
+    \caption{Solving progress of the integer solution (in blue) and the lower bound (red) along the solver iterations (x axis). Instance id=I\_7.}
     \label{fig:progress}
@@
 Table \ref{tab:multiobj} shows the obtained results for each instance.
-The base instance that was used for this comparison was the one of 21 periods and 9 tasks corresponding to id 1331.
+The base instance that was used for this comparison was the one of 21 periods and 9 tasks corresponding to id=I\_1.
 
@@
 
-  \caption{Comparison of instances where variations on the weight of its two objectives has been done. Instance id=1331. All experiments were solved to optimality.}
+  \caption{Comparison of instances where variations on the weight of its two objectives has been done. Instance I\_1. All experiments were solved to optimality.}
   \vspace{-0.5cm}
@@
       \includegraphics[width=0.5\textwidth]{./../../img/multiobjective.png}
-    \caption{Pareto diagram comparing the quality of both objectives for each instance. Instance id=1331.}
+    \caption{Pareto diagram comparing the quality of both objectives for each instance. Instance id=I\_1.}
     \label{fig:pareto}

### 2018-01-22 latex/MOSIM2018/article.tex
updated MOSIM article

@@
 \\
-     \normalsize ISAE-SUPAERO \\
-     \normalsize Université de Toulouse \\
+     \normalsize ISAE-SUPAERO, % \\
+     \normalsize Université de Toulouse, % \\
      \normalsize 31055 Toulouse cedex 4 - France\\
@@
 \rm
-{\em The Military Flight and Maintenance Planning problem is presented.
-In this problem, preventive maintenance operations are scheduled for military aircraft along with the assignment of regular missions.
-The scheduling of maintenances and mission assignments is a tedious task currently done manually by specialized personnel.
-A MIP mathematical model is formulated in order to produce better quality solutions and reduce the time it takes to find them.
+{\em A specific Flight and Maintenance Planning problem is presented in this paper, regarding special requirements from the French Airforce's real-life operative.
+In this problem, preventive maintenance operations are scheduled for military aircraft, along with the assignment of regular missions, fulfilling requirements in terms of standards and retrofits. 
 The quality of the solutions is measured via a bi-objective function that smooths both maintenances and aircraft unavailability among time periods. 
-A real world dataset provided by the French Airforce is used to test the implementation.}\\~\\
+An exact optimization approach is provided by a Mixed Integer Programming model.
+A real world dataset is used to test the implementation.
+Mono-objective computations are solved to optimality for medium size instances, allowing to compute exactly the Pareto frontier of the bi-objective optimization problems.
+An industrial result is that these two objectives do not lead to the same optimal solution, but very good compromise solutions
+can be computed thanks to the bi-objective optimization.
+}\\~\\
 {\bf KEYWORDS:}
 \rm
-{\em optimization, planning, aircraft maintenance, MIP formulation, bi-objective optimization}
+{\em optimization, Flight and Maintenance Planning, aircraft maintenance, Mixed Integer Programming, bi-objective optimization}
 \end{minipage}
@@
 
-  This paper presents a specific version of the Flight and Maintenance Problem, which we call Military Flight and Maintenance Problem.
-We provide a description of the problem, a mathematical formulation to obtain quality solutions for a bi-objective optimization function and experimental results to show both the quality of the solutions and the limitations of the technique.
+  Flight and Maintenance Planning (FMP) problems gather optimization problematics from civilian or military operational applications,
+  to schedule maintenances of aircrafts taking into account the implications of maintenance immobilizations for the demand in flight hours,
+  for a conjoint optimization among maintenance dates and flight hours.
+
+  The specific FMP  of this paper comes from the French Airforce's real-life operations, with two specificities regarding the state-of-the-art FMP.
+  On one hand, the fleet of aircraft is heterogeneous, with different standards, capacities and retrofits which are required for some missions.
+  On the other hand, there is a will to align two objectives: the smoothing of the maintenance planning in order to better use the maintenance capacities and for the organization of external contracting; and the minimization of the global unavailability of aircrafts to be able to add new missions.
+This last point raises a problematic of bi-objective optimization.
+
+  This paper investigates how to solve such new problematic, and how to take good compromise decisions if these objectives are antagonistic.
 
@@
   Based on the information provided, a number of instances were created as subsets of the main dataset in order to run different experiments.
+
+  Table \ref{tab:instance} shows the characteristics of each instance. The main differences between instances are in the number of tasks and the number of periods.
+
   The criteria for the creation of the instances was the following:
@@
         \item The base instances was the biggest instance where an integer solution was found in less than 3600 seconds: all 11 tasks and 31 periods.
-        \item Take out the task with the biggest number of required resources.
-        \item Increase the number of periods in multiples of 10 until no integer solutions can be found in 3600 seconds.
-        \item Repeat step 2.
+        \item We take out the task with the biggest number of required resources.
+        \item We increased the number of periods in multiples of 10 until no integer solutions can be found in 3600 seconds.
+        \item Repeated step 2.
     \end{enumerate}
+
+  \begin{table}
+  \begin{center}
+
+  \input{tables/instance.tex}
+
+  \caption{Instances used in the experiments with indicators of size.
+"id" is the instance; $|\mathcal{T}|$ the number of periods; "assign" the number of assignments of resources to tasks; $|\mathcal{J}|$ the number of tasks; "vars" the number of variables; "cons" the number of constraints; and "nonzeros" the number of non zero values in the matrix}
+  \vspace{-0.5cm}
+  \label{tab:instance}
+  \end{center}
+  \end{table}
 
@@
   All tests were run on a 64 bit Ubuntu 16.04 workstation with 16GB of RAM and Intel i7-6700HQ CPU \@ 2.60GHz x 8 processor.
-  Several experiments were done with the MIP formulation.
-Table \ref{tab:instance} shows the characteristics of each experiment.
-The main differences between instances is the number of tasks and the number of periods.
+  Several experiments were done with the MIP formulation, following instances created from the base dataset and shown in table \ref{tab:instance}.
 
@@
   Both weights on the objective function were set at 1. For a sensibility analysis on these two parameters, see section \ref{sec:multi}.
-
-  \begin{table}
-  \begin{center}
-
-  \input{tables/instance.tex}
-
-  \caption{Instances used in the experiments with indicators of size.
-"id" is the instance; $|\mathcal{T}|$ the number of periods; "assign" the number of assignments of resources to tasks; $|\mathcal{J}|$ the number of tasks; "vars" the number of variables; "cons" the number of constraints; and "nonzeros" the number of non zero values in the matrix}
-  \vspace{-0.5cm}
-  \label{tab:instance}
-  \end{center}
-  \end{table}
 
@@
   Table \ref{tab:results} show the details of the solving process for each instance.
-As can be seen, cuts do not improve significantly the relaxation (bound cuts is very close to the root solution).
+As can be seen, cuts do not improve significantly the relaxation ("bound cuts" is very close to the root solution).
 In fact, the original lineal relaxation (root) is quite similar to the final lower bound (bound).
 This suggests further improvement in more compact modelization could help reduce the gap faster.
+
+  An example of tighter modelization could be to replace the constraints that count the remaining usage time and elapsed time for tighter equivalents.
 
@@
 
+  \begin{figure}
+    \centering
+      \includegraphics[width=0.5\textwidth]{./../../img/num-maintenances.png}
+    \caption{Solution for instance id=I\_7: number of resources under maintenance at each period.}
+    \label{fig:num-maintenances}
+  \end{figure}
+
+    \begin{figure}
+    \centering
+      \includegraphics[width=0.5\textwidth]{./../../img/num-maintenances.png}
+    \caption{Solution for instance id=I\_7: number of unavailable resources at each period.}
+    \label{fig:num-unavailable}
+  \end{figure}
+
   The table \ref{tab:results2} show the progress of the integer solution for each instance.
@@
 
+  Figures \ref{fig:num-maintenances} and \ref{fig:num-unavailable} show the solution for the maintenance and unavailable smoothing objectives, respectively.
+
   For two, rather small, instances the solver was able to find an optimal solution before starting to branch. In these two instances, the missing values have been replaced by "-1".
 
-  These results show how the branching is not providing with much improvement in increasing the lower bound or finding better integer solutions.
-This hints into a possible reason for the difficult in solving big instances: the quantity of symmetries in the model, represented as candidates for each task: there exist a very big number of possible candidates to be potentially assigned for each task with respect with the actual needed resources for the task.
+  These results show how the branching is not providing much improvement in increasing the lower bound or finding better integer solutions.
+This hints into a possible reason for the difficult in solving big instances: the quantity of symmetries in the model, represented as candidates for each task.
+There appears to exist a very big number of possible candidates to be potentially assigned for each task with respect with the actual needed resources for the task.
 This reduces the capacity of the branching the solver does by making it explore many nodes that are really almost equivalent.
+
+  An example to break these symmetries could be to slightly change the objective function so it can measure more gradually the smoothness of the objective and help get better bounds.
+Another alternative could be to assign priorities to some combinations of resource-task or to limit the number of candidates that can be assigned to each task by some heuristic that clusters resources with tasks.
+Finally, arbitrarily preassigning (fixing) the final state (in terms of remaining usage time) to each resource could also help break those symmetries without sacrificing too much optimality.
 
@@
 
+    As new constraints are added, such as minimum durations for assignments and the possibility to "store" resources for a predefined time, it is predicted that the present instances will grow bigger.
+    This, of course, will require new and better techniques to decompose the problem into smaller problems using heuristics or to take a decomposition approach such as Column Generation.
+
   \subsection{Bi-objective analysis}
@@
 
-    In order to understand the sensibility of the weights $W_1$ and $W_2$ on the solution, a bi-objective analysis was done.
+    In order to understand the sensibility of the weights $w_1$ and $w_2$ on the solution, a bi-objective analysis was done.
 In it, both weights were changed between 0 and 1 in intervals of 0.1 so that they always summed 1.
 Table \ref{tab:multiobj} shows the obtained results for each instance.
-The base instance that was used for this comparison was the one of 21 periods and 9 tasks corresponding to id=I\_1.
+The base instance that was used for this comparison was the one of 21 periods and 10 tasks corresponding to id=I\_5.
 
-    The instances that were chosen solved up to optimality.
+    The experiments done with this instance solved up to optimality.
 
@@
 
-  \caption{Comparison of instances where variations on the weight of its two objectives has been done. Instance I\_1. All experiments were solved to optimality.}
+  \caption{Comparison of experiments where variations on the weight of its two objectives has been done. Instance id=I\_5. All experiments were solved to optimality.}
   \vspace{-0.5cm}
@@
       \includegraphics[width=0.5\textwidth]{./../../img/multiobjective.png}
-    \caption{Pareto diagram comparing the quality of both objectives for each instance. Instance id=I\_1.}
+    \caption{Pareto diagram comparing the quality of both objectives for each instance. Instance id=I\_5.}
     \label{fig:pareto}
   \end{figure}
+
+    A second analysis was done on all the previously defined instances (see \ref{tab:instance}).
+This second analysis is similar to the first one but included all of the instances instead of just picking one.
+The analysis had as objective to observe the amount of Pareto optimal points for each instance.
+
+    Table \ref{tab:multmultiobj} shows the results of this analysis.
+It can be seen that almost all of the instances have a small number of Pareto optimal solutions.
+This shows how both objectives really are quite aligned one with the other.
+
+  \begin{table}[h]
+  \begin{center}
+
+  \input{tables/multi-multiobj.tex}
+
+  \caption{Pareto optimal points found on each instance after 3600 seconds.
+All instances were analyzed.
+"First" and "Last" indicate the two extremes of the Pareto frontier.
+"\# points" indicates the number of Pareto optimal points found.
+Not all experiments were solved up to optimality.}
+  \vspace{-0.5cm}
+  \label{tab:multmultiobj}
+  \end{center}
+  \end{table}
 
@@
 
-  The results show that a direct exact method to solve the whole dataset may not be ideal.
-Although medium size instances have been solved, the complete dataset is not solved in a reasonable time for more horizons bigger than 10 periods.
-Still, the analysis on the progress of the lower bound shows that further work can be done to improve the model in order for it to have a tighter, better relaxation.
-One possibility is to replace the constraints that count the remaining usage time and elapsed time for tighter constraints.
-Another possibility is slightly changing the objective function so it can measure more gradually the smoothness of the objective and help get better bounds.
+  This paper proposed a new MIP formulation to address a specific FMP with an heterogeneous fleet of aircraft with different  capacities and retrofits which are required for some missions.
+  Two objective functions were modeled to tackle the  bi-objective optimization problematic, trying to maximize both  the aircraft availability 
+  and the smoothness of maintenances.
+  MIP convergence characteristics are analyzed on real-world instances.
+  Mono-objective computations are solved to optimality for medium size instances, allowing to compute exactly the Pareto frontier of the bi-objective optimization problems.
+  An industrial result is that these two objectives  do not lead to the same optimal solution, but very good compromise solutions
+  can be computed thanks to the bi-objective optimization.
+  For larger instances, difficulties occurred because of multiple symmetries and a poor quality of MIP generic primal heuristics.
 
-  Regarding the multi-objective nature of the problem, to confirm the true number of solutions, an analysis using epsilon constrains should follow.
-Further tests with different, bigger, instances could also reveal more information regarding the relation between the two objectives.
-In addition to this, new objectives could be included into the analysis, such as smoothing the remaining usage time among periods instead of doing it only in the last period.
-
-  Regarding the branching and symmetries: this could be handled by assigning priorities to some combinations of resource-task or by limiting the number of candidates that can be assigned to each task by some heuristic that clusters resources with tasks.
-Finally, arbitrarily preassigning (fixing) the final state (in terms of remaining usage time) to each resource could also help break those symmetries.
-
-  As new constraints are added, such as minimum durations for assignments and the possibility to "store" resources for a predefined time, it is predicted that the present instances will grow bigger.
-This, of course, will require new and better techniques to decompose the problem into smaller problems using heuristics or to take a decomposition approach such as Column Generation.
-
-  Finally, this work presents a first step useful into the solving of the Military Flight and Maintenance Planning.
-The model presented provides already good initial solutions and also the potential of solving even bigger, more complex problems.
+  A first perspective is to improve the resolution for large size instances. Heuristics are a first research direction, but also mathematical programming work to improve the exact resolution
+  with symmetry breaking or MIP reformulations. 
+  For the bi-objective optimization for large size instances, more advanced bi-objective methods such as $\epsilon$-constraints or Two-Phases-Method are likely to be investigated
+  for the computation of larger Pareto frontiers.
+  Lastly, additional constraints from the real world application can be incorporated in the model.
 

### 2018-01-22 latex/MOSIM2018/article.tex
MOSIM updated after re-read

@@
 \rm
-{\em A specific Flight and Maintenance Planning problem is presented in this paper, regarding special requirements from the French Airforce's real-life operative.
-In this problem, preventive maintenance operations are scheduled for military aircraft, along with the assignment of regular missions, fulfilling requirements in terms of standards and retrofits. 
+{\em A specific Flight and Maintenance Planning problem is presented in this paper, regarding special requirements from the French Air Force's real-life operative.
+In this problem, preventive maintenance operations are scheduled for military aircraft along with the assignment of regular missions, fulfilling requirements in terms of standards and retrofits. 
 The quality of the solutions is measured via a bi-objective function that smooths both maintenances and aircraft unavailability among time periods. 
@@
 
-  Flight and Maintenance Planning (FMP) problems gather optimization problematics from civilian or military operational applications,
-  to schedule maintenances of aircrafts taking into account the implications of maintenance immobilizations for the demand in flight hours,
+  Flight and Maintenance Planning (FMP) problems gather optimization problematics from civilian or military operational applications. 
+  They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours,
   for a conjoint optimization among maintenance dates and flight hours.
 
-  The specific FMP  of this paper comes from the French Airforce's real-life operations, with two specificities regarding the state-of-the-art FMP.
+  The specific FMP  of this paper comes from the French Air Force's real-life operations, with two specificities regarding the state-of-the-art FMP.
   On one hand, the fleet of aircraft is heterogeneous, with different standards, capacities and retrofits which are required for some missions.
-  On the other hand, there is a will to align two objectives: the smoothing of the maintenance planning in order to better use the maintenance capacities and for the organization of external contracting; and the minimization of the global unavailability of aircrafts to be able to add new missions.
+  On the other hand, there is a will to align two objectives: the smoothing of the maintenance planning in order to better use the maintenance capacities and for the organization of external contracting; and the minimization of the global unavailability of aircraft to be able to add new missions.
 This last point raises a problematic of bi-objective optimization.
@@
   \label{sec:problem}
-  The problem consists in assigning resources to predefined tasks and scheduling periodic preventive maintenances for these same resources in order to keep them ready for doing tasks in the future.
+  The problem consists in assigning resources to predefined tasks and scheduling periodic preventive maintenances for these same resources in order to keep them ready for doing (more) tasks in the future.
 In order to generalize the formulation of the model, in the description presented below, the following naming convention has been adopted: aircraft are called resources, missions are called tasks.
@@
 
-    The assignment of a resource to a task is not decided for the whole duration of the task.
+    % The assignment of a resource to a task is not decided for the whole duration of the task.
 After a minimum amount of time ($MT$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished.
@@
     Finally, tasks have a set of requisites that the resources that can be assigned to it need to comply with.
+Each task requires one and only one type of resource.
+In addition to this, specific functionalities can be demanded by the task.
 
@@
         % \item remaining storage time (see \ref{def:sto}).
-        \item type of last maintenance (see \ref{def:maint}).
+        % \item type of last maintenance (see \ref{def:maint}).
     \end{itemize}
@@
 
-    In planning tasks and maintenances, it is important to take into account the initial state of each resource.
+    In planning tasks and maintenances, it is important to take into account the initial and end state of each resource.
 This initial state can be a maintenance or an assigned task.
@@
 
-    Given that the creation of new tasks and the duration of maintenance are considered stochastic, one basic goal is to maximize the robustness of the planning by having the greatest amount of available resources at every period of the planning horizon.
+    Given that the creation of new tasks and the duration of maintenance are considered stochastic in real-life, one basic goal is to maximize the robustness of the planning by having the greatest amount of available resources at every period of the planning horizon.
 
@@
 
-  First of all, the objective function's motivation is on reducing the load on the maintenance facilities and guaranteeing the availability of the fleet instead of reducing the cost.
+  Firstly, the objective function's motivation is on reducing the load on the maintenance facilities and guaranteeing the availability of the fleet instead of reducing the cost.
 Also, the maintenance operations take considerably longer and the rules that govern the maintenances are different.
@@
 In \cite{gavranis2015exact}, an exact algorithm was presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
-Finally, in \cite{kozanidis2014heuristics}, two heuristics to solve big instances of the problem by taking consecutive decisions period by period with the help of mathematical models at each step.
+Finally, in \cite{kozanidis2014heuristics}, two heuristics are presented to solve big instances of the problem by taking consecutive decisions period by period with the help of mathematical models at each step.
 
@@
         $R_j$             & amount of resources that task $j$ needs when active. \\
-        $MT$              & minimum number of periods a resource can be assigned to a task. \\
+        % $MT$              & minimum number of periods a resource can be assigned to a task. \\
         % $aut_i$           & maximal absolute usage time for resource $i$. \\
@@
         $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - m+1\}},  ..., {t}$\}. \\
+        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
     \end{tabular}
@@
                 & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
-        & ret_{it} \in [0,me]
+        & ret_{it} \in [0,E]
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
@@
   \label{sec:data}
-  The dataset being used was provided by the French Airforce and already used by the French Defense Procurement Agency in \cite{Chastellux2017}.
+  The dataset being used was provided by the French Air Force and already used by the French Defense Procurement Agency in \cite{Chastellux2017}.
   Based on the information provided, a number of instances were created as subsets of the main dataset in order to run different experiments.
@@
     \begin{enumerate}
-        \item The base instances was the biggest instance where an integer solution was found in less than 3600 seconds: all 11 tasks and 31 periods.
+        \item The base instances was the biggest instance where an integer solution was found in less than 3600 seconds: all 11 tasks and 21 periods.
         \item We take out the task with the biggest number of required resources.
         \item We increased the number of periods in multiples of 10 until no integer solutions can be found in 3600 seconds.
-        \item Repeated step 2.
+        \item We repeated step 2.
     \end{enumerate}
@@
     There are 127 resources in total. Each one with its functionalities and past assignments.
-    The distribution of initial usage times is shown in the figure \ref{fig:histogramUsage}. It can be seen that the aircraft do not necessarily follow an horizontal line.
+    The distribution of initial usage times is shown in the figure \ref{fig:histogramUsage}.
+It can be seen that the distribution of remaining usage times does not necessarily follow an horizontal line as is expected.
 
@@
   Table \ref{tab:summary} gives an overview on the results obtained for each of the executed instances.
-As can be seen, smaller instances are solved up to optimality while medium size ones are stopped after one hour.
+As can be seen, smaller instances are solved up to optimality while medium-size ones are stopped after one hour.
 
-  Both weights on the objective function were set at 1. For a sensibility analysis on these two parameters, see section \ref{sec:multi}.
+  Both weights on the objective function were set at 1. For a sensibility analysis of these two parameters, see section \ref{sec:multi}.
+
+  It is worth noticing that the final gap between integer and lineal solutions is relatively small given that instances only a little bigger did not provide an integer solution after 3600 seconds.
 
@@
   \caption{Solution information for each tested instance.
-"id" is the instance identifier; "objective" the best integer solution found; "gap" the percentage gap between the integer solution and the lineal relaxation; "time (s)" the time limit that was chosen; and "bound" the lower bound from the lineal relaxation}
+"id" is the instance identifier; "objective" the best integer solution found; "gap" the percentage gap between the integer solution and the lineal relaxation; "time (s)" the time it took to solve or 3600; and "bound" the biggest lower bound found from the lineal relaxation}
   \vspace{-0.5cm}
@@
 
-  Table \ref{tab:results} show the details of the solving process for each instance.
+  Table \ref{tab:results} shows the details of the solving process for each instance.
 As can be seen, cuts do not improve significantly the relaxation ("bound cuts" is very close to the root solution).
@@
 
-  \begin{figure}
-    \centering
-      \includegraphics[width=0.5\textwidth]{./../../img/num-maintenances.png}
-    \caption{Solution for instance id=I\_7: number of resources under maintenance at each period.}
-    \label{fig:num-maintenances}
-  \end{figure}
+  The table \ref{tab:results2} show the progress of the integer solution for each instance.
+For more detail on this progress, figure \ref{fig:progress} shows an example of progress for instance with id=I\_7.
+It can be seen that initial solutions are quite big in comparison with the optimal due to the use of heuristics being run at the beginning of the solving process.
+In fact, for two rather small instances the solver was able to find an optimal solution before starting to branch.
+In these two instances, the missing values have been replaced by "-1".
 
-    \begin{figure}
-    \centering
-      \includegraphics[width=0.5\textwidth]{./../../img/num-maintenances.png}
-    \caption{Solution for instance id=I\_7: number of unavailable resources at each period.}
-    \label{fig:num-unavailable}
-  \end{figure}
-
-  The table \ref{tab:results2} show the progress of the integer solution for each instance.
-For more detail of this progress, figure \ref{fig:progress} shows an example of progress for instance with id=I\_7.
-It can be seen that initial solutions are quite big in comparison with the optimal.
-
-  Figures \ref{fig:num-maintenances} and \ref{fig:num-unavailable} show the solution for the maintenance and unavailable smoothing objectives, respectively.
-
-  For two, rather small, instances the solver was able to find an optimal solution before starting to branch. In these two instances, the missing values have been replaced by "-1".
-
-  These results show how the branching is not providing much improvement in increasing the lower bound or finding better integer solutions.
+  Both graphically and tabularly, results show that the branching is not providing much improvement in increasing the lower bound or finding better integer solutions.
 This hints into a possible reason for the difficult in solving big instances: the quantity of symmetries in the model, represented as candidates for each task.
@@
 
-  An example to break these symmetries could be to slightly change the objective function so it can measure more gradually the smoothness of the objective and help get better bounds.
+  An example of an alternative to break these symmetries could be to slightly change the objective function so it can measure more gradually the smoothness of the objective and help get better bounds.
 Another alternative could be to assign priorities to some combinations of resource-task or to limit the number of candidates that can be assigned to each task by some heuristic that clusters resources with tasks.
@@
 
+  As an example, figures \ref{fig:num-maintenances} and \ref{fig:num-unavailables} show the solution for the maintenance and unavailable smoothing objectives, respectively.
+
+  \begin{figure}
+    \centering
+      \includegraphics[width=0.5\textwidth]{./../../img/num-maintenances.png}
+    \caption{Solution for instance id=I\_7: number of resources under maintenance at each period.}
+    \label{fig:num-maintenances}
+  \end{figure}
+
+  \begin{figure}
+    \centering
+      \includegraphics[width=0.5\textwidth]{./../../img/num-unavailables.png}
+    \caption{Solution for instance id=I\_7: number of unavailable resources at each period.}
+    \label{fig:num-unavailables}
+  \end{figure}
+
   \subsection{Bi-objective analysis}
@@
 
-    In order to understand the sensibility of the weights $w_1$ and $w_2$ on the solution, a bi-objective analysis was done.
+    In order to understand the sensibility of the weights $W_1$ and $W_2$ on the solution, a bi-objective analysis was done.
 In it, both weights were changed between 0 and 1 in intervals of 0.1 so that they always summed 1.
-Table \ref{tab:multiobj} shows the obtained results for each instance.
-The base instance that was used for this comparison was the one of 21 periods and 10 tasks corresponding to id=I\_5.
+This was done in all of the previously defined instances (see \ref{tab:instance}).
 
-    The experiments done with this instance solved up to optimality.
+    Table \ref{tab:multmultiobj} shows the results of this analysis. It can be seen that all of the instances have a small number of Pareto optimal solutions.
 
-    In figure \ref{fig:pareto} it is possible to see that there are at least three Pareto optimal points.
-Further analysis must be done with different instances to obtain insights on the relation between the two objectives.
+    \begin{table}[h]
+    \begin{center}
 
-    The bi-objective analysis appears to indicate the existence of a small number of solutions that are Pareto optimal.
+    \input{tables/multi-multiobj.tex}
+
+    \caption{Pareto optimal points found on each instance after 3600 seconds.
+All instances were analyzed.
+"First" and "Last" indicate the two extremes of the Pareto frontier.
+"\# points" indicates the number of Pareto optimal points found.
+Not all experiments were solved up to optimality.}
+    \vspace{-0.5cm}
+    \label{tab:multmultiobj}
+    \end{center}
+    \end{table}
+
+    More particularly, an individual instance was analyzed, where an optimal solution was found in each experiment.
+    Table \ref{tab:multiobj} shows the obtained results for each experiment. The base instance that was used for this comparison was the one of 21 periods and 10 tasks corresponding to id=I\_5.
+
+    In figure \ref{fig:pareto} it is possible to see the different solutions for this instance.
+
+    This bi-objective analysis appears to indicate the existence of a small number of solutions that are Pareto optimal.
 If this is the case, an iterative approach that explores these solutions could be done.
 Improvement in the mono-objective modeling could also lead to producing non-dominated solutions.
+
+    Finally, the results show that both objectives are quite aligned one with the other.
 
@@
     \end{figure}
-
-    A second analysis was done on all the previously defined instances (see \ref{tab:instance}).
-This second analysis is similar to the first one but included all of the instances instead of just picking one.
-The analysis had as objective to observe the amount of Pareto optimal points for each instance.
-
-    Table \ref{tab:multmultiobj} shows the results of this analysis.
-It can be seen that almost all of the instances have a small number of Pareto optimal solutions.
-This shows how both objectives really are quite aligned one with the other.
-
-  \begin{table}[h]
-  \begin{center}
-
-  \input{tables/multi-multiobj.tex}
-
-  \caption{Pareto optimal points found on each instance after 3600 seconds.
-All instances were analyzed.
-"First" and "Last" indicate the two extremes of the Pareto frontier.
-"\# points" indicates the number of Pareto optimal points found.
-Not all experiments were solved up to optimality.}
-  \vspace{-0.5cm}
-  \label{tab:multmultiobj}
-  \end{center}
-  \end{table}
 

### 2018-01-28 latex/MOSIM2018/article.tex
added Olgas corrections and better graphs

@@
 \rm
-{\em A specific Flight and Maintenance Planning problem is presented in this paper, regarding special requirements from the French Air Force's real-life operative.
-In this problem, preventive maintenance operations are scheduled for military aircraft along with the assignment of regular missions, fulfilling requirements in terms of standards and retrofits. 
-The quality of the solutions is measured via a bi-objective function that smooths both maintenances and aircraft unavailability among time periods. 
+{\em A specific Flight and Maintenance Planning problem is presented in this. 
+In this problem, preventive maintenance operations are scheduled for military aircraft along with the assignment of regular missions. 
+The quality of the solutions is measured via a bi-objective function that smooths both maintenance operations and aircraft unavailability among time periods. 
 An exact optimization approach is provided by a Mixed Integer Programming model.
-A real world dataset is used to test the implementation.
+A real world dataset provided by the French Air Force is used to test the developed optimization approach.
 Mono-objective computations are solved to optimality for medium size instances, allowing to compute exactly the Pareto frontier of the bi-objective optimization problems.
-An industrial result is that these two objectives do not lead to the same optimal solution, but very good compromise solutions
-can be computed thanks to the bi-objective optimization.
+The tests show that these two objectives do not lead to the same optimal solution, but very good compromise solutions can be found thanks to the bi-objective optimization.
 }\\~\\
@@
 
-  Flight and Maintenance Planning (FMP) problems gather optimization problematics from civilian or military operational applications. 
+  Flight and Maintenance Planning (FMP) problems raise civilian or military operational applications. 
   They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours,
@@
 
-  The specific FMP  of this paper comes from the French Air Force's real-life operations, with two specificities regarding the state-of-the-art FMP.
-  On one hand, the fleet of aircraft is heterogeneous, with different standards, capacities and retrofits which are required for some missions.
-  On the other hand, there is a will to align two objectives: the smoothing of the maintenance planning in order to better use the maintenance capacities and for the organization of external contracting; and the minimization of the global unavailability of aircraft to be able to add new missions.
-This last point raises a problematic of bi-objective optimization.
+  The specific FMP  of this paper comes from the French Air Force's operations. It has two specificities regarding the classical FMP considered in the literature.
+  On one hand, the fleet of aircraft is heterogeneous, with different standards, capacities and retrofits.
+  On the other hand, there is a will to align two objectives: the smoothing of the maintenance operations and the minimization of the global unavailability of aircraft to be able to add new missions.
+This last point calls for a bi-objective optimization.
 
@@
 
-  Section \ref{sec:problem} describes with detail the problem, its different elements, conditions and objectives.
-We also make a review of previous work done in the subject in section \ref{sec:soa} where we compare the addressed problems with our own and explain the differences.
-
-  Later, in section \ref{sec:data}, the specific dataset at hand is described.
-Section \ref{sec:model} presents a complete MIP formulation to solve the described problem and section \ref{sec:results} shows the commented results obtained while solving several instances of the problem based on the provided dataset.
-
-  Finally, section \ref{sec:conclusions} mentions the conclusions obtained from the experimentation and the possible direction of further work.
+  Section \ref{sec:problem} presents a detailed description of the problem. An analysis of the previous work on FMP is given in section \ref{sec:soa}.
+  Section \ref{sec:model} presents a new MIP formulation for the considered problem. 
+  Section \ref{sec:data} introduces the real case study.
+  Section \ref{sec:results} discusses the obtained results.
+  Finally, section \ref{sec:conclusions} provides conclusions and directions of further work.
 
@@
   \label{sec:problem}
-  The problem consists in assigning resources to predefined tasks and scheduling periodic preventive maintenances for these same resources in order to keep them ready for doing (more) tasks in the future.
-In order to generalize the formulation of the model, in the description presented below, the following naming convention has been adopted: aircraft are called resources, missions are called tasks.
+  The problem consists in assigning resources to predefined tasks and scheduling periodic preventive maintenances for these same resources.
+For the seek of simplicity, the following convention has been adopted: aircraft are called resources, missions are called tasks.
 
@@
     There is a fixed set of $j \in \mathcal{J}$ tasks to be accomplished over an horizon of time divided into $t \in \mathcal{T}$ discrete periods.
-In order to be accomplished, these tasks require the assignment of a specific number of resources $R_j$ each period of time the task is active.
-Tasks have both a fixed start and end time that define the set of periods of time it is active.
+For their execution, these tasks require the assignment of a specific number of resources $R_j$ each period of time the task is active.
+The start and end periods for each task are known and a task is considered active between its start and end period.
 
-    During each assigned period, tasks use the assigned resources for a time $H_j$.
+    During each period, tasks consume an amount of time equal to $H_j$ from each of its assigned resources.
 
@@
 After a minimum amount of time ($MT$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished.
-    The total number of resources being used at any given time in a specific task can never be less than the required $R_j$.
+    The total number of resources being used at any given time in a specific task should be equal to $R_j$.
 
-    Finally, tasks have a set of requisites that the resources that can be assigned to it need to comply with.
-Each task requires one and only one type of resource.
-In addition to this, specific functionalities can be demanded by the task.
+    Each task requires one and only one type of resource which should comply with task requirements.
 
@@
 Each resource can only be assigned to a single task in any given period.
-These resources suffer from wear and tear and require regular maintenances during their lifetime.
-The need for these maintenances is calculated based on two indicators every resource has.
+These resources suffer from wear and tear and require regular maintenance operations during their lifetime.
+The need for maintenance is calculated based on two indicators.
 
-    We will call "remaining elapsed time" (or $ret_{it}$) to the amount of time (measured in time periods) that needs to pass in order for the resource $i$ to need a maintenance at any given period $t$.
-In a similar way, we will call "remaining usage time" (or $rut_{it}$) to the amount of time that the resource $i$ needs to be used in order for it to need a maintenance at any given period $t$.
+    The first one is called "remaining elapsed time" (or $ret_{it}$).
+It expresses the amount of time (measured in time periods) after which the resource cannot be used anymore and has to undergo a maintenance operation.
+Its value is calculated for each resource $i$ and each time period $t$.
+In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the resource $i$ needs to be used in order for it to need a maintenance at any given period $t$.
 
@@
 
-    At any given period, including at the start of the planning period, each resources has a specific status given by:
+    At any given period, including at the start of the planning horizon, each resources has a specific status given by remaining usage time and remaining elapsed time.
 
-    \begin{itemize}
-        \item remaining usage time.
-        \item remaining elapsed time.
+    % \begin{itemize}
         % \item remaining absolute elapsed time.
@@
         % \item type of last maintenance (see \ref{def:maint}).
-    \end{itemize}
+    % \end{itemize}
 
@@
 
-    Maintenances have a fix duration of $M$ periods.
+    Each maintenance operation has a fix duration of $M$ periods.
 
-    After receiving a maintenance, a resource restores its remaining elapsed time and remaining usage time back to their max values $E$ and $H$ respectively.
+    After a maintenance operation, a resource restores its remaining elapsed time and remaining usage time to their max values $E$ and $H$ respectively.
 
-    Finally, resources are organized into families or groups. Each resource inside a family or group shares the same types of maintenances and, usually, is able to do the same type of tasks.
+    % Finally, resources are organized into families or groups. Each resource inside a family or group shares the same types of maintenances and, usually, is able to do the same type of tasks.
 
@@
 
-    As a summary, the following are the possible logical states that a resource can be in:
+    As a summary, the following are the possible logical states that a resource can be in: Assigned to a task (see \ref{def:task}); Under maintenance (see \ref{def:maint}); Available.
 
-      \begin{itemize}
-          \item Assigned to a task (see \ref{def:task}).
-          \item Under maintenance (see \ref{def:maint}).
+      % \begin{itemize}
           % \item Under storage (see \ref{def:sto}).
             % \item Obsolete (see \ref{def:res}).
-          \item Available.
-      \end{itemize}
+      % \end{itemize}
 
-    Figure \ref{fig:solution} shows an extraction of an assignment of missions (red) and maintenances (blue) to aircraft (codes A1 - A9).
+    % Figure \ref{fig:solution} shows an extraction of an assignment of missions (red) and maintenances (blue) to aircraft (codes A1 - A9).
 
-      \begin{figure}
-        \centering
-          \includegraphics[width=0.5\textwidth]{./../../img/calendar.png}
-        \caption{Visual representation of part of the solution to an instance of the problem showing maintenances and assignments.
+    %   \begin{figure}
+    %     \centering
+    %       \includegraphics[width=0.5\textwidth]{./../../img/calendar.png}
+    %     \caption{Visual representation of part of the solution to an instance of the problem showing maintenances and assignments.
 Resources are listed in the vertical axis and the horizontal axis defines the time.
 Red periods correspond to task assignments while blue ones to maintenance assignments.}
-        \label{fig:solution}
-      \end{figure}
+    %     \label{fig:solution}
+    %   \end{figure}
 
-  \subsection{Horizon}
+  \subsection{Time}
     \label{def:hor}
@@
 
-    On the other hand, the remaining used and elapsed times need to be assigned to each resource at the beginning of the planning period.
+    The remaining used and elapsed times are assigned to each resource at the beginning of the planning horizon.
 
-    Finally, the end state of each resource needs to be addressed too.
-This implies guaranteeing that the remaining (elapsed, used) time of resources is sustainable for future tasks and the final states of resources are not too skewed.
+    Finally, the state of each resource at the end of the planning horizon, its remaining (elapsed, used) time, needs to be defined.
+      % needs to be addressed too. This implies guaranteeing that the remaining (elapsed, used) time of resources is sustainable for future tasks and the final states of resources are not too skewed.
 
@@
 
-    The are multiple objectives that need to be taken into account. In this study two will be addressed.
+    In this study, the following objectives are considered.
 
@@
     Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of resources under maintenance over the planning horizon.
-
-    This is summarized by the following two sentences:
-
-    \begin{enumerate}
-        \item Minimize the maximum number of unavailable resources in any given period.
-        \item Minimize the maximum number of resources under maintenance in any given period.
-    \end{enumerate}
 
@@
 
-  The temporal scope of the problem is also bigger, making the normal planning horizon bigger and each period bigger too: months instead of days.
-On the other hand, the physical scope, while present in the commercial planning, is not present in the military one: aircraft are assumed to return to the origin every time.
+  The temporal scope of the problem is also larger, making the normal planning horizon larger and each period larger too: months instead of days.
 
-  In \cite{Kozanidis2008}, a model was made for the Hellenic Air Force.
-It assigned flight hours to aircraft and scheduled flexible maintenance periods based on a "remaining usage time" rule.
-The objectives were to maximize this "remaining usage time" in the aircraft as well as to smooth the availability of aircraft during the planning period.
-Instances of 24 aircraft and 6 monthly periods were tested with Mixed Integer Programming (MIP) formulations and heuristics.
+  In the following, the studies on the Military Flight and Maintenance Planning are discussed.
 
-  In \cite{Cho2011}, the objective was to smooth the maintenances as much as possible.
-Randomly generated instances of 15 aircraft and 520 bi-daily periods were tested with MIP formulations.
-The final state of aircraft was taken into account in order to distribute as much as possible the "remaining usage time" of aircraft.
+  \cite{Kozanidis2008} presents a model for the Hellenic Air Force.
+It assigns flight hours to aircraft and scheduled flexible maintenance periods based on a "remaining usage time" rule.
+The objectives are to maximize this "remaining usage time" for the aircraft as well as to smooth the availability of aircraft during the planning horizon.
+Instances of 24 aircraft and 6 monthly periods were solved with Mixed Integer Programming (MIP) formulations and heuristics.
 
-  In \cite{Verhoeff2015}, a model for the Royal Netherlands Air Force was developed.
+  In \cite{Cho2011}, the objective is to smooth the maintenance operations.
+Randomly generated instances of 15 aircraft and 520 bi-daily periods were solved with MIP formulations.
+The final state of aircraft is taken into account in order to distribute as much as possible the "remaining usage time" of aircraft.
+
+  In \cite{Verhoeff2015}, a model for the Royal Netherlands Air Force is developed.
 Here, three criteria for improving the planning robustness are presented and applied in a MIP model: availability, serviceability and sustainability.
 
-  With respect to heuristic methods for solving this problem, both \cite{Kozanidis2008} and \cite{Cho2011} already presented alternatives to the MIP formulation for solving bigger instances with reduced scope.
-In \cite{steiner2006heuristic}, a constructive, stepped heuristic approach is described to obtain balanced maintenances for the Swiss Air Force.
+  In \cite{gavranis2015exact}, an exact algorithm was presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
+
+  Both \cite{Kozanidis2008} and \cite{Cho2011} develop heuristic alternatives to the MIP formulation for solving larger problem instances.
+In \cite{steiner2006heuristic}, a constructive, multi-stage heuristic approach is described to obtain balanced maintenances for the Swiss Air Force.
 Both calendar-based maintenances actions (CBMA) and usage-based maintenance actions (UBMA) are used.
-In \cite{gavranis2015exact}, an exact algorithm was presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
-Finally, in \cite{kozanidis2014heuristics}, two heuristics are presented to solve big instances of the problem by taking consecutive decisions period by period with the help of mathematical models at each step.
+Finally, in \cite{kozanidis2014heuristics}, two heuristics are presented to solve large instances of the problem by taking consecutive decisions period by period with the help of mathematical models at each step.
 
-  Regarding operational-level planning, a more detailed implementation is found in \cite{safaei2011workforce}, where the specific conditions during the maintenances operations are taken into account to schedule the maintenance jobs, such as the workforce skills and availability.
+  % Regarding operational-level planning, a more detailed implementation can be found in \cite{safaei2011workforce}, where the specific conditions during the maintenance operations are taken into account to schedule the maintenance jobs, such as the workforce skills and availability.
 
-  The concept of mission is not introduced as part of these formulations, only a demand of flight hours is given for all aircraft in total.
+  It should be noted that previous formulations only consider an aggregate demand in flight hours for all aircraft.
+  The aircraft are assumed to be homogeneous and interchangeable.
+  The use of aggregate demand makes it impossible to optimize the task assignments for aircraft.
+   % The concept of mission is not introduced as part of these formulations, only a demand of flight hours is given for all aircraft in total.
 The aircraft themselves have been assumed as homogeneous and interchangeable.
@@
 Here, cycles, composed of one maintenance followed by mission assignments were used.
-Instances of 6 - 70 aircraft and 60 monthly periods were tested, founding close to optimal solutions for small to medium instances using MIP formulations and heuristics based on these MIP formulations.
+Instances of 6 - 70 aircraft and 60 monthly periods were solved.
+Close to optimal solutions for small to medium instances using MIP formulations and heuristics based on these MIP formulations were found.
 
-  Here, an alternative model is presented that simplifies the formulation and should permit the inclusion of the rest of constraints without sacrificing optimality.
+  In the following, an alternative model is presented. This model is then tested in a case study.
 
@@
 
+  \subsection{Sets}
+
+    \begin{tabular}{p{5mm}p{70mm}}
+        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
+        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
+        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
+        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
+        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
+    \end{tabular}
+
     \begin{tabular}{p{8mm}p{67mm}}
-        $H_j$             & amount of usage time for resources used in task $j$. \\
-        $R_j$             & amount of resources that task $j$ needs when active. \\
+        $H_j$             & amount of resource time required by task $j$. \\
+        $R_j$             & number of resources required by task $j$. \\
         % $MT$              & minimum number of periods a resource can be assigned to a task. \\
@@
         % $S$               & remaining storage time after (certain) maintenances. \\
-        $W_1$               & weight of smoothness for maintenances. \\
-        $W_2$               & weight of smoothness for unavailabilities. \\
-        $N_t$             & number of resources in already-planned maintenance in period $t$.\\
-        $D_t$             & number of resources that need to be assigned in total in period $t$. \\
-        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning period. \\
-        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning period. \\
-        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning period. \\
-        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning period. \\
-    \end{tabular}
-
-  \subsection{Sets}
-
-    \begin{tabular}{p{5mm}p{70mm}}
-        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ will be active. \\
-        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
-        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to each task $j$. \\
-        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
+        $W_1$               & weight of the first objective in the objective function. \\
+        $W_2$               & weight of the second objective in the objective function. \\
+        $N_t$             & number of resources in already-planned maintenances in period $t$ at the beginning of the planning horizon.\\
+        $D_t$             & number of resources to be assigned in total in period $t$. \\
+        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
+        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
+        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
+        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
     \end{tabular}
@@
 
-     The decision variables manage the previously defined resources.
+     The following decision variables define a solution.
 
     \begin{tabular}{p{8mm}p{67mm}}
-        $a_{jti}$   &  1 if task $j \in J$ in period $t\in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
-        $m_{it}$    &  1 if resource $i \in I$ starts maintenance in period $t \in \mathcal{T}$, 0 otherwise. \\
+        $a_{jti}$   &  =1 if task $j \in J$ in period $t\in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
+        $m_{it}$    &  =1 if resource $i \in I$ starts maintenance in period $t \in \mathcal{T}$, 0 otherwise. \\
         $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
-        $ret_{it}$  &  remaining elapsed time (integer) for resource $i$ after period $t \in \mathcal{T}0$. \\  
+        $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
         $u_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
@@
 
-    Note that  $a_{jti}$ and $m_{it}$ are initially set up to 0 for all resources already in maintenance at the beginning of the planning period for the remaining time periods of maintenance.
-The remaining usage time for each resource at the beginning of the planning period is used to initialize $rut_{i0}$.
+    Note that  $a_{jti}$ and $m_{it}$ are initially set up to 0 for all resources already in maintenance at the beginning of the planning horizon for the remaining time periods of maintenance.
+The remaining usage time for each resource at the beginning of the planning horizon is used to initialize $rut_{i0}$.
 
@@
     Similarly, the remaining elapsed time is defined by (\ref{eq:ret_upper})-(\ref{eq:ret_initial}) and its limits by (\ref{eq:ret_lower})-(\ref{eq:me}). 
-    Finally, \ref{eq:min_ret} and \ref{eq:min_rut} guarantee that resources have, globally, the same number of remaining used and elapsed times that at the beginning.
+    Finally, constraints \ref{eq:min_ret} and \ref{eq:min_rut} guarantee that resources have, globally, the same amount of remaining used and elapsed times that at the beginning.
 
@@
   \label{sec:data}
-  The dataset being used was provided by the French Air Force and already used by the French Defense Procurement Agency in \cite{Chastellux2017}.
+  The dataset being used was provided by the French Air Force and already used in \cite{Chastellux2017}.
   Based on the information provided, a number of instances were created as subsets of the main dataset in order to run different experiments.
@@
 
-  The criteria for the creation of the instances was the following:
+  The criteria for the creation of the instances were the following:
 
     \begin{enumerate}
-        \item The base instances was the biggest instance where an integer solution was found in less than 3600 seconds: all 11 tasks and 21 periods.
+        \item The original instance was the largest instance where an integer solution was found in less than 3600 seconds: it contains 11 tasks and 21 periods in total.
         \item We take out the task with the biggest number of required resources.
@@
   \caption{Instances used in the experiments with indicators of size.
-"id" is the instance; $|\mathcal{T}|$ the number of periods; "assign" the number of assignments of resources to tasks; $|\mathcal{J}|$ the number of tasks; "vars" the number of variables; "cons" the number of constraints; and "nonzeros" the number of non zero values in the matrix}
+"id" is the instance; $|\mathcal{T}|$ the number of periods; "assign" the number of assignments of resources to tasks; $|\mathcal{J}|$ the number of tasks; "vars" the number of variables; "cons" the number of constraints; and "nonzeros" the number of non zero values in the matrix.}
   \vspace{-0.5cm}
@@
 
-    There are a total of 11 tasks.
-    Most tasks demand between 1 and 8 resources per month. The biggest and second biggest tasks demand 50 and 20 resources respectively each month.
-    Candidates are calculated based on each task's requirements matching the candidate's capabilities.
+    The dataset includes 11 tasks.
+Most of them demand between 1 and 8 resources per period, which is equal to one month.
+The most loaded and second most loaded tasks demand 50 and 20 resources respectively per period.
+    Candidates resources are calculated based on each task's requirements matching the candidate's capabilities.
 
@@
 
-    All maintenances take 6 months (periods).
-Each maintenance provides with 1000 hours of usage time to the resource and a resource can be without maintenance for a maximum of 60 consecutive periods.
+    All maintenance operations last exactly 6 periods.
+Each maintenance provides with 1000 hours of usage time to the resource.
+After it, a resource can be without maintenance for a maximum of 60 consecutive periods.
 The nominative capacity for maintenance is 18 resources at any given time.
@@
 
-    There are 127 resources in total. Each one with its functionalities and past assignments.
-    The distribution of initial usage times is shown in the figure \ref{fig:histogramUsage}.
-It can be seen that the distribution of remaining usage times does not necessarily follow an horizontal line as is expected.
+    There are 127 resources in total. Each one has its functionalities and past assignments.
 
-    \begin{figure}
-      \centering
-        \includegraphics[width=0.5\textwidth]{./../../img/initial_used.png}
-      \caption{Distribution of Usage Times among resources at the beginning of the planning horizon}
-      \label{fig:histogramUsage}
-    \end{figure}
+    % \begin{figure}
+    %   \centering
+    %     \includegraphics[width=0.5\textwidth]{./../../img/initial_used.png}
+    %   \caption{Distribution of Usage Times among resources at the beginning of the planning horizon}
+    %   \label{fig:histogramUsage}
+    % \end{figure}
 
@@
 
+  \section{Numerical tests}
+    \label{sec:tests}
+
     All tests were run on a 64 bit Ubuntu 16.04 workstation with 16GB of RAM and Intel i7-6700HQ CPU \@ 2.60GHz x 8 processor.
-  Several experiments were done with the MIP formulation, following instances created from the base dataset and shown in table \ref{tab:instance}.
+    Several experiments were done with the MIP formulation on problem instances presented in table \ref{tab:instance}.
 
-  Table \ref{tab:summary} gives an overview on the results obtained for each of the executed instances.
-As can be seen, smaller instances are solved up to optimality while medium-size ones are stopped after one hour.
+    Table \ref{tab:summary} gives an overview of the obtained results. As can be seen, smaller instances are solved up to optimality while medium-size ones are stopped after one hour.
 
-  Both weights on the objective function were set at 1. For a sensibility analysis of these two parameters, see section \ref{sec:multi}.
-
-  It is worth noticing that the final gap between integer and lineal solutions is relatively small given that instances only a little bigger did not provide an integer solution after 3600 seconds.
+    For this calculation, both weights on the objective function were set at 1. A sensibility analysis of these two parameters is provided in section \ref{sec:multi}.
 
@@
 
-  \caption{Solution information for each tested instance.
-"id" is the instance identifier; "objective" the best integer solution found; "gap" the percentage gap between the integer solution and the lineal relaxation; "time (s)" the time it took to solve or 3600; and "bound" the biggest lower bound found from the lineal relaxation}
+    \caption{Information on the solution of each tested instance: "id" is the instance identifier; "objective" is the value of the objective function; "gap" is the percentage gap between the integer solution and the linear relaxation; "time (s)" is the solution time or 3600; and "bound" is the greatest lower bound found from the linear relaxation.}
     \vspace{-0.5cm}
@@
     Table \ref{tab:results} shows the details of the solving process for each instance.
-As can be seen, cuts do not improve significantly the relaxation ("bound cuts" is very close to the root solution).
-In fact, the original lineal relaxation (root) is quite similar to the final lower bound (bound).
-This suggests further improvement in more compact modelization could help reduce the gap faster.
+As it can be seen, cuts do not improve significantly the relaxation (the value of "bound cuts" is very close to the root solution).
+In fact, the original linear relaxation (root) is quite similar to the final lower bound (bound).
+This suggests that further improvement in modelization would help to reduce the gap faster.
 
-  An example of tighter modelization could be to replace the constraints that count the remaining usage time and elapsed time for tighter equivalents.
+    An example of a tighter modelization could be to replace the constraints that count the remaining usage time and elapsed time for tighter equivalents.
 
-  The biggest amount of cuts that were used were 'Implied bound' and 'Mixed integer rounding', depending on the instance.
-Being these two generic cuts, it appears the solver has not detected any particular structure as part of this problem.
+    The greatest number of used was of 'Implied bound' and 'Mixed integer rounding', depending on the instance.
+Since these are two generic-type cuts, it can be concluded that the solver has not detected any particular structure as part of this problem.
 
@@
 
-  The table \ref{tab:results2} show the progress of the integer solution for each instance.
+    Table \ref{tab:results2} shows the progress of the integer solution for each instance.
 For more detail on this progress, figure \ref{fig:progress} shows an example of progress for instance with id=I\_7.
@@
 In fact, for two rather small instances the solver was able to find an optimal solution before starting to branch.
-In these two instances, the missing values have been replaced by "-1".
 
-  Both graphically and tabularly, results show that the branching is not providing much improvement in increasing the lower bound or finding better integer solutions.
-This hints into a possible reason for the difficult in solving big instances: the quantity of symmetries in the model, represented as candidates for each task.
-There appears to exist a very big number of possible candidates to be potentially assigned for each task with respect with the actual needed resources for the task.
-This reduces the capacity of the branching the solver does by making it explore many nodes that are really almost equivalent.
+    The obtained results show that the branching is not providing much improvement in increasing the lower bound or finding better integer solutions.
+The quantity of symmetries in the model, represented as candidates for each task could be a reason the difficulty in solving large instances.
+It may exist a very large number of possible candidates to be potentially assigned for each task with respect to the requirements in resources for each task.
+This reduces the branching capacity for the solver, which tries to explore many nodes that are almost equivalent in reality.
 
@@
 Another alternative could be to assign priorities to some combinations of resource-task or to limit the number of candidates that can be assigned to each task by some heuristic that clusters resources with tasks.
-Finally, arbitrarily preassigning (fixing) the final state (in terms of remaining usage time) to each resource could also help break those symmetries without sacrificing too much optimality.
+Finally, arbitrarily preassigning (fixing) the final state (in terms of remaining usage time) to each resource could also help break those symmetries without important loss on the objective function value.
 
@@
 
-  As new constraints are added, such as minimum durations for assignments and the possibility to "store" resources for a predefined time, it is predicted that the present instances will grow bigger. 
-  This, of course, will require new and better techniques to decompose the problem into smaller problems using heuristics or to take a decomposition approach such as Column Generation.
-
-  As an example, figures \ref{fig:num-maintenances} and \ref{fig:num-unavailables} show the solution for the maintenance and unavailable smoothing objectives, respectively.
+    Figures \ref{fig:num-maintenances} and \ref{fig:num-unavailables} show an example distribution of maintenance operations and unavailable numbers of resources per period, respectively.
 
@@
 
-    In order to understand the sensibility of the weights $W_1$ and $W_2$ on the solution, a bi-objective analysis was done.
-In it, both weights were changed between 0 and 1 in intervals of 0.1 so that they always summed 1.
-This was done in all of the previously defined instances (see \ref{tab:instance}).
+    The further calculation was done in order to analyze the impact of the weights $W_1$ and $W_2$ on the solution.
+In it, both weights were changed between 0 and 1 with pace of 0.1 under the condition that $W_1 + W_2 = 1 $.
+This was done for all previously defined instances.
 
-    Table \ref{tab:multmultiobj} shows the results of this analysis. It can be seen that all of the instances have a small number of Pareto optimal solutions.
+    Table \ref{tab:multmultiobj} presents the obtained results of this analysis. It can be seen that all instances have a small number of Pareto optimal solutions.
 
@@
 
-    \caption{Pareto optimal points found on each instance after 3600 seconds.
+    \caption{Non-dominated points found on each instance after 3600 seconds.
 All instances were analyzed.
@@
     More particularly, an individual instance was analyzed, where an optimal solution was found in each experiment.
-    Table \ref{tab:multiobj} shows the obtained results for each experiment. The base instance that was used for this comparison was the one of 21 periods and 10 tasks corresponding to id=I\_5.
+    Table \ref{tab:multiobj} shows the obtained results. The instance used consisted of 21 periods and 10 tasks corresponding to instance with id=I\_5.
 
-    In figure \ref{fig:pareto} it is possible to see the different solutions for this instance.
-
-    This bi-objective analysis appears to indicate the existence of a small number of solutions that are Pareto optimal.
-If this is the case, an iterative approach that explores these solutions could be done.
-Improvement in the mono-objective modeling could also lead to producing non-dominated solutions.
-
-    Finally, the results show that both objectives are quite aligned one with the other.
-
+    For all solved instances, a small number of non dominated solutions was observed.
     \begin{table}[h]
@@
 
-    \caption{Comparison of experiments where variations on the weight of its two objectives has been done. Instance id=I\_5. All experiments were solved to optimality.}
+    \caption{Values of $u_{max}$ (unavailable) and $m_{max}$ (maint) for different combinations of $W_{1}$ and $W_{2}$. Instance id=I\_5. All experiments were solved to optimality.}
     \vspace{-0.5cm}
@@
         \includegraphics[width=0.5\textwidth]{./../../img/multiobjective.png}
-      \caption{Pareto diagram comparing the quality of both objectives for each instance. Instance id=I\_5.}
+      \caption{Pareto diagram for instance id=I\_5.}
       \label{fig:pareto}
@@
 
-  This paper proposed a new MIP formulation to address a specific FMP with an heterogeneous fleet of aircraft with different  capacities and retrofits which are required for some missions.
-  Two objective functions were modeled to tackle the  bi-objective optimization problematic, trying to maximize both  the aircraft availability 
-  and the smoothness of maintenances.
-  MIP convergence characteristics are analyzed on real-world instances.
-  Mono-objective computations are solved to optimality for medium size instances, allowing to compute exactly the Pareto frontier of the bi-objective optimization problems.
-  An industrial result is that these two objectives do not lead to the same optimal solution, but very good compromise solutions
-  can be computed thanks to the bi-objective optimization.
-  For larger instances, difficulties occurred because of multiple symmetries and a poor quality of MIP generic primal heuristics.
+  This paper proposed a new MIP formulation to address a specific FMP with an heterogeneous fleet of aircraft.
+  Two objective functions were chosen to address the French Air Force requirements.
+  The first one to smooth maintenance operations and the second one to aircraft availability.
+  The characteristics of MIP convergence were analyzed on real-world instances.
+  Mono-objective problem instances were solved to optimality for medium size instances. Then, exact Pareto frontiers for bi-objective formulation were calculated for these instances. 
+  The use of these two objectives does not lead to the same optimal solution but very good compromise solutions can be found due to the bi-objective optimization.
+  For larger instances, difficulties in finding optimal solutions occurred because of multiple symmetries and a poor quality of MIP generic primal heuristics.
 
-  A first perspective is to improve the resolution for large size instances. Heuristics are a first research direction, but also mathematical programming work to improve the exact resolution
-  with symmetry breaking or MIP reformulations. 
-  For the bi-objective optimization for large size instances, more advanced bi-objective methods such as $\epsilon$-constraints or Two-Phases-Method are likely to be investigated
-  for the computation of larger Pareto frontiers.
+  The first perspective is to improve the resolution of large size instances.
+Heuristics could be implemented, but also mathematical programming can be improved with symmetry breaking or MIP reformulations.
+  For the bi-objective optimization for large size instances, more advanced bi-objective methods such as $\epsilon$-constraints or Two-Phases-Method should be investigated for the computation of larger Pareto frontiers.
   Lastly, additional constraints from the real world application can be incorporated in the model.

### 2018-01-28 latex/MOSIM2018/article.tex
Corrected some errors in tables mentioned by Nicolas
small other changes that were missing from MOSIM article. Re-executed multi-multi table.

@@
   The aircraft are assumed to be homogeneous and interchangeable.
-  The use of aggregate demand makes it impossible to optimize the task assignments for aircraft.
+  The use of aggregate demand makes it impossible to take into account the task requirements for aircraft.
    % The concept of mission is not introduced as part of these formulations, only a demand of flight hours is given for all aircraft in total.
@@
     All maintenance operations last exactly 6 periods.
-Each maintenance provides with 1000 hours of usage time to the resource.
+Each maintenance sets the remaining usage time of a resource to 1000 hours.
 After it, a resource can be without maintenance for a maximum of 60 consecutive periods.
@@
 For more detail on this progress, figure \ref{fig:progress} shows an example of progress for instance with id=I\_7.
-It can be seen that initial solutions are quite big in comparison with the optimal due to the use of heuristics being run at the beginning of the solving process.
+It can be seen that initial solutions are quite far from the optimal one due to the use of heuristics at the beginning of the solving process.
 In fact, for two rather small instances the solver was able to find an optimal solution before starting to branch.

### 2018-01-29 latex/MOSIM2018/article.tex
additional corrections and recreation of correct multiobjective table and graph with rounding
more corrections
corrected a surprising amount of grammatical errors

@@
 \rm
-{\em A specific Flight and Maintenance Planning problem is presented in this. 
+{\em A specific Flight and Maintenance Planning problem is presented. 
 In this problem, preventive maintenance operations are scheduled for military aircraft along with the assignment of regular missions. 
@@
 
-  Flight and Maintenance Planning (FMP) problems raise civilian or military operational applications. 
-  They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours,
-  for a conjoint optimization among maintenance dates and flight hours.
+  Flight and Maintenance Planning (FMP) problems raise in civilian or military operational applications. 
+  They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.
 
@@
 
-  This paper investigates how to solve such new problematic, and how to take good compromise decisions if these objectives are antagonistic.
+  This paper investigates how to solve such new optimization problem, and how to take good compromise decisions if these objectives are antagonistic.
 
@@
 
-    Each task requires one and only one type of resource which should comply with task requirements.
+    Each task requires one and only one type of resource which, in addition, should comply with additional task requirements.
 
@@
 Its value is calculated for each resource $i$ and each time period $t$.
-In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the resource $i$ needs to be used in order for it to need a maintenance at any given period $t$.
+In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the resource $i$ can be used before needing a maintenance operation at any given period $t$.
 
@@
 
-    At any given period, including at the start of the planning horizon, each resources has a specific status given by remaining usage time and remaining elapsed time.
+    At any given period, including at the start of the planning horizon, each resource has a specific status given by remaining usage time and remaining elapsed time.
 
@@
 
-    Maintenances are the process by which resources that have reached a limit in their usage can return to a state where they can continue to be used in tasks.
+    Maintenances operations are the process by which resources that have reached a limit in some indicator can return to a state where they can continue to be used in tasks.
 
@@
 
-    As a summary, the following are the possible logical states that a resource can be in: Assigned to a task (see \ref{def:task}); Under maintenance (see \ref{def:maint}); Available.
+    As a summary, the following are the possible logical states that a resource can be in: assigned to a task (see \ref{def:task}); under maintenance (see \ref{def:maint}); available.
 
@@
 
-  The Military Flight and Maintenance Planning problem considered assigns missions and schedules maintenances to aircraft.
+  The Military Flight and Maintenance Planning problem considered assigns missions and schedules maintenance operations to aircraft.
 In this problem, aircraft are needed to comply with a series of planned missions while at the same time needing to receive frequent preventive maintenances in order to be ready for new missions.
 
-  It is a variant of the Flight and Maintenance Planning problem, where maintenances and flights are scheduled to commercial aircraft.
+  It is a variant of the Flight and Maintenance Planning problem, where maintenance operations and flights are scheduled to commercial aircraft.
 The main differences are the fact that flights are not modeled geographically: all aircraft are assumed to exit the main depot and return to the same depot at the end of the flight.
@@
 
-  The temporal scope of the problem is also larger, making the normal planning horizon larger and each period larger too: months instead of days.
+  The temporal scope of the problem is also larger, making the normal planning horizon and each period larger too: years instead of months and months instead of days, respectively.
 
@@
 
-  \cite{Kozanidis2008} presents a model for the Hellenic Air Force.
-It assigns flight hours to aircraft and scheduled flexible maintenance periods based on a "remaining usage time" rule.
+  In \cite{Kozanidis2008}, a model for the Hellenic Air Force is presented.
+It assigns flight hours to aircraft and schedules flexible maintenance periods based on a "remaining usage time" rule.
 The objectives are to maximize this "remaining usage time" for the aircraft as well as to smooth the availability of aircraft during the planning horizon.
@@
 
-  In \cite{gavranis2015exact}, an exact algorithm was presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
+  In \cite{gavranis2015exact}, an exact algorithm is presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
 
@@
   \label{sec:model}
-  \subsection{Parameters}
 
@@
     \begin{tabular}{p{5mm}p{70mm}}
+        % $\mathcal{I}$     &   \\
+        % $\mathcal{T}$     &   \\
+        % $\mathcal{J}$     &   \\
         $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
@@
     \end{tabular}
+
+  \subsection{Parameters}
 
@@
     \begin{tabular}{p{8mm}p{67mm}}
-        $a_{jti}$   &  =1 if task $j \in J$ in period $t\in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
-        $m_{it}$    &  =1 if resource $i \in I$ starts maintenance in period $t \in \mathcal{T}$, 0 otherwise. \\
+        $a_{jti}$   &  =1 if task $j \in J$ in period $t \in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
+        $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
         $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
@@
     Similarly, the remaining elapsed time is defined by (\ref{eq:ret_upper})-(\ref{eq:ret_initial}) and its limits by (\ref{eq:ret_lower})-(\ref{eq:me}). 
-    Finally, constraints \ref{eq:min_ret} and \ref{eq:min_rut} guarantee that resources have, globally, the same amount of remaining used and elapsed times that at the beginning.
+    Finally, constraints \ref{eq:min_ret} and \ref{eq:min_rut} guarantee that resources have, globally, the same amount of remaining used and elapsed times at the beginning and at the end of the planning horizon.
 
@@
         \item The original instance was the largest instance where an integer solution was found in less than 3600 seconds: it contains 11 tasks and 21 periods in total.
-        \item We take out the task with the biggest number of required resources.
+        \item We took out the task with the biggest number of required resources.
         \item We increased the number of periods in multiples of 10 until no integer solutions can be found in 3600 seconds.
@@
 
-  \caption{Instances used in the experiments with indicators of size.
-"id" is the instance; $|\mathcal{T}|$ the number of periods; "assign" the number of assignments of resources to tasks; $|\mathcal{J}|$ the number of tasks; "vars" the number of variables; "cons" the number of constraints; and "nonzeros" the number of non zero values in the matrix.}
+  \caption{Instances used in the experiments with indicators of size: "id" is the instance; $|\mathcal{T}|$ is the number of periods; "assign" is the number of assignments of resources to tasks; $|\mathcal{J}|$ is the number of tasks; "vars" the number of variables; "cons" is the number of constraints; and "nonzeros" is the number of non zero values in the matrix.}
   \vspace{-0.5cm}
@@
 The most loaded and second most loaded tasks demand 50 and 20 resources respectively per period.
-    Candidates resources are calculated based on each task's requirements matching the candidate's capabilities.
+    Candidate resources are calculated based on each task's requirements matching the candidate's capabilities.
 
@@
 
-    There are 127 resources in total. Each one has its functionalities and past assignments.
+    There are 127 resources in total. Each one has its own functionalities and past assignments.
 
@@
 
-  \section{Numerical tests}
+  \subsection{Numerical tests}
     \label{sec:tests}
@@
 
-    The greatest number of used was of 'Implied bound' and 'Mixed integer rounding', depending on the instance.
+    The greatest number of cuts used was of 'Implied bound' and 'Mixed integer rounding', depending on the instance.
 Since these are two generic-type cuts, it can be concluded that the solver has not detected any particular structure as part of this problem.
@@
 
-    \caption{Solution details on the progress of the lower bound.
-"id" shows the name of the instance; "root" the relaxation before cuts; "b.
-cuts" the relaxation after the cuts; "bound" the relaxation at the time limit; "cuts (\#)" the number of cuts; and "cuts (s)" the time the cuts took in seconds.}
+    \caption{Solution details on the progress of the lower bound: "id" is the name of the instance; "root" is the relaxation before cuts; "b.
+cuts is the relaxation after the cuts; "bound" is the relaxation at the time limit; "cuts (\#)" is the number of cuts; and "cuts (s)" is the time the cuts took in seconds.}
     \vspace{-0.5cm}
@@
 For more detail on this progress, figure \ref{fig:progress} shows an example of progress for instance with id=I\_7.
-It can be seen that initial solutions are quite far from the optimal one due to the use of heuristics at the beginning of the solving process.
+It can be seen that initial solutions are quite far from the optimal ones due to the use of heuristics at the beginning of the solving process.
 In fact, for two rather small instances the solver was able to find an optimal solution before starting to branch.
@@
     The obtained results show that the branching is not providing much improvement in increasing the lower bound or finding better integer solutions.
-The quantity of symmetries in the model, represented as candidates for each task could be a reason the difficulty in solving large instances.
-It may exist a very large number of possible candidates to be potentially assigned for each task with respect to the requirements in resources for each task.
-This reduces the branching capacity for the solver, which tries to explore many nodes that are almost equivalent in reality.
+The quantity of symmetries in the model, represented as candidates for each task, could be a reason the difficulty in solving large instances.
+It may exist a very large number of possible candidates to be potentially assigned for each task with respect to the requirements of resources for each task.
+This reduces the branching capacity of the solver, which tries to explore many nodes that are almost equivalent in reality.
 
@@
 
-    \caption{Solution details on the progress of the integer solution.
-"id" shows the name of the instance; "first" is the first solution integer found; "sol.
-cuts" the solution after the cuts; "last" is the last solution found at the time limit.}
+    \caption{Solution details on the progress of the integer solution: "id" is the name of the instance; "first" is the first solution integer found; "sol.
+cuts" is the solution after the cuts; "last" is the last solution found at the time limit.}
     \vspace{-0.5cm}
@@
 
-    Table \ref{tab:multmultiobj} presents the obtained results of this analysis. It can be seen that all instances have a small number of Pareto optimal solutions.
+    Table \ref{tab:multmultiobj} presents the obtained results of this analysis. It can be seen that all instances have a small number of non-dominated solutions.
 
@@
     \caption{Non-dominated points found on each instance after 3600 seconds.
-All instances were analyzed.
-"First" and "Last" indicate the two extremes of the Pareto frontier.
-"\# points" indicates the number of Pareto optimal points found.
+All instances were analyzed: "First" and "Last" indicate the two extremes of the calculated Pareto frontier; "\# points" indicates the number of Pareto optimal points found.
 Not all experiments were solved up to optimality.}
@@
   Two objective functions were chosen to address the French Air Force requirements.
-  The first one to smooth maintenance operations and the second one to aircraft availability.
+  The first one was to smooth maintenance operations and the second one was to increase aircraft availability.
   The characteristics of MIP convergence were analyzed on real-world instances.

### 2018-01-30 latex/MOSIM2018/article.tex
small changes in result tables and some small changes in article.

@@
 %\usepackage{timesnew}
-\usepackage{harvard}
+% \usepackage{harvard}
+\usepackage[]{natbib}
 \usepackage{amsmath}
@@
 
-  Flight and Maintenance Planning (FMP) problems raise in civilian or military operational applications. 
+  Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications. 
   They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.
@@
 
-  % Although the former has been studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015}, it has not received as much attention as the latter.
+  % Although the former has been studied in literature in \citet{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015}, it has not received as much attention as the latter.
 In the following, the problem with already scheduled missions is considered.
@@
 
-  In \cite{Kozanidis2008}, a model for the Hellenic Air Force is presented.
+  In \citet{Kozanidis2008}, a model for the Hellenic Air Force is presented.
 It assigns flight hours to aircraft and schedules flexible maintenance periods based on a "remaining usage time" rule.
@@
 
-  In \cite{Cho2011}, the objective is to smooth the maintenance operations.
+  In \citet{Cho2011}, the objective is to smooth the maintenance operations.
 Randomly generated instances of 15 aircraft and 520 bi-daily periods were solved with MIP formulations.
@@
 
-  In \cite{Verhoeff2015}, a model for the Royal Netherlands Air Force is developed.
+  In \citet{Verhoeff2015}, a model for the Royal Netherlands Air Force is developed.
 Here, three criteria for improving the planning robustness are presented and applied in a MIP model: availability, serviceability and sustainability.
 
-  In \cite{gavranis2015exact}, an exact algorithm is presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
+  In \citet{gavranis2015exact}, an exact algorithm is presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
 
-  Both \cite{Kozanidis2008} and \cite{Cho2011} develop heuristic alternatives to the MIP formulation for solving larger problem instances.
-In \cite{steiner2006heuristic}, a constructive, multi-stage heuristic approach is described to obtain balanced maintenances for the Swiss Air Force.
+  Both \citet{Kozanidis2008} and \citet{Cho2011} develop heuristic alternatives to the MIP formulation for solving larger problem instances.
+In \citet{steiner2006heuristic}, a constructive, multi-stage heuristic approach is described to obtain balanced maintenances for the Swiss Air Force.
 Both calendar-based maintenances actions (CBMA) and usage-based maintenance actions (UBMA) are used.
-Finally, in \cite{kozanidis2014heuristics}, two heuristics are presented to solve large instances of the problem by taking consecutive decisions period by period with the help of mathematical models at each step.
+Finally, in \citet{kozanidis2014heuristics}, two heuristics are presented to solve large instances of the problem by taking consecutive decisions period by period with the help of mathematical models at each step.
 
-  % Regarding operational-level planning, a more detailed implementation can be found in \cite{safaei2011workforce}, where the specific conditions during the maintenance operations are taken into account to schedule the maintenance jobs, such as the workforce skills and availability.
+  % Regarding operational-level planning, a more detailed implementation can be found in \citet{safaei2011workforce}, where the specific conditions during the maintenance operations are taken into account to schedule the maintenance jobs, such as the workforce skills and availability.
 
@@
 
-  Lastly, in \cite{Chastellux2016} a mission-centered modeled is presented.
+  Lastly, in \citet{Chastellux2016} a mission-centered modeled is presented.
 Here, cycles, composed of one maintenance followed by mission assignments were used.
@@
   \label{sec:data}
-  The dataset being used was provided by the French Air Force and already used in \cite{Chastellux2017}.
+  The dataset being used was provided by the French Air Force and already used in \citet{Chastellux2017}.
   Based on the information provided, a number of instances were created as subsets of the main dataset in order to run different experiments.

### 2018-01-31 latex/MOSIM2018/article.tex
last changes to MOSIM before sending it

@@
 \begin{tabular}{c}
-\bf \normalsize {Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin} \\
+\bf \normalsize {Franco Peschiera, Nicolas Dupin, Olga Battaïa, Alain Haït} \\
 \\
@@
      \normalsize 31055 Toulouse cedex 4 - France\\
-     \normalsize {\{franco.peschiera,alain.hait,olga.battaia,nicolas.dupin\}@isae-supaero.fr}
+     \normalsize {\{franco.peschiera,nicolas.dupin,olga.battaia,alain.hait\}@isae-supaero.fr}
      \\
@@
     The further calculation was done in order to analyze the impact of the weights $W_1$ and $W_2$ on the solution.
-In it, both weights were changed between 0 and 1 with pace of 0.1 under the condition that $W_1 + W_2 = 1 $.
-This was done for all previously defined instances.
-
+    In it, both weights were changed between 0 and 1 with pace of 0.1 under the condition that $W_1 + W_2 = 1 $. This was done for all previously defined instances.
+    This method does not a priori guarantee to compute exactly Pareto Fronts: optimal weighted computations can only prove that some points guarantee supported solutions.
+    However, a posteriori analysis can conclude that the bi-objective convex relaxation guarantees the completeness of the Pareto Front.
+By doing this analysis, exact extreme points can be computed (and thus the Nadir and ideal points), and some intermediate points found for all the intermediate integer values.
     Table \ref{tab:multmultiobj} presents the obtained results of this analysis. It can be seen that all instances have a small number of non-dominated solutions.
@@
 
-    More particularly, an individual instance was analyzed, where an optimal solution was found in each experiment.
+    An individual case study is illustrated in Figure \ref{fig:pareto}, where an optimal solution was found in each experiment.
     Table \ref{tab:multiobj} shows the obtained results. The instance used consisted of 21 periods and 10 tasks corresponding to instance with id=I\_5.
+    Computations $W10$ and $W09$ prove that $(15,68)$ is an extreme point of the Pareto Front.
+    Computations $W00$ and $W01$ prove that $(20,66)$ is the other extreme point of the Pareto Front. This implies the ideal and Nadir points are, respectively, $(15,66)$ and $(20,68)$.
+    Computation $W03$ proves that $(16,67)$ is also a non dominated point.
+    Having all the discrete points in $[\![66,68]\!]$ for the $m_{max}$ objective, it can be concluded that no other non dominated points exists.
+    In this case, weighted computations prove the completeness of the convex relaxation, which is also exactly the Pareto front of non-dominated solutions: no unsupported solution exists in this case.
 
-    For all solved instances, a small number of non dominated solutions was observed.
     \begin{table}[h]

### 2018-02-14 R/roadef_presentation/roadef_presentation.Rmd
Added R files with ROADEF presentation and graphs code

---
title: "ROADEF"
author: "Franco Peschiera"
date: "February 14, 2018"
output: 
  slidy_presentation:
    css: ./../../css/baobab_MH.css
    incremental: yes
  beamer_presentation:
    incremental: yes
nocite: '@*'
bibliography: MFMP.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("../")
source('scripts/params.R')
source('scripts/import_results.R')
exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
gantt <- print_solution(exp_directory, max_resources = 30)
```

## Introduction

## Problem

**Informally**:

Assign both missions and maintenance operations to a fleet of aircraft in order to maximize availability and minimize costs. Missions have particular needs in terms of aircraft and time.

## Example

<div style="font-size: 20px;">
```{r}
gantt
```
</div>

## Problem

**Formally**:

A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.

Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period. Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used. The number of resources in use for each period $d_t$ is also known.

## Relevance and origin

The problem was born out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet.

In fact, a similar subject has been presented as a challenge in a open competition: Défit Opti-plan.

## State of the art

* In @Cho2011, US Army aircraft was assigned daily operations over a year to aircraft.
* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours. 
* In @Verhoeff2015, monthly assignments were also done, and several objectives were taken into account: availability, serviceability and sustainability.
* In @Chastellux2017, a first version of the model was done, assigning missions instead of flight hours.

## State of the art (2)

The present model assigns missions to aircraft, permitting to add more complex functionality, such as minimum assignment durations.

This model also features several improvements in the MIP modelling that permit greater instance sizes.

## Model: sets

|Set|Description|
|-----|--------------------------------------------------------------------|
|    $\mathcal{T}_j$     |  time periods $t \in \mathcal{T}$ in which task $j$ is active. |
|    $\mathcal{J}_t$    |  tasks $j \in \mathcal{J}$ to be realized in period $t$. |
|    $\mathcal{I}_j$     |  resources $i \in \mathcal{I}$ that can be assigned to task $j$. |
|    $\mathcal{O}_i$     |  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. |
|    $\mathcal{T}^{s}_t$ |  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. |

## Model: parameters

Only *main parameters* are shown

|Parameters|Description|
|--------------|-------------------------------------------------|
| $H_j$ |  amount of **resource time required** by task $j$. |
| $R_j$ |  number of **resources required** by task $j$. |
| $M$   |  **maintenance duration** in number of periods. |
| $E$   |  remaining **elapsed time** after a maintenance. |
| $H$   |  remaining **usage time** after a maintenance. |

<!-- ## Model: parameters (2) -->

<!-- |Parameters|Description| -->
<!-- |--------------|----------------------------------------------------------------------| -->
<!-- | $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.| -->
<!-- | $D_t$              | number of resources to be assigned in total in period $t$. | -->
<!-- | $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. | -->
<!-- | $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. | -->
<!-- | $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. | -->
<!-- | $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. | -->

## Model: variables

| Variable   |  Description |
|-----------|------------------------------------------------------- |
| $a_{jti}$   |  =1 if **task** $j \in J$ in period $t \in \mathcal{T}_j$ **is realized with resource** $i \in \mathcal{I}_j$, 0 otherwise. |
| $m_{it}$    |  =1 if resource $i \in I$ **starts a maintenance** operation in period $t \in \mathcal{T}$, 0 otherwise. |
| $rut_{it}$  |  **remaining usage time** (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. |
| $ret_{it}$  |  **remaining elapsed time** (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. |
| $u_{max}$   |  maximal number (integer) of **unavailable resources** in any period. |
| $m_{max}$   |  maximal number (integer) of **resources in maintenance** in any period. |

## Model: constraints (1)

$$
\begin{align}
&\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
&t \in \mathcal{T} \\
&\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
&t \in \mathcal{T}\\
&\sum_{i \in \mathcal{I}_j} a_{jti} = R_j
&j \in \mathcal{J}, t \in \mathcal{T}_j\\
&\sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
& t \in \mathcal{T}, i \in \mathcal{I}
\end{align}
$$

## Model: constraints (2)

Only usage time is shown. Elapsed times are treated in the same way.

$$
\begin{align}
& rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
& rut_{i0} = Rut^{Init}_i
        & i \in \mathcal{I}\\
& rut_{it} \geq H m_{it}
        & t \in \mathcal{T}, i \in \mathcal{I}\\ 
& rut_{it} \in [0,H]
        & t \in \mathcal{T}, i \in \mathcal{I} \\
& \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
        & t = |\mathcal{T}|
\end{align}
$$

## Results

## Multi-objective

## Next steps

## References

### 2018-02-14 R/scripts/graphs.R

@@ -1,0 +1,19 @@
+source('scripts/params.R')
+
+exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
+print_solution(exp_directory, max_resources = 30)
+
+
+    
+#       id   start     end   content group                                         style
+#    <int>   <chr>   <chr>    <fctr> <chr>                                         <chr>
+#  1     1 2015-01 2016-01      <NA>  D601               background-color:;border-color:
+#  2     4 2016-01 2016-02 [ 0,   4)  D601 background-color:#FFFFB2;border-color:#FFFFB2
+#  3     5 2016-02    <NA> [48,1200]  D601 background-color:#BD0026;border-color:#BD0026
+#  4    49 2015-01 2016-02 [ 0,   4)  D602 background-color:#FFFFB2;border-color:#FFFFB2
+#  5    50 2016-02 2016-03 [ 4,  23)  D602 background-color:#FECC5C;border-color:#FECC5C
+#  6    51 2016-03 2016-04 [48,1200]  D602 background-color:#BD0026;border-color:#BD0026
+#  7    52 2016-04 2016-05 [24,  48)  D602 background-color:#F03B20;border-color:#F03B20
+#  8    53 2016-05 2016-07 [ 4,  23)  D602 background-color:#FECC5C;border-color:#FECC5C
+#  9    55 2016-07 2016-08 [24,  48)  D602 background-color:#F03B20;border-color:#F03B20
+# 10    56 2016-08 2016-09 [ 4,  23)  D602 background-color:#FECC5C;border-color:#FECC5C

### 2018-02-14 R/scripts/import_results.R

@@ -9,22 +9,93 @@ library(timevis)
 library(Hmisc)
 library(RColorBrewer)
 library(htmlwidgets)
+library(jsonlite)
 
-collapse_states <- function(table, column_name){
+collapse_states <- function(table){
     
-    series <- table[[column_name]]
+    # we complete missing months
+    months <- table$month
+    resources <- table %>% extract2('UNIT')
+    min_date <- months %>% min %>% paste0('-01') %>% as.Date()
+    max_date <- months %>% max %>% paste0('-01') %>% as.Date()
+    total_rows <- 
+        seq(from= min_date, to=max_date, "months") %>% 
+        as.character() %>% 
+        str_sub(end='7') %>% 
+        CJ(month= ., UNIT=resources)
     
     table %>% 
-        mutate(temp = series) %>% 
+        full_join(total_rows) %>% 
+        mutate(state = if_else(is.na(state), "", state),
+               temp = state) %>% 
         arrange(UNIT, month) %>% 
-        group_by(UNIT) %>% 
+        group_by(UNIT) %>%  
         mutate(prev_temp= temp %>% lag,
                first_row= row_number()==1,
                change = temp != prev_temp) %>% 
         filter(first_row | change) %>% 
-        mutate(post_month = month %>% lead %>% paste0("-01") %>% ymd() %>% add(0) %>% format("%Y-%m")) %>% 
+        mutate(post_month = month %>% lead %>% paste0("-01") %>% ymd() %>% add(0) %>% format("%Y-%m"),
+               post_month = post_month %>% if_else(is.na(.), max_date %>% ymd() %>% format("%Y-%m"), .)) %>% 
         select(-temp, -prev_temp, -first_row, -change) %>% 
-        ungroup()
+        ungroup() %>% filter(state != "")
+}
+
+print_solution <- function(exp_directory, max_resources=NULL){
+    solution_path = exp_directory %>% paste0('data_out.json')
+    input_path = exp_directory %>% paste0('data_in.json')
+    
+    solution <- read_json(solution_path)
+    input <- read_json(input_path)
+    
+    task_hours <-
+        input %>% 
+        extract2('tasks') %>% 
+        lapply("[[", 'consumption') %>% 
+        bind_rows() %>% 
+        gather(key='state', value = 'hours') %>% 
+        bind_rows(data.table(state='M', hours=0)) %>% 
[…]

### 2018-02-14 R/diagrams/heuristic.Rmd
Added markdown with heuristic documentation

---
title: "heuristic1"
author: "Franco Peschiera"
date: "January 9, 2018"
output:
    html_document:
        # css: ./../../css/baobab.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic idea

The algorithm consists on a greedy logic to fill, one mission at a time, with aircraft until the needed amount is reached. It tries to use each aircraft as much as possible until it reaches its maximum flight time. When this happens: the aircraft is granted a maintenance, if possible and as soon as possible. When the aircraft has all its slots full or can't be assigned a maintenance and needs one, the algorithm continues with the next aircraft.

## Pseudocode

With the candidate and the task we do the following:

    for every task:
        1. calculate candidates.
        for each candidate until we run out of candidates or we satisfy task's needs:
            1. find available periods.
            Until we run out of periods:
                1. we check if we can assign it (checking remaining time and elapsed time). 
                    * If we can't: we tag the aircraft as "in need of a maintenance'.
                    * if we can: we assign the task.
                2. After assigning a task: we update the remaining and elapsed times for the aircraft.

For assigning maintenances is more or less similar:

    if the aircraft is tagged as 'in need of a maintenance':
        1. we check if we can reserve 6 continous months of maintenance.
            * if we can: we do it (and continue assigining tasks to the aircraft).
            * if we cannot: we go to the next candidate.
        2. After assigning a maintenance: we update the remaining and elapsed times for the aircraft.

<!-- <br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br> -->
## Main python code

I think the code gives an idea of the logic behind it. I only include the code for each mission so we would have to execute this function for each mission (or task).

`task` is the mission to satisfy.

* **candidates**: a list with the universe of aircraft that can be assigned to that task.
* **rem_resources**: remaining aircraft that need to be assigned at each period of the task's duration.
* **periods_task**: intersection of available slots for the aircraft (periods without taks or maintenance) and the task's period's needs.

```python

def fill_mission(self, task):
    dtype = 'U7'
    # get task candidates: we copy the list.
    candidates = self.instance.data['tasks'][task]['candidates'][:]
    # get task periods to satisfy:
    rem_resources = \
        aux.dicttup_to_dictdict(
            self.instance.get_task_period_needs()
        )[task]

    while len(rem_resources) > 0 and len(candidates) > 0:
        # candidate = 'A124'
        candidate = candidates[0]
        # get free periods for candidate
        periods_task = \
            self.get_free_starts(candidate, np.fromiter(rem_resources, dtype=dtype))

        # consider eliminating the resource from the list.
        # if all its periods are 'used'
        if len(periods_task) == 0:
            candidates.pop(0)
            continue
        print('candidate={}\ntask={}\nperiods={}'.format(candidate, task, periods_task))
        maint_period = ''
        # TODO: this is still wrong:
        for start, end in periods_task:
            for period in aux.get_months(start, end):
                result = self.check_set_task(candidate, period, task)
                if not result:
                    maint_period = period
                    break
                rem_resources[period] -= 1
            self.update_time_usage(candidate, period, self.get_next_maintenance(candidate, period), 'rut')

        if maint_period != '':
            # find soonest period to start maintenance:
            maint_start = self.get_soonest_maint(candidate, maint_period)
            if maint_start is None:
                # this means that we cannot assign tasks BUT
                # we cannot assign maintenance either :/
                # we should then take out the candidate:
                candidates.pop(0)
                continue
            maint_end = aux.shift_month(maint_start, 5)
            periods_maint = aux.get_months(maint_start,  maint_end)
            for period in periods_maint:
                self.set_maint(candidate, period)
            self.update_time_maint(maint_start, maint_end)
            next_month = aux.get_next_month(maint_end)
            self.update_time_usage(candidate, next_month, 'ret')
            self.update_time_usage(candidate, next_month, 'rut')

```

### 2018-02-15 R/roadef_presentation/roadef_presentation.Rmd
Changes in ROADEF presentation

@@
 ---
-title: "ROADEF"
-author: "Franco Peschiera"
+title: "Maintenance planning on French military aircraft operations"
+author: "Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin"
 date: "February 14, 2018"
 output: 
+  beamer_presentation:
+    theme: "Madrid"
+    toc: true
+    slide_level: 2
+    incremental: yes
+    fig_height: 4
   slidy_presentation:
@@
     incremental: yes
-  beamer_presentation:
-    incremental: yes
+    footer: "ISAE-SUPAERO, Université de Toulouse, France"
 nocite: '@*'
@@
 source('scripts/import_results.R')
+# library(png)
+# library(grid)
 exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
 gantt <- print_solution(exp_directory, max_resources = 30)
+show_gantt <- knitr::opts_knit$get("rmarkdown.pandoc.to")=="slidy"
 ```
@@
 ## Introduction
+
+Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications.
+
+They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.
 
@@
 
-<div style="font-size: 20px;">
-```{r}
+```{r, echo=FALSE, out.width='100%'}
+if (show_gantt){
+    htmltools::HTML('<div style="font-size: 20px;width:100%">')
     gantt    
+    htmltools::HTML("</div>")
+} else {
+    knitr::include_graphics('./../../img/calendar.png')
+}
+
 ```
-</div>
 
@@
 
-A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
+* A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
 Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
 Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.
-
-Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
+* Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
 Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
@@
 
-The problem was born out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet.
+This thesis was started out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet while controlling cost and maintenance capacity.
 
@@
 
-* In @Cho2011, US Army aircraft was assigned daily operations over a year to aircraft.
+* In @Cho2011, US Army aircraft were assigned daily operations over a year to aircraft.
 * In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours. 
-* In @Verhoeff2015, monthly assignments were also done, and several objectives were taken into account: availability, serviceability and sustainability.
-* In @Chastellux2017, a first version of the model was done, assigning missions instead of flight hours.
+* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and sustainability.
+* In @Chastellux2017, an initial version of the model was done, assigning missions instead of only flight hours.
 
@@
 
-The present model assigns missions to aircraft, permitting to add more complex functionality, such as minimum assignment durations.
-
-This model also features several improvements in the MIP modelling that permit greater instance sizes.
+* The present model deals with an heterogeneus fleet, with different standards and capacities.
+* Multiple objectives have been incorporated in order to satisfy the French Air Force's specific needs.
+* This model also features several improvements in the MIP modelling that permit greater instance sizes.
 
@@
 |-----|--------------------------------------------------------------------|
+|    $\mathcal{I}$       |  resources. |
+|    $\mathcal{T}$       |  time periods inside planning horizon. |
+|    $\mathcal{J}$       |  tasks. |
 |    $\mathcal{T}_j$     |  time periods $t \in \mathcal{T}$ in which task $j$ is active. |
@@
 
-Only *main parameters* are shown
+(Only **main parameters** are shown).
 
@@
 
-<!-- ## Model: parameters (2) -->
+## Model: parameters (2)
 
-<!-- |Parameters|Description| -->
-<!-- |--------------|----------------------------------------------------------------------| -->
-<!-- | $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.| -->
-<!-- | $D_t$              | number of resources to be assigned in total in period $t$. | -->
-<!-- | $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. | -->
-<!-- | $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. | -->
-<!-- | $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. | -->
-<!-- | $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. | -->
+(**secondary parameters**).
+
+|Parameters|Description|
+|--------------|----------------------------------------------------------------------|
+| $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.|
+| $D_t$              | number of resources to be assigned in total in period $t$. |
+| $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. |
+| $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. |
+| $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. |
+| $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. |
 
@@
 
-$$
+<div class="incremental">
+
+The two components of the objective function: the max number of maintenance and the max number of unavailable aircraft.
 \begin{align}
@@
     &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max}
-&t \in \mathcal{T}\\
+    &t \in \mathcal{T}
+\end{align}
+
+Mission's needs and incompatibility of several tasks or maintenance in the same period.
+\begin{align}
 &\sum_{i \in \mathcal{I}_j} a_{jti} = R_j
@@
 \end{align}
-$$
+
+</div>
 
@@
 
-Only usage time is shown. Elapsed times are treated in the same way.
-
-$$
+Time usage of an aircraft depending on the assignment to a task.
+<!-- <div class="incremental" style="margin-left: 4em; position: relative">  -->
+<!--   <div style="position: absolute; left: 0; top: 0;background-color:white;width:100%"> -->
 \begin{align}
@@
 \end{align}
-$$
+
+  <!-- </div>  -->
+  <!-- <div style="position: absolute; left: 0; top: 0;background-color:white;width:100%"> -->
+
+## Model: constraints (3)
+
+Elapsed times are treated in the same way but consumption is always 1.
+\begin{align}
+& ret_{it} \leq ret_{it-1} + E m_{it} - 1 & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
+& ret_{i0} = Ret^{Init}_i
+        & i \in \mathcal{I}\\
+& ret_{it} \geq E m_{it}
+        & t \in \mathcal{T}, i \in \mathcal{I}\\
+& ret_{it} \in [0,E]
+        & t \in \mathcal{T}, i \in \mathcal{I} \\
+& \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
+        & t = |\mathcal{T}|
+\end{align}
+
+## Model: constraints (4)
+
+Visualization of each constraints for one resource.
+
+```{r out.width="250px", echo=FALSE}
+# img <- readPNG("./../../img/balance2.png")
+#  grid.raster(img)
+knitr::include_graphics("./../../img/balance2.png")
+```
 
@@
 
-## Multi-objective
+The following instances were tested:
+
+| id | $\|\mathcal{J}\|$ |$\|\mathcal{T}\|$ |  assign |  objective  |  time (s) |   bound |
+|-----|------------------|------------------|---------| ------------| -----------| --------|
+| I_0 |                9 |               11 |     310 |        62.0 |        0.7 |    62.0 |
+| I_1 |                9 |               21 |     650 |        63.0 |       68.7 |    63.0 |
+| I_2 |                9 |               31 |     990 |        63.0 |     3600.1 |    62.0 |
+| I_3 |                9 |               41 |    1249 |        64.0 |     3603.9 |    61.7 |
+| I_4 |               10 |               11 |     530 |        82.0 |        0.9 |    82.0 |
+| I_5 |               10 |               21 |    1070 |        83.0 |      144.0 |    83.0 |
+| I_6 |               10 |               31 |    1610 |        83.0 |     3600.1 |    82.0 |
+| I_7 |               10 |               41 |    2069 |        84.0 |     3609.1 |    81.8 |
+| I_8 |               11 |               11 |    1080 |       139.0 |      530.6 |   139.0 |
+| I_9 |               11 |               21 |    2120 |       149.0 |     3600.0 |   139.9 |
+
+## Multi-objective (1)
+
+Comparison of two objectives: max availability or maintenance capacity.
+Different weights were used for the two objectives: $W_1, W_2 \in [0..1]$ with pace of 0.1
+```{r out.width="200pt", echo=FALSE}
+# img <- readPNG()
+#  grid.raster(img)
+knitr::include_graphics("./../../img/multiobjective.png")
+
+```
+
+Pareto diagram for instance id=I_5.
+
+## Multi-objective (2)
+
+* Analysis for all instances: not many pareto optimal points available.
+* Between 2 and 4 for all instances
+* The objectives are quite aligned one with the other.
 
@@
 
+* **Try different objectives**: availability and maintenances.
+* **Try more linent objectives**: relax the `max` in the objective.
+* **Break symmetries**: clustering candidates, fixing end availabilities.
+* **Add more constraints**: minimal duration of assignments, storage of aircraft.
+
 ## References

### 2018-02-15 R/scripts/import_results.R

@@ -74,9 +74,10 @@ print_solution <- function(exp_directory, max_resources=NULL){
         collapse_states() %>% 
         inner_join(task_hours) %>% 
         mutate(id= c(1:nrow(.)),
-               style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color)
+               style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color),
+               content = if_else(state=='M', 'M', sprintf('%s (%sh)', state, hours))
         ) %>% 
-        select(id, start= month, end= post_month, content = state, group= UNIT, style)
+        select(id, start= month, end= post_month, content, group= UNIT, style)
     
     if (max_resources %>% is.null %>% not){
         resources <- states %>% distinct(group) %>% slice(1:max_resources)

### 2018-02-16 R/roadef_presentation/roadef_presentation.Rmd
new version of ppt presented to Olga friday. Need to still take out many slides.

@@
     footer: "ISAE-SUPAERO, Université de Toulouse, France"
+    slide_level: 2
+    toc: true
 nocite: '@*'
@@
 
+# Introduction
+
 ## Introduction
@@
 
-## Problem
+# Problem
 
-**Informally**:
+## Problem (informally)
 
@@
 
-## Example
+## Problem (example)
 
@@
 
-## Problem
-
-**Formally**:
+## Problem (formally)
 
@@
 
-## State of the art
+# State of the art
+
+## State of the art (1)
 
@@
 * This model also features several improvements in the MIP modelling that permit greater instance sizes.
+
+# Model
 
@@
 
-## Model: parameters
-
-(Only **main parameters** are shown).
+## Model: main parameters
 
@@
 
-## Model: parameters (2)
-
-(**secondary parameters**).
+## Model: more parameters
 
@@
 
-## Model: constraints (1)
+## Model: main constraints
+
+Requirements per task:
+```{r out.width="250px", echo=FALSE}
+knitr::include_graphics("./../../img/tasks_gantt.png")
+```
+\pause
+
+Max maintenances:
+```{r out.width="150px", echo=FALSE}
+knitr::include_graphics("./../../img/num-maintenances.png")
+```
+
+## Model: main constraints
 
@@
 
+\pause
+
 Mission's needs and incompatibility of several tasks or maintenance in the same period.
@@
 
-## Model: constraints (2)
+## Model: flow constraints
+
+Visualization of accumulated time constraints for one resource.
+
+```{r out.width="250px", echo=FALSE}
+# img <- readPNG("./../../img/balance2.png")
+#  grid.raster(img)
+knitr::include_graphics("./../../img/balance2.png")
+```
+
+## Model: flow constraints
 
@@
 
-## Model: constraints (3)
+## Model: flow constraints
 
@@
 
-## Model: constraints (4)
+# Results
 
-Visualization of each constraints for one resource.
-
-```{r out.width="250px", echo=FALSE}
-# img <- readPNG("./../../img/balance2.png")
-#  grid.raster(img)
-knitr::include_graphics("./../../img/balance2.png")
-```
-
-## Results
+## Experiments
 
@@
 
+# Next steps
 ## Next steps
@@
 
+# References
 ## References
+
+<!-- \bibliographystyle{dcu} -->
+<!-- \bibliography{./../biblio/MFMP} -->

### 2018-02-16 R/scripts/import_results.R

@@ -40,6 +40,31 @@ collapse_states <- function(table){
         ungroup() %>% filter(state != "")
 }
 
+print_tasks <- function(exp_directory){
+    input_path = exp_directory %>% paste0('data_in.json')
+    input <- read_json(input_path)
+    tasks <-
+        input %>% 
+        extract2('tasks') %>% 
+        lapply("[", c('consumption', 'num_resource', 'start', 'end')) %>% 
+        bind_rows(.id="group") %>% 
+        mutate(id= c(1:nrow(.)),
+               content = sprintf('%s aircraft (%sh)', num_resource, consumption),
+               color = RColorBrewer::brewer.pal(5, "YlOrRd")[c(1, 2)],
+               style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color),
+        )
+    config <- list(
+        stack = FALSE,
+        editable = TRUE,
+        align = "center",
+        orientation = "top",
+        snap = NULL,
+        margin = 0
+    )
+    groups <- tasks %>% distinct(group) %>% rename(id= group) %>% mutate(content= id)
+    timevis(tasks, groups= groups, options= config, width="100%")
+}
+
 print_solution <- function(exp_directory, max_resources=NULL){
     solution_path = exp_directory %>% paste0('data_out.json')
     input_path = exp_directory %>% paste0('data_in.json')

### 2018-02-19 R/roadef_presentation/roadef_presentation.Rmd
Added changes in ROADEF presentation after last test.

@@
 title: "Maintenance planning on French military aircraft operations"
-author: "Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin"
-date: "February 14, 2018"
+author: |
+    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
+    | \tiny (franco.peschiera@isae-supaero.fr)
+institute: "ISAE SUPAERO"
+date: "February 22, 2018"
 output: 
   beamer_presentation:
-    theme: "Madrid"
+    theme: "Berlin"
     toc: true
@@
     fig_height: 4
+    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
+    includes:
+        in_header: header.tex
   slidy_presentation:
@@
 
-# Introduction
+<!-- # Introduction -->
 
-## Introduction
+<!-- ## Introduction -->
 
-Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications.
+<!-- Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications.  -->
 
-They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.
+<!-- They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.
+-->
 
@@
 
-Assign both missions and maintenance operations to a fleet of aircraft in order to maximize availability and minimize costs. Missions have particular needs in terms of aircraft and time.
+Assign both missions and maintenance operations to a fleet of aircraft in order to maximize availability and minimize costs.
+Missions have fixed start and end times and have particular needs in terms of aircraft and time.
 
@@
 
-```{r, echo=FALSE, out.width='100%'}
+```{r, echo=FALSE, out.width='70%'}
+
 if (show_gantt){
@@
 
-## Problem (formally)
+## Problem
 
@@
 
-## Relevance and origin
+<!-- ## Relevance and origin -->
 
-This thesis was started out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet while controlling cost and maintenance capacity.
+<!-- This thesis was started out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet while controlling cost and maintenance capacity. -->
 
-In fact, a similar subject has been presented as a challenge in a open competition: Défit Opti-plan.
+<!-- In fact, a similar subject has been presented as a challenge in a open competition: Défit Opti-plan. -->
 
@@
 
-* In @Cho2011, US Army aircraft were assigned daily operations over a year to aircraft.
-* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours. 
-* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and sustainability.
+* FMP: Flight and Maintenance Planning problem.
+* In @Cho2011, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
+* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
+* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
 * In @Chastellux2017, an initial version of the model was done, assigning missions instead of only flight hours.
@@
 
-## Model: sets
+<!-- ## Model: sets -->
 
-|Set|Description|
-|-----|--------------------------------------------------------------------|
-|    $\mathcal{I}$       |  resources. |
-|    $\mathcal{T}$       |  time periods inside planning horizon. |
-|    $\mathcal{J}$       |  tasks. |
-|    $\mathcal{T}_j$     |  time periods $t \in \mathcal{T}$ in which task $j$ is active. |
-|    $\mathcal{J}_t$    |  tasks $j \in \mathcal{J}$ to be realized in period $t$. |
-|    $\mathcal{I}_j$     |  resources $i \in \mathcal{I}$ that can be assigned to task $j$. |
-|    $\mathcal{O}_i$     |  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. |
-|    $\mathcal{T}^{s}_t$ |  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. |
+<!-- |Set|Description| -->
+<!-- |-----|--------------------------------------------------------------------| -->
+<!-- |    $\mathcal{I}$       |  resources. | -->
+<!-- |    $\mathcal{T}$       |  time periods inside planning horizon. | -->
+<!-- |    $\mathcal{J}$       |  tasks. | -->
+<!-- |    $\mathcal{T}_j$     |  time periods $t \in \mathcal{T}$ in which task $j$ is active. | -->
+<!-- |    $\mathcal{J}_t$    |  tasks $j \in \mathcal{J}$ to be realized in period $t$. | -->
+<!-- |    $\mathcal{I}_j$     |  resources $i \in \mathcal{I}$ that can be assigned to task $j$. | -->
+<!-- |    $\mathcal{O}_i$     |  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. | -->
+<!-- |    $\mathcal{T}^{s}_t$ |  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. | -->
 
-## Model: main parameters
+<!-- ## Model: main parameters -->
 
-|Parameters|Description|
-|--------------|-------------------------------------------------|
-| $H_j$ |  amount of **resource time required** by task $j$. |
-| $R_j$ |  number of **resources required** by task $j$. |
-| $M$   |  **maintenance duration** in number of periods. |
-| $E$   |  remaining **elapsed time** after a maintenance. |
-| $H$   |  remaining **usage time** after a maintenance. |
+<!-- |Parameters|Description| -->
+<!-- |--------------|-------------------------------------------------| -->
+<!-- | $H_j$ |  amount of **resource time required** by task $j$. | -->
+<!-- | $R_j$ |  number of **resources required** by task $j$. | -->
+<!-- | $M$   |  **maintenance duration** in number of periods. | -->
+<!-- | $E$   |  remaining **elapsed time** after a maintenance. | -->
+<!-- | $H$   |  remaining **usage time** after a maintenance. | -->
 
-## Model: more parameters
+<!-- ## Model: more parameters -->
 
-|Parameters|Description|
-|--------------|----------------------------------------------------------------------|
-| $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.|
-| $D_t$              | number of resources to be assigned in total in period $t$. |
-| $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. |
-| $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. |
-| $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. |
-| $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. |
+<!-- |Parameters|Description| -->
+<!-- |--------------|----------------------------------------------------------------------| -->
+<!-- | $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.| -->
+<!-- | $D_t$              | number of resources to be assigned in total in period $t$. | -->
+<!-- | $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. | -->
+<!-- | $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. | -->
+<!-- | $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. | -->
+<!-- | $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. | -->
 
-## Model: variables
+<!-- ## Model: variables -->
 
-| Variable   |  Description |
-|-----------|------------------------------------------------------- |
-| $a_{jti}$   |  =1 if **task** $j \in J$ in period $t \in \mathcal{T}_j$ **is realized with resource** $i \in \mathcal{I}_j$, 0 otherwise. |
-| $m_{it}$    |  =1 if resource $i \in I$ **starts a maintenance** operation in period $t \in \mathcal{T}$, 0 otherwise. |
-| $rut_{it}$  |  **remaining usage time** (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. |
-| $ret_{it}$  |  **remaining elapsed time** (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. |
-| $u_{max}$   |  maximal number (integer) of **unavailable resources** in any period. |
-| $m_{max}$   |  maximal number (integer) of **resources in maintenance** in any period. |
+<!-- | Variable   |  Description | -->
+<!-- |-----------|------------------------------------------------------- | -->
+<!-- | $a_{jti}$   |  =1 if **task** $j \in J$ in period $t \in \mathcal{T}_j$ **is realized with resource** $i \in \mathcal{I}_j$, 0 otherwise. | -->
+<!-- | $m_{it}$    |  =1 if resource $i \in I$ **starts a maintenance** operation in period $t \in \mathcal{T}$, 0 otherwise. | -->
+<!-- | $rut_{it}$  |  **remaining usage time** (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. | -->
+<!-- | $ret_{it}$  |  **remaining elapsed time** (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. | -->
+<!-- | $u_{max}$   |  maximal number (integer) of **unavailable resources** in any period. | -->
+<!-- | $m_{max}$   |  maximal number (integer) of **resources in maintenance** in any period. | -->
 
-## Model: main constraints
+<!-- ## Model: main constraints -->
 
-Requirements per task:
-```{r out.width="250px", echo=FALSE}
-knitr::include_graphics("./../../img/tasks_gantt.png")
-```
-\pause
+<!-- Requirements per task: -->
+<!-- ```{r out.width="250px", echo=FALSE} -->
+<!-- knitr::include_graphics("./../../img/tasks_gantt.png") -->
+<!-- ``` -->
+<!-- \pause -->
 
-Max maintenances:
-```{r out.width="150px", echo=FALSE}
-knitr::include_graphics("./../../img/num-maintenances.png")
-```
+<!-- Max maintenances: -->
+<!-- ```{r out.width="150px", echo=FALSE} -->
+<!-- knitr::include_graphics("./../../img/num-maintenances.png") -->
+<!-- ``` -->
 
-## Model: main constraints
+<!-- ## Model: main constraints -->
+
+## Model: objective function
 
 <div class="incremental">
+
+\begin{align}
+    & \text{Min}\; W_1 m_{max} + W_2 u_{max}
+\end{align}
+
+\pause
 
@@
 
-\pause
+</div>
+
+## Model: main constraints
 
@@
 
-</div>
+<!-- ## Model: flow constraints -->
 
-## Model: flow constraints
+<!-- Visualization of accumulated time constraints for one resource. -->
 
-Visualization of accumulated time constraints for one resource.
-
-```{r out.width="250px", echo=FALSE}
-# img <- readPNG("./../../img/balance2.png")
-#  grid.raster(img)
-knitr::include_graphics("./../../img/balance2.png")
-```
+<!-- ```{r out.width="250px", echo=FALSE} -->
+<!-- # img <- readPNG("./../../img/balance2.png") -->
+<!-- #  grid.raster(img) -->
+<!-- knitr::include_graphics("./../../img/balance2.png") -->
+<!-- ``` -->
 
@@
 
-Elapsed times are treated in the same way but consumption is always 1.
+Elapsed times are treated in the same way but consumption is expressed in time periods.
 \begin{align}
@@
 
+## Sample solution progress
+
+The analysis of the solution progress for all instances shows that:
+
+* Initial solutions are quite far from the linear relaxation.
+* Branching does not improve significantly the lower bound.
+
+\pause
+```{r out.width="150pt", echo=FALSE}
+knitr::include_graphics("./../../img/progress.png.png")
+```
+
+*(Progress diagram for instance id=I_7).*
+
 ## Multi-objective (1)
@@
 Comparison of two objectives: max availability or maintenance capacity.
-Different weights were used for the two objectives: $W_1, W_2 \in [0..1]$ with pace of 0.1
-```{r out.width="200pt", echo=FALSE}
-# img <- readPNG()
-#  grid.raster(img)
+Different weights were used for the two objectives: $W_1, W_2 \in [0..1]$ with pace of 0.1.
+\pause
+```{r out.width="180pt", echo=FALSE}
 knitr::include_graphics("./../../img/multiobjective.png")
-
 ```
 
-Pareto diagram for instance id=I_5.
+*(Pareto diagram for instance id=I_5).*
 
@@
 
-* Analysis for all instances: not many pareto optimal points available.
-* Between 2 and 4 for all instances
+* Analysis for all instances: not many pareto optimal points available (between two and four).
 * The objectives are quite aligned one with the other.
@@
 
-* **Try different objectives**: availability and maintenances.
+* **Try different objective formulations**: availability - maintenability - future state.
 * **Try more linent objectives**: relax the `max` in the objective.
@@
 # References
-## References
+## References / Questions?
 

### 2018-02-27 R/roadef_presentation/roadef_presentation.Rmd
changes adding images and R diagrams for OptiPlan proposal. Also, the proposal in txt

@@
 Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
-Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.
 * Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
-Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
-The number of resources in use for each period $d_t$ is also known.
+* Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
+* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H$ units.
 

### 2018-02-27 R/diagrams/architecture.R

@@ -1,0 +1,55 @@
+library(visNetwork)
+library(tidyverse)
+library(data.table)
+library(magrittr)
+
+img_path <- "/home/pchtsp/Documents/projects/OPTIMA"
+# img_path <- "./.."
+names_nodes = c('pc', 'server', 'db', 'model')
+images_name <- c('', 'server.png', 'database.png', 'model.png') %>% sprintf('%s/img/%s',img_path, .)
+
+nodes <- data.table(
+    label = names_nodes
+    ) %>% 
+    mutate(shape = "icon", id = row_number()
+           # image= images_name
+           ,icon.code = c('f108', 'f233', 'f1c0', 'f013')
+           ,icon.color= "black"
+           )
+    
+edges_data <- data.table(
+    from= c('pc', 'server', 'server', 'server', 'model'),
+    to= c('server', 'db', 'model', 'pc', 'server')
+)
+
+    # c('postgreSQL', 'model', 'Excel file', 'Excel template', 'csv / json', 'html')
+
+
+# data %<>% bind_rows()
+# data %<>% extract2(1)
+
+# nodes <- 
+#     data %>% 
+#     full_join(., select(., from), by=c("to"= 'from')) %>% 
+#     select(-from) %>% 
+#     distinct %>% 
+#     mutate(id= row_number(),
+#            # shadow = TRUE
+#     ) %>% 
+#     rename(label=to)
+
+edges <- 
+    edges_data %>% 
+    inner_join(nodes, by=c('from'='label')) %>% 
+    inner_join(nodes, by=c('to'='label')) %>% 
+    select(from= id.x, to=id.y)
+
+# head(nodes)
+
+# edges <- data
+# 
+visNetwork(nodes, edges, height = "800px", width = "800px") %>% 
+    visNodes(shapeProperties = list(useBorderWithImage = TRUE))
+    
+# visNetwork(nodes, edges, width = "100%") %>% visHierarchicalLayout
+

### 2018-02-27 R/diagrams/functionalities.R

@@ -1,0 +1,78 @@
+library(visNetwork)
+library(tidyverse)
+library(data.table)
+library(magrittr)
+
+data = 
+    list(
+        data.table(
+            from = 'Gérer cas'
+            ,to = c('Créer cas', 'Copier cas', 'Eliminer cas', 'Publier cas')
+            ,color="#abd9e9"
+            ,value=1
+        ),
+        data.table(
+            from = 'Rapports'
+            ,to = c('Visualisation', 'Comparaison des cas', 'autres rapports')
+            ,color="#abd9e9"
+            ,value=1
+        ),
+        data.table(
+            from = "Données d'entrée"
+            ,to = c('Editer données', 'Ajouter historiques')
+            ,color="#abd9e9"
+            ,value=1
+        ),
+        data.table(
+            from = "Planification"
+            ,to = c('Editer solution', 'Vérificateur automatique')
+            ,color="#abd9e9"
+            ,value=1
+        ),
+        data.table(
+            from = "Optimisateur"
+            ,to = c('Configurer résolution', 'Génerer solution')
+            ,color="#abd9e9"
+            ,value=1
+        ),
+        data.table(
+            from = "Intégration"
+            ,to = c('Importer Excel', 'Exporter fichiers')
+            ,color="#abd9e9"
+            ,value=1
+        ),
+        data.table(
+            from = 'Application'
+            ,to = c('Rapports', 'Gérer cas', "Données d'entrée", 
+                    'Planification', 'Optimisateur', 'Intégration')
+            ,shape = "circle"
+            ,value=100
+        )
+    ) 
+
+data %<>% bind_rows()
+# data %<>% extract2(1)
+
+nodes <- 
+    data %>% 
+    full_join(., select(., from), by=c("to"= 'from')) %>% 
+    select(-from) %>% 
+    distinct %>% 
+    mutate(id= row_number(),
+           # shadow = TRUE
+           ) %>% 
+    rename(label=to)
+
+edges <- 
+    data %>% 
+    inner_join(nodes, by=c('from'='label')) %>% 
+    inner_join(nodes, by=c('to'='label')) %>% 
+    select(from= id.x, to=id.y)
+
+# head(nodes)
+
+# edges <- data
+# 
+visNetwork(nodes, edges, height = "800px", width = "800px") 
+# visNetwork(nodes, edges, width = "100%") %>% visHierarchicalLayout
+

### 2018-03-01 R/OptiPlan_presentation/proposal.Rmd
changed name of OptiPlan document, deleted big gantt in html

# Optiplan proposal

## Introduction

In this document we present the problem as the team understands it. The information used was the previous knowledge the team and the OptiPlan meeting in Paris.

Following, a technical proposal is presented that includes the functionalities a possible complete application would include based on the exchanges made during the OptiPlan day in Paris and from previous experiences the team has in building planning applications for industry.

Furthermore, the team's credentials are presented, mentioning the different projects and research that is relevant to the present project.

Finally, the technologies and tools to be used are presented, describing the use and justifying the choice at each point.

## Summary

The project presented is a web-based modern interactive application. It uses a mathematical model to generate new solutions, permits different users to access it with different access rights and provides powerful visualization to show, compare, and analyze plannings. It integrates with existing data sources (Excel file), new proposed ones (Excel template) and possible existing ones (other internal systems) as well as simple output (csv, json). Finally, it permits the use of cases to store information, making it easy to test, experiment and compare different scenarios.

We believe this team is uniquely well-positioned to tackle this project because of the following reasons:

* The team has more than 40 years of experience in the Operations Research field, which is the theory behind this problem.
* The team has years of experience in building planning and optimization software applications for a wide range of big industrial companies.
* The team has vast experience in both scientific articles and commercial projects related to planning and maintenance.
* The team has unique knowledge of the specific problem at hand, because a PhD thesis with this subject was started in October 2017 in collaboration with the DGA. Work was presented in ROADEF 2018 and is going to be presented in MOSIM 2018.

## Problem

<!-- TODO: translate into non-mathematical -->

The problem is considered a Flight and Maintenance Planning problem. Work has been already done by \citet{Kozanidis2008}, \citet{Cho2011}, \citet{Verhoeff2015} and \citet{Chastellux2016}, among others.

The problem consists in assigning aircraft to predefined missions and scheduling periodic preventive maintenances for these same aircraft.

Missions can only be understood as a budget of hours per month or per year. In this document we present the most general formulation possible, where missions are assigned to aircraft for one or more months.

### Missions

There is a fixed set of $j \in \mathcal{J}$ missions to be accomplished over an horizon of time divided into $t \in \mathcal{T}$ discrete periods. For their execution, these missions require the assignment of a specific number of aircraft $R_j$ each period of time the mission is active. The start and end periods for each mission are known and a mission is considered active between its start and end period.

During each period, missions consume an amount of time equal to $H_j$ hours from each of its assigned aircraft.

The assignment of an aircraft to a mission is not decided for the whole duration of the mission. After a minimum amount of time $MT_j$, an aircraft can be freed and exchanged for another one, even if the mission it is assigned to has not finished. 
The total number of aircraft being used at any given time in a specific mission should always be equal to $R_j$.

Each mission requires one and only one type of aircraft which, in addition, should comply with additional mission requirements.

### Aircraft

There is a set $i \in \mathcal{I}$ of available aircraft that are assigned to missions in order to accomplish them. Each aircraft can only be assigned to a single mission in any given period. These aircraft suffer from wear and tear and require regular maintenance operations during their lifetime. The need for maintenance is calculated based on two indicators.

The first one is called "remaining elapsed time" (or $ret_{it}$). It expresses the amount of time (measured in time periods) after which the aircraft cannot be used anymore and has to undergo a maintenance operation. Its value is calculated for each aircraft $i$ and each time period $t$. In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the aircraft $i$ can be used before needing a maintenance operation at any given period $t$.

Additionally, after an absolute amount of time and/or usage ($aet_i$ or $aut_i$), the aircraft becomes obsolete. There is no way to reverse this process.

At any given period, including at the start of the planning horizon, each aircraft has a specific status given by remaining usage time and remaining elapsed time.

### Maintenances

Maintenances operations are the process by which aircraft that have reached a limit in some indicator can return to a state where they can continue to be used in missions.

Each maintenance operation has a fix duration of $M$ periods.

After a maintenance operation, an aircraft restores its remaining elapsed time and remaining usage time to their max values $E$ (~60 months) and $H$ (~1200 hours) respectively.

There are also special maintenances that improve the aircraft capabilities (their ability to be assigned to new missions). These maintenances will be decided by the user and fixed for the application so they can be taken into account when assigning the rest of maintenances.

### Storages

Following the rule of remaining elapsed time, even if an aircraft is not being used, it still needs to have a maintenance after a given amount of time has passed. In order to avoid this problem, the aircraft can be put into a storage state.

A aircraft in this state has to be kept for a minimum time of $sm$ periods. While in this state it cannot receive maintenance or be assigned to any mission.

Every aircraft has the capacity to be stored and this capacity is measured in a number of periods $sc$. In order for an aircraft to restore its remaining storage capacity, it needs to receive a specific maintenance. Similar to the remaining elapsed time, the aircraft recovers its storage capacity up to a certain level $S$.

### Possible states

As a summary, the following are the possible logical states that an aircraft can be in:

* Assigned to a mission.
* Receiving maintenance.
* Under storage.
* Available.

### Time

In planning missions and maintenances, it is important to take into account the initial and end state of each aircraft. This initial state can be a maintenance or an assigned mission. If an aircraft is already under maintenance, it needs to continue in this state for its remaining maintenance time. Tasks' assignments should be taken into account in a similar manner.

For planned maintenances whose starting time is close to the beginning of the planning horizon, they should also be considered fixed.

<!-- TODO: diagram -->

The remaining used and elapsed times are assigned to each aircraft at the beginning of the planning horizon.

Finally, the state of each aircraft at the end of the planning horizon, its remaining (elapsed, used) time, needs to be defined and taken into account.

### Objectives

The following objectives are considered, although more could be potentially added.

Given that the creation of new missions and the duration of maintenance are considered stochastic in real-life, one basic goal is to maximize the robustness of the planning by having the greatest amount of available aircraft at every period of the planning horizon.

Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of aircraft under maintenance over the planning horizon. Maintenances should also be controlled so it does not pass the maximum capacity.

### Proposal

A summary of the application's functionalities is shown in the following image:

![](../../img/functionalities.png)

### Usage

<!-- TODO: flowcharts! -->

### Cases and experimentation

Given the strategic nature of the application, it will be case-centric. This means that it will support saving all the relevant data (input and output) in a case. This way, different experiments could be done by copying and modifying an existing case. This is similar as working with multiple spreadsheets, each one with its own name, while sharing the same logic in one place.

The following figure shows an hypothetical set of cases already saved:

The following figure represents the possible interactions between cases. Case 1 was published for the month of March. During the month, it received some modifications, resulting in cases 2 and 3. Finally, Case 3 was (re)published for March.
Later, Case 4 was created and modified before publication to Case 5. Case 5 was published for April.

![](../../img/cases.png)

The available functionality will be the following:

1. Create a new empty case.
2. Copy an existing case.
3. Delete a case.
4. Compare two cases.
5. Publish a case.
6. Export a case.

#### Contents of a case

A case consists on all the information needed to construct a planning and one possible planning built from this information. This unit of storage permits to easily integrate with current data structures and run, compare different experiments.

More explicitly:

* Set of aircraft with their characteristics and historic information on them until a certain date.
* Set of missions or flight hours to do in the planning period.
* Assignments of aircraft to operations during the planning horizon.
* Future maintenance schedules.
* Configuration such as maintenance duration, flight hours after a maintenance, etc.

#### Comparison of cases

This functionality will potentially work in several levels depending on the final needs of the client.
One option would be to visualize the differences in input data for two cases.
Another option would be to visualize the differences in the resulting planning for the two cases.
Finally, it could be interesting to compare specific KPI for two or more cases, in order to choose the best scenario.

#### Published case

A situation that is common in the creation of plannings is the fact that several different cases may exist but it is still important to identify which one is the 'active' one. In order to deal with this, a simple publication system will be included in the application, where the 'planner' (see *Users* section) can identify the case that should be shared with others.

### Graphic User Interface (GUI)

All interaction from the user with the application will be done via an HTML application.

Besides showing data in interactive tables, specific alternatives can be incorporated to make the usage easier.

Here are some examples of available functionality:

Here, an interactive gantt is used to present and potentially modify a maintenance plan:

<!-- TODO: gantt example -->

Here, an interactive graph is used to present the usage of aircraft:

<!-- TODO: rut example -->

Here, an interactive table is used to present the usage of aircraft:

<!-- TODO: DT example -->

### Data integration

The application will include its own data storage through a database. All information used by the application will be stored here. In addition to this system, the application will permit several ways to import and export data from it.

#### Importing data

A new case can be created from the following options:

* The Excel file that is currently being used.
* An new yet-to-be-created Excel template file.

In addition to this, the historic data could be potential read from another system, if available.

#### Exporting data

The data in a case will be able to be exported in the following ways:

* In the *current Excel* format (or one that is similar).
* Into the new Excel template that is will be created.
* csv and/ or json files if needed.

### Architecture and deployment

The application will be deployed via a client-server architecture. This way, it would be easy to share the application given that it will only require having a modern web browser installed (Google Chrome).

The server should be installed somewhere inside the client's infrastructure.

The default requirements for the server are:

* Ubuntu Server 64 bits.
* Cores >= 4
* RAM >= 8GB
* Hard drive >= 500 GB

Depending on the future integration, additional data flows will be included in the design. The only known data flow is between the existing Excel file and the application.

The following figure shows a summary of the architecture to be implemented and the data flows between components.

Users will interact via web browsers with the server. In order to load new data they can upload a new Excel file with the relevant data or edit an existing case via the application.

The server will be in charge of solving and generating new planning solutions and storing all relevant information in its database.

Users will be able to recover visually the results. Also, they will be able to export the results into csv, Excel and json formats depending on their needs.

![Proposed architecture with data flows between components](../../img/architecture.png)

This client-server architecture provides several advantages:

* Powerful graphics can be incorporated via open source javascript libraries.
* It is easier to maintain as a clear separation of concerns is done between GUI, logic and model.
* No installation of software besides a modern web browser is required.
* Multiple users can be easily incorporated.
* Integration with other systems can be done with minimal changes to the existing IT.
* Powerful resources can be allocated where needed (server) instead of the users pcs.

### Users

Two possible roles are identified for the users of this application.

The first one will be called "planner". This user is in charge of creating (manually or automatically) different plannings.
The second one will be called "viewer". This user will be able to visualize the results of the planning created and stored inside the application. 
Depending on the needs of the client, several "viewer" profiles could be created in order to share only parts of the stored information with each profile.

## Team

<!-- TODO: names, CV and résumés-->

* Operations research experts.
* Familiar with maintenance planning.
* With experience building applications for industrial clients.
* Direct experience with the problem at hand.

### Alain Haït

### Olga Battaïa

### Franco Peschiera

Franco has been working in the Operations Research industry for almost 7 years. During which he has been the Chief Technical Officer for a Spanish startup that builds software applications with mathematical optimization or simulation models inside for several big Spanish and international companies. Examples of clients have been: Repsol and CLH (oil logistics, marketing, human resources), General Motors, L.I. Lilly and Air Liquide (production planning), Alstom and ASM (logistics and transportation) among many others.

Franco started a PhD at ISAE-SUPAERO university in October 2017 on the topic of military aircraft maintenance planning.

## Technologies and techniques

* **Mathematical Modeling** to solve the problem.
* **Python** for modeling, heuristics and data integration.
* **R** for statistics, possible forecasting and data visualization.
* **PostgreSQL** for storing data.
* **HTML5** for reports, visualization.

### Mathematical modeling

Mathematical modeling is a technique that is part of the branch of Operations Research in mathematics. It permits the construction of a set of equations to represent a specific well-defined situation and the use of advanced state of the art programs to search for and find the best solution that takes respects these equations. These programs (solvers) are available both in commercial and open source (free).

Advantages of using this technique compared to other alternatives are that it can prove the optimality of a solution, it can incorporate multiple objectives easily and it is possible to easily fix certain decisions that are already taken, making it possible to have interactive user-model use cases. With little modifications, it can be reused to take into account different situations.

Other functionality also includes having time limits and other more complex stop conditions.

### Python

python permits at the same time rapid prototyping as well as good code syntax. It offer a extremely wide offer of libraries and, in particular, it includes the best APIs to solver and non commercial solvers for Mixed Integer Programming problems.

It also excels at web developing with libraries such as flask and jinja templates.

### R

R is the state of the art in statistics and forecasting modelling. It includes several libraries for data manipulation, visualization and, as a programming language, good integrations with other languages and software.

Finally, it offers specially good integration with html-based applications via the [Shiny](https://shiny.rstudio.com/) technology.

### PostgreSQL

PostgreSQL is one of the most well-known open-source database alternatives. It includes many functionalities that are not present in other database engines (namely MySQL) such as:

* NoSQL storing (for cases, for example).
* window-functions integrated with R.

### HTML5

HTML 5 is the most recent change to the HTML that incorporates graphical functionality. This standard has permited a multitude of open sourced libraries that offer very powerful interactive graphics that are easily incorporated into webpages.

Examples of these libraries are: [leaflet](http://leafletjs.com/), [visjs](visjs.org), [d3js](https://d3js.org/) among many others.

### 2018-03-01 R/OptiPlan_presentation/proposal.Rmd
New version of Optiplan document. Added resources for document: img, diagrams in R

@@
-# Optiplan proposal
+---
+title: "Optiplan proposal"
+date: "March 11, 2018"
+output: 
+  html_document:
+    css: ./../../css/baobab_docs.css
+    toc: true
+  beamer_presentation:
+    theme: "Berlin"
+    toc: true
+    slide_level: 3
+    incremental: yes
+    fig_height: 4
+    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
+    includes:
+        in_header: header.tex
+  slidy_presentation:
+    css: ./../../css/baobab_MH.css
+    incremental: yes
+    footer: "ISAE-SUPAERO, Université de Toulouse, France"
+    slide_level: 4
+    toc: true
+nocite: '@*'
+bibliography: MFMP.bib
+---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE)
+setwd("../")
+source('scripts/params.R')
+source('scripts/import_results.R')
+source('diagrams/cases.R')
+source("diagrams/architecture.R")
+source("diagrams/functionalities.R")
+source('diagrams/exampleDT.R')
+exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
+gantt <- print_solution(exp_directory, max_resources = 5, width="100%", height="200px")
+graphs_rem <- graph_remaining(exp_directory)
+table_rem <- table_remaining(exp_directory)
+html_output <- knitr::opts_knit$get("rmarkdown.pandoc.to") %in% c("slidy", 'html')
+```
 
@@
 
-<!-- TODO: translate into non-mathematical -->
-
-The problem is considered a Flight and Maintenance Planning problem.
-Work has been already done by \citet{Kozanidis2008}, \citet{Cho2011}, \citet{Verhoeff2015} and \citet{Chastellux2016}, among others.
+The problem is considered a Flight and Maintenance Planning problem. Work has been already done by @Kozanidis2008, @Cho2011, @Verhoeff2015 and @Chastellux2016, among others.
 
@@
 
-### Proposal
+## Proposal
 
@@
 
-![](../../img/functionalities.png)
+```{r, echo=FALSE, out.width='70%'}
+if (html_output){
+    graph_functionalities()
+} else {
+    knitr::include_graphics('../../img/functionalities.png')
+}
+```
 
@@
 
-![](../../img/cases.png)
+```{r, echo=FALSE, out.width='70%'}
+if (html_output){
+    graph_cases()
+} else {
+    knitr::include_graphics('../../img/cases.png')
+}
+```
 
@@
 
-Here, an interactive gantt is used to present and potentially modify a maintenance plan:
+Here, an interactive gantt is used to present and potentially modify a maintenance plan.
+It is possible to edit the maintenances and assignments tasks (change the size or move it) by double clicking on them and dragging.
+Also, it is possible to add new assignments by double clicking.
 
-<!-- TODO: gantt example -->
+```{r, echo=FALSE, out.width='70%'}
 
-Here, an interactive graph is used to present the usage of aircraft:
+if (html_output){
+    gantt
+} else {
+    knitr::include_graphics('./../../img/calendar.png')
+}
+```
 
-<!-- TODO: rut example -->
+Here, an interactive graph is used to present the evolution of the status of a single aircraft over time:
 
-Here, an interactive table is used to present the usage of aircraft:
+```{r, echo=FALSE, out.width='90%'}
 
-<!-- TODO: DT example -->
+if (html_output){
+    graphs_rem
+} else {
+    knitr::include_graphics('./../../img/remaining.png')
+}
+```
+
+Here, an interactive table is used to present some data:
+
+```{r, echo=FALSE, out.width='90%'}
+
+if (html_output){
+    table_rem
+}
+```
 
@@
 
-![Proposed architecture with data flows between components](../../img/architecture.png)
+```{r, echo=FALSE, out.width='70%'}
+
+if (html_output){
+    graph_architecture()
+} else {
+    knitr::include_graphics('../../img/architecture.png')
+}
+```
+
+<!-- ![Proposed architecture with data flows between components]() -->
 
@@
 Examples of these libraries are: [leaflet](http://leafletjs.com/), [visjs](visjs.org), [d3js](https://d3js.org/) among many others.
+
+### References

### 2018-03-01 R/roadef_presentation/roadef_presentation.Rmd
New version of Optiplan document. Added resources for document: img, diagrams in R

@@
 source('scripts/import_results.R')
-# library(png)
-# library(grid)
 exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')

### 2018-03-01 R/diagrams/architecture.R

@@ -3,42 +3,46 @@ library(tidyverse)
 library(data.table)
 library(magrittr)
 
-# img_path <- "/home/pchtsp/Documents/projects/OPTIMA"
-# img_path <- "./.."
-names_nodes = c('planner', 'viewer', 'serveur', 'base de données', 'modèle')
-# images_name <- c('', 'serveur.png', 'database.png', 'model.png') %>% sprintf('%s/img/%s',img_path, .)
-
-nodes <- data.table(
-    label = names_nodes
+graph_architecture <- function(){
+    # img_path <- "/home/pchtsp/Documents/projects/OPTIMA"
+    # img_path <- "./.."
+    names_nodes = c('planner', 'viewer', 'serveur', 'base de données', 'modèle')
+    # images_name <- c('', 'serveur.png', 'database.png', 'model.png') %>% sprintf('%s/img/%s',img_path, .)
+    
+    nodes <- data.table(
+        label = names_nodes
     ) %>% 
-    mutate(shape = "icon", id = row_number()
-           # image= images_name
-           ,icon.code = c('f007', 'f007', 'f233', 'f1c0', 'f013')
-           ,icon.color= c('orange', 'blue', rep("black", 3))
-           ,icon.size = 80
-           ,font.size = 25
-           )
+        mutate(shape = "icon", id = row_number()
+               # image= images_name
+               ,icon.code = c('f007', 'f007', 'f233', 'f1c0', 'f013')
+               ,icon.color= c('orange', 'blue', rep("black", 3))
+               ,icon.size = 80
+               ,font.size = 25
+        )
     
-edges_data <- data.table(
-    from= c('planner', 'serveur', 'serveur', 'serveur', 'serveur', 'serveur', 'serveur')
-    ,to= c( 'serveur', 'base de données', 'modèle', 'planner', 'viewer', 'planner', 'serveur')
-    ,arrows= c(rep('to', 1), '', '', rep('to', 3), '')
-    ,label= c('Excel', 'postgreSQL', 'python', 'R+html5', 'R+html5', 'csv+json', 'python')
-)
-
-edges <- 
-    edges_data %>% 
-    inner_join(nodes, by=c('from'='label')) %>% 
-    inner_join(nodes, by=c('to'='label')) %>% 
-    mutate(from= id.x, to=id.y) %>% 
-    select(-ends_with(".x")) %>% 
-    select(-ends_with(".y")) %>% 
-    mutate(color='lightblue', 
[…]

### 2018-03-01 R/diagrams/cases.R

@@ -3,40 +3,43 @@ library(tidyverse)
 library(data.table)
 library(magrittr)
 
-base_nodes <- paste('Case', 1:5)
-names_nodes = c(base_nodes, 'Cases', 'Publié mars', 'Publié avril')
-
-nodes <- data.table(
-    label = names_nodes
-) %>% 
-    mutate(shape = 'icon'
-           ,id = row_number()
-           # image= images_name
-           ,icon.code = if_else(label=='Cases', 'f187', 'f07b')
-           ,icon.color= c(rep("black", 6), rep('red', 2))
-           ,icon.size = if_else(label=='Cases', 120, 80)
-           ,font.size = if_else(label=='Cases', 30, 20)
-    )
-
-edges_data <- data.table(
-    from = c('Case 1', 'Case 3', 'Case 5', paste('Case', 1:2), "Case 4")
-    ,to = c('Publié mars', 'Publié mars', 'Publié avril', paste('Case', 2:3), "Case 5")
-    ,arrows= c('', '', '', rep('to', 3))
-) %>% 
-    bind_rows(CJ(to=base_nodes, from=c('Cases')))
-
-
-edges <- 
-    edges_data %>% 
-    inner_join(nodes, by=c('from'='label')) %>% 
-    inner_join(nodes, by=c('to'='label')) %>% 
-    mutate(from= id.x, to=id.y) %>% 
-    mutate(color='lightblue', 
-           font.size = 10,
-           value = 0.01,
-           length = 300)
-
-visNetwork(nodes, edges, height = "800px", width = "800px") %>% 
-    visNodes(shapeProperties = list(useBorderWithImage = TRUE))
-
-
+graph_cases <- function(){
+    base_nodes <- paste('Cas', 1:5)
+    names_nodes = c(base_nodes, 'Cas dépôt', 'Publié mars', 'Publié avril')
+    
+    nodes <- data.table(
+        label = names_nodes
+    ) %>% 
+        mutate(shape = 'icon'
+               ,id = row_number()
+               # image= images_name
+               ,icon.code = if_else(label=='Cas dépôt', 'f187', 'f07b')
+               ,icon.color= c(rep("black", 6), rep('red', 2))
+               ,icon.size = if_else(label=='Cas dépôt', 120, 80)
+               ,font.size = if_else(label=='Cas dépôt', 30, 20)
+        )
+    
+    edges_data <- data.table(
[…]

### 2018-03-01 R/diagrams/exampleDT.R

@@ -1,0 +1,59 @@
+library(DT)
+library(tidyverse)
+library(data.table)
+library(magrittr)
+library(ggplot2)
+library(plotly)
+
+treat_data <- function(exp_directory, aircraft_to_graph){
+    solution_path = exp_directory %>% paste0('data_out.json')
+    
+    solution <- read_json(solution_path)
+    
+    evolution_rut <- 
+        solution %>% 
+        extract2("aux") %>% 
+        extract2("rut") %>% 
+        bind_rows(.id="aircraft") %>% 
+        gather(key='Mois', value = 'Temps', -aircraft) %>% 
+        filter(aircraft == 'A31') %>% 
+        mutate(type='heures de vol')
+    
+    evolution_ret <- 
+        solution %>% 
+        extract2("aux") %>% 
+        extract2("ret") %>% 
+        bind_rows(.id="aircraft") %>% 
+        gather(key='Mois', value = 'Temps', -aircraft) %>% 
+        filter(aircraft == aircraft_to_graph) %>% 
+        mutate(type='disponibilité (mois)')
+    
+    data <- bind_rows(evolution_ret, evolution_rut) %>% 
+        mutate(Mois= Mois %>% paste0("-01") %>% as.POSIXct())
+    
+}
+
+table_remaining <- function(exp_directory, aircraft_to_graph='A31'){
+    data <- treat_data(exp_directory, aircraft_to_graph)
+    
+    data %<>% mutate(Mois = Mois %>% format("%Y-%m"))
+    
+    DT::datatable(data, editable = TRUE, filter='none')
+    
+    }
+
+    
+graph_remaining <- function(exp_directory, aircraft_to_graph='A31'){
+    data <- treat_data(exp_directory, aircraft_to_graph)
+    
+    p <- ggplot(data, aes(x= Mois, y= Temps, group='aircraft')) + 
+        geom_line() + 
+        facet_grid(type~ ., scales = "free_y") +
+        theme_minimal() + 
+        scale_x_datetime()
+    
+    ggplotly(p)
+    
+}
+
+

### 2018-03-01 R/diagrams/functionalities.R

@@ -3,76 +3,80 @@ library(tidyverse)
 library(data.table)
 library(magrittr)
 
-data = 
-    list(
-        data.table(
-            from = 'Gérer cas'
-            ,to = c('Créer cas', 'Copier cas', 'Eliminer cas', 'Publier cas')
-            ,color="#abd9e9"
-            ,value=1
-        ),
-        data.table(
-            from = 'Rapports'
-            ,to = c('Visualisation', 'Comparaison des cas', 'autres rapports')
-            ,color="#abd9e9"
-            ,value=1
-        ),
-        data.table(
-            from = "Données d'entrée"
-            ,to = c('Editer données', 'Ajouter historiques')
-            ,color="#abd9e9"
-            ,value=1
-        ),
-        data.table(
-            from = "Planification"
-            ,to = c('Editer solution', 'Vérificateur automatique')
-            ,color="#abd9e9"
-            ,value=1
-        ),
-        data.table(
-            from = "Optimisateur"
-            ,to = c('Configurer résolution', 'Génerer solution')
-            ,color="#abd9e9"
-            ,value=1
-        ),
-        data.table(
-            from = "Intégration"
-            ,to = c('Importer Excel', 'Exporter fichiers')
-            ,color="#abd9e9"
-            ,value=1
-        ),
-        data.table(
-            from = 'Application'
-            ,to = c('Rapports', 'Gérer cas', "Données d'entrée", 
-                    'Planification', 'Optimisateur', 'Intégration')
-            ,shape = "circle"
-            ,value=100
-        )
-    ) 
-
-data %<>% bind_rows()
-# data %<>% extract2(1)
-
-nodes <- 
-    data %>% 
-    full_join(., select(., from), by=c("to"= 'from')) %>% 
-    select(-from) %>% 
-    distinct %>% 
-    mutate(id= row_number(),
-           # shadow = TRUE
-           ) %>% 
-    rename(label=to)
-
-edges <- 
-    data %>% 
-    inner_join(nodes, by=c('from'='label')) %>% 
-    inner_join(nodes, by=c('to'='label')) %>% 
-    select(from= id.x, to=id.y)
-
-# head(nodes)
-
-# edges <- data
-# 
-visNetwork(nodes, edges, height = "800px", width = "800px") 
-# visNetwork(nodes, edges, width = "100%") %>% visHierarchicalLayout
-
+graph_functionalities <- function(){
+    data = 
+        list(
+            data.table(
[…]

### 2018-03-01 R/scripts/import_results.R

@@ -65,7 +65,7 @@ print_tasks <- function(exp_directory){
     timevis(tasks, groups= groups, options= config, width="100%")
 }
 
-print_solution <- function(exp_directory, max_resources=NULL){
+print_solution <- function(exp_directory, max_resources=NULL, ...){
     solution_path = exp_directory %>% paste0('data_out.json')
     input_path = exp_directory %>% paste0('data_in.json')
     
@@ -105,7 +105,7 @@ print_solution <- function(exp_directory, max_resources=NULL){
         select(id, start= month, end= post_month, content, group= UNIT, style)
     
     if (max_resources %>% is.null %>% not){
-        resources <- states %>% distinct(group) %>% slice(1:max_resources)
+        resources <- states %>% distinct(group) %>% sample_n(max_resources)
         states <- states %>% inner_join(resources)
     }
     
@@ -118,10 +118,11 @@ print_solution <- function(exp_directory, max_resources=NULL){
         align = "center",
         orientation = "top",
         snap = NULL,
-        margin = 0
+        margin = 0,
+        zoomable= FALSE
     )
     
-    timevis(states, groups= groups, options= config, width="100%")
+    timevis(states, groups= groups, options= config, ...)
 }
 
 if (FALSE){

### 2018-03-05 R/OptiPlan_presentation/proposal.Rmd
updated proposal with planning. Added french version. slighlty changed css

@@
 
+<!-- TODO: check if all points from the rules are covered -->
+
 ## Introduction
@@
 
-The problem consists in assigning aircraft to predefined missions and scheduling periodic preventive maintenances for these same aircraft.
+The problem consists in assigning aircraft to predefined missions and scheduling periodic preventive maintenance for these same aircraft.
 
@@
 
-Maintenances operations are the process by which aircraft that have reached a limit in some indicator can return to a state where they can continue to be used in missions.
+Maintenance operations are the process by which aircraft that have reached a limit in some indicator can return to a state where they can continue to be used in missions.
 
@@
 
-There are also special maintenances that improve the aircraft capabilities (their ability to be assigned to new missions).
-These maintenances will be decided by the user and fixed for the application so they can be taken into account when assigning the rest of maintenances.
+There are also special maintenance operations that improve the aircraft capabilities (their ability to be assigned to new missions).
+These operations will be decided by the user and fixed for the application so they can be taken into account when assigning the rest of maintenance operations.
 
@@
 
-In planning missions and maintenances, it is important to take into account the initial and end state of each aircraft.
+In planning missions and maintenance operations, it is important to take into account the initial and end state of each aircraft.
 This initial state can be a maintenance or an assigned mission.
@@
 
-For planned maintenances whose starting time is close to the beginning of the planning horizon, they should also be considered fixed.
-
-<!-- TODO: diagram -->
+For planned maintenance operations whose starting time is close to the beginning of the planning horizon, they should also be considered fixed.
 
@@
 Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of aircraft under maintenance over the planning horizon.
-Maintenances should also be controlled so it does not pass the maximum capacity.
+Maintenance operations should also be controlled so it does not pass the maximum capacity.
 
@@
 
-### Usage
+    Graph showing functionalities for the application
 
-<!-- TODO: flowcharts! -->
+### Optimizer
+
+The model built to generate new solutions will comply with the requirements described below.
+
+#### RQ001: missions hours
+
+Each month, all missions need to be satisfied by assigning enough flight hours to each aircraft.
+
+#### RQ002: assigned aircraft
+
+Each month, all missions need to have a minimum number of aircraft assigned.
+
+#### RQ003: maintenance needs
+
+Each aircraft needs to always have enough flight hours to be able to do the required missions it is assigned to.
+In order to recover its flight hours (to around 1200 hours), a maintenance operation is needed.
+
+Similarly, after certain amount of time has passed, aircraft need a maintenance even if they did not reach this flight hour limit. This amount of time is around 5 years.
+
+#### RQ004: maintenance duration
+
+Maintenance operations take around 5 months and once an aircraft has started the maintenance operation, it needs to continue it. It cannot do any operation while in maintenance.
+
+#### RQ005: availability
+
+At each month, it is important to have the biggest possible number of available aircraft ready to engage in a new mission.
+
+#### RQ006: maintenance capacity
+
+The maintenance operations to be planned for all aircraft need to be done in the most smooth way possible, avoiding specific months were there are many aircraft under maintenance and other months were there are but a few aircraft under maintenance.
+
+In other words, it is important to have the lowest possible number of available aircraft under maintenance at each month.
+
+#### RQ007: end state
+
+At the end of the planning horizon, the aircraft need to be in the best state possible.
+This implies that, in average, they need to have a big budget of flight hours remaining and a big number of months before needing a maintenance.
+
+#### RQ008: storage
+
+The storage can be decided by the model in order to avoid the deterioration of the aircraft.
+The minimum period to store an aircraft is 6 months and while in this state it is unavailable both for maintenance and for missions.
+
+#### RQ009: fixing parts of the planning
+
+There are several reasons for the user to want to fix some part of a previous planning.
+For example, there are some maintenance operations that are already decided and need to take place in the future.
+There are some aircraft that are already in a mission and need to stay there for some months at least.
+Also, storage can be already decided or in process.
+
+Another reason for fixing maintenance operations could be for planned improvement operations to aircraft.
+These special type of maintenance operations are decided many months in advance for specific selected aircraft.
+
+These fixed assignments will be communicated to the planning generator in order for it to not change them.
+
+#### RQ010: configuration
+
+Available options will be, by default:
+
+* **Maximal time**: in case a solution is needed before proving optimality.
+* **Maximal GAP**: in order to apply a maximum allowed distance from the best obtained solution to the hypothetical best possible solution.
+This distance can be a proportional difference or an absolute difference.
+* **First and last month**: in order to define the horizon of planning to plan.
+* **Objective weights**: for the different objectives to optimize, to give preference to one or another objective.
+* **Maintenance parameters**: maintenance duration (in months), maximum flight hours, maximum elapsed time, and other maintenance-related parameters.
 
@@
 ```
+
+    Example showing interaction between cases and publication of cases.
 
@@
 
-Besides showing data in interactive tables, specific alternatives can be incorporated to make the usage easier.
+Besides showing data in interactive tables, specific alternatives can be incorporated to make the usage easier and more potent.
 
-Here are some examples of available functionality:
+Below are some examples of available functionality.
 
 Here, an interactive gantt is used to present and potentially modify a maintenance plan.
-It is possible to edit the maintenances and assignments tasks (change the size or move it) by double clicking on them and dragging.
+It is possible to edit the maintenance operations and assignments tasks (change the size or move it) by double clicking on them and dragging.
 Also, it is possible to add new assignments by double clicking.
@@
 
-Here, an interactive graph is used to present the evolution of the status of a single aircraft over time:
+    Interactive gantt example
+
+Here, an interactive graph is used to present the evolution of the status of a single aircraft over time. It is possible to filter, see the data in details and export as an image.
 
@@
 
-Here, an interactive table is used to present some data:
+    Interactive graph example
+
+Here, an interactive table is used to present some data. It is possible to edit the contents, filter, re-order data, etc.
 
@@
 ```
+
+    Interactive table example
 
@@
 
+This exporting capabilities will also permit the sharing of partial information to other actors in Excel or flat files.
+
 ### Architecture and deployment
@@
 
-The default requirements for the server are:
+The minimal hardware requirements for the server are:
 
@@
 * Hard drive >= 500 GB
+
+The minimal software requirements for the server are:
+
+* R + RStudio + Shiny
+* python >= 3.6
+* PostgreSQL
+
+Additional software could be needed depending on the final specifications. For example, in case the model is way to big for open source solvers, a commercial solver might be needed.
 
@@
     graph_architecture()
+
 } else {
@@
 
-<!-- ![Proposed architecture with data flows between components]() -->
+    Proposed architecture with data flows between components
 
@@
 * Integration with other systems can be done with minimal changes to the existing IT.
-* Powerful resources can be allocated where needed (server) instead of the users pcs.
+* Powerful resources can be allocated where needed (server) instead of the users PCs.
+
+In addition to this, there is not third-party software to buy.
+All components are open source.
+The only possible exception could be the mathematical solver, depending on the size of the problem and the needs of speed.
 
@@
 
-* Operations research experts.
-* Familiar with maintenance planning.
-* With experience building applications for industrial clients.
-* Direct experience with the problem at hand.
-
 ### Alain Haït
@@
 
+An industrial engineer, mathematician by profession and a programmer by heart.
+
 Franco has been working in the Operations Research industry for almost 7 years.
-During which he has been the Chief Technical Officer for a Spanish startup that builds software applications with mathematical optimization or simulation models inside for several big Spanish and international companies.
+During which he has been the owner and Chief Technical Officer for a Spanish startup (www.baobabsoluciones.es) building software applications with mathematical optimization or simulation models inside for several big Spanish and international companies.
 Examples of clients have been: Repsol and CLH (oil logistics, marketing, human resources), General Motors, L.I.
-Lilly and Air Liquide (production planning), Alstom and ASM (logistics and transportation) among many others.
+Lilly and Air Liquide (production planning), RATP, Alstom and ASM (logistics and transportation) among many others.
+
+These projects involved the following incomplete list of tasks: the specification of the client"s requirements, the construction of a mathematical or simulation model, the programming of a graphical user interface, deployment and integration with the IT systems, documentation and training for users.
 
 Franco started a PhD at ISAE-SUPAERO university in October 2017 on the topic of military aircraft maintenance planning.
+
+## Planning
+
+Here, the required tasks and their order are presented.
+The main risk on the present planning is the incomplete information about:
+
+* the specific problem to be solved.
+* the available IT infrastructure to work with.
+* availability of potential users to discuss example solutions and test and evaluate functionality.
+
+### Phase 1: requirements
+
+In this phase, several in-person and telephone meetings will be scheduled.
+
+The requisites for this phase are: **None**
+
+The deliverable of this phase will be:
+
+* a document with an agreed description with the minimal requirements for the prototype and 
+* a draft of possible additional requirements to be specified later on.
+
+The duration of this phase is estimated to be: **3 weeks**
+
+### Phase 2: first iteration
+
+In this phase, a first version of the prototype will be constructed and deployed.
+
+The requisites for this phase are:
+
+* the list with minimal requirements.
+* a deployment alternative to share the application with the users.
+
+The deliverable of this phase will be:
+
+* an prototype application that users can use and test.
+* example solutions for users to analyze.
+
+The duration of this phase is estimated to be: **5 weeks**
+
+### Phase 3: second iteration
+
+In this phase, a final version of the prototype will be constructed and deployed.
+
+The requisites for this phase are:
+
+* a list with changes to make to the first version based on the feedback from users.
+* a list with additional functionality that has high priority.
+
+The deliverable of this phase will be:
+
+* an application that users can test.
+* technical documentation.
+* User's guide.
+
+The duration of this phase is estimated to be: **5 weeks**
 
@@
 R is the state of the art in statistics and forecasting modelling.
-It includes several libraries for data manipulation, visualization and, as a programming language, good integrations with other languages and software.
+It includes several libraries for data manipulation, visualization and, as a programming language, good integration with other languages and software.
 
-Finally, it offers specially good integration with html-based applications via the [Shiny](https://shiny.rstudio.com/) technology.
+Finally, it offers specially good integration with html-based applications via the [Shiny](https://shiny.rstudio.com/gallery/) technology, which will be used in this project.
 
@@
 HTML 5 is the most recent change to the HTML that incorporates graphical functionality.
-This standard has permited a multitude of open sourced libraries that offer very powerful interactive graphics that are easily incorporated into webpages.
+This standard has allowed a multitude of open sourced libraries that offer very powerful interactive graphics that are easily incorporated into webpages.
 
@@
 
-### References
+## Disclaimers
+
+1. The proposal presented above is intended for a complete application delivered as a part of a complete project to DGA.
+1.
+The prototype to be delivered will not include the full functionality mentioned in this document but will show examples of the most important pieces of functionality following a meeting with the future users and decision makers.
+1. Integration with other systems is out of the scope of this proposal since more information is needed regarding the existing systems to be able to estimate workload.
+1.
+It is assumed that DGA will provide the hardware requirements together with the access rights and permissions to install all the components of the software.
+Alternatively, the software could be installed in the cloud upon agreement.
+1.
+Even though mathematical programming provides the capacity to prove the optimality of a solution, the application does not guarantee finding it en every case.
+The quality of the solutions and the time to obtain them will depend on many factors; the size of the instance being the main one.
+
+## References

### 2018-03-05 R/OptiPlan_presentation/proposal_fr.Rmd
updated proposal with planning. Added french version. slighlty changed css

---
title: "Optiplan proposal"
date: "March 11, 2018"
output: 
  html_document:
    css: ./../../css/baobab_docs.css
    toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 3
    incremental: yes
    fig_height: 4
    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
    includes:
        in_header: header.tex
  slidy_presentation:
    css: ./../../css/baobab_MH.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    slide_level: 4
    toc: true
nocite: '@*'
bibliography: MFMP.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("../")
source('scripts/params.R')
source('scripts/import_results.R')
source('diagrams/cases.R')
source("diagrams/architecture.R")
source("diagrams/functionalities.R")
source('diagrams/exampleDT.R')
exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
gantt <- print_solution(exp_directory, max_resources = 5, width="100%", height="200px")
graphs_rem <- graph_remaining(exp_directory)
table_rem <- table_remaining(exp_directory)
html_output <- knitr::opts_knit$get("rmarkdown.pandoc.to") %in% c("slidy", 'html')
```

<!-- TODO: check if all points from the rules are covered -->

## Introduction

Dans ce document, nous présentons d'abord le problème tel que l'équipe le comprend. Les sources d'information sont la connaissance préliminaire de l'équipe et la réunion OptiPlan qui a eu lieu à Paris.

Une proposition technique est présentée qui inclut les fonctionnalités qu'une éventuelle application complète inclurait sur la base des échanges effectués lors de la journée OptiPlan à Paris et des expériences précédentes de l'équipe dans la construction d'applications d'aide à la planification pour l'industrie.

Ensuite, les références de l'équipe sont présentées, mentionnant les différents projets et recherches pertinents liés au présent projet.

Une planification de temps est présenté pour le développement du logiciel prototype, avec les livrables par étape.

Enfin, les technologies et les outils logiciels à utiliser sont présentés, décrivant l'utilisation et justifiant le choix à chaque point.

## Résumé

Le projet présenté est une application web interactive. Il utilise un modèle mathématique pour générer de nouvelles solutions, permet à différents utilisateurs d'y accéder avec différents droits d'accès et offre une puissante visualisation pour afficher, comparer et analyser les planifications. Il s'intègre aux sources de données existantes (fichier Excel), aux nouvelles proposées (modèle Excel) et a la possibilité de s’intégrer aux sources existantes (autres systèmes internes) ainsi qu'à la sortie simple (csv, json). Enfin, il permet l'utilisation de cas pour stocker des informations, ce qui facilite les essaies, l'expérimentation et la comparaison de différents scénarios.

Nous pensons que cette équipe est particulièrement bien placée pour aborder ce projet pour les raisons suivantes:

* L'équipe somme plus de 40 ans d'expérience dans le domaine de la recherche opérationnelle, qui est la théorie de bas derrière ce problème.
* L'équipe possède des années d'expérience dans la conception d'applications logicielles de planification et d'optimisation pour un large ensemble de grandes entreprises industrielles.
* L'équipe possède une vaste expérience produisant articles scientifiques et projets commerciaux liés à la planification et à la maintenance.
* L'équipe a une connaissance unique du problème spécifique, car une thèse de doctorat sur ce sujet a été lancée en octobre 2017. Les travaux ont été présentés au ROADEF 2018 et seront présentés dans la conférence MOSIM 2018.

## Problème

Le problème est considéré comme un problème de planification de vol et de maintenance (Flight and Maintenance Planning problem). Travail a déjà été fait par @Kozanidis2008, @Cho2011, @Verhoeff2015, entre autres.

Le problème consiste à affecter des avions aux missions prédéfinies et à programmer des opérations de maintenance préventive périodique pour ces mêmes avions.

Les missions peuvent aussi être comprises comme un budget d'heures par mois ou par an. Dans ce document, nous présentons la formulation la plus générale possible, où les missions sont affectés aux avions pour un ou plusieurs mois.

### Missions

Il y a un ensemble fixe de $j \in \mathcal{J}$ missions à accomplir sur un horizon de temps divisé en $t \in \mathcal{T}$ périodes discrètes. Pour leur exécution, ces missions nécessitent l'attribution d'un nombre spécifique d'aéronefs $R_j$ chaque période de temps la mission est active. Les périodes de début et de fin de chaque mission sont connues et une mission est considérée comme active dans les mois entre sa période de début et sa fin.

Au cours de chaque période, les missions consomment une quantité de temps égale à $H_j$ heures de chacun de ses avions affectés.

L'affectation d'un avion à une mission n'est pas décidée pour toute la durée de la mission. Après un minimum de temps $MT_j$, un aéronef peut être libéré et échangé par un autre, même si la mission à laquelle il est affecté n'a pas terminée. Le nombre total d'avions affectés à un moment donné à chaque mission doit toujours être égal à $R_j$.

Chaque mission nécessite un et un seul type d'aéronef qui, de plus, doit satisfaire à des exigences de mission supplémentaires.

### Avion

Il y a un ensemble $i \in \mathcal{I}$ des avions disponibles à être affectés aux missions. Chaque avion ne peut être affecté qu'à une seule mission au cours d'une période donnée. Ces avions souffrent d'usure et nécessitent des opérations de maintenance régulières pendant leur durée de vie. Le besoin de maintenance est calculé sur la base de deux indicateurs.

<!-- TODO: traduire ret et rut. -->

Le premier est appelé "temps écoulé restant" (ou $ret_{it}$). Il exprime la durée (mesurée en périodes) après laquelle l'aéronef ne peut plus être utilisé et doit subir une opération de maintenance. Sa valeur est calculée pour chaque avion   $i$ et chaque période de temps $t$. De la même manière, "temps d'utilisation restant" (ou $rut_{it}$) est utilisé pour mesurer la quantité de temps que l'avion $i$ peut être utilisé avant d'avoir besoin d'une opération de maintenance à une période donnée $t$.

De plus, après une quantité absolue de temps et / ou d'utilisation ($aet_i$ ou $aut_i$), l'aéronef devient obsolète. Il n'y a aucun moyen d'inverser ce processus.

<!-- TODO: traduire ret et rut. -->
À toute période donnée, y compris au début de l'horizon de planification, chaque aéronef a un statut spécifique donné par le temps d'utilisation restant et le temps écoulé restant.

### Maintenances

<!-- TODO: traduire ret et rut. -->
Les opérations de maintenance sont le processus par lequel les aéronefs qui ont atteint une limite dans certains indicateurs peuvent revenir à un état où ils peuvent continuer à être utilisés dans les missions.

Chaque opération de maintenance a une durée fixe de $M$ périodes.

Après une opération de maintenance, un aéronef restaure le temps restant et le temps d'utilisation restant à leurs valeurs maximales $E$ (~ 60 mois) et $H$ (~ 1200 heures) respectivement.

Il existe également des opérations de maintenance spéciales qui améliorent les capacités de l'avion (ils améliorent leur capacité à être affecté à de nouvelles missions). Ces opérations seront décidées par l'utilisateur et fixées pour l'application afin qu'elles puissent être prises en compte lors de l'attribution du reste des opérations de maintenance.

### Stockages

Conformément à la règle du temps écoulé, même si un aéronef n'est pas utilisé, il doit encore subir un opération de maintenance après un certain temps. Afin d'éviter ce problème, l'aéronef peut être mis dans un état de stockage.

Un aéronef dans cet état doit être gardé pendant un temps minimum de $sm$ périodes. Dans cet état, il ne peut pas recevoir de maintenance ou être affecté à une mission.

Chaque avion a une certaine capacité d'être stocké et cette capacité est mesurée en un nombre de périodes $sc$. Pour qu'un aéronef restaure sa capacité de stockage restante, il doit recevoir une opération de maintenance spécifique où l'avion récupère sa capacité de stockage jusqu'à un certain niveau $S$.

### États possibles

En résumé, voici les états logiques possibles d'un aéronef:

* Assigné à une mission.
* Réception de maintenance
* Sous stockage
* Disponible.

### Temps

Lors de la planification des missions et des opérations de maintenance, il est important de prendre en compte l'état initial et final de chaque aéronef. Cet état initial peut être une maintenance ou une mission assignée. Si un aéronef est déjà en maintenance, il doit continuer dans cet état pour son temps de maintenance restant. Les missions affectés doivent être prises en compte de la même manière.

Pour les opérations de maintenance planifiées dont l'heure de début est proche du début de l'horizon de planification, elles doivent également être considérées comme fixes.

<!-- TODO: traduire ret et rut. -->
Les temps restants utilisés et écoulés sont assignés à chaque aéronef au début de l'horizon de planification.

<!-- TODO: traduire ret et rut. -->
Enfin, l'état de chaque aéronef à la fin de l'horizon de planification, son temps restant (écoulé, utilisé), doit être défini et pris en compte.

### Objectifs

Les objectifs suivants sont considérés, bien que d'autres pourraient potentiellement être ajoutés.

Étant donné que la possibilité de nouvelles missions et la durée de l'opération de maintenance sont considérées stochastiques dans la vie réelle, un objectif fondamental est de maximiser la robustesse de la planification en ayant la plus grande quantité d'avions disponibles à chaque période de l'horizon de planification.

Compte tenu de la capacité limitée de maintenance et de son coût, un autre objectif consiste à lisser le plus possible le nombre d'avions sous maintenance sur l'horizon de planification. Les opérations de maintenance doivent également être contrôlées afin de ne pas dépasser la capacité maximale.

### Proposition

Un résumé des fonctionnalités de l'application est montré dans l'image suivante:

```{r, echo=FALSE, out.width='70%'}
if (html_output){
    graph_functionalities()
} else {
    knitr::include_graphics('../../img/functionalities.png')
}
```

    Graphique montrant les fonctionnalités de l'application

### Optimiseur

Le modèle conçu pour générer de nouvelles solutions sera conforme aux exigences décrites ci-dessous.

#### RQ001: heures de missions

Chaque mois, toutes les missions doivent être satisfaites en affectant suffisamment d'heures de vol à chaque avion.

#### RQ002: aéronefs assignés

Chaque mois, toutes les missions doivent avoir un nombre minimum d'avions affectés.

#### RQ003: besoins de maintenance

Chaque avion doit toujours avoir suffisamment d'heures de vol pour pouvoir effectuer les missions requises. Afin de récupérer des heures de vol (environ 1200 heures), une opération de maintenance est nécessaire.

De même, après un certain temps, les aéronefs ont besoin d'un entretien même s'ils n'ont pas atteint cette limite d'heures de vol. Cette durée est d'environ 5 ans.

#### RQ004: durée de maintenance

Les opérations de maintenance prennent environ 5 mois et une fois qu'un avion a commencé l'opération de maintenance, il doit la continuer. Il ne peut pas être affectés aux missions pendant la durée de la opération de maintenance.

#### RQ005: disponibilité

Chaque mois, il est important d'avoir le plus grand nombre possible d'avions disponibles prêts à s'engager dans une nouvelle mission.

#### RQ006: capacité de maintenance

Les opérations de maintenance à planifier pour tous les avions doivent être effectuées de la manière la plus lissé possible, en évitant avoir des mois spécifiques avec un grand quantité des maintenances.

En d'autres termes, il est important d'avoir le nombre le plus bas possible d'avions sous maintenance à chaque mois.

#### RQ007: état final

À la fin de l'horizon de planification, l'avion doit être dans le meilleur état possible. Cela implique que, en moyenne, les avions doivent avoir un gros budget d'heures de vol restant et un grand nombre de mois avant d'avoir besoin d'un entretien.

#### RQ008: stockage

Le stockage peut être décidé par le modèle afin d'éviter la détérioration de l'aéronef. La période minimale de stockage d'un avion est de 6 mois et, dans cet état, l'avion n'est pas disponible ni pour la maintenance ni pour les missions.

#### RQ009: fixation des parties de la planification

Il existe plusieurs raisons pour lesquelles l'utilisateur a besoin de fixer une partie d'une planification précédente. Par exemple, certaines opérations de maintenance sont déjà décidées et doivent avoir lieu à l'avenir dans de dates spécifiques. Certains avions sont déjà en mission ou sous maintenance et doivent y rester au moins quelques mois dans cet état. En outre, le stockage peut être déjà décidé ou en cours.

Une autre raison de la fixation des opérations de maintenance pourrait être pour les opérations d'amélioration planifiées aux aéronefs. Ces types spéciaux d'opérations de maintenance sont décidés plusieurs mois à l'avance pour certains avions sélectionnés.

Ces assignations fixes seront communiquées au optimiseur afin qu'il ne les change pas.

#### RQ010: configuration

Les options disponibles seront, par défaut:

<!-- rut, ret -->
* **Temps maximum**: dans le cas où une solution est nécessaire avant de prouver l'optimalité.
* **Maximal GAP**: afin d'appliquer une distance maximum autorisée de la meilleure solution obtenue à la meilleure solution hypothétique possible. Cette distance peut être une différence proportionnelle ou une différence absolue.
* **Premier et dernier mois a planifier**: afin de définir l'horizon de planification à planifier.
* **Poids objectifs**: pour optimiser les différents objectifs, privilégier l'un ou l'autre objectif.
* **Paramètres de maintenance**: durée de maintenance (en mois), heures de vol maximum, temps maximum écoulé et autres paramètres liés à la maintenance.

### Cas et expérimentation

Étant donné la nature stratégique de l'application, elle sera centrée sur les cas. Cela signifie qu'il permettra de sauvegarder toutes les données pertinentes (entrée et sortie) dans le même cas. De cette façon, différentes expériences pourraient être faites en copiant et en modifiant un cas existant. Ceci est similaire à travailler avec plusieurs feuilles de calcul, chacune avec son propre nom, tout en partageant la même logique en un seul endroit.

La figure suivante montre un ensemble hypothétique de cas déjà enregistrés. Elle représente les interactions possibles entre les cas. Le cas 1 a été publié pour le mois de mars. Au cours du mois, il a reçu quelques modifications, ce qui a entraîné les cas 2 et 3. Enfin, le cas 3 a été re-publié pour le mois de mars. Plus tard, le cas 4 a été créé et modifié avant sa publication dans le cas 5. Le cas 5 a été publié en avril.

```{r, echo=FALSE, out.width='70%'}
if (html_output){
    graph_cases()
} else {
    knitr::include_graphics('../../img/cases.png')
}
```

    Exemple montrant l'interaction entre les cas et la publication des cas.

Les fonctionnalités disponibles seront les suivantes:

* Créez un nouveau cas.
* Copier un cas existant.
* Supprimer un cas.
* Comparez deux cas.
* Publier un cas.
* Exporter un cas.

#### Contenu d'une cas

Un cas consiste en toutes les informations nécessaires pour construire une planification et une planification possible construite à partir de cette information. Cette unité de stockage permet de s'intégrer facilement aux structures de données actuelles et de les exécuter, de comparer différentes expériences.

Plus explicitement:

* Ensemble d'avions avec leurs caractéristiques et informations historiques sur eux jusqu'à une certaine date.
* Ensemble de missions ou d'heures de vol à faire dans la période de planification.
* (pre)Affectation d'aéronefs à des opérations pendant l'horizon de planification.
* Futurs calendriers de maintenance.
* Configuration comme la durée de maintenance, les heures de vol après une maintenance, etc.

#### Comparaison des cas

Cette fonctionnalité peut potentiellement avoir plusieurs utilisations en fonction des besoins finaux du client. Une option serait de visualiser les différences dans les données d'entrée pour deux cas. Une autre option consisterait à visualiser les différences dans la planification résultante pour les deux cas. Enfin, il pourrait être intéressant de comparer des KPI (Key Performance Indicators) spécifiques pour deux cas, afin de sélectionner le meilleur scénario.

#### Publications de cas

Une situation courante dans la création de planifications est le fait que plusieurs cas différents peuvent exister mais il est toujours important d'identifier lequel est le 'actif'. Pour faire face à cela, un système de publication simple sera inclus dans l'application, où le «planificateur» (voir la section "Utilisateurs") peut identifier le cas qui devrait être partagé avec d'autres.

### Interface graphique utilisateur (GUI)

Toute interaction de l'utilisateur avec l'application se fera via une application HTML.

En plus de montrer des données dans des tableaux interactifs, des alternatives spécifiques peuvent être incorporées pour rendre l'utilisation plus facile et plus efficace.

Voici quelques exemples de fonctionnalités disponibles.

Ici, un gantt interactif est utilisé pour présenter et éventuellement modifier un plan de maintenance. Il est possible d'éditer les tâches de maintenance et d'assignation (changer la taille ou la déplacer) en double-cliquant dessus et en faisant glisser. En outre, il est possible d'ajouter de nouvelles affectations en double-cliquant.

```{r, echo=FALSE, out.width='70%'}

if (html_output){
    gantt
} else {
    knitr::include_graphics('./../../img/calendar.png')
}
```

    Exemple de Gantt interactif

Ici, un graphique interactif est utilisé pour présenter l'évolution du statut d'un seul avion au fil du temps. Il est possible de filtrer, voir les données dans les détails et exporter en tant qu'image.

```{r, echo=FALSE, out.width='90%'}

if (html_output){
    graphs_rem
} else {
    knitr::include_graphics('./../../img/remaining.png')
}
```

    Exemple de graphique interactif

Ici, une table interactive est utilisée pour présenter certaines données. Il est possible d'éditer le contenu, de filtrer, de réorganiser les données, etc.

```{r, echo=FALSE, out.width='90%'}

if (html_output){
    table_rem
}
```

    Exemple de table interactive

### Intégration de données

L'application inclura son propre stockage de données à travers une base de données. Toutes les informations utilisées par l'application seront stockées ici. En plus de ce système, l'application permettra plusieurs façons d'importer et d'exporter des données à partir de ce système.

## Importer des données

Un nouveau cas peut être créé à partir des options suivantes:

* Le fichier Excel actuellement utilisé.
* Un nouveau fichier modèle Excel à créer.
* Les données historiques pourraient être lues par un autre système, si disponible.

## Exporter des données

Les données dans un cas pourront être exportées de la manière suivante:

* Dans le format actuel de Excel (ou un format similaire).
* Dans le nouveau modèle Excel qui sera créé.
* Fichiers csv et / ou json si nécessaire.
* Ces capacités d'exportation permettront également de partager des informations partielles avec d'autres acteurs dans des fichiers Excel ou plats.

## Architecture et déploiement

L'application sera déployée via une architecture client-serveur. De cette façon, il serait facile de partager l'application étant donné que cela nécessitera seulement l'installation d'un navigateur web moderne (Google Chrome).

Le serveur doit être installé quelque part dans l'infrastructure du client.

La configuration matérielle minimale requise pour le serveur est la suivante:

* Ubuntu Server 64 bits.
* Noyaux> = 4
* RAM> = 8 Go
* Disque dur> = 500 Go

Les exigences logicielles minimales pour le serveur sont:

* R + Shiny Server
* python> = 3.6
* PostgreSQL

Des logiciels supplémentaires pourraient être nécessaires en fonction des spécifications finales. Par exemple, dans le cas où le modèle est trop gros pour les solveurs de source libre, un solveur commercial pourrait être nécessaire.

En fonction de la future intégration, des flux de données supplémentaires seront inclus dans la conception. Le seul flux de données connu se situe entre le fichier Excel existant et l'application.

La figure suivante présente un résumé de l'architecture à mettre en œuvre et des flux de données entre les composants.

```{r, echo=FALSE, out.width='70%'}

if (html_output){
    graph_architecture()

} else {
    knitr::include_graphics('../../img/architecture.png')
}
```

    Architecture proposée avec flux de données entre composants

Les utilisateurs interagiront via les navigateurs Web avec le serveur. Pour charger de nouvelles données, ils peuvent télécharger un nouveau fichier Excel avec les données pertinentes ou éditer un dossier existant via l'application.

Le serveur sera chargé de résoudre et de générer de nouvelles solutions de planification et de stocker toutes les informations pertinentes dans sa base de données.

Les utilisateurs seront en mesure de récupérer visuellement les résultats. En outre, ils seront en mesure d'exporter les résultats en formats csv, Excel et json en fonction de leurs besoins.

Cette architecture client-serveur offre plusieurs avantages:

* Des graphiques puissants peuvent être incorporés via des bibliothèques javascript.
* Il est plus facile à maintenir car une séparation claire des préoccupations est faite entre l'interface graphique, la logique d'utilisation et le modèle.
* Aucune installation de logiciel additionnel à un navigateur Web moderne n'est requise.
* Plusieurs utilisateurs peuvent être facilement incorporés.
* L'intégration avec d'autres systèmes peut être effectuée avec des changements minimes à la structure existante.
* Des ressources puissantes peuvent être allouées si nécessaire au niveau du serveur au lieu des devoir changer les ordinateurs des utilisateurs.
* En plus de cela, il n'y a pas de logiciel tiers à acheter. Tous les composants sont open source. La seule exception possible pourrait être le solveur mathématique, en fonction de la taille du problème et des besoins de performance.

## Utilisateurs

Deux rôles possibles sont identifiés pour les utilisateurs de cette application.

Le premier s'appellera "planificateur". Cet utilisateur est en charge de créer (manuellement ou automatiquement) des plannings différents. Le second s'appellera "spectateur". Cet utilisateur pourra visualiser les résultats de la planification créée et stockée dans l'application. En fonction des besoins du client, plusieurs profils "spectateur" pourraient être créés afin de ne partager que des parties des informations stockées avec chaque profil.

## Équipe

## Alain Haït

## Olga Battaïa

## Franco Peschiera

Un ingénieur industriel, mathématicien de profession et programmeur par cœur.

Franco a travaillé dans l'industrie de la recherche opérationnelle depuis près de 7 ans. Au cours de ce temps il a été le propriétaire et le directeur technique pour une startup espagnole (www.baobabsoluciones.es), en construisant des applications logicielles avec des modèles d'optimisation mathématique ou de simulation à l'intérieur pour plusieurs grandes entreprises espagnoles et internationales. Exemples de clients sont: Repsol et CLH (logistique pétrolière, marketing, ressources humaines), General Motors, LI Lilly et Air Liquide (planning de production), RATP, Alstom et ASM (logistique et transport) parmi tant d'autres.

Ces projets comportaient la liste suivante de tâches: la spécification des besoins du client, la construction d'un modèle mathématique ou de simulation, la programmation d'une interface utilisateur graphique, le déploiement et l'intégration avec les systèmes informatiques, la documentation et la formation des utilisateurs.

Franco a commencé un doctorat à l'université ISAE-SUPAERO en octobre 2017 sur le thème de la planification de la maintenance des avions militaires.

## Planification

Ici, les tâches requises avec leur ordre sont présentées. Le principal risque sur la planification actuelle est l'information incomplète sur:

* le problème spécifique à résoudre.
* l'infrastructure informatique disponible pour travailler avec.
* la disponibilité des utilisateurs potentiels pour discuter des exemples de solutions et tester et évaluer les fonctionnalités.

### Phase 1: exigences

Au cours de cette phase, plusieurs réunions en personne et par téléphone seront prévues.

Les conditions requises pour cette phase sont: *Aucun*.

Le livrable de cette phase sera:

* un document avec une description convenue avec les exigences minimales pour le prototype et
* un projet d'exigences supplémentaires éventuelles à préciser ultérieurement.

La durée de cette phase est estimée à: *3 semaines*.

### Phase 2: première itération

Dans cette phase, une première version du prototype sera construite et déployée.

Les conditions requises pour cette phase sont:

* la liste avec des exigences minimales.
* une alternative de déploiement pour partager l'application avec les utilisateurs.

Le livrable de cette phase sera:

* une application prototype que les utilisateurs peuvent utiliser et tester.
* exemples de solutions pour les utilisateurs à analyser.

La durée de cette phase est estimée à: *5 semaines*.

### Phase 3: deuxième itération

Dans cette phase, une version finale du prototype sera construite et déployée.

Les conditions requises pour cette phase sont:

* une liste avec les changements à faire à la première version basée sur les commentaires des utilisateurs.
* une liste avec des fonctionnalités supplémentaires à haute priorité.

Le livrable de cette phase sera:

* une application que les utilisateurs peuvent tester.
* documentation technique.
* Guide de l'utilisateur.

La durée de cette phase est estimée à: *5 semaines*.

## Technologies et techniques

* **Modélisation mathématique**: pour générer des planifications.
* **Python**: pour la modélisation, l'heuristique et l'intégration de données.
* **R**: pour les statistiques, possibles prévisions et la visualisation des données.
* **PostgreSQL**: pour stocker des données.
* **HTML5**: pour les rapports, et les visualisations.

### Modélisation mathématique

La modélisation mathématique est une technique qui fait partie de la branche de la recherche opérationnelle en mathématiques. Il permet la construction d'un ensemble d'équations pour représenter une situation spécifique bien définie et il permet l'utilisation de logiciels de avancés pour rechercher et trouver la meilleure solution qui respecte ces équations. Ces logiciels (solveurs) sont disponibles à la fois en version commerciale et open source (gratuite).

Les avantages de l'utilisation de cette technique par rapport à d'autres alternatives sont qu'elle peut prouver l'optimalité d'une solution, elle peut facilement incorporer plusieurs objectifs et il est possible de fixer facilement certaines décisions déjà prises, rendant possible l'utilisation interactive du modèle par l'utilisateur. Avec peu de modifications, le modèle peut aussi être réutilisé pour prendre en compte différentes situations qui ne soient si différentes de la situation originelle.

D'autres fonctionnalités incluent également des limites de temps et d'autres conditions d'arrêt plus complexes.

### Python

Python permet le prototypage rapide ainsi qu'une bonne syntaxe de code. Il offre une extrêmement large ensemble des bibliothèques et, en particulier, il inclut les meilleures API pour les solveurs commerciaux et non commerciaux pour les problèmes d'optimisation.

Il excelle également au développement de Web avec des bibliothèques connus.

### R

R est l'état de l'art en matière de statistiques et de modélisation prévisionnelle. Il comprend plusieurs bibliothèques pour la manipulation des données, la visualisation et, en tant que langage de programmation, une bonne intégration avec d'autres langages et logiciels.

Enfin, il offre une intégration particulièrement bonne avec les applications html via le module [Shiny](https://shiny.rstudio.com/gallery/), qui sera utilisée dans ce projet.

### PostgreSQL

PostgreSQL est l'une des alternatives de base de données open-source les plus connues. Il comprend nombreuses fonctionnalités qui ne sont pas présentes dans d'autres moteurs de base de données (à savoir MySQL) tels que:

* NoSQL stockage.
* Fonctions de fenêtre intégrées à R.

### HTML5

HTML 5 est le changement le plus récent au HTML qui intègre des fonctionnalités graphiques. Cette norme a permis une multitude de bibliothèques à source ouverte qui offrent des graphiques interactifs très puissants qui sont facilement incorporés dans les pages Web.

Des exemples de ces bibliothèques sont:  parmi beaucoup d'autres.

### Avertissements

1. La proposition présentée ci-dessus est destinée à une application complète livrée dans le cadre d'un projet complet à la DGA.
1. Le prototype à livrer n'inclura pas toutes les fonctionnalités mentionnées dans ce document mais montrera des exemples des fonctionnalités les plus importantes suite à une réunion avec les futurs utilisateurs et décideurs.
1. L'intégration avec d'autres systèmes est hors de la portée de cette proposition car plus d'informations sont nécessaires concernant les systèmes existants pour pouvoir estimer la charge de travail.
1. Il est supposé que DGA fournira les exigences hardware ainsi que les droits d'accès et les autorisations pour installer tous les composants du logiciel. Alternativement, le logiciel pourrait être installé dans le cloud après accord.
1. Même si la programmation mathématique fournit la capacité de prouver l'optimalité d'une solution, l'application ne garantit pas la trouver dans tous les cas. La qualité des solutions et le temps nécessaire pour les obtenir dépendront de nombreux facteurs; la taille de l'instance étant la principale.

## References

### 2018-03-05 R/diagrams/fixed_planning.R

@@ -1,0 +1,43 @@
+library(timevis)
+library(tidyverse)
+library(data.table)
+library(magrittr)
+
+config <- list(
+    stack = FALSE,
+    editable = TRUE,
+    align = "center",
+    orientation = "top",
+    snap = NULL,
+    margin = 0,
+    zoomable= FALSE
+)
+
+data.table(
+    id = 'A1'
+    ,start= c('2018-01', '2018-10', '2019-05')
+    ,end= c('2018-01', '2018-10', '2019-05')
+)
+
+states <- 
+    solution %>% 
+    extract2('state') %>% 
+    bind_rows(.id = "UNIT") %>% 
+    gather(key = 'month', value = 'state', -UNIT) %>% 
+    filter(state %>% is.na %>% not) %>% 
+    bind_rows(tasks) %>% 
+    collapse_states() %>% 
+    inner_join(task_hours) %>% 
+    mutate(id= c(1:nrow(.)),
+           style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color),
+           content = if_else(state=='M', 'M', sprintf('%s (%sh)', state, hours))
+    ) %>% 
+    select(id, start= month, end= post_month, content, group= UNIT, style)
+
+if (max_resources %>% is.null %>% not){
+    resources <- states %>% distinct(group) %>% sample_n(max_resources)
+    states <- states %>% inner_join(resources)
+}
+
+groups_c <- states %>% distinct(group) %>% unlist
+groups <- data.table(id= groups_c, content= groups_c)

### 2018-03-07 R/OptiPlan_presentation/proposal_fr.Rmd
Added new images, new version of document Optiplan. Compatible with PDF and HTML output

@@
 title: "Optiplan proposal"
-date: "March 11, 2018"
+date: "Mars 11, 2018"
+nocite: '@*'
 output:
+  pdf_document:
+    toc: yes
+    # toc_depth: 2
+    # template: ./../../latex/templates/baobab.tex
+    template: isae-report-template_pandoc.tex
+    # top_level_division: part
+    keep_tex: true
+    pandoc_args: ["--top-level-division=part"]
   html_document:
     css: ./../../css/baobab_docs.css
-    toc: true
+    toc: yes
   beamer_presentation:
-    theme: "Berlin"
-    toc: true
-    slide_level: 3
-    incremental: yes
     fig_height: 4
-    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
     includes:
       in_header: header.tex
+    incremental: yes
+    slide_level: 3
+    theme: Berlin
+    toc: yes
   slidy_presentation:
     css: ./../../css/baobab_MH.css
+    footer: ISAE-SUPAERO, Université de Toulouse, France
     incremental: yes
-    footer: "ISAE-SUPAERO, Université de Toulouse, France"
     slide_level: 4
-    toc: true
-nocite: '@*'
+    toc: yes
 bibliography: MFMP.bib
+editor_options: 
+  chunk_output_type: console
 ---
@@
 source('diagrams/exampleDT.R')
+source('diagrams/gantt_tasks.R')
 exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
+exp_directory2 = PATHS[['experiments']] %>% paste0('201802061201/')
 gantt <- print_solution(exp_directory, max_resources = 5, width="100%", height="200px")
+gantt_tasks <- print_tasks(exp_directory2, width="100%", height="200px", max_tasks = 5)
 graphs_rem <- graph_remaining(exp_directory)
@@
 * L'équipe a une connaissance unique du problème spécifique, car une thèse de doctorat sur ce sujet a été lancée en octobre 2017.
-Les travaux ont été présentés au ROADEF 2018 et seront présentés dans la conférence MOSIM 2018.
+Les travaux ont été présentés déjà à la conference ROADEF 2018 et seront présentés dans la conférence MOSIM 2018.
 
@@
 
-Le problème est considéré comme un problème de planification de vol et de maintenance (Flight and Maintenance Planning problem).
+L'Armée de l'Air dispose d'une flotte d'avions Mirage 2000 de différents types, avec des fonctionnalités particulaires et avec un état d'opération spécifique donné par certains indicateurs temporelles d'usage.
+
+Au même temps, il existe un besoin opérationnel généré par les différents missions que la flotte doit être capable de remplir.
+Ce besoin peut être formulé comme un nombre de missions à satisfaire ou un budget d'heures de vol a atteindre à chaque période.
+
+Finalement, il existe de contraintes opératifs du au fait que les avions ont besoin de maintenances régulières et possibles stockages pour être conservés en bon état et, de cette façon, garantir la capacité future de la flotte.
+
+Le problème consiste, donc, à affecter des avions aux missions prédéfinies et à programmer des opérations de maintenance préventive pour ces mêmes avions de façon d'avoir le plus grand quantité d'avions disponibles en lissant et réduisant les opérations de maintenance le plus possible dans le temps, sans impacter la capacité de la flotte présent ou future.
+
+Dans la littérature, le problème est considéré comme un problème de planification de vol et de maintenance (Flight and Maintenance Planning problem).
 Travail a déjà été fait par @Kozanidis2008, @Cho2011, @Verhoeff2015, entre autres.
 
-Le problème consiste à affecter des avions aux missions prédéfinies et à programmer des opérations de maintenance préventive périodique pour ces mêmes avions.
+### Avions
 
-Les missions peuvent aussi être comprises comme un budget d'heures par mois ou par an.
-Dans ce document, nous présentons la formulation la plus générale possible, où les missions sont affectés aux avions pour un ou plusieurs mois.
+Il y a un ensemble $i \in \mathcal{I}$ des avions disponibles à être affectés aux missions.
+Chaque avion ne peut être affecté qu'à une seule mission au cours d'une période donnée.
+Ces avions souffrent d'usure et nécessitent des opérations de maintenance régulières pendant leur durée de vie.
+Le besoin de maintenance est calculé sur la base de deux indicateurs.
+
+Le premier est appelé "temps écoulé restant" (ou $ret_{it}$).
+Il exprime la durée (mesurée en périodes) après laquelle l'aéronef ne peut plus être utilisé et doit subir une opération de maintenance.
+Sa valeur est calculée pour chaque avion   $i$ et chaque période de temps $t$.
+De la même manière, "temps d'utilisation restant" (ou $rut_{it}$) est utilisé pour mesurer la quantité de temps que l'avion $i$ peut être utilisé avant d'avoir besoin d'une opération de maintenance à une période donnée $t$.
+
+De plus, après une quantité absolue de temps et / ou d'utilisation ($aet_i$ ou $aut_i$), l'aéronef devient obsolète. Il n'y a aucun moyen d'inverser ce processus.
+
+À toute période donnée, y compris au début de l'horizon de planification, chaque aéronef a un statut spécifique donné par le temps d'utilisation restant et le temps écoulé restant.
 
 ### Missions
+
+Dans ce document, nous présentons la formulation la plus générale possible, où les missions sont affectés aux avions pour un ou plusieurs mois.
 
@@
 
-### Avion
+```{r, echo=FALSE, out.width='90%'}
+if (html_output){
+    gantt_tasks
+} else {
+    knitr::include_graphics('../../img/tasks_gantt.png')
+}
+```
 
-Il y a un ensemble $i \in \mathcal{I}$ des avions disponibles à être affectés aux missions.
-Chaque avion ne peut être affecté qu'à une seule mission au cours d'une période donnée.
-Ces avions souffrent d'usure et nécessitent des opérations de maintenance régulières pendant leur durée de vie.
-Le besoin de maintenance est calculé sur la base de deux indicateurs.
-
-<!-- TODO: traduire ret et rut. -->
-
-Le premier est appelé "temps écoulé restant" (ou $ret_{it}$).
-Il exprime la durée (mesurée en périodes) après laquelle l'aéronef ne peut plus être utilisé et doit subir une opération de maintenance.
-Sa valeur est calculée pour chaque avion   $i$ et chaque période de temps $t$.
-De la même manière, "temps d'utilisation restant" (ou $rut_{it}$) est utilisé pour mesurer la quantité de temps que l'avion $i$ peut être utilisé avant d'avoir besoin d'une opération de maintenance à une période donnée $t$.
-
-De plus, après une quantité absolue de temps et / ou d'utilisation ($aet_i$ ou $aut_i$), l'aéronef devient obsolète. Il n'y a aucun moyen d'inverser ce processus.
-
-<!-- TODO: traduire ret et rut. -->
-À toute période donnée, y compris au début de l'horizon de planification, chaque aéronef a un statut spécifique donné par le temps d'utilisation restant et le temps écoulé restant.
+    Graphique montrant possibles missions et ses besoins.
 
@@
 
-<!-- TODO: traduire ret et rut. -->
 Les opérations de maintenance sont le processus par lequel les aéronefs qui ont atteint une limite dans certains indicateurs peuvent revenir à un état où ils peuvent continuer à être utilisés dans les missions.
@@
 
-Après une opération de maintenance, un aéronef restaure le temps restant et le temps d'utilisation restant à leurs valeurs maximales $E$ (~ 60 mois) et $H$ (~ 1200 heures) respectivement.
+Après une opération de maintenance, un aéronef restaure le temps écoulé restant et le temps d'utilisation restant à leurs valeurs maximales $E$ (~ 60 mois) et $H$ (~ 1200 heures) respectivement.
 
@@
 
-<!-- TODO: traduire ret et rut. -->
-Les temps restants utilisés et écoulés sont assignés à chaque aéronef au début de l'horizon de planification.
+Les temps (écoulé et d'utilisation) restants sont assignés à chaque aéronef au début de l'horizon de planification.
 
-<!-- TODO: traduire ret et rut. -->
-Enfin, l'état de chaque aéronef à la fin de l'horizon de planification, son temps restant (écoulé, utilisé), doit être défini et pris en compte.
+Enfin, l'état de chaque aéronef à la fin de l'horizon de planification, ses temps restant (écoulé, utilisé), doivent être défini et pris en compte.
 
@@
 
-Compte tenu de la capacité limitée de maintenance et de son coût, un autre objectif consiste à lisser le plus possible le nombre d'avions sous maintenance sur l'horizon de planification.
+Compte tenu de la capacité limitée pour faire opérations de maintenance et de son coût, un autre objectif consiste à lisser le plus possible le nombre d'avions sous maintenance dans l'horizon de planification.
 Les opérations de maintenance doivent également être contrôlées afin de ne pas dépasser la capacité maximale.
 
-### Proposition
+## Proposition
 
@@
 
-```{r, echo=FALSE, out.width='70%'}
+```{r, echo=FALSE, out.width='90%'}
 if (html_output){
@@
 
-    Graphique montrant les fonctionnalités de l'application
+    Graphique montrant les fonctionnalités de l'application.
 
@@
 
-Chaque mois, toutes les missions doivent être satisfaites en affectant suffisamment d'heures de vol à chaque avion.
+Chaque mois, toutes les missions doivent être satisfaites grâce à l'affectation des suffisamment d'heures de vol à chaque avion.
 
@@
 
-Chaque avion doit toujours avoir suffisamment d'heures de vol pour pouvoir effectuer les missions requises.
+Chaque avion doit toujours avoir suffisamment de "temps d'utilisation restant" pour pouvoir effectuer les missions requises.
 Afin de récupérer des heures de vol (environ 1200 heures), une opération de maintenance est nécessaire.
@@
 
-Les opérations de maintenance à planifier pour tous les avions doivent être effectuées de la manière la plus lissé possible, en évitant avoir des mois spécifiques avec un grand quantité des maintenances.
+Les opérations de maintenance à planifier pour tous les avions doivent être effectuées de la manière la plus lissé possible, en évitant avoir des mois spécifiques avec un grand quantité des opérations de maintenance.
 
@@
 À la fin de l'horizon de planification, l'avion doit être dans le meilleur état possible.
-Cela implique que, en moyenne, les avions doivent avoir un gros budget d'heures de vol restant et un grand nombre de mois avant d'avoir besoin d'un entretien.
+Cela implique que, en moyenne, les avions doivent avoir un gros budget d'heures de vol restant (temps d'utilisation restant) et un grand nombre de mois (temps écoulé restant) avant d'avoir besoin d'un entretien.
 
@@
 Le stockage peut être décidé par le modèle afin d'éviter la détérioration de l'aéronef.
-La période minimale de stockage d'un avion est de 6 mois et, dans cet état, l'avion n'est pas disponible ni pour la maintenance ni pour les missions.
+Le nombre de périodes minimale de stockage d'un avion est de 6 mois et, dans cet état, l'avion n'est pas disponible ni pour recevoir un maintenance ni pour faire une mission.
 
@@
 
-Il existe plusieurs raisons pour lesquelles l'utilisateur a besoin de fixer une partie d'une planification précédente.
+Il existe plusieurs raisons pour lesquelles l'utilisateur peut avoir besoin de fixer une partie d'une planification précédente.
 Par exemple, certaines opérations de maintenance sont déjà décidées et doivent avoir lieu à l'avenir dans de dates spécifiques.
@@
 * **Temps maximum**: dans le cas où une solution est nécessaire avant de prouver l'optimalité.
-* **Maximal GAP**: afin d'appliquer une distance maximum autorisée de la meilleure solution obtenue à la meilleure solution hypothétique possible.
+* **Maximal GAP**: afin d'autoriser une distance maximal entre la meilleure solution obtenue et la meilleure solution hypothétique possible.
 Cette distance peut être une différence proportionnelle ou une différence absolue.
 * **Premier et dernier mois a planifier**: afin de définir l'horizon de planification à planifier.
-* **Poids objectifs**: pour optimiser les différents objectifs, privilégier l'un ou l'autre objectif.
-* **Paramètres de maintenance**: durée de maintenance (en mois), heures de vol maximum, temps maximum écoulé et autres paramètres liés à la maintenance.
+* **Poids des objectifs**: pour optimiser les différents objectifs et privilégier un ou autre objectif.
+* **Paramètres de maintenance**: durée de maintenance (en mois), temps d'utilisation maximal, temps écoulé maximal et d'autres paramètres liés à la maintenance.
 
@@
 Enfin, le cas 3 a été re-publié pour le mois de mars.
-Plus tard, le cas 4 a été créé et modifié avant sa publication dans le cas 5.
+Plus tard, le cas 4 a été créé et modifié sans être publié dans le cas 5.
 Le cas 5 a été publié en avril.
@@
 
-* Créez un nouveau cas.
+* Créer un nouveau cas.
 * Copier un cas existant.
 * Supprimer un cas.
-* Comparez deux cas.
+* Comparer deux cas.
 * Publier un cas.
@@
 
-Un cas consiste en toutes les informations nécessaires pour construire une planification et une planification possible construite à partir de cette information.
-Cette unité de stockage permet de s'intégrer facilement aux structures de données actuelles et de les exécuter, de comparer différentes expériences.
+Un cas consiste, au même temps, de toutes les informations nécessaires pour construire une planification et une planification réalisable construite à partir de cette information.
 
-Plus explicitement:
+Cette unité de stockage permet de s'intégrer facilement aux structures de données actuelles.
 
-* Ensemble d'avions avec leurs caractéristiques et informations historiques sur eux jusqu'à une certaine date.
-* Ensemble de missions ou d'heures de vol à faire dans la période de planification.
-* (pre)Affectation d'aéronefs à des opérations pendant l'horizon de planification.
+Plus explicitement, un cas consiste en:
+
+* L'ensemble d'avions avec leurs caractéristiques et leurs informations historiques jusqu'à une certaine date.
+* L'ensemble de missions ou d'heures de vol à faire dans la période de planification.
+* Une pre-affectation d'aéronefs à des opérations pendant l'horizon de planification.
 * Futurs calendriers de maintenance.
 * Configuration comme la durée de maintenance, les heures de vol après une maintenance, etc.
+* Une planification (généré manuellement ou avec l'Optimiseur).
 
@@
 
-Une situation courante dans la création de planifications est le fait que plusieurs cas différents peuvent exister mais il est toujours important d'identifier lequel est le 'actif'.
+Une situation courante dans la création de planifications est le fait que plusieurs cas différents peuvent exister mais il est toujours important d'identifier lequel est le cas 'actif'.
 Pour faire face à cela, un système de publication simple sera inclus dans l'application, où le «planificateur» (voir la section "Utilisateurs") peut identifier le cas qui devrait être partagé avec d'autres.
@@
 
-En plus de montrer des données dans des tableaux interactifs, des alternatives spécifiques peuvent être incorporées pour rendre l'utilisation plus facile et plus efficace.
+En plus de montrer des données dans des tableaux interactifs, des alternatives graphiques spécifiques peuvent être incorporées pour rendre l'utilisation plus facile et plus efficace.
 
@@
 Ici, un gantt interactif est utilisé pour présenter et éventuellement modifier un plan de maintenance.
-Il est possible d'éditer les tâches de maintenance et d'assignation (changer la taille ou la déplacer) en double-cliquant dessus et en faisant glisser.
+Il est possible d'éditer les opérations de maintenance et l’affection aux missions (changer la taille ou la déplacer) en double-cliquant dessus et en faisant glisser.
 En outre, il est possible d'ajouter de nouvelles affectations en double-cliquant.
@@
 } else {
-    knitr::include_graphics('./../../img/calendar.png')
+    knitr::include_graphics('./../../img/calendar2.png')
 }
@@
 
-    Exemple de Gantt interactif
+    Exemple de Gantt interactif.
 
-Ici, un graphique interactif est utilisé pour présenter l'évolution du statut d'un seul avion au fil du temps.
-Il est possible de filtrer, voir les données dans les détails et exporter en tant qu'image.
+Ici, un graphique interactif est utilisé pour présenter l'évolution du statut d'un seul avion au fil du temps. Il est possible de filtrer, faire zoom et exporter en tant qu'image.
 
@@
 
-    Exemple de graphique interactif
+    Exemple de graphique interactif.
 
@@
     table_rem
+} else {
+    knitr::include_graphics('./../../img/table.png')
 }
@@
 
-    Exemple de table interactive
+    Exemple de table interactive.
 
@@
 
-L'application inclura son propre stockage de données à travers une base de données.
+L'application inclura son propre stockage de données à travers d'une base de données.
 Toutes les informations utilisées par l'application seront stockées ici.
@@
 
-## Importer des données
+#### Importer des données
 
@@
 
-## Exporter des données
+#### Exporter des données
 
@@
 
-* Dans le format actuel de Excel (ou un format similaire).
-* Dans le nouveau modèle Excel qui sera créé.
+* Dans le format d l'Excel actuel (ou un format similaire).
+* Dans le nouveau template Excel qui sera créé.
 * Fichiers csv et / ou json si nécessaire.
@@
 
-## Architecture et déploiement
+### Architecture et déploiement
 
@@
 
-Le serveur doit être installé quelque part dans l'infrastructure du client.
+Le serveur devra être installé quelque part dans l'infrastructure du client.
 
@@
 
-Les exigences logicielles minimales pour le serveur sont:
+Les exigences logicielles minimales pour le serveur sont les suivants:
 
@@
 Des logiciels supplémentaires pourraient être nécessaires en fonction des spécifications finales.
-Par exemple, dans le cas où le modèle est trop gros pour les solveurs de source libre, un solveur commercial pourrait être nécessaire.
+Par exemple, dans le cas où le modèle soit trop gros pour les solveurs de source libre, un solveur commercial pourrait être nécessaire.
 
@@
 
-    Architecture proposée avec flux de données entre composants
+    Architecture proposée avec flux de données entre composants.
 
 Les utilisateurs interagiront via les navigateurs Web avec le serveur.
-Pour charger de nouvelles données, ils peuvent télécharger un nouveau fichier Excel avec les données pertinentes ou éditer un dossier existant via l'application.
+Pour charger de nouvelles données, ils peuvent charger un nouveau fichier Excel avec les données pertinentes ou éditer un cas existant via l'application.
 
@@
 
-* Des graphiques puissants peuvent être incorporés via des bibliothèques javascript.
-* Il est plus facile à maintenir car une séparation claire des préoccupations est faite entre l'interface graphique, la logique d'utilisation et le modèle.
+* Des graphiques puissants peuvent être incorporés via des bibliothèques javascript et HTML.
+* Il est plus facile à maintenir car une séparation claire des fonctionnalités est faite entre l'interface graphique, la logique d'utilisation et l'Optimiseur.
 * Aucune installation de logiciel additionnel à un navigateur Web moderne n'est requise.
@@
 Tous les composants sont open source.
-La seule exception possible pourrait être le solveur mathématique, en fonction de la taille du problème et des besoins de performance.
+La seule exception possible pourrait être le solveur mathématique, en fonction de la taille du problème et les besoins de performance.
 
-## Utilisateurs
+### Utilisateurs
 
@@
 Cet utilisateur pourra visualiser les résultats de la planification créée et stockée dans l'application.
-En fonction des besoins du client, plusieurs profils "spectateur" pourraient être créés afin de ne partager que des parties des informations stockées avec chaque profil.
+En fonction des besoins du client, plusieurs profils "spectateur" pourraient être créés afin de partager quelques parties des informations stockées avec chaque profil.
 
@@
 
-## Alain Haït
+### Alain Haït
 
-## Olga Battaïa
+### Olga Battaïa
 
-## Franco Peschiera
+### Franco Peschiera
 
@@
 Au cours de ce temps il a été le propriétaire et le directeur technique pour une startup espagnole (www.baobabsoluciones.es), en construisant des applications logicielles avec des modèles d'optimisation mathématique ou de simulation à l'intérieur pour plusieurs grandes entreprises espagnoles et internationales.
-Exemples de clients sont: Repsol et CLH (logistique pétrolière, marketing, ressources humaines), General Motors, LI Lilly et Air Liquide (planning de production), RATP, Alstom et ASM (logistique et transport) parmi tant d'autres.
+Exemples de clients sont: Repsol et CLH (logistique pétrolière, marketing, ressources humaines), General Motors, LI Lilly et Air Liquide (planning de production), RATP, Alstom et ASM (logistique et transport) parmi plusieurs autres.
 
@@
 
-Franco a commencé un doctorat à l'université ISAE-SUPAERO en octobre 2017 sur le thème de la planification de la maintenance des avions militaires.
+Franco a commencé un doctorat à l'université ISAE-SUPAERO en octobre 2017 sur le sujet de la planification de la maintenance des avions militaires.
 
@@
 * le problème spécifique à résoudre.
-* l'infrastructure informatique disponible pour travailler avec.
-* la disponibilité des utilisateurs potentiels pour discuter des exemples de solutions et tester et évaluer les fonctionnalités.
+* l'infrastructure informatique disponible du client.
+* la disponibilité des utilisateurs potentiels pour discuter des exemples de solutions et tester et évaluer les fonctionnalités faits.
+
+On présente une proposition de planification basé en expériences précédents. La planification dépendra de la voie de communication finalement choisi et la disponibilité du client.
 
@@
 
-Au cours de cette phase, plusieurs réunions en personne et par téléphone seront prévues.
+Au cours de cette phase, des réunions en personne et par téléphone seront prévues.
 
@@
 
-* un document avec une description convenue avec les exigences minimales pour le prototype et
-* un projet d'exigences supplémentaires éventuelles à préciser ultérieurement.
+* un document avec une description accordé sur les exigences minimales pour le prototype et
+* un document d'exigences additionnelles à préciser ultérieurement.
 
@@
 Dans cette phase, une version finale du prototype sera construite et déployée.
+En fonction de niveau d'implication du client, le types de changements à réaliser et l'urgence, cette phase pourrait être coupé en deux phases de la moitié de durée.
 
@@
 * une liste avec les changements à faire à la première version basée sur les commentaires des utilisateurs.
-* une liste avec des fonctionnalités supplémentaires à haute priorité.
+* une liste avec des fonctionnalités supplémentaires à ajouter.
 
@@
 
-* une application que les utilisateurs peuvent tester.
-* documentation technique.
-* Guide de l'utilisateur.
+* une nouvelle version de l'application que les utilisateurs pourront utiliser.
+* la documentation technique.
+* la Guide de l'utilisateur.
 
 La durée de cette phase est estimée à: *5 semaines*.
+
+### Profils et engagement
+
+Il y aura trois profils dans ce projet:
+
+* Profil 1: Consultant Senior spécialisé en Recherche Opérationnelle (RO).
+* Profil 2: Développeur Senior spécialisé en RO.
+* Profil 3: Chef de projet spécialise en RO.
+
+Dans le tableau ci-dessous sont les engagements hebdomadaires estimés de profils:
+
+| Phase | Durée | Profil 1 | Profil 2 | Profil 3
+|-------|----------|----------|----------|---------|
+|Phase 1| 3 semaines| 4 h    | 30 h    | 8 h     |
+|Phase 2| 5 semaines| 2 h    | 30 h    | 8 h     |
+|Phase 3| 5 semaines| 2 h    | 30 h    | 8 h     |
+
+### Estimation économique
+
+Basé en la définition de profils et les engagements déjà présente, le montant correspondant au projet de développement du prototype est de 55.600€.
 
@@
 * **Modélisation mathématique**: pour générer des planifications.
+    * **Solveur mathématique**: pour générer les planifications. Cela pourrait être CBC, GUROBI ou CPLEX en fonction de besoins du modèle.
 * **Python**: pour la modélisation, l'heuristique et l'intégration de données.
@@
 Il permet la construction d'un ensemble d'équations pour représenter une situation spécifique bien définie et il permet l'utilisation de logiciels de avancés pour rechercher et trouver la meilleure solution qui respecte ces équations.
-Ces logiciels (solveurs) sont disponibles à la fois en version commerciale et open source (gratuite).
+Ces logiciels (modélisateurs et solveurs) sont disponibles à la fois en version commerciale et open source (gratuite).
 
@@
 
-Il excelle également au développement de Web avec des bibliothèques connus.
+Il excelle également au développement de Web avec des bibliothèques très connus.
 
@@
 
-Enfin, il offre une intégration particulièrement bonne avec les applications html via le module [Shiny](https://shiny.rstudio.com/gallery/), qui sera utilisée dans ce projet.
+Enfin, il offre une intégration particulièrement bonne avec les applications HTML via le module [Shiny](https://shiny.rstudio.com/gallery/), qui sera utilisée dans ce projet.
 
@@
 
-Des exemples de ces bibliothèques sont:  parmi beaucoup d'autres.
+Des exemples de ces bibliothèques sont: [leaflet](http://leafletjs.com/), [visjs](visjs.org), [d3js](https://d3js.org/) parmi beaucoup d'autres.
 
-### Avertissements
+## Avertissements
 
-1. La proposition présentée ci-dessus est destinée à une application complète livrée dans le cadre d'un projet complet à la DGA.
+1. La proposition présentée ci-dessus est destinée à une application complète livrée dans le cadre d'un projet complet au client.
 1.
@@
 1.
-Il est supposé que DGA fournira les exigences hardware ainsi que les droits d'accès et les autorisations pour installer tous les composants du logiciel.
-Alternativement, le logiciel pourrait être installé dans le cloud après accord.
+Il est supposé que le client fournira les exigences hardware ainsi que les droits d'accès et les autorisations pour installer tous les composants du logiciel.
+Alternativement, le logiciel pourrait être installé dans le cloud.
 1.
-Même si la programmation mathématique fournit la capacité de prouver l'optimalité d'une solution, l'application ne garantit pas la trouver dans tous les cas.
+Même si la programmation mathématique offre la capacité de prouver l'optimalité d'une solution, l'application ne garantit pas la trouver dans tous les cas.
 La qualité des solutions et le temps nécessaire pour les obtenir dépendront de nombreux facteurs; la taille de l'instance étant la principale.
 
-## References
+## Références

### 2018-03-07 R/diagrams/gantt_tasks.R

@@ -1,0 +1,41 @@
+library(tidyverse)
+library(readxl)
+library(magrittr)
+library(stringr)
+library(zoo)
+library(data.table)
+library(lubridate)
+library(timevis)
+library(Hmisc)
+library(jsonlite)
+
+print_tasks <- function(exp_directory, max_tasks, ...){
+    input_path = exp_directory %>% paste0('data_in.json')
+    input <- read_json(input_path)
+    tasks <-
+        input %>% 
+        extract2('tasks') %>% 
+        lapply("[", c('consumption', 'num_resource', 'start', 'end')) %>% 
+        bind_rows(.id="group") %>% 
+        sample_n(max_tasks) %>% 
+        arrange(consumption) %>% 
+        mutate(id= c(1:nrow(.)),
+               content = sprintf('%s aéronefs (%sh)', num_resource, consumption),
+               color = RColorBrewer::brewer.pal(5, "YlOrRd")[rep_len(1:5, n())],
+               style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color),
+        ) %>% 
+        slice(sample(1:n()))
+        
+    config <- list(
+        stack = FALSE,
+        editable = TRUE,
+        align = "center",
+        orientation = "top",
+        snap = NULL,
+        margin = 0,
+        zoomable= FALSE
+    )
+    
+    groups <- tasks %>% distinct(group) %>% rename(id= group) %>% mutate(content= id)
+    timevis(tasks, groups= groups, options= config, ...)
+}

### 2018-03-07 R/scripts/import_results.R

@@ -40,31 +40,6 @@ collapse_states <- function(table){
         ungroup() %>% filter(state != "")
 }
 
-print_tasks <- function(exp_directory){
-    input_path = exp_directory %>% paste0('data_in.json')
-    input <- read_json(input_path)
-    tasks <-
-        input %>% 
-        extract2('tasks') %>% 
-        lapply("[", c('consumption', 'num_resource', 'start', 'end')) %>% 
-        bind_rows(.id="group") %>% 
-        mutate(id= c(1:nrow(.)),
-               content = sprintf('%s aircraft (%sh)', num_resource, consumption),
-               color = RColorBrewer::brewer.pal(5, "YlOrRd")[c(1, 2)],
-               style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color),
-        )
-    config <- list(
-        stack = FALSE,
-        editable = TRUE,
-        align = "center",
-        orientation = "top",
-        snap = NULL,
-        margin = 0
-    )
-    groups <- tasks %>% distinct(group) %>% rename(id= group) %>% mutate(content= id)
-    timevis(tasks, groups= groups, options= config, width="100%")
-}
-
 print_solution <- function(exp_directory, max_resources=NULL, ...){
     solution_path = exp_directory %>% paste0('data_out.json')
     input_path = exp_directory %>% paste0('data_in.json')

### 2018-03-12 R/OptiPlan_presentation/proposal_fr.Rmd
Document presented in 2018-03-11 for Optiplan.

@@
 title: "Optiplan proposal"
+author: "ISAE-SUPAERO, Université de Toulouse"
 date: "Mars 11, 2018"
@@
 output:
+  html_document:
+    css: ./../../css/baobab_docs.css
+    toc: yes
+  beamer_presentation:
+    pandoc_args: ["--slide-level=3", "--top-level-division=part"]
+    # fig_height: 4
+    # includes:
+    #   in_header: header.tex
+    # incremental: yes
+    # slide_level: 3
+    theme: Madrid
+    # toc: yes
   pdf_document:
@@
     pandoc_args: ["--top-level-division=part"]
-  html_document:
-    css: ./../../css/baobab_docs.css
-    toc: yes
-  beamer_presentation:
-    fig_height: 4
-    includes:
-      in_header: header.tex
-    incremental: yes
-    slide_level: 3
-    theme: Berlin
-    toc: yes
   slidy_presentation:
@@
     footer: ISAE-SUPAERO, Université de Toulouse, France
-    incremental: yes
-    slide_level: 4
+    # incremental: yes
+    slide_level: 3
+    pandoc_args: ["--slide-level=3"]
     toc: yes
@@
 exp_directory2 = PATHS[['experiments']] %>% paste0('201802061201/')
-gantt <- print_solution(exp_directory, max_resources = 5, width="100%", height="200px")
+gantt <- print_solution(exp_directory, max_resources = 10, width="100%", height="500px")
 gantt_tasks <- print_tasks(exp_directory2, width="100%", height="200px", max_tasks = 5)
@@
 ## Introduction 
+### Introduction
 
-Dans ce document, nous présentons d'abord le problème tel que l'équipe le comprend.
-Les sources d'information sont la connaissance préliminaire de l'équipe et la réunion OptiPlan qui a eu lieu à Paris.
+Dans ce document, nous présentons tout d’abord le problème tel que l’équipe le comprend.
+Les sources d’information sont la connaissance préliminaire de l’équipe et la réunion OptiPlan qui a eu lieu à Paris.
 
-Une proposition technique est présentée qui inclut les fonctionnalités qu'une éventuelle application complète inclurait sur la base des échanges effectués lors de la journée OptiPlan à Paris et des expériences précédentes de l'équipe dans la construction d'applications d'aide à la planification pour l'industrie.
+Notre proposition technique inclut des fonctionnalités qu’une éventuelle application complète comprendrait.
+Dans le temps imparti au projet nous projetons de réaliser un prototype basé sur cette proposition technique
 
-Ensuite, les références de l'équipe sont présentées, mentionnant les différents projets et recherches pertinents liés au présent projet.
+## Présentation de l'entreprise
+### Présentation de l'entreprise
 
-Une planification de temps est présenté pour le développement du logiciel prototype, avec les livrables par étape.
+Nom: ISAE-SUPAERO
+Activité: école d’ingénieurs.
 
-Enfin, les technologies et les outils logiciels à utiliser sont présentés, décrivant l'utilisation et justifiant le choix à chaque point.
+Grand école du secteur aéronautique et spatial, l'ISAE-SUPAERO développe depuis 2007 une recherche propre portée par ses département.
+Le département d'ingénierie des systèmes complexes (DISC), par son groupe Systèmes Décisionnels, travaille sur les outils de décision appliqués aux systèmes industriels, en particulier pour la production et la maintenance dans l'industrie aérospatiale.
 
-## Résumé
+## Savoir-faire
+### Compétences clés
 
-Le projet présenté est une application web interactive.
-Il utilise un modèle mathématique pour générer de nouvelles solutions, permet à différents utilisateurs d'y accéder avec différents droits d'accès et offre une puissante visualisation pour afficher, comparer et analyser les planifications.
-Il s'intègre aux sources de données existantes (fichier Excel), aux nouvelles proposées (modèle Excel) et a la possibilité de s’intégrer aux sources existantes (autres systèmes internes) ainsi qu'à la sortie simple (csv, json).
-Enfin, il permet l'utilisation de cas pour stocker des informations, ce qui facilite les essaies, l'expérimentation et la comparaison de différents scénarios.
+Nous pensons que cette équipe est particulièrement bien placée pour
+aborder ce projet pour les raisons suivantes:
 
-Nous pensons que cette équipe est particulièrement bien placée pour aborder ce projet pour les raisons suivantes:
+L’équipe possède une grande expérience en recherche opérationnelle:
 
-* L'équipe somme plus de 40 ans d'expérience dans le domaine de la recherche opérationnelle, qui est la théorie de bas derrière ce problème.
-* L'équipe possède des années d'expérience dans la conception d'applications logicielles de planification et d'optimisation pour un large ensemble de grandes entreprises industrielles.
-* L'équipe possède une vaste expérience produisant articles scientifiques et projets commerciaux liés à la planification et à la maintenance.
-* L'équipe a une connaissance unique du problème spécifique, car une thèse de doctorat sur ce sujet a été lancée en octobre 2017.
-Les travaux ont été présentés déjà à la conference ROADEF 2018 et seront présentés dans la conférence MOSIM 2018.
+* recherche scientifique, publication d’articles
+* projets de recherche en collaboration avec l’industrie
+* conception d’applications logicielles en planification et optimisation pour des entreprises
+
+L’équipe a déjà encadré des travaux de recherche sur le sujet de la maintenance aéronautique.
+En outre, elle a une connaissance du problème en question car une thèse de doctorat sur ce sujet a été lancée en octobre 2017.
+Les travaux ont été présentés à la conférence ROADEF 2018 et seront présentés dans la conférence MOSIM 2018
+
+### Alain Haït
+
+Ingénieur en mécanique, Alain Haït possède une thèse de robotique préparée au LAAS CNRS.
+Enseignant-chercheur à l’INP Toulouse, il participé à la conception et la mise en place du département de génie industriel de l’ENSIACET.
+Chercheur visiteur au département de mathématiques et génie industriel de l’École Polytechnique de Montréal pendant deux ans, il a co-édité un livre sur les méthodes de planification et d’ordonnancement de la production avec contraintes liées aux ressources humaines, avec les professeurs François Soumis, Vincent Giard et Pierre Baptiste.
+Alain Haït est certifié CPIM de l’APICS.
+Spécialisé dans les méthodes de planification, il a encadré une thèse sur la planification de la maintenance d’hélicoptères civils.
+
+Professeur à l’ISAE-SUPAERO depuis 2007, il dirige depuis trois ans le département d’ingénierie des systèmes complexes (DISC) qui couvre les thèmes mathématiques appliquées, réseaux de communication, systèmes embarqués, ingénierie système et systèmes décisionnels.
+
+### Olga Battaïa
+
+Professeur à l'ISAE-SUPAERO et responsable de la formation en Génie Industriel, Olga Battaïa est une chercheuse internationalement reconnue dans le domaine de l'optimisation des systèmes industriels.
+Sa thèse de doctorat soutenue en 2007 à l’Ecole des Mines de Saint-Etienne a été primée du prix des meilleures thèses par le GdR MACS.
+Olga Battaïa est auteure de plus de 170 publications scientifiques, parmi lesquelles plus de 50 articles dans les revues internationales et 57 articles dans les proceedings des conférences internationales.
+Elle a participé à la réalisation de nombreux projets de recherche nationaux et internationaux.
+Elle a été invitée à plusieurs reprises par des universités étrangères pour donner des conférences scientifiques : Linköping University (2016), University of Siegen (2016), Politecnico di Milano, 2014 (Italie), Tshwane University of Technology, 2014 (Afrique du Sud), Lebanese University, 2013 (Liban) MIT, 2011 (USA), University of Michigan, 2011 (USA).
+En 2018, elle est devenue la première femme française nommée Membre Associé de la prestigieuse Académie Internationale de Recherche en Productique (CIRP).
+
+### Franco Peschiera
+
+Franco Peschiera a travaillé dans l'industrie de la recherche opérationnelle depuis près de 7 ans.
+Au cours de ce temps il a été le propriétaire et le directeur technique pour une startup espagnole (www.baobabsoluciones.es), en construisant des applications logicielles avec des modèles d'optimisation mathématique ou de simulation à l'intérieur pour plusieurs grandes entreprises espagnoles et internationales.
+
+Ces projets comportaient la liste suivante de tâches: la spécification des besoins du client, la construction d'un modèle mathématique ou de simulation, la programmation d'une interface utilisateur graphique, le déploiement et l'intégration avec les systèmes informatiques, la documentation et la formation des utilisateurs.
+
+Franco a commencé un doctorat à l'université ISAE-SUPAERO en octobre 2017 sur le sujet de la planification de la maintenance des avions militaires.
+
+## Motivations
+### Motivations
+
+Nous sommes déjà engagé dans une thèse sur le sujet de planification de maintenances des aéronefs pour la DGA.
+
+Le challenge OptiPlan va nous permettre avoir interaction directe avec la DGA pendant que on développe des meilleurs modèles pour avoir une meilleure compréhension du problème et résoudre des possibles doutes, ainsi que de recevoir de données de qualité pour évaluer nos modèles.
+
+Le challenge va nous permettre recevoir feedback de la part DGA sur les résultats qu'on obtienne. Le retourne d'information permettra ensuite améliorer encore plus les résultats.
+
+Finalement, le challenge permettra valider l'utilité des modèles et techniques employés dans une application complète et réel.
+
+## Expérience
+
+### Plannification logistique et RH
+
+**Clients**: Repsol et CLH.
+
+**Problematiques**:
+
+* Génération des horaires et affectation du personnel. 
+* Planification stratégique des réseaux de transport et production au niveau national.
+
+**Réponses**:
+
+* Logiciel de gestion de ressources humaines utilisé tous les jours.
+* Logiciel de optimisation mensuelle logistique.
+
+```{r, echo=FALSE, out.width='50%'}
+knitr::include_graphics('../../img/CLH.png')
+```
+
+### Plannification de la production
+
+**Clients**: General Motors, LI Lilly, Air Liquide, Repsol.
+
+**Problematiques**:
+
+* Planification de la production.
+* Détection de problèmes d'approvisionnement.
+
+**Réponses**:
+
+* Logiciel de simulation horaire de la production en temps réelle.
+* Logiciel optimisation de la production hebdomadaire.
+* Logiciel de planification mensuel de la production.
+
+```{r, echo=FALSE, out.width='50%'}
+knitr::include_graphics('../../img/Lilly.png')
+```
+
+### Logistique et transport
+
+**Clients**: Alstom, ASM et IO.
+
+**Problematiques**:
+
+* Simulation de trains et réseaux métro.
+* Optimisation de tournées de véhicules.
+
+**Réponses**:
+
+* Logiciel de simulation de réseaux de trains.
+* Logiciel d'optimisation de transports.
+
+```{r, echo=FALSE, out.width='50%'}
+knitr::include_graphics('../../img/IO.png')
+```
 
@@
 
-L'Armée de l'Air dispose d'une flotte d'avions Mirage 2000 de différents types, avec des fonctionnalités particulaires et avec un état d'opération spécifique donné par certains indicateurs temporelles d'usage.
+L’Armée de l’Air dispose d’une flotte d’avions Mirage 2000 de différents types, avec des fonctionnalités particulières et avec un état d’opération spécifique donné par certains indicateurs temporels d’usage.
 
-Au même temps, il existe un besoin opérationnel généré par les différents missions que la flotte doit être capable de remplir.
-Ce besoin peut être formulé comme un nombre de missions à satisfaire ou un budget d'heures de vol a atteindre à chaque période.
+En même temps, il existe un besoin opérationnel généré par les différentes missions que la flotte doit être capable de remplir.
+Ce besoin peut être formulé comme un nombre de missions à satisfaire ou un budget d’heures de vol a atteindre à chaque période.
 
-Finalement, il existe de contraintes opératifs du au fait que les avions ont besoin de maintenances régulières et possibles stockages pour être conservés en bon état et, de cette façon, garantir la capacité future de la flotte.
+Finalement, il existe des contraintes opérationnelles du au fait que les avions ont besoin de maintenances régulières pour être conservés en bon état et  garantir la capacité future de la flotte.
 
-Le problème consiste, donc, à affecter des avions aux missions prédéfinies et à programmer des opérations de maintenance préventive pour ces mêmes avions de façon d'avoir le plus grand quantité d'avions disponibles en lissant et réduisant les opérations de maintenance le plus possible dans le temps, sans impacter la capacité de la flotte présent ou future.
+Le problème consiste donc à affecter des avions aux missions prédéfinies et à programmer des opérations de maintenance préventive pour ces mêmes avions de façon à avoir la plus grand quantité d’avions disponibles en lissant et réduisant les opérations de maintenance le plus possible dans le temps, sans impacter la capacité présente ou future de la flotte.
 
 Dans la littérature, le problème est considéré comme un problème de planification de vol et de maintenance (Flight and Maintenance Planning problem).
-Travail a déjà été fait par @Kozanidis2008, @Cho2011, @Verhoeff2015, entre autres.
+Des recherches dans ce domaine ont été réalisées par @Kozanidis2008, @Cho2011, @Verhoeff2015 et @Chastellux2016, entre autres.
 
@@
 
-Il y a un ensemble $i \in \mathcal{I}$ des avions disponibles à être affectés aux missions.
-Chaque avion ne peut être affecté qu'à une seule mission au cours d'une période donnée.
-Ces avions souffrent d'usure et nécessitent des opérations de maintenance régulières pendant leur durée de vie.
-Le besoin de maintenance est calculé sur la base de deux indicateurs.
+Il y a un ensemble $i \in \mathcal{I}$ d’avions disponibles qui peuvent être affectés aux missions.
+Chaque avion ne peut être affecté qu’à une seule mission au cours d’une période donnée.
+Ces avions souffrent d’usure et nécessitent des opérations de maintenance régulières tout au long de leur durée de vie.
+Le besoin de maintenance est calculé sur la base de deux indicateurs:
 
-Le premier est appelé "temps écoulé restant" (ou $ret_{it}$).
-Il exprime la durée (mesurée en périodes) après laquelle l'aéronef ne peut plus être utilisé et doit subir une opération de maintenance.
+* Le temps écoulé restant ($ret_{it}$): durée (mesurée en périodes) après laquelle l'aéronef ne peut plus être utilisé et doit subir une opération de maintenance.
 Sa valeur est calculée pour chaque avion $i$ et chaque période de temps $t$.
-De la même manière, "temps d'utilisation restant" (ou $rut_{it}$) est utilisé pour mesurer la quantité de temps que l'avion $i$ peut être utilisé avant d'avoir besoin d'une opération de maintenance à une période donnée $t$.
+* Le temps d’utilisation restant ($rut_{it}$): durée d’utilisation d’un avion $i$ avant d’avoir besoin d'une opération de maintenance à une période donnée $t$.
 
-De plus, après une quantité absolue de temps et / ou d'utilisation ($aet_i$ ou $aut_i$), l'aéronef devient obsolète. Il n'y a aucun moyen d'inverser ce processus.
+De plus, après une certaines durée ou temps d’utilisation ($aet_i$ or $aut_i$), l’aéronef devient obsolète. Il n’y a aucun moyen d’inverser ce processus.
 
-À toute période donnée, y compris au début de l'horizon de planification, chaque aéronef a un statut spécifique donné par le temps d'utilisation restant et le temps écoulé restant.
+À toute période donnée, y compris au début de l’horizon de planification, chaque aéronef a un statut spécifique donné par le temps d’utilisation restant et le temps écoulé restant.
 
@@
 Il y a un ensemble fixe de $j \in \mathcal{J}$ missions à accomplir sur un horizon de temps divisé en $t \in \mathcal{T}$ périodes discrètes.
-Pour leur exécution, ces missions nécessitent l'attribution d'un nombre spécifique d'aéronefs $R_j$ chaque période de temps la mission est active.
-Les périodes de début et de fin de chaque mission sont connues et une mission est considérée comme active dans les mois entre sa période de début et sa fin.
+Pour leur exécution, ces missions nécessitent l’attribution d’un nombre spécifique d’aéronefs $R_j$ à chaque période active (c’est-à-dire entre le début et la fin de la mission).
+Les périodes de début et de fin de chaque mission sont connues.
 
-Au cours de chaque période, les missions consomment une quantité de temps égale à $H_j$ heures de chacun de ses avions affectés.
+Au cours de chaque période, les missions consomment une quantité de temps égale à $H_j$ heures de chacun des avions qui lui sont affectés.
 
-L'affectation d'un avion à une mission n'est pas décidée pour toute la durée de la mission.
-Après un minimum de temps $MT_j$, un aéronef peut être libéré et échangé par un autre, même si la mission à laquelle il est affecté n'a pas terminée.
-Le nombre total d'avions affectés à un moment donné à chaque mission doit toujours être égal à $R_j$.
+Chaque mission nécessite un et un seul type d’aéronef qui, de plus, doit satisfaire à des exigences de mission supplémentaires.
 
-Chaque mission nécessite un et un seul type d'aéronef qui, de plus, doit satisfaire à des exigences de mission supplémentaires.
+-----
 
@@
 
-    Graphique montrant possibles missions et ses besoins.
+    Exemples de missions avec leurs besoins en aéronefs.
 
@@
 
-Les opérations de maintenance sont le processus par lequel les aéronefs qui ont atteint une limite dans certains indicateurs peuvent revenir à un état où ils peuvent continuer à être utilisés dans les missions.
+Les opérations de maintenance sont le processus par lequel les aéronefs qui ont atteint une limite dans certains indicateurs peuvent revenir à un état où ils peuvent à nouveau être utilisés en mission.
 
@@
 
-Après une opération de maintenance, un aéronef restaure le temps écoulé restant et le temps d'utilisation restant à leurs valeurs maximales $E$ (~ 60 mois) et $H$ (~ 1200 heures) respectivement.
+Après une opération de maintenance, un aéronef restaure le temps écoulé restant et le temps d’utilisation restant à leurs valeurs maximales respectives $E$ (~ 60 mois) et $H$ (~ 1200 heures).
 
-Il existe également des opérations de maintenance spéciales qui améliorent les capacités de l'avion (ils améliorent leur capacité à être affecté à de nouvelles missions).
-Ces opérations seront décidées par l'utilisateur et fixées pour l'application afin qu'elles puissent être prises en compte lors de l'attribution du reste des opérations de maintenance.
+Il existe également des opérations de maintenance spéciales qui améliorent les capacités de l’avion (ils améliorent leur capacité à être affectés à de nouvelles missions).
+Ces opérations seront décidées par l’utilisateur et fixées dans l’application afin qu’elles puissent être prises en compte lors de l’attribution du reste des opérations de maintenance.
 
@@
 
-Conformément à la règle du temps écoulé, même si un aéronef n'est pas utilisé, il doit encore subir un opération de maintenance après un certain temps.
-Afin d'éviter ce problème, l'aéronef peut être mis dans un état de stockage.
+Conformément à la règle du temps écoulé, même si un aéronef n’est pas utilisé, il doit encore subir une opération de maintenance après un certain temps.
+Afin d’éviter ce problème, l’aéronef peut être mis dans un état de stockage.
 
@@
 
-Chaque avion a une certaine capacité d'être stocké et cette capacité est mesurée en un nombre de périodes $sc$.
-Pour qu'un aéronef restaure sa capacité de stockage restante, il doit recevoir une opération de maintenance spécifique où l'avion récupère sa capacité de stockage jusqu'à un certain niveau $S$.
+Chaque avion a une certaine limite de temps de stockage.
+Cette limite est mesurée en nombre de périodes $sc$.
+Pour qu’un aéronef restaure sa limite de stockage à son maximum $S$, il doit recevoir une opération de maintenance spécifique.
 
@@
 
-* Assigné à une mission.
-* Réception de maintenance
-* Sous stockage
-* Disponible.
+- en mission
+- en maintenance
+- en stockage
+- disponible
 
@@
 
-Lors de la planification des missions et des opérations de maintenance, il est important de prendre en compte l'état initial et final de chaque aéronef.
+Lors de la planification des missions et des opérations de maintenance, il est important de prendre en compte l’état initial et final de chaque aéronef.
 Cet état initial peut être une maintenance ou une mission assignée.
-Si un aéronef est déjà en maintenance, il doit continuer dans cet état pour son temps de maintenance restant.
+Si un aéronef est déjà en maintenance, il doit continuer dans cet état pendant le reste de sa durée de maintenance.
 Les missions affectés doivent être prises en compte de la même manière.
 
-Pour les opérations de maintenance planifiées dont l'heure de début est proche du début de l'horizon de planification, elles doivent également être considérées comme fixes.
+Pour les opérations de maintenance planifiées dont l’heure de début est proche du début de l’horizon de planification, elles doivent également être considérées comme fixes.
 
-Les temps (écoulé et d'utilisation) restants sont assignés à chaque aéronef au début de l'horizon de planification.
+Les temps (écoulé et d’utilisation) restants sont assignés à chaque aéronef au début de l’horizon de planification.
 
-Enfin, l'état de chaque aéronef à la fin de l'horizon de planification, ses temps restant (écoulé, utilisé), doivent être défini et pris en compte.
+Enfin, l’état de chaque aéronef à la fin de l’horizon de planification, ses temps restant (écoulé, utilisé), doivent être défini et pris en compte.
 
@@
 
-Les objectifs suivants sont considérés, bien que d'autres pourraient potentiellement être ajoutés.
+Deux objectifs principaux ont été considérés  bien que d’autres pourraient potentiellement être ajoutés.
 
-Étant donné que la possibilité de nouvelles missions et la durée de l'opération de maintenance sont considérées stochastiques dans la vie réelle, un objectif fondamental est de maximiser la robustesse de la planification en ayant la plus grande quantité d'avions disponibles à chaque période de l'horizon de planification.
+Étant donné que le déclenchement de nouvelles missions et la durée des opérations de maintenance sont en partie imprévisible dans la vie réelle, un objectif fondamental est de maximiser la robustesse de la planification en ayant la plus grande quantité d’avions disponibles à chaque période de l’horizon de planification.
 
-Compte tenu de la capacité limitée pour faire opérations de maintenance et de son coût, un autre objectif consiste à lisser le plus possible le nombre d'avions sous maintenance dans l'horizon de planification.
+Compte tenu de la capacité limitée pour réaliser des opérations de maintenance et de leurs coûts, un autre objectif consiste à lisser le nombre d’avions en maintenance sur l’horizon de planification.
 Les opérations de maintenance doivent également être contrôlées afin de ne pas dépasser la capacité maximale.
@@
 ## Proposition
+
+### Résumé
+
+Le projet présenté est une application web interactive.
+Il utilise un modèle mathématique pour générer de nouvelles solutions, permet à différents utilisateurs d'y accéder avec différents droits d'accès et offre une puissante visualisation pour afficher, comparer et analyser les planifications.
+Il s'intègre aux sources de données existantes (fichier Excel), aux nouvelles proposées (modèle Excel) et a la possibilité de s’intégrer aux sources existantes (autres systèmes internes) ainsi qu'à la sortie simple (csv, json).
+Enfin, il permet l'utilisation de cas pour stocker des informations, ce qui facilite les essaies, l'expérimentation et la comparaison de différents scénarios.
 
@@
 
-Chaque mois, toutes les missions doivent être satisfaites grâce à l'affectation des suffisamment d'heures de vol à chaque avion.
+Chaque mois, toutes les missions doivent être satisfaites grâce à l’affectation de suffisamment d’heures de vol à chaque avion.
 
@@
 
-Chaque mois, toutes les missions doivent avoir un nombre minimum d'avions affectés.
+Chaque mois, toutes les missions doivent avoir un nombre exact d’avions affectés.
 
@@
 
-Chaque avion doit toujours avoir suffisamment de "temps d'utilisation restant" pour pouvoir effectuer les missions requises.
+Chaque avion doit toujours avoir suffisamment de “temps d’utilisation restant” pour pouvoir effectuer les missions requises.
 Afin de récupérer des heures de vol (environ 1200 heures), une opération de maintenance est nécessaire.
+Après un certain temps, les aéronefs ont besoin d’un entretien même s’ils n’ont pas atteint leur limite d’heures de vol. Ce temps est d’environ 5 ans.
 
-De même, après un certain temps, les aéronefs ont besoin d'un entretien même s'ils n'ont pas atteint cette limite d'heures de vol. Cette durée est d'environ 5 ans.
+-----
 
@@
 
-Les opérations de maintenance prennent environ 5 mois et une fois qu'un avion a commencé l'opération de maintenance, il doit la continuer.
-Il ne peut pas être affectés aux missions pendant la durée de la opération de maintenance.
+Les opérations de maintenance prennent environ 5 mois et une fois qu’un avion a commencé l’opération de maintenance, il doit la terminer.
+Il ne peut pas être affectés aux missions pendant la durée de l'opération.
 
@@
 
-Chaque mois, il est important d'avoir le plus grand nombre possible d'avions disponibles prêts à s'engager dans une nouvelle mission.
+Chaque mois, il est important d’avoir le plus grand nombre possible d’avions disponibles prêts à s’engager dans une nouvelle mission.
 
@@
 
-Les opérations de maintenance à planifier pour tous les avions doivent être effectuées de la manière la plus lissé possible, en évitant avoir des mois spécifiques avec un grand quantité des opérations de maintenance.
+Les opérations de maintenance à planifier pour tous les avions doivent être effectuées de la manière la plus lissée possible, en évitant d'avoir des mois spécifiques avec une grande quantité d’opérations de maintenance.
 
-En d'autres termes, il est important d'avoir le nombre le plus bas possible d'avions sous maintenance à chaque mois.
+En d’autres termes, il est important d’avoir le moins d’avions possible en maintenance chaque mois.
+
+-----
 
@@
 
-À la fin de l'horizon de planification, l'avion doit être dans le meilleur état possible.
-Cela implique que, en moyenne, les avions doivent avoir un gros budget d'heures de vol restant (temps d'utilisation restant) et un grand nombre de mois (temps écoulé restant) avant d'avoir besoin d'un entretien.
+À la fin de l’horizon de planification, chaque avion doit être dans le meilleur état possible.
+Cela implique qu’en moyenne, les avions doivent avoir si possible un budget important d’heures de vol et de mois restants.
 
@@
 
-Le stockage peut être décidé par le modèle afin d'éviter la détérioration de l'aéronef.
-Le nombre de périodes minimale de stockage d'un avion est de 6 mois et, dans cet état, l'avion n'est pas disponible ni pour recevoir un maintenance ni pour faire une mission.
+Le stockage peut être décidé par le modèle afin d’éviter la détérioration de l’aéronef.
+Le nombre minimal de périodes de stockage d’un avion est de 6 mois et, dans cet état, l’avion n’est pas disponible ni pour recevoir une maintenance ni pour réaliser une mission.
+
+-----
 
@@
 
-Il existe plusieurs raisons pour lesquelles l'utilisateur peut avoir besoin de fixer une partie d'une planification précédente.
-Par exemple, certaines opérations de maintenance sont déjà décidées et doivent avoir lieu à l'avenir dans de dates spécifiques.
-Certains avions sont déjà en mission ou sous maintenance et doivent y rester au moins quelques mois dans cet état.
+Il existe plusieurs raisons pour lesquelles l’utilisateur peut avoir besoin de fixer une partie d’une planification précédente.
+Par exemple, certaines opérations de maintenance sont déjà décidées et doivent avoir lieu à l’avenir à des dates spécifiques.
+Certains avions sont déjà en mission ou en maintenance et doivent y rester.
 En outre, le stockage peut être déjà décidé ou en cours.
 
-Une autre raison de la fixation des opérations de maintenance pourrait être pour les opérations d'amélioration planifiées aux aéronefs.
-Ces types spéciaux d'opérations de maintenance sont décidés plusieurs mois à l'avance pour certains avions sélectionnés.
+Une autre raison de la fixation des opérations de maintenance pourrait être une opération d’amélioration planifiée sur un aéronef.
+Ce type d’opérations de maintenance est décidé plusieurs mois à l’avance pour certains avions sélectionnés.
 
-Ces assignations fixes seront communiquées au optimiseur afin qu'il ne les change pas.
+Ces assignations fixes seront communiquées au modèle d'optimisation afin qu’il ne les change pas.
+
+-----
 
@@
 
-<!-- rut, ret -->
-* **Temps maximum**: dans le cas où une solution est nécessaire avant de prouver l'optimalité.
-* **Maximal GAP**: afin d'autoriser une distance maximal entre la meilleure solution obtenue et la meilleure solution hypothétique possible.
+* **Temps maximum**: dans le cas où une solution est nécessaire rapidement sans prouver l’optimalité.
+* **Gap maximum**: afin d’autoriser une distance maximale entre la meilleure solution obtenue et la meilleure solution hypothétique possible.
 Cette distance peut être une différence proportionnelle ou une différence absolue.
-* **Premier et dernier mois a planifier**: afin de définir l'horizon de planification à planifier.
-* **Poids des objectifs**: pour optimiser les différents objectifs et privilégier un ou autre objectif.
-* **Paramètres de maintenance**: durée de maintenance (en mois), temps d'utilisation maximal, temps écoulé maximal et d'autres paramètres liés à la maintenance.
+* **Premier et dernier mois à planifier**: afin de définir l’horizon de planification.
+* **Poids des objectifs**: pour optimiser les différents objectifs et privilégier un objectif ou un autre.
+* **Paramètres de maintenance**: durée de maintenance (en mois), temps d’utilisation maximal, temps écoulé maximal et autres paramètres liés à la maintenance.
 
@@
 
-Étant donné la nature stratégique de l'application, elle sera centrée sur les cas.
-Cela signifie qu'il permettra de sauvegarder toutes les données pertinentes (entrée et sortie) dans le même cas.
-De cette façon, différentes expériences pourraient être faites en copiant et en modifiant un cas existant.
-Ceci est similaire à travailler avec plusieurs feuilles de calcul, chacune avec son propre nom, tout en partageant la même logique en un seul endroit.
+Étant donné la nature stratégique de l’application, elle sera centrée sur des études de cas.
+Cela signifie qu’il permettra de sauvegarder toutes les données pertinentes (entrée et sortie) dans un même ensemble ou cas.
+De cette façon, différentes expériences pourront être réalisé en copiant et en modifiant un cas existant.
+La gestion de cas est similaire à l'utilisation de plusieurs feuilles de calcul pour étudier plusieurs scénarios différents.
 
-La figure suivante montre un ensemble hypothétique de cas déjà enregistrés.
-Elle représente les interactions possibles entre les cas.
-Le cas 1 a été publié pour le mois de mars.
-Au cours du mois, il a reçu quelques modifications, ce qui a entraîné les cas 2 et 3.
-Enfin, le cas 3 a été re-publié pour le mois de mars.
-Plus tard, le cas 4 a été créé et modifié sans être publié dans le cas 5.
-Le cas 5 a été publié en avril.
+La figure suivante montre un exemple de cas déjà enregistrés. Elle représente les interactions possibles entre les cas pendant une période de quelques mois.
+
+1. Le cas 1 a été publié pour le mois de mars.
+2. Au cours du mois, il a reçu quelques modifications, ce qui a engendré les cas 2 et 3.
+3. Enfin, le cas 3 a été re-publié pour le mois de mars.
+4. Plus tard, le cas 4 a été créé et modifié sans être publié.
+5. Finalement, Le cas 5 a été crée à partir du cas 4 et publié en avril.
+
+-----
 
@@
 
-    Exemple montrant l'interaction entre les cas et la publication des cas.
+    Exemple montrant la création et publication de les cas.
 
@@
 
-#### Contenu d'une cas
+-----
 
-Un cas consiste, au même temps, de toutes les informations nécessaires pour construire une planification et une planification réalisable construite à partir de cette information.
+#### Contenu d'un cas
 
-Cette unité de stockage permet de s'intégrer facilement aux structures de données actuelles.
+Un «cas» est une unité de stockage qui permet de s’intégrer facilement aux structures de données actuelles. Il contient toutes les informations nécessaires pour construire une planification:
 
-Plus explicitement, un cas consiste en:
+* L’ensemble d’avions avec leurs caractéristiques et leurs informations historiques jusqu’à une certaine date.
+* L’ensemble de missions ou d’heures de vol à réaliser dans la période de planification.
+* Une pre-affectation d’aéronefs à des opérations pendant l’horizon de planification.
+* Les futurs calendriers de maintenance.
+* La configuration : la durée de maintenance, les heures de vol après une maintenance, etc.
+* Une planification (générée manuellement ou avec l'application).
 
-* L'ensemble d'avions avec leurs caractéristiques et leurs informations historiques jusqu'à une certaine date.
-* L'ensemble de missions ou d'heures de vol à faire dans la période de planification.
-* Une pre-affectation d'aéronefs à des opérations pendant l'horizon de planification.
-* Futurs calendriers de maintenance.
-* Configuration comme la durée de maintenance, les heures de vol après une maintenance, etc.
-* Une planification (généré manuellement ou avec l'Optimiseur).
+-----
 
@@
 Cette fonctionnalité peut potentiellement avoir plusieurs utilisations en fonction des besoins finaux du client.
-Une option serait de visualiser les différences dans les données d'entrée pour deux cas.
+Une option pourrait être de visualiser les différences dans les données d’entrée pour deux cas.
 Une autre option consisterait à visualiser les différences dans la planification résultante pour les deux cas.
@@
 
+-----
+
 #### Publications de cas
 
-Une situation courante dans la création de planifications est le fait que plusieurs cas différents peuvent exister mais il est toujours important d'identifier lequel est le cas 'actif'.
-Pour faire face à cela, un système de publication simple sera inclus dans l'application, où le «planificateur» (voir la section "Utilisateurs") peut identifier le cas qui devrait être partagé avec d'autres.
+Une situation courante dans la création de planifications est le fait que plusieurs cas différents peuvent exister mais il est toujours important d’identifier lequel est le cas ‘actif’.
+Pour faire face à cela, un système de publication simple sera inclus dans l’application, où le «planificateur» (voir la section “Utilisateurs”) peut identifier le cas qui devra être partagé avec les autres utilisateurs.
 
@@
 
-Toute interaction de l'utilisateur avec l'application se fera via une application HTML.
+Toute interaction de l’utilisateur avec l’application se fera via une application HTML.
+En plus de montrer des données dans des tableaux interactifs, des alternatives graphiques spécifiques peuvent être incorporées pour rendre l’utilisation plus facile et plus efficace.
 
-En plus de montrer des données dans des tableaux interactifs, des alternatives graphiques spécifiques peuvent être incorporées pour rendre l'utilisation plus facile et plus efficace.
-
-Voici quelques exemples de fonctionnalités disponibles.
-
-Ici, un gantt interactif est utilisé pour présenter et éventuellement modifier un plan de maintenance.
-Il est possible d'éditer les opérations de maintenance et l’affection aux missions (changer la taille ou la déplacer) en double-cliquant dessus et en faisant glisser.
-En outre, il est possible d'ajouter de nouvelles affectations en double-cliquant.
+Ici, un Gantt interactif est utilisé pour présenter et éventuellement modifier un plan de maintenance.
+Il est possible d’éditer les opérations de maintenance et l’affection aux missions (changer la taille ou la déplacer) en double-cliquant dessus et en faisant glisser.
+En outre, il est possible d’ajouter de nouvelles affectations en double-cliquant.
 
@@
 
-Ici, un graphique interactif est utilisé pour présenter l'évolution du statut d'un seul avion au fil du temps. Il est possible de filtrer, faire zoom et exporter en tant qu'image.
+-----
+
+Ici, un graphique interactif est utilisé pour présenter l’évolution du statut d’un seul avion au fil du temps. Il est possible de filtrer, zoomer et exporter en tant qu’image.
 
@@
     Exemple de graphique interactif.
+
+-----
 
@@
 
-L'application inclura son propre stockage de données à travers d'une base de données.
-Toutes les informations utilisées par l'application seront stockées ici.
-En plus de ce système, l'application permettra plusieurs façons d'importer et d'exporter des données à partir de ce système.
+L’application inclura son propre stockage de données à travers une base de données.
+Toutes les informations utilisées par l’application seront stockées ici.
+En plus de ce système, l’application permettra plusieurs façons d’importer et d’exporter des données à partir de l'application.
 
@@
 
+-----
+
 #### Exporter des données
@@
 
-* Dans le format d l'Excel actuel (ou un format similaire).
+* Dans le format Excel actuel (ou un format similaire).
 * Dans le nouveau template Excel qui sera créé.
-* Fichiers csv et / ou json si nécessaire.
-* Ces capacités d'exportation permettront également de partager des informations partielles avec d'autres acteurs dans des fichiers Excel ou plats.
+* Dans un fichier csv et / ou json si nécessaire.
+
+Ces capacités d’exportation permettront également de partager des informations partielles avec d’autres acteurs dans les formats mentionnés précédemment.
 
@@
 
-L'application sera déployée via une architecture client-serveur.
-De cette façon, il serait facile de partager l'application étant donné que cela nécessitera seulement l'installation d'un navigateur web moderne (Google Chrome).
+L’application sera déployée via une architecture client-serveur.
+De cette façon, il sera facile de partager l’application étant donné que cela nécessitera seulement l’installation d’un navigateur web moderne (Google Chrome).
 
-Le serveur devra être installé quelque part dans l'infrastructure du client.
-
-La configuration matérielle minimale requise pour le serveur est la suivante:
-
-* Ubuntu Server 64 bits.
-* Noyaux> = 4
-* RAM> = 8 Go
-* Disque dur> = 500 Go
-
-Les exigences logicielles minimales pour le serveur sont les suivants:
-
-* R + Shiny Server
-* python> = 3.6
-* PostgreSQL
+Le serveur devra être installé quelque part dans l’infrastructure du client.
 
 Des logiciels supplémentaires pourraient être nécessaires en fonction des spécifications finales.
-Par exemple, dans le cas où le modèle soit trop gros pour les solveurs de source libre, un solveur commercial pourrait être nécessaire.
+Par exemple, si le modèle est trop gros pour les solveurs en logiciel libre, un solveur commercial pourrait être nécessaire.
 
-En fonction de la future intégration, des flux de données supplémentaires seront inclus dans la conception. Le seul flux de données connu se situe entre le fichier Excel existant et l'application.
+En fonction de la future intégration, des flux de données supplémentaires seront inclus dans la conception.
+Le seul flux de données confirmé pour le moment se situe entre le fichier Excel existant et l’application.
 
-La figure suivante présente un résumé de l'architecture à mettre en œuvre et des flux de données entre les composants.
+La figure suivante présente un résumé de l’architecture à mettre en œuvre et des flux de données entre les composants.
+Les deux types d’utilisateurs vont communiquer avec l’application pour générer les plannings ou les voir.
+L’application a son propre moteur (modèle) pour générer des solutions et son propre stockage (base de données).
+Les technologies sont mentionnées sur chaque arc pour indiquer là où elles seront utilisées.
+
+-----
 
@@
 
-    Architecture proposée avec flux de données entre composants.
+    Architecture proposée avec flux de données.
+
+-----
 
 Les utilisateurs interagiront via les navigateurs Web avec le serveur.
-Pour charger de nouvelles données, ils peuvent charger un nouveau fichier Excel avec les données pertinentes ou éditer un cas existant via l'application.
+Pour charger de nouvelles données, ils pourront charger un nouveau fichier Excel avec les données pertinentes ou éditer un cas existant via l’application.
 
@@
 
-Les utilisateurs seront en mesure de récupérer visuellement les résultats. En outre, ils seront en mesure d'exporter les résultats en formats csv, Excel et json en fonction de leurs besoins.
+Les utilisateurs seront en mesure de visualiser les résultats. En outre, ils seront en mesure d’exporter les résultats en formats csv, Excel et json en fonction de leurs besoins.
+
+-----
 
@@
 
-* Des graphiques puissants peuvent être incorporés via des bibliothèques javascript et HTML.
-* Il est plus facile à maintenir car une séparation claire des fonctionnalités est faite entre l'interface graphique, la logique d'utilisation et l'Optimiseur.
-* Aucune installation de logiciel additionnel à un navigateur Web moderne n'est requise.
-* Plusieurs utilisateurs peuvent être facilement incorporés.
-* L'intégration avec d'autres systèmes peut être effectuée avec des changements minimes à la structure existante.
-* Des ressources puissantes peuvent être allouées si nécessaire au niveau du serveur au lieu des devoir changer les ordinateurs des utilisateurs.
-* En plus de cela, il n'y a pas de logiciel tiers à acheter.
+* Des graphiques puissants pouvant être incorporés via des bibliothèques javascript et HTML.
+* Une facilité de maintenance liée à une séparation claire entre les fonctionnalités de l’interface graphique, de la logique d’utilisation et du modèle d'optimisation.
+* Un navigateur Web moderne comme seul besoin logiciel.
+* Facilité pour incorporer plusieurs utilisateurs.
+* L’intégration avec d’autres systèmes peut être effectuée avec des changements minimes à la structure existante.
+* Possibilité d'allouer des ressources supplémentaires au niveau du serveur au lieu de devoir changer les ordinateurs des utilisateurs.
+* Absence de logiciel tiers à acheter.
 Tous les composants sont open source.
-La seule exception possible pourrait être le solveur mathématique, en fonction de la taille du problème et les besoins de performance.
+La seule exception possible pourrait être le solveur mathématique, en fonction de la taille du problème et des besoins en performance.
 
@@
 
-Le premier s'appellera "planificateur".
+Le premier s’appellera “planificateur”.
 Cet utilisateur est en charge de créer (manuellement ou automatiquement) des plannings différents.
-Le second s'appellera "spectateur".
-Cet utilisateur pourra visualiser les résultats de la planification créée et stockée dans l'application.
-En fonction des besoins du client, plusieurs profils "spectateur" pourraient être créés afin de partager quelques parties des informations stockées avec chaque profil.
-
-## Équipe
-
-### Alain Haït
-
-### Olga Battaïa
-
-### Franco Peschiera
-
-Un ingénieur industriel, mathématicien de profession et programmeur par cœur.
-
-Franco a travaillé dans l'industrie de la recherche opérationnelle depuis près de 7 ans.
-Au cours de ce temps il a été le propriétaire et le directeur technique pour une startup espagnole (www.baobabsoluciones.es), en construisant des applications logicielles avec des modèles d'optimisation mathématique ou de simulation à l'intérieur pour plusieurs grandes entreprises espagnoles et internationales.
-Exemples de clients sont: Repsol et CLH (logistique pétrolière, marketing, ressources humaines), General Motors, LI Lilly et Air Liquide (planning de production), RATP, Alstom et ASM (logistique et transport) parmi plusieurs autres.
-
-Ces projets comportaient la liste suivante de tâches: la spécification des besoins du client, la construction d'un modèle mathématique ou de simulation, la programmation d'une interface utilisateur graphique, le déploiement et l'intégration avec les systèmes informatiques, la documentation et la formation des utilisateurs.
-
-Franco a commencé un doctorat à l'université ISAE-SUPAERO en octobre 2017 sur le sujet de la planification de la maintenance des avions militaires.
+Le second s’appellera “spectateur”.
+Cet utilisateur pourra visualiser la planification créée et stockée dans l’application.
+En fonction des besoins du client, plusieurs profils “spectateur” pourraient être créés afin de partager certaines parties des informations stockées avec chaque profil.
 
@@
 
-Ici, les tâches requises avec leur ordre sont présentées. Le principal risque sur la planification actuelle est l'information incomplète sur:
+Ici, les tâches requises avec leur ordre sont présentées. Le principal risque sur la planification actuelle est l’information incomplète sur:
 
 * le problème spécifique à résoudre.
-* l'infrastructure informatique disponible du client.
-* la disponibilité des utilisateurs potentiels pour discuter des exemples de solutions et tester et évaluer les fonctionnalités faits.
+* l’infrastructure informatique disponible du client.
+* la disponibilité des utilisateurs potentiels pour discuter des exemples de solutions et tester et évaluer les fonctionnalités.
 
-On présente une proposition de planification basé en expériences précédents. La planification dépendra de la voie de communication finalement choisi et la disponibilité du client.
+On présente une proposition de planification basée sur les expériences précédentes. La planification dépendra de la voie de communication finalement choisie et de la disponibilité du client.
 
@@
 
-* un document avec une description accordé sur les exigences minimales pour le prototype et
-* un document d'exigences additionnelles à préciser ultérieurement.
+* un document avec une description validée par le client des exigences minimales du prototype
+* un document d’exigences additionnelles à préciser ultérieurement.
 
@@
 
-* la liste avec des exigences minimales.
-* une alternative de déploiement pour partager l'application avec les utilisateurs.
+* une liste d'exigences minimales.
+* une possibilité de déploiement pour partager l’application avec les utilisateurs.
 
 Le livrable de cette phase sera:
-
 * une application prototype que les utilisateurs peuvent utiliser et tester.
-* exemples de solutions pour les utilisateurs à analyser.
+* des exemples de solutions à analyser par les utilisateurs.
 
@@
 Dans cette phase, une version finale du prototype sera construite et déployée.
-En fonction de niveau d'implication du client, le types de changements à réaliser et l'urgence, cette phase pourrait être coupé en deux phases de la moitié de durée.
+En fonction du niveau d’implication du client, du types de changements à réaliser et de l’urgence, cette phase pourrait être coupée en deux sous phases avec des livrables distincts.
 
@@
 
-* une liste avec les changements à faire à la première version basée sur les commentaires des utilisateurs.
-* une liste avec des fonctionnalités supplémentaires à ajouter.
+* une liste des changements à réaliser sur la première version basée sur les commentaires des utilisateurs.
+* une liste de fonctionnalités supplémentaires à ajouter.
 
@@
 
-* une nouvelle version de l'application que les utilisateurs pourront utiliser.
+* une nouvelle version de l’application que les utilisateurs pourront utiliser.
 * la documentation technique.
-* la Guide de l'utilisateur.
+* la Guide de l’utilisateur.
 
-La durée de cette phase est estimée à: *5 semaines*.
+La durée de cette phase est estimée à: *5 semaines* .
 
@@
 
-Dans le tableau ci-dessous sont les engagements hebdomadaires estimés de profils:
+Dans le tableau ci-dessous sont les engagements hebdomadaires estimés des profils:
 
@@
 
-Basé en la définition de profils et les engagements déjà présente, le montant correspondant au projet de développement du prototype est de 55.600€.
+Basé sur la définition de profils et des engagements déjà présentés, le montant correspondant au projet de développement du prototype est de 55.600€.
 
@@
 * **Modélisation mathématique**: pour générer des planifications.
-    * **Solveur mathématique**: pour générer les planifications. Cela pourrait être CBC, GUROBI ou CPLEX en fonction de besoins du modèle.
-* **Python**: pour la modélisation, l'heuristique et l'intégration de données.
+    * **Solveur mathématique**: pour générer les planifications. Cela pourrait être CBC, GUROBI ou CPLEX en fonction des besoins du modèle.
+* **Python**: pour la modélisation, l’heuristique et l’intégration de données.
 * **R**: pour les statistiques, possibles prévisions et la visualisation des données.
@@
 
-La modélisation mathématique est une technique qui fait partie de la branche de la recherche opérationnelle en mathématiques.
-Il permet la construction d'un ensemble d'équations pour représenter une situation spécifique bien définie et il permet l'utilisation de logiciels de avancés pour rechercher et trouver la meilleure solution qui respecte ces équations.
-Ces logiciels (modélisateurs et solveurs) sont disponibles à la fois en version commerciale et open source (gratuite).
+La modélisation mathématique est une technique qui permet la construction d’un ensemble d’équations pour représenter une situation spécifique bien définie et communique avec les logiciels avancés pour rechercher et trouver la meilleure solution qui respecte ces équations.
+Ces logiciels (modélisateurs et solveurs) sont disponibles à la fois en version commerciale et libre (gratuite).
 
-Les avantages de l'utilisation de cette technique par rapport à d'autres alternatives sont qu'elle peut prouver l'optimalité d'une solution, elle peut facilement incorporer plusieurs objectifs et il est possible de fixer facilement certaines décisions déjà prises, rendant possible l'utilisation interactive du modèle par l'utilisateur.
-Avec peu de modifications, le modèle peut aussi être réutilisé pour prendre en compte différentes situations qui ne soient si différentes de la situation originelle.
+Les avantages de l’utilisation de cette technique par rapport à d’autres alternatives sont qu’elle peut prouver l’optimalité d’une solution, qu'elle peut facilement incorporer plusieurs objectifs et qu'il est possible de fixer facilement certaines décisions déjà prises, rendant possible l’utilisation interactive du modèle par l’utilisateur.
+Avec peu de modifications, le modèle peut aussi être réutilisé pour prendre en compte différentes situations similaires à la situation originelle.
 
-D'autres fonctionnalités incluent également des limites de temps et d'autres conditions d'arrêt plus complexes.
+D’autres fonctionnalités incluent également des limites de temps et d’autres conditions d’arrêt plus complexes.
 
@@
 
-Python permet le prototypage rapide ainsi qu'une bonne syntaxe de code.
-Il offre une extrêmement large ensemble des bibliothèques et, en particulier, il inclut les meilleures API pour les solveurs commerciaux et non commerciaux pour les problèmes d'optimisation.
+Python permet le prototypage rapide ainsi qu’une bonne syntaxe de code.
+Il offre un extrêmement large ensemble de bibliothèques et, en particulier, il inclut les meilleures API pour les solveurs commerciaux et non commerciaux pour les problèmes d’optimisation.
 
-Il excelle également au développement de Web avec des bibliothèques très connus.
+Il excelle également au développement  Web avec des bibliothèques très connus.
 
@@
 
-R est l'état de l'art en matière de statistiques et de modélisation prévisionnelle.
-Il comprend plusieurs bibliothèques pour la manipulation des données, la visualisation et, en tant que langage de programmation, une bonne intégration avec d'autres langages et logiciels.
+R est l’état de l’art en matière de statistiques et de modélisation prévisionnelle.
+Il comprend plusieurs bibliothèques pour la manipulation des données, la visualisation et, en tant que langage de programmation, une bonne intégration avec d’autres langages et logiciels.
 
@@
 
-PostgreSQL est l'une des alternatives de base de données open-source les plus connues.
-Il comprend nombreuses fonctionnalités qui ne sont pas présentes dans d'autres moteurs de base de données (à savoir MySQL) tels que:
+PostgreSQL est l’une des alternatives de base de données open-source les plus connues.
+Il comprend nombreuses fonctionnalités qui ne sont pas présentes dans d’autres moteurs de base de données (à savoir MySQL) tels que:
 
-* NoSQL stockage.
+* stockage NoSQL.
 * Fonctions de fenêtre intégrées à R.
@@
 
-## Avertissements
+## Remarques
+### Remarques
 
-1. La proposition présentée ci-dessus est destinée à une application complète livrée dans le cadre d'un projet complet au client.
-1.
-Le prototype à livrer n'inclura pas toutes les fonctionnalités mentionnées dans ce document mais montrera des exemples des fonctionnalités les plus importantes suite à une réunion avec les futurs utilisateurs et décideurs.
-1. L'intégration avec d'autres systèmes est hors de la portée de cette proposition car plus d'informations sont nécessaires concernant les systèmes existants pour pouvoir estimer la charge de travail.
-1.
-Il est supposé que le client fournira les exigences hardware ainsi que les droits d'accès et les autorisations pour installer tous les composants du logiciel.
+1. La proposition présentée ci-dessus est destinée à une application complète livrée dans le cadre d’un projet complet au client.
+2.
+Le prototype à livrer n’inclura pas toutes les fonctionnalités mentionnées dans ce document mais montrera des exemples des fonctionnalités les plus importantes suite à une réunion avec les futurs utilisateurs et décideurs.
+3.
+L’intégration avec d’autres systèmes n'est pas inclus dans le cadre de cette proposition car plus d’informations sont nécessaires concernant les systèmes existants pour pouvoir estimer la charge de travail.
+4.
+Il est supposé que le client fournira les exigences hardware ainsi que les droits d’accès et les autorisations pour installer tous les composants du logiciel.
 Alternativement, le logiciel pourrait être installé dans le cloud.
-1.
-Même si la programmation mathématique offre la capacité de prouver l'optimalité d'une solution, l'application ne garantit pas la trouver dans tous les cas.
-La qualité des solutions et le temps nécessaire pour les obtenir dépendront de nombreux facteurs; la taille de l'instance étant la principale.
+5.
+Même si la programmation mathématique offre la capacité de prouver l’optimalité d’une solution, l’application ne garantit pas de la trouver dans tous les cas de figure.
+La qualité des solutions et le temps nécessaire pour les obtenir dépendront de nombreux facteurs; la taille de l’instance étant la principale.
 
-## Références
+## Bibliographie
+### Bibliographie

### 2018-03-20 R/OptiPlan_presentation/proposal_fr.Rmd
Take tex generation out of compilation for pdf

@@
     # top_level_division: part
-    keep_tex: true
+    # keep_tex: true
     pandoc_args: ["--top-level-division=part"]

### 2018-04-03 latex/MOSIM2018/article.tex
Added changes in ROADEF presentation after last test.
changes to MOSIM article based on reviewers' commentaries

@@
 The quality of the solutions is measured via a bi-objective function that smooths both maintenance operations and aircraft unavailability among time periods. 
-An exact optimization approach is provided by a Mixed Integer Programming model.
-A real world dataset provided by the French Air Force is used to test the developed optimization approach.
+A real world dataset provided by the French Air Force is used to test an exact optimization approach based on a Mixed Integer Programming model.
 Mono-objective computations are solved to optimality for medium size instances, allowing to compute exactly the Pareto frontier of the bi-objective optimization problems.
@@
     \begin{tabular}{p{5mm}p{70mm}}
-        % $\mathcal{I}$     &   \\
-        % $\mathcal{T}$     &   \\
-        % $\mathcal{J}$     &   \\
+        % $\mathcal{I}$     &  resources. \\
+        % $\mathcal{T}$     &  time periods inside planning horizon. \\
+        % $\mathcal{J}$     &  tasks. \\
         $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
@@
     where weights $W_1$ and $W_2$ are chosen by the decision maker.
+$W_1$ penalizes the maximum number of aircraft in maintenance among all periods.
+$W_2$ penalizes the maximum number of unavailable aircraft among all periods.
     The following constraints are used in the model:       
@@
 Heuristics could be implemented, but also mathematical programming can be improved with symmetry breaking or MIP reformulations.
+A matheuristic method derived from the existing MIP formulation can also be used to improve the performance.
+
   For the bi-objective optimization for large size instances, more advanced bi-objective methods such as $\epsilon$-constraints or Two-Phases-Method should be investigated for the computation of larger Pareto frontiers.

### 2018-04-03 R/OptiPlan_presentation/proposal.Rmd
Re-organized R functions and files

@@
 setwd("../")
-source('scripts/params.R')
-source('scripts/import_results.R')
+source('functions/params.R')
+source('functions/import_results.R')
 source('diagrams/cases.R')

### 2018-04-03 R/dga_diaporama/dga.Rmd
Re-organized R functions and files

@@
 output:
-  beamer_presentation:
-    incremental: yes
   slidy_presentation:
     css: ./../../css/baobab.css
+    incremental: yes
+  beamer_presentation:
     incremental: yes
@@
 knitr::opts_chunk$set(echo = FALSE)
+setwd("../")
+source('scripts/params.R')
+source('scripts/import_results.R')
+source('diagrams/cases.R')
+source("diagrams/architecture.R")
+source("diagrams/functionalities.R")
+source('diagrams/exampleDT.R')
+source('diagrams/gantt_tasks.R')
+exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
+exp_directory2 = PATHS[['experiments']] %>% paste0('201802061201/')
+gantt <- print_solution(exp_directory, max_resources = 10, width="100%", height="500px")
+gantt_tasks <- print_tasks(exp_directory2, width="100%", height="200px", max_tasks = 5)
+graphs_rem <- graph_remaining(exp_directory)
+table_rem <- table_remaining(exp_directory)
+html_output <- knitr::opts_knit$get("rmarkdown.pandoc.to") %in% c("slidy", 'html')
 ```
@@
 
-* Péruvien et français.
 * Ingénieur Industriel de l'Universidad Politecnica de Madrid.
@@
 * Un problème d'affectation des missions et planification des maintenances des aéronefs militaires.
-* Résolu par techniques comme PLNE, méta-heuristiques ou heuristiques.
+* Résolu par techniques comme PLNE, méta-heuristiques ou heuristiques. Ou une combinaison des trois.
 
@@
 
-![Maintenances](./../../img/calendar.png)
+```{r, echo=FALSE, out.width='90%'}
+if (html_output){
+    gantt_tasks
+} else {
+    knitr::include_graphics('../../img/tasks_gantt.png')
+}
+```
 ```{r echo=FALSE}
-#htmltools::includeHTML("./../../img/calendar.html")
 ```
@@
 <!-- <div class="col2"> -->
-![Maintenances](./../../img/maintenances.png)
+<!-- ![Maintenances](./../../img/maintenances.png) -->
+```{r, echo=FALSE, out.width='70%'}
+
+if (html_output){
+    graph_architecture()
+
+} else {
+    knitr::include_graphics('../../img/maintenances.png')
+}
+```
 
 # Exemple (3)
-![Affectations](./../../img/unavailable.png)
-```{r echo=FALSE}
-# htmltools::includeHTML("./../../img/maintenances.html")
-# htmltools::includeHTML("./../../img/unavailable.html")
+<!-- ![Affectations](./../../img/unavailable.png) -->
+```{r, echo=FALSE, out.width='70%'}
+
+if (html_output){
+    graph_architecture()
+
+} else {
+    knitr::include_graphics('../../img/unavailable.png')
+}
 ```
-<!-- </div> -->
 
@@
 * Domaine scientifique: l'optimisation combinatoire.
-* Pour trouver des bonnes solutions aux problèmes il faut souvent chercher la structure du problème, sa complexité et ses similitudes / differences avec d'autres problèmes.
+* Pour trouver des bonnes solutions aux problèmes il faut souvent chercher la structure du problème, sa complexité et ses similitudes / différences avec d'autres problèmes.
 

### 2018-04-03 R/diagrams/fixed_planning.R

@@ -1,43 +1,0 @@
-library(timevis)
-library(tidyverse)
-library(data.table)
-library(magrittr)
-
-config <- list(
-    stack = FALSE,
-    editable = TRUE,
-    align = "center",
-    orientation = "top",
-    snap = NULL,
-    margin = 0,
-    zoomable= FALSE
-)
-
-data.table(
-    id = 'A1'
-    ,start= c('2018-01', '2018-10', '2019-05')
-    ,end= c('2018-01', '2018-10', '2019-05')
-)
-
-states <- 
-    solution %>% 
-    extract2('state') %>% 
-    bind_rows(.id = "UNIT") %>% 
-    gather(key = 'month', value = 'state', -UNIT) %>% 
-    filter(state %>% is.na %>% not) %>% 
-    bind_rows(tasks) %>% 
-    collapse_states() %>% 
-    inner_join(task_hours) %>% 
-    mutate(id= c(1:nrow(.)),
-           style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color),
-           content = if_else(state=='M', 'M', sprintf('%s (%sh)', state, hours))
-    ) %>% 
-    select(id, start= month, end= post_month, content, group= UNIT, style)
-
-if (max_resources %>% is.null %>% not){
-    resources <- states %>% distinct(group) %>% sample_n(max_resources)
-    states <- states %>% inner_join(resources)
-}
-
-groups_c <- states %>% distinct(group) %>% unlist
-groups <- data.table(id= groups_c, content= groups_c)

### 2018-04-03 R/scripts/import_results.R

@@ -1,156 +1,0 @@
-library(tidyverse)
-library(readxl)
-library(magrittr)
-library(stringr)
-library(zoo)
-library(data.table)
-library(lubridate)
-library(timevis)
-library(Hmisc)
-library(RColorBrewer)
-library(htmlwidgets)
-library(jsonlite)
-
-collapse_states <- function(table){
-    
-    # we complete missing months
-    months <- table$month
-    resources <- table %>% extract2('UNIT')
-    min_date <- months %>% min %>% paste0('-01') %>% as.Date()
-    max_date <- months %>% max %>% paste0('-01') %>% as.Date()
-    total_rows <- 
-        seq(from= min_date, to=max_date, "months") %>% 
-        as.character() %>% 
-        str_sub(end='7') %>% 
-        CJ(month= ., UNIT=resources)
-    
-    table %>% 
-        full_join(total_rows) %>% 
-        mutate(state = if_else(is.na(state), "", state),
-               temp = state) %>% 
-        arrange(UNIT, month) %>% 
-        group_by(UNIT) %>%  
-        mutate(prev_temp= temp %>% lag,
-               first_row= row_number()==1,
-               change = temp != prev_temp) %>% 
-        filter(first_row | change) %>% 
-        mutate(post_month = month %>% lead %>% paste0("-01") %>% ymd() %>% add(0) %>% format("%Y-%m"),
-               post_month = post_month %>% if_else(is.na(.), max_date %>% ymd() %>% format("%Y-%m"), .)) %>% 
-        select(-temp, -prev_temp, -first_row, -change) %>% 
-        ungroup() %>% filter(state != "")
-}
-
-print_solution <- function(exp_directory, max_resources=NULL, ...){
-    solution_path = exp_directory %>% paste0('data_out.json')
-    input_path = exp_directory %>% paste0('data_in.json')
-    
-    solution <- read_json(solution_path)
-    input <- read_json(input_path)
-    
-    task_hours <-
-        input %>% 
-        extract2('tasks') %>% 
-        lapply("[[", 'consumption') %>% 
-        bind_rows() %>% 
-        gather(key='state', value = 'hours') %>% 
-        bind_rows(data.table(state='M', hours=0)) %>% 
-        mutate(bucket= hours %>% as.integer %>% multiply_by(-1) %>% cut2(g= 4),
-               color = RColorBrewer::brewer.pal(4, "RdYlGn")[bucket])
-    
-    tasks <- 
-        solution %>% 
[…]

### 2018-04-03 R/scripts/params.R

@@ -1,12 +1,0 @@
-library(magrittr)
-
-path_root = '/home/pchtsp/Documents/projects/'
-path_results = path_root %>% paste0("OPTIMA_documents/results/")
-
-PATHS = list(
-    root= path_root
-    ,results = path_results
-    ,experiments = path_results %>% paste0("experiments/")
-    ,img = path_root %>% paste0("OPTIMA/img/")
-    ,latex = path_root %>% paste0("OPTIMA/latex/")
-)

### 2018-04-03 R/dga_diaporama/dga.Rmd
updated DGA presentation

--- 2017-12-20 R/dga_diaporama/dga.Rmd
@@
 author: "Franco Peschiera"
-date: '`r format(Sys.time(), "%Y-%m-%d")`'
-bibliography: MFMP.bib
-nocite: "@*"
+# date: "`r format(Sys.time(), "%Y-%m-%d")`"
 output:
   slidy_presentation:
+    css: ./../../css/baobab_slidy.css
+    # css: ./../../css/baobab.css
     incremental: yes
-    css: "./../../css/baobab.css"
+  beamer_presentation:
+    incremental: yes
+nocite: '@*'
+bibliography: MFMP.bib
 ---
@@
 knitr::opts_chunk$set(echo = FALSE)
+setwd("../")
+source('functions/params.R')
+source('functions/import_results.R')
+source('functions/exampleDT.R')
+source('functions/gantt_tasks.R')
+exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
+exp_directory2 = PATHS[['experiments']] %>% paste0('201802061201/')
+gantt <- print_solution(exp_directory, max_resources = 10, width="100%", height="500px")
+# gantt_tasks <- print_tasks(exp_directory2, width="100%", height="200px", max_tasks = 5)
+graphs_rem <- graph_remaining(exp_directory)
+table_rem <- table_remaining(exp_directory)
+html_output <- knitr::opts_knit$get("rmarkdown.pandoc.to") %in% c("slidy", 'html')
 ```
@@
 
+```{r, echo=FALSE, out.width='90%'}
+if (html_output){
+    gantt
+} else {
+    knitr::include_graphics('../../img/tasks_gantt.png')
+}
+```
 ```{r echo=FALSE}
-htmltools::includeHTML("./../../img/calendar.html")
 ```
@@
 
-<style>
-  .col2 {
-    columns: 2 200px;         /* number of columns and width in pixels*/
-    -webkit-columns: 2 200px; /* chrome, safari */
-    -moz-columns: 2 200px;    /* firefox */
+<!-- ![Maintenances](./../../img/maintenances.png) -->
+```{r, echo=FALSE, out.width='70%'}
+
+if (html_output){
+    # graph_architecture()
+    knitr::include_graphics('../../img/maintenances.png')
+} else {
+    knitr::include_graphics('../../img/maintenances.png')
 }
-</style>
+```
 
-<div class="col2">
-```{r echo=FALSE}
-htmltools::includeHTML("./../../img/maintenances.html")
-htmltools::includeHTML("./../../img/unavailable.html")
+# Exemple (3)
+<!-- ![Affectations](./../../img/unavailable.png) -->
+```{r, echo=FALSE, out.width='70%'}
+
+if (html_output){
+    # graph_architecture()
+    knitr::include_graphics('../../img/unavailable.png')
+} else {
+    knitr::include_graphics('../../img/unavailable.png')
+}
 ```
-</div>
 

### 2018-04-12 latex/CLAIO2018/article.tex
very preliminary version of CLAIO2018 article

--- 2018-01-30 latex/MOSIM2018/article.tex
@@
-\documentclass[a4paper,twocolumn,fleqn]{article}
-\textwidth 17cm \textheight 247mm
-\topmargin -4mm
-\hoffset -9mm \voffset -14mm
-%\setlength{\topmargin}{-0.8truecm}
-\setlength{\oddsidemargin}{0.3cm}
-\setlength{\evensidemargin}{0.3cm}
-\setlength{\columnsep}{8mm}
-\setlength{\parindent}{0mm}
-\setlength{\parskip}{-2.0ex}
-\setlength{\mathindent}{0mm}
-\flushbottom
+%========================================================
+% File:         claio_paper.tex, template for CLAIO 2015 conference paper
+% Based on:     CLAIO 2014 paper template
+%========================================================
+\documentclass[letterpaper,11pt]{article}
 
-\usepackage{epsfig}
-%\usepackage{timesnew}
-% \usepackage{harvard}
-\usepackage[]{natbib}
+% added by Roger
+% \usepackage{psfig,epsfig}
+\usepackage{layout}
+% added by J. Vera
+\usepackage[pdftex]{graphics}  % In case you want to use PDFTeX
+\usepackage{graphicx}
+\renewcommand{\baselinestretch}{1.0}
 \usepackage{amsmath}
-\usepackage{booktabs}
+\usepackage[latin1]{inputenc}
+\usepackage[frenchb,english]{babel}
+%
+%       Text dimensions.
+%       PLEASE, DO NOT CHANGE THE TEXT DIMENSIONS.
+%
+\setlength{\textwidth}{6.5in}
+\setlength{\textheight}{8.5in}
+\setlength{\oddsidemargin}{0.0in}
+\setlength{\evensidemargin}{0.0in}
+\setlength{\topmargin}{0.0in}
+\setlength{\headsep}{0.1in}
+\setlength{\headheight}{0.4in}
 
-\usepackage[latin1]{inputenc}
-\usepackage{aeguill}
-\usepackage[frenchb,english]{babel}
-\usepackage{caption}
+\pagestyle{empty}
 
-\setlength{\parskip}{2ex} \pagestyle{myheadings}
-\markright{\hspace*{3.8cm} \textit{MOSIM18 - June 27-29, 2018 - Toulouse - France}}
+%
+%  Title definition
+%
+\title{\bf MIP formulation for the optimization of maintenance planning on French military aircraft operations}
 
-\renewcommand{\thepage}{}
-\renewcommand{\refname}{REFERENCES}
+\author{%
+ISAE-SUPAERO, Université de Toulouse, France\\%
+Franco Peschiera\\%
+\textit{franco.peschiera@isae-supaero}\\[3mm]
+Nicolas Dupin\\%
+\textit{nicolas.dupin@isae-supaero}\\[3mm]
+Olga Battaïa\\%
+\textit{olga.battaia@isae-supaero}\\[3mm]
+Alain Haït\\%
+\textit{alain.hait@isae-supaero}
+}
 
-\makeatletter
-\renewcommand\section{\@startsection{section}{1}{\z@}%
-                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
-                       {2\p@ \@plus 0\p@ \@minus 0\p@}%
-                       {\normalsize\textbf}}
+\date{ }
 
-\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
-                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
-                       {2\p@ \@plus 0\p@ \@minus 0\p@}%
-                       {\normalsize\textbf}}
-
-\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
-                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
-                       {1\p@ \@plus 0\p@ \@minus 0\p@}%
-                       {\normalsize\itshape\bfseries}}
-\makeatother
-
+%
+%  Document starts here
+%
 \begin{document}
 
-\title{ \vspace*{-25mm}
-\centering
-\fbox{\normalsize
-\begin{minipage}{17cm}
-\centering
-\em \small{$12^{th}$ International Conference on MOdeling, Optimization and SIMlation - MOSIM18 - June 27-29 2018  \\ Toulouse - France "The rise of connected systems in industry and services"}
-\end{minipage} }\\
-\vspace*{6mm}
-{\Large \textbf{Bi-objective MIP formulation for the optimization of maintenance planning on French military aircraft operations}}}
+\maketitle
 
-\author{
-\begin{tabular}{c}
-\bf \normalsize {Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin} \\
-\\
-     \normalsize ISAE-SUPAERO, % \\
-     \normalsize Université de Toulouse, % \\
-     \normalsize 31055 Toulouse cedex 4 - France\\
-     \normalsize {\{franco.peschiera,alain.hait,olga.battaia,nicolas.dupin\}@isae-supaero.fr}
-     \\
-\end{tabular}
-}
+\thispagestyle{empty}
 
-\date{\begin{minipage}{17cm}
-\normalsize
-{\bf  ABSTRACT:}
-\rm
+\begin{abstract}
+
 {\em A specific Flight and Maintenance Planning problem is presented. 
@@
 The quality of the solutions is measured via a bi-objective function that smooths both maintenance operations and aircraft unavailability among time periods. 
-An exact optimization approach is provided by a Mixed Integer Programming model.
-A real world dataset provided by the French Air Force is used to test the developed optimization approach.
-Mono-objective computations are solved to optimality for medium size instances, allowing to compute exactly the Pareto frontier of the bi-objective optimization problems.
-The tests show that these two objectives do not lead to the same optimal solution, but very good compromise solutions can be found thanks to the bi-objective optimization.
-}\\~\\
-{\bf KEYWORDS:}
-\rm
-{\em optimization, Flight and Maintenance Planning, aircraft maintenance, Mixed Integer Programming, bi-objective optimization}
-\end{minipage}
+A real world dataset provided by the French Air Force is used to test an exact optimization approach based on a Mixed Integer Programming model.
 }
-\maketitle
+\vspace{2em}
 
-% Here I put the list of things we will leave for another article:
-% 1. stochasticity
-% 2. symmetry breaking inside model.
-% 3. storage?
+\noindent\textit{Keywords:}
+optimization;
+Flight and Maintenance Planning;
+aircraft maintenance;
+Mixed Integer Programming.
+
+\end{abstract}
 
 \section{Introduction}
+\label{sec:intro}
 
@@
     On the other hand, there is a will to align two objectives: the smoothing of the maintenance operations and the minimization of the global unavailability of aircraft to be able to add new missions.
-This last point calls for a bi-objective optimization.
+
+    This case of the FMP has been already studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015,Seif2018}.
 
@@
 
-  The article is structured as follows:
+\section{Problem Description}
+\label{sec:model}
+% TODO: make this compatible with parameters.
 
-  Section \ref{sec:problem} presents a detailed description of the problem. An analysis of the previous work on FMP is given in section \ref{sec:soa}.
-  Section \ref{sec:model} presents a new MIP formulation for the considered problem. 
-  Section \ref{sec:data} introduces the real case study.
-  Section \ref{sec:results} discusses the obtained results.
-  Finally, section \ref{sec:conclusions} provides conclusions and directions of further work.
+    The problem consists in assigning military aircraft to a given set of missions while scheduling the maintenance operations over a time horizon.
+    In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks".
+A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
+Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
+Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.
+Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
+Set $I_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
+The number of resources in use for each period $d_t$ is also known.
 
-\section{Problem statement}
-  \label{sec:problem}
-  The problem consists in assigning resources to predefined tasks and scheduling periodic preventive maintenances for these same resources.
-For the seek of simplicity, the following convention has been adopted: aircraft are called resources, missions are called tasks.
+    Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
+A resource cannot be used for a task $j$ if its remaining usage time is less than usage $h_j$ required by task $j$.
+A maintenance operation assigns to the resource exactly $H$ units of remaining usage time.
+The remaining usage time not used before the maintenance operation is lost.
 
-  \subsection{Tasks}
-    \label{def:task}
+    A maintenance operation takes exactly $m$ periods and cannot be interrupted: during this time the resource cannot be assigned to any task.
+Some resources can be in maintenance at the beginning of the planning horizon, $n_t$ is used for the number of aircraft in such planned maintenance per period $t$ and defined only for the first $m-1$ time periods.
+In the next section, we present a new mathematical model for this problem.
 
-    There is a fixed set of $j \in \mathcal{J}$ tasks to be accomplished over an horizon of time divided into $t \in \mathcal{T}$ discrete periods.
-For their execution, these tasks require the assignment of a specific number of resources $R_j$ each period of time the task is active.
-The start and end periods for each task are known and a task is considered active between its start and end period.
-
-    During each period, tasks consume an amount of time equal to $H_j$ from each of its assigned resources.
-
-    % The assignment of a resource to a task is not decided for the whole duration of the task.
-After a minimum amount of time ($MT$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished.
-    The total number of resources being used at any given time in a specific task should be equal to $R_j$.
-
-    Each task requires one and only one type of resource which, in addition, should comply with additional task requirements.
-
-  \subsection{Resources}
-    \label{def:res}
-
-    There is a set $i \in \mathcal{I}$ of available resources that are assigned to tasks in order to accomplish them.
-Each resource can only be assigned to a single task in any given period.
-These resources suffer from wear and tear and require regular maintenance operations during their lifetime.
-The need for maintenance is calculated based on two indicators.
-
-    The first one is called "remaining elapsed time" (or $ret_{it}$).
-It expresses the amount of time (measured in time periods) after which the resource cannot be used anymore and has to undergo a maintenance operation.
-Its value is calculated for each resource $i$ and each time period $t$.
-In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the resource $i$ can be used before needing a maintenance operation at any given period $t$.
-
-    % Additionally, after an absolute amount of time and/or usage ($aet_i$ or $aut_i$), the resource becomes obsolete. There is no way to reverse this process.
-
-    At any given period, including at the start of the planning horizon, each resource has a specific status given by remaining usage time and remaining elapsed time.
-
-    % \begin{itemize}
-        % \item remaining absolute elapsed time.
-        % \item remaining absolute usage time.
-        % \item remaining storage time (see \ref{def:sto}).
-        % \item type of last maintenance (see \ref{def:maint}).
-    % \end{itemize}
-
-  \subsection{Maintenances}
-    \label{def:maint}
-
-    Maintenances operations are the process by which resources that have reached a limit in some indicator can return to a state where they can continue to be used in tasks.
-
-    % Each maintenance belongs to a specific type of maintenance $m \in \mathcal{M}$.
-These types of maintenance differentiate between each other by having potentially different characteristics such as: different durations, the provision of new functionalities for the resource or the restoration of the storage capacity (see \ref{def:sto}).
-The maintenance type is not a decision to be made since there is a specific sequence of maintenance types a resources needs to follow according to the resource type.
-
-    Each maintenance operation has a fix duration of $M$ periods.
-
-    After a maintenance operation, a resource restores its remaining elapsed time and remaining usage time to their max values $E$ and $H$ respectively.
-
-    % Finally, resources are organized into families or groups. Each resource inside a family or group shares the same types of maintenances and, usually, is able to do the same type of tasks.
-
-  % \subsection{Storages}
-  %   \label{def:sto}
-
-  %   Following the rule of remaining elapsed time, even if a resource is not being used, it still needs to have a maintenance after a given amount of time has passed.
-In order to avoid this problem, the resource can be put into a storage state.
-
-  %   A resource in this states has to be kept for a minimum time of $sm$ periods. While in this state it cannot receive maintenance or be assigned any task.
-
-  %   Every resource has the capacity to be stored and this capacity is measured in a number of periods $sc$.
-In order for a resource to restore its remaining storage capacity, it needs to receive a specific maintenance (see \ref{def:maint}).
-Similar to the remaining elapsed time, after these maintenances, the resource recovers its storage capacity up to a certain level $S$.
-
-  \subsection{Possible states}
-
-    As a summary, the following are the possible logical states that a resource can be in: assigned to a task (see \ref{def:task}); under maintenance (see \ref{def:maint}); available.
-
-      % \begin{itemize}
-          % \item Under storage (see \ref{def:sto}).
-            % \item Obsolete (see \ref{def:res}).
-      % \end{itemize}
-
-    % Figure \ref{fig:solution} shows an extraction of an assignment of missions (red) and maintenances (blue) to aircraft (codes A1 - A9).
-
-    %   \begin{figure}
-    %     \centering
-    %       \includegraphics[width=0.5\textwidth]{./../../img/calendar.png}
-    %     \caption{Visual representation of part of the solution to an instance of the problem showing maintenances and assignments.
-Resources are listed in the vertical axis and the horizontal axis defines the time.
-Red periods correspond to task assignments while blue ones to maintenance assignments.}
-    %     \label{fig:solution}
-    %   \end{figure}
-
-  \subsection{Time}
-    \label{def:hor}
-
-    In planning tasks and maintenances, it is important to take into account the initial and end state of each resource.
-This initial state can be a maintenance or an assigned task.
-If a resource is under maintenance, it needs to continue in this state for its remaining maintenance time.
-Tasks' assignments should be taken into account in a similar manner.
-
-    The remaining used and elapsed times are assigned to each resource at the beginning of the planning horizon.
-
-    Finally, the state of each resource at the end of the planning horizon, its remaining (elapsed, used) time, needs to be defined.
-      % needs to be addressed too. This implies guaranteeing that the remaining (elapsed, used) time of resources is sustainable for future tasks and the final states of resources are not too skewed.
-
-  \subsection{Objectives}
-
-    In this study, the following objectives are considered.
-
-    Given that the creation of new tasks and the duration of maintenance are considered stochastic in real-life, one basic goal is to maximize the robustness of the planning by having the greatest amount of available resources at every period of the planning horizon.
-
-    Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of resources under maintenance over the planning horizon.
-
-    As it will be shown in the mathematical formulation, these objectives are quite related one with the other: the more resources are in maintenance in a given period, the more unavailable resources will be had.
-
-\section{Related work}
-  \label{sec:soa}
-
-  % TODO: add FMP biblio?
-
-  % The Military Flight and Maintenance Planning Problem considered here aims to assign missions and maintenance tasks to military aircraft.
-It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically since a round-trip to the base is assumed for each flight.
-
-  % It also includes different objectives and constraints.
-
-  % Although the former has been studied in literature in \citet{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015}, it has not received as much attention as the latter.
-In the following, the problem with already scheduled missions is considered.
-
-  The Military Flight and Maintenance Planning problem considered assigns missions and schedules maintenance operations to aircraft.
-In this problem, aircraft are needed to comply with a series of planned missions while at the same time needing to receive frequent preventive maintenances in order to be ready for new missions.
-
-  It is a variant of the Flight and Maintenance Planning problem, where maintenance operations and flights are scheduled to commercial aircraft.
-The main differences are the fact that flights are not modeled geographically: all aircraft are assumed to exit the main depot and return to the same depot at the end of the flight.
-
-  Other differences include the following.
-
-  Firstly, the objective function's motivation is on reducing the load on the maintenance facilities and guaranteeing the availability of the fleet instead of reducing the cost.
-Also, the maintenance operations take considerably longer and the rules that govern the maintenances are different.
-
-  The temporal scope of the problem is also larger, making the normal planning horizon and each period larger too: years instead of months and months instead of days, respectively.
-
-  In the following, the studies on the Military Flight and Maintenance Planning are discussed.
-
-  In \citet{Kozanidis2008}, a model for the Hellenic Air Force is presented.
-It assigns flight hours to aircraft and schedules flexible maintenance periods based on a "remaining usage time" rule.
-The objectives are to maximize this "remaining usage time" for the aircraft as well as to smooth the availability of aircraft during the planning horizon.
-Instances of 24 aircraft and 6 monthly periods were solved with Mixed Integer Programming (MIP) formulations and heuristics.
-
-  In \citet{Cho2011}, the objective is to smooth the maintenance operations.
-Randomly generated instances of 15 aircraft and 520 bi-daily periods were solved with MIP formulations.
-The final state of aircraft is taken into account in order to distribute as much as possible the "remaining usage time" of aircraft.
-
-  In \citet{Verhoeff2015}, a model for the Royal Netherlands Air Force is developed.
-Here, three criteria for improving the planning robustness are presented and applied in a MIP model: availability, serviceability and sustainability.
-
-  In \citet{gavranis2015exact}, an exact algorithm is presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
-
-  Both \citet{Kozanidis2008} and \citet{Cho2011} develop heuristic alternatives to the MIP formulation for solving larger problem instances.
-In \citet{steiner2006heuristic}, a constructive, multi-stage heuristic approach is described to obtain balanced maintenances for the Swiss Air Force.
-Both calendar-based maintenances actions (CBMA) and usage-based maintenance actions (UBMA) are used.
-Finally, in \citet{kozanidis2014heuristics}, two heuristics are presented to solve large instances of the problem by taking consecutive decisions period by period with the help of mathematical models at each step.
-
-  % Regarding operational-level planning, a more detailed implementation can be found in \citet{safaei2011workforce}, where the specific conditions during the maintenance operations are taken into account to schedule the maintenance jobs, such as the workforce skills and availability.
-
-  It should be noted that previous formulations only consider an aggregate demand in flight hours for all aircraft.
-  The aircraft are assumed to be homogeneous and interchangeable.
-  The use of aggregate demand makes it impossible to take into account the task requirements for aircraft.
-   % The concept of mission is not introduced as part of these formulations, only a demand of flight hours is given for all aircraft in total.
-The aircraft themselves have been assumed as homogeneous and interchangeable.
-Due to this aggregation of demand, there are no minimal time assignments for aircraft to missions.
-Missions are important in order to guarantee that aircraft have the required capabilities needed for the missions they are assigned to.
-
-  % Additionally, the possibility of storage of aircraft is not present in the previous work.
-
-  Lastly, in \citet{Chastellux2016} a mission-centered modeled is presented.
-Here, cycles, composed of one maintenance followed by mission assignments were used.
-Instances of 6 - 70 aircraft and 60 monthly periods were solved.
-Close to optimal solutions for small to medium instances using MIP formulations and heuristics based on these MIP formulations were found.
-
-  In the following, an alternative model is presented. This model is then tested in a case study.
-
-\section{Optimization model}
-  \label{sec:model}
+\section{Solution Method}
+\label{sec:method}
 
@@
 
-    \begin{tabular}{p{5mm}p{70mm}}
-        % $\mathcal{I}$     &   \\
-        % $\mathcal{T}$     &   \\
-        % $\mathcal{J}$     &   \\
+    \begin{tabular}{p{5mm}p{165mm}}
+
         $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
@@
 
-    \begin{tabular}{p{8mm}p{67mm}}
+    \begin{tabular}{p{8mm}p{162mm}}
         $H_j$             & amount of resource time required by task $j$. \\
         $R_j$             & number of resources required by task $j$. \\
-        % $MT$              & minimum number of periods a resource can be assigned to a task. \\
-        % $aut_i$           & maximal absolute usage time for resource $i$. \\
-        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
         $M$               & maintenance duration in number of periods. \\
@@
         $H$               & remaining usage time after a maintenance. \\
-        % $S$               & remaining storage time after (certain) maintenances. \\
         $W_1$             & weight of the first objective in the objective function. \\
@@
 
-    \begin{tabular}{p{8mm}p{67mm}}
+    \begin{tabular}{p{8mm}p{162mm}}
         $a_{jti}$   &  =1 if task $j \in J$ in period $t \in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
@@
 
-    The objective is to simultaneously minimize the maximum number of maintenances and the maximum number of unavailable aircraft.
+    The objective is to minimize the total number of maintenances done and while preserving the aircraft ready at the end of the planning period.
 
@@
     \end{align}
-    where weights $W_1$ and $W_2$ are chosen by the decision maker. 
-    The following constraints are used in the model:       
+
     \begin{align}
@@
           & t \in \mathcal{T} \label{eq:capacity1}\\             
-               % maximum capacity2:                
-        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} m_{it'} \leq m_{max}
-        % & t =m, ..., |\mathcal{T}|  \label{eq:capacity2}\\
-               %avail
-       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \notag \\
-           &\hspace*{10mm}+ D_t\leq u_{max} 
+       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
         &t \in \mathcal{T} \label{eq:avalaibility1}\\                
-               % maximum capacity2:                
-        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} m_{it'} + D_t\leq u_{max} 
-        % & t =m, ..., |\mathcal{T}|  \label{eq:avalaibility2}\\
         & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
@@
         % remaining used time
-         & rut_{it} \leq rut_{it-1} + H m_{it} \notag \\ 
-           &\hspace*{4mm}- \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j \notag \\
-                && t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
+         & rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
+                & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
         & rut_{i0} = Rut^{Init}_i
@@
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\
-        & ret_{it} \leq ret_{it-1} - 1 \notag \\
-          &\hspace*{4mm}+ E m_{it}
+        & ret_{it} \leq ret_{it-1} - 1 + E m_{it}
                 & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
@@
 
-\section{Datasets}
-  \label{sec:data}
-  The dataset being used was provided by the French Air Force and already used in \citet{Chastellux2017}.
-  Based on the information provided, a number of instances were created as subsets of the main dataset in order to run different experiments.
-
-  Table \ref{tab:instance} shows the characteristics of each instance. The main differences between instances are in the number of tasks and the number of periods.
-
-  The criteria for the creation of the instances were the following:
-
-    \begin{enumerate}
-        \item The original instance was the largest instance where an integer solution was found in less than 3600 seconds: it contains 11 tasks and 21 periods in total.
-        \item We took out the task with the biggest number of required resources.
-        \item We increased the number of periods in multiples of 10 until no integer solutions can be found in 3600 seconds.
-        \item We repeated step 2.
-    \end{enumerate}
-
-  \begin{table}
-  \begin{center}
-
-  \input{tables/instance.tex}
-
-  \caption{Instances used in the experiments with indicators of size: "id" is the instance; $|\mathcal{T}|$ is the number of periods; "assign" is the number of assignments of resources to tasks; $|\mathcal{J}|$ is the number of tasks; "vars" the number of variables; "cons" is the number of constraints; and "nonzeros" is the number of non zero values in the matrix.}
-  \vspace{-0.5cm}
-  \label{tab:instance}
-  \end{center}
-  \end{table}
-
-  \subsection{Tasks}
-
-    The dataset includes 11 tasks.
-Most of them demand between 1 and 8 resources per period, which is equal to one month.
-The most loaded and second most loaded tasks demand 50 and 20 resources respectively per period.
-    Candidate resources are calculated based on each task's requirements matching the candidate's capabilities.
-
-    Some tasks have a limited duration, meaning they start and end during the planning horizon. Others are active during the whole planning horizon.
-
-  \subsection{Maintenances}
-
-    All maintenance operations last exactly 6 periods.
-Each maintenance sets the remaining usage time of a resource to 1000 hours.
-After it, a resource can be without maintenance for a maximum of 60 consecutive periods.
-The nominative capacity for maintenance is 18 resources at any given time.
-
-  \subsection{Resources}
-
-    There are 127 resources in total. Each one has its own functionalities and past assignments.
-
-    % \begin{figure}
-    %   \centering
-    %     \includegraphics[width=0.5\textwidth]{./../../img/initial_used.png}
-    %   \caption{Distribution of Usage Times among resources at the beginning of the planning horizon}
-    %   \label{fig:histogramUsage}
-    % \end{figure}
-
-% \section{Heuristics}
-
-\section{Results}
-  \label{sec:results}
-
-  \subsection{Numerical tests}
-    \label{sec:tests}
-
-    All tests were run on a 64 bit Ubuntu 16.04 workstation with 16GB of RAM and Intel i7-6700HQ CPU \@ 2.60GHz x 8 processor.
-    Several experiments were done with the MIP formulation on problem instances presented in table \ref{tab:instance}.
-
-    Table \ref{tab:summary} gives an overview of the obtained results. As can be seen, smaller instances are solved up to optimality while medium-size ones are stopped after one hour.
-
-    For this calculation, both weights on the objective function were set at 1. A sensibility analysis of these two parameters is provided in section \ref{sec:multi}.
-
-    \begin{table}
-    \begin{center}
-
-    \input{tables/summary.tex}
-
-    \caption{Information on the solution of each tested instance: "id" is the instance identifier; "objective" is the value of the objective function; "gap" is the percentage gap between the integer solution and the linear relaxation; "time (s)" is the solution time or 3600; and "bound" is the greatest lower bound found from the linear relaxation.}
-    \vspace{-0.5cm}
-    \label{tab:summary}
-    \end{center}
-    \end{table}
-
-    Table \ref{tab:results} shows the details of the solving process for each instance.
-As it can be seen, cuts do not improve significantly the relaxation (the value of "bound cuts" is very close to the root solution).
-In fact, the original linear relaxation (root) is quite similar to the final lower bound (bound).
-This suggests that further improvement in modelization would help to reduce the gap faster.
-
-    An example of a tighter modelization could be to replace the constraints that count the remaining usage time and elapsed time for tighter equivalents.
-
-    The greatest number of cuts used was of 'Implied bound' and 'Mixed integer rounding', depending on the instance.
-Since these are two generic-type cuts, it can be concluded that the solver has not detected any particular structure as part of this problem.
-
-    \begin{table}
-    \begin{center}
-
-    \input{tables/results.tex}
-
-    \caption{Solution details on the progress of the lower bound: "id" is the name of the instance; "root" is the relaxation before cuts; "b.
-cuts is the relaxation after the cuts; "bound" is the relaxation at the time limit; "cuts (\#)" is the number of cuts; and "cuts (s)" is the time the cuts took in seconds.}
-    \vspace{-0.5cm}
-    \label{tab:results}
-    \end{center}
-    \end{table}
-
-    Table \ref{tab:results2} shows the progress of the integer solution for each instance.
-For more detail on this progress, figure \ref{fig:progress} shows an example of progress for instance with id=I\_7.
-It can be seen that initial solutions are quite far from the optimal ones due to the use of heuristics at the beginning of the solving process.
-In fact, for two rather small instances the solver was able to find an optimal solution before starting to branch.
-
-    The obtained results show that the branching is not providing much improvement in increasing the lower bound or finding better integer solutions.
-The quantity of symmetries in the model, represented as candidates for each task, could be a reason the difficulty in solving large instances.
-It may exist a very large number of possible candidates to be potentially assigned for each task with respect to the requirements of resources for each task.
-This reduces the branching capacity of the solver, which tries to explore many nodes that are almost equivalent in reality.
-
-    An example of an alternative to break these symmetries could be to slightly change the objective function so it can measure more gradually the smoothness of the objective and help get better bounds.
-Another alternative could be to assign priorities to some combinations of resource-task or to limit the number of candidates that can be assigned to each task by some heuristic that clusters resources with tasks.
-Finally, arbitrarily preassigning (fixing) the final state (in terms of remaining usage time) to each resource could also help break those symmetries without important loss on the objective function value.
-
-    \begin{table}
-    \begin{center}
-
-    \input{tables/results2.tex}
-
-    \caption{Solution details on the progress of the integer solution: "id" is the name of the instance; "first" is the first solution integer found; "sol.
-cuts" is the solution after the cuts; "last" is the last solution found at the time limit.}
-    \vspace{-0.5cm}
-    \label{tab:results2}
-    \end{center}
-    \end{table}
-
-    \begin{figure}
-      \centering
-        \includegraphics[width=0.5\textwidth]{./../../img/progress.png}
-      \caption{Solving progress of the integer solution (in blue) and the lower bound (red) along the solver iterations (x axis). Instance id=I\_7.}
-      \label{fig:progress}
-    \end{figure}
-
-    Figures \ref{fig:num-maintenances} and \ref{fig:num-unavailables} show an example distribution of maintenance operations and unavailable numbers of resources per period, respectively.
-
-    \begin{figure}
-      \centering
-        \includegraphics[width=0.5\textwidth]{./../../img/num-maintenances.png}
-      \caption{Solution for instance id=I\_7: number of resources under maintenance at each period.}
-      \label{fig:num-maintenances}
-    \end{figure}
-
-    \begin{figure}
-      \centering
-        \includegraphics[width=0.5\textwidth]{./../../img/num-unavailables.png}
-      \caption{Solution for instance id=I\_7: number of unavailable resources at each period.}
-      \label{fig:num-unavailables}
-    \end{figure}
-
-  \subsection{Bi-objective analysis}
-  \label{sec:multi}
-
-    The further calculation was done in order to analyze the impact of the weights $W_1$ and $W_2$ on the solution.
-In it, both weights were changed between 0 and 1 with pace of 0.1 under the condition that $W_1 + W_2 = 1 $.
-This was done for all previously defined instances.
-
-    Table \ref{tab:multmultiobj} presents the obtained results of this analysis. It can be seen that all instances have a small number of non-dominated solutions.
-
-    \begin{table}[h]
-    \begin{center}
-
-    \input{tables/multi-multiobj.tex}
-
-    \caption{Non-dominated points found on each instance after 3600 seconds.
-All instances were analyzed: "First" and "Last" indicate the two extremes of the calculated Pareto frontier; "\# points" indicates the number of Pareto optimal points found.
-Not all experiments were solved up to optimality.}
-    \vspace{-0.5cm}
-    \label{tab:multmultiobj}
-    \end{center}
-    \end{table}
-
-    More particularly, an individual instance was analyzed, where an optimal solution was found in each experiment.
-    Table \ref{tab:multiobj} shows the obtained results. The instance used consisted of 21 periods and 10 tasks corresponding to instance with id=I\_5.
-
-    For all solved instances, a small number of non dominated solutions was observed.
-    \begin{table}[h]
-    \begin{center}
-
-    \input{tables/multiobj.tex}
-
-    \caption{Values of $u_{max}$ (unavailable) and $m_{max}$ (maint) for different combinations of $W_{1}$ and $W_{2}$. Instance id=I\_5. All experiments were solved to optimality.}
-    \vspace{-0.5cm}
-    \label{tab:multiobj}
-    \end{center}
-    \end{table}
-
-    \begin{figure}
-      \centering
-        \includegraphics[width=0.5\textwidth]{./../../img/multiobjective.png}
-      \caption{Pareto diagram for instance id=I\_5.}
-      \label{fig:pareto}
-    \end{figure}
-
-\section{Conclusions and further work}
+%
+%   Conclusions
+%
+\section{Conclusions}
 \label{sec:conclusions}
@@
   The characteristics of MIP convergence were analyzed on real-world instances.
-  Mono-objective problem instances were solved to optimality for medium size instances. Then, exact Pareto frontiers for bi-objective formulation were calculated for these instances. 
-  The use of these two objectives does not lead to the same optimal solution but very good compromise solutions can be found due to the bi-objective optimization.
+  Mono-objective problem instances were solved to optimality for medium size instances.
 For larger instances, difficulties in finding optimal solutions occurred because of multiple symmetries and a poor quality of MIP generic primal heuristics.
@@
 Heuristics could be implemented, but also mathematical programming can be improved with symmetry breaking or MIP reformulations.
-  For the bi-objective optimization for large size instances, more advanced bi-objective methods such as $\epsilon$-constraints or Two-Phases-Method should be investigated for the computation of larger Pareto frontiers.
+A matheuristic method derived from the existing MIP formulation can also be used to improve the performance.
+
   Lastly, additional constraints from the real world application can be incorporated in the model.
 
-\bibliographystyle{dcu}
+%
+%   References
+%
+\bibliographystyle{plain}
 \bibliography{./../biblio/MFMP}
-
 \end{document}

### 2018-04-18 latex/CLAIO2018/summary.tex
renamed file summary.tex

A specific Flight and Maintenance Planning problem is addressed. In this problem, preventive maintenance operations are scheduled for military aircraft along with the assignment of regular, pre-scheduled, missions.

More generally, a series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Each period usually represents a month. Since all tasks are already scheduled, each task has periods $T_j \subset \mathcal{T}$ in which they will be realized. Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ hours at each period. Set $I_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task.

Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time. A maintenance operation takes exactly $m$ periods and cannot be interrupted. It assigns to the resource exactly $H$ units of remaining usage time. The remaining usage time not used before the maintenance operation is lost.

Work has already been done in military aircraft planning by \cite{Kozanidis2008, Cho2011},

There are several objectives to consider. Reducing the maximum amount of resources under maintenance; maximizing the availability of resources; reducing the total number of maintenance operations.

### 2018-04-27 latex/notes/FMP_variants.tex
Added file with variants of models to see complexity of FMP

\documentclass[a4paper,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

%opening
\title{Variants on the Flight and Maintenance Planning problem}
\author{}

\setlength{\topmargin}{-2.1cm}
\setlength{\textwidth}{15.8cm}
\setlength{\textheight}{26.3cm}
\setlength{\oddsidemargin}{-0.1cm}
\setlength{\evensidemargin}{-0.1cm}

\def\II{{\mathcal{I}}}
\def\TT{{\mathcal{T}}}
\def\KK{{\mathcal{K}}}

\usepackage{graphicx}
\usepackage{amssymb}

\begin{document}

\maketitle

\section{Crew scheduling variant}

    \subsection{Hypothesis}

    \begin{itemize}
     \item fixed maintenance duration.
     \item fixed duration between maintenances.
     \item homogeneous fleet.
     \item objective: minimize the number of aircraft.
    \end{itemize}

    \subsection{Notations}

    \textbf{Sets}

    \begin{tabular}{ll}
    $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
    $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
    \end{tabular}

    \vskip 0.3cm

    \subsection{Parameters}

    \begin{tabular}{ll}
        $D_{off}$ & Number of consecutive non-working days. \\
        $D_{on}$ & Number of consecutive working days. \\
        $\mathcal{A}_{it}$ & Pattern $i$ covers period $t$ (circulant matrix built from $D_{on}$ and $D_{off}$).\\
        $b_{t}$ & Need of resources at period $t$. \\
    \end{tabular}

    \vskip 0.3cm

    Example of A for $D_{on}=4$ and $D_{off}=2$

    $$
    \begin{bmatrix}
        1 & 0 & 0 & \dots  & A_{i1} \\
        1 & 1 & 0 & \dots  & A_{i2} \\
        1 & 1 & 1 & \dots  & A_{i3} \\
        1 & 1 & 1 & \dots  & A_{i4} \\
        0 & 1 & 1 & \dots  & A_{i5} \\
        0 & 0 & 1 & \dots  & A_{i6} \\
        1 & 0 & 0 & \dots  & A_{i6} \\
        \vdots & \vdots & \vdots & \ddots & \vdots \\
        A_{1t} & A_{2t} & A_{3t} & \dots  & A_{nm}
    \end{bmatrix}
    $$

    \subsection{Variables}

    \begin{itemize}
     \item $x_i$: number of patterns of type $i$ to be included in the solution.
    \end{itemize}

    \subsection{Model}

    \begin{align}
        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i
    \end{align}

    \begin{align}
        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq d_j & t \in \mathcal{T} \\
        & x_{i} \geq 0,\; integer \\
    \end{align}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Set covering variant}

    \subsection{Hypothesis}

    \begin{itemize}
     \item only 1 aircraft needed per month.
     \item objective: minimize the number of aircraft.
    \end{itemize}

    \subsection{Notations}

    \textbf{Sets}

    \begin{tabular}{ll}
    $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
    $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
    \end{tabular}

    \vskip 0.3cm

    \subsection{Parameters}

    \begin{tabular}{ll}
        $\mathcal{A}_{it}$ & Pattern $i$ covers period $t$.\\
    \end{tabular}

    \vskip 0.3cm

    \subsection{Variables}

    \begin{itemize}
     \item $x_i$: is pattern $i$ in the solution or not.
    \end{itemize}

    \subsection{Model}

    \begin{align}
        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i
    \end{align}

    \begin{align}
        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq 1 & t \in \mathcal{T} \\
        & x_{i} \geq 0,\; binary \\
    \end{align}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Scheduling variant}

    \subsection{Hypothesis}

    \begin{itemize}
     \item durée de maintenance définie
     \item fenêtres de temps pour commencer les maintenances
     \item objectif: lisser les maintenances dans les fenêtres de temps prédéfinies
    \end{itemize}

    \subsection{Notations}

    \noindent{
    \begin{tabular}{ll}
        $t\in \TT = [\![1, T]\!]$ & Discrétisation temporelle de l'horizon de planigfication. \\
        $i \in \II = [\![1, I]\!]$ & Ensemble des avions.\\
        $k \in \KK_i = [\![1, K_i]\!]$ & Ensemble des maintenances à planifier successivement.\\ 
    \end{tabular}
    }

    \vskip 0.3cm

    \textbf{Paramètres}

    \begin{tabular}{ll}
        $\mathbf{Da}_{i,k}$ & Durée nominale de la maintenance $k$ de l'avion $i$.\\
        $\mathbf{l}_{i}$ & Durée mininale du cycle de vol de l'avion $i$.\\
        $\mathbf{{To}}_{i,k}$ & Pas de temps le plus tôt pour commencer la maintenance $k$ de l'avion $i$.\\
        $\mathbf{{Ta}}_{i,k}$ & Pas de temps le plus tard pour terminer la maintenance $k$ de l'avion $i$.\\
    \end{tabular}

    % \caption{Definition and notations the challenge ROADEF}\label{contraintesROADEF}
    % \end{table}

    \subsection{Définition des variables}

    \begin{itemize}
     \item $d_{i,k,t}\in \{0,1\}$; variables binaires pour indiquer les dates d'arrêts, "step varaibles" 
    $d_{i,k,t}=1$ si l'arrêt du cycle $k$ du réacteur nucléaire $i$ a été commencé avant la semaine $t$.
    $d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \in \{0,1\}$ vaut $1$ uniquement pendant a maintenance $(i,k)$.

     \item $m,M\geqslant 0$: variables continues pour linéariser la fonction objectif
    \end{itemize}

    \subsection{Formulation PLNE}

    \begin{eqnarray}
    \displaystyle \min_{d,m,M} & M-m
     & \\
    \forall i , k, t, & d_{i,k,t-1}\leqslant d_{i,k,t} & \label{PANprecedence}\\
    \forall i , k, & d_{i,k,\mathbf{To}_{i,k}-1}\leqslant 0 & \label{PANtw0} \\
    \forall i , k, & d_{i,k,\mathbf{Ta}_{i,k}}\geqslant 1 & \label{PANtw1}\\
    \forall i , k, t, & d_{i,k,t}\leqslant d_{i,k-1,t-l_i-\mathbf{Da}_{i,k}} & \label{PANprecedence2}\\
    \forall t, & \displaystyle \sum_{i,k} d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \leqslant M & \label{maxMaint}\\
    \forall t, & \displaystyle \sum_{i,k} d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \geqslant m & \label{minMaint}\\
    & d \in \{0,1\}^N, m,M \geqslant 0
    \end{eqnarray}

    (\ref{PANprecedence}): contrainte induite par la définition ``step variables".

    (\ref{PANtw0}), (\ref{PANtw1}): contraintes de fenêtres de temps (écrit explicitement, ou sinon définir uniquement les variables $d_{i,k,t}$
    dans les fenêtres de temps)

    (\ref{PANprecedence2}): durée minimale d'un cycle de vol $l\geqslant 0$, contrainte de succession des cycles.

    (\ref{maxMaint}), (\ref{minMaint}): contraites pour linéariser les $\min$ et $\max$ de la focntion objectif de lissage

    \subsection{Questions ouvertes}

    \begin{itemize}
    %\item Avec $l=0$ , y a t'il un problème de référence proche ? Ordonnancement, lissage de charge ?
    \item Avec $l=0$ , y a t'il un problème de référence proche ? Ordonnancement, lissage de charge ?
    \item Est-ce un problème NP-complet? Peut on le prouver ?
     \item Cas particulier $\mathbf{Ta}_{i,k}-\mathbf{To}_{i,k}=1$? (pour lissage local)
     \item Autres fonctions objectifs?
     \item version robuste avec aléas de prolongation de durée de maintenance?
    \end{itemize}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Crew scheduling: smooth maintenances}

    \subsection{Hypothesis}

    \begin{itemize}
     \item fixed maintenance duration.
     \item range to start each maintenance.
     \item min distance between maintenances.
     \item objective: minimize the max number of maintenances.
    \end{itemize}

    \subsection{Notations}

    \textbf{Sets}

    \begin{tabular}{ll}
    $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
    $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
    \end{tabular}

    \vskip 0.3cm

    \subsection{Parameters}

    \begin{tabular}{ll}
        $\mathcal{A}_{it}$ & Pattern $i$ has a maintenance in $t$.\\
    \end{tabular}

    \vskip 0.3cm

    Here, A needs to be created taking into account:
    1. The range of start of maintenances.
    2. The fixed duration of maintenances.
    3. The minimal distance between maintenances.

    \subsection{Variables}

    \begin{itemize}
     \item $x_i$: number of patterns of type $i$ to be included in the solution.
     \item $M$: maximum number of maintenance among all periods.
    \end{itemize}

    \subsection{Model}

    \begin{align}
        & \text{Min}\; M
    \end{align}

    \begin{align}
        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \leq M & t \in \mathcal{T} \\
        & x_{i} \geq 0,\; integer \\
    \end{align}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Scheduling variant 2}

    As modelled in Rieck2012.

    \subsection{Hypothesis}

    \begin{itemize}
     \item fixed maintenance duration.
     \item min distance between maintenances that.
     \item each maintenance uses an amount of resource.
     \item objective: minimize the max number of maintenances.
    \end{itemize}

      \subsection{Notations}

    \noindent{
    \begin{tabular}{ll}
        $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
        $(i, j) \in \mathcal{A} = [\![1, v]\!]$ & Set of consecutive (i, j) maintenance operations.\\
        $i = 0$ & Artificial starting node.\\
        $i = n+1$ & Artificial ending node.\\
    \end{tabular}
    }

    \vskip 0.3cm

    The set $\mathcal{A}$ we will construct it by deciding in advance the maintenances $i$ that share the same aircraft.

    \subsection{Parameters}

    \begin{tabular}{ll}
        $p_{i}$ & Duration of maintenance $k$ of resource $i$.\\
        ${\delta}_{ij}$ & Minimum duration between two maintenances $i$ and $j$.\\
        $\mathbf{d}$ & Last period on the horizon.\\
        $r_{i}$ & Resources taken by maintenance $i$.\\
        % $\mathbf{{To}}_{i,k}$ & Minimum starting period for maintenance $k$ of resource $i$.\\
        % $\mathbf{{Ta}}_{i,k}$ & Maximum starting period for maintenance $k$ of resource $i$.\\
    \end{tabular}

    \subsection{Variables}

    \begin{itemize}
     \item $S_i$: the start of maintenance operation $i$.
    \end{itemize}

    \subsection{Model}

    $f(S)$ needs to be defined still. But includes a way to smooth the use of resoures by the activities.

    \begin{align}
        f(S)
    \end{align}

    \begin{align}
        & S_j - S_i \geq \delta_{ij} & (i, j) \in \mathcal{A} \\
        & S_0 = 0 & \\
        & S_{n+1} = \mathbf{d}  \\
        & S_{i} \geq 0, i \in \mathcal{V} \\
    \end{align}

\end{document}

### 2018-05-02 latex/notes/FMP_variants.tex
Add new bibliographie and notes on models. And CLAIO summary.

@@
 \usepackage{amssymb}
+\usepackage{chngcntr}
+
+\counterwithin*{equation}{section}
+\counterwithin*{equation}{subsection}
 
@@
 \section{Crew scheduling variant}
+
+    As presented in \cite{Brunner2013}.
 
@@
     \begin{align}
-        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i
+        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i \label{numPatterns}
     \end{align}
@@
     \begin{align}
-        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq d_j & t \in \mathcal{T} \\
+        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq d_j & t \in \mathcal{T} \label{demand} \\
         & x_{i} \geq 0,\; integer \\
     \end{align}
+
+    (\ref{numPatterns}) minimizes the number of workers. (\ref{demand}) defines the need to have at least a given number of workers at any given time.
 
@@
     \begin{align}
-        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq 1 & t \in \mathcal{T} \\
+        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq 1 & t \in \mathcal{T} \label{demand} \\
         & x_{i} \geq 0,\; binary \\
     \end{align}
+
+    (\ref{numPatterns}) minimizes the number of workers. (\ref{demand}) defines the need to have at least one worker working at any given time.
 
@@
     Here, A needs to be created taking into account:
-    1. The range of start of maintenances.
-    2. The fixed duration of maintenances.
-    3. The minimal distance between maintenances.
+
+    \begin{enumerate}
+        \item The range of start of maintenances.
+        \item The fixed duration of maintenances.
+        \item The minimal distance between maintenances.
+    \end{enumerate}
 
@@
     \begin{align}
-        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \leq M & t \in \mathcal{T} \\
+        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \leq M & t \in \mathcal{T} \label{numMaint} \\
         & x_{i} \geq 0,\; integer \\
     \end{align}
+
+        (\ref{numMaint}) defines the $M$ variable as the maximum number of maintenances at any period of time.
 
@@
 
-    As modelled in Rieck2012.
+    As modelled in \cite{Rieck2012} and \cite{Neumann2000}.
 
@@
      \item fixed maintenance duration.
-     \item min distance between maintenances that.
-     \item each maintenance uses an amount of resource.
+     \item min and max durations between maintenances that belong to the same aircraft.
+     \item each maintenance uses an amount of resource. In this case 1.
      \item objective: minimize the max number of maintenances.
+     \item There are two artificial activities representing the start and end of the horizon.
     \end{itemize}
@@
 
+    \textbf{Sets}
+
     \noindent{
@@
         $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
-        $(i, j) \in \mathcal{A} = [\![1, v]\!]$ & Set of consecutive (i, j) maintenance operations.\\
+        $(i, j) \in \mathcal{E} = [\![1, v]\!]$ & Set of consecutive (i, j) maintenance operations, i.e. consecutive maintenances belonging to the same aircraft.\\
+        $k \in \mathcal{R}$ & Types of resources that maintenance can take. We only have one type.\\
         $i = 0$ & Artificial starting node.\\
@@
     \begin{tabular}{ll}
-        $p_{i}$ & Duration of maintenance $k$ of resource $i$.\\
-        ${\delta}_{ij}$ & Minimum duration between two maintenances $i$ and $j$.\\
+        $p_i$ & Duration of maintenance $i$.\\
         $\mathbf{d}$ & Last period on the horizon.\\
-        $r_{i}$ & Resources taken by maintenance $i$.\\
+        $r_{ik}$ & Resources of type $k$ taken by maintenance $i$. In this case it's always 1 per maintenance.\\
+        $d^{max}_{ij}$ & Maximum duration between two maintenances $i$ and $j$.\\
+        $d^{min}_{ij}$ & Minimum duration between two maintenances $i$ and $j$.\\
+        ${\delta}_{ij}$ & Precedence duration arcs between two maintenances $i$ and $j$ based on $d^{max}$ and $d^{min}$.\\
+        $R_k$ & total maintenance capacity at each period for resource $k$.\\
+        $c_k$ & cost per unit of investment on resource $k$.
         % $\mathbf{{To}}_{i,k}$ & Minimum starting period for maintenance $k$ of resource $i$.\\
@@
 
-    \begin{itemize}
-     \item $S_i$: the start of maintenance operation $i$.
-    \end{itemize}
+    \begin{tabular}{ll}
+        $S_i$ & the start of maintenance operation $i$.\\
+        $\mathcal{A}(S^{\mathcal{C}}, t)$ & set of maintenance operations active at moment $t$.\\
+    \end{tabular}
 
@@
 
-    $f(S)$ needs to be defined still. But includes a way to smooth the use of resoures by the activities.
+    The following formulation is taken from \cite{Neumann2000}.
 
     \begin{align}
-        f(S)
+        \min f(\mathbf{R}(S)) = \sum_{k \in \mathcal{R}} c_k \max_{t \in \mathcal{T}} r_k(S^{\mathcal{C}}, t)
     \end{align}
@@
     \begin{align}
-        & S_j - S_i \geq \delta_{ij} & (i, j) \in \mathcal{A} \\
-        & S_0 = 0 & \\
-        & S_{n+1} = \mathbf{d}  \\
-        & S_{i} \geq 0, i \in \mathcal{V} \\
+        & S_j - S_i \geq \delta_{ij} & (i, j) \in \mathcal{E}  \label{precedence}\\
+        & S_0 = 0 &  \label{firstjob}\\
+        & S_{n+1} \leq \mathbf{d}   \label{lastjob}\\
+        & S_{i} \in \mathbb{Z}_{\geq 0} & i \in \mathcal{V}  \label{integervars}\\
+        & r_k(S^{\mathcal{C}}, t) = \sum_{i \in \mathcal{A}(S^{\mathcal{C}}, t)} r_{ik}  \label{consumptiondef}\\
+        & r_k(S^{\mathcal{C}}, t) \leq R_k \label{consumptionmax}\\
     \end{align}
 
+    (\ref{precedence}) constraints define the relative minimum and maximum starting times of tasks.
+    (\ref{firstjob}) and (\ref{lastjob}) constraints define the fictitious start and end jobs.
+    (\ref{integervars}) define the range for the starting time variables: integer positive.
+    (\ref{consumptiondef}) defines the used quantity for each type of resource depending on the starting times and durations and (\ref{consumptionmax}) defines the maximum amount of resources of each type that can be used at any given time.
+
+% As it has been shown by Bartusch et al. (1988),
+% the feasible region S (without the integrality
+% constraints S i 2 Z, i 2 V ) of R   is generally dis-
+% connected and represents the union of (convex)
+% polytopes whose number grows exponentially in n.
+% The decision problem whether or not S 6 ;
+
+% \bibliographystyle{dcu}
+\bibliographystyle{alpha}
+\bibliography{./../biblio/Scheduling,./../biblio/NRP}
+
 \end{document}

### 2018-05-17 latex/notes/FMP_variants.tex
add new notes to variants
biblio and complexity analysis

@@
 
-    As presented in \cite{Brunner2013}.
+    As presented in \cite{Brunner2013}. There, they prove the complexity by comparing the crew scheduling problem with the circulant problem. Below I do the same with our problem.
 
@@
 
+    \subsection{Reductions}
+
+    \begin{itemize}
+        \item The missions consume 0 hours of aircraft.
+        \item The minimum time between maintenances is equal to the maximum.
+        \item All missions can use any aircraft.
+    \end{itemize}
+
     \subsection{Notations}
@@
     \begin{tabular}{ll}
-    $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
+    $t \in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
     $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
@@
     \begin{tabular}{ll}
-        $D_{off}$ & Number of consecutive non-working days. \\
-        $D_{on}$ & Number of consecutive working days. \\
+        $D^{off}$ & Number of consecutive non-working days. \\
+        $D^{on}$ & Number of consecutive working days. \\
         $\mathcal{A}_{it}$ & Pattern $i$ covers period $t$ (circulant matrix built from $D_{on}$ and $D_{off}$).\\
@@
         & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq d_j & t \in \mathcal{T} \label{demand} \\
-        & x_{i} \geq 0,\; integer \\
+        & x_{i} \geq 0,\, x_{i} \in \mathbb{Z} \\
     \end{align}
@@
         & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq 1 & t \in \mathcal{T} \label{demand} \\
-        & x_{i} \geq 0,\; binary \\
+        & x_{i} \in \{0, 1\} \\
     \end{align}
@@
      \item each maintenance uses an amount of resource. In this case 1.
+     \item there is a maintenance capacity to respect at every period.
      \item objective: minimize the max number of maintenances.
      \item There are two artificial activities representing the start and end of the horizon.
+    \end{itemize}
+
+    \subsection{Reductions}
+
+    \begin{itemize}
+        \item The missions consume 0 hours of aircraft.
+        \item The missions require 0 aircraft at each period.
+    \end{itemize}
+
+    \subsection{Limitations}
+
+    \begin{itemize}
+        \item Resource usage is always 1.
+        \item There is only one type of resource.
+        \item All durations are the same.
     \end{itemize}
@@
     \begin{tabular}{ll}
-        $p_i$ & Duration of maintenance $i$.\\
+        $p_i$ & Duration of maintenance $i$. In this case it's always p=6.\\
         $\mathbf{d}$ & Last period on the horizon.\\
-        $r_{ik}$ & Resources of type $k$ taken by maintenance $i$. In this case it's always 1 per maintenance.\\
+        $r_{ik}$ & Resources of type $k$ taken by maintenance $i$. In this case it's always 1.\\
         $d^{max}_{ij}$ & Maximum duration between two maintenances $i$ and $j$.\\
@@
 
-% As it has been shown by Bartusch et al. (1988),
-% the feasible region S (without the integrality
-% constraints S i 2 Z, i 2 V ) of R   is generally dis-
-% connected and represents the union of (convex)
-% polytopes whose number grows exponentially in n.
-% The decision problem whether or not S 6 ;
+    \subsection{Possible prove}
+
+    \cite{Blazewicz1983} proved that the following problem is NP-Hard:
+
+    $$P2 | res111, chain, p_j = 1 / C_{max}$$ is NP-hard via the 3 partition problem.
+
+    This problem is, from what I understand, the following:
+
+    \begin{itemize}
+        \item The number of machines is 2.
+        \item There is only one type of resource.
+        \item The resource size is 1.
+        \item The consumption of the resource by each job is upper bounded by 1.
+        \item The duration of each job is 1.
+        \item There is a chain precedence between jobs: a job can have at most one predecessor and one successor.
+    \end{itemize}
+
+    Here they use the machine and resource concept.
+This could be understood as there being TWO resource types.
+Not sure how to really come around this.
+Is the same, for example, as doing the following:
+
+    $$P| res1R1, chain, p_j = 1 / C_{max}$$ Where I have an unlimited number of machines and a quantity R of available resource?
+
+    Although they minimize the maximum completion time, the decision problem 'is there a planning that uses X number of resources and finishes before period Y' is the same as ours.
+When we pass to the optimization problem we just choose the number of resources instead of the time.
+
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-05-24 R/diagrams/heuristic/heuristic.Rmd
R diagrams for heuristic

--- 2018-02-14 R/diagrams/heuristic.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+source("heuristic.R")
+source("../../functions/params.R")
+
+knitr::opts_chunk$set(echo = FALSE)
+path <- paste0(PATHS['r_project'], 'diagrams/heuristic/')
 ```
@@
                 2. After assigning a task: we update the remaining and elapsed times for the aircraft.
+            1. If we tagged the aircraft as in need of maintenances: we update.
 
@@
 
-    if the aircraft is tagged as 'in need of a maintenance':
     1. we check if we can reserve 6 continous months of maintenance.
@@
 
-<!-- <br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br><br> -->
-## Main python code
+## Some graphs
 
-I think the code gives an idea of the logic behind it. I only include the code for each mission so we would have to execute this function for each mission (or task).
+### Main loop
 
-`task` is the mission to satisfy.
+```{r, echo=FALSE}
+graph_solve(paste0(path, 'heur_solve.csv'), height=500)
+```
 
-* **candidates**: a list with the universe of aircraft that can be assigned to that task.
-* **rem_resources**: remaining aircraft that need to be assigned at each period of the task's duration.
-* **periods_task**: intersection of available slots for the aircraft (periods without taks or maintenance) and the task's period's needs.
+This loop just goes over each task and tries to fill each one. Particularly important is the order in which we choose to assign the tasks.
 
-```python
+### Fill mission
 
-def fill_mission(self, task):
-    dtype = 'U7'
-    # get task candidates: we copy the list.
-    candidates = self.instance.data['tasks'][task]['candidates'][:]
-    # get task periods to satisfy:
-    rem_resources = \
-        aux.dicttup_to_dictdict(
-            self.instance.get_task_period_needs()
-        )[task]
+```{r, echo=FALSE}
+graph_fill_mission(paste0(path, 'heur_fill_mission.csv'), height=700)
+```
 
-    while len(rem_resources) > 0 and len(candidates) > 0:
-        # candidate = 'A124'
-        candidate = candidates[0]
-        # get free periods for candidate
-        periods_task = \
-            self.get_free_starts(candidate, np.fromiter(rem_resources, dtype=dtype))
+This is the most important loop. It gets a task as input.
+It iterates over the candidates, finding available periods to assign the task.
+It assigns all possible periods to the task.
+At the end, it updates the remaining used and elapsed times.
+If it finds a period where the candidate cannot get a task because of flight hours or calendar, it marks the candidate as "in need of maintenance" and tries to assign a maintenance in the correct spot or as soon as possible.
 
-        # consider eliminating the resource from the list.
-        # if all its periods are 'used'
-        if len(periods_task) == 0:
-            candidates.pop(0)
-            continue
-        print('candidate={}\ntask={}\nperiods={}'.format(candidate, task, periods_task))
-        maint_period = ''
-        # TODO: this is still wrong:
-        for start, end in periods_task:
-            for period in aux.get_months(start, end):
-                result = self.check_set_task(candidate, period, task)
-                if not result:
-                    maint_period = period
-                    break
-                rem_resources[period] -= 1
-            self.update_time_usage(candidate, period, self.get_next_maintenance(candidate, period), 'rut')
+### Assign maintenance
 
-        if maint_period != '':
-            # find soonest period to start maintenance:
-            maint_start = self.get_soonest_maint(candidate, maint_period)
-            if maint_start is None:
-                # this means that we cannot assign tasks BUT
-                # we cannot assign maintenance either :/
-                # we should then take out the candidate:
-                candidates.pop(0)
-                continue
-            maint_end = aux.shift_month(maint_start, 5)
-            periods_maint = aux.get_months(maint_start,  maint_end)
-            for period in periods_maint:
-                self.set_maint(candidate, period)
-            self.update_time_maint(maint_start, maint_end)
-            next_month = aux.get_next_month(maint_end)
-            self.update_time_usage(candidate, next_month, 'ret')
-            self.update_time_usage(candidate, next_month, 'rut')
+```{r, echo=FALSE}
+graph_find_assign_maintenance(paste0(path, 'heur_maints.csv'), height=700)
+```
 
-```
+This loop takes as input a resource and a period in need of maintenance.
+With this information, it tries to find the soonest possible period to get the maintenance.
+Sometimes it doesn't exist, sometimes it is too late.
+In this case it exits.
+
+If not, it could be the case that there is a maintenance already scheduled around this dates.
+In this case, it tries to advance (move) the maintenance in order to do it sooner.
+If it fails the swap, then it does not assign any maintenance.
+
+Finally, if it's possible it assign the maintenance and updates the remaining used and elapsed times for the periods of the maintenance and the following ones.

### 2018-05-24 R/functions/params.R

@@ -9,4 +9,5 @@ PATHS = list(
     ,experiments = path_results %>% paste0("experiments/")
     ,img = path_root %>% paste0("OPTIMA/img/")
     ,latex = path_root %>% paste0("OPTIMA/latex/")
+    ,r_project = path_root %>% paste0("OPTIMA/R/")
 )

### 2018-05-29 R/OptiPlan_presentation/proposal_fr.Rmd
Added ret analysis and heuristic explanation.

--- 2018-03-20 R/OptiPlan_presentation/proposal_fr.Rmd
@@
 source("diagrams/functionalities.R")
-source('diagrams/exampleDT.R')
-source('diagrams/gantt_tasks.R')
+source('functions/exampleDT.R')
+source('functions/gantt_tasks.R')
 exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
@@
 Afin de récupérer des heures de vol (environ 1200 heures), une opération de maintenance est nécessaire.
-Après un certain temps, les aéronefs ont besoin d’un entretien même s’ils n’ont pas atteint leur limite d’heures de vol. Ce temps est d’environ 5 ans.
+Après un certain temps, les aéronefs ont besoin d’un entretien même s’ils n’ont pas atteint leur limite d’heures de vol.
+Ce temps est d’environ 5 ans.
 

### 2018-05-29 R/diagrams/heuristic/heuristic.Rmd
Added ret analysis and heuristic explanation.

@@
 ---
-title: "heuristic1"
-author: "Franco Peschiera"
+title: "Greeedy heuristic"
 date: "January 9, 2018"
@@
     html_document:
-        # css: ./../../css/baobab.css
+        toc: yes
+        css: ./../../../css/baobab_docs.css
 ---
@@
 source("heuristic.R")
-source("../../functions/params.R")
+functions_dir <- '../../functions/'
+
+source(paste0(functions_dir, "params.R"))
+source(paste0(functions_dir, 'import_results.R'))
+
+exp_directory = PATHS[['experiments']] %>% paste0('201805241334/')
 
 knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(message = FALSE)
+knitr::opts_chunk$set(warning = FALSE)
 path <- paste0(PATHS['r_project'], 'diagrams/heuristic/')
@@
 
-With the candidate and the task we do the following:
-
     for every task:
-        1. calculate candidates.
-        for each candidate until we run out of candidates or we satisfy task's needs:
-            1. find available periods.
-            Until we run out of periods:
-                1. we check if we can assign it (checking remaining time and elapsed time). 
-                    * If we can't: we tag the aircraft as "in need of a maintenance'.
-                    * if we can: we assign the task.
-                2. After assigning a task: we update the remaining and elapsed times for the aircraft.
-            1. If we tagged the aircraft as in need of maintenances: we update.
+        Calculate candidates.
+        Repeat:
+            Sort candidates.
+            Get first candidate
+            Find available periods.
+            Repeat:
+                Get first period.
+                Can we assign task?
+                    no: we tag it "in need of maintenance'.
+                    yes: we assign the task.
+                After assigning a task: we update the counters.
+            If we tagged the aircraft as in need of maintenances: 
+                Execute maintenance assignment
 
@@
 
-    1. we check if we can reserve 6 continous months of maintenance.
-        * if we can: we do it (and continue assigining tasks to the aircraft).
-        * if we cannot: we go to the next candidate.
-    2. After assigning a maintenance: we update the remaining and elapsed times for the aircraft.
+    good candidate slot maintenance?
+        yes: we give maintenance in slot.
+        no: we go to the next candidate.
+    if maintenance was assigned:
+        Update the remaining and elapsed times for the aircraft.
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 graph_solve(paste0(path, 'heur_solve.csv'), height=500)
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE}
 graph_fill_mission(paste0(path, 'heur_fill_mission.csv'), height=700)
@@
 Finally, if it's possible it assign the maintenance and updates the remaining used and elapsed times for the periods of the maintenance and the following ones.
+
+## Example planning
+
+```{r, echo=FALSE}
+print_solution(exp_directory)
+```

### 2018-05-29 R/functions/import_results.R

@@ -84,7 +84,6 @@ get_ret <- function(exp_directory, time='used'){
     }
     solution <- read_json(solution_path)
     input <- read_json(input_path)
-    # input$resources[[1]] %>% names
     initial <- sapply(input$resources, '[[', 'initial_used')
     max_rem <- input$parameters[['max_used_time']]
     previous_month <- input$parameters$start %>% paste0('-01') %>% as_date %>% subtract(ddays(1)) %>% floor_date('month') %>% format('%Y-%m')
@@ -207,7 +206,7 @@ print_solution <- function(exp_directory, max_resources=NULL, ...){
 
     states <- get_states(exp_directory)
     
-    timevis_from_states(states, max_resources=NULL, ...)
+    timevis_from_states(states, max_resources=max_resources, ...)
 }
 
 if (FALSE){

### 2018-05-29 R/scripts/graphs.R

@@ -1,11 +1,10 @@
 source('functions/params.R')
 source('functions/import_results.R')
 
-exp_directory = PATHS[['experiments']] %>% paste0('201805031113/')
+exp_directory = PATHS[['experiments']] %>% paste0('201805241334/')
 print_solution(exp_directory)
 
-
-    
+ 
 #       id   start     end   content group                                         style
 #    <int>   <chr>   <chr>    <fctr> <chr>                                         <chr>
 #  1     1 2015-01 2016-01      <NA>  D601               background-color:;border-color:

### 2018-05-29 R/scripts/ret_analysis.R

@@ -1,0 +1,68 @@
+library(ggplot2)
+library(magrittr)
+source('functions/params.R')
+source('functions/import_results.R')
+
+exp_directory = PATHS[['experiments']] %>% paste0('201805241334/')
+tasks <- get_tasks(exp_directory)
+params <- get_parameters(exp_directory)
+params <- params$value %>% magrittr::set_names(params$name)
+month_range <- seq.Date(from=params['start'] %>% paste0('-01') %>% as_date, 
+                        to=params['end'] %>% paste0('-01') %>% as_date,
+                        by="months")
+task_period <- 
+    tasks %>% 
+    mutate_at(c('start', 'end'), function(x) x %>% paste0('-01') %>% as_date) %>% 
+    merge(data.table(period=month_range)) %>% 
+    filter(start<=period & end>= period) %>% 
+    group_by(period, type_resource) %>% 
+    summarise_at(c('num_resource', 'consumption'), sum)
+
+input_path = exp_directory %>% paste0('data_in.json')
+input <- read_json(input_path)
+rut_init <- sapply(input$resources, '[[', 'initial_used')
+ret_init <- sapply(input$resources, '[[', 'initial_elapsed') %>% data.frame(ret_init=., resource=names(.), stringsAsFactors = F)
+type_resource <- sapply(input$resources, '[[', 'capacities') %>% sapply('[[', 1) %>% data.frame(type_resource=., resource=names(.), stringsAsFactors = F)
+
+tot_resource <- type_resource %>% group_by(type_resource) %>% summarise(tot_resource=n())
+
+maintenances <- 
+    lapply(
+        0:4,
+        function(x) {
+            data.table(period=month_range) %>% mutate(period_maint = period %m+% months(x))
+        }
+    ) %>% bind_rows() %>% arrange(period)
+
+ret_info <- 
+    ret_init %>% 
+    inner_join(type_resource) %>% 
+    mutate(start=params['start'] %>% paste0('-01') %>% as_date,
+           period = start %m+% months(ret_init - 1)) %>% 
+    group_by(type_resource, period) %>% 
+    summarise(maint_ret=n()) %>% 
+    inner_join(maintenances) %>% 
+    mutate(period = period_maint) %>% 
+    group_by(type_resource, period) %>% 
+    summarise_at(c('maint_ret'), sum)
+
+resource_avail <- 
+    task_period %>% 
[…]

### 2018-06-01 R/Dassault_presentation/ppt_v2_fr.Rmd
added dassault ppt, heuristic documentation

---
title: "Planification de maintenances des aéronefs militaires"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa
institute: "ISAE SUPAERO"
date: "June 06, 2018"
output: 
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    keep_tex: true
    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
    # includes:
    #     in_header: header.tex
  slidy_presentation:
    css: ./../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    slide_level: 2
    toc: true
nocite: '@*'
bibliography: MFMP.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("../")
source('functions/params.R')
source('functions/import_results.R')
source('diagrams/thesis_gantt.R')
exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
gantt <- print_solution(exp_directory, max_resources = 30)
gantt_th <- gantt_thesis()
show_gantt <- knitr::opts_knit$get("rmarkdown.pandoc.to")=="slidy"
```

# Introduction

## Planification Thèse
<!-- |Phase  | Libellé                        | Mois     | -->
<!-- |-------|--------| --------| -->
<!-- |1      | État de l’art FMP              | 1 – 8    | -->
<!-- |       | Travaux antérieurs             | 1 – 8    | -->
<!-- |2      | Nouveaux modèles mathématiques | 9 – 16   | -->
<!-- |3      | Incertitude                    | 17 – 24  | -->
<!-- |4      | Cas d’étude Armée de l’Air     | 25 – 30  | -->
<!-- |5      | Prototype logiciel             | 25 – 30  | -->
<!-- |6      | Mémoire de thèse               | 31 – 36  | -->

```{r, echo=FALSE, out.width='100%'}

if (show_gantt){
    # htmltools::HTML('<div style="font-size: 20px;width:100%">')
    gantt_th
    # htmltools::HTML("</div>")
} else {
    knitr::include_graphics('./../../img/gantt_thesis.png')
}

```

# Problème

## Avions

Une flotte hétérogène d'avions avec les caractéristiques suivantes:

* Temps écoulé restant.
* Des heures de vol restant.
* Fonctionnalités.

## Missions

Un ensemble de missions à réaliser, chaque une avec:

* Dates de début et fin.
* Un besoin mensuelle d'avions.
* Une consommation d'heures par avion-mois affecté.
* Des fonctionnalités requise.

## Maintenances

Chaque maintenance suit les règles suivantes:

* La *durée est de 6 mois* consécutifs.
* Le nombre maximal *d'heures de vol* entre deux maintenances est *1000h*.
* Le nombre maximal de *mois calendrier* entre deux maintenances est *60 mois*.

## États possibles

En résumé, voici les états logiques possibles d'un aéronef:

- En mission.
- En maintenance.
- En stockage.
- Disponible.

## Objectifs

* Maximiser la disponibilité.
* Minimiser le coût (réduisant le nombre de maintenances).

## Problème (exemple)

```{r, echo=FALSE, out.width='100%'}

if (show_gantt){
    # htmltools::HTML('<div style="font-size: 20px;width:100%">')
    gantt
    # htmltools::HTML("</div>")
} else {
    knitr::include_graphics('./../../img/calendar.png')
}

```

## Contraintes /  Règles / Objectifs

* RQ001: heures de missions
* RQ002: aéronefs assignés
* RQ003: besoins de maintenance
* RQ004: durée de maintenance
* RQ005: disponibilité
* RQ006: capacité de maintenance
* RQ007: état final
* RQ008: stockage

# État de l'art

## État de l'art (1)

* FMP: Flight and Maintenance Planning problem.
* En @Cho2011, avions de l'Armée de l'air états-unienne ont été affectés à des opérations quotidiens pour minimiser le nombre maximal de maintenances.
* En @Kozanidis2008, des heures de vol et des maintenances ont été affectées aux avions de la flotte grecque.
* En @Verhoeff2015, différents objectifs ont été essayé pour les affectations d'heures de vol et maintenances.

## État de l'art (2)

Les travaux précédents:

* ont étudié de flottes homogènes: nous avons de flottes hétérogènes.
* n'ont pas le mêmes contraintes (stockages, 60 périodes entre maintenances) ou les mêmes objectifs (maximiser disponibilité).
* n'affectent pas les avions aux missions.

# Modèle
## Modèle: fonction objectif

<div class="incremental">

\begin{align}
    & \text{Min}\; W_1 m_{max} + W_2 u_{max}
\end{align}

\pause

Minimiser le nombre maximal de maintenances et le nombre maximal de non-disponibilités.

\begin{align}
    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
    &t \in \mathcal{T} \\
    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max}
    &t \in \mathcal{T}
\end{align}

</div>

## Modèle: contraintes principales

Besoins des missions et incompatibilités entre les missions et les maintenances.

\begin{align}
&\sum_{i \in \mathcal{I}_j} a_{jti} = R_j
&j \in \mathcal{J}, t \in \mathcal{T}_j\\
&\sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1
& t \in \mathcal{T}, i \in \mathcal{I}
\end{align}

## Modèle: contraintes de flux

Calcule d'heures de vol restantes pour chaque avion:

\begin{align}
& rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
& rut_{i0} = Rut^{Init}_i
        & i \in \mathcal{I}\\
& rut_{it} \geq H m_{it}
        & t \in \mathcal{T}, i \in \mathcal{I}\\
& rut_{it} \in [0,H]
        & t \in \mathcal{T}, i \in \mathcal{I} \\
& \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
        & t = |\mathcal{T}|
\end{align}

# Résultats

## Expériences

Des instances ont déjà été testés:

| id | $\|\mathcal{J}\|$ |$\|\mathcal{T}\|$ |  assign |  objective  |  time (s) |   bound |
|-----|------------------|------------------|---------| ------------| -----------| --------|
| I_0 |                9 |               11 |     310 |        62.0 |        0.7 |    62.0 |
| I_1 |                9 |               21 |     650 |        63.0 |       68.7 |    63.0 |
| I_2 |                9 |               31 |     990 |        63.0 |     3600.1 |    62.0 |
| I_3 |                9 |               41 |    1249 |        64.0 |     3603.9 |    61.7 |
| I_4 |               10 |               11 |     530 |        82.0 |        0.9 |    82.0 |
| I_5 |               10 |               21 |    1070 |        83.0 |      144.0 |    83.0 |
| I_6 |               10 |               31 |    1610 |        83.0 |     3600.1 |    82.0 |
| I_7 |               10 |               41 |    2069 |        84.0 |     3609.1 |    81.8 |
| I_8 |               11 |               11 |    1080 |       139.0 |      530.6 |   139.0 |
| I_9 |               11 |               21 |    2120 |       149.0 |     3600.0 |   139.9 |

# Prochains pas
## Prochains pas

* **L'amélioration du modèle mathématique**: casser les symétries, reformuler les objectifs.
* **Solutions techniques alternatives**: essayer d'autres méthodes de résolution.
* **Continuer le dialogue avec les utilisateurs**: ajouter des nouvelles contraintes, valider les résultats.

# Références
## Références / Questions?

### 2018-06-01 R/diagrams/heuristic/heuristic_2.Rmd
added dassault ppt, heuristic documentation

rut = remaining usage time = remaining flight hours.
ret = remaining elapsed time = remaining periods.

------

**The base logic** (probably similar to what's in the students' PIE code):

Sort *periods* from lowest to greatest.
For each period:
    Get *tasks* that need at least one aircraft in that period.
    Sort *tasks* according to the number of candidates they have.
    For each task in *tasks*:
        Get available *candidates* that can do the task.
        Sort *candidates* according to 'tie methodology' (largest $rut / ret$ first).
        For each candidate:
            Try to assign candidate to task in period.
                If possible: update the candidate's rut and task remaining needs.
                If not possible: give a maintenance to the candidate.
            if no more remaining needs => go to the next task.

**The additional logic** (that should be executed at the end of the each period loop):

We simulate the next X (10?) periods only by using ret to generate maintenances (since we're using the 'tie methodology', this could be "good enough").

if this simulation:

* shows periods with too many maintenances: 
    * we try to advance some of those maintenances to the present. 
    * (How many?) The number of maintenances to advance may depend on how many maintenances we can actually do and how many months we can move the maintenances.
* shows periods with non-availabilities for missions: 
    * we try to advance some maintenances in that period from an aircraft with the same capabilities to the present. 
    * (How many?) This can be the same logic than the one before.

Other comments on these 'moves':

* We should first try to advance the aircraft that have the least ret.
* We need to check that the changed periods continue to be feasible.
* The changed periods are the ones at each extreme of the maintenance and the number will be two times the size of the move.

### 2018-06-01 R/roadef_presentation/roadef_presentation.Rmd
added dassault ppt, heuristic documentation

@@
 setwd("../")
-source('scripts/params.R')
-source('scripts/import_results.R')
+source('functions/params.R')
+source('functions/import_results.R')
 exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')

### 2018-06-01 R/diagrams/thesis_gantt.R

@@ -1,0 +1,36 @@
+library(tidyverse)
+library(data.table)
+library(timevis)
+library(lubridate)
+
+gantt_thesis <- function(...){
+    start_month <- as.Date('2017-10-01')
+    
+    data <- 
+        data.table(
+            Phase=c(1, 1, 2, 3, 4, 5, 6)
+            ,Libelle=c('État de l’art FMP','Travaux antérieurs','Nouveaux modèles',
+                       'Incertitude','Cas d’étude','Prototype logiciel','Mémoire de thèse')
+            ,month_start=c(1, 1, 9, 17, 25, 25, 31)
+            ,month_end= c(8, 8, 16, 24, 30, 30, 36)
+        ) %>% 
+        mutate(start= start_month %m+% months(month_start-1),
+               end= start_month %m+% months(month_end),
+               group= Phase,
+               id= c(1:nrow(.)),
+               content = Libelle,
+               color = RColorBrewer::brewer.pal(5, "YlOrRd")[rep_len(1:5, n())],
+               style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color),
+        )
+    config <- list(
+        editable = TRUE,
+        align = "center",
+        orientation = "top",
+        snap = NULL,
+        margin = 0,
+        zoomable= FALSE
+    )
+    
+    groups <- data %>% distinct(group) %>% rename(id= group) %>% mutate(content= id)
+    timevis(data, groups= groups, options= config, ...)    
+}

### 2018-06-01 R/scripts/graphs.R

@@ -4,7 +4,7 @@ source('functions/import_results.R')
 exp_directory = PATHS[['experiments']] %>% paste0('201805241334/')
 print_solution(exp_directory)
 
- 
+
 #       id   start     end   content group                                         style
 #    <int>   <chr>   <chr>    <fctr> <chr>                                         <chr>
 #  1     1 2015-01 2016-01      <NA>  D601               background-color:;border-color:

### 2018-06-22 R/Dassault_presentation/Dassault_20180606_v2.tex
new documents and presnetations or re-organised documents

\documentclass[ignorenonframetext,]{beamer}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
\usetheme[]{Berlin}
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\newif\ifbibliography
\hypersetup{
            pdftitle={Planification de maintenances des aéronefs militaires},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{caption}
% These lines are needed to make table captions work with longtable:
\makeatletter
\def\fnum@table{\tablename~\thetable}
\makeatother

% Prevent slide breaks in the middle of a paragraph:
\widowpenalties 1 10000
\raggedbottom

\AtBeginPart{
  \let\insertpartnumber\relax
  \let\partname\relax
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \let\insertsectionnumber\relax
    \let\sectionname\relax
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \let\insertsubsectionnumber\relax
  \let\subsectionname\relax
  \frame{\subsectionpage}
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

\title{Planification de maintenances des aéronefs militaires}
\author{Franco Peschiera, Alain Haït, Olga Battaïa}
\institute{ISAE SUPAERO}
\date{June 06, 2018}

\begin{document}
\frame{\titlepage}

\begin{frame}
\tableofcontents[hideallsubsections]
\end{frame}

\section{Introduction}\label{introduction}

\begin{frame}{Planification Thèse}

\includegraphics[width=1\linewidth]{./../../img/gantt_thesis}

\end{frame}

\section{Problème}\label{probleme}

\begin{frame}{Avions}

Une flotte hétérogène d'avions avec les caractéristiques suivantes:

\begin{itemize}[<+->]
\tightlist
\item
  Temps écoulé restant.
\item
  Des heures de vol restant.
\item
  Fonctionnalités.
\end{itemize}

\end{frame}

\begin{frame}{Missions}

Un ensemble de missions à réaliser, chaque une avec:

\begin{itemize}[<+->]
\tightlist
\item
  Dates de début et fin.
\item
  Un besoin mensuelle d'avions.
\item
  Une consommation d'heures par avion-mois affecté.
\item
  Des fonctionnalités requise.
\end{itemize}

\end{frame}

\begin{frame}{Maintenances}

Chaque maintenance suit les règles suivantes:

\begin{itemize}[<+->]
\tightlist
\item
  La \emph{durée est de 6 mois} consécutifs.
\item
  Le nombre maximal \emph{d'heures de vol} entre deux maintenances est
  \emph{1000h}.
\item
  Le nombre maximal de \emph{mois calendrier} entre deux maintenances
  est \emph{60 mois}.
\end{itemize}

\end{frame}

\begin{frame}{États possibles}

En résumé, voici les états logiques possibles d'un aéronef:

\begin{itemize}[<+->]
\tightlist
\item
  En mission.
\item
  En maintenance.
\item
  En stockage.
\item
  Disponible.
\end{itemize}

\end{frame}

\begin{frame}{Objectifs}

\begin{itemize}[<+->]
\tightlist
\item
  Maximiser la disponibilité.
\item
  Minimiser le coût (réduisant le nombre de maintenances).
\end{itemize}

\end{frame}

\begin{frame}{Problème (exemple)}

\includegraphics[width=1\linewidth]{./../../img/calendar}

\end{frame}

\begin{frame}{Contraintes / Règles / Objectifs}

\begin{itemize}[<+->]
\tightlist
\item
  RQ001: heures de missions
\item
  RQ002: aéronefs assignés
\item
  RQ003: besoins de maintenance
\item
  RQ004: durée de maintenance
\item
  RQ005: disponibilité
\item
  RQ006: capacité de maintenance
\item
  RQ007: état final
\item
  RQ008: stockage
\end{itemize}

\end{frame}

\section{État de l'art}\label{etat-de-lart}

\begin{frame}{État de l'art (1)}

\begin{itemize}[<+->]
\tightlist
\item
  FMP: Flight and Maintenance Planning problem.
\item
  En Cho (2011), avions de l'Armée de l'air états-unienne ont été
  affectés à des opérations quotidiens pour minimiser le nombre maximal
  de maintenances.
\item
  En Kozanidis (2008), des heures de vol et des maintenances ont été
  affectées aux avions de la flotte grecque.
\item
  En Verhoeff, Verhagen, and Curran (2015), différents objectifs ont été
  essayé pour les affectations d'heures de vol et maintenances.
\end{itemize}

\end{frame}

\begin{frame}{État de l'art (2)}

Les travaux précédents:

\begin{itemize}[<+->]
\tightlist
\item
  ont étudié de flottes homogènes: nous avons de flottes hétérogènes.
\item
  n'ont pas le mêmes contraintes (stockages, 60 périodes entre
  maintenances) ou les mêmes objectifs (maximiser disponibilité).
\item
  n'affectent pas les avions aux missions.
\end{itemize}

\end{frame}

\section{Modèle}\label{modele}

\begin{frame}{Modèle: fonction objectif}

\begin{align}
    & \text{Min}\; W_1 m_{max} + W_2 u_{max}
\end{align}

\pause

Minimiser le nombre maximal de maintenances et le nombre maximal de
non-disponibilités.

\begin{align}
    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
    &t \in \mathcal{T} \\
    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max}
    &t \in \mathcal{T}
\end{align}

\end{frame}

\begin{frame}{Modèle: contraintes principales}

Besoins des missions et incompatibilités entre les missions et les
maintenances.

\begin{align}
&\sum_{i \in \mathcal{I}_j} a_{jti} = R_j
&j \in \mathcal{J}, t \in \mathcal{T}_j\\
&\sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1
& t \in \mathcal{T}, i \in \mathcal{I}
\end{align}

\end{frame}

\begin{frame}{Modèle: contraintes de flux}

Calcule d'heures de vol restantes pour chaque avion:

\begin{align}
& rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
& rut_{i0} = Rut^{Init}_i
        & i \in \mathcal{I}\\
& rut_{it} \geq H m_{it}
        & t \in \mathcal{T}, i \in \mathcal{I}\\
& rut_{it} \in [0,H]
        & t \in \mathcal{T}, i \in \mathcal{I} \\
& \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
        & t = |\mathcal{T}|
\end{align}

\end{frame}

\section{Résultats}\label{resultats}

\begin{frame}{Expériences}

Des instances ont déjà été testés:

\begin{longtable}[]{@{}lllllll@{}}
\toprule
id & \(\|\mathcal{J}\|\) & \(\|\mathcal{T}\|\) & assign & objective &
time (s) & bound\tabularnewline
\midrule
\endhead
I\_0 & 9 & 11 & 310 & 62.0 & 0.7 & 62.0\tabularnewline
I\_1 & 9 & 21 & 650 & 63.0 & 68.7 & 63.0\tabularnewline
I\_2 & 9 & 31 & 990 & 63.0 & 3600.1 & 62.0\tabularnewline
I\_3 & 9 & 41 & 1249 & 64.0 & 3603.9 & 61.7\tabularnewline
I\_4 & 10 & 11 & 530 & 82.0 & 0.9 & 82.0\tabularnewline
I\_5 & 10 & 21 & 1070 & 83.0 & 144.0 & 83.0\tabularnewline
I\_6 & 10 & 31 & 1610 & 83.0 & 3600.1 & 82.0\tabularnewline
I\_7 & 10 & 41 & 2069 & 84.0 & 3609.1 & 81.8\tabularnewline
I\_8 & 11 & 11 & 1080 & 139.0 & 530.6 & 139.0\tabularnewline
I\_9 & 11 & 21 & 2120 & 149.0 & 3600.0 & 139.9\tabularnewline
\bottomrule
\end{longtable}

\end{frame}

\section{Prochains pas}\label{prochains-pas}

\begin{frame}{Prochains pas}

\begin{itemize}[<+->]
\tightlist
\item
  \textbf{L'amélioration du modèle mathématique}: casser les symétries,
  reformuler les objectifs.
\item
  \textbf{Solutions techniques alternatives}: essayer d'autres méthodes
  de résolution.
\item
  \textbf{Continuer le dialogue avec les utilisateurs}: ajouter des
  nouvelles contraintes, valider les résultats.
\end{itemize}

\end{frame}

\section{Références}\label{references}

\begin{frame}{Références / Questions?}

\hypertarget{refs}{}
\hypertarget{ref-Cho2011}{}
Cho, P. 2011. ``Optimal Scheduling of Fighter Aircraft Maintenance.''
PhD thesis, Massachusetts Institute of Technology.

\hypertarget{ref-Kozanidis2008}{}
Kozanidis, G. 2008. ``A Multiobjective Model for Maximizing Fleet
Availability under the Presence of Flight and Maintenance
Requirements.'' \emph{Journal of Advanced Transportation} 43 (2):
155--82.

\hypertarget{ref-Verhoeff2015}{}
Verhoeff, M., W. J C Verhagen, and R. Curran. 2015. ``Maximizing
operational readiness in military aviation by optimizing flight and
maintenance planning.'' \emph{Transportation Research Procedia} 10
(July). Elsevier B.V.: 941--50.
doi:\href{https://doi.org/10.1016/j.trpro.2015.09.048}{10.1016/j.trpro.2015.09.048}.

\end{frame}

\end{document}

### 2018-06-22 R/Dassault_presentation/ppt_v2.Rmd
new documents and presnetations or re-organised documents

--- 2018-06-01 R/roadef_presentation/roadef_presentation.Rmd
@@
 author: |
-    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
-    | \tiny (franco.peschiera@isae-supaero.fr)
+    | Franco Peschiera, Alain Haït, Olga Battaïa
 institute: "ISAE SUPAERO"
-date: "February 22, 2018"
+date: "June 06, 2018"
 output: 
@@
     # footer: "ISAE-SUPAERO, Université de Toulouse, France"
-    includes:
-        in_header: header.tex
+    # includes:
+    #     in_header: header.tex
   slidy_presentation:
@@
 
-<!-- # Introduction -->
+# Introduction
 
-<!-- ## Introduction -->
+## PhD plan
 
-<!-- Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications.  -->
+|Phase  | Name                        | Month     |
+|-------|--------| --------|
+|1      | Review state of the art FMP              | 1 – 8    |
+|       | Review previous work                  | 1 – 8    |
+|2      | Build new mathematical models | 9 – 16   |
+|3      | Include stochasticity                    | 17 – 24  |
+|4      | Prepare case study: Armée de l’Air     | 25 – 30  |
+|5      | Build prototype app             | 25 – 30  |
+|6      | Write PhD thesis               | 31 – 36  |
 
-<!-- They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.
--->
+<!-- |Phase  | Libellé                        | Mois     | -->
+<!-- |-------|--------| --------| -->
+<!-- |1      | Etat de l’art FMP              | 1 – 8    | -->
+<!-- |       | Travaux antérieurs             | 1 – 8    | -->
+<!-- |2      | Nouveaux modèles mathématiques | 9 – 16   | -->
+<!-- |3      | Incertitude                    | 17 – 24  | -->
+<!-- |4      | Cas d’étude Armée de l’Air     | 25 – 30  | -->
+<!-- |5      | Prototype logiciel             | 25 – 30  | -->
+<!-- |6      | Mémoire de thèse               | 31 – 36  | -->
 
@@
 
-Assign both missions and maintenance operations to a fleet of aircraft in order to maximize availability and minimize costs.
+Assign both **missions and maintenance** operations to a fleet of aircraft in order to **maximize availability** and **minimize costs**.
 Missions have fixed start and end times and have particular needs in terms of aircraft and time.
@@
 * A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H$ units.
-
-<!-- ## Relevance and origin -->
-
-<!-- This thesis was started out of the need of the french Air Force to maximize the availability of their Mirage 2000 fleet while controlling cost and maintenance capacity. -->
-
-<!-- In fact, a similar subject has been presented as a challenge in a open competition: Défit Opti-plan. -->
 
@@
 
-## Model: flow constraints
-
-Elapsed times are treated in the same way but consumption is expressed in time periods.
-\begin{align}
-& ret_{it} \leq ret_{it-1} + E m_{it} - 1 & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
-& ret_{i0} = Ret^{Init}_i
-        & i \in \mathcal{I}\\
-& ret_{it} \geq E m_{it}
-        & t \in \mathcal{T}, i \in \mathcal{I}\\
-& ret_{it} \in [0,E]
-        & t \in \mathcal{T}, i \in \mathcal{I} \\
-& \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
-        & t = |\mathcal{T}|
-\end{align}
-
 # Results
@@
 
-## Sample solution progress
-
-The analysis of the solution progress for all instances shows that:
-
-* Initial solutions are quite far from the linear relaxation.
-* Branching does not improve significantly the lower bound.
-
-\pause
-```{r out.width="150pt", echo=FALSE}
-knitr::include_graphics("./../../img/progress.png.png")
-```
-
-*(Progress diagram for instance id=I_7).*
-
-## Multi-objective (1)
-
-Comparison of two objectives: max availability or maintenance capacity.
-Different weights were used for the two objectives: $W_1, W_2 \in [0..1]$ with pace of 0.1.
-\pause
-```{r out.width="180pt", echo=FALSE}
-knitr::include_graphics("./../../img/multiobjective.png")
-```
-
-*(Pareto diagram for instance id=I_5).*
-
-## Multi-objective (2)
-
-* Analysis for all instances: not many pareto optimal points available (between two and four).
-* The objectives are quite aligned one with the other.
-
 # Next steps

### 2018-06-22 R/Dassault_presentation/ppt_v2_fr.Rmd
new documents and presnetations or re-organised documents

@@
 
-Une flotte hétérogène d'avions avec les caractéristiques suivantes:
+Une flotte hétérogène d'avions, dont chacun a les caractéristiques suivantes :
 
-* Temps écoulé restant.
-* Des heures de vol restant.
+* Butée calendrier.
+* Butée horaire.
 * Fonctionnalités.
@@
 
-Un ensemble de missions à réaliser, chaque une avec:
+Un ensemble de missions à réaliser, chacune avec :
 
 * Dates de début et fin.
-* Un besoin mensuelle d'avions.
-* Une consommation d'heures par avion-mois affecté.
-* Des fonctionnalités requise.
+* Un nombre nécessaire d'avions par mois.
+* Une consommation d'heures par avion et par mois.
+* Des fonctionnalités requises.
 
@@
 
-Chaque maintenance suit les règles suivantes:
+Chaque maintenance suit les règles suivantes :
 
 * La *durée est de 6 mois* consécutifs.
-* Le nombre maximal *d'heures de vol* entre deux maintenances est *1000h*.
-* Le nombre maximal de *mois calendrier* entre deux maintenances est *60 mois*.
+* Le nombre maximal d'heures de vol entre deux maintenances est *1000h*.
+* Le nombre maximal de mois entre deux maintenances est *60 mois*.
 
@@
 
-En résumé, voici les états logiques possibles d'un aéronef:
+En résumé, voici les états logiques possibles d'un aéronef :
 
@@
 * Maximiser la disponibilité.
-* Minimiser le coût (réduisant le nombre de maintenances).
+* Minimiser le coût (en réduisant le nombre de maintenances).
 
@@
 
-* RQ001: heures de missions
-* RQ002: aéronefs assignés
-* RQ003: besoins de maintenance
-* RQ004: durée de maintenance
-* RQ005: disponibilité
-* RQ006: capacité de maintenance
-* RQ007: état final
-* RQ008: stockage
+* RQ001: heures de missions.
+* RQ002: aéronefs affectés.
+* RQ003: besoins de maintenance.
+* RQ004: durée de maintenance.
+* RQ005: disponibilité.
+* RQ006: capacité de maintenance.
+* RQ007: état final.
+* RQ008: stockage.
 
@@
 
-* FMP: Flight and Maintenance Planning problem.
-* En @Cho2011, avions de l'Armée de l'air états-unienne ont été affectés à des opérations quotidiens pour minimiser le nombre maximal de maintenances.
-* En @Kozanidis2008, des heures de vol et des maintenances ont été affectées aux avions de la flotte grecque.
-* En @Verhoeff2015, différents objectifs ont été essayé pour les affectations d'heures de vol et maintenances.
+> FMP: Flight and Maintenance Planning problem.
+
+* Minimisation du nombre maximal de maintenances. Affectations quotidiennes. États-Unis, @Cho2011.
+* Maximisation de disponibilité de vol. Affectations mensuelles Grèce, @Kozanidis2008.
+* Minimisation du nombre maximal de maintenances. Affectations mensuelles. Pays-Bas. @Verhoeff2015.
 
@@
 
-* ont étudié de flottes homogènes: nous avons de flottes hétérogènes.
-* n'ont pas le mêmes contraintes (stockages, 60 périodes entre maintenances) ou les mêmes objectifs (maximiser disponibilité).
+* ont étudié de flottes homogènes: nous avons une flotte hétérogène.
+* n'ont ni le mêmes contraintes ni les mêmes objectifs.
 * n'affectent pas les avions aux missions.
@@
 
-Calcule d'heures de vol restantes pour chaque avion:
+Calcul des heures de vol restantes pour chaque avion:
 
@@
 
-Des instances ont déjà été testés:
+Des instances ont déjà été testées:
 
@@
 
-# Prochains pas
-## Prochains pas
+# Perspectives
+## Perspectives
 
@@
 * **Solutions techniques alternatives**: essayer d'autres méthodes de résolution.
-* **Continuer le dialogue avec les utilisateurs**: ajouter des nouvelles contraintes, valider les résultats.
+* **Continuer le dialogue avec les utilisateurs**: ajouter de nouvelles contraintes, valider les résultats.
 

### 2018-06-22 R/Dassault_presentation/proposal_fr.Rmd
new documents and presnetations or re-organised documents

--- 2018-03-20 R/OptiPlan_presentation/proposal_fr.Rmd
@@
 ---
-title: "Optiplan proposal"
+title: "OPTIMA: Preésentation du problème"
 author: "ISAE-SUPAERO, Université de Toulouse"
-date: "Mars 11, 2018"
-nocite: '@*'
+date: "Juin 06, 2018"
+# nocite: '@*'
 output:
-  html_document:
-    css: ./../../css/baobab_docs.css
-    toc: yes
+  slidy_presentation:
+    css: ./../../css/baobab_slidy.css
+    footer: ISAE-SUPAERO, Université de Toulouse, France
+    # incremental: yes
+    # slide_level: 3
+    # pandoc_args: ["--slide-level=3"]
+    # toc: yes
   beamer_presentation:
@@
     pandoc_args: ["--top-level-division=part"]
-  slidy_presentation:
-    css: ./../../css/baobab_MH.css
-    footer: ISAE-SUPAERO, Université de Toulouse, France
-    # incremental: yes
-    slide_level: 3
-    pandoc_args: ["--slide-level=3"]
+  html_document:
+    css: ./../../css/baobab_docs.css
     toc: yes
-bibliography: MFMP.bib
 editor_options: 
@@
 source("diagrams/functionalities.R")
-source('diagrams/exampleDT.R')
-source('diagrams/gantt_tasks.R')
+source('functions/exampleDT.R')
+source('functions/gantt_tasks.R')
 exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
 exp_directory2 = PATHS[['experiments']] %>% paste0('201802061201/')
-gantt <- print_solution(exp_directory, max_resources = 10, width="100%", height="500px")
+
+states <- get_states(exp_directory) %>% mutate(content = str_extract(content, "^M|(O\\d)"))
+gantt <- timevis_from_states(states, max_resources=10, width="100%", height="500px")
 gantt_tasks <- print_tasks(exp_directory2, width="100%", height="200px", max_tasks = 5)
@@
 
-<!-- TODO: check if all points from the rules are covered -->
+### Problème
 
-## Introduction 
-### Introduction
-
-Dans ce document, nous présentons tout d’abord le problème tel que l’équipe le comprend.
-Les sources d’information sont la connaissance préliminaire de l’équipe et la réunion OptiPlan qui a eu lieu à Paris.
-
-Notre proposition technique inclut des fonctionnalités qu’une éventuelle application complète comprendrait.
-Dans le temps imparti au projet nous projetons de réaliser un prototype basé sur cette proposition technique
-
-## Présentation de l'entreprise
-### Présentation de l'entreprise
-
-Nom: ISAE-SUPAERO
-Activité: école d’ingénieurs.
-
-Grand école du secteur aéronautique et spatial, l'ISAE-SUPAERO développe depuis 2007 une recherche propre portée par ses département.
-Le département d'ingénierie des systèmes complexes (DISC), par son groupe Systèmes Décisionnels, travaille sur les outils de décision appliqués aux systèmes industriels, en particulier pour la production et la maintenance dans l'industrie aérospatiale.
-
-## Savoir-faire
-### Compétences clés
-
-Nous pensons que cette équipe est particulièrement bien placée pour
-aborder ce projet pour les raisons suivantes:
-
-L’équipe possède une grande expérience en recherche opérationnelle:
-
-* recherche scientifique, publication d’articles
-* projets de recherche en collaboration avec l’industrie
-* conception d’applications logicielles en planification et optimisation pour des entreprises
-
-L’équipe a déjà encadré des travaux de recherche sur le sujet de la maintenance aéronautique.
-En outre, elle a une connaissance du problème en question car une thèse de doctorat sur ce sujet a été lancée en octobre 2017.
-Les travaux ont été présentés à la conférence ROADEF 2018 et seront présentés dans la conférence MOSIM 2018
-
-### Alain Haït
-
-Ingénieur en mécanique, Alain Haït possède une thèse de robotique préparée au LAAS CNRS.
-Enseignant-chercheur à l’INP Toulouse, il participé à la conception et la mise en place du département de génie industriel de l’ENSIACET.
-Chercheur visiteur au département de mathématiques et génie industriel de l’École Polytechnique de Montréal pendant deux ans, il a co-édité un livre sur les méthodes de planification et d’ordonnancement de la production avec contraintes liées aux ressources humaines, avec les professeurs François Soumis, Vincent Giard et Pierre Baptiste.
-Alain Haït est certifié CPIM de l’APICS.
-Spécialisé dans les méthodes de planification, il a encadré une thèse sur la planification de la maintenance d’hélicoptères civils.
-
-Professeur à l’ISAE-SUPAERO depuis 2007, il dirige depuis trois ans le département d’ingénierie des systèmes complexes (DISC) qui couvre les thèmes mathématiques appliquées, réseaux de communication, systèmes embarqués, ingénierie système et systèmes décisionnels.
-
-### Olga Battaïa
-
-Professeur à l'ISAE-SUPAERO et responsable de la formation en Génie Industriel, Olga Battaïa est une chercheuse internationalement reconnue dans le domaine de l'optimisation des systèmes industriels.
-Sa thèse de doctorat soutenue en 2007 à l’Ecole des Mines de Saint-Etienne a été primée du prix des meilleures thèses par le GdR MACS.
-Olga Battaïa est auteure de plus de 170 publications scientifiques, parmi lesquelles plus de 50 articles dans les revues internationales et 57 articles dans les proceedings des conférences internationales.
-Elle a participé à la réalisation de nombreux projets de recherche nationaux et internationaux.
-Elle a été invitée à plusieurs reprises par des universités étrangères pour donner des conférences scientifiques : Linköping University (2016), University of Siegen (2016), Politecnico di Milano, 2014 (Italie), Tshwane University of Technology, 2014 (Afrique du Sud), Lebanese University, 2013 (Liban) MIT, 2011 (USA), University of Michigan, 2011 (USA).
-En 2018, elle est devenue la première femme française nommée Membre Associé de la prestigieuse Académie Internationale de Recherche en Productique (CIRP).
-
-### Franco Peschiera
-
-Franco Peschiera a travaillé dans l'industrie de la recherche opérationnelle depuis près de 7 ans.
-Au cours de ce temps il a été le propriétaire et le directeur technique pour une startup espagnole (www.baobabsoluciones.es), en construisant des applications logicielles avec des modèles d'optimisation mathématique ou de simulation à l'intérieur pour plusieurs grandes entreprises espagnoles et internationales.
-
-Ces projets comportaient la liste suivante de tâches: la spécification des besoins du client, la construction d'un modèle mathématique ou de simulation, la programmation d'une interface utilisateur graphique, le déploiement et l'intégration avec les systèmes informatiques, la documentation et la formation des utilisateurs.
-
-Franco a commencé un doctorat à l'université ISAE-SUPAERO en octobre 2017 sur le sujet de la planification de la maintenance des avions militaires.
-
-## Motivations
-### Motivations
-
-Nous sommes déjà engagé dans une thèse sur le sujet de planification de maintenances des aéronefs pour la DGA.
-
-Le challenge OptiPlan va nous permettre avoir interaction directe avec la DGA pendant que on développe des meilleurs modèles pour avoir une meilleure compréhension du problème et résoudre des possibles doutes, ainsi que de recevoir de données de qualité pour évaluer nos modèles.
-
-Le challenge va nous permettre recevoir feedback de la part DGA sur les résultats qu'on obtienne. Le retourne d'information permettra ensuite améliorer encore plus les résultats.
-
-Finalement, le challenge permettra valider l'utilité des modèles et techniques employés dans une application complète et réel.
-
-## Expérience
-
-### Plannification logistique et RH
-
-**Clients**: Repsol et CLH.
-
-**Problematiques**:
-
-* Génération des horaires et affectation du personnel. 
-* Planification stratégique des réseaux de transport et production au niveau national.
-
-**Réponses**:
-
-* Logiciel de gestion de ressources humaines utilisé tous les jours.
-* Logiciel de optimisation mensuelle logistique.
-
-```{r, echo=FALSE, out.width='50%'}
-knitr::include_graphics('../../img/CLH.png')
-```
-
-### Plannification de la production
-
-**Clients**: General Motors, LI Lilly, Air Liquide, Repsol.
-
-**Problematiques**:
-
-* Planification de la production.
-* Détection de problèmes d'approvisionnement.
-
-**Réponses**:
-
-* Logiciel de simulation horaire de la production en temps réelle.
-* Logiciel optimisation de la production hebdomadaire.
-* Logiciel de planification mensuel de la production.
-
-```{r, echo=FALSE, out.width='50%'}
-knitr::include_graphics('../../img/Lilly.png')
-```
-
-### Logistique et transport
-
-**Clients**: Alstom, ASM et IO.
-
-**Problematiques**:
-
-* Simulation de trains et réseaux métro.
-* Optimisation de tournées de véhicules.
-
-**Réponses**:
-
-* Logiciel de simulation de réseaux de trains.
-* Logiciel d'optimisation de transports.
-
-```{r, echo=FALSE, out.width='50%'}
-knitr::include_graphics('../../img/IO.png')
-```
-
-## Problème
-
-L’Armée de l’Air dispose d’une flotte d’avions Mirage 2000 de différents types, avec des fonctionnalités particulières et avec un état d’opération spécifique donné par certains indicateurs temporels d’usage.
+On dispose d’une flotte d’avions de différents types, avec des fonctionnalités particulières et avec un état d’opération spécifique donné par certains indicateurs temporels d’usage.
 
@@
 
-Dans la littérature, le problème est considéré comme un problème de planification de vol et de maintenance (Flight and Maintenance Planning problem).
-Des recherches dans ce domaine ont été réalisées par @Kozanidis2008, @Cho2011, @Verhoeff2015 et @Chastellux2016, entre autres.
+-----
+
+```{r, echo=FALSE, out.width='100%'}
+if (html_output){
+    gantt
+}
+```
 
@@
 Chaque mission nécessite un et un seul type d’aéronef qui, de plus, doit satisfaire à des exigences de mission supplémentaires.
-
------
-
-```{r, echo=FALSE, out.width='90%'}
-if (html_output){
-    gantt_tasks
-} else {
-    knitr::include_graphics('../../img/tasks_gantt.png')
-}
-```
-
-    Exemples de missions avec leurs besoins en aéronefs.
 
@@
 
-## Proposition
-
-### Résumé
-
-Le projet présenté est une application web interactive.
-Il utilise un modèle mathématique pour générer de nouvelles solutions, permet à différents utilisateurs d'y accéder avec différents droits d'accès et offre une puissante visualisation pour afficher, comparer et analyser les planifications.
-Il s'intègre aux sources de données existantes (fichier Excel), aux nouvelles proposées (modèle Excel) et a la possibilité de s’intégrer aux sources existantes (autres systèmes internes) ainsi qu'à la sortie simple (csv, json).
-Enfin, il permet l'utilisation de cas pour stocker des informations, ce qui facilite les essaies, l'expérimentation et la comparaison de différents scénarios.
-
-Un résumé des fonctionnalités de l'application est montré dans l'image suivante:
-
-```{r, echo=FALSE, out.width='90%'}
-if (html_output){
-    graph_functionalities()
-} else {
-    knitr::include_graphics('../../img/functionalities.png')
-}
-```
-
-    Graphique montrant les fonctionnalités de l'application.
-
 ### Optimiseur
@@
 
-Les options disponibles seront, par défaut:
+Des possibles options sont, par défaut, les suivantes:
 
@@
 
-### Cas et expérimentation
+### Techniques de résolution
 
-Étant donné la nature stratégique de l’application, elle sera centrée sur des études de cas.
-Cela signifie qu’il permettra de sauvegarder toutes les données pertinentes (entrée et sortie) dans un même ensemble ou cas.
-De cette façon, différentes expériences pourront être réalisé en copiant et en modifiant un cas existant.
-La gestion de cas est similaire à l'utilisation de plusieurs feuilles de calcul pour étudier plusieurs scénarios différents.
-
-La figure suivante montre un exemple de cas déjà enregistrés. Elle représente les interactions possibles entre les cas pendant une période de quelques mois.
-
-1. Le cas 1 a été publié pour le mois de mars.
-2. Au cours du mois, il a reçu quelques modifications, ce qui a engendré les cas 2 et 3.
-3. Enfin, le cas 3 a été re-publié pour le mois de mars.
-4. Plus tard, le cas 4 a été créé et modifié sans être publié.
-5. Finalement, Le cas 5 a été crée à partir du cas 4 et publié en avril.
-
------
-
-```{r, echo=FALSE, out.width='70%'}
-if (html_output){
-    graph_cases()
-} else {
-    knitr::include_graphics('../../img/cases.png')
-}
-```
-
-    Exemple montrant la création et publication de les cas.
-
-Les fonctionnalités disponibles seront les suivantes:
-
-* Créer un nouveau cas.
-* Copier un cas existant.
-* Supprimer un cas.
-* Comparer deux cas.
-* Publier un cas.
-* Exporter un cas.
-
------
-
-#### Contenu d'un cas
-
-Un «cas» est une unité de stockage qui permet de s’intégrer facilement aux structures de données actuelles. Il contient toutes les informations nécessaires pour construire une planification:
-
-* L’ensemble d’avions avec leurs caractéristiques et leurs informations historiques jusqu’à une certaine date.
-* L’ensemble de missions ou d’heures de vol à réaliser dans la période de planification.
-* Une pre-affectation d’aéronefs à des opérations pendant l’horizon de planification.
-* Les futurs calendriers de maintenance.
-* La configuration : la durée de maintenance, les heures de vol après une maintenance, etc.
-* Une planification (générée manuellement ou avec l'application).
-
------
-
-#### Comparaison des cas
-
-Cette fonctionnalité peut potentiellement avoir plusieurs utilisations en fonction des besoins finaux du client.
-Une option pourrait être de visualiser les différences dans les données d’entrée pour deux cas.
-Une autre option consisterait à visualiser les différences dans la planification résultante pour les deux cas.
-Enfin, il pourrait être intéressant de comparer des KPI (Key Performance Indicators) spécifiques pour deux cas, afin de sélectionner le meilleur scénario.
-
------
-
-#### Publications de cas
-
-Une situation courante dans la création de planifications est le fait que plusieurs cas différents peuvent exister mais il est toujours important d’identifier lequel est le cas ‘actif’.
-Pour faire face à cela, un système de publication simple sera inclus dans l’application, où le «planificateur» (voir la section “Utilisateurs”) peut identifier le cas qui devra être partagé avec les autres utilisateurs.
-
-### Interface graphique utilisateur (GUI)
-
-Toute interaction de l’utilisateur avec l’application se fera via une application HTML.
-En plus de montrer des données dans des tableaux interactifs, des alternatives graphiques spécifiques peuvent être incorporées pour rendre l’utilisation plus facile et plus efficace.
-
-Ici, un Gantt interactif est utilisé pour présenter et éventuellement modifier un plan de maintenance.
-Il est possible d’éditer les opérations de maintenance et l’affection aux missions (changer la taille ou la déplacer) en double-cliquant dessus et en faisant glisser.
-En outre, il est possible d’ajouter de nouvelles affectations en double-cliquant.
-
-```{r, echo=FALSE, out.width='70%'}
-
-if (html_output){
-    gantt
-} else {
-    knitr::include_graphics('./../../img/calendar2.png')
-}
-```
-
-    Exemple de Gantt interactif.
-
------
-
-Ici, un graphique interactif est utilisé pour présenter l’évolution du statut d’un seul avion au fil du temps. Il est possible de filtrer, zoomer et exporter en tant qu’image.
-
-```{r, echo=FALSE, out.width='90%'}
-
-if (html_output){
-    graphs_rem
-} else {
-    knitr::include_graphics('./../../img/remaining.png')
-}
-```
-
-    Exemple de graphique interactif.
-
------
-
-Ici, une table interactive est utilisée pour présenter certaines données. Il est possible d'éditer le contenu, de filtrer, de réorganiser les données, etc.
-
-```{r, echo=FALSE, out.width='90%'}
-
-if (html_output){
-    table_rem
-} else {
-    knitr::include_graphics('./../../img/table.png')
-}
-```
-
-    Exemple de table interactive.
-
-### Intégration de données
-
-L’application inclura son propre stockage de données à travers une base de données.
-Toutes les informations utilisées par l’application seront stockées ici.
-En plus de ce système, l’application permettra plusieurs façons d’importer et d’exporter des données à partir de l'application.
-
-#### Importer des données
-
-Un nouveau cas peut être créé à partir des options suivantes:
-
-* Le fichier Excel actuellement utilisé.
-* Un nouveau fichier modèle Excel à créer.
-* Les données historiques pourraient être lues par un autre système, si disponible.
-
------
-
-#### Exporter des données
-
-Les données dans un cas pourront être exportées de la manière suivante:
-
-* Dans le format Excel actuel (ou un format similaire).
-* Dans le nouveau template Excel qui sera créé.
-* Dans un fichier csv et / ou json si nécessaire.
-
-Ces capacités d’exportation permettront également de partager des informations partielles avec d’autres acteurs dans les formats mentionnés précédemment.
-
-### Architecture et déploiement
-
-L’application sera déployée via une architecture client-serveur.
-De cette façon, il sera facile de partager l’application étant donné que cela nécessitera seulement l’installation d’un navigateur web moderne (Google Chrome).
-
-Le serveur devra être installé quelque part dans l’infrastructure du client.
-
-Des logiciels supplémentaires pourraient être nécessaires en fonction des spécifications finales.
-Par exemple, si le modèle est trop gros pour les solveurs en logiciel libre, un solveur commercial pourrait être nécessaire.
-
-En fonction de la future intégration, des flux de données supplémentaires seront inclus dans la conception.
-Le seul flux de données confirmé pour le moment se situe entre le fichier Excel existant et l’application.
-
-La figure suivante présente un résumé de l’architecture à mettre en œuvre et des flux de données entre les composants.
-Les deux types d’utilisateurs vont communiquer avec l’application pour générer les plannings ou les voir.
-L’application a son propre moteur (modèle) pour générer des solutions et son propre stockage (base de données).
-Les technologies sont mentionnées sur chaque arc pour indiquer là où elles seront utilisées.
-
------
-
-```{r, echo=FALSE, out.width='70%'}
-
-if (html_output){
-    graph_architecture()
-
-} else {
-    knitr::include_graphics('../../img/architecture.png')
-}
-```
-
-    Architecture proposée avec flux de données.
-
------
-
-Les utilisateurs interagiront via les navigateurs Web avec le serveur.
-Pour charger de nouvelles données, ils pourront charger un nouveau fichier Excel avec les données pertinentes ou éditer un cas existant via l’application.
-
-Le serveur sera chargé de résoudre et de générer de nouvelles solutions de planification et de stocker toutes les informations pertinentes dans sa base de données.
-
-Les utilisateurs seront en mesure de visualiser les résultats. En outre, ils seront en mesure d’exporter les résultats en formats csv, Excel et json en fonction de leurs besoins.
-
------
-
-Cette architecture client-serveur offre plusieurs avantages:
-
-* Des graphiques puissants pouvant être incorporés via des bibliothèques javascript et HTML.
-* Une facilité de maintenance liée à une séparation claire entre les fonctionnalités de l’interface graphique, de la logique d’utilisation et du modèle d'optimisation.
-* Un navigateur Web moderne comme seul besoin logiciel.
-* Facilité pour incorporer plusieurs utilisateurs.
-* L’intégration avec d’autres systèmes peut être effectuée avec des changements minimes à la structure existante.
-* Possibilité d'allouer des ressources supplémentaires au niveau du serveur au lieu de devoir changer les ordinateurs des utilisateurs.
-* Absence de logiciel tiers à acheter.
-Tous les composants sont open source.
-La seule exception possible pourrait être le solveur mathématique, en fonction de la taille du problème et des besoins en performance.
-
-### Utilisateurs
-
-Deux rôles possibles sont identifiés pour les utilisateurs de cette application.
-
-Le premier s’appellera “planificateur”.
-Cet utilisateur est en charge de créer (manuellement ou automatiquement) des plannings différents.
-Le second s’appellera “spectateur”.
-Cet utilisateur pourra visualiser la planification créée et stockée dans l’application.
-En fonction des besoins du client, plusieurs profils “spectateur” pourraient être créés afin de partager certaines parties des informations stockées avec chaque profil.
-
-## Planification
-
-Ici, les tâches requises avec leur ordre sont présentées. Le principal risque sur la planification actuelle est l’information incomplète sur:
-
-* le problème spécifique à résoudre.
-* l’infrastructure informatique disponible du client.
-* la disponibilité des utilisateurs potentiels pour discuter des exemples de solutions et tester et évaluer les fonctionnalités.
-
-On présente une proposition de planification basée sur les expériences précédentes. La planification dépendra de la voie de communication finalement choisie et de la disponibilité du client.
-
-### Phase 1: exigences
-
-Au cours de cette phase, des réunions en personne et par téléphone seront prévues.
-
-Les conditions requises pour cette phase sont: *Aucun*.
-
-Le livrable de cette phase sera:
-
-* un document avec une description validée par le client des exigences minimales du prototype
-* un document d’exigences additionnelles à préciser ultérieurement.
-
-La durée de cette phase est estimée à: *3 semaines*.
-
-### Phase 2: première itération
-
-Dans cette phase, une première version du prototype sera construite et déployée.
-
-Les conditions requises pour cette phase sont:
-
-* une liste d'exigences minimales.
-* une possibilité de déploiement pour partager l’application avec les utilisateurs.
-
-Le livrable de cette phase sera:
-* une application prototype que les utilisateurs peuvent utiliser et tester.
-* des exemples de solutions à analyser par les utilisateurs.
-
-La durée de cette phase est estimée à: *5 semaines*.
-
-### Phase 3: deuxième itération
-
-Dans cette phase, une version finale du prototype sera construite et déployée.
-En fonction du niveau d’implication du client, du types de changements à réaliser et de l’urgence, cette phase pourrait être coupée en deux sous phases avec des livrables distincts.
-
-Les conditions requises pour cette phase sont:
-
-* une liste des changements à réaliser sur la première version basée sur les commentaires des utilisateurs.
-* une liste de fonctionnalités supplémentaires à ajouter.
-
-Le livrable de cette phase sera:
-
-* une nouvelle version de l’application que les utilisateurs pourront utiliser.
-* la documentation technique.
-* la Guide de l’utilisateur.
-
-La durée de cette phase est estimée à: *5 semaines* .
-
-### Profils et engagement
-
-Il y aura trois profils dans ce projet:
-
-* Profil 1: Consultant Senior spécialisé en Recherche Opérationnelle (RO).
-* Profil 2: Développeur Senior spécialisé en RO.
-* Profil 3: Chef de projet spécialise en RO.
-
-Dans le tableau ci-dessous sont les engagements hebdomadaires estimés des profils:
-
-| Phase | Durée | Profil 1 | Profil 2 | Profil 3
-|-------|----------|----------|----------|---------|
-|Phase 1| 3 semaines| 4 h    | 30 h    | 8 h     |
-|Phase 2| 5 semaines| 2 h    | 30 h    | 8 h     |
-|Phase 3| 5 semaines| 2 h    | 30 h    | 8 h     |
-
-### Estimation économique
-
-Basé sur la définition de profils et des engagements déjà présentés, le montant correspondant au projet de développement du prototype est de 55.600€.
-
-## Technologies et techniques
-
-* **Modélisation mathématique**: pour générer des planifications.
-    * **Solveur mathématique**: pour générer les planifications. Cela pourrait être CBC, GUROBI ou CPLEX en fonction des besoins du modèle.
-* **Python**: pour la modélisation, l’heuristique et l’intégration de données.
-* **R**: pour les statistiques, possibles prévisions et la visualisation des données.
-* **PostgreSQL**: pour stocker des données.
-* **HTML5**: pour les rapports, et les visualisations.
-
-### Modélisation mathématique
-
-La modélisation mathématique est une technique qui permet la construction d’un ensemble d’équations pour représenter une situation spécifique bien définie et communique avec les logiciels avancés pour rechercher et trouver la meilleure solution qui respecte ces équations.
+La **modélisation mathématique** est une technique qui permet la construction d’un ensemble d’équations pour représenter une situation spécifique bien définie et communique avec les logiciels avancés pour rechercher et trouver la meilleure solution qui respecte ces équations.
 Ces logiciels (modélisateurs et solveurs) sont disponibles à la fois en version commerciale et libre (gratuite).
 
-Les avantages de l’utilisation de cette technique par rapport à d’autres alternatives sont qu’elle peut prouver l’optimalité d’une solution, qu'elle peut facilement incorporer plusieurs objectifs et qu'il est possible de fixer facilement certaines décisions déjà prises, rendant possible l’utilisation interactive du modèle par l’utilisateur.
-Avec peu de modifications, le modèle peut aussi être réutilisé pour prendre en compte différentes situations similaires à la situation originelle.
-
-D’autres fonctionnalités incluent également des limites de temps et d’autres conditions d’arrêt plus complexes.
+D'autres techniques aussi disponibles sont la **programation par contraintes** et les **heuristiques**.
 
@@
 Des exemples de ces bibliothèques sont: [leaflet](http://leafletjs.com/), [visjs](visjs.org), [d3js](https://d3js.org/) parmi beaucoup d'autres.
-
-## Remarques
-### Remarques
-
-1. La proposition présentée ci-dessus est destinée à une application complète livrée dans le cadre d’un projet complet au client.
-2.
-Le prototype à livrer n’inclura pas toutes les fonctionnalités mentionnées dans ce document mais montrera des exemples des fonctionnalités les plus importantes suite à une réunion avec les futurs utilisateurs et décideurs.
-3.
-L’intégration avec d’autres systèmes n'est pas inclus dans le cadre de cette proposition car plus d’informations sont nécessaires concernant les systèmes existants pour pouvoir estimer la charge de travail.
-4.
-Il est supposé que le client fournira les exigences hardware ainsi que les droits d’accès et les autorisations pour installer tous les composants du logiciel.
-Alternativement, le logiciel pourrait être installé dans le cloud.
-5.
-Même si la programmation mathématique offre la capacité de prouver l’optimalité d’une solution, l’application ne garantit pas de la trouver dans tous les cas de figure.
-La qualité des solutions et le temps nécessaire pour les obtenir dépendront de nombreux facteurs; la taille de l’instance étant la principale.
-
-## Bibliographie
-### Bibliographie

### 2018-06-22 latex/notes/complexite/FMP_variants.tex
new documents and presnetations or re-organised documents

--- 2018-05-17 latex/notes/FMP_variants.tex
@@
         \item All missions can use any aircraft.
+    \end{itemize}
+
+    \subsection{Limitations}
+
+    \begin{itemize}
+        \item There is not starting period.
     \end{itemize}
@@
      \item There are two artificial activities representing the start and end of the horizon.
+     \item maintenance operations consume also resources of the missions.
     \end{itemize}
@@
         \item The missions consume 0 hours of aircraft.
-        \item The missions require 0 aircraft at each period.
+        %\item The missions require 0 aircraft at each period.
+        \item Each aircraft can only be used for one mission.
     \end{itemize}
@@
         \item Resource usage is always 1.
-        \item There is only one type of resource.
+        %\item There is only one type of resource.
         \item All durations are the same.
@@
 
+\section{Three partition problem}
+
+    \subsection{Hypothesis}
+
+    \begin{itemize}
+     \item assigned maintenances at regular intervals of M.
+     \item We need to assign tasks in between maintenances.
+     \item Each task assignment needs M/3 consecutif periods.
+     \item Each task assignment has an flight hours consumption.
+    \end{itemize}
+
+    \subsection{Reductions}
+
+    \begin{itemize}
+        \item Missions have a fixed duration.
+        \item Maintenances are evenly spaced.
+        %\item The missions require 0 aircraft at each period.
+        % \item Each aircraft can only be used for one mission.
+    \end{itemize}
+
+    % \subsection{Limitations}
+
+    % \begin{itemize}
+    %     \item Resource usage is always 1.
+    %     %\item There is only one type of resource.
+    %     \item All durations are the same.
+    % \end{itemize}
+
+    \subsection{Notations}
+
+    \textbf{Sets}
+
+    \noindent{
+    \begin{tabular}{ll}
+        $i \in \II = [\![1, I]\!]$ & Set of aircraft.\\
+        $k \in \KK_i = [\![1, K_i]\!]$ & Set of maintenances for aircraft $i$.\\ 
+        $j \in J$ & Set of missions.\\ 
+        % $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
+        % $k \in \mathcal{R}$ & Types of resources that maintenance can take. We only have one type.\\
+        % $i = 0$ & Artificial starting node.\\
+        % $i = n+1$ & Artificial ending node.\\
+    \end{tabular}
+    }
+
+    This doesn't make sense because the tasks are already defined. We cannot chose when it executes.
+
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % \bibliographystyle{dcu}

### 2018-06-22 latex/notes/model/model.tex
new documents and presnetations or re-organised documents

--- 2018-01-30 latex/MOSIM2018/article.tex
@@
-\documentclass[a4paper,twocolumn,fleqn]{article}
+\documentclass[a4paper,onecolumn,fleqn]{article}
 \textwidth 17cm \textheight 247mm
@@
 
-\title{ \vspace*{-25mm}
-\centering
-\fbox{\normalsize
-\begin{minipage}{17cm}
-\centering
-\em \small{$12^{th}$ International Conference on MOdeling, Optimization and SIMlation - MOSIM18 - June 27-29 2018  \\ Toulouse - France "The rise of connected systems in industry and services"}
-\end{minipage} }\\
-\vspace*{6mm}
-{\Large \textbf{Bi-objective MIP formulation for the optimization of maintenance planning on French military aircraft operations}}}
-
-\author{
-\begin{tabular}{c}
-\bf \normalsize {Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin} \\
-\\
-     \normalsize ISAE-SUPAERO, % \\
-     \normalsize Université de Toulouse, % \\
-     \normalsize 31055 Toulouse cedex 4 - France\\
-     \normalsize {\{franco.peschiera,alain.hait,olga.battaia,nicolas.dupin\}@isae-supaero.fr}
-     \\
-\end{tabular}
-}
-
-\date{\begin{minipage}{17cm}
-\normalsize
-{\bf  ABSTRACT:}
-\rm
-{\em A specific Flight and Maintenance Planning problem is presented. 
-In this problem, preventive maintenance operations are scheduled for military aircraft along with the assignment of regular missions. 
-The quality of the solutions is measured via a bi-objective function that smooths both maintenance operations and aircraft unavailability among time periods. 
-An exact optimization approach is provided by a Mixed Integer Programming model.
-A real world dataset provided by the French Air Force is used to test the developed optimization approach.
-Mono-objective computations are solved to optimality for medium size instances, allowing to compute exactly the Pareto frontier of the bi-objective optimization problems.
-The tests show that these two objectives do not lead to the same optimal solution, but very good compromise solutions can be found thanks to the bi-objective optimization.
-}\\~\\
-{\bf KEYWORDS:}
-\rm
-{\em optimization, Flight and Maintenance Planning, aircraft maintenance, Mixed Integer Programming, bi-objective optimization}
-\end{minipage}
-}
-\maketitle
-
-% Here I put the list of things we will leave for another article:
-% 1. stochasticity
-% 2. symmetry breaking inside model.
-% 3. storage?
-
-\section{Introduction}
-
-  Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications. 
-  They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.
-
-  The specific FMP of this paper comes from the French Air Force's operations. It has two specificities regarding the classical FMP considered in the literature.
-  On one hand, the fleet of aircraft is heterogeneous, with different standards, capacities and retrofits.
-  On the other hand, there is a will to align two objectives: the smoothing of the maintenance operations and the minimization of the global unavailability of aircraft to be able to add new missions.
-This last point calls for a bi-objective optimization.
-
-  This paper investigates how to solve such new optimization problem, and how to take good compromise decisions if these objectives are antagonistic.
-
-  The article is structured as follows:
-
-  Section \ref{sec:problem} presents a detailed description of the problem. An analysis of the previous work on FMP is given in section \ref{sec:soa}.
-  Section \ref{sec:model} presents a new MIP formulation for the considered problem. 
-  Section \ref{sec:data} introduces the real case study.
-  Section \ref{sec:results} discusses the obtained results.
-  Finally, section \ref{sec:conclusions} provides conclusions and directions of further work.
-
-\section{Problem statement}
-  \label{sec:problem}
-  The problem consists in assigning resources to predefined tasks and scheduling periodic preventive maintenances for these same resources.
-For the seek of simplicity, the following convention has been adopted: aircraft are called resources, missions are called tasks.
-
-  \subsection{Tasks}
-    \label{def:task}
-
-    There is a fixed set of $j \in \mathcal{J}$ tasks to be accomplished over an horizon of time divided into $t \in \mathcal{T}$ discrete periods.
-For their execution, these tasks require the assignment of a specific number of resources $R_j$ each period of time the task is active.
-The start and end periods for each task are known and a task is considered active between its start and end period.
-
-    During each period, tasks consume an amount of time equal to $H_j$ from each of its assigned resources.
-
-    % The assignment of a resource to a task is not decided for the whole duration of the task.
-After a minimum amount of time ($MT$), a resource can be freed and exchanged for another one, even if the task it is assigned to has not finished.
-    The total number of resources being used at any given time in a specific task should be equal to $R_j$.
-
-    Each task requires one and only one type of resource which, in addition, should comply with additional task requirements.
-
-  \subsection{Resources}
-    \label{def:res}
-
-    There is a set $i \in \mathcal{I}$ of available resources that are assigned to tasks in order to accomplish them.
-Each resource can only be assigned to a single task in any given period.
-These resources suffer from wear and tear and require regular maintenance operations during their lifetime.
-The need for maintenance is calculated based on two indicators.
-
-    The first one is called "remaining elapsed time" (or $ret_{it}$).
-It expresses the amount of time (measured in time periods) after which the resource cannot be used anymore and has to undergo a maintenance operation.
-Its value is calculated for each resource $i$ and each time period $t$.
-In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the resource $i$ can be used before needing a maintenance operation at any given period $t$.
-
-    % Additionally, after an absolute amount of time and/or usage ($aet_i$ or $aut_i$), the resource becomes obsolete. There is no way to reverse this process.
-
-    At any given period, including at the start of the planning horizon, each resource has a specific status given by remaining usage time and remaining elapsed time.
-
-    % \begin{itemize}
-        % \item remaining absolute elapsed time.
-        % \item remaining absolute usage time.
-        % \item remaining storage time (see \ref{def:sto}).
-        % \item type of last maintenance (see \ref{def:maint}).
-    % \end{itemize}
-
-  \subsection{Maintenances}
-    \label{def:maint}
-
-    Maintenances operations are the process by which resources that have reached a limit in some indicator can return to a state where they can continue to be used in tasks.
-
-    % Each maintenance belongs to a specific type of maintenance $m \in \mathcal{M}$.
-These types of maintenance differentiate between each other by having potentially different characteristics such as: different durations, the provision of new functionalities for the resource or the restoration of the storage capacity (see \ref{def:sto}).
-The maintenance type is not a decision to be made since there is a specific sequence of maintenance types a resources needs to follow according to the resource type.
-
-    Each maintenance operation has a fix duration of $M$ periods.
-
-    After a maintenance operation, a resource restores its remaining elapsed time and remaining usage time to their max values $E$ and $H$ respectively.
-
-    % Finally, resources are organized into families or groups. Each resource inside a family or group shares the same types of maintenances and, usually, is able to do the same type of tasks.
-
-  % \subsection{Storages}
-  %   \label{def:sto}
-
-  %   Following the rule of remaining elapsed time, even if a resource is not being used, it still needs to have a maintenance after a given amount of time has passed.
-In order to avoid this problem, the resource can be put into a storage state.
-
-  %   A resource in this states has to be kept for a minimum time of $sm$ periods. While in this state it cannot receive maintenance or be assigned any task.
-
-  %   Every resource has the capacity to be stored and this capacity is measured in a number of periods $sc$.
-In order for a resource to restore its remaining storage capacity, it needs to receive a specific maintenance (see \ref{def:maint}).
-Similar to the remaining elapsed time, after these maintenances, the resource recovers its storage capacity up to a certain level $S$.
-
-  \subsection{Possible states}
-
-    As a summary, the following are the possible logical states that a resource can be in: assigned to a task (see \ref{def:task}); under maintenance (see \ref{def:maint}); available.
-
-      % \begin{itemize}
-          % \item Under storage (see \ref{def:sto}).
-            % \item Obsolete (see \ref{def:res}).
-      % \end{itemize}
-
-    % Figure \ref{fig:solution} shows an extraction of an assignment of missions (red) and maintenances (blue) to aircraft (codes A1 - A9).
-
-    %   \begin{figure}
-    %     \centering
-    %       \includegraphics[width=0.5\textwidth]{./../../img/calendar.png}
-    %     \caption{Visual representation of part of the solution to an instance of the problem showing maintenances and assignments.
-Resources are listed in the vertical axis and the horizontal axis defines the time.
-Red periods correspond to task assignments while blue ones to maintenance assignments.}
-    %     \label{fig:solution}
-    %   \end{figure}
-
-  \subsection{Time}
-    \label{def:hor}
-
-    In planning tasks and maintenances, it is important to take into account the initial and end state of each resource.
-This initial state can be a maintenance or an assigned task.
-If a resource is under maintenance, it needs to continue in this state for its remaining maintenance time.
-Tasks' assignments should be taken into account in a similar manner.
-
-    The remaining used and elapsed times are assigned to each resource at the beginning of the planning horizon.
-
-    Finally, the state of each resource at the end of the planning horizon, its remaining (elapsed, used) time, needs to be defined.
-      % needs to be addressed too. This implies guaranteeing that the remaining (elapsed, used) time of resources is sustainable for future tasks and the final states of resources are not too skewed.
-
-  \subsection{Objectives}
-
-    In this study, the following objectives are considered.
-
-    Given that the creation of new tasks and the duration of maintenance are considered stochastic in real-life, one basic goal is to maximize the robustness of the planning by having the greatest amount of available resources at every period of the planning horizon.
-
-    Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of resources under maintenance over the planning horizon.
-
-    As it will be shown in the mathematical formulation, these objectives are quite related one with the other: the more resources are in maintenance in a given period, the more unavailable resources will be had.
-
-\section{Related work}
-  \label{sec:soa}
-
-  % TODO: add FMP biblio?
-
-  % The Military Flight and Maintenance Planning Problem considered here aims to assign missions and maintenance tasks to military aircraft.
-It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically since a round-trip to the base is assumed for each flight.
-
-  % It also includes different objectives and constraints.
-
-  % Although the former has been studied in literature in \citet{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015}, it has not received as much attention as the latter.
-In the following, the problem with already scheduled missions is considered.
-
-  The Military Flight and Maintenance Planning problem considered assigns missions and schedules maintenance operations to aircraft.
-In this problem, aircraft are needed to comply with a series of planned missions while at the same time needing to receive frequent preventive maintenances in order to be ready for new missions.
-
-  It is a variant of the Flight and Maintenance Planning problem, where maintenance operations and flights are scheduled to commercial aircraft.
-The main differences are the fact that flights are not modeled geographically: all aircraft are assumed to exit the main depot and return to the same depot at the end of the flight.
-
-  Other differences include the following.
-
-  Firstly, the objective function's motivation is on reducing the load on the maintenance facilities and guaranteeing the availability of the fleet instead of reducing the cost.
-Also, the maintenance operations take considerably longer and the rules that govern the maintenances are different.
-
-  The temporal scope of the problem is also larger, making the normal planning horizon and each period larger too: years instead of months and months instead of days, respectively.
-
-  In the following, the studies on the Military Flight and Maintenance Planning are discussed.
-
-  In \citet{Kozanidis2008}, a model for the Hellenic Air Force is presented.
-It assigns flight hours to aircraft and schedules flexible maintenance periods based on a "remaining usage time" rule.
-The objectives are to maximize this "remaining usage time" for the aircraft as well as to smooth the availability of aircraft during the planning horizon.
-Instances of 24 aircraft and 6 monthly periods were solved with Mixed Integer Programming (MIP) formulations and heuristics.
-
-  In \citet{Cho2011}, the objective is to smooth the maintenance operations.
-Randomly generated instances of 15 aircraft and 520 bi-daily periods were solved with MIP formulations.
-The final state of aircraft is taken into account in order to distribute as much as possible the "remaining usage time" of aircraft.
-
-  In \citet{Verhoeff2015}, a model for the Royal Netherlands Air Force is developed.
-Here, three criteria for improving the planning robustness are presented and applied in a MIP model: availability, serviceability and sustainability.
-
-  In \citet{gavranis2015exact}, an exact algorithm is presented and tested on instances of 6 periods and between 10 and 200 aircraft for the special case of maximizing the residual flight time availability or total remaining usage time.
-
-  Both \citet{Kozanidis2008} and \citet{Cho2011} develop heuristic alternatives to the MIP formulation for solving larger problem instances.
-In \citet{steiner2006heuristic}, a constructive, multi-stage heuristic approach is described to obtain balanced maintenances for the Swiss Air Force.
-Both calendar-based maintenances actions (CBMA) and usage-based maintenance actions (UBMA) are used.
-Finally, in \citet{kozanidis2014heuristics}, two heuristics are presented to solve large instances of the problem by taking consecutive decisions period by period with the help of mathematical models at each step.
-
-  % Regarding operational-level planning, a more detailed implementation can be found in \citet{safaei2011workforce}, where the specific conditions during the maintenance operations are taken into account to schedule the maintenance jobs, such as the workforce skills and availability.
-
-  It should be noted that previous formulations only consider an aggregate demand in flight hours for all aircraft.
-  The aircraft are assumed to be homogeneous and interchangeable.
-  The use of aggregate demand makes it impossible to take into account the task requirements for aircraft.
-   % The concept of mission is not introduced as part of these formulations, only a demand of flight hours is given for all aircraft in total.
-The aircraft themselves have been assumed as homogeneous and interchangeable.
-Due to this aggregation of demand, there are no minimal time assignments for aircraft to missions.
-Missions are important in order to guarantee that aircraft have the required capabilities needed for the missions they are assigned to.
-
-  % Additionally, the possibility of storage of aircraft is not present in the previous work.
-
-  Lastly, in \citet{Chastellux2016} a mission-centered modeled is presented.
-Here, cycles, composed of one maintenance followed by mission assignments were used.
-Instances of 6 - 70 aircraft and 60 monthly periods were solved.
-Close to optimal solutions for small to medium instances using MIP formulations and heuristics based on these MIP formulations were found.
-
-  In the following, an alternative model is presented. This model is then tested in a case study.
-
 \section{Optimization model}
@@
 
-    \begin{tabular}{p{5mm}p{70mm}}
-        % $\mathcal{I}$     &   \\
-        % $\mathcal{T}$     &   \\
-        % $\mathcal{J}$     &   \\
+    \begin{tabular}{p{5mm}p{170mm}}
+        % $\mathcal{I}$     &  resources. \\
+        % $\mathcal{T}$     &  time periods inside planning horizon. \\
+        % $\mathcal{J}$     &  tasks. \\
         $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
@@
         $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
+        $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
+        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{m}+1\}},  ..., {t}$\}. \\
+        $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{M}+1\}},  ..., \max{\{1, t - Ret^{m}+1\}}$\}. \\
     \end{tabular}
@@
 
-    \begin{tabular}{p{8mm}p{67mm}}
+    \begin{tabular}{p{8mm}p{167mm}}
         $H_j$             & amount of resource time required by task $j$. \\
@@
         $H$               & remaining usage time after a maintenance. \\
-        % $S$               & remaining storage time after (certain) maintenances. \\
+        $S$               & Storage duration in number of periods. \\
         $W_1$             & weight of the first objective in the objective function. \\
@@
         $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
+        $Ret^{M}$         & maximum number of periods between two consecutif maintenances. \\
+        $Ret^{m}$         & minimum number of periods between two consecutif maintenances. \\
     \end{tabular}
@@
 
-    \begin{tabular}{p{8mm}p{67mm}}
+    \begin{tabular}{p{8mm}p{167mm}}
         $a_{jti}$   &  =1 if task $j \in J$ in period $t \in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
         $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
+        $s_{it}$    &  =1 if resource $i \in I$ starts a storage operation in period $t \in \mathcal{T}$, 0 otherwise. \\
         $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
@@
                %avail
-       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \notag \\
-           &\hspace*{10mm}+ D_t\leq u_{max} 
+       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
         &t \in \mathcal{T} \label{eq:avalaibility1}\\
@@
                 & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
+        % & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} \leq 1 
         & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
@@
 Tasks' resource requirements are defined by (\ref{eq:taskres}).
-Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task or maintenance operation at the same period.
+Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task, a maintenance operation or a storage operation at the same period.
     \begin{align}
         % remaining used time
-         & rut_{it} \leq rut_{it-1} + H m_{it} \notag \\ 
-           &\hspace*{4mm}- \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j \notag \\
-                && t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
+         & rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
         & rut_{i0} = Rut^{Init}_i
@@
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\               
-        & ret_{it} \leq ret_{it-1} - 1 \notag \\
-          &\hspace*{4mm}+ E m_{it}
-                & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
-        & ret_{i0} = Ret^{Init}_i
-                & i \in \mathcal{I} \label{eq:ret_initial}\\
-        & ret_{it} \geq E m_{it}
-                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
-        & ret_{it} \in [0,E]
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
-        & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
-              & t = |\mathcal{T}| \label{eq:min_ret}\\
+        % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}
+        %         & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
+        % & ret_{i0} = Ret^{Init}_i
+        %         & i \in \mathcal{I} \label{eq:ret_initial}\\
+        % & ret_{it} \geq E m_{it}
+        %         & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
+        % & ret_{it} \in [0,E]
+        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
+        % & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
+        %       & t = |\mathcal{T}| \label{eq:min_ret}\\
         & \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
-              & t = |\mathcal{T}| \label{eq:min_rut}
+              & t = |\mathcal{T}| \label{eq:min_rut} \\
+        & m_{it'} + m_{it} \leq 1
+          & t, \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
+        & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \leq  m_{it}
+          & t, \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\ 
     \end{align}
+
+    % \begin{align}
+    % & m_{it'} + m_{it} \leq 1
+    %   & t, \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_better}\\ 
+    % & \sum_{t' \in \mathcal{T}^{M}_t - \mathcal{T}^{m}_t} m_{it'} \leq  m_{it}
+    %   & t, \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_better}\\ 
+    % \end{align}
+
         % These constraints calculate the balances of hours for each resource.
     The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}). 
-    Similarly, the remaining elapsed time is defined by (\ref{eq:ret_upper})-(\ref{eq:ret_initial}) and its limits by (\ref{eq:ret_lower})-(\ref{eq:me}). 
-    Finally, constraints \ref{eq:min_ret} and \ref{eq:min_rut} guarantee that resources have, globally, the same amount of remaining used and elapsed times at the beginning and at the end of the planning horizon.
+    The remaining elapsed time is defined by (\ref{eq:ret_min})-(\ref{eq:ret_max}).
 
-\section{Datasets}
-  \label{sec:data}
-  The dataset being used was provided by the French Air Force and already used in \citet{Chastellux2017}.
-  Based on the information provided, a number of instances were created as subsets of the main dataset in order to run different experiments.
-
-  Table \ref{tab:instance} shows the characteristics of each instance. The main differences between instances are in the number of tasks and the number of periods.
-
-  The criteria for the creation of the instances were the following:
-
-    \begin{enumerate}
-        \item The original instance was the largest instance where an integer solution was found in less than 3600 seconds: it contains 11 tasks and 21 periods in total.
-        \item We took out the task with the biggest number of required resources.
-        \item We increased the number of periods in multiples of 10 until no integer solutions can be found in 3600 seconds.
-        \item We repeated step 2.
-    \end{enumerate}
-
-  \begin{table}
-  \begin{center}
-
-  \input{tables/instance.tex}
-
-  \caption{Instances used in the experiments with indicators of size: "id" is the instance; $|\mathcal{T}|$ is the number of periods; "assign" is the number of assignments of resources to tasks; $|\mathcal{J}|$ is the number of tasks; "vars" the number of variables; "cons" is the number of constraints; and "nonzeros" is the number of non zero values in the matrix.}
-  \vspace{-0.5cm}
-  \label{tab:instance}
-  \end{center}
-  \end{table}
-
-  \subsection{Tasks}
-
-    The dataset includes 11 tasks.
-Most of them demand between 1 and 8 resources per period, which is equal to one month.
-The most loaded and second most loaded tasks demand 50 and 20 resources respectively per period.
-    Candidate resources are calculated based on each task's requirements matching the candidate's capabilities.
-
-    Some tasks have a limited duration, meaning they start and end during the planning horizon. Others are active during the whole planning horizon.
-
-  \subsection{Maintenances}
-
-    All maintenance operations last exactly 6 periods.
-Each maintenance sets the remaining usage time of a resource to 1000 hours.
-After it, a resource can be without maintenance for a maximum of 60 consecutive periods.
-The nominative capacity for maintenance is 18 resources at any given time.
-
-  \subsection{Resources}
-
-    There are 127 resources in total. Each one has its own functionalities and past assignments.
-
-    % \begin{figure}
-    %   \centering
-    %     \includegraphics[width=0.5\textwidth]{./../../img/initial_used.png}
-    %   \caption{Distribution of Usage Times among resources at the beginning of the planning horizon}
-    %   \label{fig:histogramUsage}
-    % \end{figure}
-
-% \section{Heuristics}
-
-\section{Results}
-  \label{sec:results}
-
-  \subsection{Numerical tests}
-    \label{sec:tests}
-
-    All tests were run on a 64 bit Ubuntu 16.04 workstation with 16GB of RAM and Intel i7-6700HQ CPU \@ 2.60GHz x 8 processor.
-    Several experiments were done with the MIP formulation on problem instances presented in table \ref{tab:instance}.
-
-    Table \ref{tab:summary} gives an overview of the obtained results. As can be seen, smaller instances are solved up to optimality while medium-size ones are stopped after one hour.
-
-    For this calculation, both weights on the objective function were set at 1. A sensibility analysis of these two parameters is provided in section \ref{sec:multi}.
-
-    \begin{table}
-    \begin{center}
-
-    \input{tables/summary.tex}
-
-    \caption{Information on the solution of each tested instance: "id" is the instance identifier; "objective" is the value of the objective function; "gap" is the percentage gap between the integer solution and the linear relaxation; "time (s)" is the solution time or 3600; and "bound" is the greatest lower bound found from the linear relaxation.}
-    \vspace{-0.5cm}
-    \label{tab:summary}
-    \end{center}
-    \end{table}
-
-    Table \ref{tab:results} shows the details of the solving process for each instance.
-As it can be seen, cuts do not improve significantly the relaxation (the value of "bound cuts" is very close to the root solution).
-In fact, the original linear relaxation (root) is quite similar to the final lower bound (bound).
-This suggests that further improvement in modelization would help to reduce the gap faster.
-
-    An example of a tighter modelization could be to replace the constraints that count the remaining usage time and elapsed time for tighter equivalents.
-
-    The greatest number of cuts used was of 'Implied bound' and 'Mixed integer rounding', depending on the instance.
-Since these are two generic-type cuts, it can be concluded that the solver has not detected any particular structure as part of this problem.
-
-    \begin{table}
-    \begin{center}
-
-    \input{tables/results.tex}
-
-    \caption{Solution details on the progress of the lower bound: "id" is the name of the instance; "root" is the relaxation before cuts; "b.
-cuts is the relaxation after the cuts; "bound" is the relaxation at the time limit; "cuts (\#)" is the number of cuts; and "cuts (s)" is the time the cuts took in seconds.}
-    \vspace{-0.5cm}
-    \label{tab:results}
-    \end{center}
-    \end{table}
-
-    Table \ref{tab:results2} shows the progress of the integer solution for each instance.
-For more detail on this progress, figure \ref{fig:progress} shows an example of progress for instance with id=I\_7.
-It can be seen that initial solutions are quite far from the optimal ones due to the use of heuristics at the beginning of the solving process.
-In fact, for two rather small instances the solver was able to find an optimal solution before starting to branch.
-
-    The obtained results show that the branching is not providing much improvement in increasing the lower bound or finding better integer solutions.
-The quantity of symmetries in the model, represented as candidates for each task, could be a reason the difficulty in solving large instances.
-It may exist a very large number of possible candidates to be potentially assigned for each task with respect to the requirements of resources for each task.
-This reduces the branching capacity of the solver, which tries to explore many nodes that are almost equivalent in reality.
-
-    An example of an alternative to break these symmetries could be to slightly change the objective function so it can measure more gradually the smoothness of the objective and help get better bounds.
-Another alternative could be to assign priorities to some combinations of resource-task or to limit the number of candidates that can be assigned to each task by some heuristic that clusters resources with tasks.
-Finally, arbitrarily preassigning (fixing) the final state (in terms of remaining usage time) to each resource could also help break those symmetries without important loss on the objective function value.
-
-    \begin{table}
-    \begin{center}
-
-    \input{tables/results2.tex}
-
-    \caption{Solution details on the progress of the integer solution: "id" is the name of the instance; "first" is the first solution integer found; "sol.
-cuts" is the solution after the cuts; "last" is the last solution found at the time limit.}
-    \vspace{-0.5cm}
-    \label{tab:results2}
-    \end{center}
-    \end{table}
-
-    \begin{figure}
-      \centering
-        \includegraphics[width=0.5\textwidth]{./../../img/progress.png}
-      \caption{Solving progress of the integer solution (in blue) and the lower bound (red) along the solver iterations (x axis). Instance id=I\_7.}
-      \label{fig:progress}
-    \end{figure}
-
-    Figures \ref{fig:num-maintenances} and \ref{fig:num-unavailables} show an example distribution of maintenance operations and unavailable numbers of resources per period, respectively.
-
-    \begin{figure}
-      \centering
-        \includegraphics[width=0.5\textwidth]{./../../img/num-maintenances.png}
-      \caption{Solution for instance id=I\_7: number of resources under maintenance at each period.}
-      \label{fig:num-maintenances}
-    \end{figure}
-
-    \begin{figure}
-      \centering
-        \includegraphics[width=0.5\textwidth]{./../../img/num-unavailables.png}
-      \caption{Solution for instance id=I\_7: number of unavailable resources at each period.}
-      \label{fig:num-unavailables}
-    \end{figure}
-
-  \subsection{Bi-objective analysis}
-  \label{sec:multi}
-
-    The further calculation was done in order to analyze the impact of the weights $W_1$ and $W_2$ on the solution.
-In it, both weights were changed between 0 and 1 with pace of 0.1 under the condition that $W_1 + W_2 = 1 $.
-This was done for all previously defined instances.
-
-    Table \ref{tab:multmultiobj} presents the obtained results of this analysis. It can be seen that all instances have a small number of non-dominated solutions.
-
-    \begin{table}[h]
-    \begin{center}
-
-    \input{tables/multi-multiobj.tex}
-
-    \caption{Non-dominated points found on each instance after 3600 seconds.
-All instances were analyzed: "First" and "Last" indicate the two extremes of the calculated Pareto frontier; "\# points" indicates the number of Pareto optimal points found.
-Not all experiments were solved up to optimality.}
-    \vspace{-0.5cm}
-    \label{tab:multmultiobj}
-    \end{center}
-    \end{table}
-
-    More particularly, an individual instance was analyzed, where an optimal solution was found in each experiment.
-    Table \ref{tab:multiobj} shows the obtained results. The instance used consisted of 21 periods and 10 tasks corresponding to instance with id=I\_5.
-
-    For all solved instances, a small number of non dominated solutions was observed.
-    \begin{table}[h]
-    \begin{center}
-
-    \input{tables/multiobj.tex}
-
-    \caption{Values of $u_{max}$ (unavailable) and $m_{max}$ (maint) for different combinations of $W_{1}$ and $W_{2}$. Instance id=I\_5. All experiments were solved to optimality.}
-    \vspace{-0.5cm}
-    \label{tab:multiobj}
-    \end{center}
-    \end{table}
-
-    \begin{figure}
-      \centering
-        \includegraphics[width=0.5\textwidth]{./../../img/multiobjective.png}
-      \caption{Pareto diagram for instance id=I\_5.}
-      \label{fig:pareto}
-    \end{figure}
-
-\section{Conclusions and further work}
-  \label{sec:conclusions}
-
-  This paper proposed a new MIP formulation to address a specific FMP with an heterogeneous fleet of aircraft.
-  Two objective functions were chosen to address the French Air Force requirements.
-  The first one was to smooth maintenance operations and the second one was to increase aircraft availability.
-  The characteristics of MIP convergence were analyzed on real-world instances.
-  Mono-objective problem instances were solved to optimality for medium size instances. Then, exact Pareto frontiers for bi-objective formulation were calculated for these instances. 
-  The use of these two objectives does not lead to the same optimal solution but very good compromise solutions can be found due to the bi-objective optimization.
-  For larger instances, difficulties in finding optimal solutions occurred because of multiple symmetries and a poor quality of MIP generic primal heuristics.
-
-  The first perspective is to improve the resolution of large size instances.
-Heuristics could be implemented, but also mathematical programming can be improved with symmetry breaking or MIP reformulations.
-  For the bi-objective optimization for large size instances, more advanced bi-objective methods such as $\epsilon$-constraints or Two-Phases-Method should be investigated for the computation of larger Pareto frontiers.
-  Lastly, additional constraints from the real world application can be incorporated in the model.
-
-\bibliographystyle{dcu}
-\bibliography{./../biblio/MFMP}
+    Finally, constraints \ref{eq:min_rut} guarantee that resources have, globally, the same amount of remaining used and elapsed times at the beginning and at the end of the planning horizon.
 

### 2018-06-22 R/diagrams/thesis_gantt.R

@@ -8,18 +8,19 @@ gantt_thesis <- function(...){
     
     data <- 
         data.table(
-            Phase=c(1, 1, 2, 3, 4, 5, 6)
+            Phase=c(1, 1, 2, 3, 4, 5, 6, 7)
             ,Libelle=c('État de l’art FMP','Travaux antérieurs','Nouveaux modèles',
-                       'Incertitude','Cas d’étude','Prototype logiciel','Mémoire de thèse')
-            ,month_start=c(1, 1, 9, 17, 25, 25, 31)
-            ,month_end= c(8, 8, 16, 24, 30, 30, 36)
+                       'Incertitude','Cas d’étude','Prototype logiciel','Mémoire de thèse',
+                       'Collaboration Dassault')
+            ,month_start=c(1, 1, 9, 17, 25, 25, 31, 9)
+            ,month_end= c(8, 8, 16, 24, 30, 30, 36, 30)
         ) %>% 
         mutate(start= start_month %m+% months(month_start-1),
                end= start_month %m+% months(month_end),
                group= Phase,
                id= c(1:nrow(.)),
                content = Libelle,
-               color = RColorBrewer::brewer.pal(5, "YlOrRd")[rep_len(1:5, n())],
+               color = RColorBrewer::brewer.pal(7, "YlOrRd")[rep_len(1:7, n())],
                style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color),
         )
     config <- list(

### 2018-06-25 R/presentations/Dassault_presentation/Dassault_20180606_v3.tex
Moved presentations to dedicated directory

--- 2018-06-22 R/Dassault_presentation/Dassault_20180606_v2.tex
@@
 
-Une flotte hétérogène d'avions avec les caractéristiques suivantes:
+Une flotte hétérogène d'avions, dont chacun a les caractéristiques
+suivantes :
 
@@
 \item
-  Temps écoulé restant.
+  Butée calendrier.
 \item
-  Des heures de vol restant.
+  Butée horaire.
 \item
@@
 
-Un ensemble de missions à réaliser, chaque une avec:
+Un ensemble de missions à réaliser, chacune avec :
 
@@
 \item
-  Un besoin mensuelle d'avions.
+  Un nombre nécessaire d'avions par mois.
 \item
-  Une consommation d'heures par avion-mois affecté.
+  Une consommation d'heures par avion et par mois.
 \item
-  Des fonctionnalités requise.
+  Des fonctionnalités requises.
 \end{itemize}
@@
 
-Chaque maintenance suit les règles suivantes:
+Chaque maintenance suit les règles suivantes :
 
@@
 \item
-  Le nombre maximal \emph{d'heures de vol} entre deux maintenances est
+  Le nombre maximal d'heures de vol entre deux maintenances est
   \emph{1000h}.
 \item
-  Le nombre maximal de \emph{mois calendrier} entre deux maintenances
-  est \emph{60 mois}.
+  Le nombre maximal de mois entre deux maintenances est \emph{60 mois}.
 \end{itemize}
@@
 \item
-  Minimiser le coût (réduisant le nombre de maintenances).
+  Minimiser le coût (en réduisant le nombre de maintenances).
 \end{itemize}
@@
 \item
-  RQ001: heures de missions
+  RQ001: heures de missions.
 \item
-  RQ002: aéronefs assignés
+  RQ002: aéronefs affectés.
 \item
-  RQ003: besoins de maintenance
+  RQ003: besoins de maintenance.
 \item
-  RQ004: durée de maintenance
+  RQ004: durée de maintenance.
 \item
-  RQ005: disponibilité
+  RQ005: disponibilité.
 \item
-  RQ006: capacité de maintenance
+  RQ006: capacité de maintenance.
 \item
-  RQ007: état final
+  RQ007: état final.
 \item
-  RQ008: stockage
+  RQ008: stockage.
 \end{itemize}
@@
 
+\begin{quote}
+FMP: Flight and Maintenance Planning problem.
+\end{quote}
+
 \begin{itemize}[<+->]
@@
 \item
-  FMP: Flight and Maintenance Planning problem.
+  Minimisation du nombre maximal de maintenances. Affectations
+  quotidiennes. États-Unis, Cho (2011).
 \item
-  En Cho (2011), avions de l'Armée de l'air états-unienne ont été
-  affectés à des opérations quotidiens pour minimiser le nombre maximal
-  de maintenances.
+  Maximisation de disponibilité de vol.~Affectations mensuelles Grèce,
+  Kozanidis (2008).
 \item
-  En Kozanidis (2008), des heures de vol et des maintenances ont été
-  affectées aux avions de la flotte grecque.
-\item
-  En Verhoeff, Verhagen, and Curran (2015), différents objectifs ont été
-  essayé pour les affectations d'heures de vol et maintenances.
+  Minimisation du nombre maximal de maintenances. Affectations
+  mensuelles. Pays-Bas. Verhoeff, Verhagen, and Curran (2015).
 \end{itemize}
@@
 \item
-  ont étudié de flottes homogènes: nous avons de flottes hétérogènes.
+  ont étudié de flottes homogènes: nous avons une flotte hétérogène.
 \item
-  n'ont pas le mêmes contraintes (stockages, 60 périodes entre
-  maintenances) ou les mêmes objectifs (maximiser disponibilité).
+  n'ont ni le mêmes contraintes ni les mêmes objectifs.
 \item
@@
 
-\section{Prochains pas}\label{prochains-pas}
+\section{Perspectives}\label{perspectives}
 
-\begin{frame}{Prochains pas}
+\begin{frame}{Perspectives}
 
@@
 \item
-  \textbf{Continuer le dialogue avec les utilisateurs}: ajouter des
+  \textbf{Continuer le dialogue avec les utilisateurs}: ajouter de
   nouvelles contraintes, valider les résultats.

### 2018-06-25 R/presentations/OptiPlan_presentation_20180327/proposal_fr.Rmd
Moved presentations to dedicated directory

---
title: Optiplan proposal
author: ISAE-SUPAERO, Université de Toulouse
date: Mars 27, 2018
output:
  slidy_presentation:
    # css: ./../../css/baobab_MH.css
    # css: ./../../css/baobab_docs.css
    footer: ISAE-SUPAERO, Université de Toulouse, France
    duration: 10
    # incremental: yes
    # slide_level: 3
    # pandoc_args: ["--slide-level=3"]
    # font_adjustment: +4
    toc: yes
  beamer_presentation:
    pandoc_args: ["--slide-level=3", "--top-level-division=part"]
    # fig_height: 4
    # includes:
    #   in_header: header.tex
    # incremental: yes
    # slide_level: 3
    theme: Madrid
    # toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("../")
source('diagrams/cases.R')
source("diagrams/architecture.R")
source("diagrams/functionalities.R")
html_output <- knitr::opts_knit$get("rmarkdown.pandoc.to") %in% c("slidy", 'html')
```

<!-- dfg -->
<!-- ```{r, echo=FALSE, out.width='80%'} -->
<!-- knitr::include_graphics('../../img/logos-defi-optiplan.jpg') -->
<!-- ``` -->

<!-- ## Présentation de l'entreprise -->

<!-- Nom: ISAE-SUPAERO -->
<!-- Activité: école d’ingénieurs. -->

<!-- Grand école du secteur aéronautique et spatial, l'ISAE-SUPAERO développe depuis 2007 une recherche propre portée par ses département. Le département d'ingénierie des systèmes complexes (DISC), par son groupe Systèmes Décisionnels, travaille sur les outils de décision appliqués aux systèmes industriels, en particulier pour la production et la maintenance dans l'industrie aérospatiale. -->

<!-- ```{r, echo=FALSE, out.width='50%'} -->
<!-- knitr::include_graphics('../../img/logo-isae-supaero.png') -->
<!-- ``` -->

## Savoir-faire

* recherche scientifique, publication d’articles
* projets de recherche en collaboration avec l’industrie
* conception d’applications logicielles en planification et optimisation pour des entreprises

L’équipe a déjà encadré des travaux de recherche sur le sujet de la maintenance aéronautique. En outre, elle a une connaissance du problème en question car **une thèse doctorale** sur ce sujet a été lancée en octobre 2017. Les travaux ont été présentés à la conférence **ROADEF 2018** et seront présentés dans la conférence **MOSIM 2018**.

## Expérience: Génération des horaires du personnel

```{r, echo=FALSE, out.width='80%'}
knitr::include_graphics('../../img/CLH.png')
```

## Fonctionalitées

```{r, echo=FALSE}
if (html_output){
    graph_functionalities()
} else {
    knitr::include_graphics('../../img/functionalities.png')
}
```

## Moteur de planification: Modélisation mathématique

* Notre point fort c'est la modélisation efficace.
* Nous avons dèjà des modèles faits pour ce problèm.
* Des alternatives d'hybridations vont être étudiées.

\begin{align}
& \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
  & t \in \mathcal{T} \label{eq:capacity1}\\
& \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
  & t \in \mathcal{T} \label{eq:avalaibility1}\\
& \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
        & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
& \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
        & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
\end{align}

<!-- ## Configuration -->

<!-- Les options disponibles seront: -->

<!-- * **Temps maximum**: dans le cas où une solution est nécessaire rapidement sans prouver l’optimalité. -->
<!-- * **Gap maximum**: afin d’autoriser une distance maximale entre la meilleure solution obtenue et la meilleure solution hypothétique possible. Cette distance peut être une différence proportionnelle ou une différence absolue. -->
<!-- * **Premier et dernier mois à planifier**: afin de définir l’horizon de planification. -->
<!-- * **Poids des objectifs**: pour optimiser les différents objectifs et privilégier un objectif ou un autre. -->
<!-- * **Paramètres de maintenance**: durée de maintenance (en mois), temps d’utilisation maximal, temps écoulé maximal et autres paramètres liés à la maintenance. -->

## Cas et expérimentation

Les fonctionnalités disponibles seront les suivantes:

* Créer un nouveau cas.
* Copier un cas existant.
* Supprimer un cas.
* Comparer deux cas.
* Publier un cas.
* Exporter un cas.

```{r, echo=FALSE}
if (html_output){
    graph_cases()
} else {
    knitr::include_graphics('../../img/cases.png')
}
```

## Technologies et techniques

Tout les technologies à utiliser seront open source.\*

* **Modélisation mathématique**: pour modéliser les besoins des planifications.
* **Solveur mathématique**: pour générer les planifications.
* **Python**: pour la modélisation, l’heuristique et l’intégration de données.
* **R**: pour les statistiques, possibles prévisions et la visualisation des données.
* **PostgreSQL**: pour stocker des données.
* **HTML5**: pour les rapports, et les visualisations.

\* Le solveur est conditioné aux données et la taille du problème. Cela pourrait être CBC (libre), GUROBI ou CPLEX (payants) en fonction des besoins du modèle.

```{r, echo=FALSE, results='asis', out.width='15%'}
paths <- 
    c('postgresql.svg', 'Rlogo.png', 'python.png', 
      'coinor.jpg', 'html5.png') %>% sprintf('../../img/%s', .)
knitr::include_graphics(paths)

```

### 2018-06-25 R/presentations/dga_diaporama/dga.Rmd
Moved presentations to dedicated directory

--- 2018-04-03 R/dga_diaporama/dga.Rmd
@@
 author: "Franco Peschiera"
-# date: "`r format(Sys.time(), "%Y-%m-%d")`"
+date: "17 avril, 2018"
 output:
+  beamer_presentation:
+    theme: "Berlin"
+    # toc: true
+    slide_level: 1
+    incremental: yes
+    fig_height: 4
+    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
+    includes:
+        in_header: header.tex
   slidy_presentation:
@@
     # css: ./../../css/baobab.css
-    incremental: yes
-  beamer_presentation:
     incremental: yes
@@
 } else {
-    knitr::include_graphics('../../img/tasks_gantt.png')
+    knitr::include_graphics('../../img/calendar2.png')
 }

### 2018-06-25 R/presentations/MOSIM_presentation/presentation.Rmd
MOSIM presentation added

--- 2018-03-01 R/roadef_presentation/roadef_presentation.Rmd
@@
     | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
-    | \tiny (franco.peschiera@isae-supaero.fr)
 institute: "ISAE SUPAERO"
-date: "February 22, 2018"
+date: "June 27, 2018"
 output: 
@@
 bibliography: MFMP.bib
+editor_options: 
+  chunk_output_type: console
 ---
@@
 knitr::opts_chunk$set(echo = FALSE)
-setwd("../")
-source('scripts/params.R')
-source('scripts/import_results.R')
+path_to_R_dir <- "./../../"
+setwd(path_to_R_dir)
+source('functions/params.R')
+source('functions/import_results.R')
 exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
@@
 } else {
-    knitr::include_graphics('./../../img/calendar.png')
+    knitr::include_graphics('../img/calendar.png' %>% paste0(path_to_R_dir, .))
 }
@@
 ```{r out.width="150pt", echo=FALSE}
-knitr::include_graphics("./../../img/progress.png.png")
+knitr::include_graphics("../img/progress.png.png" %>% paste0(path_to_R_dir, .))
 ```
@@
 ```{r out.width="180pt", echo=FALSE}
-knitr::include_graphics("./../../img/multiobjective.png")
+knitr::include_graphics("../img/multiobjective.png" %>% paste0(path_to_R_dir, .))
 ```

### 2018-06-26 R/presentations/MOSIM_presentation/presentation.Rmd
ppt MOSIM

@@
 
-* The present model deals with an heterogeneus fleet, with different standards and capacities.
+* The present model deals with an heterogeneous fleet, with different standards and capacities.
 * Multiple objectives have been incorporated in order to satisfy the French Air Force's specific needs.
@@
 
-* Analysis for all instances: not many pareto optimal points available (between two and four).
+* Analysis for all instances: not many Pareto optimal points available (between two and four).
 * The objectives are quite aligned one with the other.
 
-# Next steps
+# Current and next steps
+
+## Current steps
+
+* **Linent objectives**: minimized total number of maintenances with a hard maximum on total maintenances per month.
+* **Multiple solvers**: GUROBI, CPLEX, CBC, CPO, CHOCO.
+
 ## Next steps
@@
 * **Try different objective formulations**: availability - maintenability - future state.
-* **Try more linent objectives**: relax the `max` in the objective.
 * **Break symmetries**: clustering candidates, fixing end availabilities.
 * **Add more constraints**: minimal duration of assignments, storage of aircraft.
+* **Compare solvers exhaustively**: to determine when each one can be advantageous.
 

### 2018-07-05 R/presentations/MOSIM_presentation/presentation.Rmd
MOSIM presentations last modifications

@@
 source('functions/import_results.R')
-exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
-gantt <- print_solution(exp_directory, max_resources = 30)
+# exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
+# gantt <- print_solution(exp_directory, max_resources = 30)
 show_gantt <- knitr::opts_knit$get("rmarkdown.pandoc.to")=="slidy"
@@
 * Multiple objectives have been incorporated in order to satisfy the French Air Force's specific needs.
+* Asignments to missions are done instead of assignments of flight hours.
 * This model also features several improvements in the MIP modelling that permit greater instance sizes.
@@
 
-* **Try different objective formulations**: availability - maintenability - future state.
+* **Add useful cuts**: for example limiting maintenance possibilities.
 * **Break symmetries**: clustering candidates, fixing end availabilities.
 * **Add more constraints**: minimal duration of assignments, storage of aircraft.
-* **Compare solvers exhaustively**: to determine when each one can be advantageous.
+* **Add robustness**: duration of maintenances, new /different missions, unavailable aircraft.
 

### 2018-07-05 R/presentations/journee_disc/presentation.Rmd

---
title: "Optimisation des maintenances et des missions des aéronefs militaires"
author: "Franco Peschiera"
date: "Juillet 06, 2018"
output:
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
---

## Présentation

* **Date de début et de fin de thèse**: 16/10/2017 - 16/11/2020
* **Ecole Doctorale**: Aéronautique et Astronautique
* **Encadrants**: Alain Haït, Olga Battaïa, Nicolas Dupin.
* **Publications 1re auteur**: ROADEF (France), MOSIM (France), CLAIO (Pérou)
* **Communications**: ROADEF (France), OPTIPLAN (défi DGA), MOSIM (France)
* **Financement**: ISAE + DGA

## Principaux travaux faits

* Spécification du problème.
* État de l'art.
* Formalisation du modèle.
* Expérimentation.

## Axes de recherche prévus pour 2018/2019

* Complexité du problème.
* Comparaison entre différentes méthodes de résolution.
* Incertitudes.

### 2018-07-10 R/presentations/MOSIM_presentation/presentation.Rmd
adapted MOSIM html presentation.

@@
 source('functions/import_results.R')
-# exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
-# gantt <- print_solution(exp_directory, max_resources = 30)
+exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
+gantt <- print_solution(exp_directory, max_resources = 30)
 show_gantt <- knitr::opts_knit$get("rmarkdown.pandoc.to")=="slidy"

### 2018-07-10 R/presentations/OptiPlan_presentation/proposal_fr.Rmd
adapted MOSIM html presentation.

--- 2018-03-20 R/OptiPlan_presentation/proposal_fr.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE)
-setwd("../")
-source('scripts/params.R')
-source('scripts/import_results.R')
+setwd("./../")
+source('functions/params.R')
+source('functions/import_results.R')
 source('diagrams/cases.R')

### 2018-07-10 R/appShiny/server.R

@@ -6,6 +6,7 @@ library(lubridate)
 library(data.table)
 
 completed_experiments <- function(path){
+    # browser()
     names <- path %>% list.files()
     with_solution <- path %>% 
         list.files(full.names = T) %>% 

### 2018-07-10 latex/notes/complexite/FMP_variants.tex

@@
 \def\KK{{\mathcal{K}}}
+\def\JJ{{\mathcal{J}}}
 
@@
 
-    \subsection{Possible prove}
+    \subsection{Possible proof}
 
@@
      \item Each task assignment needs M/3 consecutif periods.
-     \item Each task assignment has an flight hours consumption.
+     \item Each task assignment has a flight hours consumption.
     \end{itemize}
@@
         \item Maintenances are evenly spaced.
-        %\item The missions require 0 aircraft at each period.
+        \item All maintenances start at the same time. Maybe only one maintenance per Aircraft.
+        % \item The missions require 0 aircraft at each period.
         % \item Each aircraft can only be used for one mission.
     \end{itemize}
+
+    \subsection{Limitations}
+
+    \begin{itemize}
+        \item The time at which each task is done matters.
+        %\item There is only one type of resource.
+        % \item All durations are the same.
+    \end{itemize}
+
+    \subsection{Notations}
+
+    \textbf{Sets}
+
+    \noindent{
+    \begin{tabular}{ll}
+        $i \in \II = [\![1, I]\!]$ & Set of aircraft.\\
+        % $k \in \KK_i = [\![1, K_i]\!]$ & Set of maintenances for aircraft $i$.\\ 
+        $j \in J$ & Set of missions.\\ 
+        % $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
+        % $k \in \mathcal{R}$ & Types of resources that maintenance can take. We only have one type.\\
+        % $i = 0$ & Artificial starting node.\\
+        % $i = n+1$ & Artificial ending node.\\
+    \end{tabular}
+    }
+
+    \subsection{Parameters}
+
+    \begin{tabular}{ll}
+    	$h_j$ & Flight hour consumption for each mission.\\
+    	$M$ & Duration of maintenances.\\
+    	$H$ & Flight hours given by maintenance.\\
+        % $p_i$ & Duration of maintenance $i$. In this case it's always p=6.\\
+        % $\mathbf{d}$ & Last period on the horizon.\\
+        % $r_{ik}$ & Resources of type $k$ taken by maintenance $i$. In this case it's always 1.\\
+        % $d^{max}_{ij}$ & Maximum duration between two maintenances $i$ and $j$.\\
+        % $d^{min}_{ij}$ & Minimum duration between two maintenances $i$ and $j$.\\
+        % ${\delta}_{ij}$ & Precedence duration arcs between two maintenances $i$ and $j$ based on $d^{max}$ and $d^{min}$.\\
+        % $R_k$ & total maintenance capacity at each period for resource $k$.\\
+        % $c_k$ & cost per unit of investment on resource $k$.
+        % $\mathbf{{To}}_{i,k}$ & Minimum starting period for maintenance $k$ of resource $i$.\\
+        % $\mathbf{{Ta}}_{i,k}$ & Maximum starting period for maintenance $k$ of resource $i$.\\
+    \end{tabular}
+
+    In theory, with this configuration we have a Three Partition problem. The problem lies that each task has a fixed time. They're not tasks, they're demand.
+
+    % This doesn't make sense because the tasks are already defined. We cannot chose when it executes.
+
+    % I'll continue with the idea nonetheless.
+    % Imagine we preassign the aircraft a group of tasks. We need to decide at which moment we want to assign the tasks in the predefined maintenances.
+    % In this way, we need to come up with an assignment of tasks for each cycle that has a total of 1000 flight hours.
+
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Three partition problem v2}
+
+    Inspired by \cite{Caprara2003}.
+
+    \subsection{Hypothesis}
+
+    \begin{itemize}
+     \item all maintenances are already decided for all aircraft and they will happen in month 4.
+     \item all aircraft come from receiving a maintenance so are full in availability and flight hours.
+     \item each maintenance gives C flight hours.
+     \item objective: comply with all the missions in time and arrive with 0 flight hours to the maintenance.
+     \item each mission lasts for 3 months, needs 1 aircraft each month.
+     \item There is no minimal assignment duration for missions.
+    \end{itemize}
+
+    % \subsection{Reductions}
+
+    % \begin{itemize}
+    %     \item The missions consume 0 hours of aircraft.
+    %     %\item The missions require 0 aircraft at each period.
+    %     \item Each aircraft can only be used for one mission.
+    % \end{itemize}
 
@@
     \begin{tabular}{ll}
-        $i \in \II = [\![1, I]\!]$ & Set of aircraft.\\
-        $k \in \KK_i = [\![1, K_i]\!]$ & Set of maintenances for aircraft $i$.\\ 
-        $j \in J$ & Set of missions.\\ 
-        % $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
-        % $k \in \mathcal{R}$ & Types of resources that maintenance can take. We only have one type.\\
-        % $i = 0$ & Artificial starting node.\\
-        % $i = n+1$ & Artificial ending node.\\
+        $i \in \II = [\![1, M]\!]$ & Set of aircraft fleets. Each fleet has 3 aircraft.\\
+        $j \in \mathcal{J} = [\![1, 3M]\!]$ & Set of missions.\\
+        $t \in \TT = [\![1, 3]\!]$ & Set of periods.\\
+        $p \in \mathcal{P} = [\![1, 3]\!]$ & Set of aircraft per fleet.\\
+        $A_{ip}$ & Aircraft number $p$ of fleet $i$. \\
     \end{tabular}
@@
 
-    This doesn't make sense because the tasks are already defined. We cannot chose when it executes.
+    \vskip 0.3cm
+
+    \subsection{Parameters}
+
+    \begin{tabular}{ll}
+        $M$ & Number of fleets.\\
+        $3M$ & Number of aircraft and missions.\\
+        $D$ & Months between maintenances. Fixed to three.\\
+        $h_j$ & Flight hour consumption for each mission.\\
+        $C$ & Flight hours given by maintenance.\\
+    \end{tabular}
+
+    \subsection{Variables}
+
+    \begin{tabular}{ll}
+        $F_{ij}$ & Mission $j$ is done by fleet $i$. Binary. \\
+        $a_{ipjt}$ & Mission $j$ is done by aircraft number $p$ from fleet $i$ in period $t$ (auxiliary). Binary. \\
+    \end{tabular}
+
+    \subsection{Model}
+
+    It's a satisfaction problem where we want to use exactly all the available hours to do missions.
+We could also calculate the "waste hours just before the maintenance" and then it could be a minimisation problem.
+
+    \begin{align}
+        & \sum_{t \in \mathcal{T}, j \in \mathcal{J}} a_{ipjt} \times h_j = C & (i \in \mathcal{I}, p \in \mathcal{P}) \label{hours}\\
+        & \sum_{i \in \mathcal{I}, p \in \mathcal{P}} a_{ipjt} = 1 & (j \in \mathcal{J}, t \in \mathcal{T}) \label{demand-TPP}\\
+        & F_{ij} \leq \sum_{p \in \mathcal{P}, t \in \mathcal{T}} a_{ipjt} & (i \in \mathcal{I}, j \in \mathcal{J}) \label{auxiliary1}\\
+        & F_{ij} \geq a_{ipjt} & (i \in \mathcal{I}, p \in \mathcal{P}, t \in \mathcal{T}, j \in \mathcal{J}) \label{auxiliary2}\\
+        & \sum_{i \in \mathcal{I}} F_{ij} = 1 &  (j \in \mathcal{J}) \label{auxiliary3}\\
+        & F_ij, a_{ipjt} \geq 0,\, F_{ij}, a_{ipjt} \in \mathbb{B}
+    \end{align}
+
+    (\ref{hours}) constraints force the number of hours each aircraft uses to be exactly the number of available hours.
+    (\ref{demand-TPP}) constraints force the solution to have exactly 1 aircraft assigned to each mission at each period.
+    (\ref{auxiliary1}) and (\ref{auxiliary2}) constraints tie the fleet assignment with the single aircraft assignment.
+    (\ref{auxiliary3}) controls that each mission is only assigned to one fleet.
+
+    \subsection{Possible proof}
+
+    Each fleet is assigned three missions. These missions need to sum the available amount of flight hours (C).
+    Because we need to satisfy the assignment of aircraft each month, each aircraft belonging to the fleet will have a different combination of the three missions (three permutations in a cycle: ABC, CBA, BCA).
+
+    For example if the missions with flight hours (60, 30, 10) are assigned to fleet 1 and the missions with flight hours (50, 40, 10) are assigned to fleet "3M".
+In this case C=100.
+One could have the following solution:
+
+    $$
+    \begin{bmatrix}
+        A_{11} & 60 & 30 & 10 \\
+        A_{12} & 10 & 60 & 30 \\
+        A_{13} & 30 & 10 & 60 \\\hline
+        \vdots & \vdots & \vdots & \vdots \\\hline
+        A_{n1} & H_i & H_j & H_k \\
+        A_{n2} & H_k & H_i & H_j \\
+        A_{n3} & H_j & H_k & H_i \\\hline
+        \vdots & \vdots & \vdots & \vdots \\\hline
+        A_{3M1} & 50 & 40 & 10  \\
+        A_{3M2} & 10 & 50 & 40  \\
+        A_{3M3} & 40 & 10 & 50  \\\hline
+    \end{bmatrix}
+    $$
+
+    We need to comply with having at least one aircraft assigned to each mission per month. Since we have 3M aircraft, we can do this exactly by assigning all slots.
+
+    The solution to this problem should give a solution to the Three Partition Problem.
+It would just suffice to take the three missions assigned to each fleet (the order is irrelevant) and use this as a solution of M groups (corresponding to M fleets).
+In other words, the contents of the variable $F_{ij}$.
+
+    It's true that in the original problem we do not have the concept of fleet... but maybe we can reformulate the definitions to make it even easier.
+
+    I think we could get the fleet part out if we would guarantee that each aircraft has an "alldifferent" assignment of three missions.
+This would guarantee that the solution of the problem would still be equivalent to the fleet one.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % \bibliographystyle{dcu}
 \bibliographystyle{alpha}
-\bibliography{./../biblio/Scheduling,./../biblio/NRP}
+\bibliography{./../../biblio/Scheduling,./../../biblio/NRP}
 

### 2018-07-12 latex/notes/complexite/FMP_variants.tex
More detail to complexite demonstration based on crew scheduling.

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Problem presentation}
+
+    The problem consists in assigning military aircraft to a given set of missions while scheduling the maintenance operations over a time horizon.
+
+    In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks".
+A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
+Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
+Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.
+Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
+Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
+The number of resources in use for each period $d_t$ is also known.
+
+    Each resource can only be assigned to a single task in any given period.
+These resources suffer from wear and tear and require regular maintenance operations during their lifetime.
+The need for maintenance is calculated based on two indicators.
+
+    The first one is called "remaining elapsed time" (or $ret_{it}$).
+It expresses the amount of time (measured in time periods) after which the resource cannot be used any more and has to undergo a maintenance operation.
+Its value is calculated for each resource $i$ and each time period $t$.
+In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the resource $i$ can be used before needing a maintenance operation at any given period $t$.
+
+    Each maintenance operation has a fix duration of $M$ periods and cannot be interrupted: during this time the resource cannot be assigned to any task.
+After a maintenance operation, a resource restores its remaining elapsed time and remaining usage time to their maximum values $E$ and $H$ respectively.
+
+    Some resources can be in maintenance at the beginning of the planning horizon, $n_t$ is used for the number of aircraft in such planned maintenance per period $t$ and defined only for the first $m-1$ time periods.
+
+    Given that the creation of new tasks and the duration of maintenance are considered stochastic in real-life, one basic goal is to maximize the robustness of the planning by having the greatest amount of available resources at every period of the planning horizon.
+
+    Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of resources under maintenance over the planning horizon.
+
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Crew scheduling variant}
@@
     \begin{align}
-        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i \label{numPatterns}
+        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i \label{numPatterns2}
     \end{align}
@@
     \begin{align}
-        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq 1 & t \in \mathcal{T} \label{demand} \\
+        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq 1 & t \in \mathcal{T} \label{demand2} \\
         & x_{i} \in \{0, 1\} \\
@@
 
-    (\ref{numPatterns}) minimizes the number of workers. (\ref{demand}) defines the need to have at least one worker working at any given time.
+    (\ref{numPatterns2}) minimizes the number of workers. (\ref{demand2}) defines the need to have at least one worker working at any given time.
 
@@
 
+\section{3 Dimensional matching}
+
+Here I want to prove that... I need three dimensions. Resource-months, Mission-months and cycles.
+
+I want to decide if a task-month is assigned to a resource-month inside the cycle.
+
+Possibilities are constrained by saying.
+Every misision-month needs to be covered.
+Every resource-month needs to be covered.
+
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{other notes}
+
+If I take the shifts models that were polynomial...
+given that they become hard when adding limits on the number of shifts per planning horizon and if I have flight hours that vary between "shifts"; this means that I can traduce the flight hours in this...
+maybe?
+see: 2.4.3 from \cite{Smet2015}.
+
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Crew scheduling variant 2}
+
+    Based on an existing problem formulated in \cite{Smet2015} and proved in \cite{Brunner2013}.
+
+    The objective of the problem discussed by \cite{Brunner2013} is to minimise the size of the workforce.
+Apart from coverage requirements, expressed as a minimum ($b_{t}$), there are three other constraints.
+The first constraint limits the number of days worked to a maximum value ($\overline{D}^{work}$).
+The second and third constraints define a range on the number of consecutive days worked ($\underline{D}^{on} - \overline{D}^{on}$) and days-off ($\underline{D}^{off} - \overline{D}^{off}$), respectively.
+
+    All missions can use any aircraft (homogeneous fleet). 
+    Even if we have a fixed (not a range) number of consecutive non-working days, the original proof was made with both ranges being fixed.
+
+    Now, a simplified description of the original problem is done that matches this problem:
+
+    For each period $t$ with demand $b_{t}$, we create a task $j$ with $T_j = \{t\}$ that requires $r_{j}$ resources.
+    For each worker $i$ we have a corresponding resource $i$.
+    Since the workforce has the same skills, we assume $O_i = \{j \in \mathca{J}\}$.
+
+    A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
+Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
+Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.
+Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
+Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
+The number of resources in use for each period $d_t$ is also known.
+
+    Each resource can only be assigned to a single task in any given period.
+These resources suffer from wear and tear and require regular maintenance operations during their lifetime.
+The need for maintenance is calculated based on two indicators.
+
+    The first one is called "remaining elapsed time" (or $ret_{it}$).
+It expresses the amount of time (measured in time periods) after which the resource cannot be used any more and has to undergo a maintenance operation.
+Its value is calculated for each resource $i$ and each time period $t$.
+In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the resource $i$ can be used before needing a maintenance operation at any given period $t$.
+
+    Each maintenance operation has a fix duration of $M$ periods and cannot be interrupted: during this time the resource cannot be assigned to any task.
+After a maintenance operation, a resource restores its remaining elapsed time and remaining usage time to their maximum values $E$ and $H$ respectively.
+
+    Some resources can be in maintenance at the beginning of the planning horizon, $n_t$ is used for the number of aircraft in such planned maintenance per period $t$ and defined only for the first $m-1$ time periods.
+
+    Given that the creation of new tasks and the duration of maintenance are considered stochastic in real-life, one basic goal is to maximize the robustness of the planning by having the greatest amount of available resources at every period of the planning horizon.
+
+    Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of resources under maintenance over the planning horizon.
+
+    The notation follows closely the \cite{Smet2015} 2.6.1 example.
+
+    \subsection{Hypothesis}
+
+    \begin{itemize}
+         \item objective: minimize the number of aircraft to maximize availability.
+    \end{itemize}
+
+    \subsection{Reductions}
+
+    \begin{itemize}
+        \item The missions consume 0 hours of aircraft.
+        \item The set of $j \in \mathcal{J}$ tasks
+        \item No minimum duration for mission assignments.
+    \end{itemize}
+
+    \subsection{Notations}
+
+    \textbf{Sets}
+
+    \begin{tabular}{ll}
+        $t \in \mathcal{T}$ & Planning horizon periods. \\
+        $i \in \mathcal{I}$ & Resources. \\
+    \end{tabular}
+
+    \vskip 0.3cm
+
+    \subsection{Parameters}
+
+    \begin{tabular}{ll}
+        $\overline{D}^{work}$ & Maximum number of total working days in period. \\
+        $D^{off}$ & Number of consecutive non-working days. \\
+        $\overline{D}^{on}$ & Maximum number of consecutive working days. \\
+        $\underline{D}^{on}$ & Minimum number of consecutive working days. \\
+        $b_{t}$ & Need of resources at period $t$. \\
+    \end{tabular}
+
+    \vskip 0.3cm
+
+    \subsection{Variables}
+
+    \begin{itemize}
+        \item $x_{it}$: resource $i$ is working on period $t$.
+        \item $y_{i}$: resource $i$ is working on at least one period.
+    \end{itemize}
+
+    \subsection{Model}
+
+    \begin{align}
+        & \text{Min}\; \sum_{i \in \mathcal{I}} y_{i} \label{min_num_resources}
+    \end{align}
+
+    \begin{align}
+        & \sum_{i \in \mathcal{I}} x_{it} \geq b_t & t \in \mathcal{T} \label{demand3} \\
+        & \sum_{t \in \mathcal{T}} x_{it} \leq \overline{D}^{work} & i \in \mathcal{I} \label{max_work} \\
+        & \sum_{t' = t}^{\overline{D}^{on}+t-1} x_{it'} \leq \overline{D}^{on} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_work_days} \\
+        & \sum_{t' = t}^{\underline{D}^{on}+t-1} x_{it'} \geq \underline{D}^{on} (x_{it} - x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_work_days} \\
+        & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \leq D^{off} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_no_work_days} \\
+        & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \geq D^{off} (- x_{it} + x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_no_work_days} \\
+        & y_{i} \geq x_{it} & i \in \mathcal{I}, t \in \mathcal{T} \label{num_resources} \\
+        & x_{it}, y_{i} \in \{0, 1\} & i \in \mathcal{I}, t \in \mathcal{T} \label{binary}
+    \end{align}
+
+    Constraints (\ref{demand3}) guarantee the necessary number of workers is available to work. 
+    Constraints ((\ref{max_work}) checks that the total number of working days for each worker is less than the upper limit.
+    Constraints (\ref{max_work_days} - \ref{min_work_days}) control the resources meet the range of working days.
+Constraints (\ref{max_no_work_days} - \ref{min_no_work_days}) control the resources meet the fixed number of non-working days.
+    Finally constraints (\ref{num_resources}) quantify the number of necessary workers in the planning to be minimized in (\ref{min_num_resources}).
+
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % \bibliographystyle{dcu}

### 2018-07-13 latex/notes/complexite/FMP_variants.tex

@@
 
-    The objective of the problem discussed by \cite{Brunner2013} is to minimise the size of the workforce.
-Apart from coverage requirements, expressed as a minimum ($b_{t}$), there are three other constraints.
+    \subsection{Base problem}
+
+    The problem is described in \cite{Smet2015}, almost verbatim, as the following:
+
+    The objective of the problem discussed by \cite{Brunner2013} is to minimize the size of the workforce.
+Apart from coverage requirements, expressed as a minimum ($b_{k}$), there are three other constraints.
 The first constraint limits the number of days worked to a maximum value ($\overline{D}^{work}$).
@@
 
-    All missions can use any aircraft (homogeneous fleet). 
-    Even if we have a fixed (not a range) number of consecutive non-working days, the original proof was made with both ranges being fixed.
+    \subsection{Reduction}
 
-    Now, a simplified description of the original problem is done that matches this problem:
+    Now, a simplified description of our original problem is done so that it complies with the above one.
+For reference, see the original description of the problem at the beginning of this document.
+It's an adaption of the MOSIM description.
 
-    For each period $t$ with demand $b_{t}$, we create a task $j$ with $T_j = \{t\}$ that requires $r_{j}$ resources.
+    For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
+    For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $r_{j} = b_{t}| t \in T_j$ resources. This implies that $|J_t| = 1$ and $d_t = b_{t}$
     For each worker $i$ we have a corresponding resource $i$.
-    Since the workforce has the same skills, we assume $O_i = \{j \in \mathca{J}\}$.
+    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$ and $a_{j} = \{i \in \mathcal{I}\}$.
+    All resources start the planning horizon have just exited a maintenance in the previous period. This implies $n_t = 0$.
+    The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$.
+This is assumption is valid because the original proof was done with fixed values instead of ranges.
+    The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
+    The usage time for each resource by any task will be $h_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
+The minimum number of maintenances for each resource will be given by $\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$.
 
-    A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
-Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
-Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.
-Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
-Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
-The number of resources in use for each period $d_t$ is also known.
-
-    Each resource can only be assigned to a single task in any given period.
-These resources suffer from wear and tear and require regular maintenance operations during their lifetime.
-The need for maintenance is calculated based on two indicators.
-
-    The first one is called "remaining elapsed time" (or $ret_{it}$).
-It expresses the amount of time (measured in time periods) after which the resource cannot be used any more and has to undergo a maintenance operation.
-Its value is calculated for each resource $i$ and each time period $t$.
-In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the resource $i$ can be used before needing a maintenance operation at any given period $t$.
-
-    Each maintenance operation has a fix duration of $M$ periods and cannot be interrupted: during this time the resource cannot be assigned to any task.
-After a maintenance operation, a resource restores its remaining elapsed time and remaining usage time to their maximum values $E$ and $H$ respectively.
-
-    Some resources can be in maintenance at the beginning of the planning horizon, $n_t$ is used for the number of aircraft in such planned maintenance per period $t$ and defined only for the first $m-1$ time periods.
-
-    Given that the creation of new tasks and the duration of maintenance are considered stochastic in real-life, one basic goal is to maximize the robustness of the planning by having the greatest amount of available resources at every period of the planning horizon.
-
-    Given the limited amount of maintenance capacity and its cost, another goal is to smooth as much as possible the number of resources under maintenance over the planning horizon.
-
-    The notation follows closely the \cite{Smet2015} 2.6.1 example.
+    The model notation follows closely the \cite{Smet2015} 2.6.1 example.
 
@@
          \item objective: minimize the number of aircraft to maximize availability.
-    \end{itemize}
-
-    \subsection{Reductions}
-
-    \begin{itemize}
-        \item The missions consume 0 hours of aircraft.
-        \item The set of $j \in \mathcal{J}$ tasks
-        \item No minimum duration for mission assignments.
     \end{itemize}
@@
         $i \in \mathcal{I}$ & Resources. \\
+        $j \in \mathcal{J}$ & tasks. 
     \end{tabular}
-
-    \vskip 0.3cm
 
@@
     \begin{tabular}{ll}
-        $\overline{D}^{work}$ & Maximum number of total working days in period. \\
-        $D^{off}$ & Number of consecutive non-working days. \\
-        $\overline{D}^{on}$ & Maximum number of consecutive working days. \\
-        $\underline{D}^{on}$ & Minimum number of consecutive working days. \\
-        $b_{t}$ & Need of resources at period $t$. \\
+        $m$ & Minimum number of maintenances per resource in planning horizon. \\
+        $M$ & Duration of maintenance. \\
+        $E$ & Maximum elapsed time before assigning a maintenance. \\
+        $e$ & Minimum elapsed time before assigning a maintenance. \\
+        $r_{j}$ & Need of resources for task $j$. \\
     \end{tabular}
+
+    (here we could, in theory just copy the same model than we had in MOSIM).
 
@@
 
-    \subsection{Variables}
+    % \subsection{Variables}
 
-    \begin{itemize}
-        \item $x_{it}$: resource $i$ is working on period $t$.
-        \item $y_{i}$: resource $i$ is working on at least one period.
-    \end{itemize}
+    % \begin{itemize}
+    %     \item $x_{it}$: resource $i$ is working on period $t$.
+    %     \item $y_{i}$: resource $i$ is working on at least one period.
+    % \end{itemize}
 
-    \subsection{Model}
+    % \subsection{Model}
 
-    \begin{align}
-        & \text{Min}\; \sum_{i \in \mathcal{I}} y_{i} \label{min_num_resources}
-    \end{align}
+    % \begin{align}
+    %     & \text{Min}\; \sum_{i \in \mathcal{I}} y_{i} \label{min_num_resources}
+    % \end{align}
 
-    \begin{align}
-        & \sum_{i \in \mathcal{I}} x_{it} \geq b_t & t \in \mathcal{T} \label{demand3} \\
-        & \sum_{t \in \mathcal{T}} x_{it} \leq \overline{D}^{work} & i \in \mathcal{I} \label{max_work} \\
-        & \sum_{t' = t}^{\overline{D}^{on}+t-1} x_{it'} \leq \overline{D}^{on} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_work_days} \\
-        & \sum_{t' = t}^{\underline{D}^{on}+t-1} x_{it'} \geq \underline{D}^{on} (x_{it} - x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_work_days} \\
-        & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \leq D^{off} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_no_work_days} \\
-        & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \geq D^{off} (- x_{it} + x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_no_work_days} \\
-        & y_{i} \geq x_{it} & i \in \mathcal{I}, t \in \mathcal{T} \label{num_resources} \\
-        & x_{it}, y_{i} \in \{0, 1\} & i \in \mathcal{I}, t \in \mathcal{T} \label{binary}
-    \end{align}
+    % \begin{align}
+    %     & \sum_{i \in \mathcal{I}} x_{it} \geq b_t & t \in \mathcal{T} \label{demand3} \\
+    %     & \sum_{t \in \mathcal{T}} x_{it} \leq \overline{D}^{work} & i \in \mathcal{I} \label{max_work} \\
+    %     & \sum_{t' = t}^{\overline{D}^{on}+t-1} x_{it'} \leq \overline{D}^{on} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_work_days} \\
+    %     & \sum_{t' = t}^{\underline{D}^{on}+t-1} x_{it'} \geq \underline{D}^{on} (x_{it} - x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_work_days} \\
+    %     & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \leq D^{off} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_no_work_days} \\
+    %     & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \geq D^{off} (- x_{it} + x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_no_work_days} \\
+    %     & y_{i} \geq x_{it} & i \in \mathcal{I}, t \in \mathcal{T} \label{num_resources} \\
+    %     & x_{it}, y_{i} \in \{0, 1\} & i \in \mathcal{I}, t \in \mathcal{T} \label{binary}
+    % \end{align}
 
-    Constraints (\ref{demand3}) guarantee the necessary number of workers is available to work. 
-    Constraints ((\ref{max_work}) checks that the total number of working days for each worker is less than the upper limit.
-    Constraints (\ref{max_work_days} - \ref{min_work_days}) control the resources meet the range of working days.
+    % Constraints (\ref{demand3}) guarantee the necessary number of workers is available to work. 
+    % Constraints ((\ref{max_work}) checks that the total number of working days for each worker is less than the upper limit.
+    % Constraints (\ref{max_work_days} - \ref{min_work_days}) control the resources meet the range of working days.
 Constraints (\ref{max_no_work_days} - \ref{min_no_work_days}) control the resources meet the fixed number of non-working days.
-    Finally constraints (\ref{num_resources}) quantify the number of necessary workers in the planning to be minimized in (\ref{min_num_resources}).
+    % Finally constraints (\ref{num_resources}) quantify the number of necessary workers in the planning to be minimized in (\ref{min_num_resources}).
 

### 2018-07-13 latex/notes/complexite/FMP_variants.tex
last changes to complexity analysis

@@
 Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.
-Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
-Set $a_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
-The number of resources in use for each period $d_t$ is also known.
+Each task requires a certain number $R_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $H_j$ in each period.
+Set $I_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
+The number of resources in use for each period $D_t$ is also known.
 
@@
     For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
-    For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $r_{j} = b_{t}| t \in T_j$ resources. This implies that $|J_t| = 1$ and $d_t = b_{t}$
+    For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $R_{j} = b_{t}| t \in T_j$ resources.
+This implies that $J_t = \{j\}| t \in T_j$, $|J_t|=1$ and $D_t = b_{t}$.
+
     For each worker $i$ we have a corresponding resource $i$.
-    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$ and $a_{j} = \{i \in \mathcal{I}\}$.
-    All resources start the planning horizon have just exited a maintenance in the previous period. This implies $n_t = 0$.
-    The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$.
-This is assumption is valid because the original proof was done with fixed values instead of ranges.
+    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$.
+    All resources start the planning horizon have just exited a maintenance in the previous period. This implies $N_t = 0$.
+    The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$. This assumption is valid because the original proof was done with fixed values instead of ranges.
+
     The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
-    The usage time for each resource by any task will be $h_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
-The minimum number of maintenances for each resource will be given by $\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$.
+    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
+The minimum number of maintenances for each resource will be given by $M_{min}=\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$.
+(Alternatively, the limit could be measured as a minimum number of maintenance months $M_{min}=|\mathcal{T}| - \overline{D}^{work}$).
 
-    The model notation follows closely the \cite{Smet2015} 2.6.1 example.
+    % \subsection{Hypothesis}
 
-    \subsection{Hypothesis}
-
-    \begin{itemize}
-         \item objective: minimize the number of aircraft to maximize availability.
-    \end{itemize}
+    % \begin{itemize}
+    %      \item objective: minimize the number of aircraft to maximize availability.
+    % \end{itemize}
 
@@
     \begin{tabular}{ll}
-        $t \in \mathcal{T}$ & Planning horizon periods. \\
-        $i \in \mathcal{I}$ & Resources. \\
-        $j \in \mathcal{J}$ & tasks. 
+        $\mathcal{T}$ & Planning horizon periods. \\
+        $\mathcal{I}$ & Resources. \\
+        $\mathcal{J}$ & tasks.  \\
+        $\mathcal{T}_j$ &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
+        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
+        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
+        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
+        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
     \end{tabular}
@@
     \begin{tabular}{ll}
-        $m$ & Minimum number of maintenances per resource in planning horizon. \\
+        $M_{min}$ & Minimum number of maintenances per resource in planning horizon. \\
         $M$ & Duration of maintenance. \\
         $E$ & Maximum elapsed time before assigning a maintenance. \\
-        $e$ & Minimum elapsed time before assigning a maintenance. \\
-        $r_{j}$ & Need of resources for task $j$. \\
+        $E_{min}$ & Minimum elapsed time before assigning a maintenance. \\
+        $R_{j}$ & Need of resources for task $j$. \\
     \end{tabular}
 
-    (here we could, in theory just copy the same model than we had in MOSIM).
+    % (here we could, in theory just copy the same model than we had in MOSIM).
 
@@
 
-    % \subsection{Variables}
+     \subsection{Variables}
 
-    % \begin{itemize}
-    %     \item $x_{it}$: resource $i$ is working on period $t$.
-    %     \item $y_{i}$: resource $i$ is working on at least one period.
-    % \end{itemize}
+        \begin{tabular}{ll}
+            $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
+            $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
+        \end{tabular}
 
-    % \subsection{Model}
+    \subsection{Model}
 
+        The following modifications need to be done to the original problem described in MOSIM
+
+        \begin{align}
+            & ret_{it} \in [E_{min},E] 
+                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
+            & \sum_{t \in \mathcal{T}} m_{it} \geq M_{min} 
+                & i \in \mathcal{I} \label{eq:min_maintenace}
+        \end{align}
+
+        Constraints (\ref{eq:me}), which already existed, now have a specific minimum value instead of just 0.
+Constraints (\ref{eq:min_maintenace}) are new.
+They put a minimum number of maintenances for each resource.
+This had been considered as a possible cut to the original problem and does not imply a big change to the solutions.
+
+        The alternative to constraints (\ref{eq:min_maintenace}), in case we want to control the months under maintenance and not the number of maintenance operations, are constraints (\ref{eq:min_maintenace2}).
+
+        \begin{align}
+            & \sum_{t \in \mathcal{T}} \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} \geq M_{min}
+                & i \in \mathcal{I} \label{eq:min_maintenace2}\\
+        \end{align}
+
+        % \begin{align}
+        %     % maximum capacity1:
+        %     & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
+        %       & t \in \mathcal{T} \label{eq:capacity1}\\
+        %    %avail
+        %    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
+        %     &t \in \mathcal{T} \label{eq:avalaibility1}\\
+        %    % maximum capacity2:                
+        %     & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
+        %             & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
+        %     & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
+        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
+        % \end{align}
+
+        % \begin{align}
+        %     % remaining used time
+        %      & rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j \notag \\
+        %             && t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
+        %     & rut_{i0} = Rut^{Init}_i
+        %            & i \in \mathcal{I} \label{eq:rut_initial}\\
+        %     & rut_{it} \geq H m_{it}
+        %             & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
+        %     & rut_{it} \in [0,H]
+        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\
+        %     & ret_{it} \leq ret_{it-1} - 1 + E m_{it}
+        %             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
+        %     & ret_{i0} = Ret^{Init}_i
+        %             & i \in \mathcal{I} \label{eq:ret_initial}\\
+        %     & ret_{it} \geq E m_{it}
+        %             & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
+        %     & ret_{it} \in [0,E]
+        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
+        %     & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
+        %           & t = |\mathcal{T}| \label{eq:min_ret}\\
+        %     & \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
+        %           & t = |\mathcal{T}| \label{eq:min_rut}
+        % \end{align}
     % \begin{align}
@@
 
+\section{Fixed interval scheduling}
+
+    From \cite{Smet2015}, \cite{Kolen2007} and \cite{Kovalyov2007}.
+
+    \subsection{Base problem}
+
+    \cite{Smet2015} calls it "shift minimization personnel task scheduling problem. When it's not related to workers and in a more general way, it's usually called "Fixed Interval Scheduling Problem".
+
+    Here is the description of \cite{Smet2015}:
+
+    Let $J = {1, ..., n}$ be the set of tasks to be assigned and $E = {1, ..., m}$ the set of employees.
+Each task $j \in J$ has a duration $u_j$ , a start time $s_j$ and a finish time $f_j$ = $s_j + u_j$ .
+Each employee e has a set of tasks $J_e \in J$ that he/she can perform.
+Similarly, for each task $j$, a set $E_j \in E$ exists, which contains all employees that can perform task $j$.
+Both $T_e$ and $E_j$ are defined based on qualifications, time windows of tasks and availabilities of employees.
+
+    \subsection{Reduction}
+
+    Now, a simplified description of our original problem is done so that it complies with the above one.
+For reference, see the original description of the problem at the beginning of this document.
+It's an adaption of the MOSIM description.
+
+    For each employee $e \in E$, we create an analogous resource $i \in I$ which we will use indistinctly.
+
+    For each task $j \in J$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$.
+Start times and finish times define the moment when the task is active: $\mathcal{T}_j = t \in \{s_j ...
+f_j\}$.
+The need of resources for each task will be constant $R_{j} = 1$.
+The number of resources working at any given time can $D_t = |\mathcal{J}_t|$ can be calculated.
+The relationship between tasks and employees is also equivalent: $\mathcal{O}_i = J_i$ and $\mathcal{I}_j = E_j$.
+
+    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
+    For maintenances, we can just take them out of the equation completely. For this we just need to put: $E_{min} = E = M = N_t = 0$ and $Ret^{Init}_{i} = |\mathcal{T}| + 1$.
+
+    \subsection{Notations}
+
+    \textbf{Sets}
+
+    \begin{tabular}{ll}
+        $\mathcal{T}$       & Planning horizon periods. \\
+        $\mathcal{I}$       & Resources. \\
+        $\mathcal{J}$       & tasks.  \\
+        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
+        $\mathcal{J}_t$     &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
+        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
+        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
+        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
+    \end{tabular}
+
+    \subsection{Parameters}
+
+    \begin{tabular}{ll}
+        $M_{min}$        & Minimum number of maintenances per resource in planning horizon. \\
+        $M$              & Duration of maintenance. \\
+        $E$              & Maximum elapsed time before assigning a maintenance. \\
+        $E_{min}$        & Minimum elapsed time before assigning a maintenance. \\
+        $R_{j}$          & Need of resources for task $j$. \\
+        $H$              & remaining usage time after a maintenance. \\
+        $D_t$            & number of resources to be assigned in total in period $t$. \\
+        $A^min_j$        & minimum number of consecutive periods to assign any resource to task $j$. \\
+        $Ret^{Init}_{i}$ & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
+    \end{tabular}
+
+     \subsection{Variables}
+
+        \begin{tabular}{ll}
+            $at^{init}_{it}$    &  =1 if resource $i \in I$ starts a mission assignment in period $t \in \mathcal{T}$, 0 otherwise. \\
+        \end{tabular}
+
+    \subsection{Model}
+
+    The model would need to have constraints of minimum assignment.
+For this, we would have to implement a new variable for missions that would be similar to the one in maintenances (a 'start' of mission assignment) and then tie this new variable to the regular assignments.
+
+        \begin{align}
+            &  at^{init}_{it} \geq at_{it}
+                & i \in \mathcal{I} \label{eq:min_maintenace2}\\
+        \end{align}
+
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % \bibliographystyle{dcu}

### 2018-07-13 latex/notes/complexite/FMP_variants.tex
updated notes on complexity

@@
 
-    Let $J = {1, ..., n}$ be the set of tasks to be assigned and $E = {1, ..., m}$ the set of employees.
-Each task $j \in J$ has a duration $u_j$ , a start time $s_j$ and a finish time $f_j$ = $s_j + u_j$ .
-Each employee e has a set of tasks $J_e \in J$ that he/she can perform.
-Similarly, for each task $j$, a set $E_j \in E$ exists, which contains all employees that can perform task $j$.
-Both $T_e$ and $E_j$ are defined based on qualifications, time windows of tasks and availabilities of employees.
+    Let $P = {1, ..., n}$ be the set of tasks to be assigned and $E = {1, ..., m}$ the set of employees.
+Each task $p \in P$ has a duration $u_p$ , a start time $s_p$ and a finish time $f_p$ = $s_p + u_p$ .
+Each employee $e$ has a set of tasks $P_e \in P$ that he/she can perform.
+Similarly, for each task $p$, a set $E_p \in E$ exists, which contains all employees that can perform task $p$.
+Both $T_e$ and $E_p$ are defined based on qualifications, time windows of tasks and availabilities of employees.
+
+    The reference problem has the following characteristics.
+
+    $\mathcal{P}_{ref}$:
+
+    \begin{tabular}{ll}
+        $\mathcal{E}$       & Employees. \\
+        $\mathcal{P}$       & tasks.  \\
+        $u_p$     			&  Duration of task $p$. \\
+        $s_p$     			&  Starting times $p$. \\
+        $f_p$     			&  finish time of task $p$. \\
+        $S_e$     &  Set of tasks an employee $e$ can perform. \\
+        $E_p$ &  Set of employees that can perform task $p$. \\
+    \end{tabular}
 
@@
 
-    For each employee $e \in E$, we create an analogous resource $i \in I$ which we will use indistinctly.
+    % For each employee $e \in E$, we create an analogous resource $i \in I$ which we will use indistinctly.
 
-    For each task $j \in J$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$.
+    % For each task $j \in J$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$.
 Start times and finish times define the moment when the task is active: $\mathcal{T}_j = t \in \{s_j ...
@@
 
-    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
-    For maintenances, we can just take them out of the equation completely. For this we just need to put: $E_{min} = E = M = N_t = 0$ and $Ret^{Init}_{i} = |\mathcal{T}| + 1$.
+    % The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
+    % For maintenances, we can just take them out of the equation completely. For this we just need to put: $E_{min} = E = M = N_t = 0$ and $Ret^{Init}_{i} = |\mathcal{T}| + 1$.
 
-    \subsection{Notations}
+    % An instance of our problem is the following:
+
+    % \subsection{Notations}
+
+    The FMP problem has the following characteristics.
+
+    $\mathcal{P}_{FMP}$:
+
+    \vskip 0.3cm
 
@@
 
-    \begin{tabular}{ll}
-        $\mathcal{T}$       & Planning horizon periods. \\
-        $\mathcal{I}$       & Resources. \\
-        $\mathcal{J}$       & tasks.  \\
-        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
-        $\mathcal{J}_t$     &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
-        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
-        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
+    \begin{tabular}{l|l|l}
+    	\textbf{FMP}    &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
+        $\mathcal{T}$    &  Planning horizon. & $\{\min_p({s_p}) ... \max_p({f_p})\}$ \\
+        $\mathcal{I}$    &  Resources. & $ \mathcal{E}$ \\
+        $\mathcal{J}$    &  Tasks.  & $ P$ \\
+        $\mathcal{T}_j$  &  time periods in which task $j$ is active. & $t \in \{s_p ... f_p\}$ \\
+        $\mathcal{J}_t$  &  tasks to be realized in period $t$. & $j | t \in \{s_p ... f_p\}$ \\
+        $\mathcal{I}_j$  &  resources that can be assigned to task $j$. & $E_p$\\
+        $\mathcal{O}_i$  &  tasks for which resource $i$ can be used. & $P_i$ \\
+        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ that control de maintenance starts. & $\{\}$ \\
     \end{tabular}
 
-    \subsection{Parameters}
+    \vskip 0.3cm
 
-    \begin{tabular}{ll}
-        $M_{min}$        & Minimum number of maintenances per resource in planning horizon. \\
-        $M$              & Duration of maintenance. \\
-        $E$              & Maximum elapsed time before assigning a maintenance. \\
-        $E_{min}$        & Minimum elapsed time before assigning a maintenance. \\
-        $R_{j}$          & Need of resources for task $j$. \\
-        $H$              & remaining usage time after a maintenance. \\
-        $D_t$            & number of resources to be assigned in total in period $t$. \\
-        $A^min_j$        & minimum number of consecutive periods to assign any resource to task $j$. \\
-        $Ret^{Init}_{i}$ & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
+    \textbf{Parameters}
+
+    \begin{tabular}{l|l|l}
+    	\textbf{FMP}    &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
+        $D_t$            & number of resources to be assigned in total in period $t$ & $|\mathcal{J}_t|$\\
+        $A^{min}_j$      & minimum number of consecutive periods for task $j$. & $u_p$ \\
+        $Ret^{Init}_{i}$ & remaining elapsed time for resource $i$ at the start. & $|\mathcal{T}| + 1$ \\
+        $R_{j}$          & Need of resources for task $j$ & $1$ \\
+        $H_j$             & amount of resource time required by task $j$. & $ 0$ \\
+        $M$               & maintenance duration in number of periods. & $ 0$ \\
+        $E$               & remaining elapsed time after a maintenance. & $ 0$ \\
+        $H$               & remaining usage time after a maintenance.& $ 0$ \\
+        $N_t$             & number of resources in already-planned maintenances in period $t$ at the start. & $0$ \\
+        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. & $ 0$ \\
+        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. & $ 0$\\
+        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. & $ 0$ \\
     \end{tabular}
-
-     \subsection{Variables}
-
-        \begin{tabular}{ll}
-            $at^{init}_{it}$    &  =1 if resource $i \in I$ starts a mission assignment in period $t \in \mathcal{T}$, 0 otherwise. \\
-        \end{tabular}
-
-    \subsection{Model}
-
-    The model would need to have constraints of minimum assignment.
-For this, we would have to implement a new variable for missions that would be similar to the one in maintenances (a 'start' of mission assignment) and then tie this new variable to the regular assignments.
-
-        \begin{align}
-            &  at^{init}_{it} \geq at_{it}
-                & i \in \mathcal{I} \label{eq:min_maintenace2}\\
-        \end{align}
 

### 2018-07-16 latex/notes/complexite/FMP_variants.tex
modified the demonstration of the crew scheduling problem and un-commented the explanation of the two demonstration. It's still missing some parameters configuration and the objective function explanation for both.

@@
 
+    $\mathcal{P}_{ref}$:
+
+    \begin{tabular}{ll}
+        $\mathcal{E}$          & Employees. \\
+        $\mathcal{K}$          & Planning horizon. \\
+        $\overline{D}^{work}$  & Maximum number of working days at the of the planning horizon. \\
+        $\underline{D}^{on}$   & Minimum number of consecutive working days for each employee. \\
+        $\overline{D}^{on}$    & Maximum number of consecutive working days for each employee. \\
+        $\underline{D}^{off}$  & Minimum number of consecutive non working days for each employee. \\
+        $\overline{D}^{off}$   & Maximum number of consecutive non working days for each employee. \\
+        $b_k$                  & Minimum number of workers to be assigned in period $k$. \\
+    \end{tabular}
+
     \subsection{Reduction}
@@
 
-    For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
-    For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $R_{j} = b_{t}| t \in T_j$ resources.
+        \begin{tabular}{l|l|l}
+        \textbf{FMP}        &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
+        $\mathcal{T}$       &  Planning horizon. & $ \mathcal{K} $ \\
+        $\mathcal{I}$       &  Resources. & $ \mathcal{E} $ \\
+        $\mathcal{J}$       &  Tasks.  & $ \{1 .. |\mathcal{K}|\} $ \\
+        $\mathcal{T}_j$     &  time periods in which task $j$ is active. & $ t | t = j $ \\
+        $\mathcal{J}_t$     &  tasks to be realized in period $t$. & $ j | j = t $ \\
+        $\mathcal{I}_j$     &  resources that can be assigned to task $j$. & $ \mathcal{I} $\\
+        $\mathcal{O}_i$     &  tasks for which resource $i$ can be used. & $ \mathcal{J} $ \\
+        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ that control de maintenance starts. & $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ \\
+    \end{tabular}
+
+    \vskip 0.3cm
+
+    \textbf{Parameters}
+
+    \begin{tabular}{l|l|l}
+        \textbf{FMP}      &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
+        $D_t$             & number of resources to be assigned in total in period $t$ & $ b_k $ \\
+        $A^{min}_j$       & minimum number of consecutive periods for task $j$. & $ 1 $ \\
+        $R_{j}$           & Need of resources for task $j$ & $ b_k $ \\
+        $H_j$             & amount of resource time required by task $j$. & $ 0 $ \\
+        $M$               & maintenance duration in number of periods. & $ \underline{D}^{off} $ \\
+        $E$               & remaining elapsed time after a maintenance. & $ \overline{D}^{on} $ \\
+        $E^{min}$         & minimum elapsed time before a maintenance. & $ \underline{D}^{on} $ \\
+        $H$               & remaining usage time after a maintenance. & $ 0 $ \\
+        $N_t$             & number of resources in already-planned maintenances in period $t$ at the start. & $ 0 $ \\
+        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start. & $ TODO $ \\
+        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. & $ 0 $ \\
+        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. & $ 0 $\\
+        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. & $ 0 $ \\
+    \end{tabular}
+
+    % For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
+    % For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $R_{j} = b_{t}| t \in T_j$ resources.
 This implies that $J_t = \{j\}| t \in T_j$, $|J_t|=1$ and $D_t = b_{t}$.
 
-    For each worker $i$ we have a corresponding resource $i$.
-    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$.
-    All resources start the planning horizon have just exited a maintenance in the previous period. This implies $N_t = 0$.
-    The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$. This assumption is valid because the original proof was done with fixed values instead of ranges.
+    % For each worker $i$ we have a corresponding resource $i$.
+    % Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$.
+    % All resources start the planning horizon have just exited a maintenance in the previous period. This implies $N_t = 0$.
+    % The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$. This assumption is valid because the original proof was done with fixed values instead of ranges.
 
-    The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
-    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
+    % The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
+    % The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
 The minimum number of maintenances for each resource will be given by $M_{min}=\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$.
@@
     % \end{itemize}
-
-    \subsection{Notations}
-
-    \textbf{Sets}
-
-    \begin{tabular}{ll}
-        $\mathcal{T}$ & Planning horizon periods. \\
-        $\mathcal{I}$ & Resources. \\
-        $\mathcal{J}$ & tasks.  \\
-        $\mathcal{T}_j$ &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
-        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
-        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
-        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
-    \end{tabular}
-
-    \subsection{Parameters}
-
-    \begin{tabular}{ll}
-        $M_{min}$ & Minimum number of maintenances per resource in planning horizon. \\
-        $M$ & Duration of maintenance. \\
-        $E$ & Maximum elapsed time before assigning a maintenance. \\
-        $E_{min}$ & Minimum elapsed time before assigning a maintenance. \\
-        $R_{j}$ & Need of resources for task $j$. \\
-    \end{tabular}
-
-    % (here we could, in theory just copy the same model than we had in MOSIM).
-
-    \vskip 0.3cm
-
-     \subsection{Variables}
-
-        \begin{tabular}{ll}
-            $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
-            $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
-        \end{tabular}
-
-    \subsection{Model}
-
-        The following modifications need to be done to the original problem described in MOSIM
-
-        \begin{align}
-            & ret_{it} \in [E_{min},E] 
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
-            & \sum_{t \in \mathcal{T}} m_{it} \geq M_{min} 
-                & i \in \mathcal{I} \label{eq:min_maintenace}
-        \end{align}
-
-        Constraints (\ref{eq:me}), which already existed, now have a specific minimum value instead of just 0.
-Constraints (\ref{eq:min_maintenace}) are new.
-They put a minimum number of maintenances for each resource.
-This had been considered as a possible cut to the original problem and does not imply a big change to the solutions.
-
-        The alternative to constraints (\ref{eq:min_maintenace}), in case we want to control the months under maintenance and not the number of maintenance operations, are constraints (\ref{eq:min_maintenace2}).
-
-        \begin{align}
-            & \sum_{t \in \mathcal{T}} \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} \geq M_{min}
-                & i \in \mathcal{I} \label{eq:min_maintenace2}\\
-        \end{align}
-
-        % \begin{align}
-        %     % maximum capacity1:
-        %     & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
-        %       & t \in \mathcal{T} \label{eq:capacity1}\\
-        %    %avail
-        %    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
-        %     &t \in \mathcal{T} \label{eq:avalaibility1}\\
-        %    % maximum capacity2:                
-        %     & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
-        %             & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
-        %     & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
-        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
-        % \end{align}
-
-        % \begin{align}
-        %     % remaining used time
-        %      & rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j \notag \\
-        %             && t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
-        %     & rut_{i0} = Rut^{Init}_i
-        %            & i \in \mathcal{I} \label{eq:rut_initial}\\
-        %     & rut_{it} \geq H m_{it}
-        %             & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
-        %     & rut_{it} \in [0,H]
-        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\
-        %     & ret_{it} \leq ret_{it-1} - 1 + E m_{it}
-        %             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
-        %     & ret_{i0} = Ret^{Init}_i
-        %             & i \in \mathcal{I} \label{eq:ret_initial}\\
-        %     & ret_{it} \geq E m_{it}
-        %             & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
-        %     & ret_{it} \in [0,E]
-        %             & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
-        %     & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
-        %           & t = |\mathcal{T}| \label{eq:min_ret}\\
-        %     & \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
-        %           & t = |\mathcal{T}| \label{eq:min_rut}
-        % \end{align}
-    % \begin{align}
-    %     & \text{Min}\; \sum_{i \in \mathcal{I}} y_{i} \label{min_num_resources}
-    % \end{align}
-
-    % \begin{align}
-    %     & \sum_{i \in \mathcal{I}} x_{it} \geq b_t & t \in \mathcal{T} \label{demand3} \\
-    %     & \sum_{t \in \mathcal{T}} x_{it} \leq \overline{D}^{work} & i \in \mathcal{I} \label{max_work} \\
-    %     & \sum_{t' = t}^{\overline{D}^{on}+t-1} x_{it'} \leq \overline{D}^{on} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_work_days} \\
-    %     & \sum_{t' = t}^{\underline{D}^{on}+t-1} x_{it'} \geq \underline{D}^{on} (x_{it} - x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_work_days} \\
-    %     & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \leq D^{off} & i \in \mathcal{I}, t \in \mathcal{T} \label{max_no_work_days} \\
-    %     & \sum_{t' = t}^{D^{off}+t-1} (1 - x_{it'}) \geq D^{off} (- x_{it} + x_{it-1}) & i \in \mathcal{I}, t \in \mathcal{T} \label{min_no_work_days} \\
-    %     & y_{i} \geq x_{it} & i \in \mathcal{I}, t \in \mathcal{T} \label{num_resources} \\
-    %     & x_{it}, y_{i} \in \{0, 1\} & i \in \mathcal{I}, t \in \mathcal{T} \label{binary}
-    % \end{align}
-
-    % Constraints (\ref{demand3}) guarantee the necessary number of workers is available to work. 
-    % Constraints ((\ref{max_work}) checks that the total number of working days for each worker is less than the upper limit.
-    % Constraints (\ref{max_work_days} - \ref{min_work_days}) control the resources meet the range of working days.
-Constraints (\ref{max_no_work_days} - \ref{min_no_work_days}) control the resources meet the fixed number of non-working days.
-    % Finally constraints (\ref{num_resources}) quantify the number of necessary workers in the planning to be minimized in (\ref{min_num_resources}).
 
@@
 
-    % For each employee $e \in E$, we create an analogous resource $i \in I$ which we will use indistinctly.
+    For each employee $e \in E$, we create an analogous resource $i \in I$ which we will use indistinctly.
 
-    % For each task $j \in J$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$.
+    For each task $j \in J$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$.
 Start times and finish times define the moment when the task is active: $\mathcal{T}_j = t \in \{s_j ...
@@
 
-    % The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
-    % For maintenances, we can just take them out of the equation completely. For this we just need to put: $E_{min} = E = M = N_t = 0$ and $Ret^{Init}_{i} = |\mathcal{T}| + 1$.
-
-    % An instance of our problem is the following:
-
-    % \subsection{Notations}
+    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
+    For maintenances, we can just take them out of the equation completely. For this we just need to put: $E_{min} = E = M = N_t = 0$ and $Ret^{Init}_{i} = |\mathcal{T}| + 1$.
 

### 2018-07-16 latex/notes/complexite/FMP_variants.tex
small changes to complexity format

--- 2018-07-13 latex/notes/complexite/FMP_variants.tex
@@
 
-    $\mathcal{P}_{ref}$:
+    $i \in \mathcal{P}_{ref}$:
 
@@
 
-    $\mathcal{P}_{FMP}$:
+    $I(i \in \mathcal{P}_{FMP})$:
 
@@
     \begin{tabular}{l|l|l}
-    	\textbf{FMP}    &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
+    	\textbf{FMP}    &  \textbf{Meaning} & \textbf{I(i)} \\ \hline
         $\mathcal{T}$    &  Planning horizon. & $\{\min_p({s_p}) ... \max_p({f_p})\}$ \\
@@
         $\mathcal{J}$    &  Tasks.  & $ P$ \\
-        $\mathcal{T}_j$  &  time periods in which task $j$ is active. & $t \in \{s_p ... f_p\}$ \\
+        $\mathcal{T}_j$  &  time periods in which task $j=p$ is active. & $t \in \{s_p ... f_p\}$ \\
         $\mathcal{J}_t$  &  tasks to be realized in period $t$. & $j | t \in \{s_p ... f_p\}$ \\
-        $\mathcal{I}_j$  &  resources that can be assigned to task $j$. & $E_p$\\
-        $\mathcal{O}_i$  &  tasks for which resource $i$ can be used. & $P_i$ \\
+        $\mathcal{I}_j$  &  resources that can be assigned to task $j=p$. & $E_p$\\
+        $\mathcal{O}_i$  &  tasks for which resource $i=e$ can be used. & $P_e$ \\
         $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ that control de maintenance starts. & $\{\}$ \\
@@
     \begin{tabular}{l|l|l}
-    	\textbf{FMP}    &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
+    	\textbf{FMP}    &  \textbf{Meaning} & \textbf{I(i)} \\ \hline
         $D_t$            & number of resources to be assigned in total in period $t$ & $|\mathcal{J}_t|$\\
-        $A^{min}_j$      & minimum number of consecutive periods for task $j$. & $u_p$ \\
+        $A^{min}_j$      & minimum number of consecutive periods for task $j=p$. & $u_p$ \\
         $Ret^{Init}_{i}$ & remaining elapsed time for resource $i$ at the start. & $|\mathcal{T}| + 1$ \\
@@
         $H$               & remaining usage time after a maintenance.& $ 0$ \\
-        $N_t$             & number of resources in already-planned maintenances in period $t$ at the start. & $0$ \\
+        $N_t$             & number of resources in already-planned maintenances in period $t$. & $0$ \\
         $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. & $ 0$ \\
@@
     \end{tabular}
+
+$Q_{ref}$: For an instance $i \in \mathcal{P}_{ref}$, $\exists$ a solution with the number of resources $\leq k$:
+
+$Q_{FMP}$: For an instance $I(i \in \mathcal{P}_{ref})$, $\exists$ a solution with the number of available resources $\geq |I| - k$:
 

### 2018-07-16 latex/notes/complexite/FMP_variants.tex
New version of complexity analysis.

--- 2018-07-16 latex/notes/complexite/FMP_variants.tex
@@
 
-    The problem is described in \cite{Smet2015}, almost verbatim, as the following:
+    The problem is described in \cite{Smet2015}, almost verbatim, as follows:
 
@@
 
-    $\mathcal{P}_{ref}$:
+    \vskip 0.3cm
+
+    $i \in \mathcal{P}_{ref}$:
 
@@
 For reference, see the original description of the problem at the beginning of this document.
-It's an adaption of the MOSIM description.
+Tables (\ref{tab:sets_crew} - \ref{tab:parameters_crew}) summarize this reduction.
 
-        \begin{tabular}{l|l|l}
-        \textbf{FMP}        &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
-        $\mathcal{T}$       &  Planning horizon. & $ \mathcal{K} $ \\
-        $\mathcal{I}$       &  Resources. & $ \mathcal{E} $ \\
-        $\mathcal{J}$       &  Tasks.  & $ \{1 .. |\mathcal{K}|\} $ \\
-        $\mathcal{T}_j$     &  time periods in which task $j$ is active. & $ t | t = j $ \\
-        $\mathcal{J}_t$     &  tasks to be realized in period $t$. & $ j | j = t $ \\
-        $\mathcal{I}_j$     &  resources that can be assigned to task $j$. & $ \mathcal{I} $\\
-        $\mathcal{O}_i$     &  tasks for which resource $i$ can be used. & $ \mathcal{J} $ \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ that control de maintenance starts. & $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ \\
-    \end{tabular}
+    For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$ and a task $j \in \mathcal{J}$.
+Slightly abusing the notation, we can use the elements in these three sets indistinctly.
+Each task $j$ has $T_j = \{j=t\}$ that requires $R_{j} = b_{j=k}$ resources.
+This implies that $J_t = \{j=t\}$, $|J_t|=1$ and $D_t = b_{t=k}$.
+
+    For each worker $e$ we have a corresponding resource $i$.
+    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$.
+    All resources start have just exited a maintenance in the previous period to the start of the planning horizon. There's no previously planned maintenances, so $N_t = 0$.
+    The duration of the maintenance will be equal to $M = \underline{D}^{off}$.
+
+    The maximum remaining elapsed time between two maintenances is equal to $E = \overline{D}^{on}$. The minimum elapsed time between two maintenances is equal to $E^{min}=\underline{D}^{on}$.
+    The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
+The minimum number of maintenance months for each resource will be given by $MM_{min}=|\mathcal{T}| - \overline{D}^{work}$ and the maximum time a resource can stay in need of a maintenance before receiving it is $U^{max} = \overline{D}^{off} - \underline{D}^{off}$.
 
@@
 
-    \textbf{Parameters}
+    $I(i \in \mathcal{P}_{FMP})$:
 
+    \begin{table}
+        \begin{tabular}{l|l|p{40mm}}
+            \textbf{FMP}        &  \textbf{Meaning} & \textbf{I(i)} \\ \hline
+            $\mathcal{T}$       &  planning horizon. & $ \mathcal{K} $ \\
+            $\mathcal{I}$       &  resources. & $ \mathcal{E} $ \\
+            $\mathcal{J}$       &  tasks.  & $ \mathcal{K} $ \\
+            $\mathcal{T}_j$     &  time periods in which task $j=t$ is active. & $ j $ \\
+            $\mathcal{J}_t$     &  tasks to be realized in period $t=j$. & $ t $ \\
+            $\mathcal{I}_j$     &  resources that can be assigned to task $j$. & $ \mathcal{I} $\\
+            $\mathcal{O}_i$     &  tasks for which resource $i$ can be used. & $ \mathcal{J} $ \\
+            $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ that control the maintenance starts. & $\{ \max{\{1, t - M+1\}}, $ $ ..., {t}\}$ \\
+        \end{tabular}
+         \caption{Crew Scheduling sets translation}
+        \label{tab:sets_crew}
+    \end{table}
+
+    \vskip 0.3cm
+
+    % \textbf{Parameters}
+
+    \begin{table}
         \begin{tabular}{l|l|l}
-        \textbf{FMP}      &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
-        $D_t$             & number of resources to be assigned in total in period $t$ & $ b_k $ \\
+            \textbf{FMP}      &  \textbf{Meaning} & \textbf{I(i)} \\ \hline
+            $D_t$             & number of resources to be assigned in total in period $t=k$ & $ b_t $ \\
             $A^{min}_j$       & minimum number of consecutive periods for task $j$. & $ 1 $ \\
-        $R_{j}$           & Need of resources for task $j$ & $ b_k $ \\
+            $R_j$             & need of resources for task $j=t=k$ & $ b_j $ \\
             $H_j$             & amount of resource time required by task $j$. & $ 0 $ \\
@@
             $H$               & remaining usage time after a maintenance. & $ 0 $ \\
-        $N_t$             & number of resources in already-planned maintenances in period $t$ at the start. & $ 0 $ \\
-        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start. & $ TODO $ \\
-        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. & $ 0 $ \\
-        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. & $ 0 $\\
-        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. & $ 0 $ \\
+            $N_t$             & number of resources in planned maintenances in period $t$ at the start. & $ 0 $ \\
+            $Ret^{Init}_i$      & remaining elapsed time for resource $i$ at the start of the planning horizon. & $ \overline{D}^{on} $ \\
+            $Rut^{Init}_i$      & remaining usage time for resource $i$ at the start of the planning horizon. & $ 0 $ \\
+            $Ret^{Init}_{sum}$  & sum of remaining elapsed times at the start of the planning horizon. & $ 0 $\\
+            $Rut^{Init}_{sum}$  & sum of remaining elapsed time at the start of the planning horizon. & $ 0 $ \\
+            $MM^{min}$          & minimum number of total maintenance months for each resource. & $ |\mathcal{T}| - \overline{D}^{work} $ \\
+            $U^{max}$          & maximum unavailable months for a resource in need of maintenance. & $ \overline{D}^{off} - \underline{D}^{off} $ \\
         \end{tabular}
+         \caption{Crew Scheduling parameters translation}
+        \label{tab:parameters_crew}
+    \end{table}
 
-    % For each period $k$ from the original problem, we create a period $t \in \mathcal{T}$. We can use the elements in these two sets indistinctly.
-    % For each period $k$ with demand $b_{k}$, we create a task $j$ with $T_j = \{t\}$ that requires $R_{j} = b_{t}| t \in T_j$ resources.
-This implies that $J_t = \{j\}| t \in T_j$, $|J_t|=1$ and $D_t = b_{t}$.
+$Q_{ref}$: For an instance $i \in \mathcal{P}_{ref}$: $\exists$ a solution with the number of resources $\leq k$?
 
-    % For each worker $i$ we have a corresponding resource $i$.
-    % Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$.
-    % All resources start the planning horizon have just exited a maintenance in the previous period. This implies $N_t = 0$.
-    % The duration of the maintenance will be equal to $M = \overline{D}^{off} = \underline{D}^{off}$. This assumption is valid because the original proof was done with fixed values instead of ranges.
+$Q_{FMP}$: For an instance $I(i \in \mathcal{P}_{ref})$: $\exists$ a solution with the number of resources $\leq k$?
 
-    % The maximum remaining elapsed time after a maintenance is equal to $E = \overline{D}^{off}$. The minimum elapsed time before a maintenance is equal to $e=\underline{D}^{off}$.
-    % The usage time for each resource by any task will be $H_j = 0$ and the remaining usage time after a maintenance is $H = 0$.
-The minimum number of maintenances for each resource will be given by $M_{min}=\lceil(|\mathcal{T}| - \overline{D}^{work}) / M \rceil$.
-(Alternatively, the limit could be measured as a minimum number of maintenance months $M_{min}=|\mathcal{T}| - \overline{D}^{work}$).
-
-    % \subsection{Hypothesis}
-
-    % \begin{itemize}
-    %      \item objective: minimize the number of aircraft to maximize availability.
-    % \end{itemize}
+In this particular case, just finding a feasible solution to the Flight and Maintenance Planning Problem is equivalent as solving the Crew Scheduling Problem.
 
@@
 
-    $\mathcal{P}_{ref}$:
+    $i \in \mathcal{P}_{ref}$:
 
     \begin{tabular}{ll}
-        $\mathcal{E}$       & Employees. \\
+        $\mathcal{E}$       & employees. \\
         $\mathcal{P}$       & tasks.  \\
-        $u_p$     			&  Duration of task $p$. \\
-        $s_p$     			&  Starting times $p$. \\
+        $u_p$     			& duration of task $p$. \\
+        $s_p$     			& starting times $p$. \\
         $f_p$     			& finish time of task $p$. \\
-        $S_e$     &  Set of tasks an employee $e$ can perform. \\
-        $E_p$ &  Set of employees that can perform task $p$. \\
+        $S_e$               & set of tasks an employee $e$ can perform. \\
+        $E_p$               & set of employees that can perform task $p$. \\
     \end{tabular}
@@
 For reference, see the original description of the problem at the beginning of this document.
-It's an adaption of the MOSIM description.
+Tables (\ref{tab:sets_inter} - \ref{tab:parameters_inter}) summarize this reduction.
 
-    For each employee $e \in E$, we create an analogous resource $i \in I$ which we will use indistinctly.
+    For each employee $e \in E$, we create an analogous resource $i \in I$. We will $e$ and $i$ indistinctly.
 
-    For each task $j \in J$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$.
+    For each task $p \in P$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$. We will $p$ and $j$ indistinctly.
     Start times and finish times define the moment when the task is active: $\mathcal{T}_j = t \in \{s_j ...
@@
 The need of resources for each task will be constant $R_{j} = 1$.
-The number of resources working at any given time can $D_t = |\mathcal{J}_t|$ can be calculated.
-The relationship between tasks and employees is also equivalent: $\mathcal{O}_i = J_i$ and $\mathcal{I}_j = E_j$.
+The number of resources working at any given time will be $D_t = |\mathcal{J}_t|$.
+The relationship between tasks and employees is also equivalent: $\mathcal{O}_i = S_i$ and $\mathcal{I}_j = E_j$.
 
@@
 
-    \textbf{Sets}
+    % \textbf{Sets}
 
+    \begin{table}
         \begin{tabular}{l|l|l}
-    	\textbf{FMP}    &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
-        $\mathcal{T}$    &  Planning horizon. & $\{\min_p({s_p}) ... \max_p({f_p})\}$ \\
-        $\mathcal{I}$    &  Resources. & $ \mathcal{E}$ \\
-        $\mathcal{J}$    &  Tasks.  & $ P$ \\
-        $\mathcal{T}_j$  &  time periods in which task $j$ is active. & $t \in \{s_p ... f_p\}$ \\
-        $\mathcal{J}_t$  &  tasks to be realized in period $t$. & $j | t \in \{s_p ... f_p\}$ \\
-        $\mathcal{I}_j$  &  resources that can be assigned to task $j$. & $E_p$\\
-        $\mathcal{O}_i$  &  tasks for which resource $i$ can be used. & $P_i$ \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ that control de maintenance starts. & $\{\}$ \\
+        	\textbf{FMP}        &  \textbf{Meaning} & \textbf{I(i)} \\ \hline
+            $\mathcal{T}$       &  planning horizon. & $\{\min_p({s_p}) ... \max_p({f_p})\}$ \\
+            $\mathcal{I}$       &  resources. & $ \mathcal{E}$ \\
+            $\mathcal{J}$       &  tasks.  & $ P$ \\
+            $\mathcal{T}_j$     &  time periods in which task $j=p$ is active. & $t \in \{s_j ... f_j\}$ \\
+            $\mathcal{J}_t$     &  tasks $j=p \in \mathcal{J}$ to be realized in period $t$. & $j | t \in \{s_j ... f_j\}$ \\
+            $\mathcal{I}_j$     &  resources that can be assigned to task $j=p$. & $E_j$\\
+            $\mathcal{O}_i$     &  tasks for which resource $i=e$ can be used. & $P_i$ \\
+            $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ that control the maintenance starts. & $\{\}$ \\
         \end{tabular}
+        \caption{Interval scheduling set translation}
+        \label{tab:sets_inter}
+    \end{table}
 
@@
 
-    \textbf{Parameters}
+    \begin{table}
+        \begin{tabular}{l|l|l}
+        	\textbf{FMP}      &  \textbf{Meaning} & \textbf{I(i)} \\ \hline
+            $D_t$             & number of resources to be assigned in total in period $t$. & $ |\mathcal{J}_t| $\\
+            $A^{min}_j$       & minimum number of consecutive periods for task $j=p$. & $ u_p $ \\
+            $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start. & $ |\mathcal{T}| + 1 $ \\
+            $R_{j}$           & need of resources for task $j$ & $ 1 $ \\
+            $H_j$             & amount of resource time required by task $j$. & $ 0 $ \\
+            $M$               & maintenance duration in number of periods. & $ 0 $ \\
+            $E$               & remaining elapsed time after a maintenance. & $ 0 $ \\
+            $H$               & remaining usage time after a maintenance.& $ 0 $ \\
+            $N_t$             & number of resources in already-planned maintenances in period $t$. & $ 0 $ \\
+            $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. & $ 0 $ \\
+            $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. & $ 0 $\\
+            $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. & $ 0 $ \\
+            $MM_{min}$        & minimum number of maintenance months for each resource. & $ 0 $ \\
+            $U^{max}$         & maximum number of unavailable months for aircraft that need maintenance. & $ 0 $ \\
+        \end{tabular}
+         \caption{Interval scheduling parameters translation}
+        \label{tab:parameters_inter}
+    \end{table}
 
-    \begin{tabular}{l|l|l}
-    	\textbf{FMP}    &  \textbf{Meaning} & \textbf{f(I)} \\ \hline
-        $D_t$            & number of resources to be assigned in total in period $t$ & $|\mathcal{J}_t|$\\
-        $A^{min}_j$      & minimum number of consecutive periods for task $j$. & $u_p$ \\
-        $Ret^{Init}_{i}$ & remaining elapsed time for resource $i$ at the start. & $|\mathcal{T}| + 1$ \\
-        $R_{j}$          & Need of resources for task $j$ & $1$ \\
-        $H_j$             & amount of resource time required by task $j$. & $ 0$ \\
-        $M$               & maintenance duration in number of periods. & $ 0$ \\
-        $E$               & remaining elapsed time after a maintenance. & $ 0$ \\
-        $H$               & remaining usage time after a maintenance.& $ 0$ \\
-        $N_t$             & number of resources in already-planned maintenances in period $t$ at the start. & $0$ \\
-        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. & $ 0$ \\
-        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. & $ 0$\\
-        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. & $ 0$ \\
-    \end{tabular}
+$Q_{ref}$: For an instance $i \in \mathcal{P}_{ref}$: $\exists$ a solution with the number of resources $\leq k$?
+
+$Q_{FMP}$: For an instance $I(i \in \mathcal{P}_{ref})$: $\exists$ a solution with the number of resources $\leq k$?
+
+In this particular case, just finding a feasible solution to the Flight and Maintenance Planning Problem is equivalent as solving the Fixed Interval Scheduling Problem.
 

### 2018-07-19 latex/notes/model/model.tex
changes model formulation ton update it to current implemented model

@@
         $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
-        $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
-        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{m}+1\}},  ..., {t}$\}. \\
+        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
+        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
+        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{m}+1\}},  ..., {t}\}$. \\
         $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{M}+1\}},  ..., \max{\{1, t - Ret^{m}+1\}}$\}. \\
+        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
     \end{tabular}
@@
         $R_j$             & number of resources required by task $j$. \\
-        % $MT$              & minimum number of periods a resource can be assigned to a task. \\
+        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
         % $aut_i$           & maximal absolute usage time for resource $i$. \\
@@
         $M$               & maintenance duration in number of periods. \\
+        $M^{max}$         & maximum number of parallel maintenance operations. \\
         $E$               & remaining elapsed time after a maintenance. \\
         $H$               & remaining usage time after a maintenance. \\
-        $S$               & Storage duration in number of periods. \\
+        % $S$               & Storage duration in number of periods. \\
         $W_1$             & weight of the first objective in the objective function. \\
@@
         $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
-        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
-        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
-        $Ret^{M}$         & maximum number of periods between two consecutif maintenances. \\
-        $Ret^{m}$         & minimum number of periods between two consecutif maintenances. \\
+        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
+        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
+        $Ret^{M}$         & maximum number of periods between two consecutive maintenances. \\
+        $Ret^{m}$         & minimum number of periods between two consecutive maintenances. \\
     \end{tabular}
@@
         $a_{jti}$   &  =1 if task $j \in J$ in period $t \in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
+        $a^s_{jti}$ &  =1 if resource $i$ starts a new assignment to task $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-i)} = 0$. \\  
         $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
-        $s_{it}$    &  =1 if resource $i \in I$ starts a storage operation in period $t \in \mathcal{T}$, 0 otherwise. \\
+        % $s_{it}$    &  =1 if resource $i \in I$ starts a storage operation in period $t \in \mathcal{T}$, 0 otherwise. \\
         $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
-        $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
-        $u_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
-        $m_{max}$   &  maximal number (integer) of resources in maintenance in any period. \\
+        % $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
+        % $u_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
+        % $m_{max}$   &  maximal number (integer) of resources in maintenance in any period. \\
     \end{tabular}
@@
     \begin{align}
-        & \text{Min}\; W_1 m_{max} + W_2 u_{max}
+        & \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|}
     \end{align}
-    where weights $W_1$ and $W_2$ are chosen by the decision maker.
+
+    where we want to minimize the total number of resource hours given via maintenances and we want to maximize the final state measured in available hours of resources.
+These two objectives have the same units and can be easily compared and they ensure the resource hours are used in the most efficient way possible.
+    % where weights $W_1$ and $W_2$ are chosen by the decision maker.
 $W_1$ penalizes the maximum number of aircraft in maintenance among all periods.
@@
         % maximum capacity1:
-        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
-          & t \in \mathcal{T} \label{eq:capacity1}\\
-               % maximum capacity2:                
-        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} m_{it'} \leq m_{max}
-        % & t =m, ..., |\mathcal{T}|  \label{eq:capacity2}\\
-               %avail
+        % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
+        %   & t \in \mathcal{T} \label{eq:capacity1}\\
+        & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq M^{max} \label{eq:capacity1}\\
+        % availability
        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
         &t \in \mathcal{T} \label{eq:avalaibility1}\\
-               % maximum capacity2:                
-        %        & \sum_{t' = t - m+1}^{t} \,\, \sum_{i \in \mathcal{I}} m_{it'} + D_t\leq u_{max} 
-        % & t =m, ..., |\mathcal{T}|  \label{eq:avalaibility2}\\
+        % min assignments:
         & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
                 & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
-        % & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} \leq 1 
+        % just doing one thing at any given time:
         & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
@@
 Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task, a maintenance operation or a storage operation at the same period.
+
+    \begin{align}
+        % if start task => assign start
+        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} 
+                & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start1} \\
+      % # if we start a task, we need at least X periods of tasks assignments
+        & a^s_{jt'i} \leq a_{jti}
+        & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j, t' \in \mathcal{T}^{MT}_{jt} \label{eq:start2}
+    \end{align}
+
+    Constraints (\ref{eq:start1}) controls the definition of starting a consecutive assignment. If a task $j$ is assigned in period $t$ and was not assigned the previous period ($t-1$).
+    Constraints (\ref{eq:start2}) controls the minimum duration of a consecutive task assginment.
+If an assignment to task $j$ is started in period $t$, it needs to continue in periods $t' \in \mathcal{T}^{MT}_{jt}$.
+
     \begin{align}
         % remaining used time
-         & rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
+         & rut_{it} \leq rut_{i(t-1)} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
         & rut_{i0} = Rut^{Init}_i
@@
         %       & t = |\mathcal{T}| \label{eq:min_ret}\\
-        & \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
-              & t = |\mathcal{T}| \label{eq:min_rut} \\
+        % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
         & m_{it'} + m_{it} \leq 1
@@
         & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \leq  m_{it}
-          & t, \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\ 
+          & t, \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}
     \end{align}
@@
 
-    Finally, constraints \ref{eq:min_rut} guarantee that resources have, globally, the same amount of remaining used and elapsed times at the beginning and at the end of the planning horizon.
+    % Finally, constraints \ref{eq:min_rut} guarantee that resources have, globally, the same amount of remaining used and elapsed times at the beginning and at the end of the planning horizon.
 

### 2018-07-19 latex/notes/model/model.tex
added constraints of availability to model formulation in latex

@@
 % \usepackage{harvard}
-\usepackage[]{natbib}
+% \usepackage[]{natbib}
 \usepackage{amsmath}
@@
 \usepackage[latin1]{inputenc}
-\usepackage{aeguill}
+% \usepackage{aeguill}
 \usepackage[frenchb,english]{babel}
@@
 \setlength{\parskip}{2ex} \pagestyle{myheadings}
-\markright{\hspace*{3.8cm} \textit{MOSIM18 - June 27-29, 2018 - Toulouse - France}}
+% \markright{\hspace*{3.8cm} \textit{MOSIM18 - June 27-29, 2018 - Toulouse - France}}
 
@@
     \begin{tabular}{p{5mm}p{170mm}}
-        % $\mathcal{I}$     &  resources. \\
-        % $\mathcal{T}$     &  time periods inside planning horizon. \\
-        % $\mathcal{J}$     &  tasks. \\
+        $\mathcal{I}$     &  resources. \\
+        $\mathcal{T}$     &  time periods inside planning horizon. \\
+        $\mathcal{J}$     &  tasks. \\
+        $\mathcal{K}$     &  cluster of resources that share the same capacities to do tasks. \\
         $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
@@
         $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
+        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. \\
         $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
@@
         % $S$               & Storage duration in number of periods. \\
-        $W_1$             & weight of the first objective in the objective function. \\
-        $W_2$             & weight of the second objective in the objective function. \\
-        $N_t$             & number of resources in already-planned maintenances in period $t$ at the beginning of the planning horizon.\\
-        $D_t$             & number of resources to be assigned in total in period $t$. \\
+        % $W_1$             & weight of the first objective in the objective function. \\
+        % $W_2$             & weight of the second objective in the objective function. \\
+        % $N_t$             & number of resources in already-planned maintenances in period $t$ at the beginning of the planning horizon.\\
+        % $D_t$             & number of resources to be assigned in total in period $t$. \\
+        $AK_{kt}$         & maximum number of resources in cluster $k$ that can to be under maintenance at period $t$.\\
         $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
@@
 
-    The objective is to simultaneously minimize the maximum number of maintenances and the maximum number of unavailable aircraft.
+    The objective is to minimize the number of maintenances and maximize the end status of the resources.
 
@@
     % where weights $W_1$ and $W_2$ are chosen by the decision maker.
-$W_1$ penalizes the maximum number of aircraft in maintenance among all periods.
-$W_2$ penalizes the maximum number of unavailable aircraft among all periods.
+$W_1$ penalizes the maximum number of resources in maintenance among all periods.
+$W_2$ penalizes the maximum number of unavailable resources among all periods.
     The following constraints are used in the model:       
@@
         % availability
-       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
-        &t \in \mathcal{T} \label{eq:avalaibility1}\\
+       % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
+       %  &t \in \mathcal{T} \label{eq:avalaibility1}\\
         % min assignments:
@@
     Maintenance capacity is controlled by (\ref{eq:capacity1}).
-The number of unavailable resources is defined by (\ref{eq:avalaibility1}).
 Tasks' resource requirements are defined by (\ref{eq:taskres}).
@@
     Constraints (\ref{eq:start2}) controls the minimum duration of a consecutive task assginment.
-If an assignment to task $j$ is started in period $t$, it needs to continue in periods $t' \in \mathcal{T}^{MT}_{jt}$.
+If an assignment to task $j$ is started in period $t$, it needs to continue in the following periods: $t' \in \mathcal{T}^{MT}_{jt}$.
+
+    \begin{align}
+       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} \leq AK_{kt}
+        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster}
+    \end{align}
+
+    Constraints (\ref{eq:avalaibility-cluster}) control a minimum availability for each cluster $k$ of resources.
+A cluster is defined by the most general group of resources that is required, exclusively, by at least one task.
 
@@
     The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}).
-    The remaining elapsed time is defined by (\ref{eq:ret_min})-(\ref{eq:ret_max}).
+
+    The minimum elapsed time is defined by (\ref{eq:ret_min}). If a maintenance is done in some period $t$, we know that it cannot be done in the periods $t'$ that are close to it.
+
+    On the other hand, due to the maximum elapsed time between maintenances, constraints (\ref{eq:ret_max}) force that, if we have a maintenance on period $t$, we have to have done at least another in one far away period $t'$.
 

### 2018-07-23 latex/notes/model/model.tex
changes to model's latex formulation

@@
         % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
-        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{m}+1\}},  ..., {t}\}$. \\
-        $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - Ret^{M}+1\}},  ..., \max{\{1, t - Ret^{m}+1\}}$\}. \\
+        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - E^{m}+1\}},  ..., {t}\}$. \\
+        $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - E^{M}+1\}},  ..., \max{\{1, t - E^{m}+1\}}$\}. \\
         $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
+        % $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - E^{M}+1\}},  ..., \max{\{1, t - E^{m}+1\}}$\}. \\
     \end{tabular}
@@
   \subsection{Parameters}
+
+  % TODO: previously assigned missions.
+  % TODO: previous ret.
+  % TODO: previous maintenaces.
 
@@
         $M$               & maintenance duration in number of periods. \\
-        $M^{max}$         & maximum number of parallel maintenance operations. \\
-        $E$               & remaining elapsed time after a maintenance. \\
+        $M^{max}$         & maximum number of simultaneous maintenance operations. \\
+        % $E$               & remaining elapsed time after a maintenance. \\
+        $E^{M}$         & maximum number of periods between two consecutive maintenances. \\
+        $E^{m}$         & minimum number of periods between two consecutive maintenances. \\
         $H$               & remaining usage time after a maintenance. \\
@@
         % $D_t$             & number of resources to be assigned in total in period $t$. \\
-        $AK_{kt}$         & maximum number of resources in cluster $k$ that can to be under maintenance at period $t$.\\
+        $AK_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
         $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
         $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
+        % $A_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
         % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
         % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
-        $Ret^{M}$         & maximum number of periods between two consecutive maintenances. \\
-        $Ret^{m}$         & minimum number of periods between two consecutive maintenances. \\
     \end{tabular}
@@
 
-    The objective is to minimize the number of maintenances and maximize the end status of the resources.
+    The objective is to minimize the total number of maintenances and, at the same time, maximize the end status of the resources.
 
@@
     \end{align}
+
+    where the first expression counts all the resource hours given via maintenances and the second expression measures the amount of remaining hours at all resources at the end of the planning horizon.
 
@@
 Tasks' resource requirements are defined by (\ref{eq:taskres}).
-Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task, a maintenance operation or a storage operation at the same period.
+Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task or a maintenance operation at the same period.
 
@@
 
-    Constraints (\ref{eq:start1}) controls the definition of starting a consecutive assignment. If a task $j$ is assigned in period $t$ and was not assigned the previous period ($t-1$).
-    Constraints (\ref{eq:start2}) controls the minimum duration of a consecutive task assginment.
-If an assignment to task $j$ is started in period $t$, it needs to continue in the following periods: $t' \in \mathcal{T}^{MT}_{jt}$.
+    Constraints (\ref{eq:start1}) controls the definition of starting a consecutive assignment: the variable activates if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this task in the previous period ($t-1$).
+    Constraints (\ref{eq:start2}) controls the minimum duration of a consecutive task assignment.
+If an assignment to task $j$ is started in period $t$, it needs to continue in the following $t' \in \mathcal{T}^{MT}_{jt}$ periods.
 
@@
 
-    The minimum elapsed time is defined by (\ref{eq:ret_min}). If a maintenance is done in some period $t$, we know that it cannot be done in the periods $t'$ that are close to it.
-
-    On the other hand, due to the maximum elapsed time between maintenances, constraints (\ref{eq:ret_max}) force that, if we have a maintenance on period $t$, we have to have done at least another in one far away period $t'$.
+    The minimum and maximum elapsed time are defined by (\ref{eq:ret_min}) and (\ref{eq:ret_max}) respectively.
+Constraints (\ref{eq:ret_min}) guarantee that if a maintenance is done in some period $t$, we know that it cannot be done in the following $t' \in \mathcal{T}^{m}_t$ periods.
+Constraints (\ref{eq:ret_max}) guarantee that, if we have a maintenance in period $t$, we have to start at least one maintenance among periods $t' \in \mathcal{T}^{M}_t}$.
 

### 2018-07-24 latex/notes/model/model.tex
Finished adding remaining constraints to model.

@@
         % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
-        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - E^{m}+1\}},  ..., {t}\}$. \\
-        $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - E^{M}+1\}},  ..., \max{\{1, t - E^{m}+1\}}$\}. \\
+        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + E^{m}-1\}}\}$. \\
+        $\mathcal{T}^{M}_t$ &  time periods $t' \leq |\mathcal{T}| - E^{M}$ such that $t' \in \{ t + E^{m}-1 , ...,  t + E^{M}-1 \}$. \\
+        $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Ret^{Init}_{i} - E^M + E^m \}}\}$. \\
+        $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Ret^{Init}_{i} - E^M + E^m \}} , ...,  Ret^{Init}_{i} \}$. \\
         $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
-        % $\mathcal{T}^{M}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - E^{M}+1\}},  ..., \max{\{1, t - E^{m}+1\}}$\}. \\
     \end{tabular}
@@
   % TODO: previously assigned missions.
-  % TODO: previous ret.
-  % TODO: previous maintenaces.
 
@@
         $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
-        % $A_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
+        $A^{init}_{jti}$  & =1 if resource $i \in \mathcal{I}$ has task $j$ pre-assigned in period $t$. \\
         % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
@@
 
-    where the first expression counts all the resource hours given via maintenances and the second expression measures the amount of remaining hours at all resources at the end of the planning horizon.
+    where the first expression counts all the usage hours given to resources via maintenances and the second expression measures the amount of remaining usage hours in the resources at the end of the planning horizon.
 
-    where we want to minimize the total number of resource hours given via maintenances and we want to maximize the final state measured in available hours of resources.
-These two objectives have the same units and can be easily compared and they ensure the resource hours are used in the most efficient way possible.
-    % where weights $W_1$ and $W_2$ are chosen by the decision maker.
-$W_1$ penalizes the maximum number of resources in maintenance among all periods.
-$W_2$ penalizes the maximum number of unavailable resources among all periods.
+    We want to minimize the total number of maintenances and we want to maximize the final state measured.
+These two objectives have the same units, can be easily compared and they ensure the resource hours are used in the most efficient way possible.
+
     The following constraints are used in the model:       
@@
         % if start task => assign start
-        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} 
+        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} - A^{init}_{j(t-1)i}
                 & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start1} \\
@@
 
-    Constraints (\ref{eq:start1}) controls the definition of starting a consecutive assignment: the variable activates if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this task in the previous period ($t-1$).
-    Constraints (\ref{eq:start2}) controls the minimum duration of a consecutive task assignment.
-If an assignment to task $j$ is started in period $t$, it needs to continue in the following $t' \in \mathcal{T}^{MT}_{jt}$ periods.
+    Constraints (\ref{eq:start1}) control the definition of starting a consecutive assignment: the variable is active if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this same task in the previous period ($t-1$).
+The previous period can include the period before the start of the planning horizon ($A^{init}_{j(t-1)i}$ parameter).
+    Constraints (\ref{eq:start2}) control the minimum duration of a consecutive task assignment.
+If an assignment to task $j$ is started in period $t$, it needs to continue for the following $t' \in \mathcal{T}^{MT}_{jt}$ periods.
 
@@
 
-    Constraints (\ref{eq:avalaibility-cluster}) control a minimum availability for each cluster $k$ of resources.
-A cluster is defined by the most general group of resources that is required, exclusively, by at least one task.
+    Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability for each cluster $k$ of resources.
+A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task.
 
@@
         % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
-        & m_{it'} + m_{it} \leq 1
-          & t, \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
-        & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \leq  m_{it}
-          & t, \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}
     \end{align}
-
-    % \begin{align}
-    % & m_{it'} + m_{it} \leq 1
-    %   & t, \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_better}\\ 
-    % & \sum_{t' \in \mathcal{T}^{M}_t - \mathcal{T}^{m}_t} m_{it'} \leq  m_{it}
-    %   & t, \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_better}\\ 
-    % \end{align}
 
@@
 
+    \begin{align}
+        & m_{it'} + m_{it} \leq 1
+          & t \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
+        & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
+          & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\
+        & m_{it} = 0
+          & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I}\label{eq:ret_min_init} \\
+        & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
+          & i \in \mathcal{I}\label{eq:ret_max_init}
+    \end{align}
+
     The minimum and maximum elapsed time are defined by (\ref{eq:ret_min}) and (\ref{eq:ret_max}) respectively.
-Constraints (\ref{eq:ret_min}) guarantee that if a maintenance is done in some period $t$, we know that it cannot be done in the following $t' \in \mathcal{T}^{m}_t$ periods.
-Constraints (\ref{eq:ret_max}) guarantee that, if we have a maintenance in period $t$, we have to start at least one maintenance among periods $t' \in \mathcal{T}^{M}_t}$.
+Constraints (\ref{eq:ret_min}) guarantee that if a maintenance is done in some period $t$, we know that it cannot be done in the immediately consecutive $t' \in \mathcal{T}^{m}_t$ periods.
+Constraints (\ref{eq:ret_max}) guarantee that, if we have a maintenance in period $t$, we need to start at least one maintenance among periods $t' \in \mathcal{T}^{M}_t$.
+Constraints (\ref{eq:ret_min_init}) and (\ref{eq:ret_max_init}) control the minimum and maximum remaining elapsed time at the beginning of the planning period, respectively.
+They follow the same rules as the previous constraints for the particular case of the start of the planning horizon.
 

### 2018-08-21 latex/notes/complexite/FMP_variants.tex
updated model formulation and complexity analysis

@@
 \clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Crew scheduling variant}
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \section{Crew scheduling variant}
 
-    As presented in \cite{Brunner2013}. There, they prove the complexity by comparing the crew scheduling problem with the circulant problem. Below I do the same with our problem.
+%     As presented in \cite{Brunner2013}. There, they prove the complexity by comparing the crew scheduling problem with the circulant problem. Below I do the same with our problem.
 
-    \subsection{Hypothesis}
+%     \subsection{Hypothesis}
 
-    \begin{itemize}
-     \item fixed maintenance duration.
-     \item fixed duration between maintenances.
-     \item homogeneous fleet.
-     \item objective: minimize the number of aircraft.
-    \end{itemize}
+%     \begin{itemize}
+%      \item fixed maintenance duration.
+%      \item fixed duration between maintenances.
+%      \item homogeneous fleet.
+%      \item objective: minimize the number of aircraft.
+%     \end{itemize}
 
-    \subsection{Reductions}
+%     \subsection{Reductions}
 
-    \begin{itemize}
-        \item The missions consume 0 hours of aircraft.
-        \item The minimum time between maintenances is equal to the maximum.
-        \item All missions can use any aircraft.
-    \end{itemize}
+%     \begin{itemize}
+%         \item The missions consume 0 hours of aircraft.
+%         \item The minimum time between maintenances is equal to the maximum.
+%         \item All missions can use any aircraft.
+%     \end{itemize}
 
-    \subsection{Limitations}
+%     \subsection{Limitations}
 
-    \begin{itemize}
-        \item There is not starting period.
-    \end{itemize}
+%     \begin{itemize}
+%         \item There is not starting period.
+%     \end{itemize}
 
-    \subsection{Notations}
+%     \subsection{Notations}
 
-    \textbf{Sets}
+%     \textbf{Sets}
 
-    \begin{tabular}{ll}
-    $t \in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
-    $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
-    \end{tabular}
+%     \begin{tabular}{ll}
+%     $t \in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
+%     $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
+%     \end{tabular}
 
-    \vskip 0.3cm
+%     \vskip 0.3cm
 
-    \subsection{Parameters}
+%     \subsection{Parameters}
 
-    \begin{tabular}{ll}
-        $D^{off}$ & Number of consecutive non-working days. \\
-        $D^{on}$ & Number of consecutive working days. \\
-        $\mathcal{A}_{it}$ & Pattern $i$ covers period $t$ (circulant matrix built from $D_{on}$ and $D_{off}$).\\
-        $b_{t}$ & Need of resources at period $t$. \\
-    \end{tabular}
+%     \begin{tabular}{ll}
+%         $D^{off}$ & Number of consecutive non-working days. \\
+%         $D^{on}$ & Number of consecutive working days. \\
+%         $\mathcal{A}_{it}$ & Pattern $i$ covers period $t$ (circulant matrix built from $D_{on}$ and $D_{off}$).\\
+%         $b_{t}$ & Need of resources at period $t$. \\
+%     \end{tabular}
 
-    \vskip 0.3cm
+%     \vskip 0.3cm
 
-    Example of A for $D_{on}=4$ and $D_{off}=2$
+%     Example of A for $D_{on}=4$ and $D_{off}=2$
 
-    $$
-    \begin{bmatrix}
-        1 & 0 & 0 & \dots  & A_{i1} \\
-        1 & 1 & 0 & \dots  & A_{i2} \\
-        1 & 1 & 1 & \dots  & A_{i3} \\
-        1 & 1 & 1 & \dots  & A_{i4} \\
-        0 & 1 & 1 & \dots  & A_{i5} \\
-        0 & 0 & 1 & \dots  & A_{i6} \\
-        1 & 0 & 0 & \dots  & A_{i6} \\
-        \vdots & \vdots & \vdots & \ddots & \vdots \\
-        A_{1t} & A_{2t} & A_{3t} & \dots  & A_{nm}
-    \end{bmatrix}
-    $$
+%     $$
+%     \begin{bmatrix}
+%         1 & 0 & 0 & \dots  & A_{i1} \\
+%         1 & 1 & 0 & \dots  & A_{i2} \\
+%         1 & 1 & 1 & \dots  & A_{i3} \\
+%         1 & 1 & 1 & \dots  & A_{i4} \\
+%         0 & 1 & 1 & \dots  & A_{i5} \\
+%         0 & 0 & 1 & \dots  & A_{i6} \\
+%         1 & 0 & 0 & \dots  & A_{i6} \\
+%         \vdots & \vdots & \vdots & \ddots & \vdots \\
+%         A_{1t} & A_{2t} & A_{3t} & \dots  & A_{nm}
+%     \end{bmatrix}
+%     $$
 
-    \subsection{Variables}
+%     \subsection{Variables}
 
-    \begin{itemize}
-     \item $x_i$: number of patterns of type $i$ to be included in the solution.
-    \end{itemize}
+%     \begin{itemize}
+%      \item $x_i$: number of patterns of type $i$ to be included in the solution.
+%     \end{itemize}
 
-    \subsection{Model}
+%     \subsection{Model}
 
-    \begin{align}
-        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i \label{numPatterns}
-    \end{align}
+%     \begin{align}
+%         & \text{Min}\; \sum_{i \in \mathcal{P}} x_i \label{numPatterns}
+%     \end{align}
 
-    \begin{align}
-        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq d_j & t \in \mathcal{T} \label{demand} \\
-        & x_{i} \geq 0,\, x_{i} \in \mathbb{Z} \\
-    \end{align}
+%     \begin{align}
+%         & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq d_j & t \in \mathcal{T} \label{demand} \\
+%         & x_{i} \geq 0,\, x_{i} \in \mathbb{Z} \\
+%     \end{align}
 
-    (\ref{numPatterns}) minimizes the number of workers. (\ref{demand}) defines the need to have at least a given number of workers at any given time.
+%     (\ref{numPatterns}) minimizes the number of workers. (\ref{demand}) defines the need to have at least a given number of workers at any given time.
 
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \clearpage
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Set covering variant}
+% \section{Set covering variant}
 
-    \subsection{Hypothesis}
+%     \subsection{Hypothesis}
 
-    \begin{itemize}
-     \item only 1 aircraft needed per month.
-     \item objective: minimize the number of aircraft.
-    \end{itemize}
+%     \begin{itemize}
+%      \item only 1 aircraft needed per month.
+%      \item objective: minimize the number of aircraft.
+%     \end{itemize}
 
-    \subsection{Notations}
+%     \subsection{Notations}
 
-    \textbf{Sets}
+%     \textbf{Sets}
 
-    \begin{tabular}{ll}
-    $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
-    $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
-    \end{tabular}
+%     \begin{tabular}{ll}
+%     $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
+%     $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
+%     \end{tabular}
 
-    \vskip 0.3cm
+%     \vskip 0.3cm
 
-    \subsection{Parameters}
+%     \subsection{Parameters}
 
-    \begin{tabular}{ll}
-        $\mathcal{A}_{it}$ & Pattern $i$ covers period $t$.\\
-    \end{tabular}
+%     \begin{tabular}{ll}
+%         $\mathcal{A}_{it}$ & Pattern $i$ covers period $t$.\\
+%     \end{tabular}
 
-    \vskip 0.3cm
+%     \vskip 0.3cm
 
-    \subsection{Variables}
+%     \subsection{Variables}
 
-    \begin{itemize}
-     \item $x_i$: is pattern $i$ in the solution or not.
-    \end{itemize}
+%     \begin{itemize}
+%      \item $x_i$: is pattern $i$ in the solution or not.
+%     \end{itemize}
 
-    \subsection{Model}
+%     \subsection{Model}
 
-    \begin{align}
-        & \text{Min}\; \sum_{i \in \mathcal{P}} x_i \label{numPatterns2}
-    \end{align}
+%     \begin{align}
+%         & \text{Min}\; \sum_{i \in \mathcal{P}} x_i \label{numPatterns2}
+%     \end{align}
 
-    \begin{align}
-        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq 1 & t \in \mathcal{T} \label{demand2} \\
-        & x_{i} \in \{0, 1\} \\
-    \end{align}
+%     \begin{align}
+%         & \sum_{i \in \mathcal{P}} A_{it}x_{i} \geq 1 & t \in \mathcal{T} \label{demand2} \\
+%         & x_{i} \in \{0, 1\} \\
+%     \end{align}
 
-    (\ref{numPatterns2}) minimizes the number of workers. (\ref{demand2}) defines the need to have at least one worker working at any given time.
+%     (\ref{numPatterns2}) minimizes the number of workers. (\ref{demand2}) defines the need to have at least one worker working at any given time.
 
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \clearpage
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Scheduling variant}
+% \section{Scheduling variant}
 
-    \subsection{Hypothesis}
+%     \subsection{Hypothesis}
 
-    \begin{itemize}
-     \item durée de maintenance définie
-     \item fenêtres de temps pour commencer les maintenances
-     \item objectif: lisser les maintenances dans les fenêtres de temps prédéfinies
-    \end{itemize}
+%     \begin{itemize}
+%      \item durée de maintenance définie
+%      \item fenêtres de temps pour commencer les maintenances
+%      \item objectif: lisser les maintenances dans les fenêtres de temps prédéfinies
+%     \end{itemize}
 
-    \subsection{Notations}
+%     \subsection{Notations}
 
-    \noindent{
-    \begin{tabular}{ll}
-        $t\in \TT = [\![1, T]\!]$ & Discrétisation temporelle de l'horizon de planigfication. \\
-        $i \in \II = [\![1, I]\!]$ & Ensemble des avions.\\
-        $k \in \KK_i = [\![1, K_i]\!]$ & Ensemble des maintenances à planifier successivement.\\ 
-    \end{tabular}
-    }
+%     \noindent{
+%     \begin{tabular}{ll}
+%         $t\in \TT = [\![1, T]\!]$ & Discrétisation temporelle de l'horizon de planigfication. \\
+%         $i \in \II = [\![1, I]\!]$ & Ensemble des avions.\\
+%         $k \in \KK_i = [\![1, K_i]\!]$ & Ensemble des maintenances à planifier successivement.\\ 
+%     \end{tabular}
+%     }
 
-    \vskip 0.3cm
+%     \vskip 0.3cm
 
-    \textbf{Paramètres}
+%     \textbf{Paramètres}
 
-    \begin{tabular}{ll}
-        $\mathbf{Da}_{i,k}$ & Durée nominale de la maintenance $k$ de l'avion $i$.\\
-        $\mathbf{l}_{i}$ & Durée mininale du cycle de vol de l'avion $i$.\\
-        $\mathbf{{To}}_{i,k}$ & Pas de temps le plus tôt pour commencer la maintenance $k$ de l'avion $i$.\\
-        $\mathbf{{Ta}}_{i,k}$ & Pas de temps le plus tard pour terminer la maintenance $k$ de l'avion $i$.\\
-    \end{tabular}
+%     \begin{tabular}{ll}
+%         $\mathbf{Da}_{i,k}$ & Durée nominale de la maintenance $k$ de l'avion $i$.\\
+%         $\mathbf{l}_{i}$ & Durée mininale du cycle de vol de l'avion $i$.\\
+%         $\mathbf{{To}}_{i,k}$ & Pas de temps le plus tôt pour commencer la maintenance $k$ de l'avion $i$.\\
+%         $\mathbf{{Ta}}_{i,k}$ & Pas de temps le plus tard pour terminer la maintenance $k$ de l'avion $i$.\\
+%     \end{tabular}
 
-    % \caption{Definition and notations the challenge ROADEF}\label{contraintesROADEF}
-    % \end{table}
+%     % \caption{Definition and notations the challenge ROADEF}\label{contraintesROADEF}
+%     % \end{table}
 
-    \subsection{Définition des variables}
+%     \subsection{Définition des variables}
 
-    \begin{itemize}
-     \item $d_{i,k,t}\in \{0,1\}$; variables binaires pour indiquer les dates d'arrêts, "step varaibles" 
-    $d_{i,k,t}=1$ si l'arrêt du cycle $k$ du réacteur nucléaire $i$ a été commencé avant la semaine $t$.
-    $d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \in \{0,1\}$ vaut $1$ uniquement pendant a maintenance $(i,k)$.
+%     \begin{itemize}
+%      \item $d_{i,k,t}\in \{0,1\}$; variables binaires pour indiquer les dates d'arrêts, "step varaibles" 
+%     $d_{i,k,t}=1$ si l'arrêt du cycle $k$ du réacteur nucléaire $i$ a été commencé avant la semaine $t$.
+%     $d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \in \{0,1\}$ vaut $1$ uniquement pendant a maintenance $(i,k)$.
 
-     \item $m,M\geqslant 0$: variables continues pour linéariser la fonction objectif
-    \end{itemize}
+%      \item $m,M\geqslant 0$: variables continues pour linéariser la fonction objectif
+%     \end{itemize}
 
-    \subsection{Formulation PLNE}
+%     \subsection{Formulation PLNE}
 
-    \begin{eqnarray}
-    \displaystyle \min_{d,m,M} & M-m
-     & \\
-    \forall i , k, t, & d_{i,k,t-1}\leqslant d_{i,k,t} & \label{PANprecedence}\\
-    \forall i , k, & d_{i,k,\mathbf{To}_{i,k}-1}\leqslant 0 & \label{PANtw0} \\
-    \forall i , k, & d_{i,k,\mathbf{Ta}_{i,k}}\geqslant 1 & \label{PANtw1}\\
-    \forall i , k, t, & d_{i,k,t}\leqslant d_{i,k-1,t-l_i-\mathbf{Da}_{i,k}} & \label{PANprecedence2}\\
-    \forall t, & \displaystyle \sum_{i,k} d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \leqslant M & \label{maxMaint}\\
-    \forall t, & \displaystyle \sum_{i,k} d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \geqslant m & \label{minMaint}\\
-    & d \in \{0,1\}^N, m,M \geqslant 0
-    \end{eqnarray}
+%     \begin{eqnarray}
+%     \displaystyle \min_{d,m,M} & M-m
+%      & \\
+%     \forall i , k, t, & d_{i,k,t-1}\leqslant d_{i,k,t} & \label{PANprecedence}\\
+%     \forall i , k, & d_{i,k,\mathbf{To}_{i,k}-1}\leqslant 0 & \label{PANtw0} \\
+%     \forall i , k, & d_{i,k,\mathbf{Ta}_{i,k}}\geqslant 1 & \label{PANtw1}\\
+%     \forall i , k, t, & d_{i,k,t}\leqslant d_{i,k-1,t-l_i-\mathbf{Da}_{i,k}} & \label{PANprecedence2}\\
+%     \forall t, & \displaystyle \sum_{i,k} d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \leqslant M & \label{maxMaint}\\
+%     \forall t, & \displaystyle \sum_{i,k} d_{i,k,t}-d_{i,k,t- {\mathbf{Da}_{i,k}}} \geqslant m & \label{minMaint}\\
+%     & d \in \{0,1\}^N, m,M \geqslant 0
+%     \end{eqnarray}
 
-    (\ref{PANprecedence}): contrainte induite par la définition ``step variables".
+%     (\ref{PANprecedence}): contrainte induite par la définition ``step variables".
 
-    (\ref{PANtw0}), (\ref{PANtw1}): contraintes de fenêtres de temps (écrit explicitement, ou sinon définir uniquement les variables $d_{i,k,t}$
-    dans les fenêtres de temps)
+%     (\ref{PANtw0}), (\ref{PANtw1}): contraintes de fenêtres de temps (écrit explicitement, ou sinon définir uniquement les variables $d_{i,k,t}$
+%     dans les fenêtres de temps)
 
-    (\ref{PANprecedence2}): durée minimale d'un cycle de vol $l\geqslant 0$, contrainte de succession des cycles.
+%     (\ref{PANprecedence2}): durée minimale d'un cycle de vol $l\geqslant 0$, contrainte de succession des cycles.
 
-    (\ref{maxMaint}), (\ref{minMaint}): contraites pour linéariser les $\min$ et $\max$ de la focntion objectif de lissage
+%     (\ref{maxMaint}), (\ref{minMaint}): contraites pour linéariser les $\min$ et $\max$ de la focntion objectif de lissage
 
-    \subsection{Questions ouvertes}
+%     \subsection{Questions ouvertes}
 
-    \begin{itemize}
-    %\item Avec $l=0$ , y a t'il un problème de référence proche ? Ordonnancement, lissage de charge ?
-    \item Avec $l=0$ , y a t'il un problème de référence proche ? Ordonnancement, lissage de charge ?
-    \item Est-ce un problème NP-complet? Peut on le prouver ?
-     \item Cas particulier $\mathbf{Ta}_{i,k}-\mathbf{To}_{i,k}=1$? (pour lissage local)
-     \item Autres fonctions objectifs?
-     \item version robuste avec aléas de prolongation de durée de maintenance?
-    \end{itemize}
+%     \begin{itemize}
+%     %\item Avec $l=0$ , y a t'il un problème de référence proche ? Ordonnancement, lissage de charge ?
+%     \item Avec $l=0$ , y a t'il un problème de référence proche ? Ordonnancement, lissage de charge ?
+%     \item Est-ce un problème NP-complet? Peut on le prouver ?
+%      \item Cas particulier $\mathbf{Ta}_{i,k}-\mathbf{To}_{i,k}=1$? (pour lissage local)
+%      \item Autres fonctions objectifs?
+%      \item version robuste avec aléas de prolongation de durée de maintenance?
+%     \end{itemize}
 
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \clearpage
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Crew scheduling: smooth maintenances}
+% \section{Crew scheduling: smooth maintenances}
 
-    \subsection{Hypothesis}
+%     \subsection{Hypothesis}
 
-    \begin{itemize}
-     \item fixed maintenance duration.
-     \item range to start each maintenance.
-     \item min distance between maintenances.
-     \item objective: minimize the max number of maintenances.
-    \end{itemize}
+%     \begin{itemize}
+%      \item fixed maintenance duration.
+%      \item range to start each maintenance.
+%      \item min distance between maintenances.
+%      \item objective: minimize the max number of maintenances.
+%     \end{itemize}
 
-    \subsection{Notations}
+%     \subsection{Notations}
 
-    \textbf{Sets}
+%     \textbf{Sets}
 
-    \begin{tabular}{ll}
-    $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
-    $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
-    \end{tabular}
+%     \begin{tabular}{ll}
+%     $t\in \mathcal{T} = [\![1, m]\!]$ & Planning horizon periods. \\
+%     $i \in \mathcal{P} = [\![1, n]\!]$ & Working pattern.\\
+%     \end{tabular}
 
-    \vskip 0.3cm
+%     \vskip 0.3cm
 
-    \subsection{Parameters}
+%     \subsection{Parameters}
 
-    \begin{tabular}{ll}
-        $\mathcal{A}_{it}$ & Pattern $i$ has a maintenance in $t$.\\
-    \end{tabular}
+%     \begin{tabular}{ll}
+%         $\mathcal{A}_{it}$ & Pattern $i$ has a maintenance in $t$.\\
+%     \end{tabular}
 
-    \vskip 0.3cm
+%     \vskip 0.3cm
 
-    Here, A needs to be created taking into account:
+%     Here, A needs to be created taking into account:
 
-    \begin{enumerate}
-        \item The range of start of maintenances.
-        \item The fixed duration of maintenances.
-        \item The minimal distance between maintenances.
-    \end{enumerate}
+%     \begin{enumerate}
+%         \item The range of start of maintenances.
+%         \item The fixed duration of maintenances.
+%         \item The minimal distance between maintenances.
+%     \end{enumerate}
 
-    \subsection{Variables}
+%     \subsection{Variables}
 
-    \begin{itemize}
-        \item $x_i$: number of patterns of type $i$ to be included in the solution.
-        \item $M$: maximum number of maintenance among all periods.
-    \end{itemize}
+%     \begin{itemize}
+%         \item $x_i$: number of patterns of type $i$ to be included in the solution.
+%         \item $M$: maximum number of maintenance among all periods.
+%     \end{itemize}
 
-    \subsection{Model}
+%     \subsection{Model}
 
-    \begin{align}
-        & \text{Min}\; M
-    \end{align}
+%     \begin{align}
+%         & \text{Min}\; M
+%     \end{align}
 
-    \begin{align}
-        & \sum_{i \in \mathcal{P}} A_{it}x_{i} \leq M & t \in \mathcal{T} \label{numMaint} \\
-        & x_{i} \geq 0,\; integer \\
-    \end{align}
+%     \begin{align}
+%         & \sum_{i \in \mathcal{P}} A_{it}x_{i} \leq M & t \in \mathcal{T} \label{numMaint} \\
+%         & x_{i} \geq 0,\; integer \\
+%     \end{align}
 
-        (\ref{numMaint}) defines the $M$ variable as the maximum number of maintenances at any period of time.
+%         (\ref{numMaint}) defines the $M$ variable as the maximum number of maintenances at any period of time.
 
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \clearpage
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Scheduling variant 2}
+% \section{Scheduling variant 2}
 
-    As modelled in \cite{Rieck2012} and \cite{Neumann2000}.
+%     As modelled in \cite{Rieck2012} and \cite{Neumann2000}.
 
-    \subsection{Hypothesis}
+%     \subsection{Hypothesis}
 
-    \begin{itemize}
-     \item fixed maintenance duration.
-     \item min and max durations between maintenances that belong to the same aircraft.
-     \item each maintenance uses an amount of resource. In this case 1.
-     \item there is a maintenance capacity to respect at every period.
-     \item objective: minimize the max number of maintenances.
-     \item There are two artificial activities representing the start and end of the horizon.
-     \item maintenance operations consume also resources of the missions.
-    \end{itemize}
-
-    \subsection{Reductions}
-
-    \begin{itemize}
-        \item The missions consume 0 hours of aircraft.
-        %\item The missions require 0 aircraft at each period.
-        \item Each aircraft can only be used for one mission.
-    \end{itemize}
-
-    \subsection{Limitations}
-
-    \begin{itemize}
-        \item Resource usage is always 1.
-        %\item There is only one type of resource.
-        \item All durations are the same.
-    \end{itemize}
-
-    \subsection{Notations}
-
-    \textbf{Sets}
-
-    \noindent{
-    \begin{tabular}{ll}
-        $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
-        $(i, j) \in \mathcal{E} = [\![1, v]\!]$ & Set of consecutive (i, j) maintenance operations, i.e. consecutive maintenances belonging to the same aircraft.\\
-        $k \in \mathcal{R}$ & Types of resources that maintenance can take. We only have one type.\\
-        $i = 0$ & Artificial starting node.\\
-        $i = n+1$ & Artificial ending node.\\
-    \end{tabular}
-    }
-
-    \vskip 0.3cm
-
-    The set $\mathcal{A}$ we will construct it by deciding in advance the maintenances $i$ that share the same aircraft.
-
-    \subsection{Parameters}
-
-    \begin{tabular}{ll}
-        $p_i$ & Duration of maintenance $i$. In this case it's always p=6.\\
-        $\mathbf{d}$ & Last period on the horizon.\\
-        $r_{ik}$ & Resources of type $k$ taken by maintenance $i$. In this case it's always 1.\\
-        $d^{max}_{ij}$ & Maximum duration between two maintenances $i$ and $j$.\\
-        $d^{min}_{ij}$ & Minimum duration between two maintenances $i$ and $j$.\\
-        ${\delta}_{ij}$ & Precedence duration arcs between two maintenances $i$ and $j$ based on $d^{max}$ and $d^{min}$.\\
-        $R_k$ & total maintenance capacity at each period for resource $k$.\\
-        $c_k$ & cost per unit of investment on resource $k$.
-        % $\mathbf{{To}}_{i,k}$ & Minimum starting period for maintenance $k$ of resource $i$.\\
-        % $\mathbf{{Ta}}_{i,k}$ & Maximum starting period for maintenance $k$ of resource $i$.\\
-    \end{tabular}
-
-    \subsection{Variables}
-
-    \begin{tabular}{ll}
-        $S_i$ & the start of maintenance operation $i$.\\
-        $\mathcal{A}(S^{\mathcal{C}}, t)$ & set of maintenance operations active at moment $t$.\\
-    \end{tabular}
-
-    \subsection{Model}
-
-    The following formulation is taken from \cite{Neumann2000}.
-
-    \begin{align}
-        \min f(\mathbf{R}(S)) = \sum_{k \in \mathcal{R}} c_k \max_{t \in \mathcal{T}} r_k(S^{\mathcal{C}}, t)
-    \end{align}
-
-    \begin{align}
-        & S_j - S_i \geq \delta_{ij} & (i, j) \in \mathcal{E}  \label{precedence}\\
-        & S_0 = 0 &  \label{firstjob}\\
-        & S_{n+1} \leq \mathbf{d}   \label{lastjob}\\
-        & S_{i} \in \mathbb{Z}_{\geq 0} & i \in \mathcal{V}  \label{integervars}\\
-        & r_k(S^{\mathcal{C}}, t) = \sum_{i \in \mathcal{A}(S^{\mathcal{C}}, t)} r_{ik}  \label{consumptiondef}\\
-        & r_k(S^{\mathcal{C}}, t) \leq R_k \label{consumptionmax}\\
-    \end{align}
-
-    (\ref{precedence}) constraints define the relative minimum and maximum starting times of tasks.
-    (\ref{firstjob}) and (\ref{lastjob}) constraints define the fictitious start and end jobs.
-    (\ref{integervars}) define the range for the starting time variables: integer positive.
-    (\ref{consumptiondef}) defines the used quantity for each type of resource depending on the starting times and durations and (\ref{consumptionmax}) defines the maximum amount of resources of each type that can be used at any given time.
-
-    \subsection{Possible proof}
-
-    \cite{Blazewicz1983} proved that the following problem is NP-Hard:
-
-    $$P2 | res111, chain, p_j = 1 / C_{max}$$ is NP-hard via the 3 partition problem.
-
-    This problem is, from what I understand, the following:
-
-    \begin{itemize}
-        \item The number of machines is 2.
-        \item There is only one type of resource.
-        \item The resource size is 1.
-        \item The consumption of the resource by each job is upper bounded by 1.
-        \item The duration of each job is 1.
-        \item There is a chain precedence between jobs: a job can have at most one predecessor and one successor.
-    \end{itemize}
-
-    Here they use the machine and resource concept.
-This could be understood as there being TWO resource types.
-Not sure how to really come around this.
-Is the same, for example, as doing the following:
-
-    $$P| res1R1, chain, p_j = 1 / C_{max}$$ Where I have an unlimited number of machines and a quantity R of available resource?
-
-    Although they minimize the maximum completion time, the decision problem 'is there a planning that uses X number of resources and finishes before period Y' is the same as ours.
-When we pass to the optimization problem we just choose the number of resources instead of the time.
-
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Three partition problem}
-
-    \subsection{Hypothesis}
-
-    \begin{itemize}
-     \item assigned maintenances at regular intervals of M.
-     \item We need to assign tasks in between maintenances.
-     \item Each task assignment needs M/3 consecutif periods.
-     \item Each task assignment has a flight hours consumption.
-    \end{itemize}
-
-    \subsection{Reductions}
-
-    \begin{itemize}
-        \item Missions have a fixed duration.
-        \item Maintenances are evenly spaced.
-        \item All maintenances start at the same time. Maybe only one maintenance per Aircraft.
-        % \item The missions require 0 aircraft at each period.
-        % \item Each aircraft can only be used for one mission.
-    \end{itemize}
-
-    \subsection{Limitations}
-
-    \begin{itemize}
-        \item The time at which each task is done matters.
-        %\item There is only one type of resource.
-        % \item All durations are the same.
-    \end{itemize}
-
-    \subsection{Notations}
-
-    \textbf{Sets}
-
-    \noindent{
-    \begin{tabular}{ll}
-        $i \in \II = [\![1, I]\!]$ & Set of aircraft.\\
-        % $k \in \KK_i = [\![1, K_i]\!]$ & Set of maintenances for aircraft $i$.\\ 
-        $j \in J$ & Set of missions.\\ 
-        % $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
-        % $k \in \mathcal{R}$ & Types of resources that maintenance can take. We only have one type.\\
-        % $i = 0$ & Artificial starting node.\\
-        % $i = n+1$ & Artificial ending node.\\
-    \end{tabular}
-    }
-
-    \subsection{Parameters}
-
-    \begin{tabular}{ll}
-    	$h_j$ & Flight hour consumption for each mission.\\
-    	$M$ & Duration of maintenances.\\
-    	$H$ & Flight hours given by maintenance.\\
-        % $p_i$ & Duration of maintenance $i$. In this case it's always p=6.\\
-        % $\mathbf{d}$ & Last period on the horizon.\\
-        % $r_{ik}$ & Resources of type $k$ taken by maintenance $i$. In this case it's always 1.\\
-        % $d^{max}_{ij}$ & Maximum duration between two maintenances $i$ and $j$.\\
-        % $d^{min}_{ij}$ & Minimum duration between two maintenances $i$ and $j$.\\
-        % ${\delta}_{ij}$ & Precedence duration arcs between two maintenances $i$ and $j$ based on $d^{max}$ and $d^{min}$.\\
-        % $R_k$ & total maintenance capacity at each period for resource $k$.\\
-        % $c_k$ & cost per unit of investment on resource $k$.
-        % $\mathbf{{To}}_{i,k}$ & Minimum starting period for maintenance $k$ of resource $i$.\\
-        % $\mathbf{{Ta}}_{i,k}$ & Maximum starting period for maintenance $k$ of resource $i$.\\
-    \end{tabular}
-
-    In theory, with this configuration we have a Three Partition problem. The problem lies that each task has a fixed time. They're not tasks, they're demand.
-
-    % This doesn't make sense because the tasks are already defined. We cannot chose when it executes.
-
-    % I'll continue with the idea nonetheless.
-    % Imagine we preassign the aircraft a group of tasks. We need to decide at which moment we want to assign the tasks in the predefined maintenances.
-    % In this way, we need to come up with an assignment of tasks for each cycle that has a total of 1000 flight hours.
-
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Three partition problem v2}
-
-    Inspired by \cite{Caprara2003}.
-
-    \subsection{Hypothesis}
-
-    \begin{itemize}
-     \item all maintenances are already decided for all aircraft and they will happen in month 4.
-     \item all aircraft come from receiving a maintenance so are full in availability and flight hours.
-     \item each maintenance gives C flight hours.
-     \item objective: comply with all the missions in time and arrive with 0 flight hours to the maintenance.
-     \item each mission lasts for 3 months, needs 1 aircraft each month.
-     \item There is no minimal assignment duration for missions.
-    \end{itemize}
+%     \begin{itemize}
+%      \item fixed maintenance duration.
+%      \item min and max durations between maintenances that belong to the same aircraft.
+%      \item each maintenance uses an amount of resource. In this case 1.
+%      \item there is a maintenance capacity to respect at every period.
+%      \item objective: minimize the max number of maintenances.
+%      \item There are two artificial activities representing the start and end of the horizon.
+%      \item maintenance operations consume also resources of the missions.
+%     \end{itemize}
 
@@
 
-    \subsection{Notations}
+%     \subsection{Notations}
 
-    \textbf{Sets}
+%     \textbf{Sets}
 
-    \noindent{
-    \begin{tabular}{ll}
-        $i \in \II = [\![1, M]\!]$ & Set of aircraft fleets. Each fleet has 3 aircraft.\\
-        $j \in \mathcal{J} = [\![1, 3M]\!]$ & Set of missions.\\
-        $t \in \TT = [\![1, 3]\!]$ & Set of periods.\\
-        $p \in \mathcal{P} = [\![1, 3]\!]$ & Set of aircraft per fleet.\\
-        $A_{ip}$ & Aircraft number $p$ of fleet $i$. \\
-    \end{tabular}
-    }
+%     \noindent{
+%     \begin{tabular}{ll}
+%         $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
+%         $(i, j) \in \mathcal{E} = [\![1, v]\!]$ & Set of consecutive (i, j) maintenance operations, i.e. consecutive maintenances belonging to the same aircraft.\\
+%         $k \in \mathcal{R}$ & Types of resources that maintenance can take. We only have one type.\\
+%         $i = 0$ & Artificial starting node.\\
+%         $i = n+1$ & Artificial ending node.\\
+%     \end{tabular}
+%     }
 
-    \vskip 0.3cm
+%     \vskip 0.3cm
 
-    \subsection{Parameters}
+%     The set $\mathcal{A}$ we will construct it by deciding in advance the maintenances $i$ that share the same aircraft.
 
-    \begin{tabular}{ll}
-        $M$ & Number of fleets.\\
-        $3M$ & Number of aircraft and missions.\\
-        $D$ & Months between maintenances. Fixed to three.\\
-        $h_j$ & Flight hour consumption for each mission.\\
-        $C$ & Flight hours given by maintenance.\\
-    \end{tabular}
+%     \subsection{Parameters}
 
-    \subsection{Variables}
+%     \begin{tabular}{ll}
+%         $p_i$ & Duration of maintenance $i$. In this case it's always p=6.\\
+%         $\mathbf{d}$ & Last period on the horizon.\\
+%         $r_{ik}$ & Resources of type $k$ taken by maintenance $i$. In this case it's always 1.\\
+%         $d^{max}_{ij}$ & Maximum duration between two maintenances $i$ and $j$.\\
+%         $d^{min}_{ij}$ & Minimum duration between two maintenances $i$ and $j$.\\
+%         ${\delta}_{ij}$ & Precedence duration arcs between two maintenances $i$ and $j$ based on $d^{max}$ and $d^{min}$.\\
+%         $R_k$ & total maintenance capacity at each period for resource $k$.\\
+%         $c_k$ & cost per unit of investment on resource $k$.
+%         % $\mathbf{{To}}_{i,k}$ & Minimum starting period for maintenance $k$ of resource $i$.\\
+%         % $\mathbf{{Ta}}_{i,k}$ & Maximum starting period for maintenance $k$ of resource $i$.\\
+%     \end{tabular}
 
-    \begin{tabular}{ll}
-        $F_{ij}$ & Mission $j$ is done by fleet $i$. Binary. \\
-        $a_{ipjt}$ & Mission $j$ is done by aircraft number $p$ from fleet $i$ in period $t$ (auxiliary). Binary. \\
-    \end{tabular}
+%     \subsection{Variables}
 
-    \subsection{Model}
+%     \begin{tabular}{ll}
+%         $S_i$ & the start of maintenance operation $i$.\\
+%         $\mathcal{A}(S^{\mathcal{C}}, t)$ & set of maintenance operations active at moment $t$.\\
+%     \end{tabular}
 
-    It's a satisfaction problem where we want to use exactly all the available hours to do missions.
+%     \subsection{Model}
+
+%     The following formulation is taken from \cite{Neumann2000}.
+
+%     \begin{align}
+%         \min f(\mathbf{R}(S)) = \sum_{k \in \mathcal{R}} c_k \max_{t \in \mathcal{T}} r_k(S^{\mathcal{C}}, t)
+%     \end{align}
+
+%     \begin{align}
+%         & S_j - S_i \geq \delta_{ij} & (i, j) \in \mathcal{E}  \label{precedence}\\
+%         & S_0 = 0 &  \label{firstjob}\\
+%         & S_{n+1} \leq \mathbf{d}   \label{lastjob}\\
+%         & S_{i} \in \mathbb{Z}_{\geq 0} & i \in \mathcal{V}  \label{integervars}\\
+%         & r_k(S^{\mathcal{C}}, t) = \sum_{i \in \mathcal{A}(S^{\mathcal{C}}, t)} r_{ik}  \label{consumptiondef}\\
+%         & r_k(S^{\mathcal{C}}, t) \leq R_k \label{consumptionmax}\\
+%     \end{align}
+
+%     (\ref{precedence}) constraints define the relative minimum and maximum starting times of tasks.
+%     (\ref{firstjob}) and (\ref{lastjob}) constraints define the fictitious start and end jobs.
+%     (\ref{integervars}) define the range for the starting time variables: integer positive.
+%     (\ref{consumptiondef}) defines the used quantity for each type of resource depending on the starting times and durations and (\ref{consumptionmax}) defines the maximum amount of resources of each type that can be used at any given time.
+
+%     \subsection{Possible proof}
+
+%     \cite{Blazewicz1983} proved that the following problem is NP-Hard:
+
+%     $$P2 | res111, chain, p_j = 1 / C_{max}$$ is NP-hard via the 3 partition problem.
+
+%     This problem is, from what I understand, the following:
+
+%     \begin{itemize}
+%         \item The number of machines is 2.
+%         \item There is only one type of resource.
+%         \item The resource size is 1.
+%         \item The consumption of the resource by each job is upper bounded by 1.
+%         \item The duration of each job is 1.
+%         \item There is a chain precedence between jobs: a job can have at most one predecessor and one successor.
+%     \end{itemize}
+
+%     Here they use the machine and resource concept.
+This could be understood as there being TWO resource types.
+Not sure how to really come around this.
+Is the same, for example, as doing the following:
+
+%     $$P| res1R1, chain, p_j = 1 / C_{max}$$ Where I have an unlimited number of machines and a quantity R of available resource?
+
+%     Although they minimize the maximum completion time, the decision problem 'is there a planning that uses X number of resources and finishes before period Y' is the same as ours.
+When we pass to the optimization problem we just choose the number of resources instead of the time.
+
+% \clearpage
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \section{Three partition problem}
+
+%     \subsection{Hypothesis}
+
+%     \begin{itemize}
+%      \item assigned maintenances at regular intervals of M.
+%      \item We need to assign tasks in between maintenances.
+%      \item Each task assignment needs M/3 consecutif periods.
+%      \item Each task assignment has a flight hours consumption.
+%     \end{itemize}
+
+%     \subsection{Reductions}
+
+%     \begin{itemize}
+%         \item Missions have a fixed duration.
+%         \item Maintenances are evenly spaced.
+%         \item All maintenances start at the same time. Maybe only one maintenance per Aircraft.
+%         % \item The missions require 0 aircraft at each period.
+%         % \item Each aircraft can only be used for one mission.
+%     \end{itemize}
+
+%     \subsection{Limitations}
+
+%     \begin{itemize}
+%         \item The time at which each task is done matters.
+%         %\item There is only one type of resource.
+%         % \item All durations are the same.
+%     \end{itemize}
+
+%     \subsection{Notations}
+
+%     \textbf{Sets}
+
+%     \noindent{
+%     \begin{tabular}{ll}
+%         $i \in \II = [\![1, I]\!]$ & Set of aircraft.\\
+%         % $k \in \KK_i = [\![1, K_i]\!]$ & Set of maintenances for aircraft $i$.\\ 
+%         $j \in J$ & Set of missions.\\ 
+%         % $i \in \mathcal{V} = [\![1, n]\!]$ & Set of maintenance operations.\\
+%         % $k \in \mathcal{R}$ & Types of resources that maintenance can take. We only have one type.\\
+%         % $i = 0$ & Artificial starting node.\\
+%         % $i = n+1$ & Artificial ending node.\\
+%     \end{tabular}
+%     }
+
+%     \subsection{Parameters}
+
+%     \begin{tabular}{ll}
+%     	$h_j$ & Flight hour consumption for each mission.\\
+%     	$M$ & Duration of maintenances.\\
+%     	$H$ & Flight hours given by maintenance.\\
+%         % $p_i$ & Duration of maintenance $i$. In this case it's always p=6.\\
+%         % $\mathbf{d}$ & Last period on the horizon.\\
+%         % $r_{ik}$ & Resources of type $k$ taken by maintenance $i$. In this case it's always 1.\\
+%         % $d^{max}_{ij}$ & Maximum duration between two maintenances $i$ and $j$.\\
+%         % $d^{min}_{ij}$ & Minimum duration between two maintenances $i$ and $j$.\\
+%         % ${\delta}_{ij}$ & Precedence duration arcs between two maintenances $i$ and $j$ based on $d^{max}$ and $d^{min}$.\\
+%         % $R_k$ & total maintenance capacity at each period for resource $k$.\\
+%         % $c_k$ & cost per unit of investment on resource $k$.
+%         % $\mathbf{{To}}_{i,k}$ & Minimum starting period for maintenance $k$ of resource $i$.\\
+%         % $\mathbf{{Ta}}_{i,k}$ & Maximum starting period for maintenance $k$ of resource $i$.\\
+%     \end{tabular}
+
+%     In theory, with this configuration we have a Three Partition problem. The problem lies that each task has a fixed time. They're not tasks, they're demand.
+
+%     % This doesn't make sense because the tasks are already defined. We cannot chose when it executes.
+
+%     % I'll continue with the idea nonetheless.
+%     % Imagine we preassign the aircraft a group of tasks. We need to decide at which moment we want to assign the tasks in the predefined maintenances.
+%     % In this way, we need to come up with an assignment of tasks for each cycle that has a total of 1000 flight hours.
+
+% \clearpage
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \section{Three partition problem v2}
+
+%     Inspired by \cite{Caprara2003}.
+
+%     \subsection{Hypothesis}
+
+%     \begin{itemize}
+%      \item all maintenances are already decided for all aircraft and they will happen in month 4.
+%      \item all aircraft come from receiving a maintenance so are full in availability and flight hours.
+%      \item each maintenance gives C flight hours.
+%      \item objective: comply with all the missions in time and arrive with 0 flight hours to the maintenance.
+%      \item each mission lasts for 3 months, needs 1 aircraft each month.
+%      \item There is no minimal assignment duration for missions.
+%     \end{itemize}
+
+%     % \subsection{Reductions}
+
+%     % \begin{itemize}
+%     %     \item The missions consume 0 hours of aircraft.
+%     %     %\item The missions require 0 aircraft at each period.
+%     %     \item Each aircraft can only be used for one mission.
+%     % \end{itemize}
+
+%     % \subsection{Limitations}
+
+%     % \begin{itemize}
+%     %     \item Resource usage is always 1.
+%     %     %\item There is only one type of resource.
+%     %     \item All durations are the same.
+%     % \end{itemize}
+
+%     \subsection{Notations}
+
+%     \textbf{Sets}
+
+%     \noindent{
+%     \begin{tabular}{ll}
+%         $i \in \II = [\![1, M]\!]$ & Set of aircraft fleets. Each fleet has 3 aircraft.\\
+%         $j \in \mathcal{J} = [\![1, 3M]\!]$ & Set of missions.\\
+%         $t \in \TT = [\![1, 3]\!]$ & Set of periods.\\
+%         $p \in \mathcal{P} = [\![1, 3]\!]$ & Set of aircraft per fleet.\\
+%         $A_{ip}$ & Aircraft number $p$ of fleet $i$. \\
+%     \end{tabular}
+%     }
+
+%     \vskip 0.3cm
+
+%     \subsection{Parameters}
+
+%     \begin{tabular}{ll}
+%         $M$ & Number of fleets.\\
+%         $3M$ & Number of aircraft and missions.\\
+%         $D$ & Months between maintenances. Fixed to three.\\
+%         $h_j$ & Flight hour consumption for each mission.\\
+%         $C$ & Flight hours given by maintenance.\\
+%     \end{tabular}
+
+%     \subsection{Variables}
+
+%     \begin{tabular}{ll}
+%         $F_{ij}$ & Mission $j$ is done by fleet $i$. Binary. \\
+%         $a_{ipjt}$ & Mission $j$ is done by aircraft number $p$ from fleet $i$ in period $t$ (auxiliary). Binary. \\
+%     \end{tabular}
+
+%     \subsection{Model}
+
+%     It's a satisfaction problem where we want to use exactly all the available hours to do missions.
 We could also calculate the "waste hours just before the maintenance" and then it could be a minimisation problem.
 
-    \begin{align}
-        & \sum_{t \in \mathcal{T}, j \in \mathcal{J}} a_{ipjt} \times h_j = C & (i \in \mathcal{I}, p \in \mathcal{P}) \label{hours}\\
-        & \sum_{i \in \mathcal{I}, p \in \mathcal{P}} a_{ipjt} = 1 & (j \in \mathcal{J}, t \in \mathcal{T}) \label{demand-TPP}\\
-        & F_{ij} \leq \sum_{p \in \mathcal{P}, t \in \mathcal{T}} a_{ipjt} & (i \in \mathcal{I}, j \in \mathcal{J}) \label{auxiliary1}\\
-        & F_{ij} \geq a_{ipjt} & (i \in \mathcal{I}, p \in \mathcal{P}, t \in \mathcal{T}, j \in \mathcal{J}) \label{auxiliary2}\\
-        & \sum_{i \in \mathcal{I}} F_{ij} = 1 &  (j \in \mathcal{J}) \label{auxiliary3}\\
-        & F_ij, a_{ipjt} \geq 0,\, F_{ij}, a_{ipjt} \in \mathbb{B}
-    \end{align}
+%     \begin{align}
+%         & \sum_{t \in \mathcal{T}, j \in \mathcal{J}} a_{ipjt} \times h_j = C & (i \in \mathcal{I}, p \in \mathcal{P}) \label{hours}\\
+%         & \sum_{i \in \mathcal{I}, p \in \mathcal{P}} a_{ipjt} = 1 & (j \in \mathcal{J}, t \in \mathcal{T}) \label{demand-TPP}\\
+%         & F_{ij} \leq \sum_{p \in \mathcal{P}, t \in \mathcal{T}} a_{ipjt} & (i \in \mathcal{I}, j \in \mathcal{J}) \label{auxiliary1}\\
+%         & F_{ij} \geq a_{ipjt} & (i \in \mathcal{I}, p \in \mathcal{P}, t \in \mathcal{T}, j \in \mathcal{J}) \label{auxiliary2}\\
+%         & \sum_{i \in \mathcal{I}} F_{ij} = 1 &  (j \in \mathcal{J}) \label{auxiliary3}\\
+%         & F_ij, a_{ipjt} \geq 0,\, F_{ij}, a_{ipjt} \in \mathbb{B}
+%     \end{align}
 
-    (\ref{hours}) constraints force the number of hours each aircraft uses to be exactly the number of available hours.
-    (\ref{demand-TPP}) constraints force the solution to have exactly 1 aircraft assigned to each mission at each period.
-    (\ref{auxiliary1}) and (\ref{auxiliary2}) constraints tie the fleet assignment with the single aircraft assignment.
-    (\ref{auxiliary3}) controls that each mission is only assigned to one fleet.
+%     (\ref{hours}) constraints force the number of hours each aircraft uses to be exactly the number of available hours.
+%     (\ref{demand-TPP}) constraints force the solution to have exactly 1 aircraft assigned to each mission at each period.
+%     (\ref{auxiliary1}) and (\ref{auxiliary2}) constraints tie the fleet assignment with the single aircraft assignment.
+%     (\ref{auxiliary3}) controls that each mission is only assigned to one fleet.
 
-    \subsection{Possible proof}
+%     \subsection{Possible proof}
 
-    Each fleet is assigned three missions. These missions need to sum the available amount of flight hours (C).
-    Because we need to satisfy the assignment of aircraft each month, each aircraft belonging to the fleet will have a different combination of the three missions (three permutations in a cycle: ABC, CBA, BCA).
+%     Each fleet is assigned three missions. These missions need to sum the available amount of flight hours (C).
+%     Because we need to satisfy the assignment of aircraft each month, each aircraft belonging to the fleet will have a different combination of the three missions (three permutations in a cycle: ABC, CBA, BCA).
 
-    For example if the missions with flight hours (60, 30, 10) are assigned to fleet 1 and the missions with flight hours (50, 40, 10) are assigned to fleet "3M".
+%     For example if the missions with flight hours (60, 30, 10) are assigned to fleet 1 and the missions with flight hours (50, 40, 10) are assigned to fleet "3M".
 In this case C=100.
@@
 
-    $$
-    \begin{bmatrix}
-        A_{11} & 60 & 30 & 10 \\
-        A_{12} & 10 & 60 & 30 \\
-        A_{13} & 30 & 10 & 60 \\\hline
-        \vdots & \vdots & \vdots & \vdots \\\hline
-        A_{n1} & H_i & H_j & H_k \\
-        A_{n2} & H_k & H_i & H_j \\
-        A_{n3} & H_j & H_k & H_i \\\hline
-        \vdots & \vdots & \vdots & \vdots \\\hline
-        A_{3M1} & 50 & 40 & 10  \\
-        A_{3M2} & 10 & 50 & 40  \\
-        A_{3M3} & 40 & 10 & 50  \\\hline
-    \end{bmatrix}
-    $$
+%     $$
+%     \begin{bmatrix}
+%         A_{11} & 60 & 30 & 10 \\
+%         A_{12} & 10 & 60 & 30 \\
+%         A_{13} & 30 & 10 & 60 \\\hline
+%         \vdots & \vdots & \vdots & \vdots \\\hline
+%         A_{n1} & H_i & H_j & H_k \\
+%         A_{n2} & H_k & H_i & H_j \\
+%         A_{n3} & H_j & H_k & H_i \\\hline
+%         \vdots & \vdots & \vdots & \vdots \\\hline
+%         A_{3M1} & 50 & 40 & 10  \\
+%         A_{3M2} & 10 & 50 & 40  \\
+%         A_{3M3} & 40 & 10 & 50  \\\hline
+%     \end{bmatrix}
+%     $$
 
-    We need to comply with having at least one aircraft assigned to each mission per month. Since we have 3M aircraft, we can do this exactly by assigning all slots.
+%     We need to comply with having at least one aircraft assigned to each mission per month. Since we have 3M aircraft, we can do this exactly by assigning all slots.
 
-    The solution to this problem should give a solution to the Three Partition Problem.
+%     The solution to this problem should give a solution to the Three Partition Problem.
 It would just suffice to take the three missions assigned to each fleet (the order is irrelevant) and use this as a solution of M groups (corresponding to M fleets).
@@
 
-    It's true that in the original problem we do not have the concept of fleet... but maybe we can reformulate the definitions to make it even easier.
+%     It's true that in the original problem we do not have the concept of fleet... but maybe we can reformulate the definitions to make it even easier.
 
-    I think we could get the fleet part out if we would guarantee that each aircraft has an "alldifferent" assignment of three missions.
+%     I think we could get the fleet part out if we would guarantee that each aircraft has an "alldifferent" assignment of three missions.
 This would guarantee that the solution of the problem would still be equivalent to the fleet one.
 
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \clearpage
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{3 Dimensional matching}
+% \section{3 Dimensional matching}
 
-Here I want to prove that... I need three dimensions. Resource-months, Mission-months and cycles.
+% Here I want to prove that... I need three dimensions. Resource-months, Mission-months and cycles.
 
-I want to decide if a task-month is assigned to a resource-month inside the cycle.
+% I want to decide if a task-month is assigned to a resource-month inside the cycle.
 
-Possibilities are constrained by saying.
-Every misision-month needs to be covered.
-Every resource-month needs to be covered.
+% Possibilities are constrained by saying.
+% Every misision-month needs to be covered.
+% Every resource-month needs to be covered.
 
-\clearpage
+% \clearpage
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{other notes}
+% \section{other notes}
 
-If I take the shifts models that were polynomial...
+% If I take the shifts models that were polynomial...
 given that they become hard when adding limits on the number of shifts per planning horizon and if I have flight hours that vary between "shifts"; this means that I can traduce the flight hours in this...
@@
 
-\clearpage
+% \clearpage
 
@@
 Slightly abusing the notation, we can use the elements in these three sets indistinctly.
-Each task $j$ has $T_j = \{j=t\}$ that requires $R_{j} = b_{j=k}$ resources.
-This implies that $J_t = \{j=t\}$, $|J_t|=1$ and $D_t = b_{t=k}$.
+Each task $j$ has onyl one active period, or $T_j = \{j\}$ and requires $R_{j} = b_{j}$ resources.
+This implies that there is only one task active at each period, or $J_t = \{j\}$, $|J_t|=1$ and the quantity of aircraft being used at each period is $D_t = b_{t}$.
 
-    For each worker $e$ we have a corresponding resource $i$.
-    Since all resources can cover the demand, we assume $O_i = \{j \in \mathcal{J}\}$.
-    All resources start have just exited a maintenance in the previous period to the start of the planning horizon. There's no previously planned maintenances, so $N_t = 0$.
+    For each worker $e$ we create a corresponding resource $i$.
+    Since all resources can cover the demand, we assume the set of compatible tasks for each resource is the set of all tasks, or $O_i = \{j \in \mathcal{J}\}$.
+    All resources have just exited a maintenance in the previous period to the start of the planning horizon. There's no previously planned maintenances, so $N_t = 0$.
     The duration of the maintenance will be equal to $M = \underline{D}^{off}$.
@@
 
-    \cite{Smet2015} calls it "shift minimization personnel task scheduling problem. When it's not related to workers and in a more general way, it's usually called "Fixed Interval Scheduling Problem".
+    \cite{Smet2015} calls it "shift minimization personnel task scheduling problem".
+When it's not related to workers and in a more general way, it's usually called "Fixed Interval Scheduling Problem".
 
@@
 Both $T_e$ and $E_p$ are defined based on qualifications, time windows of tasks and availabilities of employees.
+
+    The objective is to minimize the required number of employees needed to perform all tasks.
 
@@
 
-    For each employee $e \in E$, we create an analogous resource $i \in I$. We will $e$ and $i$ indistinctly.
+    For each employee $e \in E$, we create an analogous resource $i \in I$. We will use $e$ and $i$ indistinctly.
 
-    For each task $p \in P$, we create a $j \in J$ task in our problem with duration and minimal assignment time equal to $A^{min}_j = u_j$. We will $p$ and $j$ indistinctly.
+    For each task $p \in P$, we create a $j \in J$ task in our problem with minimal assignment time equal to the duration $A^{min}_j = u_j$. We will use $p$ and $j$ indistinctly.
     Start times and finish times define the moment when the task is active: $\mathcal{T}_j = t \in \{s_j ...
@@
     For maintenances, we can just take them out of the equation completely. For this we just need to put: $E_{min} = E = M = N_t = 0$ and $Ret^{Init}_{i} = |\mathcal{T}| + 1$.
+
+    The objective function is reduced to a constant, implying the problem becomes a feasibility problem instead of an optimization problem.
 
@@
             $\mathcal{T}_j$     &  time periods in which task $j=p$ is active. & $t \in \{s_j ... f_j\}$ \\
-            $\mathcal{J}_t$     &  tasks $j=p \in \mathcal{J}$ to be realized in period $t$. & $j | t \in \{s_j ... f_j\}$ \\
+            $\mathcal{J}_t$     &  tasks $j \in \mathcal{J}$ to be realized in period $t$. & $j | t \in \{s_j ... f_j\}$ \\
             $\mathcal{I}_j$     &  resources that can be assigned to task $j=p$. & $E_j$\\
@@
             $D_t$             & number of resources to be assigned in total in period $t$. & $ |\mathcal{J}_t| $\\
-            $A^{min}_j$       & minimum number of consecutive periods for task $j=p$. & $ u_p $ \\
+            $A^{min}_j$       & minimum number of consecutive periods for task $j=p$. & $ u_j $ \\
             $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start. & $ |\mathcal{T}| + 1 $ \\

### 2018-08-21 latex/notes/model/model.tex
updated model formulation and complexity analysis

@@
         & rut_{it} \in [0,H]
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\               
+                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}              
         % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}

### 2018-08-22 R/presentations/CLAIO_presentation/presentation.Rmd
Added files for CLAIO presentation.

--- 2018-07-10 R/presentations/MOSIM_presentation/presentation.Rmd
@@
 institute: "ISAE SUPAERO"
-date: "June 27, 2018"
+date: "September 24, 2018"
 output: 
@@
 nocite: '@*'
-bibliography: MFMP.bib
+bibliography: MFMP_NRP.bib
 editor_options: 
@@
 ```
-
-<!-- # Introduction -->
-
-<!-- ## Introduction -->
-
-<!-- Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications.  -->
-
-<!-- They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.
--->
 
@@
 
+* FMP: Flight and Maintenance Planning problem.
 * A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
@@
 
-* FMP: Flight and Maintenance Planning problem.
 * In @Cho2011, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
@@
 * In @Chastellux2017, an initial version of the model was done, assigning missions instead of only flight hours.
+* In @Seif2018, a generalization was done in order to deal with different types of maintenances and an heterogeneus fleet.
 
@@
 
-* The present model deals with an heterogeneous fleet, with different standards and capacities.
-* Multiple objectives have been incorporated in order to satisfy the French Air Force's specific needs.
 * Asignments to missions are done instead of assignments of flight hours.
+* Missions need to be assigned for a minimal duration.
 * This model also features several improvements in the MIP modelling that permit greater instance sizes.
+
+## Complexity analysis
+
+Similar to the 'Fixed interval scheduling' problem.
+
+Let $P = {1, ..., n}$ be the set of tasks to be assigned and $E = {1, ..., m}$ the set of employees.
+Each task $p \in P$ has a duration $u_p$ , a start time $s_p$ and a finish time $f_p$ = $s_p + u_p$ .
+Each employee $e$ has a set of tasks $P_e \in P$ that he/she can perform.
+Similarly, for each task $p$, a set $E_p \in E$ exists, which contains all employees that can perform task $p$.
+Both $T_e$ and $E_p$ are defined based on qualifications, time windows of tasks and availabilities of employees.
+
+<!-- TODO: make a design. -->
 
@@
 
-The following instances were tested:
-
-| id | $\|\mathcal{J}\|$ |$\|\mathcal{T}\|$ |  assign |  objective  |  time (s) |   bound |
-|-----|------------------|------------------|---------| ------------| -----------| --------|
-| I_0 |                9 |               11 |     310 |        62.0 |        0.7 |    62.0 |
-| I_1 |                9 |               21 |     650 |        63.0 |       68.7 |    63.0 |
-| I_2 |                9 |               31 |     990 |        63.0 |     3600.1 |    62.0 |
-| I_3 |                9 |               41 |    1249 |        64.0 |     3603.9 |    61.7 |
-| I_4 |               10 |               11 |     530 |        82.0 |        0.9 |    82.0 |
-| I_5 |               10 |               21 |    1070 |        83.0 |      144.0 |    83.0 |
-| I_6 |               10 |               31 |    1610 |        83.0 |     3600.1 |    82.0 |
-| I_7 |               10 |               41 |    2069 |        84.0 |     3609.1 |    81.8 |
-| I_8 |               11 |               11 |    1080 |       139.0 |      530.6 |   139.0 |
-| I_9 |               11 |               21 |    2120 |       149.0 |     3600.0 |   139.9 |
-
 ## Sample solution progress
-
-The analysis of the solution progress for all instances shows that:
-
-* Initial solutions are quite far from the linear relaxation.
-* Branching does not improve significantly the lower bound.
-
-\pause
-```{r out.width="150pt", echo=FALSE}
-knitr::include_graphics("../img/progress.png.png" %>% paste0(path_to_R_dir, .))
-```
-
-*(Progress diagram for instance id=I_7).*
-
-## Multi-objective (1)
-
-Comparison of two objectives: max availability or maintenance capacity.
-Different weights were used for the two objectives: $W_1, W_2 \in [0..1]$ with pace of 0.1.
-\pause
-```{r out.width="180pt", echo=FALSE}
-knitr::include_graphics("../img/multiobjective.png" %>% paste0(path_to_R_dir, .))
-```
-
-*(Pareto diagram for instance id=I_5).*
-
-## Multi-objective (2)
-
-* Analysis for all instances: not many Pareto optimal points available (between two and four).
-* The objectives are quite aligned one with the other.
 
@@
 
-## Current steps
-
-* **Linent objectives**: minimized total number of maintenances with a hard maximum on total maintenances per month.
-* **Multiple solvers**: GUROBI, CPLEX, CBC, CPO, CHOCO.
-
 ## Next steps
 
-* **Add useful cuts**: for example limiting maintenance possibilities.
-* **Break symmetries**: clustering candidates, fixing end availabilities.
-* **Add more constraints**: minimal duration of assignments, storage of aircraft.
+* **Add more constraints**: storage of aircraft.
 * **Add robustness**: duration of maintenances, new /different missions, unavailable aircraft.

### 2018-08-22 R/presentations/CLAIO_presentation/presentation.Rmd
updated presentation for CLAIO

@@
 output: 
+  slidy_presentation:
+    css: ./../../../css/baobab_slidy.css
+    incremental: yes
+    footer: "ISAE-SUPAERO, Université de Toulouse, France"
+    # slide_level: 2
+    # toc: true
   beamer_presentation:
@@
         in_header: header.tex
-  slidy_presentation:
-    css: ./../../css/baobab_MH.css
-    incremental: yes
-    footer: "ISAE-SUPAERO, Université de Toulouse, France"
-    slide_level: 2
-    toc: true
-nocite: '@*'
 bibliography: MFMP_NRP.bib
@@
 source('functions/import_results.R')
+source('diagrams/fixed_interval.R')
 exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
-gantt <- print_solution(exp_directory, max_resources = 30)
+gantt <- print_solution(exp_directory, max_resources = 15, width='100%')
 show_gantt <- knitr::opts_knit$get("rmarkdown.pandoc.to")=="slidy"
+task_gantt <- tasks_gantt()
+task_employees <- tasks_employees()
 ```
@@
 if (show_gantt){
-    htmltools::HTML('<div style="font-size: 20px;width:100%">')
+    # htmltools::HTML('<div style="font-size: 20px;width:100%">')
+    # htmltools::div(style="font-size: 20px;width:100%", gantt)
     gantt
-    htmltools::HTML("</div>")
+    # htmltools::HTML("</div>")
 } else {
@@
 
-## Complexity analysis
+## Complexity analysis (1)
 
-Similar to the 'Fixed interval scheduling' problem.
+Similar to the 'Fixed interval scheduling' problem presented in @Smet2015.
 
@@
 
-<!-- TODO: make a design. -->
+## Complexity analysis (2)
+
+```{r, echo=FALSE, out.width='50%'}
+
+if (show_gantt){
+    htmltools::div(class = "row",
+        htmltools::div(task_gantt, class = "col-md-6")
+        ,htmltools::div(task_employees, class = "col-md-6")
+        )
+} else {
+    knitr::include_graphics(c('../img/tasks-days.png', '../img/task-employees.png') %>% paste0(path_to_R_dir, .))
+    # knitr::include_graphics('../img/task-employees.png' %>% paste0(path_to_R_dir, .))
+}
+
+```
 
 # Model
-
-<!-- ## Model: sets -->
-
-<!-- |Set|Description| -->
-<!-- |-----|--------------------------------------------------------------------| -->
-<!-- |    $\mathcal{I}$       |  resources. | -->
-<!-- |    $\mathcal{T}$       |  time periods inside planning horizon. | -->
-<!-- |    $\mathcal{J}$       |  tasks. | -->
-<!-- |    $\mathcal{T}_j$     |  time periods $t \in \mathcal{T}$ in which task $j$ is active. | -->
-<!-- |    $\mathcal{J}_t$    |  tasks $j \in \mathcal{J}$ to be realized in period $t$. | -->
-<!-- |    $\mathcal{I}_j$     |  resources $i \in \mathcal{I}$ that can be assigned to task $j$. | -->
-<!-- |    $\mathcal{O}_i$     |  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. | -->
-<!-- |    $\mathcal{T}^{s}_t$ |  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. | -->
-
-<!-- ## Model: main parameters -->
-
-<!-- |Parameters|Description| -->
-<!-- |--------------|-------------------------------------------------| -->
-<!-- | $H_j$ |  amount of **resource time required** by task $j$. | -->
-<!-- | $R_j$ |  number of **resources required** by task $j$. | -->
-<!-- | $M$   |  **maintenance duration** in number of periods. | -->
-<!-- | $E$   |  remaining **elapsed time** after a maintenance. | -->
-<!-- | $H$   |  remaining **usage time** after a maintenance. | -->
-
-<!-- ## Model: more parameters -->
-
-<!-- |Parameters|Description| -->
-<!-- |--------------|----------------------------------------------------------------------| -->
-<!-- | $N_t$              | number of resources in already-planned maintenances in period $t$ at the beginning | of the planning ho |izon.| -->
-<!-- | $D_t$              | number of resources to be assigned in total in period $t$. | -->
-<!-- | $Rut^{Init}_{i}$   | remaining usage time for resource $i$ at the start of the planning horizon. | -->
-<!-- | $Ret^{Init}_{i}$   | remaining elapsed time for resource $i$ at the start of the planning horizon. | -->
-<!-- | $Ret^{Init}_{sum}$ | sum of remaining elapsed times at the start of the planning horizon. | -->
-<!-- | $Rut^{Init}_{sum}$ | sum of remaining elapsed time at the start of the planning horizon. | -->
-
-<!-- ## Model: variables -->
-
-<!-- | Variable   |  Description | -->
-<!-- |-----------|------------------------------------------------------- | -->
-<!-- | $a_{jti}$   |  =1 if **task** $j \in J$ in period $t \in \mathcal{T}_j$ **is realized with resource** $i \in \mathcal{I}_j$, 0 otherwise. | -->
-<!-- | $m_{it}$    |  =1 if resource $i \in I$ **starts a maintenance** operation in period $t \in \mathcal{T}$, 0 otherwise. | -->
-<!-- | $rut_{it}$  |  **remaining usage time** (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. | -->
-<!-- | $ret_{it}$  |  **remaining elapsed time** (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. | -->
-<!-- | $u_{max}$   |  maximal number (integer) of **unavailable resources** in any period. | -->
-<!-- | $m_{max}$   |  maximal number (integer) of **resources in maintenance** in any period. | -->
-
-<!-- ## Model: main constraints -->
-
-<!-- Requirements per task: -->
-<!-- ```{r out.width="250px", echo=FALSE} -->
-<!-- knitr::include_graphics("./../../img/tasks_gantt.png") -->
-<!-- ``` -->
-<!-- \pause -->
-
-<!-- Max maintenances: -->
-<!-- ```{r out.width="150px", echo=FALSE} -->
-<!-- knitr::include_graphics("./../../img/num-maintenances.png") -->
-<!-- ``` -->
-
-<!-- ## Model: main constraints -->
 
@@
 # References
+
 ## References / Questions?

### 2018-08-22 R/diagrams/fixed_interval.R

@@ -1,0 +1,78 @@
+library(tidyverse)
+library(data.table)
+library(timevis)
+library(lubridate)
+library(visNetwork)
+library(magrittr)
+library(RColorBrewer)
+
+tasks_gantt <- function(...){
+    start_month <- as.Date('2017-10-01')
+    
+    data <- 
+        data.table(
+            Phase=c(1, 1, 2, 3, 4, 5, 6, 7)
+            ,Libelle=LETTERS[1:8]
+        ) %>% 
+        mutate(
+            month_start=sample(1:5, 8, replace=TRUE)
+            ,month_end= month_start + sample(1:4, 8, replace=TRUE)
+        ) %>% 
+        mutate(start= start_month %m+% months(month_start-1),
+               end= start_month %m+% months(month_end),
+               group= Phase,
+               id= c(1:nrow(.)),
+               content = Libelle,
+               color = RColorBrewer::brewer.pal(8, "YlOrRd")[rep_len(1:8, n())],
+               style= sprintf("background-color:%s;border-color:%s;font-size: 15px", color, color),
+        )
+    config <- list(
+        editable = TRUE,
+        align = "center",
+        orientation = "top",
+        snap = NULL,
+        margin = 0,
+        zoomable= FALSE
+    )
+    
+    groups <- data %>% distinct(group) %>% rename(id= group) %>% mutate(content= id)
+    timevis(data, groups= groups, options= config)
+}
+
+tasks_employees <- function(){
+    tasks <- LETTERS[1:8] %>% set_names(.)
+    employees <- 1:10
+    
+    nodes <- 
+        data.table(label=c(tasks, employees)) %>% 
+        mutate(
+            id = row_number(),
+            value = c(
+                rep(20, length(tasks))
+                ,rep(2, length(employees))
+            ),
+            color = c(
+                RColorBrewer::brewer.pal(8, "YlOrRd")
+                ,rep("gray",length(employees))
+                )
+            )
+
+    edges_data <- 
+        lapply(tasks, function(x) {
+            sample(employees, sample(2:4, 1)) %>% data.table(employee=.) 
+            }) %>% 
+        bind_rows(.id='task') %>% 
+        mutate(employee = employee %>% as.character())
+    
+    edges <- 
+        edges_data %>% 
+        inner_join(nodes, by=c('employee'='label')) %>% 
+        inner_join(nodes, by=c('task'='label')) %>% 
[…]

### 2018-08-27 latex/notes/model/model.tex
small changes in model

@@
         % $W_2$             & weight of the second objective in the objective function. \\
-        % $N_t$             & number of resources in already-planned maintenances in period $t$ at the beginning of the planning horizon.\\
+        $N_t$             & number of resources in already-planned maintenances in period $t$.\\
         % $D_t$             & number of resources to be assigned in total in period $t$. \\
@@
         $a_{jti}$   &  =1 if task $j \in J$ in period $t \in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
-        $a^s_{jti}$ &  =1 if resource $i$ starts a new assignment to task $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-i)} = 0$. \\  
+        $a^s_{jti}$ &  =1 if resource $i$ starts a new assignment to task $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-1)i} = 0$. \\  
         $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
@@
         % maximum capacity1:
-        % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
-        %   & t \in \mathcal{T} \label{eq:capacity1}\\
-        & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq M^{max} \label{eq:capacity1}\\
+        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq M^{max}
+          & t \in \mathcal{T} \label{eq:capacity1}\\
+        % & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq M^{max} \label{eq:capacity1}\\
         % availability
@@
     The minimum and maximum elapsed time are defined by (\ref{eq:ret_min}) and (\ref{eq:ret_max}) respectively.
-Constraints (\ref{eq:ret_min}) guarantee that if a maintenance is done in some period $t$, we know that it cannot be done in the immediately consecutive $t' \in \mathcal{T}^{m}_t$ periods.
+Constraints (\ref{eq:ret_min}) guarantee that if a maintenance is done in some period $t$, we know that another one cannot be done in the immediately consecutive $t' \in \mathcal{T}^{m}_t$ periods.
 Constraints (\ref{eq:ret_max}) guarantee that, if we have a maintenance in period $t$, we need to start at least one maintenance among periods $t' \in \mathcal{T}^{M}_t$.
 Constraints (\ref{eq:ret_min_init}) and (\ref{eq:ret_max_init}) control the minimum and maximum remaining elapsed time at the beginning of the planning period, respectively.
-They follow the same rules as the previous constraints for the particular case of the start of the planning horizon.
+They follow the same logic as constraints (\ref{eq:ret_min}) and (\ref{eq:ret_max}), respectively, for the particular case of the start of the planning horizon.
 

### 2018-08-27 latex/notes/model/model.tex
modifications on model after meeting Olga and Nicolas

--- 2018-08-21 latex/notes/model/model.tex
@@
 
-  \subsection{Sets}
+  \subsection{Basic sets}
 
@@
         $\mathcal{K}$     &  cluster of resources that share the same capacities to do tasks. \\
-        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
-        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
-        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
-        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. \\
-        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
-        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
-        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + E^{m}-1\}}\}$. \\
-        $\mathcal{T}^{M}_t$ &  time periods $t' \leq |\mathcal{T}| - E^{M}$ such that $t' \in \{ t + E^{m}-1 , ...,  t + E^{M}-1 \}$. \\
-        $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Ret^{Init}_{i} - E^M + E^m \}}\}$. \\
-        $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Ret^{Init}_{i} - E^M + E^m \}} , ...,  Ret^{Init}_{i} \}$. \\
-        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
     \end{tabular}
@@
         $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
-        $A^{init}_{jti}$  & =1 if resource $i \in \mathcal{I}$ has task $j$ pre-assigned in period $t$. \\
         % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
         % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
+    \end{tabular}
+
+    \subsection{Parametric sets}
+
+        \begin{tabular}{p{5mm}p{170mm}}
+        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
+        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
+        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
+        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. \\
+        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
+        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
+        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
+        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + E^{m}-1\}}\}$. \\
+        $\mathcal{T}^{M}_t$ &  time periods $t' \leq |\mathcal{T}| - E^{M}$ such that $t' \in \{ t + E^{m}-1 , ...,  t + E^{M}-1 \}$. \\
+        $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Ret^{Init}_{i} - E^M + E^m \}}\}$. \\
+        $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Ret^{Init}_{i} - E^M + E^m \}} , ...,  Ret^{Init}_{i} \}$. \\
+        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
+        $\mathcal{A}^{init}_{jt}$  & resources $i \in \mathcal{I}$ that have task $j$ pre-assigned in period $t$. \\
     \end{tabular}
@@
         % if start task => assign start
-        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} - A^{init}_{j(t-1)i}
+        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} - 1 \!1_{i \in \mathcal{A}^{init}_{j(t-1)}}
                 & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start1} \\
       % # if we start a task, we need at least X periods of tasks assignments
-        & a^s_{jt'i} \leq a_{jti}
-        & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j, t' \in \mathcal{T}^{MT}_{jt} \label{eq:start2}
+        % & a^s_{jt'i} \leq a_{jti}
+        % & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j, t' \in \mathcal{T}^{MT}_{jt} \label{eq:start2} \\
+        % better constraint:
+        & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
+        & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
     \end{align}
 
+    TODO: explain constraint (\ref{eq:start3}).
+
     Constraints (\ref{eq:start1}) control the definition of starting a consecutive assignment: the variable is active if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this same task in the previous period ($t-1$).
-The previous period can include the period before the start of the planning horizon ($A^{init}_{j(t-1)i}$ parameter).
-    Constraints (\ref{eq:start2}) control the minimum duration of a consecutive task assignment.
+The previous period can include the period before the start of the planning horizon ($\mathcal{A}^{init}_{j(t-1)}$ set).
+    Constraints (\ref{eq:start3}) control the minimum duration of a consecutive task assignment.
 If an assignment to task $j$ is started in period $t$, it needs to continue for the following $t' \in \mathcal{T}^{MT}_{jt}$ periods.

### 2018-08-28 latex/notes/model/model.tex

@@
 % \usepackage{aeguill}
-\usepackage[frenchb,english]{babel}
+\usepackage[english]{babel}
 \usepackage{caption}
@@
 
-  % TODO: previously assigned missions.
-
     \begin{tabular}{p{8mm}p{167mm}}
@@
         $N_t$             & number of resources in already-planned maintenances in period $t$.\\
+        $Nk_{kt}$         & number of resources in cluster $k$ in already-planned maintenances in period $t$.\\
         % $D_t$             & number of resources to be assigned in total in period $t$. \\
         $AK_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
+        $HK_{kt}$         & minimum number of total remaining usage time for cluster $k$ at period $t$.\\
         $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
@@
         $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
-        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. \\
+        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. One resource can belong to more than one cluster. \\
         $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
@@
         $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
-        $\mathcal{A}^{init}_{jt}$  & resources $i \in \mathcal{I}$ that have task $j$ pre-assigned in period $t$. \\
+        $\mathcal{A}^{init}_{j}$  & resources $i \in \mathcal{I}$ that have task $j$ pre-assigned in the previous period to the start of the horizon. \\
     \end{tabular}
@@
         % if start task => assign start
-        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} - 1 \!1_{i \in \mathcal{A}^{init}_{j(t-1)}}
-                & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start1} \\
+        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
+                & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\
+        & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_{j}}
+                & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\        
       % # if we start a task, we need at least X periods of tasks assignments
@@
 
-    TODO: explain constraint (\ref{eq:start3}).
+    Constraints (\ref{eq:start1}) control the definition of starting a consecutive assignment: the variable is active if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this same task in the previous period ($t-1$).
+Constraints (\ref{eq:start2}) are the particular case for the first period on the horizon where, the previous assignments are given.
 
-    Constraints (\ref{eq:start1}) control the definition of starting a consecutive assignment: the variable is active if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this same task in the previous period ($t-1$).
-The previous period can include the period before the start of the planning horizon ($\mathcal{A}^{init}_{j(t-1)}$ set).
     Constraints (\ref{eq:start3}) control the minimum duration of a consecutive task assignment.
 If an assignment to task $j$ is started in period $t$, it needs to continue for the following $t' \in \mathcal{T}^{MT}_{jt}$ periods.
+They are a stronger version of the constraint $ a^s_{jt'i} \leq a_{jti}$.
 
     \begin{align}
-       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} \leq AK_{kt}
-        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster}
+       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + Nk_{kt} \leq AK_{kt}
+        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\
+       & \sum_{i \in \mathcal{I}_k} rut_{it} \geq HK_{kt}
+        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster}
     \end{align}
 
-    Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability for each cluster $k$ of resources.
+    Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability of resources for each cluster $k$.
 A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task.
+Constraints (\ref{eq:avalaibility-hours-cluster}) guarantee that a minimum availability in terms of usage houres for each cluster $k$.
 
@@
         % remaining used time
-         & rut_{it} \leq rut_{i(t-1)} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
+         & rut_{it} \leq rut_{i(t-1)} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
         & rut_{i0} = Rut^{Init}_i
                & i \in \mathcal{I} \label{eq:rut_initial}\\
-        & rut_{it} \geq H m_{it}
-                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
+        & rut_{it} \geq H m_{it'}
+                & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\ 
         & rut_{it} \in [0,H]

### 2018-08-30 R/presentations/CLAIO_presentation/presentation.Rmd

@@
 if (show_gantt){
-    # htmltools::HTML('<div style="font-size: 20px;width:100%">')
-    # htmltools::div(style="font-size: 20px;width:100%", gantt)
     gantt
-    # htmltools::HTML("</div>")
 } else {
@@
 } else {
-    knitr::include_graphics(c('../img/tasks-days.png', '../img/task-employees.png') %>% paste0(path_to_R_dir, .))
+    c('../img/tasks-days.png', '../img/task-employees.png') %>% paste0(path_to_R_dir, .) %>% 
+    knitr::include_graphics
     # knitr::include_graphics('../img/task-employees.png' %>% paste0(path_to_R_dir, .))

### 2018-08-30 latex/notes/model/model.tex
updated model names and simulation document

@@
         $M$               & maintenance duration in number of periods. \\
-        $M^{max}$         & maximum number of simultaneous maintenance operations. \\
+        $C^{max}$         & maximum number of simultaneous maintenance operations. \\
         % $E$               & remaining elapsed time after a maintenance. \\
@@
         $E^{m}$         & minimum number of periods between two consecutive maintenances. \\
-        $H$               & remaining usage time after a maintenance. \\
+        $H^{M}$               & remaining usage time after a maintenance. \\
         % $S$               & Storage duration in number of periods. \\
@@
         $N_t$             & number of resources in already-planned maintenances in period $t$.\\
-        $Nk_{kt}$         & number of resources in cluster $k$ in already-planned maintenances in period $t$.\\
+        $N^{clust}_{kt}$         & number of resources in cluster $k$ in already-planned maintenances in period $t$.\\
         % $D_t$             & number of resources to be assigned in total in period $t$. \\
-        $AK_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
-        $HK_{kt}$         & minimum number of total remaining usage time for cluster $k$ at period $t$.\\
+        $A^{clust}_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
+        $H^{clust}_{kt}$         & minimum number of total remaining usage time for cluster $k$ at period $t$.\\
         $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
@@
     \begin{align}
-        & \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|}
+        & \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H^{M} - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|}
     \end{align}
@@
         % maximum capacity1:
-        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq M^{max}
+        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq C^{max}
           & t \in \mathcal{T} \label{eq:capacity1}\\
-        % & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq M^{max} \label{eq:capacity1}\\
+        % & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq C^{max} \label{eq:capacity1}\\
         % availability
@@
     \begin{align}
-       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + Nk_{kt} \leq AK_{kt}
+       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^{clust}_{kt} \leq A^{clust}_{kt}
         &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\
-       & \sum_{i \in \mathcal{I}_k} rut_{it} \geq HK_{kt}
+       & \sum_{i \in \mathcal{I}_k} rut_{it} \geq H^{clust}_{kt}
         &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster}
@@
         % remaining used time
-         & rut_{it} \leq rut_{i(t-1)} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
+         & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
@@
                & i \in \mathcal{I} \label{eq:rut_initial}\\
-        & rut_{it} \geq H m_{it'}
+        & rut_{it} \geq H^{M} m_{it'}
                 & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\ 
-        & rut_{it} \in [0,H]
+        & rut_{it} \in [0,H^{M}]
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}              

### 2018-09-07 python/package/model.py
updated some notes and keep advancing with iimplementation of the model and data.

@@ -12,33 +12,15 @@ def solve_model(instance, options=None):
     last_period = instance.get_param('end')
     consumption = instance.get_tasks('consumption')
     requirement = instance.get_tasks('num_resource')
-    ret_init = instance.get_initial_state("elapsed")
     rut_init = instance.get_initial_state("used")
     num_resource_maint = aux.fill_dict_with_default(instance.get_total_fixed_maintenances(), l['periods'])
-    num_resource_maint_cluster = aux.dict_to_lendict(instance.get_fixed_maintenances_cluster())
     maint_capacity = instance.get_param('maint_capacity')
     max_usage = instance.get_param('max_used_time')
+    cluster_data = instance.get_cluster_minimums()
+    c_candidates = instance.get_cluster_candidates()
 
     # TODO: fill this:
-    task_previous_res = {v: [] for v in l['tasks']}
     fixed_vars = [(a, v, t) for (a, v, t) in range(0)]
-    # TODO: fill this (up)
-
-    min_percent = 0.10
-    min_value = 1
-
-    c_needs = instance.get_cluster_needs()
-    c_candidates = instance.get_cluster_candidates()
-    c_num_candidates = aux.dict_to_lendict(c_candidates)
-    c_slack = {tup: c_num_candidates[tup[0]] - c_needs[tup] - num_resource_maint_cluster.get(tup, 0)
-               for tup in c_needs}
-    c_min = {(k, t): min(
-        c_slack[(k, t)],
-        max(
-            int(c_num_candidates[k] * min_percent),
-            min_value)
-        ) for (k, t) in c_slack
-    }
 
     # Sometimes we want to force variables to be integer.
     var_type = pl.LpContinuous
@@ -102,7 +84,9 @@ def solve_model(instance, options=None):
         if t > first_period:
             model += start_T[avt] >= task[avt] - task.get(avt_ant, 0)
         else:
-            model += start_T[avt] >= task[avt] - (a in task_previous_res[v])
+            # we check if we have the assignment in the previous period.
+            # this is stored in the fixed_vars set.
+            model += start_T[avt] >= task[avt] - (a, v, t) in fixed_vars
 
     # if we start a task in at least on earlier period, we need to assign a task
     for a, v, t1, t2 in l['avt']:
@@ -119,20 +103,16 @@ def solve_model(instance, options=None):
     # ##################################
 
     # minimum availability per cluster and period
-    # TODO: check this and add the minimum usage hours per cluster
[…]

### 2018-09-12 latex/notes/model/model.tex
updated model and simulation.
updated some notes and keep advancing with iimplementation of the model and data.
new version of model with minimul consumption

@@
         $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
+        $u^{min}$       & Absolut minimum usage time of resource when resources is not in maintenance.\\
         % $aut_i$           & maximal absolute usage time for resource $i$. \\
@@
         $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
+        $u_{it}$    &  usage time consumed for resource $i \in I$ during period $t \in \mathcal{T}$. \\        
         % $s_{it}$    &  =1 if resource $i \in I$ starts a storage operation in period $t \in \mathcal{T}$, 0 otherwise. \\
@@
   \subsection{Constraints}
+
+  % TODO: say something about fixed tasks at the beginning.
+  % TODO: say something about fixed maintenances at the beginning.
 
@@
 A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task.
-Constraints (\ref{eq:avalaibility-hours-cluster}) guarantee that a minimum availability in terms of usage houres for each cluster $k$.
+Constraints (\ref{eq:avalaibility-hours-cluster}) guarantee that a minimum availability in terms of usage hours for each cluster $k$.
 
@@
         % remaining used time
-         & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
+         & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage}\\
+         & u_{it} \geq u^{min} (1 - \sum_{t' \in \mathcal{T}^{s}_t} m_{it'})
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage}\\
+         & u_{it} \in [0, \max_j{\{H_j\}}]
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage_lower}\\
+         & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - u_{it}
             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
@@
         % These constraints calculate the balances of hours for each resource.
+    The usage time per resource and period is calculated in (\ref{eq:usage})-(\ref{eq:usage_lower}).
 The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}).

### 2018-09-17 R/presentations/CLAIO_presentation/presentation.Rmd
updated CLAIO presentation, graphs, simulated data in R and shiny app.

@@
     fig_height: 4
+    # citation_package: natbib
     # footer: "ISAE-SUPAERO, Université de Toulouse, France"
@@
 bibliography: MFMP_NRP.bib
+# nocite: '@*'
 editor_options: 
@@
 Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
-* Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
+* Each task requires a certain number $R_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $H_j$ in each period and needs to be assigned for a minimum of $MT_j$ consecutive periods if assigned.
 * Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
-* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H$ units.
+* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H^M$ units.
 
@@
 
-* In @Cho2011, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
+* In @Cho2011, US Army aircraft were assigned daily operations over a year to minimize the number of maintenances.
 * In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
 * In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
-* In @Chastellux2017, an initial version of the model was done, assigning missions instead of only flight hours.
-* In @Seif2018, a generalization was done in order to deal with different types of maintenances and an heterogeneus fleet.
+* In @Seif2018, a generalization of @Kozanidis2008 was done in order to deal with different types of maintenances and an heterogeneus fleet.
 
@@
 
-## Complexity analysis (1)
+# Complexity analysis
 
-Similar to the 'Fixed interval scheduling' problem presented in @Smet2015.
+## Complexity analysis
 
-Let $P = {1, ..., n}$ be the set of tasks to be assigned and $E = {1, ..., m}$ the set of employees.
-Each task $p \in P$ has a duration $u_p$ , a start time $s_p$ and a finish time $f_p$ = $s_p + u_p$ .
-Each employee $e$ has a set of tasks $P_e \in P$ that he/she can perform.
-Similarly, for each task $p$, a set $E_p \in E$ exists, which contains all employees that can perform task $p$.
-Both $T_e$ and $E_p$ are defined based on qualifications, time windows of tasks and availabilities of employees.
+<div class="incremental">
 
-## Complexity analysis (2)
+Reduction to the 'Fixed interval scheduling' in @Smet2015.
 
-```{r, echo=FALSE, out.width='50%'}
+* $P$ tasks with duration $u_p$ and start time $s_p$.
+* $E$ employees with a set of tasks $P_e \in P$ that each can perform.
+* Minimize the number of required employees to do all tasks.
+
+\pause
+
+```{r, echo=FALSE, out.width='45%'}
 
@@
     htmltools::div(class = "row",
-        htmltools::div(task_gantt, class = "col-md-6")
-        ,htmltools::div(task_employees, class = "col-md-6")
+        htmltools::div(task_employees, class = "col-md-6")
+        ,htmltools::div(task_gantt, class = "col-md-6")
         )
 } else {
-    c('../img/tasks-days.png', '../img/task-employees.png') %>% paste0(path_to_R_dir, .) %>% 
-    knitr::include_graphics
+    c('../img/tasks-days.png', '../img/task-employees.png') %>% 
+        paste0(path_to_R_dir, .) %>%
+        knitr::include_graphics()
     # knitr::include_graphics('../img/task-employees.png' %>% paste0(path_to_R_dir, .))
@@
 
+</div>
+
 # Model
 
-## Model: objective function
+## Model: main constraints
 
@@
 \begin{align}
-    & \text{Min}\; W_1 m_{max} + W_2 u_{max}
+& \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H^{M} - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|}
 \end{align}
+
+The objective is to minimize the total number of maintenances and, at the same time, maximize the end status of the resources.
 
@@
 
-The two components of the objective function: the max number of maintenance and the max number of unavailable aircraft.
 \begin{align}
-    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
-    &t \in \mathcal{T} \\
-    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max}
-    &t \in \mathcal{T}
+    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq C^{max}
+      & t \in \mathcal{T} \label{eq:capacity1}\\
+    & \sum_{i \in \mathcal{I}_j} a_{jti} \geq R_j
+            & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
+    & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
+            & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
 \end{align}
+
+Mission's needs and incompatibility of several tasks or maintenance in the same period.
 
@@
 
-## Model: main constraints
+## Model: minimum assignment times
 
-Mission's needs and incompatibility of several tasks or maintenance in the same period.
+<!-- <div class="incremental"> -->
+
 \begin{align}
-&\sum_{i \in \mathcal{I}_j} a_{jti} = R_j
-&j \in \mathcal{J}, t \in \mathcal{T}_j\\
-&\sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1
-& t \in \mathcal{T}, i \in \mathcal{I}
+    & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
+            & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\
+    & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_{j}}
+            & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\        
+    & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
+    & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
 \end{align}
 
-<!-- ## Model: flow constraints -->
+Each task needs a minimum continuous duration at each assignment.
 
-<!-- Visualization of accumulated time constraints for one resource. -->
+<!-- \pause -->
 
-<!-- ```{r out.width="250px", echo=FALSE} -->
-<!-- # img <- readPNG("./../../img/balance2.png") -->
-<!-- #  grid.raster(img) -->
-<!-- knitr::include_graphics("./../../img/balance2.png") -->
-<!-- ``` -->
+## Model: maintenances
+
+\begin{align}
+    & m_{it'} + m_{it} \leq 1
+      & t \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
+    & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
+      & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\
+    & m_{it} = 0
+      & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I}\label{eq:ret_min_init} \\
+    & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
+      & i \in \mathcal{I}\label{eq:ret_max_init}
+\end{align}
+
+Maintenances cannot executed too soon or too late with respect to the previous one.
+
+<!-- </div> -->
 
@@
 
-Time usage of an aircraft depending on the assignment to a task.
-<!-- <div class="incremental" style="margin-left: 4em; position: relative">  -->
-<!--   <div style="position: absolute; left: 0; top: 0;background-color:white;width:100%"> -->
+<!-- Time usage of an aircraft depending on the assignment to a task and remaining usage time per aircraft and period. -->
+
 \begin{align}
-& rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
+ & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
+    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage1}\\
+ & u_{it} \geq u^{min} (1 - \sum_{t' \in \mathcal{T}^{s}_t} m_{it'})
+    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage2}\\
+ & u_{it} \in [0, \max_j{\{H_j\}}]
+    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage_lower}\\
+ & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - u_{it}
+    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
  & rut_{i0} = Rut^{Init}_i
-        & i \in \mathcal{I}\\
-& rut_{it} \geq H m_{it}
-        & t \in \mathcal{T}, i \in \mathcal{I}\\
-& rut_{it} \in [0,H]
-        & t \in \mathcal{T}, i \in \mathcal{I} \\
-& \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
-        & t = |\mathcal{T}|
+       & i \in \mathcal{I} \label{eq:rut_initial}\\
+ & rut_{it} \geq H^{M} m_{it'}
+        & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\
+ & rut_{it} \in [0,H^{M}]
+        & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
 \end{align}
 
+## Model: availability
+
+\begin{align}
+   & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^{clust}_{kt} \leq A^{clust}_{kt}
+    &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\
+   & \sum_{i \in \mathcal{I}_k} rut_{it} \geq H^{clust}_{kt}
+    &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster}
+\end{align}
+
+Each cluster of aircraft needs a minimum level of availability in number and usage hours.
+
+<!-- # Results -->
+
+<!-- ## Experiments -->
+
+<!-- Instances have been created with the following characteristics: -->
+
+<!-- <div class="incremental"> -->
+
+<!-- * 60 - 200 resources. -->
+<!-- * 90 periods. -->
+<!-- * Tasks: 3 - 9  -->
+<!--     * 12 - 36 periods long -->
+<!--     * 2 - 5 resources needed -->
+
 <!-- </div> -->
-  <!-- <div style="position: absolute; left: 0; top: 0;background-color:white;width:100%"> -->
-
-# Results
-
-## Experiments
-
-## Sample solution progress
-
-# Current and next steps
-
-## Next steps
-
-* **Add more constraints**: storage of aircraft.
-* **Add robustness**: duration of maintenances, new /different missions, unavailable aircraft.
 
@@
 <!-- \bibliographystyle{dcu} -->
-<!-- \bibliography{./../biblio/MFMP} -->
+<!-- \bibliography{MFMP_NRP.bib} -->

### 2018-09-17 R/appShiny/server.R

@@ -6,9 +6,7 @@ library(lubridate)
 library(data.table)
 
 shinyServer(function(input, output, clientData, session) {
-    # browser()
     dir_pre <- "../"
-    # dir_pre <- "./"
     sources <-
         c("functions/import_results.R",
           "functions/params.R",

### 2018-09-17 R/scripts/graphs.R

@@ -3,9 +3,9 @@ source('functions/import_results.R')
 source('functions/exampleDT.R')
 source('functions/gantt_tasks.R')
 
-exp_directory = PATHS[['experiments']] %>% paste0('201809121748/')
+exp_directory = PATHS[['experiments']] %>% paste0('201809131542/')
 print_solution(exp_directory, width='100%')
-print_tasks(exp_directory, max_tasks = 8)
+print_tasks(exp_directory)
 
 PATHS[['experiments']] %>% paste0('201802061201/')
 

### 2018-09-18 latex/notes/model/model.tex
changes to model and simulation document.

@@
         % min assignments:
-        & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
+        & \sum_{i \in \mathcal{I}_j} a_{jti} \geq R_j
                 & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
@@
          & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage}\\
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage1}\\
          & u_{it} \geq u^{min} (1 - \sum_{t' \in \mathcal{T}^{s}_t} m_{it'})
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage}\\
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage2}\\
          & u_{it} \in [0, \max_j{\{H_j\}}]
@@
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
-        % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}
-        %         & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
-        % & ret_{i0} = Ret^{Init}_i
-        %         & i \in \mathcal{I} \label{eq:ret_initial}\\
-        % & ret_{it} \geq E m_{it}
-        %         & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
-        % & ret_{it} \in [0,E]
-        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
-        % & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
-        %       & t = |\mathcal{T}| \label{eq:min_ret}\\
-        % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
     \end{align}

### 2018-09-19 R/presentations/CLAIO_presentation/presentation.Rmd
changes to simplify model in ppt

@@
 
-Reduction to the 'Fixed interval scheduling' in @Smet2015.
+Reduction to the 'Fixed interval scheduling' in @Smet2015 **proves NP-completeness**.
 
@@
 
-## Model: minimum assignment times
+<!-- ## Model: minimum assignment times -->
 
-<!-- <div class="incremental"> -->
+<!-- <!-- <div class="incremental"> -->
 
-\begin{align}
-    & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
-            & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\
-    & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_{j}}
-            & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\        
-    & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
-    & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
-\end{align}
+<!-- \begin{align} -->
+<!--     & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} -->
+<!--             & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\ -->
+<!--     & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_{j}} -->
+<!--             & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\         -->
+<!--     & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti}  -->
+<!--     & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3} -->
+<!-- \end{align} -->
 
-Each task needs a minimum continuous duration at each assignment.
+<!-- Each task needs a minimum continuous duration at each assignment. -->
 
@@
 
-## Model: flow constraints
+## Model: time consumption
 
@@
 
-## Model: availability
+## Model: other constraints
 
-\begin{align}
-   & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^{clust}_{kt} \leq A^{clust}_{kt}
-    &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\
-   & \sum_{i \in \mathcal{I}_k} rut_{it} \geq H^{clust}_{kt}
-    &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster}
-\end{align}
+* Cluster availability.
+* Minimum duration for task assignment.
 
-Each cluster of aircraft needs a minimum level of availability in number and usage hours.
+<!-- ## Model: availability -->
 
-<!-- # Results -->
+<!-- \begin{align} -->
+<!--    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^{clust}_{kt} \leq A^{clust}_{kt} -->
+<!--     &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\ -->
+<!--    & \sum_{i \in \mathcal{I}_k} rut_{it} \geq H^{clust}_{kt} -->
+<!--     &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster} -->
+<!-- \end{align} -->
 
-<!-- ## Experiments -->
+<!-- Each cluster of aircraft needs a minimum level of availability in number and usage hours. -->
 
-<!-- Instances have been created with the following characteristics: -->
+# Results
+
+## Experiments
+
+Instances have been created with the following characteristics:
 
@@
 
-<!-- * 60 - 200 resources. -->
-<!-- * 90 periods. -->
-<!-- * Tasks: 3 - 9  -->
-<!--     * 12 - 36 periods long -->
-<!--     * 2 - 5 resources needed -->
+* 60 - 200 resources.
+* 90 periods.
+* Tasks: 3 - 9
+    * 12 - 36 periods long
+    * 2 - 5 resources needed
 

### 2018-09-19 R/presentations/CLAIO_presentation/presentation.Rmd
updated CLAIO ppt with statistics on results

@@
 source('functions/import_results.R')
+source('functions/scenario_analysis.R')
 source('diagrams/fixed_interval.R')
-exp_directory = PATHS[['experiments']] %>% paste0('201802141539/')
+all_exp_dir <- PATHS[['results']] %>% paste0('simulated_data/')
+exp_directory <- all_exp_dir %>% paste0('2_task/201809180941/')
 gantt <- print_solution(exp_directory, max_resources = 15, width='100%')
@@
 task_employees <- tasks_employees()
+scen_means <- scenario_mean(all_exp_dir)
+infeasible <- scenario_get_gaps(all_exp_dir) %>% lapply(function(scn){
+    a <- length(scn)
+    scn %>% is.na %>% sum %>% divide_by(a/100)
+} )
+
 ```
@@
 
-<!-- ## Model: minimum assignment times -->
-
-<!-- <!-- <div class="incremental"> -->
-
-<!-- \begin{align} -->
-<!--     & a^s_{jti} \geq a_{jti} - a_{j(t-1)i} -->
-<!--             & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\ -->
-<!--     & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_{j}} -->
-<!--             & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\         -->
-<!--     & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti}  -->
-<!--     & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3} -->
-<!-- \end{align} -->
-
-<!-- Each task needs a minimum continuous duration at each assignment. -->
-
-<!-- \pause -->
-
 ## Model: maintenances
@@
 
-# Results
+# Results and perspectives
 
@@
 
-Instances have been created with the following characteristics:
-
-<!-- <div class="incremental"> -->
+Instances have been generated with the following main characteristics:
 
@@
 * 90 periods.
-* Tasks: 3 - 9
+* 3 - 9 tasks:
     * 12 - 36 periods long
@@
 
-<!-- </div> -->
+\pause
+
+* Small instances have **30.000 variables** and **20.000 constraints**.
+* Medium instances have **80.000 variables** and **50.000 constraints**.
+* Runs of 30 minutes have been done on 30 randomly generated cases per scenario and they return a gap between 20% and 40% if a solution exists.
+* CPLEX was used in an i7, quadcore Ubuntu 18.04 workstation.
+
+</div>
+
+## Statistics
+
+* `r infeasible[1]`% of instances were proved infeasible inmediately.
+* From feasible small instances, a gap of `r scen_means[1]`% was found after 30 min
+* From feasible medium instances, a gap of `r scen_means[2]`% was found after 30 min.
+
+## Perspectives
+
+* More extreme (bigger) scenarios need to be tested: more tasks and more resources.
+* A deeper analysis on the detailed solving process is envisioned. Including extending the solving time.
+* Benchmarking of solvers (gurobi, CBC, choco and CPO) is work in progress. Also: heuristics.
+* The inclusion of new possibilities, such as the planned storage of resources is pending.
+* A stochastic sensibility analysis needs to be done, including taking account of stochasticity in the input data.
 

### 2018-09-19 R/presentations/CLAIO_presentation/presentation.Rmd
second version of CLAIO ppt

@@
 
-## Model: time consumption
+<!-- ## Model: time consumption -->
 
@@
 
-\begin{align}
- & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
-    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage1}\\
- & u_{it} \geq u^{min} (1 - \sum_{t' \in \mathcal{T}^{s}_t} m_{it'})
-    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage2}\\
- & u_{it} \in [0, \max_j{\{H_j\}}]
-    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage_lower}\\
- & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - u_{it}
-    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
- & rut_{i0} = Rut^{Init}_i
-       & i \in \mathcal{I} \label{eq:rut_initial}\\
- & rut_{it} \geq H^{M} m_{it'}
-        & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\
- & rut_{it} \in [0,H^{M}]
-        & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
-\end{align}
+<!-- \begin{align} -->
+<!--  & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j  -->
+<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage1}\\ -->
+<!--  & u_{it} \geq u^{min} (1 - \sum_{t' \in \mathcal{T}^{s}_t} m_{it'}) -->
+<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage2}\\ -->
+<!--  & u_{it} \in [0, \max_j{\{H_j\}}] -->
+<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage_lower}\\ -->
+<!--  & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - u_{it} -->
+<!--     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\ -->
+<!--  & rut_{i0} = Rut^{Init}_i -->
+<!--        & i \in \mathcal{I} \label{eq:rut_initial}\\ -->
+<!--  & rut_{it} \geq H^{M} m_{it'} -->
+<!--         & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\ -->
+<!--  & rut_{it} \in [0,H^{M}] -->
+<!--         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} -->
+<!-- \end{align} -->
 
@@
 
-* Cluster availability.
+* Minimum cluster availability.
 * Minimum duration for task assignment.
+* Hour consumption of resources depending on tasks and maintenances.
 
@@
 
-Instances have been generated with the following main characteristics:
+<div class="incremental">
 
@@
 * Medium instances have **80.000 variables** and **50.000 constraints**.
-* Runs of 30 minutes have been done on 30 randomly generated cases per scenario and they return a gap between 20% and 40% if a solution exists.
+* Runs of 30 minutes have been done on 30 randomly generated cases per scenario.
 * CPLEX was used in an i7, quadcore Ubuntu 18.04 workstation.
@@
 * `r infeasible[1]`% of instances were proved infeasible inmediately.
-* From feasible small instances, a gap of `r scen_means[1]`% was found after 30 min
-* From feasible medium instances, a gap of `r scen_means[2]`% was found after 30 min.
+* From feasible small instances, a mean gap of `r scen_means[1]`% was found after 30 min.
+* From feasible medium instances, a mean gap of `r scen_means[2]`% was found after 30 min.
 
@@
 
-* More extreme (bigger) scenarios need to be tested: more tasks and more resources.
-* A deeper analysis on the detailed solving process is envisioned. Including extending the solving time.
-* Benchmarking of solvers (gurobi, CBC, choco and CPO) is work in progress. Also: heuristics.
-* The inclusion of new possibilities, such as the planned storage of resources is pending.
-* A stochastic sensibility analysis needs to be done, including taking account of stochasticity in the input data.
+* More extreme (bigger) scenarios: more tasks and more resources.
+* A deeper analysis on the detailed solving process.
+* Benchmarking of solvers (gurobi, CBC, choco and CPO).
+* Heuristics for large scale instances.
+* The inclusion of new possibilities, such as the planned storage of resources.
+* A stochastic sensibility analysis, including taking account of stochasticity in the input data.
 

### 2018-09-20 R/presentations/CLAIO_presentation/presentation.Rmd
experiments of MOSIM instances with new model. updated CLAIO ppt with new images. New R code for images.

@@
     | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
-institute: "ISAE SUPAERO"
+institute: "ISAE-SUPAERO, Université de Toulouse, France"
 date: "September 24, 2018"
 output: 
-  slidy_presentation:
-    css: ./../../../css/baobab_slidy.css
-    incremental: yes
-    footer: "ISAE-SUPAERO, Université de Toulouse, France"
-    # slide_level: 2
-    # toc: true
   beamer_presentation:
+    # keep_tex: true
     theme: "Berlin"
@@
         in_header: header.tex
+    # before_body: before_body.tex
+  slidy_presentation:
+    css: ./../../../css/baobab_slidy.css
+    incremental: yes
+    footer: "ISAE-SUPAERO, Université de Toulouse, France"
+    # slide_level: 2
+    # toc: true
 bibliography: MFMP_NRP.bib
@@
 
-Assign both missions and maintenance operations to a fleet of aircraft in order to maximize availability and minimize costs.
-Missions have fixed start and end times and have particular needs in terms of aircraft and time.
+**Assign** both **missions and maintenance** operations to a fleet of **aircraft** in order to maximize **availability** and minimize **costs**.
+**Missions** have **fixed** start and end times and have particular **needs** in terms of aircraft and time.
 
@@
 * A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
-Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
-* Each task requires a certain number $R_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $H_j$ in each period and needs to be assigned for a minimum of $MT_j$ consecutive periods if assigned.
-* Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
-* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H^M$ units.
+* Each task requires $R_{j}$ of resources for $H_j$ hours and needs to be assigned for a minimum of $MT_j$ consecutive periods.
+* Resources $i \in \mathcal{I}$ require recurrent preventive maintenance operations.
+* A maintenance operation takes exactly $m$ periods and restores the resource's remaining usage time to exactly $H^M$ units.
 
@@
 * Missions need to be assigned for a minimal duration.
-* This model also features several improvements in the MIP modelling that permit greater instance sizes.
+* Because of these differences, a new model is proposed to deal with this new problem.
 
@@
 
+## Model: variables
+
+The following variables are used in the model.
+
+\begin{tabular}{lll}
+$m_{it}$    &: maintenance start &  $|\mathcal{I}| \times |\mathcal{T}|$ \\
+$a_{jti}$   &: task assignment &  $|\mathcal{T}| \times |\mathcal{J}_t| \times |\mathcal{I}_j|$ \\
+$a^s_{jti}$ &: task assignment start &  $|\mathcal{T}| \times |\mathcal{J}_t| \times |\mathcal{I}_j|$ \\
+$u_{it}$    &: monthly usage time &  $|\mathcal{I}| \times |\mathcal{T}|$ \\
+$rut_{it}$  &: remaining usage time &  $|\mathcal{I}| \times |\mathcal{T}|$ \\
+\end{tabular}
+
 ## Model: main constraints
@@
     & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
-      & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\
-    & m_{it} = 0
-      & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I}\label{eq:ret_min_init} \\
-    & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
-      & i \in \mathcal{I}\label{eq:ret_max_init}
+      & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}
 \end{align}
 
-Maintenances cannot executed too soon or too late with respect to the previous one.
+<!-- Maintenances cannot executed too soon or too late with respect to the previous one. -->
+
+```{r, echo=FALSE, out.width='60%', fig.align="center"}
+'../img/rut_ret.png' %>% 
+    paste0(path_to_R_dir, .) %>%
+    knitr::include_graphics()
+```
 
@@
 
+## Example of a solution
+
+```{r, echo=FALSE, out.width='100%', fig.align="center"}
+'../img/example_solution1.png' %>% 
+    paste0(path_to_R_dir, .) %>%
+    knitr::include_graphics()
+```
+
+## Results on previous instances
+
+All instances are solved to 0.02% of optimality in less than one hour. Previously this was not the case.
+
+\begin{tabular}{lrrrr}
+   id &  objective &  gap (\%) &  time (s) &      bound \\
+ I\_0 & 2212136.80 &      0.01 &     29.79 & 2211841.38 \\
+ I\_1 & 4508365.88 &      0.02 &    173.11 & 4507542.47 \\
+ I\_2 & 6208008.67 &      0.02 &    409.66 & 6206896.17 \\
+ I\_3 & 7511314.89 &      0.03 &    836.08 & 7509070.16 \\
+ I\_4 & 2217358.08 &      0.02 &     37.66 & 2216944.92 \\
+ I\_5 & 4517690.48 &      0.01 &    208.25 & 4517191.09 \\
+ I\_6 & 6220420.49 &      0.01 &    331.90 & 6219541.67 \\
+ I\_7 & 7526281.03 &      0.02 &   1731.50 & 7524413.60 \\
+ I\_8 & 2228684.87 &      0.01 &     76.18 & 2228481.13 \\
+ I\_9 & 4538717.82 &      0.02 &   3332.95 & 4537837.14 \\
+\end{tabular}
+
 ## Experiments
+
+Random instances were generated to test the performance following the known instances guidelines.
 
@@
 * Small instances have **30.000 variables** and **20.000 constraints**.
-* Medium instances have **80.000 variables** and **50.000 constraints**.
+* Large instances have **80.000 variables** and **50.000 constraints**.
 * Runs of 30 minutes have been done on 30 randomly generated cases per scenario.
@@
 
-* `r infeasible[1]`% of instances were proved infeasible inmediately.
 * From feasible small instances, a mean gap of `r scen_means[1]`% was found after 30 min.
-* From feasible medium instances, a mean gap of `r scen_means[2]`% was found after 30 min.
+* From feasible large instances, a mean gap of `r scen_means[2]`% was found after 30 min.
 

### 2018-09-20 R/appShiny/server.R

@@ -10,7 +10,7 @@ shinyServer(function(input, output, clientData, session) {
     sources <-
         c("functions/import_results.R",
           "functions/params.R",
-          "functions/exampleDT.R") %>% paste0(dir_pre, .)
+          "functions/status_resources.R") %>% paste0(dir_pre, .)
     lapply(sources, source)
 
     experiments <- PATHS[['experiments']] %>% completed_experiments()

### 2018-09-20 R/functions/status_resources.R

@@ -50,13 +50,15 @@ graph_remaining <- function(exp_directory, aircraft_to_graph='A31'){
     graph_remaining_data(data)
 }
 
-graph_remaining_data <- function(data, x= 'Mois', y= 'Temps'){
+graph_remaining_data <- function(data, x= 'Mois', y= 'Temps', return_object=FALSE){
     p <- ggplot(data, aes_string(x= x, y= y, group='aircraft')) + 
         geom_line() + 
         facet_grid(type~ ., scales = "free_y") +
         theme_minimal() + 
         scale_x_datetime()
-    
+    if (return_object){
+        return(p)
+    }
     ggplotly(p)
 }
 

### 2018-09-20 R/scripts/graphs.R

@@ -1,25 +1,43 @@
 source('functions/params.R')
 source('functions/import_results.R')
-source('functions/exampleDT.R')
+source('functions/status_resources.R')
 source('functions/gantt_tasks.R')
 
-exp_directory = PATHS[['experiments']] %>% paste0('201809131542/')
-print_solution(exp_directory, width='100%')
+exp_directory = PATHS[['experiments']] %>% paste0('201801131817/')
+# exp_directory <- PATHS[['results']] %>% paste0('simulated_data/2_task/201809180941/')
+print_solution(exp_directory, width='100%', 30)
 print_tasks(exp_directory)
 
 PATHS[['experiments']] %>% paste0('201802061201/')
 
 res <- 'A20'
+res <- '85'
 
-data_input <- treat_data(experiment1, res)
+data_input <- treat_data(exp_directory, res)
 data_input_n <- 
     data_input %>% 
     rename(Periods=Mois, Time=Temps) %>% 
     mutate(type = if_else(str_detect(type, '^dispo'), 'availability (periods)', type),
            type = if_else(str_detect(type, '^heur'), 'flight hours (hours)', type))
 
-graph_remaining_data(data_input_n, x='Periods', y='Time')
+p <- graph_remaining_data(data_input_n, x='Periods', y='Time', return_object=TRUE)
+p
+m_period <- 
+    data_input_n %>% 
+    filter(Time==60) %>% 
+    arrange(Periods) %>% 
+    select(resource=aircraft, month=Periods) %>% 
+    mutate(month = month %>% str_sub(end=7), state="M") %>% 
+    collapse_states() %>% 
+    slice(1) %>% 
+    extract2('post_month') %>% 
+    paste0("-01") %>% 
+    as.POSIXct()
+ 
+x_limits <- m_period %m+% c(months(30-6), months(60-6))
 
+# p + ggplot2::geom_vline(xintercept = ), linetype=2)
+p + geom_rect(aes(xmin=x_limits[1], xmax=x_limits[2], ymin=0, ymax=Inf), alpha = .005)
 # data_states <- get_states(experiment1)
 # periods <- 
 # data_states %>% 

### 2018-09-20 python/package/instance.py

@@ -61,7 +61,8 @@ class Instance(object):
         return self.get_category('tasks', param, default_tasks)
 
     def get_resources(self, param=None):
-        return self.get_category('resources', param)
+        default_resources = {'states': {}}
+        return self.get_category('resources', param, default_resources)
 
     def get_bounds(self):
         param_data = self.get_param()
@@ -367,9 +368,13 @@ class Instance(object):
         }
 
     def get_clusters(self):
-        capacities = self.get_tasks('capacities')
-        group = {}
         present_group = 1
+        # initialized group.
+        group = {t: present_group for t in self.get_tasks()}
+        if 'capacities' not in self.get_categories()['resources']:
+            # if we're in the old instances: we do not form clusters
+            return group
+        capacities = self.get_tasks('capacities')
         for task1, cap1 in capacities.items():
             if task1 not in group:
                 group[task1] = present_group
@@ -415,6 +420,15 @@ class Instance(object):
         return {'num': c_needs_num, 'hours': c_needs_hours}
 
     def get_task_candidates(self, task=None):
+
+        # we check if these are old instances:
+        if 'capacities' not in self.get_categories()['resources']:
+            t_candidates = self.get_tasks('candidates')
+            if task is not None:
+                return t_candidates[task]
+            return t_candidates
+
+        # if they're the 'newer' instances:
         r_cap = self.get_resources('capacities')
         t_cap = self.get_tasks('capacities')
 
@@ -424,24 +438,6 @@ class Instance(object):
                 if len(set(task_caps) - set(res_caps)) == 0:
                     t_candidates[t].append(res)
 
-        # if task is not None:
-        #     t_cap = t_cap[task]
-        # t_cap_df = pd.DataFrame([(t, c) for t in t_cap for c in t_cap[t]], columns=['IdTask', 'CAP'])
-        # r_cap_df = pd.DataFrame([(t, c) for t in r_cap for c in r_cap[t]], columns=['IdResource', 'CAP'])
-        #
-        # num_capacites = t_cap_df.groupby("IdTask"). \
-        #     agg(len).reset_index()
-        # capacites_join = t_cap_df.merge(r_cap_df, on='CAP').\
-        #     groupby(['IdTask', 'IdResource']).agg(len).reset_index()
-        # mission_aircraft = \
-        #     pd.merge(capacites_join, num_capacites, on=["IdTask", "CAP"]) \
-        #         [["IdTask", "IdResource"]]
-        #
-        # t_candidates =\
[…]

### 2018-09-20 python/package/logFiles.py

@@ -51,7 +51,7 @@ class LogFile(object):
         """
         :return: tuple of length 2
         """
-        regex = r'MIP\s+-\s+{1}:\s+Objective\s+=\s+{0}\n'.format(self.numberSearch, self.wordSearch)
+        regex = r'MIP\s+-\s+{1}.*:\s+Objective\s+=\s+{0}\n'.format(self.numberSearch, self.wordSearch)
         result = self.apply_regex(regex, flags=re.MULTILINE)
         if result is None:
             return None, None

### 2018-09-20 python/package/model.py

@@ -25,8 +25,8 @@ def solve_model(instance, options=None):
 
     # In order to break some symmetries, we're gonna give a
     # (different) price for each assignment:
-    # price_assign = {(a, v): rn.random()*10 for v in l['tasks'] for a in l['candidates'][v]}
-    price_assign = {(a, v): 0 for v in l['tasks'] for a in l['candidates'][v]}
+    price_assign = {(a, v): rn.random() for v in l['tasks'] for a in l['candidates'][v]}
+    # price_assign = {(a, v): 0 for v in l['tasks'] for a in l['candidates'][v]}
 
     # Sometimes we want to force variables to be integer.
     var_type = pl.LpContinuous
@@ -48,8 +48,11 @@ def solve_model(instance, options=None):
     rut_obj_var = pl.LpVariable(name="rut_obj_var", lowBound=0, upBound=ub['rut_end'], cat=var_type)
 
     # TEMP
-    # slack_vt = pl.LpVariable.dicts(name="slack_vt", lowBound=0, indexs=l['vt'], cat=var_type)
-    # slack_at = pl.LpVariable.dicts(name="slack_at", lowBound=0, indexs=l['at'], cat=var_type)
+    slack_vt = {tup: 0 for tup in l['vt']}
+    slack_at = {tup: 0 for tup in l['at']}
+    if options.get('slack_vars', False):
+        slack_vt = pl.LpVariable.dicts(name="slack_vt", lowBound=0, indexs=l['vt'], cat=var_type)
+        slack_at = pl.LpVariable.dicts(name="slack_at", lowBound=0, indexs=l['at'], cat=var_type)
 
     # MODEL
     model = pl.LpProblem("MFMP_v0002", pl.LpMinimize)
@@ -60,10 +63,12 @@ def solve_model(instance, options=None):
     #     model += objective
     #     model += objective >= num_maint * max_usage - rut_obj_var
     # else:
-    model += num_maint * max_usage - rut_obj_var + \
-             pl.lpSum(assign * price_assign[a, v] for (a, v, t), assign in task.items()) + \
-             1
-             # pl.lpSum(slack_vt.values()) * 99999 + pl.lpSum(slack_at.values()) * 99
+    model += 100 * num_maint * max_usage - rut_obj_var + \
+             1 * pl.lpSum(assign_st * price_assign[a, v]
+                      for (a, v, t), assign_st in start_T.items() if price_assign[a, v] > 0) + \
[…]

### 2018-09-20 python/scripts/check_solutions.py

@@ -40,11 +40,14 @@ def test2():
 
 
 def test555():
-    e = '201809121056'
-    experiment = exp.Experiment.from_dir(PATHS['experiments'] + e)
+    e = 'simulated_data/2_task/201809180941/'
+    experiment = exp.Experiment.from_dir(PATHS['results'] + e)
     instance = experiment.instance
-    options = di.load_data(PATHS['experiments'] + e + '/options.json')
-    solution = md.solve_model(instance, options)
+    options = di.load_data(PATHS['results'] + e + 'options.json')
+    checks = experiment.check_solution()
+    solution = experiment.solution
+    # solution = md.solve_model(instance, options)
+    di.export_data(options['path'], solution.data, name="data_out", file_type='json')
     # experiment.solution.print_solution("/home/pchtsp/Downloads/calendar_temp1.html")
     # checks = experiment.check_solution()
     # checks.keys()

### 2018-09-20 python/scripts/exec.py

@@ -60,16 +60,16 @@ if __name__ == "__main__":
     parser = argparse.ArgumentParser(description='Solve an instance MFMP.')
     parser.add_argument('-c', dest='file', default="package.params",
                         help='config file (default: package.params)')
-    parser.add_argument('-d', '--options', dest='config_dict',  type=json.loads)
+    parser.add_argument('-d', '--options', dest='config_dict', type=json.loads)
 
     args = parser.parse_args()
     # if not os.path.exists(args.file):
     #     raise FileNotFoundError("{} was not found".format(args.file))
-    more_options = args.config_dict
 
     print('Using config file in {}'.format(args.file))
     params = importlib.import_module(args.file)
-    params.OPTIONS.update(more_options)
+    if args.config_dict:
+        params.OPTIONS.update(args.config_dict)
 
     # import package.params as params
     config_and_solve(params)

### 2018-09-20 python/scripts/graphs2.py

@@ -106,11 +106,11 @@ def get_results_table(path_abs, exp_list=None, **kwargs):
     return table
 
 
-def results_table():
+def results_table(dir_origin=path_abs, dir_dest=path_latex + 'MOSIM2018/tables/', exp_list=None):
     ################################################
     # Instances and Results table
     ################################################
-    table = get_results_table(path_abs)
+    table = get_results_table(dir_origin, exp_list=exp_list)
 
     # pp.pprint(exps)
     # input: num missions, num periods, variables, constraints, assignments, timeLimit
@@ -120,7 +120,8 @@ def results_table():
 
     table_in = table[input_cols].rename(columns=cols_rename)
     latex = table_in.to_latex(escape=False, bold_rows=True, index=False, float_format='%.0f')
-    with open(path_latex + 'MOSIM2018/tables/instance.tex', 'w') as f:
+    destination = dir_dest + 'instance.tex'
+    with open(destination, 'w') as f:
         f.write(latex)
 
     # output: number of cuts, initial relaxation, relaxation after cuts, end relaxation, cuts' time.
@@ -148,8 +149,9 @@ def results_table():
 
     table1 = table.rename(columns=cols_rename)[
         ['id', 'objective', 'gap (\%)', 'time (s)', 'bound']]
-    latex = table1.to_latex(escape=False, bold_rows=True, index=False, float_format='%.1f')
-    with open(path_latex + 'MOSIM2018/tables/summary.tex', 'w') as f:
+    latex = table1.to_latex(escape=False, bold_rows=True, index=False, float_format='%.2f')
+    destination = dir_dest + 'summary.tex'
+    with open(destination, 'w') as f:
         f.write(latex)
 
     table2 = table
@@ -158,7 +160,8 @@ def results_table():
         ['id', 'root', 'b. cuts',
          'bound', 'cuts (\#)', 'cuts (s)']]
     latex = table2.to_latex(escape=False, bold_rows=True, index=False, float_format='%.1f')
-    with open(path_latex + 'MOSIM2018/tables/results.tex', 'w') as f:
+    destination = dir_dest + 'results.tex'
+    with open(destination, 'w') as f:
         f.write(latex)
 
     table3 = table.rename(columns=cols_rename3)[
@@ -167,7 +170,8 @@ def results_table():
[…]

### 2018-09-20 python/scripts/remake_MOSIM.py

@@ -1,0 +1,46 @@
+import scripts.exec as exec
+import os
+import datetime as dt
+import time
+# import package.instance as inst
+import package.params as params
+import package.experiment as exp
+import package.model as md
+import package.data_input as di
+
+
+if __name__ == "__main__":
+
+    path = params.PATHS['results'] + 'MOSIM2018_copy/'
+    cases = [c for c in os.listdir(path) if c.startswith("2018")]
+    for case in cases:
+        # case = '201801141646'
+        output_path = path + case + '/'
+        experiment = exp.Experiment.from_dir(output_path)
+        instance = experiment.instance
+        options = di.load_data(output_path + 'options.json')
+        options['path'] = output_path
+        options['gap'] = 0.00049
+        solution = md.solve_model(instance, options=options)
+        if solution is not None:
+            di.export_data(output_path, solution.data, name="data_out", file_type='json')
+
+    # sim_data = params.OPTIONS['simulation']
+    # params.PATHS['experiments'] = params.PATHS['results'] + 'simulated_data/2_task_slack/'
+    #
+    # for num_tasks in range(2, 3):
+    #     for sim in range(30):
+    #         sim_data['seed'] = None
+    #         sim_data['num_resources'] = num_tasks * 50
+    #         sim_data['num_parallel_tasks'] = num_tasks
+    #         # we don't care about the original params object.
+    #         # so we do not copy it.
+    #         # params.OPTIONS['end_pos'] = period
+    #         # params.OPTIONS['solver'] = solver
+    #         params.OPTIONS['path'] = os.path.join(
+    #             params.PATHS['experiments'],
+    #             dt.datetime.now().strftime("%Y%m%d%H%M")
+    #         ) + '/'
+    #
+    #         exec.config_and_solve(params)
+    #         time.sleep(60)

### 2018-10-02 R/presentations/CLAIO_presentation/presentation.Rmd
updated CLAIO presentation. This was what was presented.

@@
 
-* In @Cho2011, US Army aircraft were assigned daily operations over a year to minimize the number of maintenances.
+* In @Cho2011, US Air Force aircraft were assigned daily operations over a year to minimize the number of maintenances.
 * In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
@@
 
-\begin{tabular}{lrrrr}
-   id &  objective &  gap (\%) &  time (s) &      bound \\
- I\_0 & 2212136.80 &      0.01 &     29.79 & 2211841.38 \\
- I\_1 & 4508365.88 &      0.02 &    173.11 & 4507542.47 \\
- I\_2 & 6208008.67 &      0.02 &    409.66 & 6206896.17 \\
- I\_3 & 7511314.89 &      0.03 &    836.08 & 7509070.16 \\
- I\_4 & 2217358.08 &      0.02 &     37.66 & 2216944.92 \\
- I\_5 & 4517690.48 &      0.01 &    208.25 & 4517191.09 \\
- I\_6 & 6220420.49 &      0.01 &    331.90 & 6219541.67 \\
- I\_7 & 7526281.03 &      0.02 &   1731.50 & 7524413.60 \\
- I\_8 & 2228684.87 &      0.01 &     76.18 & 2228481.13 \\
- I\_9 & 4538717.82 &      0.02 &   3332.95 & 4537837.14 \\
-\end{tabular}
+| id | $\|\mathcal{J}\|$ |$\|\mathcal{T}\|$ |  assign |  gap (\%)   |  time (s) |
+|-----|------------------|------------------|---------|-------------|-----------|
+| I_0 |                9 |               11 |     310 |      0.01 	|     29.79 |
+| I_1 |                9 |               21 |     650 |      0.02 	|    173.11 |
+| I_2 |                9 |               31 |     990 |      0.02 	|    409.66 |
+| I_3 |                9 |               41 |    1249 |      0.03 	|    836.08 |
+| I_4 |               10 |               11 |     530 |      0.02 	|     37.66 |
+| I_5 |               10 |               21 |    1070 |      0.01 	|    208.25 |
+| I_6 |               10 |               31 |    1610 |      0.01 	|    331.90 |
+| I_7 |               10 |               41 |    2069 |      0.02 	|   1731.50 |
+| I_8 |               11 |               11 |    1080 |      0.01 	|     76.18 |
+| I_9 |               11 |               21 |    2120 |      0.02 	|   3332.95 |
 
@@
 
-* Small instances have **30.000 variables** and **20.000 constraints**.
-* Large instances have **80.000 variables** and **50.000 constraints**.
+* Instances have **30.000 - 80.000 variables** and **20.000 - 50.000 constraints**.
 * Runs of 30 minutes have been done on 30 randomly generated cases per scenario.

### 2018-10-30 latex/AFIS2018/poster_FP.tex
added poster for event AFIS2018

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.1 (14/06/14)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
% 
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\usepackage[scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{ upgreek }
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{ bbold }
\usepackage{dsfont}
\usepackage{array}
\usepackage{mathtools, bm}
\usepackage{amssymb, bm}
\usepackage{amsthm} 
\usepackage{geometry} 
\usepackage{relsize,exscale}
\usepackage{multicol}

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{36in} % A0 width: 46.8in
\setlength{\paperheight}{48in} % A0 height: 33.1in
\setlength{\sepwid}{0.024\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.22\paperwidth} % Width of one column
\setlength{\twocolwid}{0.464\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.708\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables

\DeclareFontShape{OMX}{cmex}{m}{n}{
  <-7.5> cmex7
  <7.5-8.5> cmex8
  <8.5-9.5> cmex9
  <9.5-> cmex10
}{}

\SetSymbolFont{largesymbols}{normal}{OMX}{cmex}{m}{n}
\SetSymbolFont{largesymbols}{bold}  {OMX}{cmex}{m}{n}

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------
		% \begin{center}
		% 	\begin{tabular}{ccc}
		% 		\includegraphics[width=0.6\linewidth]{img/logo-isae-supaero.png}
		% 	\end{tabular}
		% \end{center}

\title{\hspace*{10ex} Planning aircraft maintenances for the French Airforce} % Poster title

\author{Franco Peschiera, Olga Battaïa, Alain Haït, Nicolas Dupin} % Author(s)

\institute{ISAE-SUPAERO} % Institution(s)

 % \footer{line1 \\ line2}
% ----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{headline}{} 
{
\begin{tikzpicture}[remember picture,overlay] 
	\node [shift={(12 cm,-8cm)}] at (current page.north west) {\includegraphics[height=10cm]{img/logo-isae-supaero.png}}; 
\end{tikzpicture} 
}

% \setbeamertemplate{footline}{} 
% {
% % \parbox{\linewidth}{\vspace*{-8pt}some text\hfill\insertshortauthor\hfill\insertpagenumber}
% Séminaire Doctoral, Rencontres académie – industrie AFIS 			5-6 décembre 2018, Université de Lorraine
% }

\addtobeamertemplate{footline}{} 
{

		\begin{tikzpicture}[remember picture,overlay] 
			\node [shift={(35 cm,5cm)}] at (current page.south west) {\includegraphics[height=8cm]{img/AFIS_logo_2.jpg}};
		\end{tikzpicture}

}

% \footer{}

    % { % these braces make the change local to the single frame
    %     \setbeamertemplate{footline}{boo \insertframenumber}
    %     \begin{frame}[t]{Frame 2}
    %         B
    %     \end{frame}
    % }

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

	\begin{column}{\sepwid}\end{column} % Empty spacer column

	\begin{column}{\onecolwid} % The first column

		%----------------------------------------------------------------------------------------
		%	OBJECTIVES
		%----------------------------------------------------------------------------------------

		\begin{alertblock}{Practical problematic}

			\begin{itemize}
			\item How to increase the Air Force's availability of aircraft?
			\item What's the maximum number of aircraft that we can keep in good shape with a given maintenance capacity?
			\end{itemize}
		\end{alertblock}

		%----------------------------------------------------------------------------------------
		%	INTRODUCTION
		%----------------------------------------------------------------------------------------

		\begin{block}{Problem}

		    The problem consists in assigning military aircraft to a given set of missions while scheduling the maintenance operations over a time horizon.

			\begin{figure}
				\includegraphics[width=1\linewidth]{img/mirage2000.jpg}
			\caption{The fleet of aircraft consists on Mirage 2000 (source: wikipedia; public domain)}
			\end{figure}

			A simplified description follows.

		    A series of $j \in \mathcal{J}$ missions are planned along a horizon divided into $t \in \mathcal{T}$ periods. Each mission requires a certain number $r_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period. Set $a_{j} \subset \mathcal{I}$ lists the aircraft that can be assigned to each mission and set $O_i \subset \mathcal{J}$ consists of missions for which aircraft $i$ can be used.

		    Aircraft require recurrent preventive maintenance operations (of fixed duration $m$) since the realization of missions diminish their remaining flight hours. An aircraft cannot be used for a mission $j$ if its remaining flight hours are less than the amount $h_j$ required by mission $j$. A maintenance operation assigns each aircraft exactly $H$ flight hours. The remaining flight hours not used before the maintenance operation are lost.

		\end{block}

	\end{column} % End of the first column

	\begin{column}{\sepwid}\end{column} % Empty spacer column

	\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)

		\setbeamercolor{block alerted title}{fg=black,bg=norange} % Change the alert block title colors
		\setbeamercolor{block alerted body}{fg=black,bg=white} % Change the alert block body colors

	\begin{column}{\onecolwid}\end{column} % Empty spacer column

	\begin{column}{\twocolwid}
			\vspace*{-5ex}
			\begin{alertblock}{}				
				\begin{itemize}
					\item Doctorant (email): Franco Peschiera (franco.peschiera@isae-supaero.fr)
					\item Directeurs de Thèse: Olga Battaïa, Alain Haït, Nicolas Dupin
					\item Laboratoire, Université: ISAE-SUPAERO
					\item Collaboration industrielle: DGA
				\end{itemize}
			\end{alertblock}

	\end{column}

	\begin{column}{\onecolwid}\end{column} % Empty spacer column

		\begin{figure}
			\includegraphics[width=1\linewidth]{img/calendar2.png}
			\caption{Partial solution showing the planning of maintenances (green) and assignments of missions (red and yellow) for some aircraft (in rows) over time periods (in columns).}
		\end{figure}

	\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column

		\begin{column}{\twocolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)

			\begin{block}{Mathematical model}

				The following model represents feasible and optimal solutions to the problem.

			    \begin{align*}
			    	& \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \\
			        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq M^{max}
			          & t \in \mathcal{T} \\
			        & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
			                & j \in \mathcal{J}, t \in \mathcal{T}_j  \\
			        & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
			                & t \in \mathcal{T}, i \in \mathcal{I}\\
			        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
			                & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \\
			        & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
			        & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}\\
			       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} \leq AK_{kt}
			        &k \in \mathcal{K}, t \in \mathcal{T} \\
			         & rut_{it} \leq rut_{i(t-1)} + H m_{it} \notag \\ 
			         	& - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
			         	& t =1, ..., \mathcal{T}, i \in \mathcal{I} \\
			        & rut_{i0} = Rut^{Init}_i
			               & i \in \mathcal{I} \\
			        & rut_{it} \geq H m_{it}
			                & t \in \mathcal{T}, i \in \mathcal{I}\\ 
			        & rut_{it} \in [0,H]
			                & t \in \mathcal{T}, i \in \mathcal{I} \\
			        & m_{it'} + m_{it} \leq 1
			          & t \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\\ 
			        & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
			          & t \in \mathcal{T}, i \in \mathcal{I}\\
			        & m_{it} = 0
			          & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I} \\
			        & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
			          & i \in \mathcal{I}
			    \end{align*}

			\end{block}

			\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
			\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks

			\begin{alertblock}{Acknowledgements}

				This thesis is half financed by DGA and takes place at ISAE-SUPAERO.

			\end{alertblock}

		\end{column} % End of column 2.1

		\end{columns} % End of the split of column 2

	\end{column} % End of the second column

	\begin{column}{\sepwid}\end{column} % Empty spacer column

	\begin{column}{\onecolwid} % The third column

			% \begin{column}{\onecolwid}\vspace{-.6in} %  (column 2.2)

			\begin{block}{Results}

				The following figures show some views on the output generated by the model.

				\begin{figure}
					\includegraphics[width=1\linewidth]{img/remaining2.png}
					\caption{Partial solution showing the remaining periods and flight hours for a single aircraft. Both values should ideally reach 0 at the same time.}
					% \label{rem}
				\end{figure}

				\begin{figure}

					\includegraphics[width=1\linewidth]{img/num-maintenances.png}
					\caption{Partial solution showing the total number of aircraft under maintenance at each period. This value should be as constant (and low) as possible.}
					% \label{maint}
				\end{figure}
			\end{block}

			% \end{column} % End of column 2.2

		%----------------------------------------------------------------------------------------
		%	CONCLUSION
		%----------------------------------------------------------------------------------------
		\begin{block}{Discipline and techniques}

			\textbf{Operations Research}

			\begin{itemize}
				\item Commonly called "The science of better".
				\item Consists of techniques to help the decision making process.
				\item Used for planning, programming, scheduling, routing, pricing in production, logistics, human resources, marketing and investment.
			\end{itemize}

			\textbf{Mathematical programming}

			\begin{itemize}
				\item Permits the modelling of a problem as a set of variables and equations.
				\item Solved via dedicated software optimized to solve this kinds of equations.
				\item Used by the industry and governments to solve real-life large-scale problems all over the world.
			\end{itemize}

		\end{block}

	%----------------------------------------------------------------------------------------

	\end{column} % End of the third column

\end{columns} % End of all the columns in the poster

% \begin{columns}[t]

% 	\begin{column}{\onecolwid}
% 		\includegraphics[width=1\linewidth]{img/AFIS_logo.jpg}
% 	\end{column} % Empty spacer column

% 	\begin{column}{\onecolwid}\end{column}
% 	\begin{column}{\threecolwid}
% 			Séminaire Doctoral, Rencontres académie – industrie AFIS\\
% 			5-6 décembre 2018, Université de Lorraine
% 	\end{column}

% \end{columns}

\end{frame} % End of the enclosing frame

\end{document}

### 2018-10-30 latex/AFIS2018/poster_FP.tex
changes to poster and added resume for AFIS 2018

@@
 		\begin{tikzpicture}[remember picture,overlay] 
-			\node [shift={(35 cm,5cm)}] at (current page.south west) {\includegraphics[height=8cm]{img/AFIS_logo_2.jpg}};
+			\node [shift={(38 cm,5cm)}] at (current page.south west) {\includegraphics[height=8cm]{img/AFIS_logo_2.jpg}};
 		\end{tikzpicture}
@@
 	\begin{column}{\twocolwid}
-			\vspace*{-5ex}
+			\vspace*{-6ex}
 			\begin{alertblock}{}				
 				\begin{itemize}
-					\item Doctorant (email): Franco Peschiera (franco.peschiera@isae-supaero.fr)
+					\item Franco Peschiera (franco.peschiera@isae-supaero.fr)
 					\item Directeurs de Thèse: Olga Battaïa, Alain Haït, Nicolas Dupin
-					\item Laboratoire, Université: ISAE-SUPAERO
+					\item Institut: ISAE-SUPAERO
 					\item Collaboration industrielle: DGA

### 2018-10-30 latex/AFIS2018/resume.tex
changes to poster and added resume for AFIS 2018

\documentclass{roadef}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{graphicx}  % Required for including images
% \usepackage{setspace}

% \usepackage{fontspec}
% This command is to use simple quotes inside math expressions:
\newcommand{\mq}[1] {`#1\textrm'}
\setlist{nosep}

\pagestyle{fancy}
\fancyhf{}

\lfoot{\includegraphics[height=1.5cm]{img/AFIS_logo_2.jpg}}
% \rfoot{\includegraphics[height=2cm]{img/AFIS_logo_2.jpg}}

% \setstretch{1.2}

\fancypagestyle{plain}{%
  \fancyhf{}
\lfoot{\enspace\includegraphics[height=1.5cm]{img/AFIS_logo_2.jpg}}

}

\begin{document}

% \footskip{1.9pt}
% Le titre du papier
\title{Planning aircraft maintenances for the French Airforce}

% Le titre court
% \def\shorttitle{Titre court}

\author{Franco Peschiera\inst{1}, Alain Haït\inst{1}, Olga Battaïa\inst{1}, Nicolas Dupin\inst{1}}

\institute{ISAE-SUPAERO, Université de Toulouse, France}

\maketitle
% \thispagestyle{empty}

% \vspace{-10em}
\begin{itemize}
	\item[] Franco Peschiera (franco.peschiera@isae-supaero.fr)
	\item[] Directeurs de Thèse: Olga Battaïa, Alain Haït, Nicolas Dupin
	\item[] Institut: ISAE-SUPAERO, Université de Toulouse, France
	\item[] Collaboration industrielle: DGA
\end{itemize}

\keywords{optimization, planning, military, maintenance}

%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

% \end{keyword}

% \end{frontmatter}

%% Title, authors and addresses
% 
\abstract{}

The Flight and Maintenance Planning (FMP) problem deals with the scheduling of maintenance operations and the assignment of flight activities to a fleet of aircraft. It has two main variants: civil (or commercial) and military. We study the latter, the lesser known variant. Previous work include short and medium term planning: \cite{Cho2011,Kozanidis2008,Verhoeff2015,marlow2017optimal,Seif2018}.

A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$. Each task requires a certain number $R_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $H_j$ in each of a series of at least $MT_j$ consecutive periods. Set $I_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used. The number of resources in use for each period $D_t$ is also known.

Each resource can only be assigned to a single task in any given period. These resources suffer from wear and tear and require regular maintenance operations during their lifetime. The need for maintenance is calculated based on two indicators. The first one is called "remaining elapsed time" (or $ret_{it}$). It expresses the amount of time (measured in time periods) after which the resource cannot be used any more and has to undergo a maintenance operation. Its value is calculated for each resource $i$ and each time period $t$. In a similar way, "remaining usage time" (or $rut_{it}$) is used to measure the amount of time that the resource $i$ can be used before needing a maintenance operation at any given period $t$.

Each maintenance operation has a fix duration of $M$ periods and cannot be interrupted: during this time the resource cannot be assigned to any task. After a maintenance operation, a resource restores its remaining elapsed time and remaining usage time to their maximum values $E$ and $H$ respectively.

The objective is to minimize the total number of maintenances and, at the same time, maximize the end status of the resources.     Maintenance capacity is controlled by (\ref{eq:capacity1}). Tasks' resource requirements are defined by (\ref{eq:taskres}). Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task or a maintenance operation at the same period. Constraints (\ref{eq:start1}) control the definition of starting a consecutive assignment: the variable is active if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this same task in the previous period ($t-1$). Constraints (\ref{eq:start3}) control the minimum duration of a consecutive task assignment. Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability of resources for each cluster $k$. A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task. The minimum and maximum elapsed time are defined by (\ref{eq:ret_min}) and (\ref{eq:ret_max})

\begin{align}
	& \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \\
    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq M^{max}
      & t \in \mathcal{T} \label{eq:capacity1}\\
    & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
            & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
    & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
            & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
    & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
            & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start1} \\
    & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
    & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}\\
   & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} \leq AK_{kt}
    &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\
     & rut_{it} \leq rut_{i(t-1)} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
     	& t =1, ..., \mathcal{T}, i \in \mathcal{I} \\
    & rut_{i0} = Rut^{Init}_i
           & i \in \mathcal{I} \\
    & rut_{it} \geq H m_{it}
            & t \in \mathcal{T}, i \in \mathcal{I}\\ 
    & rut_{it} \in [0,H]
            & t \in \mathcal{T}, i \in \mathcal{I} \\
    & m_{it'} + m_{it} \leq 1
      & t \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
    & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
      & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\
    & m_{it} = 0
      & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I} \\
    & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
      & i \in \mathcal{I}
\end{align}

\bibliographystyle{plain}
\bibliography{MFMP}

\end{document}

### 2018-11-30 R/presentations/CLAIO_presentation/presentation.Rmd

@@
         in_header: header.tex
-    # before_body: before_body.tex
   slidy_presentation:
@@
 knitr::opts_chunk$set(echo = FALSE)
+path_to_R_dir <- '.'
 path_to_R_dir <- "./../../"
@@
 all_exp_dir <- PATHS[['results']] %>% paste0('simulated_data/')
-exp_directory <- all_exp_dir %>% paste0('2_task/201809180941/')
+exp_directory <- all_exp_dir %>% paste0('task_periods_solv_2_50_GUROBI/201810121218/')
 gantt <- print_solution(exp_directory, max_resources = 15, width='100%')

### 2018-11-30 latex/CLAIO2018/article.tex

@@
 %
-\title{\bf MIP formulation for the optimization of maintenance planning on French military aircraft operations}
+\title{\bf A novel MIP formulation for the optimization problem of maintenance planning of military aircraft}
 
@@
 
-{\em A specific Flight and Maintenance Planning problem is presented. 
-In this problem, preventive maintenance operations are scheduled for military aircraft along with the assignment of regular missions. 
-The quality of the solutions is measured via a bi-objective function that smooths both maintenance operations and aircraft unavailability among time periods. 
-A real world dataset provided by the French Air Force is used to test an exact optimization approach based on a Mixed Integer Programming model.
+{\em A Flight and Maintenance Planning problem is considered for military aircraft.
+A novel Mixed-Integer Linear Programming model is developed for scheduling maintenance operations along with the assignment of regular, pre-scheduled, missions.
+This model takes into account the specific requirements of a military environment and its operational goals.
+On the contrary to previous studies, the developed model is applicable for the management of a heterogenic fleet.
+Different combinations of objective functions are discussed for multi-objective optimization and the impact of their use is analyzed on a dataset provided by French Army Forces.
+The complexity of the obtained problem formulations is studied; new procedures are developed in order to decrease the computational time required to construct efficient Pareto-fronts of alternative solutions for real-life sized instances.
+The results obtained in computational tests show the efficiency of the developed optimization approach.
 }
@@
 \end{abstract}
-
-\section{Introduction}
-\label{sec:intro}
-
-    Flight and Maintenance Planning (FMP) problems are raised in civilian or military operational applications. 
-    They schedule maintenances of aircraft taking into account the implications of maintenance immobilizations for the demand in flight hours, for a conjoint optimization among maintenance dates and flight hours.
-
-    The specific FMP of this paper comes from the French Air Force's operations. It has two specificities regarding the classical FMP considered in the literature.
-    On one hand, the fleet of aircraft is heterogeneous, with different standards, capacities and retrofits.
-    On the other hand, there is a will to align two objectives: the smoothing of the maintenance operations and the minimization of the global unavailability of aircraft to be able to add new missions.
-
-    This case of the FMP has been already studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008,Verhoeff2015,Seif2018}.
-
-    This paper investigates how to solve such new optimization problem, and how to take good compromise decisions if these objectives are antagonistic.
-
-\section{Problem Description}
-\label{sec:model}
-% TODO: make this compatible with parameters.
-
-    The problem consists in assigning military aircraft to a given set of missions while scheduling the maintenance operations over a time horizon.
-    In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks".
-A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods.
-Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
-Similarly, all the tasks to be realized in period $t$ are known and defined by set $J_t \subset \mathcal{J}$.
-Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
-Set $I_{j} \subset \mathcal{I}$ lists the resources that can be assigned to each task and set $O_i \subset \mathcal{J}$ consists of tasks for which resource $i$ can be used.
-The number of resources in use for each period $d_t$ is also known.
-
-    Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
-A resource cannot be used for a task $j$ if its remaining usage time is less than usage $h_j$ required by task $j$.
-A maintenance operation assigns to the resource exactly $H$ units of remaining usage time.
-The remaining usage time not used before the maintenance operation is lost.
-
-    A maintenance operation takes exactly $m$ periods and cannot be interrupted: during this time the resource cannot be assigned to any task.
-Some resources can be in maintenance at the beginning of the planning horizon, $n_t$ is used for the number of aircraft in such planned maintenance per period $t$ and defined only for the first $m-1$ time periods.
-In the next section, we present a new mathematical model for this problem.
-
-\section{Solution Method}
-\label{sec:method}
-
-  \subsection{Sets}
-
-    \begin{tabular}{p{5mm}p{165mm}}
-
-        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
-        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
-        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
-        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}$\}. \\
-    \end{tabular}
-
-  \subsection{Parameters}
-
-    \begin{tabular}{p{8mm}p{162mm}}
-        $H_j$             & amount of resource time required by task $j$. \\
-        $R_j$             & number of resources required by task $j$. \\
-        $M$               & maintenance duration in number of periods. \\
-        $E$               & remaining elapsed time after a maintenance. \\
-        $H$               & remaining usage time after a maintenance. \\
-        $W_1$             & weight of the first objective in the objective function. \\
-        $W_2$             & weight of the second objective in the objective function. \\
-        $N_t$             & number of resources in already-planned maintenances in period $t$ at the beginning of the planning horizon.\\
-        $D_t$             & number of resources to be assigned in total in period $t$. \\
-        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
-        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
-        $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
-        $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
-    \end{tabular}
-
-  \subsection{Variables}
-
-     The following decision variables define a solution.
-
-    \begin{tabular}{p{8mm}p{162mm}}
-        $a_{jti}$   &  =1 if task $j \in J$ in period $t \in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
-        $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
-        $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
-        $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
-        $u_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
-        $m_{max}$   &  maximal number (integer) of resources in maintenance in any period. \\
-    \end{tabular}
-
-    Note that  $a_{jti}$ and $m_{it}$ are initially set up to 0 for all resources already in maintenance at the beginning of the planning horizon for the remaining time periods of maintenance.
-The remaining usage time for each resource at the beginning of the planning horizon is used to initialize $rut_{i0}$.
-
-  \subsection{Constraints}
-
-    The objective is to minimize the total number of maintenances done and while preserving the aircraft ready at the end of the planning period.
-
-    \begin{align}
-        & \text{Min}\; W_1 m_{max} + W_2 u_{max}
-    \end{align}
-
-    \begin{align}
-        % maximum capacity1:
-        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
-          & t \in \mathcal{T} \label{eq:capacity1}\\             
-       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
-        &t \in \mathcal{T} \label{eq:avalaibility1}\\                
-        & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
-                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
-        & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
-    \end{align}
-
-    Maintenance capacity is controlled by (\ref{eq:capacity1}).
-The number of unavailable resources is defined by (\ref{eq:avalaibility1}).
-Tasks' resource requirements are defined by (\ref{eq:taskres}).
-Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task or maintenance operation at the same period.
-    \begin{align}
-        % remaining used time
-         & rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
-                & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
-        & rut_{i0} = Rut^{Init}_i
-               & i \in \mathcal{I} \label{eq:rut_initial}\\
-        & rut_{it} \geq H m_{it}
-                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rut_lower}\\ 
-        & rut_{it} \in [0,H]
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu} \\
-        & ret_{it} \leq ret_{it-1} - 1 + E m_{it}
-                & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
-        & ret_{i0} = Ret^{Init}_i
-                & i \in \mathcal{I} \label{eq:ret_initial}\\
-        & ret_{it} \geq E m_{it}
-                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
-        & ret_{it} \in [0,E]
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
-        & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
-              & t = |\mathcal{T}| \label{eq:min_ret}\\
-        & \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
-              & t = |\mathcal{T}| \label{eq:min_rut}
-    \end{align}
-        % These constraints calculate the balances of hours for each resource.
-    The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}). 
-    Similarly, the remaining elapsed time is defined by (\ref{eq:ret_upper})-(\ref{eq:ret_initial}) and its limits by (\ref{eq:ret_lower})-(\ref{eq:me}). 
-    Finally, constraints \ref{eq:min_ret} and \ref{eq:min_rut} guarantee that resources have, globally, the same amount of remaining used and elapsed times at the beginning and at the end of the planning horizon.
-
-%
-%   Conclusions
-%
-\section{Conclusions}
-\label{sec:conclusions}
-
-  This paper proposed a new MIP formulation to address a specific FMP with an heterogeneous fleet of aircraft.
-  Two objective functions were chosen to address the French Air Force requirements.
-  The first one was to smooth maintenance operations and the second one was to increase aircraft availability.
-  The characteristics of MIP convergence were analyzed on real-world instances.
-  Mono-objective problem instances were solved to optimality for medium size instances.
-For larger instances, difficulties in finding optimal solutions occurred because of multiple symmetries and a poor quality of MIP generic primal heuristics.
-
-  The first perspective is to improve the resolution of large size instances.
-Heuristics could be implemented, but also mathematical programming can be improved with symmetry breaking or MIP reformulations.
-A matheuristic method derived from the existing MIP formulation can also be used to improve the performance.
-
-  Lastly, additional constraints from the real world application can be incorporated in the model.
-
-%
-%   References
-%
-\bibliographystyle{plain}
-\bibliography{./../biblio/MFMP}
 \end{document}

### 2019-03-25 latex/notes/model/model.tex
changes to model.tex to updated and then change it to new missions.

--- 2018-08-30 latex/notes/model/model.tex
@@
 \usepackage{caption}
+\usepackage{pgfgantt}
 
@@
 
-\section{Optimization model}
+\section{Mathematical formulation}
   \label{sec:model}
 
-  \subsection{Basic sets}
+The following model provides a tight MIP formulation that solves the Military Flight and Maintenance Problem described in section \ref{sec:problem}.
+Maintenance operations will be referred as "checks".
 
-    \begin{tabular}{p{5mm}p{170mm}}
-        $\mathcal{I}$     &  resources. \\
-        $\mathcal{T}$     &  time periods inside planning horizon. \\
-        $\mathcal{J}$     &  tasks. \\
-        $\mathcal{K}$     &  cluster of resources that share the same capacities to do tasks. \\
+\subsection{Input data}
+    \subsubsection{Basic sets}
+
+        \begin{tabular}{p{5mm}p{120mm}}
+            $\mathcal{I}$     &  aircraft. \\
+            $\mathcal{T}$     &  time periods included in the planning horizon. \\
+            $\mathcal{J}$     &  missions. \\
+            $\mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
         \end{tabular}
 
-  \subsection{Parameters}
+    \subsubsection{Parameters}
 
-    \begin{tabular}{p{8mm}p{167mm}}
-        $H_j$             & amount of resource time required by task $j$. \\
-        $R_j$             & number of resources required by task $j$. \\
-        $MT_j$            & minimum number of consecutive periods a resource can be assigned to task $j$. \\
-        % $aut_i$           & maximal absolute usage time for resource $i$. \\
-        % $aet_i$           & maximal absolute elapsed time for resource $i$. \\
-        $M$               & maintenance duration in number of periods. \\
-        $C^{max}$         & maximum number of simultaneous maintenance operations. \\
-        % $E$               & remaining elapsed time after a maintenance. \\
-        $E^{M}$         & maximum number of periods between two consecutive maintenances. \\
-        $E^{m}$         & minimum number of periods between two consecutive maintenances. \\
-        $H^{M}$               & remaining usage time after a maintenance. \\
-        % $S$               & Storage duration in number of periods. \\
-        % $W_1$             & weight of the first objective in the objective function. \\
-        % $W_2$             & weight of the second objective in the objective function. \\
-        $N_t$             & number of resources in already-planned maintenances in period $t$.\\
-        $N^{clust}_{kt}$         & number of resources in cluster $k$ in already-planned maintenances in period $t$.\\
-        % $D_t$             & number of resources to be assigned in total in period $t$. \\
-        $A^{clust}_{kt}$         & maximum number of resources in cluster $k$ that can be simultaneously under maintenance at period $t$.\\
-        $H^{clust}_{kt}$         & minimum number of total remaining usage time for cluster $k$ at period $t$.\\
-        $Rut^{Init}_{i}$  & remaining usage time for resource $i$ at the start of the planning horizon. \\
-        $Ret^{Init}_{i}$  & remaining elapsed time for resource $i$ at the start of the planning horizon. \\
-        % $Ret^{Init}_{sum}$& sum of remaining elapsed times at the start of the planning horizon. \\
-        % $Rut^{Init}_{sum}$& sum of remaining elapsed time at the start of the planning horizon. \\
+        \begin{tabular}{p{10mm}p{115mm}}
+            $H_j$             & amount of flight time required by mission $j$. \\
+            $R_j$             & number of aircraft required by mission $j$. \\
+            $MT_j$            & minimum number of consecutive periods an aircraft has to be assigned to mission $j$. \\
+            $U^{min}$       & default aircraft flight time if it is not assigned to any mission nor in maintenance.\\
+            $M$               & check duration in number of periods. \\
+            $C^{max}$         & maximum number of simultaneous checks. \\
+            $E^M$         & maximum number of periods between two consecutive checks. \\
+            $E^m$         & minimum number of periods between two consecutive checks. \\
+            $H^M$               & remaining flight time after a check. \\
+            $N_t$             & number of aircraft known to be in maintenance in period $t$.\\
+            $N^K_{kt}$         & number of aircraft in cluster $k$ known to be in maintenance in period $t$.\\
+            $A^K_{kt}$         & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$.\\
+            $H^K_{kt}$         & minimum number of total remaining flight time for cluster $k$ at period $t$.\\
+            $Rft^{Init}_i$  & remaining flight time for aircraft $i$ at the start of the planning horizon. \\
+            $Rct^{Init}_i$  & remaining calendar time for aircraft $i$ at the start of the planning horizon. \\
         \end{tabular}
 
-    \subsection{Parametric sets}
+    \subsubsection{Parametric sets}
 
-        \begin{tabular}{p{5mm}p{170mm}}
-        $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which task $j$ is active. \\
-        $\mathcal{J}_t $    &  tasks $j \in \mathcal{J}$ to be realized in period $t$. \\
-        $\mathcal{I}_j$     &  resources $i \in \mathcal{I}$ that can be assigned to task $j$. \\
-        $\mathcal{I}_k$     &  resources $i \in \mathcal{I}$ that are included in cluster $k$. One resource can belong to more than one cluster. \\
-        $\mathcal{O}_i$     &  tasks $j \in \mathcal{J}$ for which resource $i$ can be used. \\
-        $\mathcal{T}^{s}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$. \\
-        % $\mathcal{T}^{st}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - S+1\}},  ..., {t}$\}. \\
-        $\mathcal{T}^{m}_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + E^{m}-1\}}\}$. \\
-        $\mathcal{T}^{M}_t$ &  time periods $t' \leq |\mathcal{T}| - E^{M}$ such that $t' \in \{ t + E^{m}-1 , ...,  t + E^{M}-1 \}$. \\
-        $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Ret^{Init}_{i} - E^M + E^m \}}\}$. \\
-        $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Ret^{Init}_{i} - E^M + E^m \}} , ...,  Ret^{Init}_{i} \}$. \\
-        $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$. \\
-        $\mathcal{A}^{init}_{j}$  & resources $i \in \mathcal{I}$ that have task $j$ pre-assigned in the previous period to the start of the horizon. \\
+        \begin{tabular}{p{5mm}p{120mm}}
+            $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which mission $j$ is active. \\
+            $\mathcal{J}_t $    &  missions $j \in \mathcal{J}$ to be realized in period $t$. \\
+            $\mathcal{I}_j$     &  aircraft $i \in \mathcal{I}$ that can be assigned to mission $j$. \\
+            $\mathcal{I}_k$     &  aircraft $i \in \mathcal{I}$ that are included in cluster $k$. One aircraft can belong to more than one cluster. \\
+            $\mathcal{O}_i$     &  missions $j \in \mathcal{J}$ for which aircraft $i$ can be used. \\
+            $\mathcal{A}^{init}_j$  & aircraft $i \in \mathcal{I}$ that have mission $j$ pre-assigned in the previous period to the start of the planning horizon. \\
         \end{tabular}
+
+    \subsubsection{Time-related parametric sets}
+
+        Several intermediate sets have been defined based on the input data in order to simplify constraint formulation.
+
+        \begin{tabular}{p{8mm}p{117mm}}
+            $\mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
+            $\mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^m-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
+            $\mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^M - M$ such that $t' \in \{ t + M + E^m-1 , ...,  t + M + E^M-1 \}$ (figure \ref{fig:gantt_windows}a). \\
+            $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^M + E^m \}}\}$ (figure \ref{fig:gantt_windows}b). \\
+            $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^M + E^m \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b). \\
+            $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
+            $\mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $(t_1, t_2), \in \mathcal{T}_j$ such that $t_2 \ge t_1 + MT_j-1$. \\
+        \end{tabular}
+
+    \begin{figure}
+        \centering
+        % \includegraphics[width=\linewidth]{img/gantt_timewindows_all.png}
+        \begin{tikzpicture}
+            \input{gantts/time_windows.tex}
+        \end{tikzpicture}
+        \begin{tikzpicture}
+            \input{gantts/time_windows_init.tex}
+        \end{tikzpicture}
+        \begin{tikzpicture}
+            \input{gantts/time_windows_mission.tex}
+        \end{tikzpicture}
+        \caption{Example showing the maintenance-related time-parametric sets for aircraft $i$.
+(a) $\mathcal{T}_{t'}^{s}$ refers to the previous $M$ periods to period $t'$.
+$\mathcal{T}_{t}^{m}$ refers to the periods where a check cannot be planned after starting a check in period $t$.
+Finally, $\mathcal{T}_{t}^{M}$ refers to the periods where a check needs to be scheduled, after starting a check in period $t$.
+(b) $Rct_{i}^{Init}$ is equal to 40, meaning a check has to be planned between months number 10 and 39.
+(c) $\mathcal{T}_{jt}^{MT}$ refers to the periods where the assignment of mission $j$ needs to be kept.
+In this case, the size is 6.}
+        \label{fig:gantt_windows}
+    \end{figure}
 
@@
 
-     The following decision variables define a solution.
+    The following decision variables control the assignment of missions and checks to aircraft.
 
-    \begin{tabular}{p{8mm}p{167mm}}
-        $a_{jti}$   &  =1 if task $j \in J$ in period $t \in \mathcal{T}_j$ is realized with resource $i \in \mathcal{I}_j$, 0 otherwise. \\  
-        $a^s_{jti}$ &  =1 if resource $i$ starts a new assignment to task $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-1)i} = 0$. \\  
-        $m_{it}$    &  =1 if resource $i \in I$ starts a maintenance operation in period $t \in \mathcal{T}$, 0 otherwise. \\
-        % $s_{it}$    &  =1 if resource $i \in I$ starts a storage operation in period $t \in \mathcal{T}$, 0 otherwise. \\
-        $rut_{it}$  &  remaining usage time (continuous) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
-        % $ret_{it}$  &  remaining elapsed time (integer) for resource $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
-        % $u_{max}$   &  maximal number (integer) of unavailable resources in any period. \\
-        % $m_{max}$   &  maximal number (integer) of resources in maintenance in any period. \\
+    \begin{tabular}{p{8mm}p{117mm}}
+        $a_{jtt'i}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ in period $t$ and finishes it at period $t'$. \\  
+        $m_{it}$    &  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise. \\
     \end{tabular}
 
-    Note that $a_{jti}$ and $m_{it}$ are initially set up to 0 for all resources already in maintenance at the beginning of the planning horizon for the remaining time periods of maintenance.
-The remaining usage time for each resource at the beginning of the planning horizon is used to initialize $rut_{i0}$.
+    The following decision variables control the used and remaining flight time in aircraft.
 
-  \subsection{Constraints}
+    \begin{tabular}{p{8mm}p{117mm}}
+        $u_{it}$    &  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$. \\        
+        $rft_{it}$  &  remaining flight time (continuous) for aircraft $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
+    \end{tabular}
 
-    The objective is to minimize the total number of maintenances and, at the same time, maximize the end status of the resources.
+    \paragraph{Fixed values}
+
+    Note that $a_{jti}$ and $m_{it}$ are initially set up to 0 for all aircraft already in maintenance at the beginning of the planning horizon for the remaining time periods of the check.
+$N_{t}$ is calculated based on this information.
+Similarly, for aircraft that have not yet complied with their minimum mission assignment duration at the beginning of the planning horizon, $a_{jti}$ is fixed to comply with the constraints.
+
+\subsection{Objective function and constraints}
+
+    Two objectives have been studied.
+Objective (\ref{eq:objective1}) minimizes the number of checks.
+(\ref{eq:objective2}) combines the first one with the goal of maximizing the final total flight hours potential of the fleet.
 
     \begin{align}
-        & \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H^{M} - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|}
+        & \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it}  \label{eq:objective1}\\
+        & \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H^M - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
     \end{align}
 
-    where the first expression counts all the usage hours given to resources via maintenances and the second expression measures the amount of remaining usage hours in the resources at the end of the planning horizon.
-
-    We want to minimize the total number of maintenances and we want to maximize the final state measured.
-These two objectives have the same units, can be easily compared and they ensure the resource hours are used in the most efficient way possible.
+    The first term counts all the flight hours given to aircraft following checks and the second term quantifies the amount of remaining flight hours for all aircraft at the end of the planning horizon.
+These two objectives have the same units, can be easily compared and ensure the aircraft are used in the most efficient way.
+%We will not penalize the maintenances that provide a better final state for aircraft.
 
@@
         % maximum capacity1:
-        & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq C^{max}
+        & \sum_{t' \in \mathcal{T}^s_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq C^{max}
           & t \in \mathcal{T} \label{eq:capacity1}\\
-        % & \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \leq C^{max} \label{eq:capacity1}\\
-        % availability
-       % & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max} 
-       %  &t \in \mathcal{T} \label{eq:avalaibility1}\\
         % min assignments:
-        & \sum_{i \in \mathcal{I}_j} a_{jti} = R_j
-                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
+        & \sum_{i \in \mathcal{I}_j} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{jt_1t_2i} \geq R_j
+                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:missionres}\\
         % just doing one thing at any given time:
-        & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
+        & \sum_{t' \in \mathcal{T}^s_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{jt_1t_2i} \leq 1 
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
@@
     Maintenance capacity is controlled by (\ref{eq:capacity1}).
-Tasks' resource requirements are defined by (\ref{eq:taskres}).
-Constraints (\ref{eq:state}) guarantee that a resource can be used only for one task or a maintenance operation at the same period.
+The aircraft requirements of missions are defined by (\ref{eq:missionres}).
+Constraints (\ref{eq:state}) ensure that an aircraft can only be used for one mission or undergo check in the same period.
+
+%     \begin{align}
+%         % if start mission => assign start
+%         & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
+%                 & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\
+%         & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_j}
+%                 & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\
+%         & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
+%         & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
+%     \end{align}
+
+%     Constraints (\ref{eq:start1}) captures period $t$ where aircraft $i$ is firstly assigned to mission $j$ i.e.
+it was not assigned to it in period ($t - 1$).
+Constraints (\ref{eq:start2}) are introduced for the first period in the planning horizon.
+% Constraints (\ref{eq:start3}) control the minimum duration of a consecutive mission assignment.
+If aircraft $i$ was firstly assigned to mission $j$ in period $t$, it has to be assigned to it during the following $t' \in \mathcal{T}_{jt}^{MT}$ periods.
+This is a stronger version of the constraint $a_{jt'i}^{s} \leq a_{jti}$.
+
+    % To our knowledge, these constraints have not been taken into account in previous military FMP problems.
 
     \begin{align}
-        % if start task => assign start
-        & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
-                & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\
-        & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_{j}}
-                & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\        
-      % # if we start a task, we need at least X periods of tasks assignments
-        % & a^s_{jt'i} \leq a_{jti}
-        % & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j, t' \in \mathcal{T}^{MT}_{jt} \label{eq:start2} \\
-        % better constraint:
-        & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
-        & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
+       & \sum_{t' \in \mathcal{T}^s_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^K_{kt} \leq A^K_{kt}
+        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster} \\
+       & \sum_{i \in \mathcal{I}_k} rft_{it} \geq H^K_{kt}
+        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
     \end{align}
 
-    Constraints (\ref{eq:start1}) control the definition of starting a consecutive assignment: the variable is active if a resource $i$ is assigned a task $j$ in period $t$ and was not assigned this same task in the previous period ($t-1$).
-Constraints (\ref{eq:start2}) are the particular case for the first period on the horizon where, the previous assignments are given.
-
-    Constraints (\ref{eq:start3}) control the minimum duration of a consecutive task assignment.
-If an assignment to task $j$ is started in period $t$, it needs to continue for the following $t' \in \mathcal{T}^{MT}_{jt}$ periods.
-They are a stronger version of the constraint $ a^s_{jt'i} \leq a_{jti}$.
-
-    \begin{align}
-       & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^{clust}_{kt} \leq A^{clust}_{kt}
-        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\
-       & \sum_{i \in \mathcal{I}_k} rut_{it} \geq H^{clust}_{kt}
-        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster}
-    \end{align}
-
-    Constraints (\ref{eq:avalaibility-cluster}) guarantee a minimum availability of resources for each cluster $k$.
-A cluster is defined by the biggest group of resources that is required, exclusively, by at least one task.
-Constraints (\ref{eq:avalaibility-hours-cluster}) guarantee that a minimum availability in terms of usage houres for each cluster $k$.
+    Constraints (\ref{eq:serviceability-cluster}) guarantee a minimum serviceability of aircraft for each cluster $k$.
+A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission.
+Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.
 
@@
         % remaining used time
-         & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
-        & rut_{i0} = Rut^{Init}_i
-               & i \in \mathcal{I} \label{eq:rut_initial}\\
-        & rut_{it} \geq H^{M} m_{it'}
-                & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\ 
-        & rut_{it} \in [0,H^{M}]
+         & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{jt_1t_2i} H_j 
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
+         & u_{it} \geq U^{min} (1 - \sum_{t' \in \mathcal{T}^s_t} m_{it'})
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight2}\\
+         & u_{it} \in [0, \max_j{\{H_j\}}]
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight_lower}\\
+         & rft_{it} \leq rft_{i(t-1)} + H^M m_{it} - u_{it}
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rft_upper}\\
+        & rft_{i0} = Rft^{Init}_i
+               & i \in \mathcal{I} \label{eq:rft_initial}\\
+        & rft_{it} \geq H^M m_{it'}
+                & t \in \mathcal{T}, t' \in \mathcal{T}^s_t, i \in \mathcal{I}\label{eq:rft_lower}\\ 
+        & rft_{it} \in [0,H^M]
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
-        % & ret_{it} \leq ret_{it-1} - 1 + \sum_{t' \in \mathcal{T}^{st}_t} s_{it'} + E m_{it}
-        %         & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:ret_upper}\\
-        % & ret_{i0} = Ret^{Init}_i
-        %         & i \in \mathcal{I} \label{eq:ret_initial}\\
-        % & ret_{it} \geq E m_{it}
-        %         & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_lower}\\                 
-        % & ret_{it} \in [0,E]
-        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:me}\\
-        % & \sum_{i \in \mathcal{I}} ret_{it} \geq Ret^{Init}_{sum}
-        %       & t = |\mathcal{T}| \label{eq:min_ret}\\
-        % & \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|} \geq Rut^{Init}_{sum}\\
     \end{align}
 
-        % These constraints calculate the balances of hours for each resource.
-    The remaining usage time is defined by (\ref{eq:rut_upper})-(\ref{eq:rut_initial}) and its limits by (\ref{eq:rut_lower})-(\ref{eq:mu}).
+        % These constraints calculate the balances of hours for each aircraft.
+    The flight time per aircraft and period is calculated in (\ref{eq:flight1})-(\ref{eq:flight_lower}).
+The remaining flight time is defined by (\ref{eq:rft_upper})-(\ref{eq:rft_initial}) and its limits by (\ref{eq:rft_lower})-(\ref{eq:mu}).
 
@@
         & m_{it'} + m_{it} \leq 1
-          & t \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
-        & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
-          & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\
+          & t \in \mathcal{T}, t' \in \mathcal{T}^m_t, i \in \mathcal{I}\label{eq:rct_min}\\ 
+        & \sum_{t' \in \mathcal{T}^M_t} m_{it'} \geq  m_{it}
+          & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
         & m_{it} = 0
-          & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I}\label{eq:ret_min_init} \\
+          & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I}\label{eq:rct_min_init} \\
         & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
-          & i \in \mathcal{I}\label{eq:ret_max_init}
+          & i \in \mathcal{I}\label{eq:rct_max_init}
     \end{align}
 
-    The minimum and maximum elapsed time are defined by (\ref{eq:ret_min}) and (\ref{eq:ret_max}) respectively.
-Constraints (\ref{eq:ret_min}) guarantee that if a maintenance is done in some period $t$, we know that another one cannot be done in the immediately consecutive $t' \in \mathcal{T}^{m}_t$ periods.
-Constraints (\ref{eq:ret_max}) guarantee that, if we have a maintenance in period $t$, we need to start at least one maintenance among periods $t' \in \mathcal{T}^{M}_t$.
-Constraints (\ref{eq:ret_min_init}) and (\ref{eq:ret_max_init}) control the minimum and maximum remaining elapsed time at the beginning of the planning period, respectively.
-They follow the same logic as constraints (\ref{eq:ret_min}) and (\ref{eq:ret_max}), respectively, for the particular case of the start of the planning horizon.
+    The minimum and maximum calendar times are defined by (\ref{eq:rct_min}) and (\ref{eq:rct_max}) respectively.
+Constraints (\ref{eq:rct_min}) guarantee that if a check is done in some period $t$, we know that another one cannot be done in the immediately consecutive $t' \in \mathcal{T}_{t}^{m}$ periods.
+Constraints (\ref{eq:rct_max}) ensure that if a check is planned in period $t$, we need to start at least one check in periods $t' \in \mathcal{T}_{t}^{M}$.
+Constraints (\ref{eq:rct_min_init}) and (\ref{eq:rct_max_init}) control the minimum and maximum remaining calendar times respectively at the beginning of the planning period.
+They follow the same logic as constraints (\ref{eq:rct_min}) and (\ref{eq:rct_max}), respectively.
 
-    % Finally, constraints \ref{eq:min_rut} guarantee that resources have, globally, the same amount of remaining used and elapsed times at the beginning and at the end of the planning horizon.
+    To our knowledge, these constraints have not been taken into account in previous military FMP problems.
 

### 2019-03-26 latex/notes/model/model.tex
changed formulation in model.tex

@@
             $\mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
+            % $\mathcal{T}^S_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t},  ..., \min{\{|\mathcal{T}|, t + M-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
             $\mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^m-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
@@
             $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
-            $\mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $(t_1, t_2), \in \mathcal{T}_j$ such that $t_2 \ge t_1 + MT_j-1$. \\
+            $\mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{T}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
+            $\mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$. \\
         \end{tabular}
@@
     \begin{tabular}{p{8mm}p{117mm}}
-        $a_{jtt'i}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ in period $t$ and finishes it at period $t'$. \\  
-        $m_{it}$    &  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise. \\
+        $a_{ijtt'}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ in period $t$ and finishes it at period $t'$. \\  
+        $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$ and then starts the next one at period $t'$, 0 otherwise. \\
     \end{tabular}
@@
 
-    Note that $a_{jti}$ and $m_{it}$ are initially set up to 0 for all aircraft already in maintenance at the beginning of the planning horizon for the remaining time periods of the check.
+    Note that $a_{ijtt'}$ and $m_{itt'}$ are initially set up to 0 for all aircraft already in maintenance at the beginning of the planning horizon for the remaining time periods of the check.
 $N_{t}$ is calculated based on this information.
-Similarly, for aircraft that have not yet complied with their minimum mission assignment duration at the beginning of the planning horizon, $a_{jti}$ is fixed to comply with the constraints.
+Similarly, for aircraft that have not yet complied with their minimum mission assignment duration at the beginning of the planning horizon, $a_{ijtt'}$ is fixed to comply with the constraints.
 
@@
 
+    % TODO: do not count maints that start at the end
     \begin{align}
-        & \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it}  \label{eq:objective1}\\
-        & \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H^M - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
+        & \text{Min}\; \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}, i \in \mathcal{I}} m_{it_1t_2}  \label{eq:objective1}\\
+        & \text{Min}\; \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}, i \in \mathcal{I}} m_{it_1t_2} \times H^M - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
     \end{align}
@@
         % maximum capacity1:
-        & \sum_{t' \in \mathcal{T}^s_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq C^{max}
+        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} \sum_{i \in \mathcal{I}} m_{it_1t_2} + N_t \leq C^{max}
           & t \in \mathcal{T} \label{eq:capacity1}\\
         % min assignments:
-        & \sum_{i \in \mathcal{I}_j} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{jt_1t_2i} \geq R_j
+        & \sum_{i \in \mathcal{I}_j} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \geq R_j
                 & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:missionres}\\
         % just doing one thing at any given time:
-        & \sum_{t' \in \mathcal{T}^s_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{jt_1t_2i} \leq 1 
+        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \leq 1
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
@@
     \begin{align}
-       & \sum_{t' \in \mathcal{T}^s_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^K_{kt} \leq A^K_{kt}
+       & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} \sum_{i \in \mathcal{I}_k} m_{it_1t_2} + N^K_{kt} \leq A^K_{kt}
         &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster} \\
@@
         % remaining used time
-         & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{jt_1t_2i} H_j 
+         & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} H_j 
             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
-         & u_{it} \geq U^{min} (1 - \sum_{t' \in \mathcal{T}^s_t} m_{it'})
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight2}\\
-         & u_{it} \in [0, \max_j{\{H_j\}}]
+         & u_{it} \in [U^{min}, \max_j{\{H_j\}}]
             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight_lower}\\
-         & rft_{it} \leq rft_{i(t-1)} + H^M m_{it} - u_{it}
+         & rft_{it} \leq rft_{i(t-1)} + H^M \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} - u_{it}
             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rft_upper}\\
@@
                & i \in \mathcal{I} \label{eq:rft_initial}\\
-        & rft_{it} \geq H^M m_{it'}
-                & t \in \mathcal{T}, t' \in \mathcal{T}^s_t, i \in \mathcal{I}\label{eq:rft_lower}\\ 
+        & rft_{it} \geq H^M m_{it_1t_2}
+                &t \in \mathcal{T}, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}, i \in \mathcal{I}\label{eq:rft_lower}\\
         & rft_{it} \in [0,H^M]
@@
     \begin{align}
-        & m_{it'} + m_{it} \leq 1
-          & t \in \mathcal{T}, t' \in \mathcal{T}^m_t, i \in \mathcal{I}\label{eq:rct_min}\\ 
-        & \sum_{t' \in \mathcal{T}^M_t} m_{it'} \geq  m_{it}
-          & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
-        & m_{it} = 0
-          & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I}\label{eq:rct_min_init} \\
-        & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
+        % & m_{it'} + m_{it} \leq 1
+        %   & t \in \mathcal{T}, t' \in \mathcal{T}^m_t, i \in \mathcal{I}\label{eq:rct_min}\\ 
+        % & \sum_{t' \in \mathcal{T}^M_t} m_{it'} \geq  m_{it}
+        %   & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
+        & m_{it_1t_2} = 0
+          & t_1 \in \mathcal{T}^{m_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:rct_min_init} \\
+        & \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\} } m_{it_1t_2} \geq  1 
           & i \in \mathcal{I}\label{eq:rct_max_init}
@@
 
-    The minimum and maximum calendar times are defined by (\ref{eq:rct_min}) and (\ref{eq:rct_max}) respectively.
-Constraints (\ref{eq:rct_min}) guarantee that if a check is done in some period $t$, we know that another one cannot be done in the immediately consecutive $t' \in \mathcal{T}_{t}^{m}$ periods.
-Constraints (\ref{eq:rct_max}) ensure that if a check is planned in period $t$, we need to start at least one check in periods $t' \in \mathcal{T}_{t}^{M}$.
     Constraints (\ref{eq:rct_min_init}) and (\ref{eq:rct_max_init}) control the minimum and maximum remaining calendar times respectively at the beginning of the planning period.
-They follow the same logic as constraints (\ref{eq:rct_min}) and (\ref{eq:rct_max}), respectively.
-
-    To our knowledge, these constraints have not been taken into account in previous military FMP problems.
 

### 2019-03-28 latex/notes/model/model.tex
update model.

@@
     \begin{align}
+       % For now, we only want one or possibly two maintenances...
+        & \sum_{t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\} } m_{it_1t_2} =  1 
+          & i \in \mathcal{I}\label{eq:num_maint} \\
         % & m_{it'} + m_{it} \leq 1
@@
           & t_1 \in \mathcal{T}^{m_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:rct_min_init} \\
-        & \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\} } m_{it_1t_2} \geq  1 
+        & \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\} } m_{it_1t_2} =  1 
           & i \in \mathcal{I}\label{eq:rct_max_init}

### 2019-03-30 latex/notes/model/model.tex
Maintenances reformulation. Implementation still incomplete.

@@
     \begin{tabular}{p{8mm}p{117mm}}
-        $a_{ijtt'}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ in period $t$ and finishes it at period $t'$. \\  
-        $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$ and then starts the next one at period $t'$, 0 otherwise. \\
+        $a_{ijtt'}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ at the beginning of period $t$ and finishes it at the end of period $t'$. \\  
+        $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check at the beginning of period $t \in \mathcal{T}$ and then starts the next one at the beginning of period $t'$, 0 otherwise. \\
     \end{tabular}
@@
                & i \in \mathcal{I} \label{eq:rft_initial}\\
-        & rft_{it} \geq H^M m_{it_1t_2}
-                &t \in \mathcal{T}, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}, i \in \mathcal{I}\label{eq:rft_lower}\\
+        & rft_{it} \geq H^M \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2}
+                &t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rft_lower}\\
         & rft_{it} \in [0,H^M]

### 2019-03-30 python/package/model.py

@@ -52,14 +52,15 @@ class Model(exp.Experiment):
         # binary:
         start_T = pl.LpVariable.dicts(name="start_T", indexs=l['avtt2'], lowBound=0, upBound=1, cat=pl.LpInteger)
         # task = pl.LpVariable.dicts(name="task", indexs=l['avt'], lowBound=0, upBound=1, cat=pl.LpInteger)
-        start_M = pl.LpVariable.dicts(name="start_M", indexs=l['at_start'], lowBound=0, upBound=1, cat=pl.LpInteger)
+        start_M = pl.LpVariable.dicts(name="start_M", indexs=l['att_maints'], lowBound=0, upBound=1, cat=pl.LpInteger)
+        # start_M = pl.LpVariable.dicts(name="start_M", indexs=l['at_start'], lowBound=0, upBound=1, cat=pl.LpInteger)
 
         # numeric:
         rut = pl.LpVariable.dicts(name="rut", indexs=l['at0'], lowBound=0, upBound=ub['rut'], cat=var_type)
-        usage = pl.LpVariable.dicts(name="usage", indexs=l['at'], lowBound=0, upBound=ub['used_max'], cat=var_type)
+        usage = pl.LpVariable.dicts(name="usage", indexs=l['at'], lowBound=min_usage, upBound=ub['used_max'], cat=var_type)
 
         # objective function:
-        num_maint = pl.LpVariable(name="num_maint", lowBound=0, upBound=ub['num_maint'], cat=var_type)
+        # num_maint = pl.LpVariable(name="num_maint", lowBound=0, upBound=ub['num_maint'], cat=var_type)
         rut_obj_var = pl.LpVariable(name="rut_obj_var", lowBound=0, upBound=ub['rut_end'], cat=var_type)
 
         if options.get('mip_start'):
@@ -109,12 +110,10 @@ class Model(exp.Experiment):
                     if (a, t) in rut:
                         rut[a, t].setInitialValue(v)
 
-            num_maint.setInitialValue(number_maint)
-
             if options.get('fix_start', False):
                 # vars_to_fix = [start_M]
                 # vars_to_fix = [start_T, task, start_M, rut, usage, {0: rut_obj_var}, {0: num_maint}]
-                vars_to_fix = [start_T, start_M, rut, usage, {0: rut_obj_var}, {0: num_maint}]
+                vars_to_fix = [start_T, start_M, rut, usage, {0: rut_obj_var}]
                 for _vars in vars_to_fix:
[…]

### 2019-04-04 latex/notes/model/model.tex
changed formulation of model in latex.

@@
 
-    \subsubsection{Parameters}
+    \subsubsection{Mission parameters}
 
@@
             $M$               & check duration in number of periods. \\
+        \end{tabular}
+
+    \subsubsection{Maintenance parameters}
+
+        \begin{tabular}{p{10mm}p{115mm}}
             $C^{max}$         & maximum number of simultaneous checks. \\
@@
             $H^M$               & remaining flight time after a check. \\
+        \end{tabular}
+
+    \subsubsection{Fleet parameters}
+
+        \begin{tabular}{p{10mm}p{115mm}}
             $N_t$             & number of aircraft known to be in maintenance in period $t$.\\
@@
         \begin{tikzpicture}
+            \input{gantts/time_windows2.tex}
+        \end{tikzpicture}
+        \begin{tikzpicture}
             \input{gantts/time_windows_init.tex}
@@
         \end{tikzpicture}
-        \caption{Example showing the maintenance-related time-parametric sets for aircraft $i$.
+        \caption{Examples showing the maintenance-related time-parametric sets for aircraft $i$.
 (a) $\mathcal{T}_{t'}^{s}$ refers to the previous $M$ periods to period $t'$.
@@
 Finally, $\mathcal{T}_{t}^{M}$ refers to the periods where a check needs to be scheduled, after starting a check in period $t$.
-(b) $Rct_{i}^{Init}$ is equal to 40, meaning a check has to be planned between months number 10 and 39.
-(c) $\mathcal{T}_{jt}^{MT}$ refers to the periods where the assignment of mission $j$ needs to be kept.
-In this case, the size is 6.}
+(b) The maintenance assignment for aicraft $i$ is $m_{it_1t_2}=1$.
+$\mathcal{T}_A \cup \mathcal{T}_B$ represents the set of periods $t$ for which $(t_1, t_2) \in \mathcal{TTT}_t$.
+(c) $Rct_{i}^{Init}$ is equal to 55, meaning a check has to be planned between months number 25 and 54.
+The maintenance assignment for aicraft $i$ is $m_{it_1t_2}=1$ and, since it includes the last period, it only represents one operation.
+(d) $\mathcal{T}_{jt}^{MT}$ refers to the periods where the assignment of mission $j$ needs to be kept.
+In this case, the size is 6.
+The mission assignments for aicraft $i$ to mission $j$ are $a_{ijt_1t_2}=1$ and $a_{ijt'_1t'_2}=1$}
         \label{fig:gantt_windows}
@@
 
-    % TODO: do not count maints that start at the end
     \begin{align}
-        & \text{Min}\; \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}, i \in \mathcal{I}} m_{it_1t_2}  \label{eq:objective1}\\
-        & \text{Min}\; \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}, i \in \mathcal{I}} m_{it_1t_2} \times H^M - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
+        & \text{Min}\; \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}} m_{it_1t_2}  \label{eq:objective1}\\
+        & \text{Min}\; \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}} m_{it_1t_2} \times H^M - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
     \end{align}
@@
        % For now, we only want one or possibly two maintenances...
-        & \sum_{t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\} } m_{it_1t_2} =  1 
+        & \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\} } m_{it_1t_2} =  1 
           & i \in \mathcal{I}\label{eq:num_maint} \\
@@
         %   & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
+        % & m_{it_1t_2} = 0
+        %   & t_1 \in \mathcal{T}^{m_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I}\label{eq:rct_min_init} \\
         & m_{it_1t_2} =  0
-          & t_1 \in \mathcal{T}^{m_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:rct_min_init} \\
-        & \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\} } m_{it_1t_2} =  1 
-          & i \in \mathcal{I}\label{eq:rct_max_init}
+          & t_1 \notin \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}, i \in \mathcal{I} \label{eq:maint_forbid} \\
+        & m_{it_1t_2} =  0
+          & t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \notin \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I} \label{eq:maint_forbid2}
     \end{align}
 
-    Constraints (\ref{eq:rct_min_init}) and (\ref{eq:rct_max_init}) control the minimum and maximum remaining calendar times respectively at the beginning of the planning period.
+    Constraints (\ref{eq:num_maint}) limit the number of maintenance assignments per aircraft to one.
+This implies that each aircraft will get strictly one or two maintenances over the whole planning horizon.
+
+    Constraints (\ref{eq:maint_forbid} - \ref{eq:maint_forbid2}) limit the combinations of maintenance schedules that can be done during the horizon assuming that the first maintenance is dependant in the initial state of the aircraft and the second maintenance (if any) is limited by the moment at which the first maintenance was planned.
 

### 2019-04-09 latex/notes/model/model.tex
renamed some sets and params according to Robert's comments. Corrected some minor errors.

@@
 
-        \begin{tabular}{p{5mm}p{120mm}}
-            $\mathcal{I}$     &  aircraft. \\
-            $\mathcal{T}$     &  time periods included in the planning horizon. \\
-            $\mathcal{J}$     &  missions. \\
-            $\mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
+        \begin{tabular}{p{15mm}p{110mm}}
+            $i \in \mathcal{I}$     &  aircraft. \\
+            $t \in \mathcal{T}$     &  time periods included in the planning horizon. \\
+            $j \in \mathcal{J}$     &  missions. \\
+            $k \in \mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
         \end{tabular}
@@
 
-        \begin{tabular}{p{10mm}p{115mm}}
-            $H_j$             & amount of flight time required by mission $j$. \\
-            $R_j$             & number of aircraft required by mission $j$. \\
-            $MT_j$            & minimum number of consecutive periods an aircraft has to be assigned to mission $j$. \\
-            $U^{min}$       & default aircraft flight time if it is not assigned to any mission nor in maintenance.\\
-            $M$               & check duration in number of periods. \\
+        \begin{tabular}{p{15mm}p{95mm}p{15mm}}
+            $H_j$             & amount of flight time required by mission $j$. & hours. \\
+            $R_j$             & number of aircraft required by mission $j$. & aircraft. \\
+            $MT_j$            & minimum number of consecutive periods an aircraft has to be assigned to mission $j$. & periods. \\
+            $U^{min}$         & default aircraft flight time if it is not assigned to any mission nor in maintenance.& hours \\
+            $M$               & check duration in number of periods. & periods. \\
         \end{tabular}
@@
 
-        \begin{tabular}{p{10mm}p{115mm}}
-            $C^{max}$         & maximum number of simultaneous checks. \\
-            $E^M$         & maximum number of periods between two consecutive checks. \\
-            $E^m$         & minimum number of periods between two consecutive checks. \\
-            $H^M$               & remaining flight time after a check. \\
+        \begin{tabular}{p{15mm}p{95mm}p{15mm}}
+            $C^{max}$         & maximum number of simultaneous checks. & aircraft. \\
+            $E^{max}$         & maximum number of periods between two consecutive checks. & periods. \\
+            $E^{min}$         & minimum number of periods between two consecutive checks. & periods. \\
+            $H^{max}$         & remaining flight time after a check. & hours. \\
         \end{tabular}
@@
 
-        \begin{tabular}{p{10mm}p{115mm}}
-            $N_t$             & number of aircraft known to be in maintenance in period $t$.\\
-            $N^K_{kt}$         & number of aircraft in cluster $k$ known to be in maintenance in period $t$.\\
-            $A^K_{kt}$         & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$.\\
-            $H^K_{kt}$         & minimum number of total remaining flight time for cluster $k$ at period $t$.\\
-            $Rft^{Init}_i$  & remaining flight time for aircraft $i$ at the start of the planning horizon. \\
-            $Rct^{Init}_i$  & remaining calendar time for aircraft $i$ at the start of the planning horizon. \\
+        \begin{tabular}{p{15mm}p{95mm}p{15mm}}
+            $N_t$               & number of aircraft known to be in maintenance in period $t$. & aircraft. \\
+            $N^{Clust}_{kt}$    & number of aircraft in cluster $k$ known to be in maintenance in period $t$. & aircraft. \\
+            $A^{Clust}_{kt}$    & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$. & aircraft. \\
+            $H^{Clust}_{kt}$    & minimum number of total remaining flight time for cluster $k$ at period $t$. & hours. \\
+            $Rft^{Init}_i$      & remaining flight time for aircraft $i$ at the start of the planning horizon.  & hours.  \\
+            $Rct^{Init}_i$      & remaining calendar time for aircraft $i$ at the start of the planning horizon.  & periods. \\
         \end{tabular}
@@
 
-        \begin{tabular}{p{5mm}p{120mm}}
-            $\mathcal{T}_j$     &  time periods $t \in \mathcal{T}$ in which mission $j$ is active. \\
-            $\mathcal{J}_t $    &  missions $j \in \mathcal{J}$ to be realized in period $t$. \\
-            $\mathcal{I}_j$     &  aircraft $i \in \mathcal{I}$ that can be assigned to mission $j$. \\
-            $\mathcal{I}_k$     &  aircraft $i \in \mathcal{I}$ that are included in cluster $k$. One aircraft can belong to more than one cluster. \\
-            $\mathcal{O}_i$     &  missions $j \in \mathcal{J}$ for which aircraft $i$ can be used. \\
-            $\mathcal{A}^{init}_j$  & aircraft $i \in \mathcal{I}$ that have mission $j$ pre-assigned in the previous period to the start of the planning horizon. \\
+        \begin{tabular}{p{15mm}p{110mm}}
+            $t \in \mathcal{TJ}_j$     &  time periods $t \in \mathcal{T}$ in which mission $j$ is active. \\
+            $j \in \mathcal{JT}_t $    &  missions $j \in \mathcal{J}$ to be realized in period $t$. \\
+            $i \in \mathcal{IJ}_j$     &  aircraft $i \in \mathcal{I}$ that can be assigned to mission $j$. \\
+            $i \in \mathcal{IK}_k$     &  aircraft $i \in \mathcal{I}$ that are included in cluster $k$. One aircraft can belong to more than one cluster. \\
+            $j \in \mathcal{JI}_i$     &  missions $j \in \mathcal{J}$ for which aircraft $i$ can be used. \\
+            $i \in \mathcal{A}^{Init}_j$  & aircraft $i \in \mathcal{I}$ that have mission $j$ pre-assigned in the previous period to the start of the planning horizon. \\
         \end{tabular}
@@
 
-        \begin{tabular}{p{8mm}p{117mm}}
-            $\mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
+        \begin{tabular}{p{30mm}p{95mm}}
+            $t \in \mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
             % $\mathcal{T}^S_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t},  ..., \min{\{|\mathcal{T}|, t + M-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
-            $\mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^m-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
-            $\mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^M - M$ such that $t' \in \{ t + M + E^m-1 , ...,  t + M + E^M-1 \}$ (figure \ref{fig:gantt_windows}a). \\
-            $\mathcal{T}^{m_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^M + E^m \}}\}$ (figure \ref{fig:gantt_windows}b). \\
-            $\mathcal{T}^{M_{ini}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^M + E^m \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b). \\
-            $\mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
-            $\mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{T}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
-            $\mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$. \\
+            $t' \in \mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^{min}-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
+            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$ (figure \ref{fig:gantt_windows}a). \\
+            $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}b). \\
+            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b).
+\\
+            $t' \in \mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
+            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
+            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$.
+\\
         \end{tabular}
@@
     \begin{align}
-        & \text{Min}\; \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}} m_{it_1t_2}  \label{eq:objective1}\\
-        & \text{Min}\; \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}} m_{it_1t_2} \times H^M - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
+        & \text{Min}\; \sum_{i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|} m_{it_1t_2}  \label{eq:objective1}\\
+        & \text{Min}\; \sum_{i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1}} m_{it_1t_2} \times H^{max} - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
     \end{align}
@@
         % maximum capacity1:
-        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} \sum_{i \in \mathcal{I}} m_{it_1t_2} + N_t \leq C^{max}
+        & \sum_{i \in \mathcal{I}, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + N_t \leq C^{max}
           & t \in \mathcal{T} \label{eq:capacity1}\\
         % min assignments:
-        & \sum_{i \in \mathcal{I}_j} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \geq R_j
-                & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:missionres}\\
+        & \sum_{i \in \mathcal{IJ}_j, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \geq R_j
+                & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
         % just doing one thing at any given time:
-        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \leq 1
+        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \leq 1
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
@@
 %         & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
-%                 & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\
-%         & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_j}
-%                 & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\
+%                 & t =1, ..., \mathcal{T}, j \in \mathcal{JT}_t, i \in \mathcal{IJ}_j \label{eq:start1} \\
+%         & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{Init}_j}
+%                 & j \in \mathcal{J}_0, i \in \mathcal{IJ}_j \label{eq:start2} \\
 %         & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
-%         & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
+%         & j \in \mathcal{J}, t \in \mathcal{TJ}_j, i \in \mathcal{IJ}_j \label{eq:start3}
 %     \end{align}
@@
     \begin{align}
-       & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} \sum_{i \in \mathcal{I}_k} m_{it_1t_2} + N^K_{kt} \leq A^K_{kt}
+       & \sum_{i \in \mathcal{IK}_k, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt}
         &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster} \\
-       & \sum_{i \in \mathcal{I}_k} rft_{it} \geq H^K_{kt}
+       & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
         &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
@@
         % remaining used time
-         & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} H_j 
+         & u_{it} \geq \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} H_j 
             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
@@
             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight_lower}\\
-         & rft_{it} \leq rft_{i(t-1)} + H^M \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} - u_{it}
+         & rft_{it} \leq rft_{i(t-1)} + H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} - u_{it}
             & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rft_upper}\\
@@
                & i \in \mathcal{I} \label{eq:rft_initial}\\
-        & rft_{it} \geq H^M \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2}
+        & rft_{it} \geq H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2}
                 &t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rft_lower}\\
-        & rft_{it} \in [0,H^M]
+        & rft_{it} \in [0,H^{max}]
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
@@
        % For now, we only want one or possibly two maintenances...
-        & \sum_{t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\} } m_{it_1t_2} =  1 
+        & \sum_{t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup |\mathcal{T}| } m_{it_1t_2} =  1 
           & i \in \mathcal{I}\label{eq:num_maint} \\
@@
         % & m_{it_1t_2} = 0
-        %   & t_1 \in \mathcal{T}^{m_{ini}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I}\label{eq:rct_min_init} \\
+        %   & t_1 \in \mathcal{T}^{m_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I}\label{eq:rct_min_init} \\
         & m_{it_1t_2} =  0
-          & t_1 \notin \mathcal{T}^{M_{ini}}_i, t_2 \in \mathcal{T}, i \in \mathcal{I} \label{eq:maint_forbid} \\
+          & t_1 \notin \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}, i \in \mathcal{I} \label{eq:maint_forbid} \\
         & m_{it_1t_2} =  0
-          & t_1 \in \mathcal{T}^{M_{ini}}_i, t_2 \notin \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I} \label{eq:maint_forbid2}
+          & t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \notin \mathcal{T}^M_{t_1} \cup |\mathcal{T}|, i \in \mathcal{I} \label{eq:maint_forbid2}
     \end{align}

### 2019-04-12 latex/notes/model/model.tex
reformulation of flight hours in latex.

@@
 
-    The following decision variables control the used and remaining flight time in aircraft.
+    % The following decision variables control the used and remaining flight time in aircraft.
 
-    \begin{tabular}{p{8mm}p{117mm}}
-        $u_{it}$    &  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$. \\        
-        $rft_{it}$  &  remaining flight time (continuous) for aircraft $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
-    \end{tabular}
+    % \begin{tabular}{p{8mm}p{117mm}}
+    %     $u_{it}$    &  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$. \\        
+    %     $rft_{it}$  &  remaining flight time (continuous) for aircraft $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
+    % \end{tabular}
 
@@
         & \text{Min}\; \sum_{i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|} m_{it_1t_2}  \label{eq:objective1}\\
-        & \text{Min}\; \sum_{i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1}} m_{it_1t_2} \times H^{max} - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
+        % & \text{Min}\; \sum_{i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1}} m_{it_1t_2} \times H^{max} - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
     \end{align}
@@
         &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster} \\
-       & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
-        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
+       % & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
+       %  &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
     \end{align}
@@
         % remaining used time
-         & u_{it} \geq \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} H_j 
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
-         & u_{it} \in [U^{min}, \max_j{\{H_j\}}]
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight_lower}\\
-         & rft_{it} \leq rft_{i(t-1)} + H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} - u_{it}
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rft_upper}\\
-        & rft_{i0} = Rft^{Init}_i
-               & i \in \mathcal{I} \label{eq:rft_initial}\\
-        & rft_{it} \geq H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2}
-                &t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rft_lower}\\
-        & rft_{it} \in [0,H^{max}]
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
+        %  & u_{it} \geq \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} H_j 
+        %     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
+        %  & u_{it} \in [U^{min}, \max_j{\{H_j\}}]
+        %     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight_lower}\\
+        %  & rft_{it} \leq rft_{i(t-1)} + H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} - u_{it}
+        %     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rft_upper}\\
+        % & rft_{i0} = Rft^{Init}_i
+        %        & i \in \mathcal{I} \label{eq:rft_initial}\\
+        % & rft_{it} \geq H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2}
+        %         &t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rft_lower}\\
+        % & rft_{it} \in [0,H^{max}]
+        %         & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
+        & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{a1t_1}} a_{ijtt'} H'_{jtt'} + U^{min}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
+                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours1}\\
+        & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{at_1t_2}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours2}\\
+        & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{at_2|\mathcal{T}|}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours3}
     \end{align}
@@
         % These constraints calculate the balances of hours for each aircraft.
-    The flight time per aircraft and period is calculated in (\ref{eq:flight1})-(\ref{eq:flight_lower}).
+    % The flight time per aircraft and period is calculated in (\ref{eq:flight1})-(\ref{eq:flight_lower}).
 The remaining flight time is defined by (\ref{eq:rft_upper})-(\ref{eq:rft_initial}) and its limits by (\ref{eq:rft_lower})-(\ref{eq:mu}).

### 2019-04-12 latex/notes/model/model.tex

--- 2019-04-09 latex/notes/model/model.tex
@@
 
-The following model provides a tight MIP formulation that solves the Military Flight and Maintenance Problem described in section \ref{sec:problem}.
-Maintenance operations will be referred as "checks".
+The following model provides a tight MIP formulation that solves the Military Flight and Maintenance Problem. Maintenance operations will be referred as "checks".
 
@@
 
-        \begin{tabular}{p{15mm}p{110mm}}
+        \begin{tabular}{p{15mm}p{140mm}}
             $i \in \mathcal{I}$     &  aircraft. \\
@@
 
-        \begin{tabular}{p{15mm}p{95mm}p{15mm}}
+        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
             $H_j$             & amount of flight time required by mission $j$. & hours. \\
@@
 
-        \begin{tabular}{p{15mm}p{95mm}p{15mm}}
+        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
             $C^{max}$         & maximum number of simultaneous checks. & aircraft. \\
@@
 
-        \begin{tabular}{p{15mm}p{95mm}p{15mm}}
+        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
             $N_t$               & number of aircraft known to be in maintenance in period $t$. & aircraft. \\
@@
 
-        \begin{tabular}{p{15mm}p{110mm}}
+        \begin{tabular}{p{15mm}p{140mm}}
             $t \in \mathcal{TJ}_j$     &  time periods $t \in \mathcal{T}$ in which mission $j$ is active. \\
-            $j \in \mathcal{JT}_t $    &  missions $j \in \mathcal{J}$ to be realized in period $t$. \\
+            $j \in \mathcal{JT}_t$    &  missions $j \in \mathcal{J}$ to be realized in period $t$. \\
             $i \in \mathcal{IJ}_j$     &  aircraft $i \in \mathcal{I}$ that can be assigned to mission $j$. \\
@@
 
-        \begin{tabular}{p{30mm}p{95mm}}
+        \begin{tabular}{p{30mm}p{125mm}}
             $t \in \mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
@@
 \\
+            $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT_j-1$ and $t \ge t_1 + M$ and $t' < t_2$ \\
         \end{tabular}
@@
 (c) $Rct_{i}^{Init}$ is equal to 55, meaning a check has to be planned between months number 25 and 54.
-The maintenance assignment for aicraft $i$ is $m_{it_1t_2}=1$ and, since it includes the last period, it only represents one operation.
+The maintenance assignment for aircraft $i$ is $m_{it_1t_2}=1$ and, since it includes the last period, it only represents one operation.
 (d) $\mathcal{T}_{jt}^{MT}$ refers to the periods where the assignment of mission $j$ needs to be kept.
 In this case, the size is 6.
-The mission assignments for aicraft $i$ to mission $j$ are $a_{ijt_1t_2}=1$ and $a_{ijt'_1t'_2}=1$}
+The mission assignments for aircraft $i$ to mission $j$ are $a_{ijt_1t_2}=1$ and $a_{ijt'_1t'_2}=1$}
         \label{fig:gantt_windows}
     \end{figure}
+
+    \subsubsection{Consumption derived parameters}\label{subsubsec:consumption}
+
+    The following parameters are derived from the basic input parameters.
+They are used in constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_rem_hours3}).
+They are transformations of the default and mission flight hours consumption under different time circumstances.
+
+    \begin{align}
+        & U^{min}_{t_1t_2} = U^{min} \sum_{t=t_1}^{t_2} (t_2 - t) \notag \\
+        & H'_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1) \notag \\
+        & H''_{jt_1t_2t'} = \sum_{t=t_1}^{t_2} (H_j- U^{min}) (t' - t) \notag
+    \end{align}
 
@@
 
-    \begin{tabular}{p{8mm}p{117mm}}
+    \begin{tabular}{p{8mm}p{147mm}}
         $a_{ijtt'}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ at the beginning of period $t$ and finishes it at the end of period $t'$. \\  
         $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check at the beginning of period $t \in \mathcal{T}$ and then starts the next one at the beginning of period $t'$, 0 otherwise. \\
-    \end{tabular}
-
-    The following decision variables control the used and remaining flight time in aircraft.
-
-    \begin{tabular}{p{8mm}p{117mm}}
-        $u_{it}$    &  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$. \\        
-        $rft_{it}$  &  remaining flight time (continuous) for aircraft $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
+        $h_{itt'}$  &  Continuous. Sum of the monthly remaining flight hours for aircraft $i \in I$ for periods between the beginning of period $t$ and the beginning of period $t'$. \\
     \end{tabular}
@@
 
-    Two objectives have been studied.
-Objective (\ref{eq:objective1}) minimizes the number of checks.
-(\ref{eq:objective2}) combines the first one with the goal of maximizing the final total flight hours potential of the fleet.
+    Objective (\ref{eq:objective1}) minimizes the number of checks while at the same time maximizing the monthly total flight hours potential of the fleet.
 
     \begin{align}
-        & \text{Min}\; \sum_{i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|} m_{it_1t_2}  \label{eq:objective1}\\
-        & \text{Min}\; \sum_{i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1}} m_{it_1t_2} \times H^{max} - \sum_{i \in \mathcal{I}} rft_{i|\mathcal{T}|} \label{eq:objective2}
+        & \text{Min}\; \sum_{\substack{i \in \mathcal{I}, \\ t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|}} m_{it_1t_2} \times H^{max} \times (t_2 - t_1) - h_{it_1t_2}  \label{eq:objective1}
     \end{align}
 
-    The first term counts all the flight hours given to aircraft following checks and the second term quantifies the amount of remaining flight hours for all aircraft at the end of the planning horizon.
+    The first term counts all the remaining flight hours given to aircraft following checks and the second term quantifies the amount of remaining flight hours for all aircraft at each period.
 These two objectives have the same units, can be easily compared and ensure the aircraft are used in the most efficient way.
-%We will not penalize the maintenances that provide a better final state for aircraft.
 
@@
         % maximum capacity1:
-        & \sum_{i \in \mathcal{I}, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + N_t \leq C^{max}
+        & \sum_{\substack{i \in \mathcal{I}, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N_t \leq C^{max}
           & t \in \mathcal{T} \label{eq:capacity1}\\
         % min assignments:
-        & \sum_{i \in \mathcal{IJ}_j, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \geq R_j
+        & \sum_{\substack{i \in \mathcal{IJ}_j, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}}} a_{ijt_1t_2} \geq R_j
                 & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
         % just doing one thing at any given time:
-        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \leq 1
+        & \sum_{\substack{(t_1, t_2) \in \\ \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + \sum_{\substack{j \in \\ \mathcal{JT}_t \cap \mathcal{JI}_i}} \sum_{\substack{(t_1, t_2) \in \\ \mathcal{T}\mathcal{T}\mathcal{J}_{jt}}} a_{ijt_1t_2} \leq 1
                 & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
@@
 
-%     \begin{align}
-%         % if start mission => assign start
-%         & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
-%                 & t =1, ..., \mathcal{T}, j \in \mathcal{JT}_t, i \in \mathcal{IJ}_j \label{eq:start1} \\
-%         & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{Init}_j}
-%                 & j \in \mathcal{J}_0, i \in \mathcal{IJ}_j \label{eq:start2} \\
-%         & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
-%         & j \in \mathcal{J}, t \in \mathcal{TJ}_j, i \in \mathcal{IJ}_j \label{eq:start3}
-%     \end{align}
-
-%     Constraints (\ref{eq:start1}) captures period $t$ where aircraft $i$ is firstly assigned to mission $j$ i.e.
-it was not assigned to it in period ($t - 1$).
-Constraints (\ref{eq:start2}) are introduced for the first period in the planning horizon.
-% Constraints (\ref{eq:start3}) control the minimum duration of a consecutive mission assignment.
-If aircraft $i$ was firstly assigned to mission $j$ in period $t$, it has to be assigned to it during the following $t' \in \mathcal{T}_{jt}^{MT}$ periods.
-This is a stronger version of the constraint $a_{jt'i}^{s} \leq a_{jti}$.
-
-    % To our knowledge, these constraints have not been taken into account in previous military FMP problems.
-
     \begin{align}
-       & \sum_{i \in \mathcal{IK}_k, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt}
-        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster} \\
-       & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
-        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
+       & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt}
+        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}
+       % & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
+       %  &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
     \end{align}
@@
 A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission.
-Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.
+    % Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.
 
     \begin{align}
-        % remaining used time
-         & u_{it} \geq \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} H_j 
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
-         & u_{it} \in [U^{min}, \max_j{\{H_j\}}]
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight_lower}\\
-         & rft_{it} \leq rft_{i(t-1)} + H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} - u_{it}
-            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rft_upper}\\
-        & rft_{i0} = Rft^{Init}_i
-               & i \in \mathcal{I} \label{eq:rft_initial}\\
-        & rft_{it} \geq H^{max} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2}
-                &t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rft_lower}\\
-        & rft_{it} \in [0,H^{max}]
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
+        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H'_{jtt'} + U^{min}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
+                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours1}\\
+        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours2}\\
+        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours3}
     \end{align}
 
-        % These constraints calculate the balances of hours for each aircraft.
-    The flight time per aircraft and period is calculated in (\ref{eq:flight1})-(\ref{eq:flight_lower}).
-The remaining flight time is defined by (\ref{eq:rft_upper})-(\ref{eq:rft_initial}) and its limits by (\ref{eq:rft_lower})-(\ref{eq:mu}).
+        Constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}) limit the total flight hours of any given aircraft between checks.
+Constraints (\ref{eq:cycle_hours1}) takes into account periods since the beginning of the planning horizon until the first check, \ref{eq:cycle_hours2} takes into account periods between the first and second check and \ref{eq:cycle_hours3} takes into account periods between the second check and the end of the planning horizon.
+
+    \begin{align}
+        & h_{i1t_1} \geq (t_1 - 1) Rft^{Init}_i m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H''_{jtt't_1} - U^{min}_{1t_1} 
+            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours1}\\
+        & h_{it_1t_2} \geq (t_2 - t_1) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H''_{jtt't_2} - U^{min}_{t_1t_2}
+            &t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours2}\\
+        & h_{it_2|\mathcal{T}|} \geq (|\mathcal{T}|- t_2) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H''_{jtt'|\mathcal{T}|} - U^{min}_{t_2|\mathcal{T}|}
+            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours3}
+    \end{align}
+
+    Constraints (\ref{eq:cycle_rem_hours1}) - (\ref{eq:cycle_rem_hours3}) define the sum of remaining flight hours each aircraft has in between checks.
+Similarly to (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}), their difference lies in the three possible intervals produced by assigning at most two checks.
+The derived parameters used in these constraints can be found in section \ref{subsubsec:consumption}.
 
@@
        % For now, we only want one or possibly two maintenances...
-        & \sum_{t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup |\mathcal{T}| } m_{it_1t_2} =  1 
+        & \sum_{\substack{t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} \cup |\mathcal{T}|}} m_{it_1t_2} =  1 
           & i \in \mathcal{I}\label{eq:num_maint} \\
@@
 
-    Constraints (\ref{eq:maint_forbid} - \ref{eq:maint_forbid2}) limit the combinations of maintenance schedules that can be done during the horizon assuming that the first maintenance is dependant in the initial state of the aircraft and the second maintenance (if any) is limited by the moment at which the first maintenance was planned.
+    Constraints (\ref{eq:maint_forbid}) - (\ref{eq:maint_forbid2}) limit the combinations of maintenance schedules that can be done during the horizon assuming that the first maintenance is dependant in the initial state of the aircraft and the second maintenance (if any) is limited by the moment at which the first maintenance was planned.
 

### 2019-04-15 latex/notes/bounds/bounds.tex
passed bounds document into latex.

--- 2019-04-12 latex/notes/model/model.tex
@@
 
-\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
+\renewcommand\section{\@startsection{section}{2}{\z@}%
                        {-6\p@ \@plus -0\p@ \@minus -0\p@}%
@@
 
-\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
+\renewcommand\section{\@startsection{section}{3}{\z@}%
                        {-6\p@ \@plus -0\p@ \@minus -0\p@}%
@@
 
-\section{Mathematical formulation}
-  \label{sec:model}
+The idea is to define ways with which we can understand the problem, based on the initial input data.
 
-The following model provides a tight MIP formulation that solves the Military Flight and Maintenance Problem. Maintenance operations will be referred as "checks".
-
-\subsection{Input data}
-    \subsubsection{Basic sets}
+\section{Input data}
+    \subsection{Basic sets}
 
@@
 
-    \subsubsection{Mission parameters}
+    \subsection{Mission parameters}
 
@@
 
-    \subsubsection{Maintenance parameters}
+    \subsection{Maintenance parameters}
 
@@
 
-    \subsubsection{Fleet parameters}
+    \subsection{Fleet parameters}
 
@@
 
-    \subsubsection{Parametric sets}
+    \subsection{Parametric sets}
 
@@
 
-    \subsubsection{Time-related parametric sets}
+    \subsection{Time-related parametric sets}
 
@@
             $t \in \mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
-            % $\mathcal{T}^S_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t},  ..., \min{\{|\mathcal{T}|, t + M-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
             $t' \in \mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^{min}-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
@@
 
-    \begin{figure}
-        \centering
-        % \includegraphics[width=\linewidth]{img/gantt_timewindows_all.png}
-        \begin{tikzpicture}
-            \input{gantts/time_windows.tex}
-        \end{tikzpicture}
-        \begin{tikzpicture}
-            \input{gantts/time_windows2.tex}
-        \end{tikzpicture}
-        \begin{tikzpicture}
-            \input{gantts/time_windows_init.tex}
-        \end{tikzpicture}
-        \begin{tikzpicture}
-            \input{gantts/time_windows_mission.tex}
-        \end{tikzpicture}
-        \caption{Examples showing the maintenance-related time-parametric sets for aircraft $i$.
-(a) $\mathcal{T}_{t'}^{s}$ refers to the previous $M$ periods to period $t'$.
-$\mathcal{T}_{t}^{m}$ refers to the periods where a check cannot be planned after starting a check in period $t$.
-Finally, $\mathcal{T}_{t}^{M}$ refers to the periods where a check needs to be scheduled, after starting a check in period $t$.
-(b) The maintenance assignment for aicraft $i$ is $m_{it_1t_2}=1$.
-$\mathcal{T}_A \cup \mathcal{T}_B$ represents the set of periods $t$ for which $(t_1, t_2) \in \mathcal{TTT}_t$.
-(c) $Rct_{i}^{Init}$ is equal to 55, meaning a check has to be planned between months number 25 and 54.
-The maintenance assignment for aircraft $i$ is $m_{it_1t_2}=1$ and, since it includes the last period, it only represents one operation.
-(d) $\mathcal{T}_{jt}^{MT}$ refers to the periods where the assignment of mission $j$ needs to be kept.
-In this case, the size is 6.
-The mission assignments for aircraft $i$ to mission $j$ are $a_{ijt_1t_2}=1$ and $a_{ijt'_1t'_2}=1$}
-        \label{fig:gantt_windows}
-    \end{figure}
+% \section{Number and possible dates for maintenances}
+% \label{number-and-possible-dates-for-maintenances}
 
-    \subsubsection{Consumption derived parameters}\label{subsubsec:consumption}
+% The aircraft initial state already conditions the first maintenance of each aircraft.
 
-    The following parameters are derived from the basic input parameters.
-They are used in constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_rem_hours3}).
-They are transformations of the default and mission flight hours consumption under different time circumstances.
+% Moreover, using the maximum and minimum distances between maintenances (min and max size for maintenances cycles respectively), \textbf{we can precalculate the possible total number of maintenances that an aircraft will have in the horizon}.
+This number will probably be one of the following: {[}1{]}, {[}1, 2{]}, {[}2{]}, {[}1, 2, 3{]}.
+This will depend on the size of the horizon and maintenance conditions.
+
+\section{Range of number of maintenances per period and cluster}
+\label{range-of-number-of-maintenances-per-period-and-cluster}
+
+First, for each period $t$, and using the initial states of aircraft, we can calculate the minimum and maximum number of maintenances that an aircraft could have already done (\emph{ended}) before the start of the period.
+We can then aggregate this for each aircraft type to obtain a lower and an upper bound on the number of maintenances that can be finished before period $t$.
 
 \begin{align}
-        & U^{min}_{t_1t_2} = U^{min} \sum_{t=t_1}^{t_2} (t_2 - t) \notag \\
-        & H'_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1) \notag \\
-        & H''_{jt_1t_2t'} = \sum_{t=t_1}^{t_2} (H_j- U^{min}) (t' - t) \notag
+
+	& t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \forall i \in \mathcal{I}, |\mathcal{T}|
+
 \end{align}
 
-\subsection{Variables}
+Secondly, we can calculate the number of maintenances we can fit until period $t$, by looking at missions requirements in terms of number of aircraft.
+This would be done by aircraft type.
+We first take the sum of all the non-mission periods over all aircraft $i \in IK_k \forall k \in K$\$ and all periods $t' < t$.
+Then, we divide this over the duration of the maintenances ($M$).
+This would an upper bound.
 
-    The following decision variables control the assignment of missions and checks to aircraft.
+ Thirdly, we can sum all flight-hour needs of missions of type $k$ before time $t$.
+If we substract this from the initial remaining flight time of the group of aircraft, we are left with a number of hours that need to be obtained via maintenances.
+If we divide this number over the $H^{max}$ flight hours a maintenance gives, we get a lower bound on the number of maintenances we need to do for those aircraft of type $k$.
+So, in this way, we have two lower bounds and two upper bounds per period and aircraft type.
+We could just get the maximum and the minimum respectively to get some bounds on the number of maintenances until period $t$.
+{%
 
-    \begin{tabular}{p{8mm}p{147mm}}
-        $a_{ijtt'}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ at the beginning of period $t$ and finishes it at the end of period $t'$. \\  
-        $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check at the beginning of period $t \in \mathcal{T}$ and then starts the next one at the beginning of period $t'$, 0 otherwise. \\
-        $h_{itt'}$  &  Continuous. Sum of the monthly remaining flight hours for aircraft $i \in I$ for periods between the beginning of period $t$ and the beginning of period $t'$. \\
-    \end{tabular}
+\section{Mission assignment}\label{mission-assignment}}  {%
 
-    \paragraph{Fixed values}
+\section{Mission assignment for initial periods}\label{mission-assignment-for-initial-periods}}
 
-    Note that $a_{ijtt'}$ and $m_{itt'}$ are initially set up to 0 for all aircraft already in maintenance at the beginning of the planning horizon for the remaining time periods of the check.
-$N_{t}$ is calculated based on this information.
-Similarly, for aircraft that have not yet complied with their minimum mission assignment duration at the beginning of the planning horizon, $a_{ijtt'}$ is fixed to comply with the constraints.
-
-\subsection{Objective function and constraints}
-
-    Objective (\ref{eq:objective1}) minimizes the number of checks while at the same time maximizing the monthly total flight hours potential of the fleet.
-
-    \begin{align}
-        & \text{Min}\; \sum_{\substack{i \in \mathcal{I}, \\ t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|}} m_{it_1t_2} \times H^{max} \times (t_2 - t_1) - h_{it_1t_2}  \label{eq:objective1}
-    \end{align}
-
-    The first term counts all the remaining flight hours given to aircraft following checks and the second term quantifies the amount of remaining flight hours for all aircraft at each period.
-These two objectives have the same units, can be easily compared and ensure the aircraft are used in the most efficient way.
-
-    The following constraints are used in the model:
-    \begin{align}
-        % maximum capacity1:
-        & \sum_{\substack{i \in \mathcal{I}, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N_t \leq C^{max}
-          & t \in \mathcal{T} \label{eq:capacity1}\\
-        % min assignments:
-        & \sum_{\substack{i \in \mathcal{IJ}_j, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}}} a_{ijt_1t_2} \geq R_j
-                & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
-        % just doing one thing at any given time:
-        & \sum_{\substack{(t_1, t_2) \in \\ \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + \sum_{\substack{j \in \\ \mathcal{JT}_t \cap \mathcal{JI}_i}} \sum_{\substack{(t_1, t_2) \in \\ \mathcal{T}\mathcal{T}\mathcal{J}_{jt}}} a_{ijt_1t_2} \leq 1
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
-    \end{align}
-
-    Maintenance capacity is controlled by (\ref{eq:capacity1}).
-The aircraft requirements of missions are defined by (\ref{eq:missionres}).
-Constraints (\ref{eq:state}) ensure that an aircraft can only be used for one mission or undergo check in the same period.
-
-    \begin{align}
-       & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt}
-        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}
-       % & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
-       %  &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
-    \end{align}
-
-    Constraints (\ref{eq:serviceability-cluster}) guarantee a minimum serviceability of aircraft for each cluster $k$.
-A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission.
-    % Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.
-
-    \begin{align}
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H'_{jtt'} + U^{min}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
-                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours1}\\
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
-                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours2}\\
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
-                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours3}
-    \end{align}
-
-        Constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}) limit the total flight hours of any given aircraft between checks.
-Constraints (\ref{eq:cycle_hours1}) takes into account periods since the beginning of the planning horizon until the first check, \ref{eq:cycle_hours2} takes into account periods between the first and second check and \ref{eq:cycle_hours3} takes into account periods between the second check and the end of the planning horizon.
-
-    \begin{align}
-        & h_{i1t_1} \geq (t_1 - 1) Rft^{Init}_i m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H''_{jtt't_1} - U^{min}_{1t_1} 
-            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours1}\\
-        & h_{it_1t_2} \geq (t_2 - t_1) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H''_{jtt't_2} - U^{min}_{t_1t_2}
-            &t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours2}\\
-        & h_{it_2|\mathcal{T}|} \geq (|\mathcal{T}|- t_2) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H''_{jtt'|\mathcal{T}|} - U^{min}_{t_2|\mathcal{T}|}
-            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours3}
-    \end{align}
-
-    Constraints (\ref{eq:cycle_rem_hours1}) - (\ref{eq:cycle_rem_hours3}) define the sum of remaining flight hours each aircraft has in between checks.
-Similarly to (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}), their difference lies in the three possible intervals produced by assigning at most two checks.
-The derived parameters used in these constraints can be found in section \ref{subsubsec:consumption}.
-
-    \begin{align}
-       % For now, we only want one or possibly two maintenances...
-        & \sum_{\substack{t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} \cup |\mathcal{T}|}} m_{it_1t_2} =  1 
-          & i \in \mathcal{I}\label{eq:num_maint} \\
-        % & m_{it'} + m_{it} \leq 1
-        %   & t \in \mathcal{T}, t' \in \mathcal{T}^m_t, i \in \mathcal{I}\label{eq:rct_min}\\  
-        % & \sum_{t' \in \mathcal{T}^M_t} m_{it'} \geq  m_{it}
-        %   & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
-        % & m_{it_1t_2} = 0
-        %   & t_1 \in \mathcal{T}^{m_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I}\label{eq:rct_min_init} \\
-        & m_{it_1t_2} =  0
-          & t_1 \notin \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}, i \in \mathcal{I} \label{eq:maint_forbid} \\
-        & m_{it_1t_2} =  0
-          & t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \notin \mathcal{T}^M_{t_1} \cup |\mathcal{T}|, i \in \mathcal{I} \label{eq:maint_forbid2}
-    \end{align}
-
-    Constraints (\ref{eq:num_maint}) limit the number of maintenance assignments per aircraft to one.
-This implies that each aircraft will get strictly one or two maintenances over the whole planning horizon.
-
-    Constraints (\ref{eq:maint_forbid}) - (\ref{eq:maint_forbid2}) limit the combinations of maintenance schedules that can be done during the horizon assuming that the first maintenance is dependant in the initial state of the aircraft and the second maintenance (if any) is limited by the moment at which the first maintenance was planned.
-
+Specially for the initial periods, and based on the initial state, missions' hourly requirements and minimum assignment duration, it is possible to \textbf{precalculate which aircraft will not be able to start a mission in some periods}.
+We can do this only for the initial periods because we do not have the actual consumption for individual aircraft.
+In the case of having a minimum consumption, this could be used to have some lower bounds too.
 \end{document}

### 2019-04-16 latex/notes/bounds/bounds.tex
deleted bounds.txt. First version of bounds document.

@@
 \usepackage{pgfgantt}
+\usepackage{multicol}
 
 \setlength{\parskip}{2ex} \pagestyle{myheadings}
-% \markright{\hspace*{3.8cm} \textit{MOSIM18 - June 27-29, 2018 - Toulouse - France}}
 
@@
             $j \in \mathcal{J}$     &  missions. \\
-            $k \in \mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
+            $y \in \mathcal{Y}$     &  types of mission or aircraft. \\
+            $k \in \mathcal{K}$     &  clusters of aircraft that share exactly the same functionality. \\
         \end{tabular}
@@
             $M$               & check duration in number of periods. & periods. \\
+            $Y_j$             & type $y \in \mathcal{Y}$ for mission $j$. \\
         \end{tabular}
@@
             $Rct^{Init}_i$      & remaining calendar time for aircraft $i$ at the start of the planning horizon.  & periods. \\
+            $Y_i$               & type $y \in \mathcal{Y}$ for aircraft $j$. \\
         \end{tabular}
@@
             $i \in \mathcal{IJ}_j$     &  aircraft $i \in \mathcal{I}$ that can be assigned to mission $j$. \\
+            $i \in \mathcal{IY}_y$     &  aircraft $i \in \mathcal{I}$ that are included in type $y$. One aircraft cannot belong to more than one type. \\
+            $j \in \mathcal{JY}_y$     &  missions $j \in \mathcal{J}$ that are included in type $y$. One mission cannot belong to more than one type. \\
             $i \in \mathcal{IK}_k$     &  aircraft $i \in \mathcal{I}$ that are included in cluster $k$. One aircraft can belong to more than one cluster. \\
@@
         \begin{tabular}{p{30mm}p{125mm}}
-            $t \in \mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
-            $t' \in \mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^{min}-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
-            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$ (figure \ref{fig:gantt_windows}a). \\
-            $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}b). \\
-            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b).
-\\
-            $t' \in \mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
-            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
-            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$.
-\\
+            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$. \\
+            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$. \\
             $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT_j-1$ and $t \ge t_1 + M$ and $t' < t_2$ \\
@@
 
-% \section{Number and possible dates for maintenances}
-% \label{number-and-possible-dates-for-maintenances}
+\section{Range of number of maintenances until each period}
+\label{range-of-number-of-maintenances-per-period}
 
-% The aircraft initial state already conditions the first maintenance of each aircraft.
-
-% Moreover, using the maximum and minimum distances between maintenances (min and max size for maintenances cycles respectively), \textbf{we can precalculate the possible total number of maintenances that an aircraft will have in the horizon}.
-This number will probably be one of the following: {[}1{]}, {[}1, 2{]}, {[}2{]}, {[}1, 2, 3{]}.
-This will depend on the size of the horizon and maintenance conditions.
-
-\section{Range of number of maintenances per period and cluster}
-\label{range-of-number-of-maintenances-per-period-and-cluster}
-
-First, for each period $t$, and using the initial states of aircraft, we can calculate the minimum and maximum number of maintenances that an aircraft could have already done (\emph{ended}) before the start of the period.
+First, for each period $t$, and using the initial states of aircraft, we can calculate the minimum and maximum number of maintenances that an aircraft could have already \emph{ended} before the start of the period.
 We can then aggregate this for each aircraft type to obtain a lower and an upper bound on the number of maintenances that can be finished before period $t$.
 
+$TM1^{min}_i$ and $TM1^{max}_i$ represent, respectively, the minimum and maximum period for starting the first maintenance for aircraft $i$.
+Analogously, $TM2^{min}_i$ and $TM2^{max}_i$ represent the minimum and maximum period for starting the second maintenance for aircraft $i$.
+
+\paragraph{Note on the number of maintenances.} Here we are assuming a maximum number of maintenances of two.
+This is the case based on the size of the horizon we are studying (90 - 120 periods) and the minimum - maximum distance between maintenances (30 to 60 periods).
+
 \begin{align}
-
-	& t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \forall i \in \mathcal{I}, |\mathcal{T}|
-
+	& TM1^{min}_i = \min_t{\{t \in \mathcal{T}^{M_{Init}}_i\}} \\
+  & TM1^{max}_i = \max_t{\{t \in \mathcal{T}^{M_{Init}}_i\}} \\
+  & TM2^{min}_i = \min_t{\{t \in \mathcal{T}^M_{TM1^{min}_i}\}} \\
+  & TM2^{max}_i = \max_t{\{t \in \mathcal{T}^M_{TM1^{max}_i}\}} %\\
 \end{align}
 
-Secondly, we can calculate the number of maintenances we can fit until period $t$, by looking at missions requirements in terms of number of aircraft.
+Lower bounds and upper bounds for the number of maintenances possibly started ($NMS^{min}_{it}$, $NMS^{max}_{it}$) and finished ($NMF^{min}_{it}$, $NMF^{max}_{it}$) can be pre-calculated based on where the period is located in the planning horizon relative to the possible first and second maintenance possibilities and the maintenance duration.
+
+\begin{multicols}{2}
+\[
+ NMS^{min}_{it} = 
+  \begin{cases} 
+   0 & \text{if } t \le TM1^{max}_i\\
+   1 & \text{if } TM1^{max}_i < t \le TM2^{max}_i \\
+   2 & \text{if } t > TM2^{max}_i \\
+  \end{cases}
+\]
+
+\[
+ NMS^{max}_{it} = 
+  \begin{cases} 
+   0 & \text{if } t < TM1^{min}_i \\
+   1 & TM1^{min}_i \le t < TM2^{min}_i \\
+   2 & t \ge TM2^{min}_i \\
+  \end{cases}
+\]
+
+\columnbreak
+
+\[
+ NMF^{min}_{it} = 
+  \begin{cases} 
+   0 & \text{if } t \le TM1^{max}_i + M\\
+   1 & \text{if } TM1^{max}_i + M< t \le TM2^{max}_i + M\\
+   2 & \text{if } t > TM2^{max}_i + M \\
+  \end{cases}
+\]
+
+\[
+ NMF^{max}_{it} = 
+  \begin{cases} 
+   0 & \text{if } t < TM1^{min}_i + M\\
+   1 & TM1^{min}_i + M \le t < TM2^{min}_i + M\\
+   2 & t \ge TM2^{min}_i + M\\
+  \end{cases}
+\]
+
+\end{multicols}
+
+\section{Range of number of maintenances per aircraft type until each period}
+
+We can aggregate the previous values by type of fleet $y$, as can be seen in (\ref{eq:agg_maints_start}) - (\ref{eq:agg_maints_end}).
+
+\begin{align}
+  & NMSY^{b}_{yt} = \sum_{i \in \mathcal{IY}_y} NMS^{b}_{it} 
+      & t \in \mathcal{T}, y \in \mathcal{Y}, b \in \{min, max\} \label{eq:agg_maints_start} \\
+  & NMFY^{b}_{yt} = \sum_{i \in \mathcal{IY}_y} NMF^{b}_{it} 
+      & t \in \mathcal{T}, y \in \mathcal{Y}, b \in \{min, max\} \label{eq:agg_maints_end}
+\end{align}
+
+Secondly, we can pre-calculate the number of maintenances we can fit until period $t$, by looking at missions requirements in terms of number of aircraft.
 This would be done by aircraft type.
@@
 Then, we divide this over the duration of the maintenances ($M$).
-This would an upper bound.
+$JR^{Acc}_{jt}$ represents the accumulated required number of assignments of aircraft for mission $j$ until period $t$, including it.
+$YFN_{yt}$ represent the number of periods that are not assigned to a mission for all aircraft of type $y$.
+$NMFY'^{max}_{yt}$ is an upper bound on the number of maintenances that can be finished until period $t$ for all aircraft of type $y$.
+
+\begin{align}
+  & JR^{Acc}_{jt} = |\{t' \in \mathcal{TJ}_j \mid t' \le t\}| 
+      & t \in \mathcal{T}, y \in \mathcal{Y}\\
+  & YFN_{yt} = |IY_y| \times t -  \sum_{j \in \mathcal{JY}_y} R_j \times JR^{Acc}_{jt} 
+      & t \in \mathcal{T}, y \in \mathcal{Y}\\
+  & NMFY^{\prime max}_{yt} = \lfloor \frac{YFN_{yt}}{M} \rfloor 
+      & t \in \mathcal{T}, y \in \mathcal{Y}
+\end{align}
 
 Thirdly, we can sum all flight-hour needs of missions of type $k$ before time $t$.
-If we substract this from the initial remaining flight time of the group of aircraft, we are left with a number of hours that need to be obtained via maintenances.
+If we subtract this from the initial remaining flight time of the group of aircraft, we are left with a number of hours that need to be obtained via maintenances.
 If we divide this number over the $H^{max}$ flight hours a maintenance gives, we get a lower bound on the number of maintenances we need to do for those aircraft of type $k$.
+
+\begin{align}
+  & YH^{Acc}_{yt} =  \sum_{j \in \mathcal{JY}_y} H_j \times JR^{Acc}_{jt} 
+      & t \in \mathcal{T}, y \in \mathcal{Y}\\
+  & YH^{Init}_{yt} = \sum_{i \in IY_y} Rft^{Init}_i 
+      & t \in \mathcal{T}, y \in \mathcal{Y}\\
+  & NMFY^{\prime min}_{yt} =  \lceil \frac{YH^{Acc}_{yt} - YH^{Init}_{yt}}{H^{max}} \rceil
+      & t \in \mathcal{T}, y \in \mathcal{Y}
+\end{align}
+
 So, in this way, we have two lower bounds and two upper bounds per period and aircraft type.
 We could just get the maximum and the minimum respectively to get some bounds on the number of maintenances until period $t$.
-{%
+$NMFY^{\prime\prime b}_{yt}$ represents the bound $b$ in the number of total maintenances since the beginning of the planning horizon for all aircraft of type $y$ until period $t$.
 
-\section{Mission assignment}\label{mission-assignment}}  {%
+\begin{align}
+  & NMFY^{\prime\prime min}_{yt} = \max\{NMFY^{min}_{yt}, NMFY^{\prime min}_{yt}\} \\ 
+  & NMFY^{\prime\prime max}_{yt} = \min\{NMFY^{max}_{yt}, NMFY^{\prime max}_{yt}\}
+\end{align}
 
-\section{Mission assignment for initial periods}\label{mission-assignment-for-initial-periods}}
+This can be applied via a cut by the following way:
+
+\begin{align}
+  & NMFY^{\prime\prime min}_{yt} = \max\{NMFY^{min}_{yt}, NMFY^{\prime min}_{yt}\} \\ 
+  & NMFY^{\prime\prime max}_{yt} = \min\{NMFY^{max}_{yt}, NMFY^{\prime max}_{yt}\}
+\end{align}
+
+\[
+  QM^{num}_{tt'} = 
+  \begin{cases} 
+   1 & t' > t - M \\
+   2 & t' \le t - M \\
+  \end{cases} 
+\]
+
+$QM^{num}_{tt'}$ represents the possibility of counting two checks instead of one if both checks start before the period $t$, $t'$ representing the second one of those checks.
+
+\begin{align}
+  & NMFY^{\prime\prime min}_{yt} \le \sum_{\substack{i \in \mathcal{IY}_y, \\ t_1 \in \{1, \ldots, t - M\}, \\ t_2 \in \mathcal{T}^M_{t_1}}} m_{it_1t_2} \times QM^{num}_{tt_2} \le NMFY^{\prime\prime max}_{yt}
+      & t \in \mathcal{T}, y \in \mathcal{Y} \label{eq:range_maints}
+\end{align}
+
+Cuts (\ref{eq:range_maints}) limit the starts of checks of aircraft of type $y$ in order to have the number of finished checks to fall between two ranges.
+
+\section{Number of checks at the end of the horizon per aircraft}
+
+If we evaluate the range of checks per aircraft at the last period ($\mathcal{T}$), we can get one of the three options.
+If we find ourselves in the cases where there is certainty on the number of checks, it is possible to force the variables to reflect this certainty with the following cuts.
+
+\[
+[NMS^{min}_{i|\mathcal{T}|}, NMS^{max}_{i|\mathcal{T}|}] :
+  \begin{cases} 
+   m_{it_1|\mathcal{T}|} =  1
+    \,\, \forall t_1 \in \mathcal{T}^{M_{Init}}_i, i \in \mathcal{I} & \text{if } [1, 1] \\
+   m_{it_1|\mathcal{T}|} =  0 \,\, \forall t_1 \in \mathcal{T}^{M_{Init}}_i, i \in \mathcal{I} & \text{if } [2, 2] 
+  \end{cases}
+\]
+
+\section{Mission assignment for initial periods}
+\label{mission-assignment-for-initial-periods}
 
@@
 In the case of having a minimum consumption, this could be used to have some lower bounds too.
+
+For each aircraft, we calculate the last period before being able to finish a maintenance ($t^{s}_i$).
+Before this period, the sum of all assigned misions and consummed flight hours need to be less than the initial remaining flight hours.
+This is similar to the already existing constraints when assigning maintenances.
+
+\begin{align}
+    & U^{min}_{t_1t_2} = U^{min} \sum_{t=t_1}^{t_2} (t_2 - t) \notag \\
+    & H'_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1) \notag \\
+    & H''_{jt_1t_2t'} = \sum_{t=t_1}^{t_2} (H_j- U^{min}) (t' - t) \notag
+\end{align}
+
+\begin{align}
+  & t^{s}_i = TM1^{min}_i + M - 1&  \\
+  & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t^{s}_i}}} a_{ijtt'} H'_{jtt'} + U^{min}_{1t^{s}_i} \leq Rft^{Init}_i
+                  & i \in \mathcal{I}
+\end{align}
+
+% Not sure if this is redundant or not.
+On top of that, some mission assignments can already be discarded because the equivalent flight hour consumption is more than the initial remaining time.
+
+\begin{align}
+  & a_{ijtt'} = 0 & i \in \mathcal{I}, (j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{i1t^{s}_i} \land Rft^{Init}_i < H'_{jt_1t_2}
+\end{align}
+
 \end{document}

### 2019-04-16 latex/notes/bounds/bounds.tex
reformatted the bounds document.

--- 2019-04-15 latex/notes/bounds/bounds.tex
@@
 
-The idea is to define ways with which we can understand the problem, based on the initial input data.
+The purpose of this document is to present particular bounds that can be derived from the input data for a particular instance of the problem.
+These derived bounds can be used to prove the infeasibility of a given instance or to add cuts to the original mathematical formulation.
+
+The first section is a description of the input data. The following sections are reserved for specific derived bounds.
 
@@
 
-% \section{Number and possible dates for maintenances}
-% \label{number-and-possible-dates-for-maintenances}
+\section{Bounds on accumulated checks per aircraft and period} \label{sec:rangechecksaircraft}
 
-% The aircraft initial state already conditions the first maintenance of each aircraft.
+  For each period $t$, and using the aircraft initial states, we calculate the minimum and maximum number of checks that an aircraft could have already started / ended at the start of the period.
 
-% Moreover, using the maximum and minimum distances between maintenances (min and max size for maintenances cycles respectively), \textbf{we can precalculate the possible total number of maintenances that an aircraft will have in the horizon}.
-This number will probably be one of the following: {[}1{]}, {[}1, 2{]}, {[}2{]}, {[}1, 2, 3{]}.
-This will depend on the size of the horizon and maintenance conditions.
+  $TM1^{min}_i$ and $TM1^{max}_i$ represent, respectively, the minimum and maximum periods for starting the first check for aircraft $i$.
+Analogously, $TM2^{min}_i$ and $TM2^{max}_i$ represent the minimum and maximum period for starting the second check for aircraft $i$.
 
-\section{Range of number of maintenances per period and cluster}
-\label{range-of-number-of-maintenances-per-period-and-cluster}
+  \paragraph{Note on the number of checks.} Here we are assuming a maximum number of checks of two.
+Based on the size of the horizon we are studying (90 - 120 periods) and the minimum - maximum distance between checks (30 to 60 periods), these are the only possibilities.
 
-First, for each period $t$, and using the initial states of aircraft, we can calculate the minimum and maximum number of maintenances that an aircraft could have already done (\emph{ended}) before the start of the period.
-We can then aggregate this for each aircraft type to obtain a lower and an upper bound on the number of maintenances that can be finished before period $t$.
+  These limits are obtained in the following way:
 
   \begin{align}
-
-	& t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \forall i \in \mathcal{I}, |\mathcal{T}|
-
+  	& TM1^{min}_i = \min_t{\{t \in \mathcal{T}^{M_{Init}}_i\}}   \notag\\
+    & TM1^{max}_i = \max_t{\{t \in \mathcal{T}^{M_{Init}}_i\}}   \notag \\
+    & TM2^{min}_i = \min_t{\{t \in \mathcal{T}^M_{TM1^{min}_i}\}}\notag \\
+    & TM2^{max}_i = \max_t{\{t \in \mathcal{T}^M_{TM1^{max}_i}\}}\notag %\\
   \end{align}
 
-Secondly, we can calculate the number of maintenances we can fit until period $t$, by looking at missions requirements in terms of number of aircraft.
-This would be done by aircraft type.
-We first take the sum of all the non-mission periods over all aircraft $i \in IK_k \forall k \in K$\$ and all periods $t' < t$.
-Then, we divide this over the duration of the maintenances ($M$).
-This would an upper bound.
+  Lower and upper bounds for the accumulate number of checks started ($M^{Acc}S^{min}_{it}$, $M^{Acc}S^{max}_{it}$) and finished ($M^{Acc}F^{min}_{it}$, $M^{Acc}F^{max}_{it}$) can be pre-calculated based on where the period is located in the planning horizon relative to the possible first and second check ranges and the check duration.
+We will name the range $M^{Acc}S^{b}_{it}$ to represent both the lower level and upper level of the range of accumulated starts by replacing the terms $min$ and $max$ by $b$, as in $bound$.
 
- Thirdly, we can sum all flight-hour needs of missions of type $k$ before time $t$.
-If we substract this from the initial remaining flight time of the group of aircraft, we are left with a number of hours that need to be obtained via maintenances.
-If we divide this number over the $H^{max}$ flight hours a maintenance gives, we get a lower bound on the number of maintenances we need to do for those aircraft of type $k$.
-So, in this way, we have two lower bounds and two upper bounds per period and aircraft type.
-We could just get the maximum and the minimum respectively to get some bounds on the number of maintenances until period $t$.
-{%
+  \begin{multicols}{2}
+  \[
+   M^{Acc}S^{min}_{it} = 
+    \begin{cases} 
+     0 & \text{if } t \le TM1^{max}_i\\
+     1 & \text{if } TM1^{max}_i < t \le TM2^{max}_i \\
+     2 & \text{if } t > TM2^{max}_i \\
+    \end{cases}
+  \]
 
-\section{Mission assignment}\label{mission-assignment}}  {%
+  \[
+   M^{Acc}S^{max}_{it} = 
+    \begin{cases} 
+     0 & \text{if } t < TM1^{min}_i \\
+     1 & TM1^{min}_i \le t < TM2^{min}_i \\
+     2 & t \ge TM2^{min}_i \\
+    \end{cases}
+  \]
 
-\section{Mission assignment for initial periods}\label{mission-assignment-for-initial-periods}}
+  \columnbreak
+
+  \[
+   M^{Acc}F^{min}_{it} = 
+    \begin{cases} 
+     0 & \text{if } t \le TM1^{max}_i + M\\
+     1 & \text{if } TM1^{max}_i < t - M \le TM2^{max}_i \\
+     2 & \text{if } t > TM2^{max}_i + M \\
+    \end{cases}
+  \]
+
+  \[
+   M^{Acc}F^{max}_{it} = 
+    \begin{cases} 
+     0 & \text{if } t < TM1^{min}_i + M\\
+     1 & TM1^{min}_i \le t - M < TM2^{min}_i\\
+     2 & t \ge TM2^{min}_i + M\\
+    \end{cases}
+  \]
+
+  \end{multicols}
+
+\section{Bounds on accumulated checks at the end of the horizon per aircraft}
+
+  If we evaluate the range of checks per aircraft at the last period ($\mathcal{T}$), we can get one of three options: [1, 1], [1, 2], [2, 2].
+Lets call $I^{1M}$ to aircraft that fall into the [1, 1] case and $I^{2M}$ to groups aircraft that fall into the [2, 2] case.
+These two sets of aircraft can have particular cuts applied to them to guarantee the number of checks.
+
+   If we find ourselves in a case where there is certainty on the number of checks, it is possible to force the variables to reflect this certainty with the following cuts.
+
+  \begin{align}
+    & I^{1M} = \{i \in \mathcal{I} \mid M^{Acc}S^{min}_{i|\mathcal{T}|} = M^{Acc}S^{max}_{i|\mathcal{T}|} = 1\} \notag \\
+    & I^{2M} = \{i \in \mathcal{I} \mid M^{Acc}S^{min}_{i|\mathcal{T}|} = M^{Acc}S^{max}_{i|\mathcal{T}|} = 2\} \notag
+  \end{align}
+
+  \begin{align}
+    & m_{it_1|\mathcal{T}|} =  1
+      & i \in I^{1M}, t_1 \in \mathcal{T}^{M_{Init}}_i \\
+    & m_{it_1|\mathcal{T}|} =  0
+      & i \in I^{2M}, t_1 \in \mathcal{T}^{M_{Init}}_i
+  \end{align}
+
+  % \[
+  % [M^{Acc}S^{min}_{i|\mathcal{T}|}, M^{Acc}S^{max}_{i|\mathcal{T}|}] :
+  %   \begin{cases} 
+  %    m_{it_1|\mathcal{T}|} =  1
+  %     \,\, \forall t_1 \in \mathcal{T}^{M_{Init}}_i, i \in \mathcal{I} & \text{if } [1, 1] \\
+  %    m_{it_1|\mathcal{T}|} =  0 \,\, \forall t_1 \in \mathcal{T}^{M_{Init}}_i, i \in \mathcal{I} & \text{if } [2, 2] 
+  %   \end{cases}
+  % \]
+
+\section{Bounds on mission assignments at the start of the horizon per aircraft} \label{mission-assignment-for-initial-periods}
 
@@
 In the case of having a minimum consumption, this could be used to have some lower bounds too.
+
+  For each aircraft, we calculate the last period before being able to finish the first possible check ($t^{s}_i$).
+Before this period, the sum of all assigned misions and consummed flight hours need to be less than the initial remaining flight hours.
+This is a particular case of the already existing constraints controlling flight hours when assigning checks.
+
+  As with the previous constraints, it is convenient to pre-calculate certain parameters:
+
+  \begin{align}
+      & U^{min}_{t_1t_2} = U^{min} \sum_{t=t_1}^{t_2} (t_2 - t) \notag \\
+      & H'_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1) \notag \\
+      & H''_{jt_1t_2t'} = \sum_{t=t_1}^{t_2} (H_j- U^{min}) (t' - t) \notag \\
+      & t^{s}_i = TM1^{min}_i + M - 1&  \notag
+  \end{align}
+
+  The cuts in (\ref{eq:initialFlightHourConsumption}) control the consummed hours before any maintenance can be done on aircraft $i$ (before the period $t^{s}_i$).
+
+  \begin{align}
+    & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t^{s}_i}}} a_{ijtt'} H'_{jtt'} + U^{min}_{1t^{s}_i} \leq Rft^{Init}_i
+                    & i \in \mathcal{I} \label{eq:initialFlightHourConsumption}
+  \end{align}
+
+  Some mission assignments can already be discarded because the equivalent flight hour consumption is more than the initial remaining time:
+
+  \begin{align}
+    & a_{ijtt'} = 0 & i \in \mathcal{I}, (j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{i1t^{s}_i} \land Rft^{Init}_i < H'_{jt_1t_2}
+  \end{align}
+
+\section{Bounds on accumulated checks per aircraft type and period} \label{sec:rangecheckstype}
+
+  Both $M^{Acc}F^{b}_{it}$ and $M^{Acc}S^{b}_{it}$ previously calculated in section \ref{sec:rangechecksaircraft} can be aggregated by type of fleet $y$.
+
+  \begin{align}
+    & YM^{Acc}S^{b}_{yt} = \sum_{i \in \mathcal{IY}_y} M^{Acc}S^{b}_{it} 
+        & t \in \mathcal{T}, y \in \mathcal{Y}, b \in \{min, max\} \notag \\
+    & YM^{Acc}F^{b}_{yt} = \sum_{i \in \mathcal{IY}_y} M^{Acc}F^{b}_{it} 
+        & t \in \mathcal{T}, y \in \mathcal{Y}, b \in \{min, max\} \notag
+  \end{align}
+
+  By using the mission required assignments over time, we can also calculate the number of checks we can fit until period $t$: aircraft that are in a mission cannot be in maintenance.
+Since each mission demands a specific type of aircraft, this bound can be made at the aircraf type level.
+$JR^{Acc}_{jt}$ represents the accumulated required number of assignments of aircraft for mission $j$ until the end of period $t$.
+  We use this value to obtain the sum of all the non-mission periods over all aircraft of type $y$ and all periods $t' \le t$ ($IR^{Acc}_{yt}$).
+Then, we divide this over the duration of the checks ($M$) to obtain $YM^{Acc}F^{\prime max}_{yt}$, which is an upper bound on the number of checks that can be finished until period $t$ for all aircraft of type $y$.
+
+  \begin{align}
+    & JR^{Acc}_{jt} = |\{t' \in \mathcal{TJ}_j \mid t' \le t\}| 
+        & t \in \mathcal{T}, y \in \mathcal{Y}\notag\\
+    & IR^{Acc}_{yt} = |IY_y| \times t -  \sum_{j \in \mathcal{JY}_y} R_j \times JR^{Acc}_{jt} 
+        & t \in \mathcal{T}, y \in \mathcal{Y}\notag\\
+    & YM^{Acc}F^{\prime max}_{yt} = \lfloor \frac{YFN_{yt}}{M} \rfloor 
+        & t \in \mathcal{T}, y \in \mathcal{Y}\notag
+  \end{align}
+
+  By using the mission required flight hours over time, we can calculate how many checks we need until period $t$.
+We can sum all flight-hour needs of missions of type $y$ until the end of period $t$ ($YH^{Acc}_{yt}$).
+This demand of flight hours can be subtracted this from the initial remaining flight time of the group of aircraft and then divided over the $H^{max}$ flight hours a check gives.
+Thus, we get a lower bound $YM^{Acc}F^{\prime min}_{yt}$ on the number of checks we need to do for those aircraft of type $y$ until period $t$.
+
+  \begin{align}
+    & YH^{Acc}_{yt} =  \sum_{j \in \mathcal{JY}_y} H_j \times JR^{Acc}_{jt} 
+        & t \in \mathcal{T}, y \in \mathcal{Y} \notag\\
+    & YM^{Acc}F^{\prime min}_{yt} =  \lceil \frac{YH^{Acc}_{yt} - \sum_{i \in IY_y} Rft^{Init}_i }{H^{max}} \rceil
+        & t \in \mathcal{T}, y \in \mathcal{Y} \notag
+  \end{align}
+
+  So, in this way, we arrive to two lower bounds and two upper bounds per period and aircraft type.
+We then get the maximum and the minimum respectively to get some bounds on the number of checks until period $t$.
+$YM^{Acc}F^{\prime\prime b}_{yt}$ represents the bound $b$ in the number of total checks since the beginning of the planning horizon for all aircraft of type $y$ until period $t$.
+
+  \begin{align}
+    & YM^{Acc}F^{\prime\prime min}_{yt} = \max\{YM^{Acc}F^{min}_{yt}, YM^{Acc}F^{\prime min}_{yt}\}\notag\\ 
+    & YM^{Acc}F^{\prime\prime max}_{yt} = \min\{YM^{Acc}F^{max}_{yt}, YM^{Acc}F^{\prime max}_{yt}\}\notag
+  \end{align}
+
+  \[
+    QM^{num}_{tt'} = 
+    \begin{cases} 
+     1 & t' > t - M \\
+     2 & t' \le t - M \\
+    \end{cases} 
+  \]
+
+  $QM^{num}_{tt'}$ represents the possibility of counting two checks instead of one if both checks start before the period $t$. $t'$ represents the second one of those checks.
+
+  This can be applied via a cut by the following way:
+
+  \begin{align}
+    & YM^{Acc}F^{\prime\prime min}_{yt} \le \sum_{\substack{i \in \mathcal{IY}_y, \\ t_1 \in \{1, \ldots, t - M\}, \\ t_2 \in \mathcal{T}^M_{t_1}}} m_{it_1t_2} \times QM^{num}_{tt_2} \le YM^{Acc}F^{\prime\prime max}_{yt}
+        & t \in \mathcal{T}, y \in \mathcal{Y} \label{eq:range_maints}
+  \end{align}
+
+  Cuts (\ref{eq:range_maints}) limit the starts of checks of aircraft of type $y$ in order to have the number of finished checks to fall between the $YM^{Acc}F^{\prime\prime b}_{yt}$ bounds.
+
+\section{Bounds on accumulated checks per period}
+
+  The bounds on accumulated checks by type and period defined in \ref{sec:rangecheckstype} can be aggregated into the whole fleet.
+
+  \begin{align}
+    & M^{Acc}S^{b}_{t} = \sum_{y \in \mathcal{Y}} YM^{Acc}S^{\prime\prime b}_{yt}
+        & t \in \mathcal{T}, b \in \{min, max\}  \notag\\
+    & M^{Acc}F^{b}_{t} = \sum_{y \in \mathcal{Y}} YM^{Acc}F^{\prime\prime b}_{yt}
+        & t \in \mathcal{T}, b \in \{min, max\}  \notag
+  \end{align}
+
+  Additionally, the maintenance capacity together with the maintenance duration offer an upper bound on the maximum number of checks that can be finished until a given period $t$.
+
+  \begin{align}
+    & M^{Acc}F^{\prime max}_{t} = \lfloor \frac{t}{M} \rfloor \times C^{max}
+        & t \in \mathcal{T} \notag
+  \end{align}
+
 \end{document}

### 2019-04-18 latex/notes/bounds/bounds.tex
new version of bounds and model documents.

@@
 \renewcommand{\refname}{REFERENCES}
-
-\makeatletter
-\renewcommand\section{\@startsection{section}{1}{\z@}%
-                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
-                       {2\p@ \@plus 0\p@ \@minus 0\p@}%
-                       {\normalsize\textbf}}
-
-\renewcommand\section{\@startsection{section}{2}{\z@}%
-                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
-                       {2\p@ \@plus 0\p@ \@minus 0\p@}%
-                       {\normalsize\textbf}}
-
-\renewcommand\section{\@startsection{section}{3}{\z@}%
-                       {-6\p@ \@plus -0\p@ \@minus -0\p@}%
-                       {1\p@ \@plus 0\p@ \@minus 0\p@}%
-                       {\normalsize\itshape\bfseries}}
-\makeatother
 

### 2019-04-18 latex/notes/model/model.tex
new version of bounds and model documents.

@@
         \begin{tabular}{p{15mm}p{140mm}}
-            $i \in \mathcal{I}$     &  aircraft. \\
-            $t \in \mathcal{T}$     &  time periods included in the planning horizon. \\
-            $j \in \mathcal{J}$     &  missions. \\
+            $i \in \mathcal{I}$     &  aircraft. Between 15 and 60. \\
+            $t \in \mathcal{T}$     &  time periods included in the planning horizon. Between 90 and 120 months. \\
+            $j \in \mathcal{J}$     &  missions. There are between 1 and 4 active missions at any given time. \\
+        \end{tabular}
+
+    \subsubsection{Auxiliary sets}
+
+        \begin{tabular}{p{15mm}p{140mm}}
             $k \in \mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
+            $n \in \mathcal{N}$     &  maintenance cycles for aircraft. Up to 3. \\
+            $s \in \mathcal{S}$     &  slots for elastic variables. \\
         \end{tabular}
@@
             $A^{Clust}_{kt}$    & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$. & aircraft. \\
-            $H^{Clust}_{kt}$    & minimum number of total remaining flight time for cluster $k$ at period $t$. & hours. \\
+            $H^{Clust}_{kn}$    & maximum number of mean accumulated flight time for cluster $k$ at maintenance cycle $n$. & hours. \\
             $Rft^{Init}_i$      & remaining flight time for aircraft $i$ at the start of the planning horizon.  & hours.  \\
             $Rct^{Init}_i$      & remaining calendar time for aircraft $i$ at the start of the planning horizon.  & periods. \\
+        \end{tabular}
+
+    \subsubsection{Objective function weights}
+
+        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
+            $PA^{Clust}_s$   & unit penalty for violating serviceability constraint in slot $s$. & \\
+            $PH^{Clust}_s$   & unit penalty for violating sustainability constraint in slot $s$. & \\
         \end{tabular}
@@
     \begin{align}
-        & U^{min}_{t_1t_2} = U^{min} \sum_{t=t_1}^{t_2} (t_2 - t) \notag \\
-        & H'_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1) \notag \\
-        & H''_{jt_1t_2t'} = \sum_{t=t_1}^{t_2} (H_j- U^{min}) (t' - t) \notag
+        & U^{\prime}_{t_1t_2} = U^{min} (t_2 - t_1+1) \notag \\
+        & U^{\prime\prime}_{t_1t_2} = U^{min} \sum_{t=t_1-1}^{t_2} (t_2 - t) = U^{min} (t_2 - t_1+1) (t_2 - t_1) \div 2  \notag \\
+        & H^\prime_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1+1) \notag \\
+        & H^{\prime\prime}_{jt_1t_2t'} = (H_j- U^{min}) \sum_{t=t_1-1}^{t_2} (t' - t) = (H_j- U^{min}) (t_2 - t_1 + 1) (2 t' - t_1 - t_2) \div 2 \notag
     \end{align}
@@
         $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check at the beginning of period $t \in \mathcal{T}$ and then starts the next one at the beginning of period $t'$, 0 otherwise. \\
-        $h_{itt'}$  &  Continuous. Sum of the monthly remaining flight hours for aircraft $i \in I$ for periods between the beginning of period $t$ and the beginning of period $t'$. \\
+    \end{tabular}
+
+    Auxiliary variables control the status of each aircraft or group of aircraft.
+
+    \begin{tabular}{p{8mm}p{147mm}}
+        $h_{in}$  &  Continuous, positive. Sum of the accumulated monthly consumption in flight hours for aircraft $i \in I$ for maintenance cycle $n$. \\
+        $e^{AClust}_{kts}$ & Continuous, positive. Elastic variable for serviceability constraint. \\
+        $e^{HClust}_{kts}$ & Continuous, positive. Elastic variable for sustainability constraint. \\
     \end{tabular}
@@
     \begin{align}
-        & \text{Min}\; \sum_{\substack{i \in \mathcal{I}, \\ t_1 \in \mathcal{T}^{M_{Init}}_i, \\ t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|}} m_{it_1t_2} \times H^{max} \times (t_2 - t_1) - h_{it_1t_2}  \label{eq:objective1}
+        & \text{Min}\; 
+        \sum_{\substack{i \in \mathcal{I}, \\ 
+            t_1 \in \mathcal{T}^{M_{Init}}_i, \\ 
+            t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|}
+            }  
+            m_{it_1t_2} \times H^{max} 
+        + \sum_{k \in \mathcal{K}, t \in \mathcal{T}, s \in \mathcal{S}} PA^{Clust}_s e^{AClust}_{kts} 
+        + \sum_{k \in \mathcal{K}, n \in \mathcal{N}, s \in \mathcal{S}} PH^{Clust}_s e^{HClust}_{kns}
+        \label{eq:objective1}
     \end{align}
@@
     \begin{align}
-       & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt}
-        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}
-       % & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt}
-       %  &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
-    \end{align}
-
-    Constraints (\ref{eq:serviceability-cluster}) guarantee a minimum serviceability of aircraft for each cluster $k$.
-A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission.
-    % Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.
-
-    \begin{align}
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H'_{jtt'} + U^{min}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
+        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H'_{jtt'} + U^{\prime}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
                 & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours1}\\
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H'_{jtt'} + U^{\prime}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
                 & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours2}\\
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H'_{jtt'} + U^{min}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H'_{jtt'} + U^{\prime}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
                 & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours3}
@@
     \begin{align}
-        & h_{i1t_1} \geq (t_1 - 1) Rft^{Init}_i m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H''_{jtt't_1} - U^{min}_{1t_1} 
+        & h_{i1} (t_1 - 1) \geq (t_1 - 1) H^{max} (m_{it_1t_2} -1) + \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H''_{jtt't_1} + U^{\prime\prime}_{1t_1} 
             & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours1}\\
-        & h_{it_1t_2} \geq (t_2 - t_1) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H''_{jtt't_2} - U^{min}_{t_1t_2}
+        & h_{i2} (t_2 - t_1) \geq (t_2 - t_1) H^{max} (m_{it_1t_2} -1) + \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H''_{jtt't_2} + U^{\prime\prime}_{t_1t_2}
             &t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours2}\\
-        & h_{it_2|\mathcal{T}|} \geq (|\mathcal{T}|- t_2) H^{max} m_{it_1t_2} - \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H''_{jtt'|\mathcal{T}|} - U^{min}_{t_2|\mathcal{T}|}
+        & h_{i3} (|\mathcal{T}|- t_2) \geq (|\mathcal{T}|- t_2) H^{max} (m_{it_1t_2} -1) + \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H''_{jtt'|\mathcal{T}|} + U^{\prime\prime}_{t_2|\mathcal{T}|}
             & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours3}
     \end{align}
+
+    \begin{align}
+       & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt} + \sum_{s \in \mathcal{S}} e^{AClust}_{kts}
+        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}\\
+       & \sum_{i \in \mathcal{IK}_k} h_{in} \leq |{IK}_k| H^{Clust}_{kn} + \sum_{s \in \mathcal{S}} e^{HClust}_{kns}
+        &k \in \mathcal{K}, n \in \mathcal{N} \label{eq:sustainability-cluster}
+    \end{align}
+
+    Constraints (\ref{eq:serviceability-cluster}) guarantee a minimum serviceability of aircraft for each cluster $k$.
+A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission.
+    Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.
 
@@
           & i \in \mathcal{I}\label{eq:num_maint} \\
-        % & m_{it'} + m_{it} \leq 1
-        %   & t \in \mathcal{T}, t' \in \mathcal{T}^m_t, i \in \mathcal{I}\label{eq:rct_min}\\  
-        % & \sum_{t' \in \mathcal{T}^M_t} m_{it'} \geq  m_{it}
-        %   & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rct_max}\\
-        % & m_{it_1t_2} = 0
-        %   & t_1 \in \mathcal{T}^{m_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \cup \{|\mathcal{T}|\}, i \in \mathcal{I}\label{eq:rct_min_init} \\
         & m_{it_1t_2} =  0

### 2019-04-19 latex/notes/model/model.tex
changes on model.tex

@@
 \setlength{\parskip}{2ex} \pagestyle{myheadings}
-% \markright{\hspace*{3.8cm} \textit{MOSIM18 - June 27-29, 2018 - Toulouse - France}}
 
@@
         \begin{tabular}{p{15mm}p{125mm}p{15mm}}
-            $PA^{Clust}_s$   & unit penalty for violating serviceability constraint in slot $s$. & \\
-            $PH^{Clust}_s$   & unit penalty for violating sustainability constraint in slot $s$. & \\
+            $PA_s$   & unit penalty for violating serviceability constraint in slot $s$. & \\
+            $PH_s$   & unit penalty for violating sustainability constraint in slot $s$. & \\
+            $PC_s$           & unit penalty for violating capacity constraint in slot $s$. & \\
         \end{tabular}
@@
         $h_{in}$  &  Continuous, positive. Sum of the accumulated monthly consumption in flight hours for aircraft $i \in I$ for maintenance cycle $n$. \\
-        $e^{AClust}_{kts}$ & Continuous, positive. Elastic variable for serviceability constraint. \\
-        $e^{HClust}_{kts}$ & Continuous, positive. Elastic variable for sustainability constraint. \\
+        $e^{A}_{kts}$ & Continuous, positive. Elastic variable for serviceability constraint (\ref{eq:serviceability-cluster}). \\
+        $e^{H}_{kts}$ & Continuous, positive. Elastic variable for sustainability constraint (\ref{eq:sustainability-cluster}). \\
+        $e^{C}_{ts}$ & Continuous, positive. Elastic variable for capacity constraint (\ref{eq:capacity1}). \\
     \end{tabular}
@@
 
-    Objective (\ref{eq:objective1}) minimizes the number of checks while at the same time maximizing the monthly total flight hours potential of the fleet.
+    Objective (\ref{eq:objective1}) minimizes the deviations number of checks while at the same time maximizing the monthly total flight hours potential of the fleet.
 
@@
         & \text{Min}\; 
-        \sum_{\substack{i \in \mathcal{I}, \\ 
-            t_1 \in \mathcal{T}^{M_{Init}}_i, \\ 
-            t_2 \in \mathcal{T}^M_{t_1} - |\mathcal{T}|}
+        \sum_{\substack{
+                k \in \mathcal{K}, \\ t \in \mathcal{T}, \\ s \in \mathcal{S}
                 }
-            m_{it_1t_2} \times H^{max} 
-        + \sum_{k \in \mathcal{K}, t \in \mathcal{T}, s \in \mathcal{S}} PA^{Clust}_s e^{AClust}_{kts} 
-        + \sum_{k \in \mathcal{K}, n \in \mathcal{N}, s \in \mathcal{S}} PH^{Clust}_s e^{HClust}_{kns}
+            } PA_s e^{A}_{kts} 
+        + \sum_{\substack{
+                k \in \mathcal{K}, \\ n \in \mathcal{N}, \\ s \in \mathcal{S}
+                } 
+            } PH_s e^{H}_{kns}
+        + \sum_{t \in \mathcal{T}, s \in \mathcal{S}} PC_s e^{C}_{ts}
         \label{eq:objective1}
     \end{align}
-
-    The first term counts all the remaining flight hours given to aircraft following checks and the second term quantifies the amount of remaining flight hours for all aircraft at each period.
-These two objectives have the same units, can be easily compared and ensure the aircraft are used in the most efficient way.
 
@@
         % maximum capacity1:
-        & \sum_{\substack{i \in \mathcal{I}, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N_t \leq C^{max}
+        & \sum_{\substack{i \in \mathcal{I}, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N_t \leq C^{max} + \sum_{s \in \mathcal{S}} e^{C}_{ts}
                 & t \in \mathcal{T} \label{eq:capacity1}\\
@@
     \begin{align}
-       & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt} + \sum_{s \in \mathcal{S}} e^{AClust}_{kts}
+       & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt} + \sum_{s \in \mathcal{S}} e^{A}_{kts}
         &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}\\
-       & \sum_{i \in \mathcal{IK}_k} h_{in} \leq |{IK}_k| H^{Clust}_{kn} + \sum_{s \in \mathcal{S}} e^{HClust}_{kns}
+       & \sum_{i \in \mathcal{IK}_k} h_{in} \leq |{IK}_k| H^{Clust}_{kn} + \sum_{s \in \mathcal{S}} e^{H}_{kns}
         &k \in \mathcal{K}, n \in \mathcal{N} \label{eq:sustainability-cluster}

### 2019-04-22 latex/notes/model/model.tex
changes to model and bounds docs.

@@
             $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$ (figure \ref{fig:gantt_windows}a). \\
-            $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}b). \\
-            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}b).
+            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$  (figure \ref{fig:gantt_windows}b).
 \\
-            $t' \in \mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}c). \\
-            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$. \\
-            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$.
+            $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT_j-1$ and $t \ge t_1 + M$ and $t' < t_2$  (figure \ref{fig:gantt_windows}b).\\
+            $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}c). \\
+            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}c).
 \\
-            $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT_j-1$ and $t \ge t_1 + M$ and $t' < t_2$ \\
+            $t' \in \mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}d). \\
+            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$ (figure \ref{fig:gantt_windows}d).
+\\
         \end{tabular}
@@
 $\mathcal{T}_A \cup \mathcal{T}_B$ represents the set of periods $t$ for which $(t_1, t_2) \in \mathcal{TTT}_t$.
+Also, $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ since the assignment is bounded by $t_1 + M$ and $t_2$.
 (c) $Rct_{i}^{Init}$ is equal to 55, meaning a check has to be planned between months number 25 and 54.
@@
 In this case, the size is 6.
-The mission assignments for aircraft $i$ to mission $j$ are $a_{ijt_1t_2}=1$ and $a_{ijt'_1t'_2}=1$}
+The mission assignments for aircraft $i$ to mission $j$ are $a_{ijt_1t_2}=1$ and $a_{ijt'_1t'_2}=1$ and $(t'_1, t'_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt''}$.}
         \label{fig:gantt_windows}

### 2019-04-22 latex/notes/model/model.tex
take out cycles from flight hours accounting

--- 2019-04-19 latex/notes/model/model.tex
@@
             $i \in \mathcal{IK}_k$     &  aircraft $i \in \mathcal{I}$ that are included in cluster $k$. One aircraft can belong to more than one cluster. \\
+            $j \in \mathcal{JK}_k$     &  missions $j \in \mathcal{I}$ that are included in cluster $k$. One mission can only belong to one cluster. \\
             $j \in \mathcal{JI}_i$     &  missions $j \in \mathcal{J}$ for which aircraft $i$ can be used. \\
@@
         & H^\prime_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1+1) \notag \\
-        & H^{\prime\prime}_{jt_1t_2t'} = (H_j- U^{min}) \sum_{t=t_1-1}^{t_2} (t' - t) = (H_j- U^{min}) (t_2 - t_1 + 1) (2 t' - t_1 - t_2) \div 2 \notag
+        & H^{\prime\prime}_{jt_1t_2t'} = (H_j- U^{min}) \sum_{t=t_1-1}^{t_2} (t' - t) = (H_j- U^{min}) (t_2 - t_1 + 1) (2 t' - t_1 - t_2) \div 2 \notag \\
+        & H^{Clust}_{k} = \sum_{j \in \mathcal{JK}_k} H_j
     \end{align}
@@
         $e^{A}_{kts}$ & Continuous, positive. Elastic variable for serviceability constraint (\ref{eq:serviceability-cluster}). \\
-        $e^{H}_{kts}$ & Continuous, positive. Elastic variable for sustainability constraint (\ref{eq:sustainability-cluster}). \\
+        $e^{H}_{ks}$ & Continuous, positive. Elastic variable for sustainability constraint (\ref{eq:sustainability-cluster}). \\
         $e^{C}_{ts}$ & Continuous, positive. Elastic variable for capacity constraint (\ref{eq:capacity1}). \\
@@
         + \sum_{\substack{
-                k \in \mathcal{K}, \\ n \in \mathcal{N}, \\ s \in \mathcal{S}
+                k \in \mathcal{K}, \\ s \in \mathcal{S}
                 } 
-            } PH_s e^{H}_{kns}
+            } PH_s e^{H}_{ks}
         + \sum_{t \in \mathcal{T}, s \in \mathcal{S}} PC_s e^{C}_{ts}
@@
         &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}\\
-       & \sum_{i \in \mathcal{IK}_k} h_{in} \leq |{IK}_k| H^{Clust}_{kn} + \sum_{s \in \mathcal{S}} e^{H}_{kns}
-        &k \in \mathcal{K}, n \in \mathcal{N} \label{eq:sustainability-cluster}
+       & \sum_{i \in \mathcal{IK}_k, n \in \mathcal{N}} h_{in} \leq H^{Clust}_{k} + \sum_{s \in \mathcal{S}} e^{H}_{ks}
+        &k \in \mathcal{K} \label{eq:sustainability-cluster}
     \end{align}

### 2019-08-19 R/scripts/resultsNPS.Rmd
added markdown for results reporting.

---
title: "ResultsNPS"
author: "Franco Peschiera"
date: "August 8, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
use_python('C:/Python36/python.exe')
use_virtualenv('C:/Users/pchtsp/Documents/projects/optima/python/venv/', required = TRUE)
current_dir = getwd()
print(current_dir)
# Sys.getenv('PYTHONHOME')
# # Sys.setenv(PYTHONPATH='C:/Python36/DLLs')
# py_available(TRUE)
py_discover_config()

```

```{python}
import pandas as pd
import numpy as np

import sys, os
path = os.path.join(r.current_dir, './../../python/')
print(path)
sys.path.insert(1, path)

import package.batch as ba
# import package.params as params

# from rpy2.robjects import pandas2ri
# import rpy2.robjects.lib.ggplot2 as ggplot2
# import rpy2.robjects as ro

# import stochastic.params as sto_params

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2019-08-20 R/scripts/resultsNPS.Rmd
automatic rmarkdown report using python.

@@
   chunk_output_type: console
+params:
+   size: medium
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
 library(reticulate)
-use_python('C:/Python36/python.exe')
-use_virtualenv('C:/Users/pchtsp/Documents/projects/optima/python/venv/', required = TRUE)
+library(tidyverse)
+library(knitr)
+library(magrittr)
+library(ggplot2)
+use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)
 current_dir = getwd()
-print(current_dir)
-# Sys.getenv('PYTHONHOME')
-# # Sys.setenv(PYTHONPATH='C:/Python36/DLLs')
-# py_available(TRUE)
+# print(current_dir)
 py_discover_config()
+quant_max <- 0.95
+size <- params$size
+aux_compare <- function(table){
+    table %>% 
+    gather(key='indicator', value="value", -experiment) %>% 
+    spread(experiment, value) %>% 
+    mutate(dif_abs = cuts - base,
+           dif_perc = (dif_abs/ base*100) %>% round(2))
+
+}
 
@@
 import sys, os
-path = os.path.join(r.current_dir, './../../python/')
-print(path)
+opt_path = './'
+opt_path = '../'
+rel_path_python = '{}../python/'.format(opt_path)
+path = os.path.join(r.current_dir, rel_path_python)
+# print(path)
 sys.path.insert(1, path)
@@
 import package.batch as ba
-# import package.params as params
+import package.params as params
 
-# from rpy2.robjects import pandas2ri
-# import rpy2.robjects.lib.ggplot2 as ggplot2
-# import rpy2.robjects as ro
+import stochastic.params as sto_params
+import scripts.compare_stochastic as compare_sto
 
-# import stochastic.params as sto_params
+# This is read now as a paramer in the YAML (above).
+# size = 'small'  # choose small, medium or large
+df = compare_sto.get_instances[r.size]()
+df = df.reset_index()
+
+name_change = dict(experiment=[0, 1], experiment2=["cuts", "base"])
+name_change = pd.DataFrame(name_change)
+df = df.merge(name_change, on='experiment').drop('experiment', axis=1).rename(columns=dict(experiment2='experiment'))
 
@@
 
-## R Markdown
+## Nomenclature
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+* `cuts` model with added learning constraints.
+* `base` model without any added constraints.
+* `dif_abs` absolute distance between the two models for the shown quantity.
+* `dif_rel` relative distance between the two models for the shown quantity. Always using (cuts-base)/base.
+* `time_mean` mean time.
+* `time_medi` median time.
+* `errors_q95` number of errors in the percentile 95%.
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+## General statistics
 
-```{r cars}
-summary(cars)
+The following table shows the general statistics about the instances solved.
+
+```{r stats_general}
+
+summary_stats <- 
+    py$df %>% 
+    mutate(sol_code = if_else(is.na(sol_code), 2, sol_code)) %>% 
+    group_by(experiment) %>%
+    summarise(Infeasible = sum(sol_code==-1),
+              IntegerFeasible=sum(sol_code==2),
+              IntegerInfeasible=sum(sol_code==0),
+              Optimal=sum(sol_code==1),
+              total_instances = n()) %>% 
+  aux_compare
+
+summary_stats %>% kable
+
 ```
 
-## Including Plots
+## Solution quality
 
-You can also embed plots, for example:
+### Optimality degradation
 
-```{r pressure, echo=FALSE}
-plot(pressure)
+We compare the times where both models return an **optimal solution**.
+
+```{r quality_degr}
+
+t1 <- 
+    py$df %>% 
+    filter(sol_code==1) %>% 
+    group_by(instance) %>%
+    mutate(min_value = min(best_solution),
+           dist_min = best_solution - min_value,
+           dist_min_perc = (best_solution - min_value)/abs(min_value)*100
+           )
+t1_rel <- 
+    t1 %>% 
+    select(instance, experiment, dist_min_perc) %>% 
+    spread(key=experiment, value=dist_min_perc) %>% 
+    drop_na() %>% use_series(cuts)
+
+t1_abs <- 
+    t1 %>% 
+    select(instance, experiment, dist_min) %>% 
+    spread(key=experiment, value=dist_min) %>% 
+    drop_na() %>% use_series(cuts)
+
+quality_degr_quant <- t1_rel %>% quantile()
+quality_degr_quant95 <- t1_rel %>% quantile(quant_max)
+quality_degr_nrow <- t1_rel %>% length()
+quality_degr_quant95_abs <- t1_abs %>% quantile(quant_max)
+
 ```
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+For `r quant_max*100`\% of the `r quality_degr_nrow` (`r (quant_max*quality_degr_nrow) %>% floor`) instances that comply with this condition, the loss in the objective function was `r quality_degr_quant95 %>% round(2)`\% or less from the optimal.
+This corresponds to an absolute value of `r quality_degr_quant95_abs %>% round(2)`.
+As a reference, the absolute gap tolerance used in the solving was of 10.
+
+```{r}
+t1_rel_filt <- t1_rel[t1_rel<20]
+filtered_quant <- (1 - (t1_rel_filt %>% length) / (t1_rel %>% length))*100
+```
+
+The following graph shows the distribution of the gap from the `cuts` model with the `base` model. `r filtered_quant %>% round(2)`\% of outliers were taken out from the right tail.
+
+```{r}
+qplot(t1_rel[t1_rel<20], xlab='Relative gap (in %) among optimal solutions.')
+```
+
+### Mean quality comparison
+
+If we compare among all cases where an **integer solution** was found in the two models, we can obtain an estimate of quality performance for the 1 hour solving time.
+
+```{r quality_perf}
+
+quality_perf <-
+    py$df %>% 
+    filter(sol_code>=1) %>%
+    select(instance, experiment, best_solution) %>% 
+    spread(experiment, best_solution) %>% 
+    drop_na() %>% 
+    mutate(dif_perc = ((cuts-base)/ base*100) %>% round(2))
+
+  mean_dif <- quality_perf$dif_perc %>%  mean
+  per_cut <- sum(quality_perf$dif_perc<0)/length(quality_perf$dif_perc)
+
+quality_perf_nrow <- quality_perf %>% distinct(instance) %>% nrow
+
+```
+
+```{r}
+filtered_q_perf <- quality_perf %>% filter(between(dif_perc, -5, 5))
+filtered_quant <- (1 - (filtered_q_perf %>% nrow)/ (quality_perf %>% nrow))*100
+```
+The following graph shows the distribution of relative gaps between the best integer solutions found in each of the two models.
+A negative relative gap means the `cuts` model outperformed the `base` model in that particular instance.
+Both extreme tails (`r filtered_quant %>% round(2)`\%̀) have been taken out to help better visualize it.
+
+```{r}
+qplot(filtered_q_perf$dif_perc, xlab='Relative gap (in %) among integer solutions.')
+```
+
+Of the `r quality_perf_nrow` instances were an **integer solution** was found in both models, the average relative gap between the `cuts` model and the `base` model was of `r mean_dif %>% round(2)`\%.
+This means that, in average solutions of the `cuts` model were `r (-mean_dif) %>% round(2)`\% better than the `base` model.
+
+In addition, in `r (per_cut*100) %>% round(2)`\% of instances (`r (quality_perf_nrow*per_cut) %>% floor` of `r quality_perf_nrow`) the `cuts` model outperform the solution from the base model.
+
+## Performance analysis
+
+### Average solving time
+
+Just like in the mean quality comparison, we compare among cases where both models found a **integer solution**.
+
+```{r time_perf}
+
+comparison_table <- 
+    py$df %>% 
+    filter(sol_code>=1) %>% 
+    group_by(instance) %>%
+    filter(n()==2) %>% 
+    ungroup() %>% 
+    select(instance, experiment, time) %>% 
+    spread(experiment, time) %>% 
+    arrange(base) %>% 
+    mutate(instance = row_number()) %>% 
+    gather(key = 'experiment',  value='time', -instance)
+
+comparison_table_reorder <- 
+    comparison_table %>% 
+    group_by(experiment) %>% 
+    arrange(experiment, time) %>% 
+    mutate(percentage = row_number()/n()*100)
+
+mean_times <- 
+    comparison_table %>% 
+    group_by(experiment) %>% 
+    summarise(time_mean = mean(time), 
+              time_medi = median(time)) %>% 
+    aux_compare
+
+```
+
+The difference in average and median times can be seen in the following table:
+
+```{r}
+    mean_times %>% kable
+```
+
+In order to get more intuition in the performance per instance, we present a couple of graphs.
+
+The following graph shows the solving times in seconds for each of the `r t1 %>% distinct(instance) %>% nrow` instances were an integer solution was found in both cases.
+In red are the times for the base model.
+In green the times for the model with the added cuts.
+
+```{r time_perf_graph} 
+    ggplot(data=comparison_table, aes(x=instance, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5)
+```
+
+If we reorder both models from fastest to slowest, we lose the comparison of individual instances but we gain more insight on the distribution of solving times.
+
+```{r time_perf_graph2} 
+    ggplot(data=comparison_table_reorder, aes(x=percentage, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5) + ggplot2::xlab('instance percentage')
+```
+
+### Average solving time to optimality
+
+Here, we measure the time it took for each model to reach an **optimal solution.** We compare among instances where both models obtained and optimal solution.
+
+```{r time_perf_optim}
+
+comparison_table <- 
+    py$df %>% 
+    filter(sol_code==1) %>% 
+    group_by(instance) %>% 
+    filter(n()==2) %>% 
+    select(experiment, instance, time) %>%
+    group_by(experiment) %>% 
+    summarise(time_mean = mean(time), 
+              time_medi = median(time)) %>% 
+    aux_compare
+
+```
+
+```{r time_perf_optim_graph}
+    comparison_table %>% kable
+```
+
+## Feasiblility analysis
+
+There are two indicators to measure if a model is too constrained. The first one is the fact the model is infeasible. The second one is the accomplishement of the soft constraints in the solution.
+
+```{r infeasible_list}
+
+infeasible_instances <- py$df %>% filter(sol_code==-1) %>% distinct(instance)
+```
+
+### Infeasible solutions
+
+We take all `r infeasible_instances %>% nrow` **infeasible solutions** from the cuts model and obtain statistics on their solution on the base model.
+
+The below table shows the status obtained in the base model.
+The number of infeasible solutions is very close in both models and for all instances that were identified by infeasible in the `cuts` model, the `base` model was unable to find an integer solution.
+
+```{r infeasible}
+infeasible_stats <- 
+    py$df %>% 
+    inner_join(infeasible_instances) %>% 
+    filter(experiment=="base") %>% 
+    summarise(Infeasible = sum(sol_code==-1),
+              IntegerFeasible=sum(sol_code==2),
+              IntegerInfeasible=sum(sol_code==0),
+              Optimal=sum(sol_code==1),
+              total_instances = n()) %>% 
+    gather(key='status', 'number') %>% 
+  filter(number>0)
+
+infeasible_stats %>% kable
+```
+
+### Soft constraints performance
+
+Regarding the accomplishment of soft constraints, we can analyze for mutually **integer solutions**, the quantity of violations in their solutions.
+
+```{r soft_constraints}
+errors_stats <- 
+    py$df %>% 
+    filter(sol_code>=1) %>% 
+    mutate(errors = replace_na(errors, 0)) %>% 
+    group_by(experiment) %>% 
+    summarise(errors_mean = mean(errors),
+              errors_q95 = quantile(errors, quant_max),
+              errors_medi = median(errors)) %>% 
+    aux_compare
+
+```
+
+The following table shows the difference in soft constraints errors.
+
+```{r soft_constraints_table}
+errors_stats %>% kable
+```

### 2019-08-20 python/scripts/compare_stochastic.py

@@ -4,10 +4,6 @@ import numpy as np
 import package.batch as ba
 import package.params as params
 
-from rpy2.robjects import pandas2ri
-import rpy2.robjects.lib.ggplot2 as ggplot2
-import rpy2.robjects as ro
-
 import stochastic.params as sto_params
 
 #####################
@@ -40,36 +36,47 @@ def compare_perc_0_1(table):
     return (table[0] - table[1])/table[1] *100
 
 
-if __name__ == '__main__':
+def get_large_instances():
+    exp_list = \
+        ['IT000125_20190730', 'IT000125_20190801']
+    df = get_df_comparison(exp_list)
+    df = df[df.scenario == 'numparalleltasks_3']
+    return df
 
-    ro.pandas2ri.activate()
+def get_medium_instances():
+    exp_list = \
+        ['IT000125_20190730', 'IT000125_20190729']
 
-    # get info to analyze.
+    df = get_df_comparison(exp_list)
+    df = df[df.scenario == 'numparalleltasks_2']
+    return df
 
+def get_small_instances():
     exp_list = \
         ['IT000125_20190725',
          'IT000125_20190716']
 
     df = get_df_comparison(exp_list)
+    return df
 
-    # 2 tasks
-    exp_list = \
-        ['IT000125_20190730', 'IT000125_20190729']
+get_instances = dict(small= get_small_instances,
+                     medium= get_medium_instances,
+                     large= get_large_instances)
 
-    df = get_df_comparison(exp_list)
+if __name__ == '__main__':
 
-    df = df[df.scenario == 'numparalleltasks_2']
+    from rpy2.robjects import pandas2ri
+    import rpy2.robjects.lib.ggplot2 as ggplot2
+    import rpy2.robjects as ro
 
-    # 3 tasks
-    exp_list = \
-        ['IT000125_20190730', 'IT000125_20190801']
+    ro.pandas2ri.activate()
 
-    df = get_df_comparison(exp_list)
-
-    df = df[df.scenario == 'numparalleltasks_3']
+    # get info to analyze.
+    # df = get_small_instances()
+    # df = get_medium_instances()
+    df = get_large_instances()
 
     df[df.index.get_level_values('experiment')==1]
-
     # GENERAL STATS
     # df.agg('mean')[['time', 'best_solution']]
     df.groupby(['experiment', 'status']).agg('count')['name']

### 2019-08-26 R/presentations/NPS_Work/presentation.Rmd
added presentation done in baobab.

---
title: "What I've been doing at NPS for three months"
author: "Franco Peschiera"
date: "MM, 2019"
output:
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    slide_level: 2
    toc: true
---

## Problem

## Problem (informally)

Assign both missions and maintenance operations to a fleet of aircraft in order to maximize availability and minimize costs. Missions have fixed start and end times and have particular needs in terms of aircraft and time.

## Problem (example)

```{r, echo=FALSE, out.width='70%'}
knitr::include_graphics('./../../../img/calendar3.png')

```

## Problem

* A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. 
* Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
* Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H$ units.

## Old variables

$a_{jti}$   &  =1 if mission $j \in J$ in period $t \in \mathcal{T}_j$ is realized with aircraft $i \in \mathcal{I}_j$, 0 otherwise.  
$a^s_{jti}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-1)i} = 0$.  
$m_{it}$    &  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise.

```{r, echo=FALSE, out.width='70%'}
knitr::include_graphics('./../../../img/calendar3.png')

```

## New variables

$a_{ijtt'}$ & has value one if aircraft $i$ starts an assignment to mission $j$ at the beginning of period $t$ and finishes at the end of period $t'$, zero otherwise.  
$m_{itt'}$  & has value one if aircraft $i$ starts a check at the beginning of period $t$ and then starts the next check at the beginning of period $t'$ or does not have a second check ($t'=T$), zero otherwise.

```{r, echo=FALSE, out.width='70%'}
knitr::include_graphics('./../../../img/calendar3.png')

```

## Distance between maintenance

* The distance between maintenance has a maximum of 60 periods.
* Depending on the amount of missions to do in a planning horizon, this can be shorter.
* The more room we leave as possible distance between maintenances, the more maintenance variables we have.

```{r, echo=FALSE, out.width='40%'}
imgs = c('mean_consum_vs_maints_dell_20190515_all.png', 'mean_consum_vs_cycle_2M_min_dell_20190515_all.png')
knitr::include_graphics(imgs)

```

## Forecasting + Optim

1. We train a statistical model to predict the expected distances between maintenances for any given instance.
2. We use this information to limit the possible combinations of two maintenances for the fleet.

## Benefits

1. **Performance**: a smaller model is easier to solve.
2. **User feedback**: direct feedback about the solution without needing to solve any model.

### 2019-09-09 R/presentations/NPS_Work/presentation.Rmd
new changes on reports and corrected superquantiles.

@@
 author: "Franco Peschiera"
-date: "MM, 2019"
+date: "Agusut, 2019"
 output:

### 2019-09-09 R/scripts/resultsNPS.Rmd
new changes on reports and corrected superquantiles.

@@
 date: "August 8, 2019"
-output: html_document
+output:
+  html_document: default
+  pdf_document: default
+params:
+  size: medium
 editor_options:
   chunk_output_type: console
-params:
-   size: medium
 ---
@@
 opt_path = './'
-opt_path = '../'
+# opt_path = '../'
 rel_path_python = '{}../python/'.format(opt_path)
@@
 # size = 'small'  # choose small, medium or large
-df = compare_sto.get_instances[r.size]()
-df = df.reset_index()
+# df = compare_sto.get_instances[r.size]()
+
+################# manual alternative...
+# exp_list = ['IT000125_20190808', 'IT000125_20190729']
+# df = compare_sto.get_df_comparison(exp_list)
+# df = df[df.scenario == 'numparalleltasks_2']
+
+exp_list = ['IT000125_20190808', 'IT000125_20190812']
+df = compare_sto.get_df_comparison(exp_list).reset_index()
+# df = df[df.scenario == 'numparalleltasks_3']
+df = df.query("scenario == 'numparalleltasks_3' or experiment==1")
+################# manual alternative...
 
@@
 A negative relative gap means the `cuts` model outperformed the `base` model in that particular instance.
-Both extreme tails (`r filtered_quant %>% round(2)`\%̀) have been taken out to help better visualize it.
+Both extreme tails (`r filtered_quant %>% round(2)`\%) have been taken out to help better visualize it.
 

### 2019-09-09 R/scripts/renderNPS_report.R

@@ -3,9 +3,12 @@ library(rmarkdown)
 # paste0(size,"_report_", Sys.Date(), ".html")
 options <- c('medium', 'large', 'small')
 # options <- c('medium')
+format <- 'pdf'
+format <- 'html'
 for (size in options){
     rmarkdown::render("scripts/resultsNPS.Rmd", 
-                      output_file = sprintf('NPS_report_%s.html', size),
+                      output_file = sprintf('NPS_report_%s.%s', size, format),
                       params = list(size = size), 
-                      output_options = list(self_contained = TRUE))
+                      output_options = list(self_contained = TRUE),
+                      output_format=sprintf('%s_document', format))
 }

### 2019-09-09 python/scripts/stochastic_analysis.py

@@ -85,7 +85,8 @@ def get_table(cases, _types=1):
             l_maint_date = sol_stats.get_post_2M_dist(case, _type=_type).values_l()
             l_maint_date_stat = pd.Series(l_maint_date).agg(['mean', 'max', 'min', 'sum']).tolist()
             num_maints = sol_stats.get_num_maints(case, _type=_type)
-            result.append(list(_case_name) + init_sum +
+            result.append(list(_case_name) +
+                          init_sum +
                           cons_sum +
                           airc_sum +
                           cons_min_assign + quantsw +
@@ -101,7 +102,8 @@ def get_table(cases, _types=1):
                           cycle_1M_quants +
                           l_maint_date_stat)
 
-    names = ['scenario', 'name', 'init',
+    names = ['scenario', 'name']+ \
+            ['init',
              'mean_consum', 'max_consum', 'var_consum',
              'mean_airc', 'max_airc', 'var_airc',
              'cons_min_mean', 'cons_min_max', 'quant5w', 'quant75w', 'quant9w',
@@ -137,37 +139,44 @@ def treat_table(result_tab):
     return result_tab
 
 
-# #########
-# BOUNDS:
-# #########
-def superquantiles():
-    # x_vars += ['pos_consum9', 'pos_consum5', 'quant5w', 'quant75w', 'quant9w']
-    bound_options = [
-        ('maints', True, 0.9),
-        ('cycle_2M_min', False, 0.9),
-        ('mean_dist', False, 0.9),
-        ('mean_2maint', True, 0.9),
-        ('sum_2maint', True, 0.9),
-        ('sum_2maint', False, 0.9),
-        ('mean_2maint', False, 0.9)
-    ]
-    table = result_tab
-    # table = result_tab.query('mean_consum >=180 and gap_abs <= 30')
-    # table = result_tab[result_tab.mean_consum.between(150, 300)]
-    table = result_tab.query('mean_consum >=150 and mean_consum <=250')
-    # table = result_tab.query('mean_consum >=150 and mean_consum <=250 and num_errors==0')
+def superquantiles(table, **kwargs):
+    x_vars = ['init',
+             'mean_consum', 'max_consum', 'var_consum',
[…]

### 2019-09-09 python/stochastic/model_upper_bound.py

@@ -14,6 +14,13 @@ def regression_VaR(X, Y, _lambda = 0.1, alpha = 0.95):
     :param alpha:
     :return:
     """
+    upper_bound = True
+
+    if alpha < 0.5:
+        upper_bound = False
+        alpha = 1 - alpha
+        Y *= -1
+
     y = Y.tolist()
     X_in = X
 
@@ -66,6 +73,10 @@ def regression_VaR(X, Y, _lambda = 0.1, alpha = 0.95):
 
     c_0 = ((sum_ps - alpha)*Y_aux[limit] + sum(Y_aux[limit+1:])/len_points)/(1- alpha)
     # c_0 = sum(Y_aux[limit:])/(len_points - limit)
+    if not upper_bound:
+        c_0 *= -1
+        c_out = sd.SuperDict.from_dict(c_out).vapply(lambda v: -v)
+
     return c_0, c_out
 
 # sd.SuperDict(up).vapply(pl.value)

### 2019-09-09 python/stochastic/models.py

@@ -71,62 +71,27 @@ def predict_factory(X_train, y_train, method='regression', **kwargs):
         mod = smrq.QuantReg(endog=y_train, exog=X_train)
         res = mod.fit(**kwargs)
         return res
+    elif method == 'superquantiles':
+        coef0, coefs = mub.regression_VaR(X=X_train, Y=y_train, **kwargs)
+
+        class Clf(object):
+
+            def __init__(self, coefs, coef0):
+                self.coefs = coefs
+                self.coef0 = coef0
+                self.params = dict(coefs)
+                self.params['intercept'] = coef0
+                self.params = pd.Series(self.params)
+
+            def predict(self, new_x_data):
+                return np.sum([v * new_x_data[k] for k, v in self.coefs.items()], axis=0) + self.coef0
+
+        return Clf(coefs, coef0)
     else:
         raise ValueError('method argument has no correct value.')
     clf.fit(X=X_train, y=y_train)
     return clf
 
-
-def test_superquantiles(result_tab, x_vars, predict_var, plot=True, upper_bound=True, _print=True, **kwargs):
-
-    result_tab_f =  result_tab[x_vars + [predict_var]].copy()
-    bound = 'upper'
-    if not upper_bound:
-        result_tab_f[predict_var] *= -1
-        bound = 'lower'
-    result_tab_norm = aux.normalize_variables(result_tab_f)
-    mean_std = aux.get_mean_std(result_tab_f)
-    X_norm = result_tab_norm[x_vars].copy()
-    Y_norm = result_tab_norm[predict_var].copy()
-    # X_norm = normalize_variables(X)
-    # mean_std = get_mean_std(X)
-    # Y_norm = normalize_variables(Y)
-    # mean_std_y = get_mean_std(Y)
-    X_train, X_test, y_train, y_test = train_test_split(X_norm, Y_norm, test_size=0.3)
-    coef0, coefs = mub.regression_VaR(X=X_train, Y=y_train, **kwargs)
-    if not upper_bound:
-        coef0 *= -1
-        coefs = sd.SuperDict.from_dict(coefs).vapply(lambda v: -v)
-    coefs_sans_inter = sd.SuperDict(coefs)
-    coefs.update({'intercept': coef0})
-    if not _print:
-        return coefs
-    X_out_norm = X_test.copy()
-    X_out = aux.denormalize(X_out_norm, mean_std)
[…]

### 2019-09-10 R/presentations/DISC_201909/presentation.Rmd
added presentation to DISC, many changes in R reports and graphs, some code corrections on batch instances in python

---
title: "Predicting maintenance patterns"
author: "Franco Peschiera"
date: "September, 2019"
output:
  xaringan::moon_reader:
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    slide_level: 3
    toc: true
editor_options: 
  chunk_output_type: console
---

### Structure of talk

1. Decision problem.
2. Prediction problem.
2. Current results.
3. About my ambition.

---

## Decision problem

---

### Problem (informally)

Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs. **Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours. **Maintenances operations** have specific rules that govern their frequency and capacity.

--

.center[
```{r, echo=FALSE, out.width='100%'}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
knitr::include_graphics('./../../../img/calendar3.png')

```
]
---

### Problem (pseudo formally)

* A series of $j \in \mathcal{J}$ missions are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all missions are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. 
* Each mission requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
* Resources (heterogeneous) require recurrent preventive maintenance operations since the realization of missions diminish their remaining usage time. There are also min (max) distances between maintenances.
* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H$ units. There is a limited capacity for maintenances.

--

* **Other**: soft constraints on the state of the fleet at each period and at the end, minimum assignments on missions, initial conditions, default consumption.

---

### Variables (subset)

* $a_{ijtt'}$ : has value one if aircraft $i$ starts an assignment to mission $j$ at the beginning of period $t$ and finishes at the end of period $t'$, zero otherwise.  
* $m_{ip}$ : has value one if aircraft $i$ uses pattern $p$, zero otherwise.
  * each pattern $p$ has a single feasible combination of maintenance starts for an aircraft during the whole planning (usually only two checks per aircraft).

--

.center[

```{r, echo=FALSE, out.width='100%'}
knitr::include_graphics('./../../../img/calendar3.png')

```
] 
---

## Prediction problem

---

### Distance between maintenances

.pull-left[
* The distance between maintenance has a maximum of 60 periods.
* Depending on the amount of missions to do in a planning horizon, this can be shorter.
* The more room we leave as possible distance between maintenances, the more patterns and, as a result, maintenance variables we have.

]

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('hist_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

---

### Forecasting + Optim

**We want to**:

1. Train a statistical model to predict the mean distance between maintenances for any given instance.
2. Use this information to limit the possible combinations of patterns to assign.

--

**Benefits**:

1. **Performance**: a smaller model is easier to solve.
2. **User feedback**: direct feedback about the solution without needing to solve any model.
3. **More stable solutions**: Every aircraft flies an amount that is closest to the mean of the fleet.

--

**The better we're able to predict the optimal distance between maintenances for the whole fleet, the less optimality we will lose**

---

### Implementation

* Done in python with scikit learn and statsmodels.
* Trained with 5000 small instances solved for an hour with CPLEX.
  * a 70/30 split was done in training/test.
  * solutions far from optimum were not used.
* Validated (semi) visually (not sure the good performance metric for predicting upper bounds).

--

I tested the following techniques:

* Linear regressions.
* Decision Tree regressions.
* Neural networks regressions.
* **Gradient Boosting regressions**.
* **Quantile regressions**.

--

The only two with support for quantile regression (i.e. bounds at quantiles 10% and 90%) were the last two. And the last one had the additional benefit of returning regression coefficients. So I chose that one.

---

### Model

.pull-left[
* **Technique**: *Quantile regressions* to estimate upper and lower bounds.
* **Input features**: 
  * mean flight demand per period (up to power 4), 
  * total remaining flight hours at start (init), 
  * variance of flight demand, 
  * demand of special missions, 
  * number of period where flight demand is cut in two.
* **Output features**: *mean distance between maintenances*.
]

--

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('QuantReg_mean_consum_upper_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

---

## Current results

---

### Experiments

* Number of instances: small (5000), medium (1000) and large (1000).
* Time limit at 3600 seconds.
* We seeded instance generation for better comparison.
* CPLEX running 1 thread.

Large instances have:

* 45 aircraft
* 90 months
* +20 missions (3 active missions at any given time).

---

### How good is it (performance)

Better solutions, more solutions.

.pull-left[
For instances were an integer solution was found:

* in ~85% of instances we outperformed base model.
* ~105% average gain in objective function.
* ~18% less time in average to solve.
]

--

.pull-right[
Many improvements in feasible solutions:

|indicator         | base| cuts| %|
|:-----------------|----:|----:|--------:|
|Infeasible        |   55|   71|    29.09|
|IntegerFeasible   |  183|  469|   156.28|
|IntegerInfeasible |  665|  291|   -56.24|
|Optimal           |   92|  166|    80.43|
|total_instances   |  995|  997|     0.20|
]
---

### How good is it (optimality)

For instances were an optimal solution was found (optimum degradation):
* 95% of instances had less than 4% gap with real optimal.

.center[
```{r, echo=FALSE, out.width='100%'}
imgs = c('relative_gap_optimality.png')
knitr::include_graphics(imgs)
```
*some outliers missing at right side...
]

---

## About my ambition.

---

### Present steps vs ideal steps

.pull-left[
What I have now:
```{r, out.width='100%'}
library(DiagrammeR)
mermaid("
graph TB
    i2==>|ML|s3
    subgraph Input data
    i1[raw input]-->|manual|i2[input features]
    end
    subgraph Solution
    s1[raw solution]-->s2[good patterns]
    s2-->|manual|s3[pattern features]
    s3==>|selection|s2
    end")
```
]

--

.pull-right[
What I want:
```{r, out.width='100%'}
mermaid("
graph TB
    i1==>|ML|s2
    subgraph Input data
    i1[raw input]-->|manual|i2[input features]
    end
    subgraph Solution
    s1[raw solution]-->s2[good patterns]
    s2-->|manual|s3[pattern features]
    end")
```
]

---

## Some questions

I want to be able to match assignment patterns (e.g. maintenances schedules) and unstructured input data in order to generalize a decomposition approach.

For that I need to know:

(from least ambitious to most ambitious):

* How do I train the other methods to predict bounds when they only measure mean squares (or some other function that returns the 'expected')?
* Is it possible to match input data to distance between maintenances without the need to manually generate the input features?
  * Assuming the amount of input data can, of course, change depending on the instance. I'm thinking in image recognition where images are all different size.
  * If so, how do I transform my input in such a way that can be read by some ML technique without having to obtain features? (e.g. do I need to create an image of it?)
* Is it possible to match input data to a set of patterns so that I don't even need to even tell the prediction model what characteristic of the patterns I want.

### 2019-09-10 R/scripts/resultsNPS.Rmd
added presentation to DISC, many changes in R reports and graphs, some code corrections on batch instances in python

@@
 knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
-library(reticulate)
-library(tidyverse)
-library(knitr)
-library(magrittr)
-library(ggplot2)
-use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)
-current_dir = getwd()
-# print(current_dir)
-py_discover_config()
+source('functions/graphs_NPS.R')
+
 quant_max <- 0.95
 size <- params$size
-aux_compare <- function(table){
-    table %>% 
-    gather(key='indicator', value="value", -experiment) %>% 
-    spread(experiment, value) %>% 
-    mutate(dif_abs = cuts - base,
-           dif_perc = (dif_abs/ base*100) %>% round(2))
 
-}
-
-```
-
-```{python}
-import pandas as pd
-import numpy as np
-
-import sys, os
-opt_path = './'
-# opt_path = '../'
-rel_path_python = '{}../python/'.format(opt_path)
-path = os.path.join(r.current_dir, rel_path_python)
-# print(path)
-sys.path.insert(1, path)
-
-import package.batch as ba
-import package.params as params
-
-import stochastic.params as sto_params
-import scripts.compare_stochastic as compare_sto
-
-# This is read now as a paramer in the YAML (above).
-# size = 'small'  # choose small, medium or large
-# df = compare_sto.get_instances[r.size]()
-
-################# manual alternative...
-# exp_list = ['IT000125_20190808', 'IT000125_20190729']
-# df = compare_sto.get_df_comparison(exp_list)
-# df = df[df.scenario == 'numparalleltasks_2']
-
-exp_list = ['IT000125_20190808', 'IT000125_20190812']
-df = compare_sto.get_df_comparison(exp_list).reset_index()
-# df = df[df.scenario == 'numparalleltasks_3']
-df = df.query("scenario == 'numparalleltasks_3' or experiment==1")
-################# manual alternative...
-
-name_change = dict(experiment=[0, 1], experiment2=["cuts", "base"])
-name_change = pd.DataFrame(name_change)
-df = df.merge(name_change, on='experiment').drop('experiment', axis=1).rename(columns=dict(experiment2='experiment'))
+# alternative 1:
+# exp_list <- c('IT000125_20190808', 'IT000125_20190812')
+# df_original <- compare_sto$get_df_comparison(exp_list)
+# df_original %<>% filter(scenario=='numparalleltasks_3' | experiment==1)
+# alternative 2:
+# compare_sto <- get_compare_sto()
+# df_original <- compare_sto$get_instances[[size]]()
+# df <- df_original %>% mutate(experiment=if_else(experiment==0, 'cuts', 'base'))
+# alternative 3:
+df <- do.call(params$size, args=list())
+# df <- get_4_tasks_aggressive()
 
@@
 
-summary_stats <- 
-    py$df %>% 
-    mutate(sol_code = if_else(is.na(sol_code), 2, sol_code)) %>% 
-    group_by(experiment) %>%
-    summarise(Infeasible = sum(sol_code==-1),
-              IntegerFeasible=sum(sol_code==2),
-              IntegerInfeasible=sum(sol_code==0),
-              Optimal=sum(sol_code==1),
-              total_instances = n()) %>% 
-  aux_compare
-
+summary_stats <- get_summary(df)
 summary_stats %>% kable
@@
 
-t1 <- 
-    py$df %>% 
-    filter(sol_code==1) %>% 
-    group_by(instance) %>%
-    mutate(min_value = min(best_solution),
-           dist_min = best_solution - min_value,
-           dist_min_perc = (best_solution - min_value)/abs(min_value)*100
-           )
-t1_rel <- 
-    t1 %>% 
-    select(instance, experiment, dist_min_perc) %>% 
-    spread(key=experiment, value=dist_min_perc) %>% 
-    drop_na() %>% use_series(cuts)
-
-t1_abs <- 
-    t1 %>% 
-    select(instance, experiment, dist_min) %>% 
-    spread(key=experiment, value=dist_min) %>% 
-    drop_na() %>% use_series(cuts)
+t1 <- get_quality_degr(df)
+t1_rel <- t1 %>% filter(experiment=='cuts') %>% use_series(dist_min_perc)
+t1_abs <- t1 %>% filter(experiment=='cuts') %>% use_series(dist_min)
 
@@
 ```{r}
-qplot(t1_rel[t1_rel<20], xlab='Relative gap (in %) among optimal solutions.')
+qplot(t1_rel_filt, xlab='Relative gap (in %) among optimal solutions.')
 ```
@@
 
-quality_perf <-
-    py$df %>% 
-    filter(sol_code>=1) %>%
-    select(instance, experiment, best_solution) %>% 
-    spread(experiment, best_solution) %>% 
-    drop_na() %>% 
-    mutate(dif_perc = ((cuts-base)/ base*100) %>% round(2))
-
+quality_perf <- get_quality_perf(df)
 mean_dif <- quality_perf$dif_perc %>%  mean
@@
 ```{r}
-filtered_q_perf <- quality_perf %>% filter(between(dif_perc, -5, 5))
+filtered_q_perf <- quality_perf %>% filter(between(dif_perc, -10, 10))
 filtered_quant <- (1 - (filtered_q_perf %>% nrow)/ (quality_perf %>% nrow))*100
@@
 
-comparison_table <- 
-    py$df %>% 
-    filter(sol_code>=1) %>% 
-    group_by(instance) %>%
-    filter(n()==2) %>% 
-    ungroup() %>% 
-    select(instance, experiment, time) %>% 
-    spread(experiment, time) %>% 
-    arrange(base) %>% 
-    mutate(instance = row_number()) %>% 
-    gather(key = 'experiment',  value='time', -instance)
+comparison_table <- get_time_perf_integer(df)
 
-comparison_table_reorder <- 
-    comparison_table %>% 
-    group_by(experiment) %>% 
-    arrange(experiment, time) %>% 
-    mutate(percentage = row_number()/n()*100)
+comparison_table_reorder <- get_time_perf_integer_reorder(df)
 
@@
 
-The following graph shows the solving times in seconds for each of the `r t1 %>% distinct(instance) %>% nrow` instances were an integer solution was found in both cases.
+The following graph shows the solving times in seconds for each of the `r comparison_table %>% distinct(instance) %>% nrow` instances were an integer solution was found in both cases.
 In red are the times for the base model.
@@
 
-comparison_table <- 
-    py$df %>% 
-    filter(sol_code==1) %>% 
-    group_by(instance) %>% 
-    filter(n()==2) %>% 
-    select(experiment, instance, time) %>%
-    group_by(experiment) %>% 
-    summarise(time_mean = mean(time), 
-              time_medi = median(time)) %>% 
-    aux_compare
+comparison_table <- get_time_perf_optim(df)
 
@@
 ```{r infeasible_list}
-
-infeasible_instances <- py$df %>% filter(sol_code==-1) %>% distinct(instance)
+infeasible_instances <- get_infeasible_instances(df)
+infeasible_stats <- get_infeasible_stats(df)
 ```
@@
 ```{r infeasible}
-infeasible_stats <- 
-    py$df %>% 
-    inner_join(infeasible_instances) %>% 
-    filter(experiment=="base") %>% 
-    summarise(Infeasible = sum(sol_code==-1),
-              IntegerFeasible=sum(sol_code==2),
-              IntegerInfeasible=sum(sol_code==0),
-              Optimal=sum(sol_code==1),
-              total_instances = n()) %>% 
-    gather(key='status', 'number') %>% 
-  filter(number>0)
-
 infeasible_stats %>% kable
@@
 ```{r soft_constraints}
-errors_stats <- 
-    py$df %>% 
-    filter(sol_code>=1) %>% 
-    mutate(errors = replace_na(errors, 0)) %>% 
-    group_by(experiment) %>% 
-    summarise(errors_mean = mean(errors),
-              errors_q95 = quantile(errors, quant_max),
-              errors_medi = median(errors)) %>% 
-    aux_compare
+errors_stats <- get_soft_constraints(df, quant_max)
 

### 2019-09-10 R/scripts/export_results_NPS.R

@@ -1,0 +1,72 @@
+source('functions/graphs_NPS.R')
+
+path_export_img <- '../../NPS2019/img/'
+path_export_tab <- '../../NPS2019/tab/'
+# size <- 'small'
+# df_original <- compare_sto$get_instances[[size]]()
+
+exp_list <- c('IT000125_20190808', 'IT000125_20190812')
+exp_list <- c('IT000125_20190725', 'IT000125_20190716') # small
+df_original <- compare_sto$get_df_comparison(exp_list)
+df <- df_original %>% mutate(experiment=if_else(experiment==0, 'cuts', 'base'))
+
+# summary
+summary_stats <- get_summary(df)
+path <- '%sstats_%s_%s.tex' %>% sprintf(path_export_tab, exp_list[1], exp_list[2])
+summary_stats %>% 
+    rename(Status=Indicator) %>% 
+    kable(format='latex', booktabs = TRUE, linesep="") %>% 
+    write_file(path)
+
+# summary interception:
+
+summary_int <- get_comparable_sets(df)
+path <- '%sstats_int_%s_%s.tex' %>% sprintf(path_export_tab, exp_list[1], exp_list[2])
+summary_int %>%
+    rename(Status=Indicator) %>% 
+    kable(format='latex', booktabs = TRUE, linesep="", escape=FALSE) %>% 
+    write_file(path)
+
+
+# quality degradation
+t1 <- get_quality_degr(df)
+t1_rel <- t1 %>% filter(experiment=='cuts') %>% use_series(dist_min_perc)
+t1_rel_filt <- t1_rel[t1_rel<10]
+filtered_quant <- (1 - (t1_rel_filt %>% length) / (t1_rel %>% length))*100
+length(t1_rel)
+path <- '%squality_degradation_%s_%s.png' %>% sprintf(path_export_img, exp_list[1], exp_list[2])
+qplot(t1_rel_filt, xlab='Relative gap (in %) among optimal solutions.', binwidth=0.3) + ggsave(path)
+
+# quality performance
+quality_perf <- get_quality_perf(df)
+filtered_q_perf <- quality_perf %>% filter(between(dif_perc, -7, 7))
+filtered_quant <- (1 - (filtered_q_perf %>% nrow)/ (quality_perf %>% nrow))*100
+path <- '%squality_performance_%s_%s.png' %>% sprintf(path_export_img, exp_list[1], exp_list[2])
[…]

### 2019-09-10 python/package/batch.py

@@ -70,11 +70,18 @@ class Batch(object):
         if self.logs is not None:
             return self.logs
 
+        opt_info = self.get_options()
+        try:
+            el = list(opt_info.keys())[0]
+            solver = opt_info[el]['solver']
+        except:
+            solver = 'CPLEX'
+
         self.logs = \
             self.get_instances_paths(). \
             vapply(lambda v: os.path.join(v, 'results.log')).\
             clean(func=os.path.exists). \
-            vapply(lambda v: ol.get_info_solver(v, 'CPLEX', get_progress=get_progress))
+            vapply(lambda v: ol.get_info_solver(v, solver, get_progress=get_progress))
         return self.logs
 
     def get_json(self, name):
@@ -246,6 +253,14 @@ class ZipBatch(Batch):
 
         zipobj = zipfile.ZipFile(self.path)
 
+        opt_info = self.get_options()
+        try:
+            el = list(opt_info.keys())[0]
+            solver = opt_info[el]['solver']
+        except:
+            solver = 'CPLEX'
+
+
         def _read_zip(x):
             try:
                 return zipobj.read(x)
@@ -258,7 +273,7 @@ class ZipBatch(Batch):
             vapply(_read_zip). \
             clean(). \
             vapply(lambda x: str(x, 'utf-8')). \
-            vapply(lambda v: ol.get_info_solver(v, 'CPLEX', get_progress=get_progress, content=True))
+            vapply(lambda v: ol.get_info_solver(v, solver, get_progress=get_progress, content=True))
         return self.logs
 
     def get_json(self, name):

### 2019-09-10 python/scripts/compare_stochastic.py

@@ -29,7 +29,7 @@ def get_df_comparison(exp_list):
         reset_index().\
         drop(['level_1'], axis=1).\
         rename(columns=dict(level_0='experiment')).\
-        set_index(['instance','experiment'])
+        set_index(['instance','experiment']).reset_index()
 
 
 def compare_perc_0_1(table):

### 2019-10-10 R/scripts/resultsNPS.Rmd

@@
 params:
-  size: medium
+  size: get_3_tasks_aggresive_perc_add
 editor_options:
@@
 source('functions/graphs_NPS.R')
-
+source('functions/datasets_NPS.R')
 quant_max <- 0.95
@@
 # alternative 3:
-df <- do.call(params$size, args=list())
+df <- do.call(size, args=list())
 # df <- get_4_tasks_aggressive()
@@
 per_cut <- sum(quality_perf$dif_perc<0)/length(quality_perf$dif_perc)
+q95_cut <- quality_perf$dif_perc %>% quantile(0.95)
 
@@
 In addition, in `r (per_cut*100) %>% round(2)`\% of instances (`r (quality_perf_nrow*per_cut) %>% floor` of `r quality_perf_nrow`) the `cuts` model outperform the solution from the base model.
+95\% of instances are less than `r q95_cut %>% round(2)`\% away from the best known solution.
 
@@
     comparison_table %>% 
-    group_by(experiment) %>% 
+    group_by(scenario, experiment) %>% 
     summarise(time_mean = mean(time), 

### 2019-10-11 R/presentations/DISC_201909/presentation.Rmd

@@
 
-* Number of instances: small (5000), medium (1000) and large (1000).
+* Number of instances: small (5000), medium (1000) and large (1000) and very large (1000).
 * Time limit at 3600 seconds.
@@
 
-Large instances have:
+Largest instances have:
 
-* 45 aircraft
+* 60 aircraft
 * 90 months
-* +20 missions (3 active missions at any given time).
+* ~30 missions (4 active missions at any given time).
 
@@
 
-Better solutions, more solutions.
+Faset solutions, more solutions. (Large instances aggresive cuts).
 
@@
 
-* in ~85% of instances we outperformed base model.
-* ~105% average gain in objective function.
+* +75% worse solutions (caveat).
+* most solutions are between 0-5% of best known.
 * ~18% less time in average to solve.
+* ~53% less time to "optimal".
 ]

### 2019-10-14 R/nps_reports/results.Rmd
refactored the NPS r scripts into a single directory with simpler names.

--- 2019-10-10 R/scripts/resultsNPS.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
-source('functions/graphs_NPS.R')
-source('functions/datasets_NPS.R')
+source('nps_reports/graphs_NPS.R')
+source('nps_reports/datasets_NPS.R')
 quant_max <- 0.95

### 2019-10-14 R/nps_reports/datasets.R

@@ -1,6 +1,6 @@
 library(tidyverse)
 library(reticulate)
-source('functions/graphs_NPS.R')
+source('nps_reports/graphs.R')
 
 get_python_module <- function(rel_path, name){
     use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)

### 2019-10-14 R/nps_reports/export_results.R

@@ -1,5 +1,5 @@
-source('functions/graphs_NPS.R')
-source('functions/datasets_NPS.R')
+source('nps_reports/graphs.R')
+source('nps_reports/datasets.R')
 library(stringr)
 
 path_export_img <- '../../NPS2019/img/'

### 2019-10-14 R/nps_reports/functions.R

@@ -112,7 +112,7 @@ get_time_perf_integer_reorder <- function(raw_df){
         mutate(percentage = row_number()/n()*100)
 }
 
-# TODO: aux_compare before making mean??
+# when comparing times we compare averages, not average relative differences.
 get_time_perf_optim <- function(raw_df){
     raw_df %>% 
         get_all_optimal %>% 
@@ -158,16 +158,20 @@ get_soft_constraints <- function(raw_df, quant_max, compare=TRUE){
 }
 
 get_infeasible_times <- function(raw_df){
-    # aux_compare %>% 
     raw_df %>% 
         get_all_infeasible %>% 
-        inner_join(raw_df) %>%
-        select(experiment, time) %>% 
-        gather(key='Indicator', value="value", -experiment) %>% 
-        spread(experiment, value) %>% 
-        mutate(dif_abs = cuts - base,
-               dif_perc = dif_abs/base*100
-        )
+        group_by(scenario, experiment) %>% 
+        summarise(time_mean = mean(time), 
+                  time_medi = median(time)) %>% 
+        aux_compare
+}
+
+get_variances <- function(raw_df){
+    raw_df %>% 
+        get_all_integer %>% 
+        select(instance, experiment, best_solution) %>% 
+        spread(experiment, best_solution) %>% 
+        mutate(dif_perc = ((cuts-base)/ abs(base)) %>% times_100_round)
 }
 
 get_mega_summary <- function(df){
@@ -176,6 +180,8 @@ get_mega_summary <- function(df){
     # feasability:
     # extra infeasible instances as percent of total
     # extra soft constraints violations (avg, 95%)
+    # TODO: time to detect infeasible.
+    # TODO: sum of variances.
     # performance:
     # extra feasible instances as percent of total
     # time to solve: median, avg

### 2019-10-14 R/nps_reports/results.Rmd
changes in python to get additional data from batch to compare. and some changes in R.

--- 2019-10-10 R/scripts/resultsNPS.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
-source('functions/graphs_NPS.R')
-source('functions/datasets_NPS.R')
+source('functions.R')
+source('datasets.R')
 quant_max <- 0.95
@@
 infeasible_stats <- get_infeasible_stats(df)
+infeasible_times <- get_infeasible_times(df)
 ```
@@
 
-```{r infeasible}
+```{r infeasible_stats}
 infeasible_stats %>% kable
+```
+
+In addition, the times to detect the infeaseability of a given instance have been greatly reduced (to half, approximately) as can be see in the following table:
+
+```{r infeasible_times}
+infeasible_times %>% kable
 ```
@@
 ```
+
+### Reducing variance
+
+The variance in the frequency of maintenances is a side objective that was not been taken into account previously.
+To calculate it, we have, for each fleet type, measured the variance of the distance between maintenances.
+And we have summed all variances for all fleet types.
+We do this analysis for all integer solutions obtained.
+As can be seen, in almost all of the cases there is a reduction in the variance. The mean relative difference is `r mean_var_reduction` and 95\% of instances have less than `r quant95_var_reduction`.
+
+```{r variances}
+variances <- get_variances(df)
+var_per <- variances$dif_perc
+mean_var_reduction <- var_per %>% mean
+quant95_var_reduction <- var_per %>% quantile(quant_max)
+qplot(var_per, xlab='Relative difference in variance (in % of base case) among integer solutions.')
+```

### 2019-10-14 R/nps_reports/datasets.R

@@ -1,6 +1,6 @@
 library(tidyverse)
 library(reticulate)
-source('nps_reports/graphs.R')
+source('nps_reports/functions.R')
 
 get_python_module <- function(rel_path, name){
     use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)

### 2019-10-14 R/nps_reports/export_results.R

@@ -1,4 +1,4 @@
-source('nps_reports/graphs.R')
+source('nps_reports/functions.R')
 source('nps_reports/datasets.R')
 library(stringr)
 

### 2019-10-14 python/package/batch.py

@@ -230,8 +230,8 @@ class ZipBatch(Batch):
         keys = \
             scenario_instances.\
             apply(str.split, '/').\
-                apply(tuple).\
-                filter(keys_positions)
+            apply(tuple).\
+            filter(keys_positions)
         return sd.SuperDict(zip(keys, scenario_instances))
 
     def re_make_paths(self, scenario_instances):

### 2019-10-14 python/scripts/compare_stochastic.py

@@ -5,9 +5,22 @@ import package.batch as ba
 import package.params as params
 
 import stochastic.params as sto_params
-
+import stochastic.solution_stats as sol_stats
 #####################
 # INFO
+
+
+def get_sum_variances(case, types):
+    return sum(sol_stats.get_variance_dist(case, t) for t in types)
+
+
+def get_solstats(batch):
+    cases = batch.get_cases().clean(func=lambda v: v)
+    types = cases.vapply(lambda v: v.instance.get_types())
+    variances = cases.apply(lambda k, v: {'variance': get_sum_variances(v, types[k])})
+    var_table = batch.format_df(variances).drop('name', axis=1)
+    return var_table
+
 
 def get_df_comparison(exp_list):
     raw_results_dict = {}
@@ -18,8 +31,11 @@ def get_df_comparison(exp_list):
         batch1 = ba.ZipBatch(path=params.PATHS['results'] + experiment)
         table = batch1.get_log_df()
         table_errors = batch1.get_errors_df().drop('name', axis=1)
+        sol_stats_table = get_solstats(batch1)
         table_n = \
-            table.merge(table_errors, on=['scenario', 'instance'], how='left').\
+            table.\
+            merge(table_errors, on=['scenario', 'instance'], how='left').\
+            merge(sol_stats_table, on=['scenario', 'instance'], how='left').\
             sort_values(['scenario', 'instance']).\
             reset_index(drop=True)
         raw_results_dict[name] = table_n
@@ -31,6 +47,7 @@ def get_df_comparison(exp_list):
         rename(columns=dict(level_0='experiment')).\
         set_index(['instance','experiment']).reset_index()
 
+# The following is obsolete because we now do the comparison in R:
 
 def compare_perc_0_1(table):
     return (table[0] - table[1])/table[1] *100
@@ -59,9 +76,6 @@ def get_small_instances():
     df = get_df_comparison(exp_list)
     return df
 
-get_instances = dict(small= get_small_instances,
-                     medium= get_medium_instances,
-                     large= get_large_instances)
 
 if __name__ == '__main__':
 
@@ -75,6 +89,10 @@ if __name__ == '__main__':
     # df = get_small_instances()
     # df = get_medium_instances()
     df = get_medium_instances()
+    weird = df[~np.isnan(df.best_solution) & np.isnan(df.variance)]
+    # import package.experiment as exp
+    #
[…]

### 2019-10-14 python/stochastic/solution_stats.py

@@ -1,5 +1,6 @@
 import pandas as pd
 import stochastic.instance_stats as istats
+import numpy as np
 
 
 def get_num_maints(case, _type=0):
@@ -56,6 +57,8 @@ def get_1M_2M_dist(case, _type=0, count_1M=False):
     duration = case.instance.get_param('maint_duration')
     shift = case.instance.shift_period
     limit = shift(first, duration)
+    # if an aircraft starts a maintenance in the first period;
+    # we artificially add a shadow cycle before it
     cycles_edit = cycles.clean(func=lambda v: v[0][0] == limit).vapply(lambda v: [()]+v)
     cycles.update(cycles_edit)
 
@@ -100,3 +103,8 @@ def get_post_2M_dist(case, _type=0):
         vapply(lambda v: dist(v[1], last))
 
     return last_dist
+
+
+def get_variance_dist(case, _type=0):
+    distances = get_1M_2M_dist(case, _type=0, count_1M=True)
+    return np.var(distances)

### 2019-10-14 R/scripts/resultsNPS.Rmd
cleaning comments and troublesome libraries to install in windows.

@@
 size <- params$size
-
-# alternative 1:
-# exp_list <- c('IT000125_20190808', 'IT000125_20190812')
-# df_original <- compare_sto$get_df_comparison(exp_list)
-# df_original %<>% filter(scenario=='numparalleltasks_3' | experiment==1)
-# alternative 2:
-# compare_sto <- get_compare_sto()
-# df_original <- compare_sto$get_instances[[size]]()
-# df <- df_original %>% mutate(experiment=if_else(experiment==0, 'cuts', 'base'))
-# alternative 3:
 df <- do.call(size, args=list())
-# df <- get_4_tasks_aggressive()
 

### 2019-10-14 R/nps_reports/results.Rmd
corrected report generation in python. Added code to graph variances in R.

@@
 knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
-source('functions.R')
-source('datasets.R')
+# setwd('../')
+# browser()
+source('nps_reports/functions.R')
+source('nps_reports/datasets.R')
 quant_max <- 0.95
@@
 
-### Reducing variance
-
-The variance in the frequency of maintenances is a side objective that was not been taken into account previously.
-To calculate it, we have, for each fleet type, measured the variance of the distance between maintenances.
-And we have summed all variances for all fleet types.
-We do this analysis for all integer solutions obtained.
-As can be seen, in almost all of the cases there is a reduction in the variance. The mean relative difference is `r mean_var_reduction` and 95\% of instances have less than `r quant95_var_reduction`.
-
 ```{r variances}
@@
 quant95_var_reduction <- var_per %>% quantile(quant_max)
+```
+
+### Making aircraft usage more homogeneous
+
+The variance in the frequency of maintenances is a side objective that was not been taken into account previously.
+The higher the variance, the more difference there is between amount of flight hours between one aircraft and another.
+To calculate it, we have, for each fleet type, measured the variance of the distance between maintenances.
+And we have summed all variances for all fleet types.
+We do this analysis for all integer solutions obtained.
+As can be seen, in almost all of the cases there is a reduction in the variance.
+The mean relative difference is `r mean_var_reduction %>% round(2)`\%  and `r quant_max*100`\% of instances have less than `r quant95_var_reduction`\% extra variance.
+
+```{r graph_variances}
+
 qplot(var_per, xlab='Relative difference in variance (in % of base case) among integer solutions.')

### 2019-10-14 R/nps_reports/datasets.R

@@ -4,6 +4,7 @@ source('nps_reports/functions.R')
 
 get_python_module <- function(rel_path, name){
     use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)
+    # use_condaenv('cvenv', conda='c:/Anaconda3/Scripts/conda.exe', required=TRUE)
     py_discover_config()
     sysp = import('sys')
     opt_path = ''

### 2019-10-14 R/nps_reports/export_results.R

@@ -73,6 +73,7 @@ ggplot(data=comparison_table_reorder, aes(x=percentage, y=time, color=experiment
 
 # infeasible
 infeasible_stats <- get_infeasible_stats(df)
+# infeasible_times <- get_infeasible_times(df)
 
 path <- '%sinfeas_%s_%s.tex' %>% sprintf(path_export_tab, exp_list[1], exp_list[2])
 infeasible_stats %>% 
@@ -119,7 +120,12 @@ comparison_table %>%
     summarise(time_mean = mean(time), 
               time_medi = median(time))
 
-
+# variance
+variances <- get_variances(df)
+var_per <- variances$dif_perc
+path <- '%svariance_%s_%s.png' %>% sprintf(path_export_img, exp_list[1], exp_list[2])
+qplot(var_per, xlab='Relative difference in variance among solutions.')+
+    theme(text = element_text(size=20)) + ggsave(path)
 
 # prediction models -------------------------------------------------------
 
@@ -141,7 +147,7 @@ ggplot(data=result_tab_n, aes(y=maints, x=mean_consum)) +
     ggsave(path)
 
 # quantiles
-
+# TODO: quantiles
 
 
 

### 2019-10-14 python/stochastic/solution_stats.py

@@ -64,17 +64,20 @@ def get_1M_2M_dist(case, _type=0, count_1M=False):
 
     dist = case.instance.get_dist_periods
     max_value = case.instance.get_param('max_elapsed_time')
+    size_horizon = dist(first, last) + 1
 
     if count_1M:
         # we count the second cycle.
+        # sometimes an aircraft has no maintenances done.
+        # (I thought this was impossible...)
         cycles_between = \
-            cycles.apply(lambda k, v: dist(*v[1]) + 1)
+            cycles.vapply(lambda v: dist(*v[1]) + 1 if len(v)>1 else size_horizon)
     else:
         # we only want to see the distance when
         # there is a second maintenance
         cycles_between = \
             cycles.clean(func=lambda v: len(v)==3).\
-            apply(lambda k, v: dist(*v[1]) + 1)
+            vapply(lambda v: dist(*v[1]) + 1 if len(v)>1 else size_horizon)
 
     if not len(cycles_between):
         return pd.Series(max_value)

### 2019-10-25 R/nps_reports/results.Rmd
changes in R and python to get correct logs from fixLP matheuristic.

@@
 ```{r}
-t1_rel_filt <- t1_rel[t1_rel<20]
+# this takes 10% from right side
+t1_rel_filt <- t1_rel[t1_rel<quantile(t1_rel, 0.90)]
 filtered_quant <- (1 - (t1_rel_filt %>% length) / (t1_rel %>% length))*100
@@
 ```{r}
-filtered_q_perf <- quality_perf %>% filter(between(dif_perc, -10, 10))
+
+# this takes 5% from each side
+filtered_q_perf <- quality_perf %>% filter(dif_perc %>% between(., quantile(., c(0.05)), quantile(., c(0.95))))
 filtered_quant <- (1 - (filtered_q_perf %>% nrow)/ (quality_perf %>% nrow))*100

### 2019-10-25 R/nps_reports/datasets.R

@@ -3,8 +3,9 @@ library(reticulate)
 source('nps_reports/functions.R')
 
 get_python_module <- function(rel_path, name){
-    use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)
-    # use_condaenv('cvenv', conda='c:/Anaconda3/Scripts/conda.exe', required=TRUE)
+    # use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)
+    use_condaenv('cvenv', conda='c:/Anaconda3/Scripts/conda.exe', required=TRUE)
+    # c:\Anaconda3\envs\cvenv\Scripts\pip.exe install -r requirements
     py_discover_config()
     sysp = import('sys')
     opt_path = ''
@@ -26,16 +27,28 @@ get_result_tab <- function(name){
     result_tab
 }
 
-get_quantile_prev <- function(){
-    # TODO
+get_generic_compare <- function(dataset_list, scenario_filter=NULL, exp_names=NULL){
+    compare_sto <- get_compare_sto()
+    if (exp_names %>% is.null){
+        # This assumes the dataset_list has only two datasets!
+        # (which is the most likely option)
+        exp_names <- list('cuts', 'base')
+    }
+    df_original <- compare_sto$get_df_comparison(dataset_list)
+    dataset_names <- dataset_list %>% lapply(function(x) x)
+    df_original %<>% 
+        mutate(dataset=dataset_names[experiment+1] %>% unlist,
+               experiment=exp_names[experiment+1] %>% unlist)
+    if (scenario_filter %>% is.null){
+        return(df_original)
+    }
+    tab_filter <- data.frame(scenario=scenario_filter, stringsAsFactors =FALSE)
+    df_original %>% semi_join(tab_filter)
 }
 
 get_1_tasks <- function(){
-    compare_sto <- get_compare_sto()
-    exp_list <- c('IT000125_20190725', 'IT000125_20190716')
-    df_original <- compare_sto$get_df_comparison(exp_list)
-    df_original %>% 
-        mutate(experiment=if_else(experiment==0, 'cuts', 'base'))
+    get_generic_compare(c('IT000125_20190725', 'IT000125_20190716'), 
+                        exp_names = list('cuts', 'base'))
 }
 
 get_3_tasks <- function(){
@@ -244,4 +257,10 @@ get_all_tasks_very_aggresive_percadd <- function(){
     df_original %>% 
[…]

### 2019-10-25 python/package/batch.py

@@ -66,16 +66,24 @@ class Batch(object):
         self.cases = self.get_instances_paths().vapply(load_data)
         return self.cases
 
+    def get_solver(self):
+        opt_info = self.get_options()
+        try:
+            el = list(opt_info.keys())[0]
+            engine = opt_info[el]['solver']
+            if '.' in engine:
+                engine, solver = engine.split('.')
+            else:
+                solver = engine
+        except:
+            solver = 'CPLEX'
+        return solver
+
     def get_logs(self, get_progress=False):
         if self.logs is not None:
             return self.logs
 
-        opt_info = self.get_options()
-        try:
-            el = list(opt_info.keys())[0]
-            solver = opt_info[el]['solver']
-        except:
-            solver = 'CPLEX'
+        solver = self.get_solver()
 
         self.logs = \
             self.get_instances_paths(). \
@@ -251,14 +259,7 @@ class ZipBatch(Batch):
             return self.logs
 
         zipobj = zipfile.ZipFile(self.path)
-
-        opt_info = self.get_options()
-        try:
-            el = list(opt_info.keys())[0]
-            solver = opt_info[el]['solver']
-        except:
-            solver = 'CPLEX'
-
+        solver = self.get_solver()
 
         def _read_zip(x):
             try:

### 2019-10-30 R/nps_reports/results.Rmd
added datasets and improved html reports graphs

@@
 ```{r time_perf_graph} 
-ggplot(data=comparison_table, aes(x=instance, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5)
+ggplot(data=comparison_table, aes(x=instance, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5)+ theme(text = element_text(size=20))
 ```
@@
 ```{r time_perf_graph2} 
-ggplot(data=comparison_table_reorder, aes(x=percentage, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5) + ggplot2::xlab('instance percentage')
+ggplot(data=comparison_table_reorder, aes(x=percentage, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5) + ggplot2::xlab('instance percentage') + theme(text = element_text(size=20))
 ```

### 2019-11-04 R/nps_reports/results.Rmd
prepared R reports to compare multiple experiments correctly. Boxplots.

--- 2019-10-14 R/nps_reports/results.Rmd
@@
 df <- do.call(size, args=list())
+value_filt_tails <- function(value, each_tail) value %>% between(., quantile(., c(each_tail[1])), quantile(., c(1-each_tail[2])))
 
@@
 
-We compare the times where both models return an **optimal solution**.
-
 ```{r quality_degr}
 
-t1 <- get_quality_degr(df)
-t1_rel <- t1 %>% filter(experiment=='cuts') %>% use_series(dist_min_perc)
-t1_abs <- t1 %>% filter(experiment=='cuts') %>% use_series(dist_min)
+quality_degr_all <- get_quality_degr_2(df)
 
+t1_rel <- quality_degr_all %>% filter(experiment=='cuts') %>% use_series(dif_perc)
+t1_abs <- quality_degr_all %>% filter(experiment=='cuts') %>% use_series(dif)
 quality_degr_quant <- t1_rel %>% quantile()
 quality_degr_quant95 <- t1_rel %>% quantile(quant_max)
-quality_degr_nrow <- t1_rel %>% length()
 quality_degr_quant95_abs <- t1_abs %>% quantile(quant_max)
 
+right_tail <- 0.10
+quality_degr_all_filt <- 
+  quality_degr_all %>% 
+  group_by(experiment) %>% 
+  filter(dif_perc %>% value_filt_tails(c(0, right_tail)))
+
 ```
+We compare the times where all models return an **optimal solution** (`r quality_degr_all %>% distinct(instance) %>% nrow` instances).
 
-For `r quant_max*100`\% of the `r quality_degr_nrow` (`r (quant_max*quality_degr_nrow) %>% floor`) instances that comply with this condition, the loss in the objective function was `r quality_degr_quant95 %>% round(2)`\% or less from the optimal.
+<!-- For `r quant_max*100`\% of the `r quality_degr_nrow` (`r (quant_max*quality_degr_nrow) %>% floor`) instances that comply with this condition, the loss in the objective function was `r quality_degr_quant95 %>% round(2)`\% or less from the optimal.
 This corresponds to an absolute value of `r quality_degr_quant95_abs %>% round(2)`.
 As a reference, the absolute gap tolerance used in the solving was of 10.
+-->
+
+The following boxplot shows the distribution of the gap from the models with respect to the `base` model.
 
 ```{r}
-t1_rel_filt <- t1_rel[t1_rel<20]
-filtered_quant <- (1 - (t1_rel_filt %>% length) / (t1_rel %>% length))*100
-```
+ggplot(data=quality_degr_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('% difference in objective') + theme(text = element_text(size=20)) + coord_flip()
 
-The following graph shows the distribution of the gap from the `cuts` model with the `base` model. `r filtered_quant %>% round(2)`\% of outliers were taken out from the right tail.
-
-```{r}
-qplot(t1_rel_filt, xlab='Relative gap (in %) among optimal solutions.')
 ```
@@
 
-If we compare among all cases where an **integer solution** was found in the two models, we can obtain an estimate of quality performance for the 1 hour solving time.
-
 ```{r quality_perf}
 
-quality_perf <- get_quality_perf(df)
+quality_perf_all <- get_quality_perf_2(df)
+
+quality_perf <- quality_perf_all %>% filter(experiment=='cuts')
 mean_dif <- quality_perf$dif_perc %>%  mean
@@
 
-quality_perf_nrow <- quality_perf %>% distinct(instance) %>% nrow
+```
+
+If we compare among all cases where an **integer solution** (`r quality_perf_all %>% distinct(instance) %>% nrow` instances) was found in the two models, we can obtain an estimate of quality performance for the 1 hour solving time.
+
+```{r}
+
+# this takes some % from each side
+each_tail <- 0.15
+
+quality_perf_all_filt <- 
+  quality_perf_all %>% 
+  group_by(experiment) %>% 
+  filter(dif_perc %>% value_filt_tails(c(each_tail, each_tail)))
+```
+
+The following graph shows the distribution of relative gaps between the best integer solutions found in each of the two models.
+A negative relative gap means the model outperformed the `base` model in that particular instance.
+Both extreme tails (`r (each_tail*100) %>% round(2)`\%) have been taken out to help better visualize it.
+
+```{r}
+
+ggplot(data=quality_perf_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('% difference in objective') + theme(text = element_text(size=20)) + coord_flip()
 
@@
 
-```{r}
-filtered_q_perf <- quality_perf %>% filter(between(dif_perc, -10, 10))
-filtered_quant <- (1 - (filtered_q_perf %>% nrow)/ (quality_perf %>% nrow))*100
-```
-The following graph shows the distribution of relative gaps between the best integer solutions found in each of the two models.
-A negative relative gap means the `cuts` model outperformed the `base` model in that particular instance.
-Both extreme tails (`r filtered_quant %>% round(2)`\%) have been taken out to help better visualize it.
+<!-- Of the `r quality_perf_nrow` instances were an **integer solution** was found in all model. -->
+<!-- the average relative gap between the `cuts` model and the `base` model was of `r mean_dif %>% round(2)`\%.
+This means that, in average solutions of the `cuts` model were `r (-mean_dif) %>% round(2)`\% better than the `base` model.
+-->
 
-```{r}
-qplot(filtered_q_perf$dif_perc, xlab='Relative gap (in %) among integer solutions.')
-```
-
-Of the `r quality_perf_nrow` instances were an **integer solution** was found in both models, the average relative gap between the `cuts` model and the `base` model was of `r mean_dif %>% round(2)`\%.
-This means that, in average solutions of the `cuts` model were `r (-mean_dif) %>% round(2)`\% better than the `base` model.
-
-In addition, in `r (per_cut*100) %>% round(2)`\% of instances (`r (quality_perf_nrow*per_cut) %>% floor` of `r quality_perf_nrow`) the `cuts` model outperform the solution from the base model.
+<!-- In addition, in `r (per_cut*100) %>% round(2)`\% of instances (`r (quality_perf_nrow*per_cut) %>% floor` of `r quality_perf_nrow`) the `cuts` model outperform the solution from the base model.
 95\% of instances are less than `r q95_cut %>% round(2)`\% away from the best known solution.
+-->
 
@@
 
-Just like in the mean quality comparison, we compare among cases where both models found a **integer solution**.
+Just like in the mean quality comparison, we compare among cases where all models found a **integer solution**.
 
@@
 
-The following graph shows the solving times in seconds for each of the `r comparison_table %>% distinct(instance) %>% nrow` instances were an integer solution was found in both cases.
-In red are the times for the base model.
-In green the times for the model with the added cuts.
+The following graph shows the solving times in seconds for each of the `r comparison_table %>% distinct(instance) %>% nrow` instances were an integer solution was found in all cases.
+If we reorder all models from fastest to slowest, we gain more insight on the distribution of solving times.
 
-```{r time_perf_graph} 
-ggplot(data=comparison_table, aes(x=instance, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5)
-```
-
-If we reorder both models from fastest to slowest, we lose the comparison of individual instances but we gain more insight on the distribution of solving times.
+<!-- ```{r time_perf_graph}  -->
+<!-- ggplot(data=comparison_table, aes(x=instance, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5) + theme(text = element_text(size=20)) + guides(color = guide_legend(override.aes = list(size=5))) -->
+<!-- ``` -->
 
 ```{r time_perf_graph2} 
-ggplot(data=comparison_table_reorder, aes(x=percentage, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5) + ggplot2::xlab('instance percentage')
+ggplot(data=comparison_table_reorder, aes(x=percentage, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5) + ggplot2::xlab('instance percentage') + theme(text = element_text(size=20)) + guides(color = guide_legend(override.aes = list(size=5)))
 ```
@@
 
-Here, we measure the time it took for each model to reach an **optimal solution.** We compare among instances where both models obtained and optimal solution.
+Here, we measure the time it took for each model to reach an **optimal solution.** We compare among instances where all models obtained and optimal solution.
 
@@
 
-We take all `r infeasible_instances %>% nrow` **infeasible solutions** from the cuts model and obtain statistics on their solution on the base model.
+We take all `r infeasible_instances %>% distinct(instance) %>% nrow` **infeasible solutions** from the models and obtain statistics on their solution on the base model.
 
 The below table shows the status obtained in the base model.
-The number of infeasible solutions is very close in both models and for all instances that were identified by infeasible in the `cuts` model, the `base` model was unable to find an integer solution.
+The number of infeasible solutions is very close in both models and for all instances that were identified by infeasible in the a model, the `base` model was unable to find an integer solution.
 
@@
 
-Regarding the accomplishment of soft constraints, we can analyze for mutually **integer solutions**, the quantity of violations in their solutions.
+Regarding the accomplishment of soft constraints, we can analyze for mutually **optimal solutions**, the quantity of violations in their solutions.
 
@@
 ```{r variances}
-variances <- get_variances(df)
-var_per <- variances$dif_perc
-mean_var_reduction <- var_per %>% mean
-quant95_var_reduction <- var_per %>% quantile(quant_max)
+variances_all <- get_variances(df)
+# var_per <- variances_all %>% filter(experiment=='cuts') %>% use_series(dif_perc)
+# mean_var_reduction <- var_per %>% mean
+# quant95_var_reduction <- var_per %>% quantile(quant_max)
+
+right_tail <- 0.1
+
+variances_all_filt <- 
+  variances_all %>% 
+  group_by(experiment) %>% 
+  filter(dif_perc %>% value_filt_tails(c(0, right_tail)))
+
 ```
@@
 As can be seen, in almost all of the cases there is a reduction in the variance.
-The mean relative difference is `r mean_var_reduction %>% round(2)`\%  and `r quant_max*100`\% of instances have less than `r quant95_var_reduction`\% extra variance.
 
@@
 
-qplot(var_per, xlab='Relative difference in variance (in % of base case) among integer solutions.')
+# qplot(var_per, xlab='')
+
+ggplot(data=variances_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('Difference in variance (in % of base case)') + theme(text = element_text(size=20)) + coord_flip()
+
 ```

### 2019-11-04 R/nps_reports/functions.R

@@ -15,9 +15,10 @@ filter_all_exps <- function(table){
 aux_compare <- function(raw_df){
     raw_df %>% 
         gather(key='Indicator', value="value", -experiment, -scenario) %>% 
-        spread(experiment, value) %>% 
-        mutate(dif_abs = cuts - base,
-               dif_perc = (dif_abs/ base*100) %>% round(2))
+        spread(experiment, value) 
+    # %>% 
+    #     mutate(dif_abs = cuts - base,
+    #            dif_perc = (dif_abs/ base*100) %>% round(2))
 }
 
 times_100_round <- function(value) value %>% multiply_by(100) %>% round(2)
@@ -64,23 +65,36 @@ get_all_integer <- function(raw_df){
         filter_all_exps
 }
 
+dif_perc_with_base <- function(value, base) ((value-base)/abs(base)) %>% times_100_round
+
+compare_objectives_perc <- function(df, column='best_solution'){
+    aux <- 
+        df %>% 
+        filter(experiment=='base') %>% 
+        select(instance, base=column) %>%
+        inner_join(df)
+    
+    column_value <- aux[[column]]
+    aux %>% 
+        mutate(dif_perc = dif_perc_with_base(column_value, base),
+               dif = column_value - base) %>% 
+        filter(experiment!='base')
+}
+
+get_quality_perf_2 <- function(raw_df){
+    raw_df %>% get_all_integer %>% compare_objectives_perc('best_solution')
+}
+
+get_quality_degr_2 <- function(raw_df){
+    raw_df %>% get_all_optimal %>% compare_objectives_perc('best_solution')
+}
 
 get_quality_perf <- function(raw_df){
-    raw_df %>% 
-        get_all_integer %>% 
-        select(instance, experiment, best_solution) %>% 
-        spread(experiment, best_solution) %>% 
-        mutate(dif_perc = ((cuts-base)/ abs(base)) %>% times_100_round)
+    raw_df %>% get_quality_perf_2 %>% filter(experiment=='cuts')
 }
 
 get_quality_degr <- function(raw_df){
-    raw_df %>% 
-        get_all_optimal %>% 
-        group_by(scenario, instance) %>%
-        mutate(min_value = min(best_solution),
-               dist_min = best_solution - min_value,
-               dist_min_perc = (best_solution - min_value)/abs(min_value)*100
-        )
+    raw_df %>% get_quality_degr_2 %>% filter(experiment=='cuts')
 }
 
[…]

### 2019-11-05 R/nps_reports/results.Rmd
added new, prettier and more powerful comparisons to reports.

@@
 quant_max <- 0.95
+really_small_number <- 0.000001
 size <- params$size
@@
 value_filt_tails <- function(value, each_tail) value %>% between(., quantile(., c(each_tail[1])), quantile(., c(1-each_tail[2])))
+
+scaleFUN <- function(x) sprintf("%.2f", x)
 
@@
 * `base` model without any added constraints.
-* `dif_abs` absolute distance between the two models for the shown quantity.
-* `dif_rel` relative distance between the two models for the shown quantity. Always using (cuts-base)/base.
 * `time_mean` mean time.
@@
 
-right_tail <- 0.10
+right_tail <- 0.0
 quality_degr_all_filt <- 
   quality_degr_all %>% 
-  group_by(experiment) %>% 
+  mutate(dif_perc = if_else(dif_perc<= 0, really_small_number, dif_perc)) %>%
   filter(dif_perc %>% value_filt_tails(c(0, right_tail)))
@@
 
-<!-- For `r quant_max*100`\% of the `r quality_degr_nrow` (`r (quant_max*quality_degr_nrow) %>% floor`) instances that comply with this condition, the loss in the objective function was `r quality_degr_quant95 %>% round(2)`\% or less from the optimal.
-This corresponds to an absolute value of `r quality_degr_quant95_abs %>% round(2)`.
-As a reference, the absolute gap tolerance used in the solving was of 10.
--->
-
 The following boxplot shows the distribution of the gap from the models with respect to the `base` model.
+Results are shown in logaritmic scale and so we have replace all negative values for a very small one.
 
 ```{r}
-ggplot(data=quality_degr_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('% difference in objective') + theme(text = element_text(size=20)) + coord_flip()
+ggplot(data=quality_degr_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('% difference in objective (log2 scale)') + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans='log2', labels=scaleFUN)
 
@@
 # this takes some % from each side
-each_tail <- 0.15
+each_tail <- 0
 
@@
   quality_perf_all %>% 
-  group_by(experiment) %>% 
+  mutate(dif_perc = if_else(dif_perc<= 0, really_small_number, dif_perc)) %>%
   filter(dif_perc %>% value_filt_tails(c(each_tail, each_tail)))
@@
 A negative relative gap means the model outperformed the `base` model in that particular instance.
-Both extreme tails (`r (each_tail*100) %>% round(2)`\%) have been taken out to help better visualize it.
+Results are shown in logaritmic scale and so we have replace all negative values for a very small one.
 
-```{r}
+```{r quality_perf_plot}
 
-ggplot(data=quality_perf_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('% difference in objective') + theme(text = element_text(size=20)) + coord_flip()
+ggplot(data=quality_perf_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('% difference in objective (log2 scale)') + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans='log2', labels=scaleFUN)
 
 ```
-
-<!-- Of the `r quality_perf_nrow` instances were an **integer solution** was found in all model. -->
-<!-- the average relative gap between the `cuts` model and the `base` model was of `r mean_dif %>% round(2)`\%.
-This means that, in average solutions of the `cuts` model were `r (-mean_dif) %>% round(2)`\% better than the `base` model.
--->
-
-<!-- In addition, in `r (per_cut*100) %>% round(2)`\% of instances (`r (quality_perf_nrow*per_cut) %>% floor` of `r quality_perf_nrow`) the `cuts` model outperform the solution from the base model.
-95\% of instances are less than `r q95_cut %>% round(2)`\% away from the best known solution.
--->
 
@@
 
-<!-- ```{r time_perf_graph}  -->
-<!-- ggplot(data=comparison_table, aes(x=instance, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5) + theme(text = element_text(size=20)) + guides(color = guide_legend(override.aes = list(size=5))) -->
-<!-- ``` -->
-
-```{r time_perf_graph2} 
+```{r time_perf_graph2, out.width='100%'}
 ggplot(data=comparison_table_reorder, aes(x=percentage, y=time, color=experiment)) + theme_minimal() + geom_point(size=0.5) + ggplot2::xlab('instance percentage') + theme(text = element_text(size=20)) + guides(color = guide_legend(override.aes = list(size=5)))
@@
 
-In addition, the times to detect the infeaseability of a given instance have been greatly reduced (to half, approximately) as can be see in the following table:
+In addition, the times to detect the infeaseability of a given instance can be seen in the following table:
 
@@
 Regarding the accomplishment of soft constraints, we can analyze for mutually **optimal solutions**, the quantity of violations in their solutions.
+For each model and instance, we subtract the `base` model errors.
+We show only the extra errors (in number) compared to the `base` case.
 
@@
 errors_stats <- get_soft_constraints(df, quant_max)
+errors_all <- get_soft_constraints_2(df) %>% 
+  mutate(dif= if_else(dif<=0, really_small_number, dif))
 
@@
 errors_stats %>% kable
+```
+
+And the complete distributions in logaritmic scale:
+
+```{r soft_constraints_plot}
+
+ggplot(data=errors_all, aes(x=experiment, y=dif)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('absolute difference in errors  (log2 scale)') + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans='log2', labels=scaleFUN)
+
 ```
@@
   variances_all %>% 
-  group_by(experiment) %>% 
-  filter(dif_perc %>% value_filt_tails(c(0, right_tail)))
+  mutate(dif_perc = if_else(dif_perc<= 0, really_small_number, dif_perc))
+  # group_by(experiment) %>% 
+  # filter(dif_perc %>% value_filt_tails(c(0, right_tail)))
 
@@
 We do this analysis for all integer solutions obtained.
-As can be seen, in almost all of the cases there is a reduction in the variance.
+To better visualize the results, we use a log 2 scale.
 
@@
 
-ggplot(data=variances_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('Difference in variance (in % of base case)') + theme(text = element_text(size=20)) + coord_flip()
+ggplot(data=variances_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('Difference in variance (in % of base case) (log2 scale)') + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans='log2', labels=scaleFUN)
 

### 2019-11-05 R/nps_reports/datasets.R

@@ -325,16 +325,24 @@ get_all_compare_2 <- function(){
                         scenario_filter='numparalleltasks_2')
 }
 
-get_determ_2 <- function(){
-    get_generic_compare(c('IT000125_20190729', 'IT000125_20191017', 'IT000125_20190915'),
-                        exp_names = list('base', 'determ', 'cuts_sto'),
-                        scenario_filter='numparalleltasks_2')
+get_all_2 <- function(){
+    get_generic_compare(c('IT000125_20190729', 'IT000125_20191017', 'IT000125_20190915', 'IT000125_20191030', 'IT000125_20191025'),
+                        exp_names = list('base', 'cuts_determ', 'cuts_sto', 'old', 'cuts_FlexLP'),
+                        scenario_filter='numparalleltasks_2') %>% 
+        correct_old_model
 }
 
 get_determ_3 <- function(){
     get_generic_compare(c('IT000125_20190801', 'IT000125_20191017', 'IT000125_20190917', 'IT000125_20191030'),
                         exp_names = list('base', 'cuts_determ', 'cuts_sto', 'old'),
                         scenario_filter='numparalleltasks_3') %>% 
+        correct_old_model
+}
+
+get_all_fixLP <- function(){
+    get_generic_compare(c('IT000125_20190729', 'IT000125_20191104', 'IT000125_20191023', 'IT000125_20191025', 'IT000125_20190917'),
+                        exp_names = list('base', 'FlexLP_3', 'fixLP', 'FlexLP', 'ML'),
+                        scenario_filter='numparalleltasks_2') %>% 
         correct_old_model
 }
 

### 2019-11-05 R/nps_reports/functions.R

@@ -161,8 +161,8 @@ get_soft_constraints <- function(raw_df, quant_max, compare=TRUE){
         raw_df %>% 
         get_soft_constraints_2 %>%  
         group_by(scenario, experiment) %>% 
-        summarise(errors_mean = mean(errors),
-                  errors_q95 = quantile(errors, quant_max))
+        summarise(errors_mean = mean(dif),
+                  errors_q95 = quantile(dif, quant_max))
     if (!compare){
         return(result)
     }

### 2019-11-08 R/presentations/Dassault_specs/OPTIMA_description.Rmd
dassault specs in Rmd

---
title: "OPTIMA: spécifications"
author: "ISAE SUPAERO"
date: "Août 21, 2018"
output:
  html_document:
    df_print: paged
    css: ./../../css/baobab_docs.css
    toc: yes
  pdf_document: default
  word_document: default
institute: ISAE SUPAERO
# geometry: margin=1.5in
---

## Glossaire:

* **Maintenance**: opération d'entretien (préventive) technique nécessitant l'immobilisation de l'avion concerné. Aussi appelée "visite". Les maintenances considérées sont de quatre types. Chaque type de visite a des caractéristiques spécifiques: une durée, une périodicité, une capacité, etc.
    * Visite de sécurité (VS).
    * Visite de graissage (VG).
    * Visite intermédiaire (VI).
    * Grande visite (VX ou M).
* **Heures de vol (HDV)**: nombre d'heures de l'utilisation effective d'un avion pendant un certain temps (un mois, un an).
* **Butée calendaire (BC)**: intervalle de temps calendaire de durée maximale entre deux visites consécutives du même type. Unité: jours.
* **Butée horaire (BH)**: nombre d’heures de vol autorisées entre deux visites consécutives du même type. Unité: heures (de vol).
* **Flotte**: ensemble des avions concernés par la planification. Chaque avion a ses caractéristiques techniques et son historique des maintenances et des heures de vol effectuées.
* **Livraison**: arrivée d’un nouvel avion dans la flotte. 
* **Base aérienne**: endroit où sont réalisées les visites de la flotte. Pour chaque type de visite, le nombre d’avions est limité par une capacité de maintenance de ce type.
* **Vieillissement**: le nombre d’heures effectuées par un avion depuis son entrée en service. Pour l’instant,  les indices de fatigue ne sont pas considérés.
* **Vieillissement homogène**: l’objectif recherché est d’avoir le ratio « vieillissement/temps depuis l'entrée en service » le plus homogène possible pour les différents avions de la flotte.
* **Indices de fatigue**: indices spécifiques de différents composants de l’avion dont l’analyse peut déclencher une maintenance.

## Aéronefs

Chaque avion a un état défini au moment de la planification :

* **Vieillissement**.
* **Date d'entrée en service**.

et pour chaque type de visite:

* **Date de la dernière visite**.
* **Nombre d'heures de vol après la dernière visite**.

Il est considéré qu'en dépassant 5000 HDV, l'avion ne peut plus être utilisé pour voler.
La flotte aura une taille de 60 avions au début de l'horizon de planification.

### Livraisons

Les livraisons de nouveaux avions se font par lot avec une moyenne annuelle. En partant d'une quantité annuelle, les avions arrivent par lot tout au long de l'année.

Par exemple, avec une moyenne de 10 avions livrés et un lot de taille 4, la distribution serait la suivante :

* Avril: 4 avions.
* Août: 4 avions.
* Décembre: 2 avions.

## Missions / Heures de vol

### Heures par défaut

Au moment de décider les maintenance, on suppose une charge en heures de vol à réaliser par avion. Cette demande annuelle est équivalente à $200 \frac{HDV}{avion-année} \times Q \frac{avions}{flotte}$. L'utilisation particulière de chaque avion pourra être configuré de manière indépendante et par périodé.

* Un avion effectue normalement entre 180 et 220 HDV/ an.
* La demande est repartie sur l'année en fonction des livraisons: un avion qui arrive en juillet effectue 100 HDV/ an.
* L'homogénéisation de la flotte est le critère utilisé pour décider de l'affectation réelle des heures de vol.
* Tout avion qui n'est pas en maintenance doit être utilisé pour effectuer une quantité d'heures équivalente à sa moyenne annuelle.

### Missions

Les affectations d'avions aux missions se font une fois les visites ont été décidées. Dans ce sens, les affectations doivent être compatibles avec les visites déjà présentes. Ces compatibilités sont les suivantes :

1. Un avion ne peut pas voler plus de 500 h (ou la butée horaire VS correspondante) en sommant toutes les heures résultantes des affectations, entre une visite VS et une visite M (VX). Ici, nous comptons pas les heures de vol par défaut de l'avion, vu que ces heures peuvent être réduites en fonction des heures de vol des affectations.
2. Un avion ne peut pas être affecté à une mission pendant une visite de sécurité ou une grande visite.
3. Un avion affecté à une mission réduit la capacité de maintenance en une fraction correspondant à le pourcentage que l'avion représente de la flotte. La formule serait

## Maintenances

La base aérienne dispose d'un nombre de jours ouvrés par an (200 par défaut) pour effectuer des opérations de maintenance. La distribution de ces jours ouvrés sur l'année n'est pas homogène.

Chaque type de maintenance se différencie par les paramètres suivants :

* **Durée (mois)**: durée d'affectation de l'opération en périodes.
* **BC (mois)**: nombre maximal de mois entre deux visites du même type.
* **BC (tol.)**: taille de l'interval de tolérance (en mois) de la BC.
* **BH (HDV)**: nombre maximal d'heures de vol entre deux visites du même type.
* **BH (tol)**: taille de l'interval de tolérance (en heures de vol) de la BH.
* **Utilisation**: utilisation unitère du résource de capacité.
* **Type**: Type de maintenance. Les visite du même type partage la même capacité.

|Caractéristique|VG  |VI  |VS  |VX  |
|-----------    |----|----|----|----|
|Durée (mois)   | 1  | 1  | 1  | 4 |
|BC (mois)      |8   | 16 | -  |60  |
|BC (tol.)      |2   | 2  | -  | 4  |
|BH (HDV)       |  - | -  | 500|1000|
|BH (tol.)      | -  | -  | 100|  - |
|Capacité       | 3  | 3  | 3  | 4  |
|Type           | 2  |2 | 2  | 1  |

La validation d'un seul critère de maintenance suffit pour immobiliser l'avion.

## Types de maintenances et capacités

* **Capacité**: nombre maximal d'avions en maintenance du même type à chaque moment.

Cependant, tout type de maintenance ne possède pas l'ensemble de ces paramètres. Pour les quatre types de maintenance considérés, voici les paramètres à prendre en compte :

Les trois premières types de visite (VS, VG, VI) sont faites sur chaque base aérien. La quatrième visite (VX) se fait sur l'atelier dédié de façon indépendant. Donc, la capacité des visites qui se déroulent sur la base aérien est partagée.

La visite VX (M) inclut déjà toutes les autres visites et, donc, réinitialise les BC et BH des avions pour les autres 3 visites.

Les visites sur la base récupère le potentiel au début de l'opération. Pour la visite VX, le potentiel est récupéré à la fin de la visite.

Pour la VS les clients préfère déclencher la visite toutes les 500 - 600 h de vol.

### Regroupements de visites

Il existe la possibilité de regrouper deux ou plus de visites. Un tel regroupement permet d'économiser un jour par visite regroupée. Par exemple, si les visites `VG` et `VI` sont regroupées, leur durée totale sera de $3 + 6 - 1 = 8$ jours. Si les visites `VG`, `VI` et `VS`  sont regroupées, leur durée totale sera de $3 + 6 + 4 - 2 = 11$ jours. C'est aussi possible de faire les visites `VG` et `VS` sans faire la visite `VI`.

### Calendrier de visites

La capacité de maintenance dépend des nombres de jours disponibles par mois. Les grandes visites (VX) ont un calendrier et les visites sur la base aérien ont un autre calendrier.

## Stockage

Il est possible de stocker un avion pendant un temps déterminé. L'avion stocké est immobilisé et son service est arrêté, ce qui signifie que le temps de stockage est exclu du calcul des butées calendaires.

* Le temps de stockage correspond normalement à 24 mois. Les options considérées de stockage seront les suivantes: 12 mois, 18 mois et 24 mois.
* L'avion stocké est considéré comme indisponible.
* Il n'y a pas de limite sur le nombre d'avions stockés.
* La mise en stockage d'un avion (et la sortie du stockage) prend entre 1 et 2 jours et utilise les ressources des maintenances dans la base.

## Objectifs

L'objectif principal est de maximiser la disponibilité de la flotte.

La prise en compte des incertitudes notamment sur la durée des maintenances se fera dans un deuxième temps.

L'horizon de planification est de 30 ans.

### 2019-11-12 R/presentations/Dassault_specs/OPTIMA_description.Rmd
modifications on dassault specs.

@@
     df_print: paged
-    css: ./../../css/baobab_docs.css
+    css: ./../../../css/baobab_docs.css
     toc: yes
@@
 * **Date d'entrée en service**.
+* **Type**.
 
@@
 La flotte aura une taille de 60 avions au début de l'horizon de planification.
+Le type sera utilisé pour pouvoir affecter les avions aux missions.
 
@@
 
+**Les livraisons ne font pas partis de la première version officiel (2019-12-31) du logiciel.**
+
 ## Missions / Heures de vol
@@
 
-Au moment de décider les maintenance, on suppose une charge en heures de vol à réaliser par avion.
-Cette demande annuelle est équivalente à $200 \frac{HDV}{avion-année} \times Q \frac{avions}{flotte}$.
-L'utilisation particulière de chaque avion pourra être configuré de manière indépendante et par périodé.
+Au moment de décider les calendrier de maintenances, nous supposons une charge en heures de vol à réaliser par avion.
+Cette charge peut être calculée à partir de la demande annuelle par avion (200 HDV) et sera représentée comme une demande mensuelle d'heures de vol.
 
-* Un avion effectue normalement entre 180 et 220 HDV/ an.
-* La demande est repartie sur l'année en fonction des livraisons: un avion qui arrive en juillet effectue 100 HDV/ an.
-* L'homogénéisation de la flotte est le critère utilisé pour décider de l'affectation réelle des heures de vol.
-* Tout avion qui n'est pas en maintenance doit être utilisé pour effectuer une quantité d'heures équivalente à sa moyenne annuelle.
+Tout avion qui n'est pas dans une grande visite (VX ou M) est considéré comme disponible pour voler et, donc, doit être utilisé pour effectuer une quantité d'heures équivalente à sa demande mensuelle.
+
+Le taux mensuel d'utilisation particulier par avion pourra être configuré par l'utilisateur de façon indépendante.
 
@@
 
-Les affectations d'avions aux missions se font une fois les visites ont été décidées.
+Un ensemble de missions sont connues au long de l'horizon de planification. Chaque mission contient :
+
+1. Un nombre d'avions à utiliser.\*
+2. Un type d'avion à utiliser.
+2. Un nombre d'heures de vol (HDV) à réaliser pour chaque avion à chaque période (mois) de la mission.
+3. Une date de début et une date de fin (en périodes ou mois).
+
+\* Si une mission a besoin de deux types d'avions différents, on devra diviser la mission originale en plusieurs missions indépendantes avec les mêmes informations, en fonction du nombre de types.
+
+Pendant l'affectation à une mission, l'avion ne peut pas faire les maintenances VX et VS.
+La capacité de la base pour faire de maintenances du type 2 est réduite de façon proportionnelle à la proportion des avions qui partent en fonction de la flotte entière.
+
+Exemple: une flotte de 20 avions qui a une capacité de maintenance de 40 jours-hommes voit partir 5 avions en mission. La capacité restante sera de $40 \times (20-5)/20 = 30$ jours-hommes.
+
+Les affectations d'avions aux missions se font une fois les visites ont été déjà décidées.
+Concernant les heures par défaut, elles seront réduites pour les avions qui partent en mission.
 Dans ce sens, les affectations doivent être compatibles avec les visites déjà présentes.
-Ces compatibilités sont les suivantes :
+Ces compatibilités à tenir compte sont les suivantes :
 
 1.
-Un avion ne peut pas voler plus de 500 h (ou la butée horaire VS correspondante) en sommant toutes les heures résultantes des affectations, entre une visite VS et une visite M (VX).
-Ici, nous comptons pas les heures de vol par défaut de l'avion, vu que ces heures peuvent être réduites en fonction des heures de vol des affectations.
+Un avion ne peut pas voler plus de 500 h (ou la butée horaire VS / VX correspondante) en sommant toutes les heures résultantes des affectations, entre une visite VS et une visite VX (ou M) et entre deux visites du type VX.
+Ici, nous comptons pas les heures de vol par défaut de l'avion, vu que ces heures peuvent se voir réduites en fonction des heures de vol provenant des affectations aux missions.
 2. Un avion ne peut pas être affecté à une mission pendant une visite de sécurité ou une grande visite.
-3. Un avion affecté à une mission réduit la capacité de maintenance en une fraction correspondant à le pourcentage que l'avion représente de la flotte. La formule serait
+3. Un avion ne peut pas être affecté à deux missions en même temps.
+3. Suite à la réduction de capacité de maintenance liée à la mission, il doit rester assez de capacité pour faire les visites des avions qui restent sur place.
 
@@
 
-La base aérienne dispose d'un nombre de jours ouvrés par an (200 par défaut) pour effectuer des opérations de maintenance. La distribution de ces jours ouvrés sur l'année n'est pas homogène.
+### Caractéristiques de visites
+
+Les maintenances sont les opérations qui récupèrent le potentiel des avions.
+Chaque visite doit être programmée de façon à respecter la butée horaire et la butée calendaire, en fonction du cas.
+Pour la visite VX, elle doit respecter le plus contraignant des deux.
 
@@
 * **BH (tol)**: taille de l'interval de tolérance (en heures de vol) de la BH.
-* **Utilisation**: utilisation unitère du résource de capacité.
+* **Utilisation**: utilisation unitaire du ressource de capacité.
 * **Type**: Type de maintenance. Les visite du même type partage la même capacité.
+
+Cependant, tout type de maintenance ne possède pas l'ensemble de ces paramètres. Pour les quatre types de maintenance considérés, voici les paramètres à prendre en compte :
 
@@
 |BH (tol.)      | -  | -  | 100|  - |
-|Capacité       | 3  | 3  | 3  | 4  |
+|Utilisation    |  3 | 6  |  4 |  1 |
 |Type           | 2  |2   | 2  | 1  |
@@
 La validation d'un seul critère de maintenance suffit pour immobiliser l'avion.
+La visite VX (M) inclut déjà toutes les autres visites et, donc, réinitialise les BC et BH des avions pour les autres 3 visites.
+Les visites sur la base récupère le potentiel au début de l'opération. Pour la visite VX, le potentiel est récupéré à la fin de la visite.
+Pour la VS les clients préfère déclencher la visite toutes les 500 - 600 h de vol.
 
-## Types de maintenances et capacités
+### Types de maintenances et capacités
 
-* **Capacité**: nombre maximal d'avions en maintenance du même type à chaque moment.
+Il existe deux types d'atelier de maintenance.
+Chaque maintenance ne peut être faite que dans un ces deux ateliers (celui qui corresponde au type de la maintenance).
+Chaque atelier a une capacité donnée par période mais cette capacité peut changer en fonction du période sur l'année.
+L'utilisateur aura les moyens de configurer la capacité de maintenance en fonction de chaque période et de chaque atelier.
 
-Cependant, tout type de maintenance ne possède pas l'ensemble de ces paramètres. Pour les quatre types de maintenance considérés, voici les paramètres à prendre en compte :
+|Atelier        |1   | 2 |
+|-----------    |----|---|
+|Capacité       | 66 | 4 |
 
-Les trois premières types de visite (VS, VG, VI) sont faites sur chaque base aérien.
-La quatrième visite (VX) se fait sur l'atelier dédié de façon indépendant.
-Donc, la capacité des visites qui se déroulent sur la base aérien est partagée.
+A chaque période, la capacité de chaque atelier dois être plus grande que la somme des utilisations unitaires des avions en maintenance dans l'atelier.
 
-La visite VX (M) inclut déjà toutes les autres visites et, donc, réinitialise les BC et BH des avions pour les autres 3 visites.
-
-Les visites sur la base récupère le potentiel au début de l'opération. Pour la visite VX, le potentiel est récupéré à la fin de la visite.
-
-Pour la VS les clients préfère déclencher la visite toutes les 500 - 600 h de vol.
+Les trois premiers types de visite (VS, VG, VI) sont faites sur chaque base aérien et donc, partagent la capacité. La quatrième visite (VX ou M) se fait sur l'atelier dédié de façon indépendant.
 
@@
 C'est aussi possible de faire les visites `VG` et `VS` sans faire la visite `VI`.
-
-### Calendrier de visites
-
-La capacité de maintenance dépend des nombres de jours disponibles par mois. Les grandes visites (VX) ont un calendrier et les visites sur la base aérien ont un autre calendrier.
 
@@
 
+**Le stockage ne fait pas parti de la première version officiel (2019-12-31) du logiciel.**
+
 ## Objectifs
@@
 
-L'horizon de planification est de 30 ans.
+L'horizon de planification est de 70 périodes.

### 2019-11-18 R/presentations/Dassault_specs/OPTIMA_description.Rmd
changes on specs

@@
 author: "ISAE SUPAERO"
-date: "Août 21, 2018"
+date: "`r Sys.Date()`"
 output:
@@
 
-## Glossaire:
+## Glossaire
 
-* **Maintenance**: opération d'entretien (préventive) technique nécessitant l'immobilisation de l'avion concerné.
+* **Maintenance** : opération d'entretien (préventive) technique nécessitant l'immobilisation de l'avion concerné.
 Aussi appelée "visite".
 Les maintenances considérées sont de quatre types.
-Chaque type de visite a des caractéristiques spécifiques: une durée, une périodicité, une capacité, etc.
+Chaque type de visite a des caractéristiques spécifiques : une durée, une périodicité, une capacité, etc.
     * Visite de sécurité (VS).
@@
     * Visite intermédiaire (VI).
-    * Grande visite (VX ou M).
-* **Heures de vol (HDV)**: nombre d'heures de l'utilisation effective d'un avion pendant un certain temps (un mois, un an).
-* **Butée calendaire (BC)**: intervalle de temps calendaire de durée maximale entre deux visites consécutives du même type. Unité: jours.
-* **Butée horaire (BH)**: nombre d’heures de vol autorisées entre deux visites consécutives du même type. Unité: heures (de vol).
-* **Flotte**: ensemble des avions concernés par la planification. Chaque avion a ses caractéristiques techniques et son historique des maintenances et des heures de vol effectuées.
-* **Livraison**: arrivée d’un nouvel avion dans la flotte. 
-* **Base aérienne**: endroit où sont réalisées les visites de la flotte. Pour chaque type de visite, le nombre d’avions est limité par une capacité de maintenance de ce type.
-* **Vieillissement**: le nombre d’heures effectuées par un avion depuis son entrée en service. Pour l’instant,  les indices de fatigue ne sont pas considérés.
-* **Vieillissement homogène**: l’objectif recherché est d’avoir le ratio « vieillissement/temps depuis l'entrée en service » le plus homogène possible pour les différents avions de la flotte.
-* **Indices de fatigue**: indices spécifiques de différents composants de l’avion dont l’analyse peut déclencher une maintenance.
+    * Grande visite (VX aussi appelée M).
+* **Heures de vol (HDV)** : nombre d'heures de l'utilisation effective d'un avion pendant un certain temps (un mois, un an).
+* **Butée calendaire (BC)** : intervalle de temps calendaire de durée maximale entre deux visites consécutives du même type. Unité : jours.
+* **Butée horaire (BH)** : nombre d’heures de vol autorisées entre deux visites consécutives du même type. Unité : heures (de vol).
+* **Flotte** : ensemble des avions concernés par la planification. Chaque avion a ses caractéristiques techniques et son historique des maintenances et des heures de vol effectuées.
+* **Livraison** : arrivée d’un nouvel avion dans la flotte. 
+* **Base aérienne** : endroit où sont réalisées les visites de la flotte. Pour chaque type de visite, le nombre d’avions est limité par une capacité de maintenance de ce type.
+* **Vieillissement** : le nombre d’heures effectuées par un avion depuis son entrée en service. Pour l’instant,  les indices de fatigue ne sont pas considérés.
+* **Vieillissement homogène** : l’objectif recherché est d’avoir le ratio « vieillissement/temps depuis l'entrée en service » le plus homogène possible pour les différents avions de la flotte.
+* **Indices de fatigue** : indices spécifiques de différents composants de l’avion dont l’analyse peut déclencher une maintenance.
+* **Mission** : une demande spécifique d'avions pendant une durée de temps donnée (quantifiée en périodes) et avec un nombre donné d'heures par période.
 
@@
 
-et pour chaque type de visite:
+et pour chaque type de visite :
 
@@
 
-* Avril: 4 avions.
-* Août: 4 avions.
-* Décembre: 2 avions.
+* Avril : 4 avions.
+* Août : 4 avions.
+* Décembre : 2 avions.
 
-**Les livraisons ne font pas partis de la première version officiel (2019-12-31) du logiciel.**
+**Les livraisons ne font pas partie de la première version officiel (2019-12-31) du logiciel.**
 
@@
 
-### Heures par défaut
+Les heures de vol sont utilisées pour générer une planification initiale de maintenances. Les informations des missions sont utilisées à posteriori pour affecter les avions aux missions.
 
-Au moment de décider les calendrier de maintenances, nous supposons une charge en heures de vol à réaliser par avion.
-Cette charge peut être calculée à partir de la demande annuelle par avion (200 HDV) et sera représentée comme une demande mensuelle d'heures de vol.
+### Heures de vol
+
+Au moment de décider le calendrier de maintenances, nous supposons une charge en heures de vol à réaliser par avion.
+Cette charge peut être calculée à partir de la demande annuelle par avion (200 HDV) et sera représentée comme une demande mensuelle d'heures de vol par avion.
 
@@
 
-Le taux mensuel d'utilisation particulier par avion pourra être configuré par l'utilisateur de façon indépendante.
+Le nombre d'heures de vol précis pour un avion et mois donné pourra être configuré par l'utilisateur.
 
@@
 
-1. Un nombre d'avions à utiliser.\*
+1. Un nombre d'avions à utiliser.
 2. Un type d'avion à utiliser.
-2. Un nombre d'heures de vol (HDV) à réaliser pour chaque avion à chaque période (mois) de la mission.
-3. Une date de début et une date de fin (en périodes ou mois).
+2. Un nombre d'heures de vol (HDV) à réaliser pour chaque avion et à chaque période (mois) de la mission.
+3. Une date (période) de début et une date (période) de fin.
 
-\* Si une mission a besoin de deux types d'avions différents, on devra diviser la mission originale en plusieurs missions indépendantes avec les mêmes informations, en fonction du nombre de types.
+Si une mission donnée avait besoin d'avions de deux types différents, nous devrions séparer la mission originale en plusieurs missions indépendantes, une mission par type.
+Exemple : Mission $O1$ a besoin de $N$ avions de type $1$ et $M$ avions de type $2$.
+Nous devrons, donc, créer deux missions : la mission $O1a$ qui demande $N$ avions et qui a le type $1$; et la mission $O1b$ qui demande $M$ avions et qui a le type $2$.
+Les autres caractéristiques restent les mêmes.
+
+Nous supposons que le nombre d'heures de vol à réaliser est le même pour chaque avion affecté à la mission.
 
 Pendant l'affectation à une mission, l'avion ne peut pas faire les maintenances VX et VS.
+
 La capacité de la base pour faire de maintenances du type 2 est réduite de façon proportionnelle à la proportion des avions qui partent en fonction de la flotte entière.
-
-Exemple: une flotte de 20 avions qui a une capacité de maintenance de 40 jours-hommes voit partir 5 avions en mission. La capacité restante sera de $40 \times (20-5)/20 = 30$ jours-hommes.
+Exemple : une flotte de 20 avions qui a une capacité de maintenance de 40 jours-hommes voit partir 5 avions en mission.
+La capacité restante dans la base sera de $40 \times (20-5)/20 = 30$ jours-hommes pendant la durée de cette mission.
 
 Les affectations d'avions aux missions se font une fois les visites ont été déjà décidées.
-Concernant les heures par défaut, elles seront réduites pour les avions qui partent en mission.
-Dans ce sens, les affectations doivent être compatibles avec les visites déjà présentes.
-Ces compatibilités à tenir compte sont les suivantes :
+Concernant les heures par défaut, elles seront réduites pour les avions qui partent en mission de façon à garder le nombre d'heures de vol de l'avion qui correspond à son cycle de maintenance.
+Les affectations doivent être compatibles avec les visites déjà présentes.
+Les compatibilités à tenir compte sont les suivantes :
 
+1. Un avion affecté à une mission doit être affecté pendant toute la durée de la mission.
 1.
 Un avion ne peut pas voler plus de 500 h (ou la butée horaire VS / VX correspondante) en sommant toutes les heures résultantes des affectations, entre une visite VS et une visite VX (ou M) et entre deux visites du type VX.
-Ici, nous comptons pas les heures de vol par défaut de l'avion, vu que ces heures peuvent se voir réduites en fonction des heures de vol provenant des affectations aux missions.
-2. Un avion ne peut pas être affecté à une mission pendant une visite de sécurité ou une grande visite.
+Ici, nous ne comptons pas les heures de vol par défaut de l'avion, vu que ces heures peuvent se voir réduites en fonction des heures de vol provenant des affectations aux missions.
+2. Un avion ne peut pas être affecté à une mission pendant une visite de sécurité (VS) ou une grande visite (VX ou M).
 3. Un avion ne peut pas être affecté à deux missions en même temps.
@@
 
-Les maintenances sont les opérations qui récupèrent le potentiel des avions.
-Chaque visite doit être programmée de façon à respecter la butée horaire et la butée calendaire, en fonction du cas.
+Les maintenances sont les opérations qui récupèrent le potentiel de vol des avions.
+Chaque visite doit être programmée de façon à respecter la butée horaire et /ou la butée calendaire (en fonction de la visite).
 Pour la visite VX, elle doit respecter le plus contraignant des deux.
@@
 
-* **Durée (mois)**: durée d'affectation de l'opération en périodes.
-* **BC (mois)**: nombre maximal de mois entre deux visites du même type.
-* **BC (tol.)**: taille de l'interval de tolérance (en mois) de la BC.
-* **BH (HDV)**: nombre maximal d'heures de vol entre deux visites du même type.
-* **BH (tol)**: taille de l'interval de tolérance (en heures de vol) de la BH.
-* **Utilisation**: utilisation unitaire du ressource de capacité.
-* **Type**: Type de maintenance. Les visite du même type partage la même capacité.
+* **Durée (mois)** : durée d'affectation de l'opération en périodes.
+* **BC (mois)** : nombre maximal de mois entre deux visites du même type.
+* **BC (tol.)** : taille de l'intervalle de tolérance (en mois) de la BC.
+* **BH (HDV)** : nombre maximal d'heures de vol entre deux visites du même type.
+* **BH (tol)** : taille de l'intervalle de tolérance (en heures de vol) de la BH.
+* **Utilisation** : utilisation unitaire de la ressource de capacité.
+* **Type** : Type de maintenance. Les visite du même type partage la même capacité.
 
@@
 La visite VX (M) inclut déjà toutes les autres visites et, donc, réinitialise les BC et BH des avions pour les autres 3 visites.
-Les visites sur la base récupère le potentiel au début de l'opération. Pour la visite VX, le potentiel est récupéré à la fin de la visite.
-Pour la VS les clients préfère déclencher la visite toutes les 500 - 600 h de vol.
+Les visites de type 2 récupèrent le potentiel au début de l'opération. Pour la visite VX, le potentiel est récupéré à la fin de la visite.
+Pour la VS les clients préfèrent déclencher la visite toutes les 500 - 600 HDV.
 
@@
 Chaque maintenance ne peut être faite que dans un ces deux ateliers (celui qui corresponde au type de la maintenance).
-Chaque atelier a une capacité donnée par période mais cette capacité peut changer en fonction du période sur l'année.
+Chaque atelier a une capacité par période donnée mais elle peut changer en fonction de la période de l'année.
 L'utilisateur aura les moyens de configurer la capacité de maintenance en fonction de chaque période et de chaque atelier.
@@
 
-A chaque période, la capacité de chaque atelier dois être plus grande que la somme des utilisations unitaires des avions en maintenance dans l'atelier.
+Les trois premiers types de visite (VS, VG, VI) sont faites sur chaque base aérien et donc, partagent la capacité. La quatrième visite (VX ou M) se fait sur l'atelier dédié de façon indépendant.
 
-Les trois premiers types de visite (VS, VG, VI) sont faites sur chaque base aérien et donc, partagent la capacité. La quatrième visite (VX ou M) se fait sur l'atelier dédié de façon indépendant.
+A chaque période, la capacité de l'atelier doit être plus grande que la somme des utilisations unitaires (voir *Utilisation* sur le tableau de maintenances) des avions en maintenance dans l'atelier.
+L'utilisation est décrite dans le tableau des caractéristiques des maintenances.
+
+Exemple : imaginons une période où la capacité est égale à 40 pour la maintenance de type 2.
+Nous pouvons réaliser, au maximum, 10 visites VS et aucune autre visite : $10 \times 4 = 40$.
+Alternativement, nous pouvons faire 2 visites VG, 3 visites VI et 4 visites VS : $2 \times 3 + 3 \times 6 + 4 \times 4 = 40$.
 
@@
 
-* Le temps de stockage correspond normalement à 24 mois. Les options considérées de stockage seront les suivantes: 12 mois, 18 mois et 24 mois.
+* Le temps de stockage correspond normalement à 24 mois. Les options considérées de stockage seront les suivantes : 12 mois, 18 mois et 24 mois.
 * L'avion stocké est considéré comme indisponible.
@@
 
-**Le stockage ne fait pas parti de la première version officiel (2019-12-31) du logiciel.**
+**Le stockage ne fait pas partie de la première version officiel (2019-12-31) du logiciel.**
 
@@
 
-L'objectif principal est de maximiser la disponibilité de la flotte.
+L'objectif principal est de maximiser le regroupement de visites de maintenance. Un objectif secondaire vise à retarder les visites de sécurité (VS) le plus possible.
 
-La prise en compte des incertitudes notamment sur la durée des maintenances se fera dans un deuxième temps.
-
-L'horizon de planification est de 70 périodes.
+L'horizon de planification est d'entre 50 et 70 périodes mensuelles (entre 4 et 6 ans).

### 2019-11-20 latex/ROADEF2020/article.tex
article ROADEF2020. biblio ML.

\documentclass{roadef}
\usepackage{amsmath}

% \usepackage{fontspec}
% This command is to use simple quotes inside math expressions:
\usepackage[hyphens,spaces,obeyspaces]{url}
\newcommand{\mq}[1] {`#1\textrm'}

\begin{document}

\title{An alternative MIP formulation for the Military Flight and Maintenance Planning problem}

\author{Franco Peschiera\inst{1}, Alain Ha{\"i}t\inst{1}, Olga Batta{\"i}a\inst{1}, Nicolas Dupin\inst{2}}

\institute{
ISAE-SUPAERO, Universit{\'e} de Toulouse, France \\
\email{\{franco.peschiera,alain.hait,olga.battaia\}@isae-supaero.fr}\\
Laboratoire de Recherche en Informatique, France \\
\email{dupin@lri.fr}

}

\maketitle
\thispagestyle{empty}

\keywords{optimization, planning, military, maintenance}

\section{Introduction}
    The Military Flight and Maintenance Planning Problem assigns missions and schedules maintenance operations (checks) for military aircraft. It has been studied in short, medium and long-term planning horizons (\cite{Cho2011,Kozanidis2008,Verhoeff2015,PeschieraR2018}) and it has been proved NP-Hard in its longterm variant by \cite{Peschiera2019}. The present model is an alternative formulation to that of \cite{Peschiera2019} where mission assignments and maintenance cycles are modeled as start-stop assignment.

    The problem consists in assigning military aircraft $i \in \mathcal{I}$ to a given set $j \in \mathcal{J}$ of missions while scheduling the check over a time horizon divided into $t \in \mathcal{T}$ periods. Missions have known start and end dates, each requires $R_j$ aircraft to be assigned at any given time and to fly $H^\prime_{jtt'}$ when assigned between periods $t$ and $t'$. Each aircraft $i$ can be assigned to only missions $j \in \mathcal{JI}_i$ and is assumed to fly $U^{\prime}_{tt'}$ hours between periods $t$ and $t'$. Each period $t$ has $j \in \mathcal{JT}_t$ active missions.

    Each aircraft $i$ needs a check every $H^{max}$ flight hours or less and has $Rft^{Init}_i$ remaining flight hours at the beginning of the planning horizon. Each aircraft can be assigned its first check during periods $t \in \mathcal{T}^{M_{Init}}_i$ and, if assigned a first check in period $t$, it can have its second check during periods $t \in \mathcal{T}^{M+}_{t}$. $\mathcal{TTT}_t$ is the set of check patterns that make an aircraft be in maintenance in period $t$; $\mathcal{TTJ}_{jt}$ is the set of mission $j$ patterns that make aircraft be in mission in period $t$; $\mathcal{JTT}_{itt'}$ is the set of mission patterns for aircraft $i$ between periods $t$ and $t'$.

\section{Mathematical model}

    \begin{tabular}{p{5mm}lp{130mm}}
        $a_{ijtt'}$ & =& 1 if aircraft $i$ starts mission $j$ at period $t$ and finishes at period $t'$, 0 otherwise. \\  
        $m_{itt'}$  & =& 1 if aircraft $i$ starts a check at period $t$ and then starts the next check at period $t'$, 0 otherwise. \\
    \end{tabular}

    The objective is to delay the second check as much as possible or avoid doing it altogether.
    \begin{align}
        & \text{Max}\; 
        \sum_{
                i \in \mathcal{I}, t \in \mathcal{T}^{M_{Init}}_i, t' \in \mathcal{T}^{M+}_t
            } m_{itt'} \times t'
        \label{eq:objective1}
    \end{align}

    \begin{align}
        % maximum capacity1:
        & \sum_{i \in \mathcal{I}, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} \leq C^{max} 
                & t \in \mathcal{T} \label{eq:capacity1}\\
        % min assignments:
        & \sum_{i \in \mathcal{IJ}_j, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \geq R_j
                & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
        % just doing one thing at any given time:
        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \leq 1 
                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
    \end{align}

    Constraints (\ref{eq:capacity1}) limit the number of simultaneous checks. Constraints (\ref{eq:missionres}) enforce aircraft mission requirements. Constraints (\ref{eq:state}) restrict each aircraft to at most one assignment each period.

    \begin{align}
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours1}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours2}\\
        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2T}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_2T} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours3}
    \end{align}

    Constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}) limit the total flight hours of each aircraft before the first check, between checks and after the second check.

    \begin{align}
        & \sum_{t \in \mathcal{T}^{M_{Init}}_i, t' \in \mathcal{T}^{M+}_{t}} m_{itt'} =  1 
          & i \in \mathcal{I}\label{eq:num_maint}
    \end{align}

    Constraints (\ref{eq:num_maint}) require a check pattern assignment for each aircraft.

\section{Conclusions and future work}

    The present work proposes a new model for the Military Flight and Maintenance Problem. The formulation results in better lower bounds (LP-relaxation) compared to \cite{Peschiera2019} and lends itself to pattern prediction forecasting models, which can be used to reduce the size of the model.

% La bibliographie

\bibliographystyle{plain}
% \selectlanguage{english}
\bibliography{./../biblio/MFMP,./../biblio/FrancoPeschiera}

\end{document}

### 2019-11-20 latex/ROADEF2020/article.tex
changes to ROADEF2020.

@@
 
-\author{Franco Peschiera\inst{1}, Alain Ha{\"i}t\inst{1}, Olga Batta{\"i}a\inst{1}, Nicolas Dupin\inst{2}}
+\author{Franco Peschiera\inst{1}, Alain Ha{\"i}t\inst{1}, Olga Batta{\"i}a\inst{2}, Nicolas Dupin\inst{3}}
 
@@
 ISAE-SUPAERO, Universit{\'e} de Toulouse, France \\
-\email{\{franco.peschiera,alain.hait,olga.battaia\}@isae-supaero.fr}\\
+\email{\{franco.peschiera,alain.hait\}@isae-supaero.fr}\\
+KEDGE Business School, France \\
+\email{olga.battaia@kedgebs.com} \\
 Laboratoire de Recherche en Informatique, France \\
 \email{dupin@lri.fr}
-
 }
@@
 
-    The problem consists in assigning military aircraft $i \in \mathcal{I}$ to a given set $j \in \mathcal{J}$ of missions while scheduling the check over a time horizon divided into $t \in \mathcal{T}$ periods.
-Missions have known start and end dates, each requires $R_j$ aircraft to be assigned at any given time and to fly $H^\prime_{jtt'}$ when assigned between periods $t$ and $t'$.
+    The problem consists in assigning aircraft $i \in \mathcal{I}$ to a set $j \in \mathcal{J}$ of missions while scheduling checks over a time horizon divided into $t \in \mathcal{T}$ periods.
+Missions have known start and end dates; and requires $R_j$ aircraft to fly $H^\prime_{jtt'}$ when assigned between periods $t$ and $t'$.
 Each aircraft $i$ can be assigned to only missions $j \in \mathcal{JI}_i$ and is assumed to fly $U^{\prime}_{tt'}$ hours between periods $t$ and $t'$.
@@
 
-    Each aircraft $i$ needs a check every $H^{max}$ flight hours or less and has $Rft^{Init}_i$ remaining flight hours at the beginning of the planning horizon.
-Each aircraft can be assigned its first check during periods $t \in \mathcal{T}^{M_{Init}}_i$ and, if assigned a first check in period $t$, it can have its second check during periods $t \in \mathcal{T}^{M+}_{t}$.
+    Each aircraft $i$ needs a check after $H^{max}$ flight hours or less; has $Rft^{Init}_i$ remaining flight hours at the beginning of the planning horizon; can be assigned its first check during periods $t \in \mathcal{T}^{M_{Init}}_i$ and, if assigned a first check in period $t$, it can have its second check during periods $t \in \mathcal{T}^{M+}_{t}$.
 $\mathcal{TTT}_t$ is the set of check patterns that make an aircraft be in maintenance in period $t$; $\mathcal{TTJ}_{jt}$ is the set of mission $j$ patterns that make aircraft be in mission in period $t$; $\mathcal{JTT}_{itt'}$ is the set of mission patterns for aircraft $i$ between periods $t$ and $t'$.

### 2019-11-21 R/nps_reports/results.Rmd

@@
   size: get_3_tasks_aggresive_perc_add
+  scale: log2
 editor_options:
@@
 source('nps_reports/datasets.R')
-quant_max <- 0.95
+default_tail <- 0.05
 really_small_number <- 0.000001
@@
 scaleFUN <- function(x) sprintf("%.2f", x)
+scale <- params$scale
 
@@
 
-t1_rel <- quality_degr_all %>% filter(experiment=='cuts') %>% use_series(dif_perc)
-t1_abs <- quality_degr_all %>% filter(experiment=='cuts') %>% use_series(dif)
-quality_degr_quant <- t1_rel %>% quantile()
-quality_degr_quant95 <- t1_rel %>% quantile(quant_max)
-quality_degr_quant95_abs <- t1_abs %>% quantile(quant_max)
-
-right_tail <- 0.0
+right_tail <- default_tail
 quality_degr_all_filt <- 
@@
 The following boxplot shows the distribution of the gap from the models with respect to the `base` model.
-Results are shown in logaritmic scale and so we have replace all negative values for a very small one.
+We have replace all negative values for a very small one close to zero.
+We have taken (`r right_tail*100`\%) values from the right side to better visualize the output.
 
 ```{r}
-ggplot(data=quality_degr_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('% difference in objective (log2 scale)') + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans='log2', labels=scaleFUN)
+if (scale!='identity'){
+  ylab <- sprintf('%% difference in objective (%s scale)', scale)  
+} else {
+  ylab <- '% difference in objective'
+}
+
+ggplot(data=quality_degr_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab(ylab) + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans=scale, labels=scaleFUN)
 
@@
 
-quality_perf <- quality_perf_all %>% filter(experiment=='cuts')
-mean_dif <- quality_perf$dif_perc %>%  mean
-per_cut <- sum(quality_perf$dif_perc<0)/length(quality_perf$dif_perc)
-q95_cut <- quality_perf$dif_perc %>% quantile(0.95)
-
 ```
@@
 # this takes some % from each side
-each_tail <- 0
+each_tail <- default_tail
 
@@
 A negative relative gap means the model outperformed the `base` model in that particular instance.
-Results are shown in logaritmic scale and so we have replace all negative values for a very small one.
+We have taken (`r each_tail*100`\%) values from each side to better visualize the output.
 
@@
 
-ggplot(data=quality_perf_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('% difference in objective (log2 scale)') + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans='log2', labels=scaleFUN)
+if (scale!='identity'){
+  ylab <- sprintf('%% difference in objective (%s scale)', scale)  
+} else {
+  ylab <- '% difference in objective'
+}
+
+ggplot(data=quality_perf_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab(ylab) + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans=scale, labels=scaleFUN)
 
@@
 
+```{r soft_constraints}
+right_tail <- default_tail
+errors_stats <- get_soft_constraints(df)
+errors_all <- get_soft_constraints_2(df) %>% 
+  mutate(dif= if_else(dif<=0, really_small_number, dif))
+
+errors_all_filt <- 
+  errors_all %>% 
+  filter(dif %>% value_filt_tails(c(0, right_tail)))
+
+```
+
 Regarding the accomplishment of soft constraints, we can analyze for mutually **optimal solutions**, the quantity of violations in their solutions.
@@
 We show only the extra errors (in number) compared to the `base` case.
-
-```{r soft_constraints}
-errors_stats <- get_soft_constraints(df, quant_max)
-errors_all <- get_soft_constraints_2(df) %>% 
-  mutate(dif= if_else(dif<=0, really_small_number, dif))
-
-```
+We have taken (`r right_tail*100`\%) values from he right side to better visualize the output.
 
@@
 
-And the complete distributions in logaritmic scale:
+And the complete distributions:
 
@@
 
-ggplot(data=errors_all, aes(x=experiment, y=dif)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('absolute difference in errors  (log2 scale)') + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans='log2', labels=scaleFUN)
+if (scale!='identity'){
+  ylab <- sprintf('absolute difference in errors (%s scale)', scale)
+} else {
+  ylab <- sprintf('absolute difference in errors')
+}
+
+ggplot(data=errors_all_filt, aes(x=experiment, y=dif)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab(ylab) + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans=scale, labels=scaleFUN)
 
@@
 variances_all <- get_variances(df)
-# var_per <- variances_all %>% filter(experiment=='cuts') %>% use_series(dif_perc)
-# mean_var_reduction <- var_per %>% mean
-# quant95_var_reduction <- var_per %>% quantile(quant_max)
 
@@
   variances_all %>% 
-  mutate(dif_perc = if_else(dif_perc<= 0, really_small_number, dif_perc))
-  # group_by(experiment) %>% 
-  # filter(dif_perc %>% value_filt_tails(c(0, right_tail)))
+  group_by(experiment) %>%
+  filter(dif_perc %>% value_filt_tails(c(0, right_tail)))
 
@@
 We do this analysis for all integer solutions obtained.
-To better visualize the results, we use a log 2 scale.
+To better visualize the results, we have taken `r right_tail*100`\% from the right tail.
 
@@
 
-ggplot(data=variances_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('Difference in variance (in % of base case) (log2 scale)') + theme(text = element_text(size=20)) + coord_flip() + scale_y_continuous(trans='log2', labels=scaleFUN)
+ggplot(data=variances_all_filt, aes(x=experiment, y=dif_perc)) + theme_minimal() + geom_boxplot() + xlab('Experiment') + ylab('Difference in variance (in % of base case)') + theme(text = element_text(size=20)) + coord_flip()
 

### 2019-11-21 R/presentations/Dassault_specs/OPTIMA_description.Rmd
more changes to Dassault specs

@@
 
-**Les livraisons ne font pas partie de la première version officiel (2019-12-31) du logiciel.**
+**Les livraisons ne font pas partie de la première version officielle (2019-12-31) du logiciel.**
 
@@
 
-Le nombre d'heures de vol précis pour un avion et mois donné pourra être configuré par l'utilisateur.
+Le nombre d'heures de vol précis pour un avion et période donné pourra être configuré par l'utilisateur. Trois niveaux de configuration sont disponibles :
+
+1. "Heures de vol par période" pour toute la flotte. 1 paramètre par planification.
+2. Heures de vol par période pour chaque avion. 1 paramètre pour avion.
+3. Heures de vol pour chaque avion et chaque période. 1 paramètre pour avion-mois:
+
+Exemple:
+
+```
+params.min_usage_period = 15
+```
+
+```
+avions=
+| avion| mois     |  min_usage_period    |
+|-----------    |----|----|
+| A_1  | 2019-04  |  14    |
+| A_2  | 2019-04  |  14    |
+| A_2  | 2019-04  |  14    |
+| A_2  | 2019-04  |  14    |
+
+```
+
+```
+heures_vol=
+|avion  |   heures  | min_usage_period  |
+|-----------    |----|----|             
+|A_1    | 0         |   16              |
+|A_10   |  0        |   16              |
+|A_11   |  0        |   16              |
+|A_12   |  0        |   16              |
+
+```
+
+| avion| mois     |  min_usage_period    |
+|-----------    |----|----|
+| A_1  | 2019-04  |  14    |
+| A_2  | 2019-04  |  14    |
+| A_2  | 2019-04  |  14    |
+| A_2  | 2019-04  |  14    |
 
@@
 3. Suite à la réduction de capacité de maintenance liée à la mission, il doit rester assez de capacité pour faire les visites des avions qui restent sur place.
+
+Exemple du point (2):
+<!-- TODO -->
 
@@
 
-**Le stockage ne fait pas partie de la première version officiel (2019-12-31) du logiciel.**
+**Le stockage ne fait pas partie de la première version officielle (2019-12-31) du logiciel.**
 

### 2019-11-22 R/presentations/Dassault_specs/OPTIMA_description.Rmd
changes on Dassault specs and template.

--- 2019-11-18 R/presentations/Dassault_specs/OPTIMA_description.Rmd
@@
 Si une mission donnée avait besoin d'avions de deux types différents, nous devrions séparer la mission originale en plusieurs missions indépendantes, une mission par type.
-Exemple : Mission $O1$ a besoin de $N$ avions de type $1$ et $M$ avions de type $2$.
-Nous devrons, donc, créer deux missions : la mission $O1a$ qui demande $N$ avions et qui a le type $1$; et la mission $O1b$ qui demande $M$ avions et qui a le type $2$.
+Exemple : Mission $O1$ a besoin de $N$ avions de type $1$ et $L$ avions de type $2$.
+Nous devrons, donc, créer deux missions : la mission $O1a$ qui demande $N$ avions et qui a le type $1$; et la mission $O1b$ qui demande $L$ avions et qui a le type $2$.
 Les autres caractéristiques restent les mêmes.

### 2019-11-22 latex/ROADEF2020/article.tex
changed Nicolas reference affiliation.
changes made by Nicolas. As submitted.

@@
 
-\author{Franco Peschiera\inst{1}, Alain Ha{\"i}t\inst{1}, Olga Batta{\"i}a\inst{2}, Nicolas Dupin\inst{3}}
+\author{Franco Peschiera\inst{1}, Nicolas Dupin\inst{2}, Olga Batta{\"i}a\inst{3}, Alain Ha{\"i}t\inst{1}}
 
@@
 \email{olga.battaia@kedgebs.com} \\
-Laboratoire de Recherche en Informatique, France \\
+LRI, Universit\'e Paris-Saclay, France \\
 \email{dupin@lri.fr}
@@
 
-\keywords{optimization, planning, military, maintenance}
+\keywords{optimization, planning, military aircraft, maintenance}
 
 \section{Introduction}
-    The Military Flight and Maintenance Planning Problem assigns missions and schedules maintenance operations (checks) for military aircraft.
-It has been studied in short, medium and long-term planning horizons (\cite{Cho2011,Kozanidis2008,Verhoeff2015,PeschieraR2018}) and it has been proved NP-Hard in its longterm variant by \cite{Peschiera2019}.
+    The Military Flight and Maintenance Planning (MFMP) Problem assigns missions and schedules maintenance operations (checks) for military aircraft.
+It has been studied in short, medium and long-term planning horizons (\cite{Cho2011,Kozanidis2008,Verhoeff2015,PeschieraR2018}) and it has been proved NP-Hard in its long-term variant by \cite{Peschiera2019}.
 The present model is an alternative formulation to that of \cite{Peschiera2019} where mission assignments and maintenance cycles are modeled as start-stop assignment.
@@
     The problem consists in assigning aircraft $i \in \mathcal{I}$ to a set $j \in \mathcal{J}$ of missions while scheduling checks over a time horizon divided into $t \in \mathcal{T}$ periods.
-Missions have known start and end dates; and requires $R_j$ aircraft to fly $H^\prime_{jtt'}$ when assigned between periods $t$ and $t'$.
-Each aircraft $i$ can be assigned to only missions $j \in \mathcal{JI}_i$ and is assumed to fly $U^{\prime}_{tt'}$ hours between periods $t$ and $t'$.
-Each period $t$ has $j \in \mathcal{JT}_t$ active missions.
+Missions have known start and end dates and require $R_j$ aircraft to fly $H^\prime_{jtt'}$ hours when assigned to mission $j$ between periods $t$ and $t'$.
+Each aircraft $i$ can be assigned to only missions $j \in \mathcal{J}_i$ and is assumed to fly $U^{\prime}_{tt'}$ hours between periods $t$ and $t'$ if not undergoing a check or a mission.
+Each period $t$ has $j \in \mathcal{J}_t$ active missions.
 
-    Each aircraft $i$ needs a check after $H^{max}$ flight hours or less; has $Rft^{Init}_i$ remaining flight hours at the beginning of the planning horizon; can be assigned its first check during periods $t \in \mathcal{T}^{M_{Init}}_i$ and, if assigned a first check in period $t$, it can have its second check during periods $t \in \mathcal{T}^{M+}_{t}$.
-$\mathcal{TTT}_t$ is the set of check patterns that make an aircraft be in maintenance in period $t$; $\mathcal{TTJ}_{jt}$ is the set of mission $j$ patterns that make aircraft be in mission in period $t$; $\mathcal{JTT}_{itt'}$ is the set of mission patterns for aircraft $i$ between periods $t$ and $t'$.
+    Each aircraft $i$ needs a check after $H^{max}$ flight hours or less and has $Rft^{Init}_i$ remaining flight hours at the beginning of the planning horizon.
+Only two checks can be assigned in the planing horizon: the first check is assigned during periods $t \in \mathcal{T}^{M_{Init}}_i$ and, a second check can be assigned during periods $t \in \mathcal{T}^{M+}_{t}$, if the first check was done in $t$.
+$\mathcal{T}_t$ is the set of check patterns that make an aircraft be in maintenance in period $t$; $\mathcal{J}_{jt}$ is the set of mission patterns that make aircraft be in mission $j$ in period $t$; $\mathcal{J}_{itt'}$ is the set of mission patterns for aircraft $i$ between periods $t$ and $t'$.
 
 \section{Mathematical model}
+
+    The following variables are defined in this new Integer Linear Formulation (ILP):
 
@@
 
-    The objective is to delay the second check as much as possible or avoid doing it altogether.
+    The objective is to delay the second check as much as possible.
+
+    \begin{small}
          \begin{align}
@@
             } m_{itt'} \times t'
-        \label{eq:objective1}
-    \end{align}
-
-    \begin{align}
+        \label{eq:objective1}\\
+%     \end{align}
+%         \begin{align}
         % maximum capacity1:
-        & \sum_{i \in \mathcal{I}, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} \leq C^{max} 
+        & \sum_{i \in \mathcal{I}, (t_1, t_2) \in \mathcal{T}_{t}} m_{it_1t_2} \leq C^{max} 
                 & t \in \mathcal{T} \label{eq:capacity1}\\
         % min assignments:
-        & \sum_{i \in \mathcal{IJ}_j, (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \geq R_j
+        & \sum_{i \in \mathcal{I}_j, (t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \geq R_j
                 & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
         % just doing one thing at any given time:
-        & \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{JT}_t \cap \mathcal{JI}_i} \sum_{(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}} a_{ijt_1t_2} \leq 1
-                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
-    \end{align}
-
-    Constraints (\ref{eq:capacity1}) limit the number of simultaneous checks.
-Constraints (\ref{eq:missionres}) enforce aircraft mission requirements.
-Constraints (\ref{eq:state}) restrict each aircraft to at most one assignment each period.
-
-    \begin{align}
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
+        & \sum_{(t_1, t_2) \in \mathcal{T}_{t}} m_{it_1t_2} + \sum_{j \in \mathcal{J}_t \cap \mathcal{J}_i} \sum_{(t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \leq 1 
+                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
+%     \end{align}
+% 
+% 
+%     \begin{align}
+        & \sum_{\substack{(j, t, t') \in  \mathcal{J}_{i1t_1}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
             & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours1}\\
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+        & \sum_{\substack{(j, t, t') \in  \mathcal{J}_{it_1t_2}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
             & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours2}\\
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2T}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_2T} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
-            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours3}
-    \end{align}
-
-    Constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}) limit the total flight hours of each aircraft before the first check, between checks and after the second check.
-
-    \begin{align}
+        & \sum_{\substack{(j, t, t') \in  \mathcal{J}_{it_2T}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_2T} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+            & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^{M+}_{t_1} \label{eq:cycle_hours3}\\
+%     \end{align}
+% 
+%    
+%     \begin{align}
         & \sum_{t \in \mathcal{T}^{M_{Init}}_i, t' \in \mathcal{T}^{M+}_{t}} m_{itt'} =  1 
@@
     \end{align}
-
+        \end{small}
+    Constraints (\ref{eq:capacity1}) limit the number of simultaneous checks,  (\ref{eq:missionres}) enforce aircraft mission requirements and  (\ref{eq:state}) restrict each aircraft to at most one assignment each period.
+    Constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}) limit the total flight hours of each aircraft before the first check, between checks and after the second check.
 Constraints (\ref{eq:num_maint}) require a check pattern assignment for each aircraft.
@@
 
-    The present work proposes a new model for the Military Flight and Maintenance Problem.
-The formulation results in better lower bounds (LP-relaxation) compared to \cite{Peschiera2019} and lends itself to pattern prediction forecasting models, which can be used to reduce the size of the model.
+    The present work proposes a new model for the MFMP Problem,
+    which can be seen as an extended ILP formulation  from \cite{Peschiera2019},  with a     twice-time index formulation.
+    This new formulation  induces more variables, but  results in better lower bounds (LP-relaxation) compared to \cite{Peschiera2019}.
+   Generic  primal heuristics implemented by MILP solvers  are less efficient with this new formulation, it is explained with the increasing number of variables.
+To improve both primal and dual bounds of \cite{Peschiera2019}, perspectives are to implement specific matheuristics designed for the extended MILP formulation, using also Machine Learning predictions to guide the search of primal solutions.
+
+%     and lends itself to pattern prediction forecasting models, which can be used to reduce the size of the model.
 
@@
 
-\bibliographystyle{plain}
-% \selectlanguage{english}
-\bibliography{./../biblio/MFMP,./../biblio/FrancoPeschiera}
+% \bibliographystyle{plain}
+% % \selectlanguage{english}
+% %\bibliography{./../biblio/MFMP,./../biblio/FrancoPeschiera}
+% \bibliography{MFMP,FrancoPeschiera}
+\small
+\begin{thebibliography}{1}
+
+\bibitem{Cho2011}
+P.~Cho.
+\newblock {\em Optimal Scheduling of Fighter Aircraft Maintenance}.
+\newblock PhD thesis, MIT, 2011.
+
+\bibitem{Kozanidis2008}
+G.~Kozanidis.
+\newblock {A Multiobjective Model for Maximizing Fleet Availability under the
+  Presence of Flight and Maintenance Requirements}.
+\newblock {\em Journal of Advanced Transportation}, 43(2):155--182, 2008.
+
+\bibitem{Peschiera2019}
+F. Peschiera, A. Ha{\"i}t, N. Dupin, and O. Batta{\"i}a.
+\newblock Long term planning of
+military aircraft flight and maintenance operations. 
+\newblock Technical report, ISAE-SUPAERO,
+Universit\'e de Toulouse, 2019.
+
+\bibitem{PeschieraR2018}
+F. Peschiera, A. Ha{\"i}t, N. Dupin, and O. Batta{\"i}a.
+\newblock Maintenance planning on french military aircraft operations.
+\newblock In {\em Congr{\`e}s annuel de la ROADEF}, 2018.
+
+\bibitem{Verhoeff2015}
+M.~Verhoeff, W.~J~C Verhagen, and R.~Curran.
+\newblock {Maximizing operational readiness in military aviation by optimizing
+  flight and maintenance planning}.
+\newblock {\em Transportation Research Procedia}, 941--950, 2015.
+
+\end{thebibliography}
 

### 2019-11-23 R/nps_reports/results.Rmd
added more datasets and now we show solving times for all instances.

@@
 
-Just like in the mean quality comparison, we compare among cases where all models found a **integer solution**.
+Just like in the mean quality comparison, we compare among **all** cases.
 
@@
 
-The following graph shows the solving times in seconds for each of the `r comparison_table %>% distinct(instance) %>% nrow` instances were an integer solution was found in all cases.
+The following graph shows the solving times in seconds for each of the `r comparison_table %>% distinct(instance) %>% nrow` instances.
 If we reorder all models from fastest to slowest, we gain more insight on the distribution of solving times.

### 2019-11-23 R/nps_reports/datasets.R

@@ -331,10 +331,31 @@ get_all_2 <- function(){
         correct_old_model
 }
 
+get_determ_4 <- function(){
+    get_generic_compare(c('IT000125_20190828', 'IT000125_20190917', 'IT000125_20191030', 'IT000125_20191105'),
+                        exp_names = list('base', 'stoch', 'old', 'stoch_determ'),
+                        scenario_filter='numparalleltasks_4') %>% 
+        correct_old_model
+}
+
 get_determ_3 <- function(){
     get_generic_compare(c('IT000125_20190801', 'IT000125_20191017', 'IT000125_20190917', 'IT000125_20191030', 'IT000125_20191105'),
                         exp_names = list('base', 'determ', 'stoch', 'old', 'stoch_determ'),
                         scenario_filter='numparalleltasks_3') %>% 
+        correct_old_model
+}
+
+get_determ_2 <- function(){
+    get_generic_compare(c('IT000125_20190729', 'IT000125_20191017', 'IT000125_20190915', 'IT000125_20191030', 'IT000125_20191105'),
+                        exp_names = list('base', 'determ', 'stoch', 'old', 'stoch_determ'),
+                        scenario_filter='numparalleltasks_2') %>% 
+        correct_old_model
+}
+
+get_determ_1 <- function(){
+    get_generic_compare(c('IT000125_20190716', 'IT000125_20191017', 'IT000125_20190915', 'IT000125_20191030'),
+                        exp_names = list('base', 'determ', 'stoch', 'old'),
+                        scenario_filter='numparalleltasks_1') %>% 
         correct_old_model
 }
 

### 2019-11-29 R/nps_reports/results.Rmd
reports. changes in datasets and filtering when not using logarithmic scale.

@@
   quality_degr_all %>% 
-  mutate(dif_perc = if_else(dif_perc<= 0, really_small_number, dif_perc)) %>%
+  mutate(dif_perc = if_else(dif_perc<= 0 & scale!='identity', really_small_number, dif_perc)) %>%
   filter(dif_perc %>% value_filt_tails(c(0, right_tail)))
@@
   quality_perf_all %>% 
-  mutate(dif_perc = if_else(dif_perc<= 0, really_small_number, dif_perc)) %>%
+  mutate(dif_perc = if_else(dif_perc<= 0 & scale!='identity', really_small_number, dif_perc)) %>%
   filter(dif_perc %>% value_filt_tails(c(each_tail, each_tail)))
@@
 errors_all <- get_soft_constraints_2(df) %>% 
-  mutate(dif= if_else(dif<=0, really_small_number, dif))
+  mutate(dif= if_else(dif<=0 & scale!='identity', really_small_number, dif))
 

### 2019-11-29 R/nps_reports/datasets.R

@@ -2,8 +2,8 @@ library(reticulate)
 source('nps_reports/functions.R')
 
 get_python_module <- function(rel_path, name){
-    # use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)
-    use_condaenv('cvenv', conda='c:/Anaconda3/Scripts/conda.exe', required=TRUE)
+    use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)
+    # use_condaenv('cvenv', conda='c:/Anaconda3/Scripts/conda.exe', required=TRUE)
     # c:\Anaconda3\envs\cvenv\Scripts\pip.exe install -r requirements
     py_discover_config()
     sysp = import('sys')
@@ -371,22 +371,39 @@ get_all_fixLP <- function(){
         correct_old_model
 }
 
-get_all_stoch <- function(){
-    get_generic_compare(c('IT000125_20190729', 'IT000125_20190808', 'IT000125_20191030', 'IT000125_20191125'),
-                        exp_names = list('base', 'stoch_agg', 'old', 'old_stoch'),
+get_all_stoch_2 <- function(){
+        get_generic_compare(c('IT000125_20190729', 'IT000125_20190808', 'IT000125_20190915', 'IT000125_20190917','IT000125_20191030', 'IT000125_20191125'),
+                        exp_names = list('base', 'stoch_a1', 'stoch_a1r', 'stoch_a2r', 'old', 'old_stoch_a1'),
                         scenario_filter='numparalleltasks_2') %>% 
         correct_old_model
 }
+
+get_all_stoch_4 <- function(){
+    get_generic_compare(c('IT000125_20190828', 'IT000125_20190812', 'IT000125_20190915', 'IT000125_20190917','IT000125_20191030', 'IT000125_20191125'),
+                        exp_names = list('base', 'stoch_a2', 'stoch_a1r', 'stoch_a2r', 'old', 'old_stoch_a1'),
+                        scenario_filter='numparalleltasks_4') %>% 
+        correct_old_model
+}
+
+get_min_usage_5 <- function(){
+    get_generic_compare(c('IT000125_20191122', 'IT000125_20191124'),
+                        exp_names = list('base', 'old'),
+                        scenario_filter='minusageperiod_5') %>% 
+        correct_old_model
+}
+
 
 correct_old_model <- function(data){
     # Function that discounts a fix number from the objective function in the old model.
[…]

### 2019-12-03 latex/notes/bounds/bounds.tex

@@
   	& TM1^{min}_i = \min_t{\{t \in \mathcal{T}^{M_{Init}}_i\}} \notag\\
-    & TM1^{max}_i = \max_t{\{t \in \mathcal{T}^{M_{Init}}_i\}}   \notag \\
+    & TM1^{max}_i = \min \{ \max_t{\{t \in \mathcal{T}^{M_{Init}}_i\}}, \lfloor Rft^{Init}_i \div U^{min} \rfloor \}   \notag \\
     & TM2^{min}_i = \min_t{\{t \in \mathcal{T}^M_{TM1^{min}_i}\}}\notag \\
-    & TM2^{max}_i = \max_t{\{t \in \mathcal{T}^M_{TM1^{max}_i}\}}\notag %\\
+    & TM2^{max}_i = \min \{ \max_t{\{t \in \mathcal{T}^M_{TM1^{max}_i}\}}, TM1^{max}_i + \lfloor H^{max} \div U^{min} \rfloor \}   \notag \\
   \end{align}
@@
   \begin{align}
-    & I^{1M} = \{i \in \mathcal{I} \mid M^{Acc}S^{min}_{i|\mathcal{T}|} = M^{Acc}S^{max}_{i|\mathcal{T}|} = 1\} \notag \\
-    & I^{2M} = \{i \in \mathcal{I} \mid M^{Acc}S^{min}_{i|\mathcal{T}|} = M^{Acc}S^{max}_{i|\mathcal{T}|} = 2\} \notag
+    & I^{1M} = \{i \in \mathcal{I} \mid M^{Acc}S^{max}_{i|\mathcal{T}|} = 1\} \notag \\
+    & I^{2M} = \{i \in \mathcal{I} \mid M^{Acc}S^{min}_{i|\mathcal{T}|} = 2\} \notag
   \end{align}
@@
         & t \in \mathcal{T}, y \in \mathcal{Y}\notag\\
-    & YM^{Acc}F^{\prime max}_{yt} = \lfloor \frac{YFN_{yt}}{M} \rfloor 
+    & YM^{Acc}F^{\prime max}_{yt} = \lfloor \frac{IR^{Acc}_{yt}}{M} \rfloor 
         & t \in \mathcal{T}, y \in \mathcal{Y}\notag
@@
 We can sum all flight-hour needs of missions of type $y$ until the end of period $t$ ($YH^{Acc}_{yt}$).
-This demand of flight hours can be subtracted this from the initial remaining flight time of the group of aircraft and then divided over the $H^{max}$ flight hours a check gives.
+This demand of flight hours can be subtracted from the initial remaining flight time of the group of aircraft and then divided over the $H^{max}$ flight hours each check provides.
 Thus, we get a lower bound $YM^{Acc}F^{\prime min}_{yt}$ on the number of checks we need to do for those aircraft of type $y$ until period $t$.
@@
 
-  The bounds on accumulated checks by type and period defined in \ref{sec:rangecheckstype} can be aggregated into the whole fleet.
+  The bounds on accumulated checks by type and period defined in section \ref{sec:rangecheckstype} can be aggregated into the whole fleet.
 
   \begin{align}
-    & M^{Acc}S^{b}_{t} = \sum_{y \in \mathcal{Y}} YM^{Acc}S^{\prime\prime b}_{yt}
+    & TM^{Acc}S^{b}_{t} = \sum_{y \in \mathcal{Y}} YM^{Acc}S^{\prime\prime b}_{yt}
         & t \in \mathcal{T}, b \in \{min, max\}  \notag\\
-    & M^{Acc}F^{b}_{t} = \sum_{y \in \mathcal{Y}} YM^{Acc}F^{\prime\prime b}_{yt}
+    & TM^{Acc}F^{b}_{t} = \sum_{y \in \mathcal{Y}} YM^{Acc}F^{\prime\prime b}_{yt}
         & t \in \mathcal{T}, b \in \{min, max\}  \notag
@@
   \begin{align}
-    & M^{Acc}F^{\prime max}_{t} = \lfloor \frac{t}{M} \rfloor \times C^{max}
+    & TM^{Acc}F^{\prime max}_{t} = \lfloor \frac{t}{M} \rfloor \times C^{max}
         & t \in \mathcal{T} \notag

### 2019-12-03 latex/notes/model/model.tex

--- 2019-04-22 latex/notes/model/model.tex
@@
 
-The following model provides a tight MIP formulation that solves the Military Flight and Maintenance Problem. Maintenance operations will be referred as "checks".
+The following model provides a tight MIP formulation that solves the Military Flight and Maintenance Problem.
+Maintenance operations will be referred as "checks".
+Missions are assumed to require a constant amount of flight hours per period and a constant amount of aircraft per period, when active.
+Maintenance capacity is constant overall the planning horizon.
 
@@
             $k \in \mathcal{K}$     &  cluster of aircraft that share the same functionality. \\
-            $n \in \mathcal{N}$     &  maintenance cycles for aircraft. Up to 3. \\
+            $n \in \mathcal{N}$     &  maintenance cycles for aircraft. Between 2 and 3. \\
             $s \in \mathcal{S}$     &  slots for elastic variables. \\
@@
         \begin{tabular}{p{15mm}p{125mm}p{15mm}}
-            $H_j$             & amount of flight time required by mission $j$. & hours. \\
-            $R_j$             & number of aircraft required by mission $j$. & aircraft. \\
-            $MT_j$            & minimum number of consecutive periods an aircraft has to be assigned to mission $j$. & periods. \\
-            $U^{min}$         & default aircraft flight time if it is not assigned to any mission nor in maintenance.& hours \\
+            $H_j$             & amount of flight hours required per period by mission $j$. & hours. \\
+            $R_j$             & number of aircraft required per period by mission $j$. & aircraft. \\
+            $MT^{min}_j$      & minimum number of consecutive periods an aircraft can be assigned to mission $j$. & periods. \\
+            $MT^{max}_j$      & maximum number of consecutive periods an aircraft can be assigned to mission $j$. & periods. \\
+            $U^{min}$         & default aircraft flight hours if it is not assigned to any mission nor in maintenance.& hours \\
             $M$               & check duration in number of periods. & periods. \\
@@
             $C^{max}$         & maximum number of simultaneous checks. & aircraft. \\
+            $E^{min}$         & minimum number of periods between two consecutive checks. & periods. \\
             $E^{max}$         & maximum number of periods between two consecutive checks. & periods. \\
-            $E^{min}$         & minimum number of periods between two consecutive checks. & periods. \\
-            $H^{max}$         & remaining flight time after a check. & hours. \\
+            $H^{max}$         & remaining flight hours for any aircraft after a check. & hours. \\
         \end{tabular}
@@
             $A^{Clust}_{kt}$    & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$. & aircraft. \\
-            $H^{Clust}_{kn}$    & maximum number of mean accumulated flight time for cluster $k$ at maintenance cycle $n$. & hours. \\
-            $Rft^{Init}_i$      & remaining flight time for aircraft $i$ at the start of the planning horizon.  & hours.  \\
-            $Rct^{Init}_i$      & remaining calendar time for aircraft $i$ at the start of the planning horizon.  & periods. \\
+            $H^{Clust}_{kt}$    & maximum number of mean accumulated flight hours for cluster $k$ at maintenance cycle $n$. & hours. \\
+            $Rft^{Init}_i$      & remaining flight hours for aircraft $i$ at the start of the planning horizon.  & hours.  \\
+            $Rct^{Init}_i$      & remaining calendar time until reaching $E^{max}$ for aircraft $i$ at the start of the planning horizon.  & periods. \\
         \end{tabular}
@@
         \begin{tabular}{p{30mm}p{125mm}}
-            $t \in \mathcal{T}^s_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a). \\
+            $t' \in \mathcal{T}^s_t$ & time periods $t' \in \mathcal{T}$ during which a check that ends in $t'$ is active, i.e.
+$t' \in \{ \max{\{1, t - M+1\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}a).
+\\
             % $\mathcal{T}^S_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t},  ..., \min{\{|\mathcal{T}|, t + M-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
-            $t' \in \mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^{min}-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
-            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq |\mathcal{T}| - E^{max} - M$ such that $t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$ (figure \ref{fig:gantt_windows}a). \\
-            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ such that $t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$  (figure \ref{fig:gantt_windows}b).
+            % $t' \in \mathcal{T}^m_t$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ {t}, ..., \min{\{|\mathcal{T}|, t + M + E^{min}-1\}}\}$ (figure \ref{fig:gantt_windows}a). \\
+            $t' \in \mathcal{T}^M_t$ &  time periods $t' \leq \mathcal{T}$ during which a second check can start, given that the first check started in $t$, i.e.
+$t' \in \{ t + M + E^{min}-1 , ...,  t + M + E^{max}-1 \}$ (figure \ref{fig:gantt_windows}a).
 \\
-            $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT_j-1$ and $t \ge t_1 + M$ and $t' < t_2$  (figure \ref{fig:gantt_windows}b).\\
-            $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}c). \\
-            $t \in \mathcal{T}^{M_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}c).
+            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}$ & pairs of time periods $t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}$ when a check can start such that it is active in $t$, i.e.
+$t_1 \in \mathcal{T}^s_{t} \lor t_2 \in \mathcal{T}^s_{t}$  (figure \ref{fig:gantt_windows}b).
 \\
-            $t' \in \mathcal{T}^{MT}_{jt}$ &  time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}d). \\
-            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ such that $t_2 \ge t_1 + MT_j-1$ and $t \in \{t_1, ..., t_2\}$ (figure \ref{fig:gantt_windows}d).
+            $(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}$ & all possible assignments for mission $j$ in between two checks starting at $t_1$ and $t_2$.
+Formally: triplet composed of mission $j \in \mathcal{JI}_i$ and $t, t' \in \mathcal{TJ}_j$ such that $t' \ge t + MT^{min}_j-1$ and $t \ge t_1 + M$ and $t' < t_2$  (figure \ref{fig:gantt_windows}b).\\
+            % $t \in \mathcal{T}^{m_{Init}}_i$ &  time periods $t \in \mathcal{T}$ such that $t \in \{ 1, ..., \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}}\}$ (figure \ref{fig:gantt_windows}c). \\
+            $t \in \mathcal{T}^{M_{Init}}_i$ & possible time periods $t \in \mathcal{T}$ during which aircraft $i$ can have its first check, i.e.
+$t \in \{ \max{\{0, Rct^{Init}_i - E^{max} + E^{min} \}} , ...,  Rct^{Init}_i \}$ (figure \ref{fig:gantt_windows}c).
+\\
+            % $t' \in \mathcal{T}^{MT}_{jt}$ & time periods $t' \in \mathcal{T}$ such that $t' \in \{ \max{\{1, t - MT^{min}_j\}},  ..., {t}\}$ (figure \ref{fig:gantt_windows}d). \\
+            $(t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}$ & pairs of time periods $t_1, t_2 \in \mathcal{TJ}_j$ when a mission $j$ active during $t$ can be assigned, i.e.
+$t_2 \ge t_1 + MT^{min}_j-1$ and $t \in \{t_1, ..., t_2\}$ (figure \ref{fig:gantt_windows}d).
 \\
@@
         \end{tikzpicture}
-        \begin{tikzpicture}
-            \input{gantts/time_windows_mission.tex}
-        \end{tikzpicture}
+        % \begin{tikzpicture}
+        %     \input{gantts/time_windows_mission.tex}
+        % \end{tikzpicture}
         \caption{Examples showing the maintenance-related time-parametric sets for aircraft $i$.
@@
     The following parameters are derived from the basic input parameters.
-They are used in constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_rem_hours3}).
+They are used in constraints (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}).
 They are transformations of the default and mission flight hours consumption under different time circumstances.
@@
         & U^{\prime}_{t_1t_2} = U^{min} (t_2 - t_1+1) \notag \\
-        & U^{\prime\prime}_{t_1t_2} = U^{min} \sum_{t=t_1-1}^{t_2} (t_2 - t) = U^{min} (t_2 - t_1+1) (t_2 - t_1) \div 2  \notag \\
+        % & U^{\prime\prime}_{t_1t_2} = U^{min} \sum_{t=t_1-1}^{t_2} (t_2 - t) = U^{min} (t_2 - t_1+1) (t_2 - t_1) \div 2  \notag \\
         & H^\prime_{jt_1t_2} = (H_j - U^{min}) (t_2 - t_1+1) \notag \\
@@
         $a_{ijtt'}$ &  =1 if aircraft $i$ starts a new assignment to mission $j$ at the beginning of period $t$ and finishes it at the end of period $t'$. \\  
-        $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check at the beginning of period $t \in \mathcal{T}$ and then starts the next one at the beginning of period $t'$, 0 otherwise. \\
+        $m_{itt'}$  &  =1 if aircraft $i \in I$ starts a check at the beginning of period $t \in \mathcal{T}^{M_{Init}}_i$ and then starts the next one at the beginning of period $t' \in \mathcal{T}^M_{t}$, 0 otherwise.
+\\
     \end{tabular}
@@
     \begin{tabular}{p{8mm}p{147mm}}
-        $h_{in}$  &  Continuous, positive. Sum of the accumulated monthly consumption in flight hours for aircraft $i \in I$ for maintenance cycle $n$. \\
+        $rft_{it}$  &  remaining flight time (continuous) for aircraft $i \in I$ at the end of period $t \in \mathcal{T}$. \\  
         $e^{A}_{kts}$ & Continuous, positive. Elastic variable for serviceability constraint (\ref{eq:serviceability-cluster}). \\
@@
         + \sum_{\substack{
-                k \in \mathcal{K}, \\ n \in \mathcal{N}, \\ s \in \mathcal{S}
+                k \in \mathcal{K}, \\ t \in \mathcal{T}, \\ s \in \mathcal{S}
                 } 
-            } PH_s e^{H}_{kns}
+            } PH_s e^{H}_{kts}
         + \sum_{t \in \mathcal{T}, s \in \mathcal{S}} PC_s e^{C}_{ts}
@@
     \begin{align}
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H'_{jtt'} + U^{\prime}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
-                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours1}\\
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H'_{jtt'} + U^{\prime}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
-                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours2}\\
-        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H'_{jtt'} + U^{\prime}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
-                & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_hours3}
+        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{1t_1} \leq Rft^{Init}_i + H^{max} (1 - m_{it_1t_2}) 
+                & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \label{eq:cycle_hours1}\\
+        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_1t_2} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+                & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \label{eq:cycle_hours2}\\
+        & \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{t_2|\mathcal{T}|} \leq H^{max} + H^{max} (1 - m_{it_1t_2}) 
+                & i \in \mathcal{I}, t_1 \in \mathcal{T}^{M_{Init}}_i, t_2 \in \mathcal{T}^M_{t_1} \label{eq:cycle_hours3}
     \end{align}
@@
     \begin{align}
-        & h_{i1} (t_1 - 1) \geq (t_1 - 1) H^{max} (m_{it_1t_2} -1) + \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{i1t_1}}} a_{ijtt'} H''_{jtt't_1} + U^{\prime\prime}_{1t_1} 
-            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours1}\\
-        & h_{i2} (t_2 - t_1) \geq (t_2 - t_1) H^{max} (m_{it_1t_2} -1) + \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_1t_2}}} a_{ijtt'} H''_{jtt't_2} + U^{\prime\prime}_{t_1t_2}
-            &t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours2}\\
-        & h_{i3} (|\mathcal{T}|- t_2) \geq (|\mathcal{T}|- t_2) H^{max} (m_{it_1t_2} -1) + \sum_{\substack{(j, t, t') \in \\ \mathcal{J}\mathcal{T}\mathcal{T}_{it_2|\mathcal{T}|}}} a_{ijtt'} H''_{jtt'|\mathcal{T}|} + U^{\prime\prime}_{t_2|\mathcal{T}|}
-            & t_1 \in \mathcal{T}, t_2 \in \mathcal{T}^M_{t_1}, i \in \mathcal{I}\label{eq:cycle_rem_hours3}
-    \end{align}
-
-    \begin{align}
        & \sum_{\substack{i \in \mathcal{IK}_k, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} + N^{Clust}_{kt} \leq A^{Clust}_{kt} + \sum_{s \in \mathcal{S}} e^{A}_{kts}
-        &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}\\
-       & \sum_{i \in \mathcal{IK}_k} h_{in} \leq |{IK}_k| H^{Clust}_{kn} + \sum_{s \in \mathcal{S}} e^{H}_{kns}
-        &k \in \mathcal{K}, n \in \mathcal{N} \label{eq:sustainability-cluster}
+            & k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}\\
+       & \sum_{i \in \mathcal{IK}_k} rft_{it} \geq H^{Clust}_{kt} + \sum_{s \in \mathcal{S}} e^{H}_{kts}
+            & k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
     \end{align}
@@
 A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission.
-    Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight time for each cluster $k$.
+    Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight hours for each cluster $k$.
 
-    Constraints (\ref{eq:cycle_rem_hours1}) - (\ref{eq:cycle_rem_hours3}) define the sum of remaining flight hours each aircraft has in between checks.
-Similarly to (\ref{eq:cycle_hours1}) - (\ref{eq:cycle_hours3}), their difference lies in the three possible intervals produced by assigning at most two checks.
-The derived parameters used in these constraints can be found in section \ref{subsubsec:consumption}.
+    \begin{align}
+        % remaining flight time
+         & rft_{it} \leq rft_{i(t-1)} + H^M \sum_{\substack{(t_1, t_2) \in \\ \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2} - U^{min} - \sum_{\substack{j \in \mathcal{J}_t \cap \mathcal{O}_i, \\ (t_1, t_2) \in \mathcal{T}\mathcal{T}\mathcal{J}_{jt}}} a_{ijt_1t_2} (H_j - U^{min})
+            & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rft_upper}\\
+        & rft_{i0} = Rft^{Init}_i
+               & i \in \mathcal{I} \label{eq:rft_initial}\\
+        & rft_{it} \geq H^M \sum_{\substack{(t_1, t_2) \in \\ \mathcal{T}\mathcal{T}\mathcal{T}_{t}}} m_{it_1t_2}
+                & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:rft_lower}\\ 
+        & rft_{it} \in [0,H^M]
+                & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
+    \end{align}
+
+    Constraints (\ref{eq:rft_upper}) - (\ref{eq:mu}) define the remaining flight time in each aircraft based on the planned missions and maintenances.
 

### 2019-12-05 R/nps_reports/results.Rmd
changes to dataset reading in R. Generalized so less code.

@@
   scale: log2
+  num_tasks: 2
 editor_options:
@@
 size <- params$size
-df <- do.call(size, args=list())
+df <- do.call(size, args=params)
 value_filt_tails <- function(value, each_tail) value %>% between(., quantile(., c(each_tail[1])), quantile(., c(1-each_tail[2])))

### 2019-12-05 R/nps_reports/datasets.R

@@ -52,107 +52,19 @@ get_generic_compare <- function(dataset_list, scenario_filter=NULL, exp_names=NU
     df_original %>% semi_join(tab_filter)
 }
 
+get_base <- function(num_tasks){
+    list(
+        'IT000125_20190716',
+        'IT000125_20190729',
+        'IT000125_20190801',
+        'IT000125_20190828'
+    )[[num_tasks]]
+    
+}
+
 get_1_tasks <- function(){
     get_generic_compare(c('IT000125_20190725', 'IT000125_20190716'), 
                         exp_names = list('cuts', 'base'))
-}
-
-get_3_tasks <- function(){
-    compare_sto <- get_compare_sto()
-    exp_list <- c('IT000125_20190730', 'IT000125_20190801')
-    equiv <- list('cuts', 'base')
-    df_original <- compare_sto$get_df_comparison(exp_list)
-    df_original %>% 
-        filter(scenario=='numparalleltasks_3') %>% 
-        mutate(experiment=equiv[experiment+1] %>% unlist)
-}
-
-get_3_tasks_perc_add <- function(){
-    compare_sto <- get_compare_sto()
-    exp_list <- c('IT000125_20190801', 'IT000125_20190913')
-    equiv <- list('base', 'cuts')
-    df_original <- compare_sto$get_df_comparison(exp_list)
-    df_original %>% 
-        filter(scenario=='numparalleltasks_3') %>% 
-        mutate(experiment=equiv[experiment+1] %>% unlist) %>% 
-        filter_all_exps
-}
-
-get_2_tasks <- function(){
-    compare_sto <- get_compare_sto()
-    exp_list <- c('IT000125_20190730', 'IT000125_20190729')
-    equiv <- list('cuts', 'base')
-    df_original <- compare_sto$get_df_comparison(exp_list)
-    df_original %>% 
-        filter(scenario=='numparalleltasks_2') %>% 
-        mutate(experiment=equiv[experiment+1] %>% unlist)
-}
-
-get_2_tasks_perc_add <- function(){
-    compare_sto <- get_compare_sto()
-    exp_list <- c('IT000125_20190729', 'IT000125_20190913')
-    equiv <- list('base', 'cuts')
-    df_original <- compare_sto$get_df_comparison(exp_list)
-    df_original %>% 
-        filter(scenario=='numparalleltasks_2') %>% 
-        mutate(experiment=equiv[experiment+1] %>% unlist) %>% 
-        filter_all_exps
-}
-
[…]

### 2019-12-05 R/nps_reports/results.Rmd
modified report code in R.

--- 2019-11-29 R/nps_reports/results.Rmd
@@
 author: "Franco Peschiera"
-date: "August 8, 2019"
+date: "`r Sys.Date()`"
 output:
@@
 size <- params$size
-df <- do.call(size, args=list())
-value_filt_tails <- function(value, each_tail) value %>% between(., quantile(., c(each_tail[1])), quantile(., c(1-each_tail[2])))
+df <- do.call(size, args=params)
 

### 2019-12-05 R/nps_reports/datasets.R

@@ -105,7 +105,7 @@ get_determ <- function(num_tasks, ...){
         correct_old_model
 }
 
-get_all_fixLP <- function(){
+get_all_fixLP <- function(...){
     get_generic_compare(c('IT000125_20190729', 'IT000125_20191104', 'IT000125_20191023', 'IT000125_20191025', 'IT000125_20190917'),
                         exp_names = list('base', 'FlexLP_3', 'fixLP', 'FlexLP', 'stoch'),
                         scenario_filter='numparalleltasks_2') %>% 
@@ -128,8 +128,9 @@ get_min_usage_5 <- function(){
         correct_old_model
 }
 
+correct_fun <- function(value, variable) if_else(variable %>% is.na, value, value-variable)
 
-correct_old_model <- function(data){
+correct_old_model <- function(data, get_progress=FALSE, keep_correction=FALSE){
     # Function that discounts a fix number from the objective function in the old model.
     # So it can be compared with the new one.
     # The fixed number depends on the size of the problem, the time horizon.
@@ -142,11 +143,17 @@ correct_old_model <- function(data){
         mutate(correction_value = 2*15*num_tasks*horizon) %>% 
         select(dataset, scenario, correction_value)
     
-    correct_fun <- function(value, variable) if_else(variable %>% is.na, value, value-variable)
-    
-    data %>% 
+    data_n <- data %>% 
         left_join(correction) %>%
         mutate(best_solution= best_solution %>% correct_fun(correction_value),
-               best_bound= best_bound %>% correct_fun(correction_value)) %>% 
-        select(-correction_value)
+               best_bound= best_bound %>% correct_fun(correction_value))
+    
+    if (get_progress){
+        data_n %<>% 
+            mutate(first_relaxed= first_relaxed %>% correct_fun(correction_value))
+    }
+    if (keep_correction){
+        return(data_n)
+    }
+    data_n %>% select(-correction_value)
 }

### 2019-12-05 R/nps_reports/export_results.R

@@ -11,12 +11,25 @@ path_export_tab <- '../../NPS2019/tab/'
 # exp_list <- c('IT000125_20190808', 'IT000125_20190812')
 exp_list <- c('IT000125_20190725', 'IT000125_20190716') # small
 df <- get_generic_compare(exp_list, exp_names = list('cuts', 'base'))
+
+df_fixed <- get_all_fixLP()
+
+raw_df_progress <- get_generic_compare(dataset_list = c(get_base(2), 'IT000125_20191030'),
+                                   exp_names = list('base', 'old'),
+                                   scenario_filter='numparalleltasks_%s' %>% sprintf(2),
+                                   get_progress = TRUE) %>%
+    correct_old_model(get_progress = TRUE, keep_correction = TRUE)
 # df_original <- compare_sto$get_df_comparison(exp_list)
 # df <- df_original %>% 
 #     mutate(experiment=if_else(experiment==0, 'cuts', 'base')) %>% 
 #     filter_all_exps
 
 # df <- get_4_tasks_perc_add()
+
+# compare models
+
+stats_info <- get_stats_summary(raw_df_progress) %>% bind_rows(get_summary(raw_df_progress))
+
 
 # summary
 summary_stats <- get_summary(df)
@@ -69,14 +82,30 @@ ggplot(data=comparison_table_reorder, aes(x=percentage, y=time, color=experiment
     theme_minimal() + geom_point(size=0.5) + ggplot2::xlab('instance percentage') + 
     theme(text = element_text(size=20)) + ggsave(path)
 
+comparison_table_reorder_fixLP <- get_time_perf_integer_reorder(df_fixed)
+path <- '%stime_performance_ordered_fixLP.png' %>% sprintf(path_export_img)
+ggplot(data=comparison_table_reorder, aes(x=percentage, y=time, color=experiment)) + theme_minimal() + 
+    geom_point(size=0.5) + ggplot2::xlab('instance percentage') + theme(text = element_text(size=20)) + 
+    guides(color = guide_legend(override.aes = list(size=5))) + ggsave(path)
+
 # infeasible
 infeasible_stats <- get_infeasible_stats(df)
-# infeasible_times <- get_infeasible_times(df)
 
 path <- '%sinfeas_%s_%s.tex' %>% sprintf(path_export_tab, exp_list[1], exp_list[2])
 infeasible_stats %>% 
[…]

### 2019-12-05 R/nps_reports/functions.R

@@ -5,6 +5,7 @@ library(ggplot2)
 library(data.table)
 library(stringr)
 
+value_filt_tails <- function(value, each_tail) value %>% between(., quantile(., c(each_tail[1])), quantile(., c(1-each_tail[2])))
 
 filter_all_exps <- function(table){
     num <- table %>% distinct(experiment) %>% nrow
@@ -200,6 +201,25 @@ get_variances <- function(raw_df){
         get_all_integer %>% 
         compare_objectives_perc('variance') %>% 
         filter(dif_perc %>% is.na %>% not)
+}
+
+get_stats_summary <- function(raw_df_progress){
+    raw_df_progress %>% 
+        get_stats %>% 
+        group_by(scenario, experiment) %>% 
+        summarise(nodes_mean = mean(nodes), 
+                  nodes_medi = median(nodes),
+                  frel_medi = median(first_relaxed),
+                  crel_medi = median(cuts_best_bound)
+                  ) %>% 
+        aux_compare
+}
+
+get_stats <- function(raw_df_progress){
+    raw_df_progress %>% 
+        get_all_optimal %>% 
+        mutate(cuts_best_bound= cut_info %>% lapply('[[', 'best_bound')
+               %>% as.numeric %>% correct_fun(correction_value))    
 }
 
 get_mega_summary <- function(df){

### 2019-12-05 R/nps_reports/render_report.R

@@ -1,17 +1,20 @@
 library(rmarkdown)
 # size <- 'medium'
 # options <- c('get_4_tasks_aggressive', 'get_1_tasks_maints', 'get_4_tasks', 'get_1_tasks')
-options <- list(list('size'='get_all_stoch', 'num_tasks'=3, 'scale'='identity'), 
-             list('size'='get_all_stoch', 'num_tasks'=2, 'scale'='identity'),
-             list('size'='get_all_stoch', 'num_tasks'=4, 'scale'='identity'))
+options <- list(
+    list('size'='get_all_fixLP', 'num_tasks'=2, 'scale'='log2')
+    # ,list('size'='get_all_stoch', 'num_tasks'=3, 'scale'='identity')
+    # ,list('size'='get_all_stoch', 'num_tasks'=2, 'scale'='identity')
+    # ,list('size'='get_all_stoch', 'num_tasks'=4, 'scale'='identity')
+             )
 # 'scale' = 'identity' # 'identity' or 'log2'
 # options <- paste0('get_determ_', c(1, 2, 3, 4))
 format <- 'pdf'
 format <- 'html'
 for (option in options){
     rmarkdown::render("nps_reports/results.Rmd", 
-                      output_file = sprintf('nps_reports/NPS_report_%s_%s.%s', 
-                                            option$size, option$num_tasks, format),
+                      output_file = sprintf('nps_reports/NPS_report_%s_%s_%s.%s', 
+                                            option$size, option$num_tasks, option$scale, format),
                       params = option, 
                       output_options = list(self_contained = TRUE),
                       output_format=sprintf('%s_document', format),

### 2020-02-14 R/presentations/roadef202002/presentation.Rmd
ROADEF2020 presentation. v1.

---
title: "An alternative MIP formulation for the Military Flight and Maintenance Planning problem"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
    | \tiny (franco.peschiera@isae-supaero.fr)
date: "February, 2020"
output:
  xaringan::moon_reader:
    css: ["footer-header.css", "default"]
    self_contained: false
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    includes:
        in_header: header.tex
editor_options: 
  chunk_output_type: console
---

layout: true

<div class="my-header"></div>

<div class="my-footer"><span>ISAE-SUPAERO, Université de Toulouse, France    
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
ROADEF2020</span></div>

---

### Structure of talk

1. Problem.
1. State of the art.
1. MIP formulations.
2. Machine Learning.
2. Experiments.
3. Conclusions.

---

### Problem (informally)

Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs. **Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours. **Maintenances operations** have specific rules that govern their frequency and capacity.

--

.center[
```{r, echo=FALSE, out.width='100%', warning=FALSE}
library(RefManageR)
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
knitr::include_graphics('calendar3.png')

BibOptions(check.entries = FALSE,
           bib.style = "authoryear",
           cite.style = "alphabetic",
           style = "markdown",
           hyperlink = FALSE,
           dashed = FALSE)
myBib <- ReadBib("./MFMP.bib", check = FALSE)

```
]

---

### Problem

* A series of $j \in \mathcal{J}$ missions are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all missions are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. 
* Each mission requires a certain number $r_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $h_j$ hours in each period. An aircraft assigned to a mission $j$ must be assigned for at least $MT^{min}_{j}$ and at most $MT^{max}_{j}$ consecutive periods.
* The heterogeneous fleet requires recurrent preventive maintenance operations (checks) as a result of the total flight hours since the last check. There are also min (max) distances between maintenances.
* A check takes exactly $m$ periods and cannot be interrupted. There is a limited capacity for maintenances at each period.
* The objective is to minimize the total number of checks while maximizing the status of the fleet at the end of the horizon.

--

**Other**: soft constraints on the state of the fleet at each period, initial conditions, default consumption, reduce variance of frequency of checks.

---

### State of the art

* FMP: Flight and Maintenance Planning problem.
* In `r Citep(myBib, "Cho2011")`, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
* In `r Citep(myBib, "Kozanidis2008")`, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In `r Citep(myBib, "Verhoeff2015")`, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In `r Citep(myBib, "Seif2018")`, a generalization for different types of maintenances and capacities was done.
* In `r Citep(myBib, "Peschiera2019")`, the problem was proved NP-Complete.

---

### Previous formulation

* $a_{jti}$   :  =1 if mission $j \in J$ in period $t \in \mathcal{T}_j$ is realized with aircraft $i \in \mathcal{I}_j$, 0 otherwise.  
* $a^s_{jti}$ :  =1 if aircraft $i$ starts a new assignment to mission $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-1)i} = 0$.  
* $m_{it}$    :  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise.  
* $u_{it}$    :  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$.

\begin{align}
 & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
 & u_{it} \geq U^{min} (1 - \sum_{t' \in \mathcal{T}^s_t} m_{it'})
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight2}\\
 & rft_{it} \leq rft_{i(t-1)} + H^M m_{it} - u_{it}
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:first_rft_upper}\\
& rft_{it} \in [0,H^M]
      & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:first-mu}
\end{align}

---

### New formulation

* $a_{ijtt'}$ : =1 if aircraft $i$ starts an assignment to mission $j$ at the beginning of period $t$ and finishes at the end of period $t'$, zero otherwise.  
* $m_{ip}$ : =1 if aircraft $i$ uses check pattern $p$, zero otherwise.
  * each pattern $p$ has a single feasible combination of check starts for an aircraft during the whole planning (usually only 1-2 checks per aircraft).

\begin{align}
& \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{max} + H^{max} (1 - m_{ip}) & \notag \\
  & \hspace{200px}  i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \label{eq:cycle_hours2}\\
\end{align}

---

### Formulation

\begin{align}
  & \text{Max}\;
  \sum_{i \in \mathcal{I}, p \in \mathcal{P}} m_{ip} \times W_p \\
  & \sum_{i \in \mathcal{I}, p \in \mathcal{P}_{t}} m_{ip} \leq C^{max} 
          & t \in \mathcal{T} \label{eq:capacity1}\\
  & \sum_{i \in \mathcal{I}_j, (t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \geq R_j
          & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
  & \sum_{p \in \mathcal{P}_{t}} m_{ip} + \sum_{j \in \mathcal{J}_t \cap \mathcal{J}_i} \sum_{(t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \leq 1 
          & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{max} + H^{max} (1 - m_{ip}) & \notag \\
    & & i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \\
\end{align}

---

### Distance between maintenances

.pull-left[
* The distance between maintenance has a maximum of $E^{max}$ periods.
* Depending on the instance, the optimal distance can be shorter.
* This distance conditions the total number of patterns to create.
]

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('hist_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

---

### Forecasting + Optim

**We want to**:

1. Train a statistical model to predict the mean distance between maintenances for any given instance.
2. Use this information to limit all possible combinations of patterns to generate.

--

**Benefits**:

1. **Performance**: a smaller model is easier to solve.
2. **User feedback**: direct feedback about the solution without needing to solve any model.
3. **More stable solutions**: Every aircraft flies an amount that is closest to the mean of the fleet.

--

**The better we're able to predict the optimal distance between maintenances for the whole fleet, the less optimality we will lose**

---

### Prediction model

.pull-left[
* **Technique**: *Quantile regressions* to estimate upper and lower bounds.
* **Training**: 5000 small instances.
* **Input features**: 
  * mean flight demand per period,
  * total remaining flight hours at start (init), 
  * variance of flight demand, 
  * demand of special missions, 
  * number of period where flight demand is cut in two.
* **Output features**: mean distance between maintenances.
]

--

.pull-right[
```{r, echo=FALSE, out.width='100%'}
imgs = c('QuantReg_mean_consum_upper_mean_dist_complete_IT000125_20190716.png')
knitr::include_graphics(imgs)
```
]

---

### Experiments

* Number of instances: medium (1000), large (1000) and very large (1000).
* Time limit at 3600 seconds.
* We seeded instance generation for better comparison.
* CPLEX running 1 thread.

Largest instances have 60 aircraft, 90 periods, ~30 missions (4 active missions at any given time).

1. Create forecasting model based in 5000 small instances.
2. Use forecasting model to predict bounds on distance between maintenaances: $\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$.
3. Implement the pseudo-cut:

\begin{align}
    & m_{ip} = 0 & p_{t'} - p_t < \hat{\mu}_{t'-t}^{lb} - tol \label{eq:dist_lb} \\
    & m_{ip} = 0  &  p_{t'} - p_t > \hat{\mu}_{t'-t}^{ub} + tol \label{eq:dist_ub}
\end{align}

---

### How good is it (performance)

Faster solutions, more solutions.

--

.center[
```{r, echo=FALSE, out.width='80%'}
imgs = c('time_performance_ordered_2tasks.png')
knitr::include_graphics(imgs)
```
]

---

### How good is it (optimality)

For instances were an optimal solution was found (optimum degradation):
* 95% of instances had less than 4% gap with real optimal.

.center[
```{r, echo=FALSE, out.width='80%'}
imgs = c('quality_degradation_2tasks.png')
knitr::include_graphics(imgs)
```
]

---

## Further steps

* **Predict a distribution** and sample patterns from the distribution instead of predicting patterns.
* **Warm-start Column Generation** with a selected subset of potentially good patterns.
* **Automatize prediction** so it can be easily integrated in other problems.

---

# References

```{r refs, echo=FALSE, results="asis"}
PrintBibliography(myBib)
```

### 2020-02-28 R/presentations/roadef202002/presentation.Rmd

@@
 
+```{r, echo=FALSE, warning=FALSE}
+library(RefManageR)
+knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
+
+BibOptions(check.entries = FALSE,
+           bib.style = "authoryear",
+           cite.style = "alphabetic",
+           style = "markdown",
+           hyperlink = FALSE,
+           dashed = FALSE)
+myBib <- ReadBib("./MFMP.bib", check = FALSE)
+
+```
+
 layout: true
@@
 **Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours.
-**Maintenances operations** have specific rules that govern their frequency and capacity.
+**Maintenance operations** have specific rules that govern their frequency and capacity.
 
@@
 ```{r, echo=FALSE, out.width='100%', warning=FALSE}
-library(RefManageR)
-knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
 knitr::include_graphics('calendar3.png')
-
-BibOptions(check.entries = FALSE,
-           bib.style = "authoryear",
-           cite.style = "alphabetic",
-           style = "markdown",
-           hyperlink = FALSE,
-           dashed = FALSE)
-myBib <- ReadBib("./MFMP.bib", check = FALSE)
 
@@
 
-* A series of $j \in \mathcal{J}$ missions are planned along a horizon divided into $t \in \mathcal{T}$ periods.
-Since all missions are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized.
-* Each mission requires a certain number $r_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $h_j$ hours in each period.
-An aircraft assigned to a mission $j$ must be assigned for at least $MT^{min}_{j}$ and at most $MT^{max}_{j}$ consecutive periods.
-* The heterogeneous fleet requires recurrent preventive maintenance operations (checks) as a result of the total flight hours since the last check.
-There are also min (max) distances between maintenances.
-* A check takes exactly $m$ periods and cannot be interrupted. There is a limited capacity for maintenances at each period.
-* The objective is to minimize the total number of checks while maximizing the status of the fleet at the end of the horizon.
+* A series of $j \in \mathcal{J}$ pre-scheduled missions are planned along a horizon divided into $t \in \mathcal{T}$ periods.
+* Each mission requires a certain **number** $R_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $H_j$ **hours** in each period.
+* The heterogeneous fleet requires recurrent preventive maintenance operations (**checks**).
+* A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
+* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the end of the horizon.
 
@@
 * $a_{jti}$   :  =1 if mission $j \in J$ in period $t \in \mathcal{T}_j$ is realized with aircraft $i \in \mathcal{I}_j$, 0 otherwise.  
-* $a^s_{jti}$ :  =1 if aircraft $i$ starts a new assignment to mission $j$ in period $t$. If $a_{jti} = 1$ and $a_{j(t-1)i} = 0$.  
 * $m_{it}$    :  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise.  
@@
     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight2}\\
- & rft_{it} \leq rft_{i(t-1)} + H^M m_{it} - u_{it}
+ & rft_{it} \leq rft_{i(t-1)} - u_{it} + H^M m_{it}
     & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:first_rft_upper}\\
@@
 \begin{align}
-& \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{max} + H^{max} (1 - m_{ip}) & \notag \\
+  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{M} + M (1 - m_{ip}) & \notag \\
   & \hspace{200px}  i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \label{eq:cycle_hours2}\\
 \end{align}
+
+```{r, echo=FALSE, out.width='100%', warning=FALSE}
+knitr::include_graphics('calendar3.png')
+
+```
 
@@
           & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
-  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{max} + H^{max} (1 - m_{ip}) & \notag \\
+  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^M + M (1 - m_{ip}) & \notag \\
     & & i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \\
 \end{align}
+
+---
+
+### Is this better?
+
+1. It uses 3 times the number of constraints and 3 times the number of variables.
+  * variables: 11000 => 28000.
+  * constraints: 13000 => 48000.
+1. It is still better. Better lineal relaxation, better performance.
 
@@
 .pull-left[
-* The distance between maintenance has a maximum of $E^{max}$ periods.
+* The distance between maintenance has a maximum of $E^{M}$ periods.
 * Depending on the instance, the optimal distance can be shorter.
@@
 
-### Forecasting + Optim
+### Forecasting + Optimization
 
@@
 
+--
+
 Largest instances have 60 aircraft, 90 periods, ~30 missions (4 active missions at any given time).
@@
 1. Create forecasting model based in 5000 small instances.
-2. Use forecasting model to predict bounds on distance between maintenaances: $\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$.
+2. Use forecasting model to predict bounds on distance between maintenances: $\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$.
 3. Implement the pseudo-cut:
@@
 \end{align}
+
+4. Recycling.
 
@@
 
+* **Better predictions** with better features, or predicting several characteristics of optimal solutions.
 * **Predict a distribution** and sample patterns from the distribution instead of predicting patterns.

### 2020-04-28 latex/notes/graph_heur/heuristic.tex
added document describing graph heuristic.

\documentclass[a4paper,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{longtable}
%opening
\title{Graph-based pattern sampling for solve-repair heuristic}
\author{}

\setlength{\topmargin}{-2.1cm}
\setlength{\textwidth}{15.8cm}
\setlength{\textheight}{26.3cm}
\setlength{\oddsidemargin}{-0.1cm}
\setlength{\evensidemargin}{-0.1cm}

\def\II{{\mathcal{I}}}
\def\TT{{\mathcal{T}}}
\def\KK{{\mathcal{K}}}
\def\JJ{{\mathcal{J}}}

\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{chngcntr}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\tikzset{%
  >={Latex[width=2mm,length=2mm]},
  % Specifications for style of nodes:
            base/.style = {rectangle, rounded corners, draw=black,
                           minimum width=4cm, minimum height=1cm,
                           text centered, font=\sffamily},
  activityStarts/.style = {base, fill=blue!30},
       startstop/.style = {base, fill=red!30},
    subproblem/.style = {base, fill=green!30},
         master/.style = {base, minimum width=2.5cm, fill=orange!15,
                           font=\ttfamily},
}

\counterwithin*{equation}{section}
\counterwithin*{equation}{subsection}

\begin{document}

\maketitle

\section{Introduction}
  \label{sec:model}

  The purpose of this document is to present the implemented idea of a general solve-repair technique using random and efficient pattern sampling at each iteration to solve the Military Flight and Maintenance Problem.

  The first section is a description of the input data. Then the master problem is presented. The pattern definition, representation and sampling follows. Lastly, several alternatives to subproblems are described.

\section{Input data}
    \subsection{Basic sets}

        \begin{tabular}{p{15mm}p{140mm}}
            $i \in \mathcal{I}$     &  aircraft. \\
            $t \in \mathcal{T}$     &  time periods included in the planning horizon. \\
            $j \in \mathcal{J}$     &  missions. \\
            $y \in \mathcal{Y}$     &  types of mission or aircraft. \\
            $k \in \mathcal{K}$     &  clusters of aircraft that share exactly the same functionality. \\
            $p \in \mathcal{P}$     &  all possible patterns to assign to any aircraft. \\
        \end{tabular}

    \subsection{Mission parameters}

        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
            $R_j$             & number of aircraft required by mission $j$. & aircraft. \\
        \end{tabular}

    \subsection{Maintenance parameters}

        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
            $C^{max}$         & maximum number of simultaneous checks. & aircraft. \\
        \end{tabular}

    \subsection{Fleet parameters}

        \begin{tabular}{p{15mm}p{125mm}p{15mm}}
            $N_t$               & number of aircraft known to be in maintenance in period $t$. & aircraft. \\
            $N^{Clust}_{kt}$    & number of aircraft in cluster $k$ known to be in maintenance in period $t$. & aircraft. \\
            $A^{Clust}_{kt}$    & maximum number of aircraft in cluster $k$ that can be simultaneously in maintenance in period $t$. & aircraft. \\
            $H^{Clust}_{kt}$    & minimum number of total remaining flight time for cluster $k$ at period $t$. & hours. \\
            $Y_i$               & type $y \in \mathcal{Y}$ for aircraft $j$. & type. \\
            $Rft_{itp}$           & remaining flight time for aircraft $i \in \mathcal{I}$ at the end of period $t \in \mathcal{T}$ if assigned pattern $p \in \mathcal{PI}_i$. & hours. \\
        \end{tabular}

    \subsection{Parametric sets}

        \begin{tabular}{p{15mm}p{140mm}}
          $t \in \mathcal{TJ}_j$     &  time periods $t \in \mathcal{T}$ in which mission $j$ is active. \\
          $j \in \mathcal{JT}_t$    &  missions $j \in \mathcal{J}$ to be realized in period $t$. \\
          $i \in \mathcal{IK}_k$     &  aircraft $i \in \mathcal{I}$ that are included in cluster $k$. One aircraft can belong to more than one cluster. \\
          $j \in \mathcal{JI}_i$     &  missions $j \in \mathcal{J}$ for which aircraft $i$ can be used. \\
          $i \in \mathcal{A}^{Init}_j$  & aircraft $i \in \mathcal{I}$ that have mission $j$ pre-assigned in the previous period to the start of the planning horizon. \\
          $p \in \mathcal{PI}_i$     &  patterns $p \in \mathcal{P}$ that are available for aircraft $i$. \\
          $p \in \mathcal{PT}^{M}_{it}$     &  patterns $p \in \mathcal{PI}_i$ where aircraft $i$ is in maintenance in period $t \in \mathcal{T}$. \\
          $p \in \mathcal{PT}^{J}_{itj}$     &  patterns $p \in \mathcal{PI}_i$ where aircraft $i i \in \mathcal{IJ}_j$ is assigned to mission $j \in \mathcal{J}$ in period $t \in \mathcal{T}$. \\
        \end{tabular}

    \subsection{Interval deviation and objective function parameters}

       \begin{tabular}{p{15mm}p{105mm}p{15mm}}
            $P\hspace{-0.5mm}A_s$   & penalty cost for violating serviceability constraint in interval $s$. & [$\frac{penalty}{aircraft.period}$] \\
            $P\hspace{-0.5mm}H_s$   & penalty cost for violating sustainability constraint in interval $s$. & [$\frac{penalty}{hour-period}$] \\
            $PC_s$   & penalty cost for violating capacity constraint in interval $s$. & [$\frac{penalty}{aircraft.period}$]\\
            $PM_p$    & reward per period for the start of each check in patter $p$. & [$\frac{penalty}{aircraft.period}$]\\
            $U\hspace{-0.5mm}A_s$   & maximum deviation for violating the serviceability limit in interval $s$. & [aircraft] \\
            $U\hspace{-0.5mm}H_s$   & maximum deviation for violating the sustainability limit in interval $s$. & [hours] \\
            $U\hspace{-0.5mm}C_s$   & maximum deviation for violating the maintenance capacity limit in interval $s$. & [aircraft] \\
        \end{tabular}

\section{General logic and master problem}

  The master problem is used to do the following tasks at each iteration.

  \begin{enumerate}
    \item Find good candidates of time window ($\mathcal{T}_{c}$) and aircrafts ($\mathcal{I}_{c}$) to repair.
    \item Call the subproblem to obtain a new solution.
    \item Decide whether to accept or not the new solution.
  \end{enumerate}

  The initial solution ($c=0$) consists of calling the subproblem with a time window the size of the planning horizon ($\mathcal{T}_{0} = \mathcal{T}$) and the complete set of aircraft ($\mathcal{I}_{0} = \mathcal{I}$).

  \begin{tikzpicture}[node distance=1.5cm,
    every node/.style={fill=white, font=\sffamily}, align=center]
  % Specification of nodes (position, etc.)
    \node (start)             [activityStarts]              {Start};
    \node (initSol)     [startstop, below of=start]          {Initial solution};
    \node (objFunc)      [master, below of=initSol]   {Get objective};
    \node (bestSol)      [master, below of=objFunc]   {Update best solution};
    \node (prevSol)      [master, below of=bestSol]   {Update previous solution};
    \node (getErrors)      [master, below of=prevSol]   {Get errors};
    \node (getCandidate)     [master, below of=getErrors]   {Find good window candidate};
    \node (samplePatterns)      [subproblem, below of=getCandidate]
                                                        {Sample patterns};
    \node (decidePatterns)      [subproblem, below of=samplePatterns, yshift=-1cm]
                                                                  {Decide patterns};
    \node (applyPatterns)       [subproblem, below of=decidePatterns, yshift=-1cm]
                                                                   {Apply patterns};
    % \node (ActivityDestroyed) [startstop, below of=onDestroyBlock]
    %                                                   {Activity is shut down};     
    % Specification of lines between nodes specified above
    % with aditional nodes for description 
    \draw[->]             (start) -- (initSol);
    \draw[->]     (initSol) -- (objFunc);
    \draw[->]     (objFunc) -- (bestSol);
    \draw[->]      (bestSol) -- (prevSol);
    \draw[->]      (prevSol) -- (getErrors);
    \draw[->]      (getErrors) -- (getCandidate);
    \draw[->]     (getCandidate) -- (samplePatterns);
    \draw[->]      (samplePatterns) -- (decidePatterns);
    \draw[->]      (decidePatterns) -- (applyPatterns);
    % \draw[->]       (applyPatterns) -- node {The activity is shut down by
    %                                  user or system} (onDestroyBlock);
    % \draw[->]    (onDestroyBlock) -- (ActivityDestroyed);
    % \draw[->]      (decidePatterns) -| node(priorityXMemory)
    %                                  {higher priority $\rightarrow$ more memory}
    %                                  (ActivityEnds);
    % \draw           (applyPatterns) -| (priorityXMemory);
    % \draw[->]     (ActivityEnds)  |- node [yshift=-2cm, text width=3.1cm]
                                      % {User navigates back to the activity}
                                      % (initSol);
    % \draw[->]      (applyPatterns.east) -- (objFunc.east);
  \draw[->] (applyPatterns.east) -- ++(2.6,0) -- ++(0,2) -- ++(0,2) --                
     (objFunc.east);
  \end{tikzpicture}

\section{Pattern generation}

  \subsection{Graph representation}

    A Directed Acyclic Graph (DAG) is created for each aircraft. Each node in the graph represents the state of the aircraft at a specific period in time. Thus, the node is represented by a unique combination of:

        \begin{tabular}{p{10mm}p{115mm}}
          $rft$    &  remaining flight time. \\
          $rct$    &  remaining calendar time. \\
          $t^s$    &  assignment starting period. \\
          $t^f$    &  assignment ending period. \\
          $y$      &  assignment type (check, mission, empty). \\
          $j$      &  assignment value (in the case of missions, the mission; in the case of checks, 'M'; else it is empty). \\
        \end{tabular}

    Each outbound neighbor of a node represents a state the aircraft can have on the immediately next (future) period. This guarantees that the graph will never have any cycles, since we only consider neighbors that are in the future of the current node and, as a consequence, have not been observed yet.

    The graph in Figure \ref{fig:graph_resource1} represents the set of transitions for an individual aircraft in an horizon of 10 periods. A pattern consists of a single unique path that starts at the source node at the extreme left (named "2017-12" in the Figure) and end at the sink node at the extreme right (named "2018-11" in the Figure).

    \begin{figure}
        \centering
        \includegraphics[height=0.7 \linewidth]{example_graph_1.png}
        \caption{DAG representing the transitions of a single aircraft. Each node represents a combination of an assignment at a given time. Nodes are plotted with time in the horizontal axis and the remaining flight time of the aircraft in the vertical axis. White nodes represent an empty assignment (of duration 1). Gray nodes represent a check (of duration 6). Finally, colored nodes represent a mission assignment (the assignment's duration can vary). A example node (circled in blue) is selected and their neighbors (circled in red) are shown: one inbound to the left and two outbound to the right. The selected node consists of an assignment to mission "O1" between periods "2018-02" and "2018-06".} \label{fig:graph_resource1}
    \end{figure}

  \subsection{Graph creation}

    In order to generate a graph from an aircraft, an exhaustive exploration of all possible states for a given aircraft needs to be done. This exploration starts in a node we will call source, where we apply a depth-first search (DFS), looking at each step for all the immediately consecutive neighbors of the node.

    Algorithm \ref{algo:graph-building} shows the procedure executed to create the graph for every aircraft. The logic of $get\_neighbors$ is described in Function \ref{algo:node-neighbors}. TODO: finish describing the $get\_neighbors$ function.

    \begin{algorithm}
    \DontPrintSemicolon
    \SetKwData{node}{$n$}
    \SetKwData{remaining}{$rem$}
    \SetKwData{graph}{$G$}
    \SetKwData{fix}{$f$}
    \KwData{\\$n$: current node. \\$rem$: list of remaining nodes to visit. \\$f$: fix initial states.}
    \KwResult{\\$G$: graph, represented by a mapping between nodes and a list of their neighbors.}
    \BlankLine
    \Begin{
      $\node \leftarrow get\_source\_node()$\;
      $\remaining \leftarrow [\node]$\;
      \If{$|\fix| > 0$}{
        $\remaining \leftarrow \graph[\node] \leftarrow get\_fixed\_nodes(\fix)$\;
      }
      \While{$|\remaining| > 0$}{
        $\node \leftarrow \remaining.pop()$\;
        \If{$\node \notin \graph$}{
          $\graph[\node] \leftarrow get\_neighbors(\node)$\;
          $\remaining.Add(\graph[\node]) $\;
        }
      }
      $\graph \leftarrow add\_artificial\_end\_nodes(\graph)$\label{algo:artificial}\;
    }
    \caption{Depth-first search to populate states graph}\label{algo:graph-building}
    \end{algorithm}

    \begin{algorithm}
    \DontPrintSemicolon
    \SetKwData{node}{$n$}
    \KwData{\\$n$: current node.}
    \BlankLine
    \Begin{
      \If{$\node.end ==last\_period$}{
        $return [get\_sink\_node()]$\;
      }
      $\node.Get\_maint\_neighbors()$\;
      $\node.Get\_mission\_neighbors()$\;
      $\node.Get\_nothing\_neighbors()$\;
    }
    \caption{Get neighbors from node}\label{algo:node-neighbors}
    \end{algorithm}

  \subsection{Path extraction}

    Given a Graph $G(X,V)$, we want to generate all paths from any node $X_1 \in X$ to node $X_2 \in X$. This can be achieved using a DFS and a maximum length cutoff for the search. In our case the maximum length cutoff is trivial to calculate since it equals the number of periods between the two nodes, given that a valid path never overlaps assignments at the same time.

    We obtain those two nodes in the graph, given an existing feasible solution and the time window $\mathcal{T}_{c}$.
    For the first node $X_1$ we do the following. Let $t_{init}$ be the first period in time window $\mathcal{T}_{c}$, we calculate the state of the aircraft $i \in \mathcal{I}_{c}$ in period $t_{init}$. In the case the assignment is an empty assignment, the state will be directly calculated using $rft=rft_i$, $rct=rct_i$, $t^s=t^f=t_{init}$, $y=2$, $j=''$. In the case the assignment is a check or mission assignment, one needs to find the initial period ($t^\prime_{init}$) and the end period ($t^{\prime\prime}_{init}$) for the assignment update the rest accordingly: $rft=rft_i$, $rct=rct_i$, $t^s=t^\prime_{init}$, $t^f=t^{\prime\prime}_{init}$, $y=1$, $j='1'$ if the assignment is mission '1'.

    For the second node $X_2$ this m

\section{Subproblem}

  The subproblem is a pattern-aircraft assignment problem. 
  In each iteration $c$, it takes as input the following:

  \begin{enumerate}
    \item A range of consecutive dates $\mathcal{T}_{c} \subset \mathcal{T}$ (time window).
    \item A set of aircraft $\mathcal{I}_{c} \subset \mathcal{I}$.
  \end{enumerate}

  The main steps this algorithm does are:

  \begin{enumerate}
    \item For each aircraft $i \in \mathcal{I}_{c}$, sample $N$ feasible patterns. Each pattern starts at the beginning of $\mathcal{T}_{c}$ and ends at the end of $\mathcal{T}_{c}$.
    \item For each aircraft, decide one pattern from those $N$ sampled patterns.
    \item Apply each pattern to each aircraft.
  \end{enumerate}

  Step 2 can be accomplished in more than one way. Section \ref{model} shows the use of a mathematical model, Section \ref{grasp} shows the use of a heuristic solution.

  \subsection{Assignment model}
    \label{model}

    Given a subset of samples for each aircraft, we solve an assignment problem using a mixed-integer lineal programming solver. The model is explained below.

    \subsubsection{Variables}

      The following decision variables control the assignment of missions and checks to aircraft.

      \begin{tabular}{p{8mm}p{147mm}}
          $a_{ip}$ &  =1 if aircraft $i$ uses pattern $p$. \\  
      \end{tabular}

      Auxiliary variables control the status of each aircraft or group of aircraft.

      \begin{tabular}{p{8mm}p{147mm}}
          $e^{A}_{kts}$ & Continuous, positive. Elastic variable for serviceability constraint (\ref{eq:serviceability-cluster}). \\
          $e^{H}_{kts}$ & Continuous, positive. Elastic variable for sustainability constraint (\ref{eq:sustainability-cluster}). \\
          $e^{C}_{ts}$ & Continuous, positive. Elastic variable for capacity constraint (\ref{eq:capacity1}). \\
          $e^{J}_{jts}$ & Continuous, positive. Elastic variable for mission aircraft needs constraint (\ref{eq:missionres}). \\
      \end{tabular}

    \subsubsection{Objective function and constraints}
      Objective (\ref{eq:objective1}) minimizes the deviations on elastic constraints while penalizing patterns with earlier checks.

      \begin{align}
          & \text{Min}\; 
          \sum_{\substack{
                  k \in \mathcal{K}, \\ t \in \mathcal{T}, \\ s \in \mathcal{S}
                  }
              } PA_s e^{A}_{kts} 
          + \sum_{\substack{
                  k \in \mathcal{K}, \\ t \in \mathcal{T}, \\ s \in \mathcal{S}
                  } 
              } PH_s e^{H}_{kts}
          + \sum_{t \in \mathcal{T}, s \in \mathcal{S}} PC_s e^{C}_{ts}
          - \sum_{i \in \mathcal{I}, p \in \mathcal{PI}_i} PM_p a_{ip}
          \label{eq:objective1}
      \end{align}

      The following constraints are used in the model:
      \begin{align}
          % maximum capacity1:
          & \sum_{\substack{i \in \mathcal{I}, \\ p \in \mathcal{PT}^{M}_{it}}} a_{ip} + N_t \leq C^{max} + \sum_{s \in \mathcal{S}} e^{C}_{ts}
                  & t \in \mathcal{T} \label{eq:capacity1}\\
          % min assignments:
          & \sum_{\substack{i \in \mathcal{IJ}_j, \\ p \in \mathcal{PT}^{J}_{itj}}} a_{ip} + \sum_{s \in \mathcal{S}} e^{J}_{jts} \geq R_j
                  & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
      \end{align}

      Maintenance capacity is controlled by (\ref{eq:capacity1}). The aircraft requirements of missions are defined by (\ref{eq:missionres}).

      \begin{align}
         & \sum_{\substack{i \in \mathcal{IK}_k, \\ p \in \mathcal{PT}^{M}_{it}}} a_{ip} + N^{Clust}_{kt} \leq A^{Clust}_{kt} + \sum_{s \in \mathcal{S}} e^{A}_{kts}
              & k \in \mathcal{K}, t \in \mathcal{T} \label{eq:serviceability-cluster}\\
         & \sum_{\substack{i \in \mathcal{IK}_k, \\ p \in \mathcal{PI}_i}} Rft_{itp} a_{ip} \geq H^{Clust}_{kt} - \sum_{s \in \mathcal{S}} e^{H}_{kts}
              & k \in \mathcal{K}, t \in \mathcal{T} \label{eq:sustainability-cluster}
      \end{align}

      Constraints (\ref{eq:serviceability-cluster}) guarantee a minimum serviceability of aircraft for each cluster $k$. A cluster is defined by the largest group of aircraft that is required exclusively for at least one mission. 
      Constraints (\ref{eq:sustainability-cluster}) ensure there is a minimum amount of remaining flight hours for each cluster $k$.

      \begin{align}
          & \sum_{p \in \mathcal{PI}_i} a_{ip} =  1 
            & i \in \mathcal{I}\label{eq:num_maint} \\
      \end{align}

      Constraints (\ref{eq:num_maint}) assure each aircraft will have only one assigned pattern.

  \subsection{GRASP heuristic}
    \label{grasp}

    Given a subset of samples for each aircraft, we iteratively choose one pattern for each aircraft. The probability of each pattern to be included depends on the improvement of the pattern to the objective function in its current status. Algorithm \ref{algo:sample_heur} shows this logic.

    \begin{algorithm}
    \DontPrintSemicolon
    \SetKwData{sol}{$x$}
    \SetKwData{pattern}{$p$}
    \SetKwData{patterns}{$\mathcal{P}$}
    \KwData{\\\sol: current solution. \\\patterns: set of patterns for aircraft $i$ in time window $\mathcal{T}_{c}$. \\\pattern: sampled pattern.}
    \BlankLine
    \Begin{
      \For{$i \in \mathcal{I}_{c}$}{
        $\patterns \leftarrow get\_patterns(i, \mathcal{T}_{c})$\;
        $\pattern \leftarrow sample\_pattern(\sol, \patterns)$\;
        $\sol.Apply(\pattern)$\;
      }
    }
    \caption{An heuristic to sample patterns iteratively}\label{algo:sample_heur}
    \end{algorithm}

  \subsection{Graph heuristic}
    \label{graphHeur}

    This heuristic is a variant of the previous one. Instead of obtaining a previous sample before re-sampling, we obtain directly one pattern per Aircraft. We do this by giving weights to each edge on the graph and using those weights to sample iteratively the resulting pattern for the aircraft.

    Another alternative is to generate random edges and do a shortest\_path.

\end{document}

### 2020-04-30 latex/notes/graph_heur/heuristic.tex
I forgot to add some more content to heuristic graph doc.

@@
 
+\subsection{Pattern}
+
+  We call pattern to a feasible combination of missions assignments and checks for aircraft $i$ between the beginning of period $t$ and the end of period $t^\prime$.
+
 \section{Input data}
+
     \subsection{Basic sets}
@@
     \node (initSol)     [startstop, below of=start]          {Initial solution};
-    \node (objFunc)      [master, below of=initSol]   {Get objective};
+    \node (maybeStop)      [master, below of=initSol]   {Reached time limit \\or max iterations?};
+    \node (Stop)      [activityStarts, right of=maybeStop, xshift=4cm]   {End};
+    \node (objFunc)      [master, below of=maybeStop]   {Get objective};
     \node (bestSol)      [master, below of=objFunc]   {Update best solution};
     \node (prevSol)      [master, below of=bestSol]   {Update previous solution};
-    \node (getErrors)      [master, below of=prevSol]   {Get errors};
-    \node (getCandidate)     [master, below of=getErrors]   {Find good window candidate};
+    \node (getErrors)      [master, below of=prevSol]   {Get errors $e$};
+    \node (getCandidate)     [master, below of=getErrors]   {Find candidate $\mathcal{T}_{c}$, $\mathcal{I}_{c}$};
     \node (samplePatterns)      [subproblem, below of=getCandidate]
                                                         {Sample patterns};
-    \node (decidePatterns)      [subproblem, below of=samplePatterns, yshift=-1cm]
+    \node (decidePatterns)      [subproblem, below of=samplePatterns]
                                                                   {Decide patterns};
-    \node (applyPatterns)       [subproblem, below of=decidePatterns, yshift=-1cm]
+    \node (applyPatterns)       [subproblem, below of=decidePatterns]
                                                                    {Apply patterns};
@@
     \draw[->]             (start) -- (initSol);
-    \draw[->]     (initSol) -- (objFunc);
+    \draw[->]     (initSol) -- (maybeStop);
+    \draw[->]     (maybeStop) -- (Stop);
+    \draw[->]     (maybeStop) -- (objFunc);
     \draw[->]     (objFunc) -- (bestSol);
@@
     % \draw[->]      (applyPatterns.east) -- (objFunc.east);
-  \draw[->] (applyPatterns.east) -- ++(2.6,0) -- ++(0,2) -- ++(0,2) --                
-     (objFunc.east);
+  \draw[->] (applyPatterns.west) -- ++(-2.6,0) -- ++(0,10) -- ++(0,2) --                
+     (maybeStop.west);
   \end{tikzpicture}
@@
 
-    For the second node $X_2$ this m
+    For the second node $X_2$ this process needs to be modified somewhat since we do not want to fix the $rct$ and $rft$ of the end node.
+This implies creating several artificial nodes that have $rft=-1$ and $rct=-1$ and connecting all the other nodes that share $t^s, t^f, y, j$ to these nodes.
+This is done in Line \ref{algo:artificial} of Algorithm \ref{algo:graph-building}.
+Let $t_{end}$ be the last period in time window $\mathcal{T}_{c}$.
+We then calculate the end node like so: $rft=-1$, $rct=-1$, $t^s=t^\prime_{end}$, $t^f=t^{\prime\prime}_{end}$, $y=0$, $j='M'$ for the case the assignment at the end is a check.
+
+    Using the example in figure \ref{fig:graph_resource1}, there exist 5 unique paths that start in the selected node in blue.
+Table \ref{tab:pattern_enumeration} shows all possible patterns for that aircraft that start in that node and finish in the sink.
+
+    \input{table_paths}
+
+  \subsection{Sampling paths}
+
+    The previous procedure can be trivially modified so it renders a maximum number of paths at each time, by stopping the DFS process after a number of paths is returned.
+This, nevertheless, results in a extremely biased sampling of the paths since, by construction, the DFS guarantees that consecutive paths share most of the path.
+In addition to this, by default, most DFS implementations use the same sequence of choices given the same graph.
+This approach is, thus, limited.
+
+    The first question that arises is how to take unbiased samples of paths in a graph, in between two nodes.
+This is possible and can be done in a DAG by first pre-calculating the number of paths that pass through each node in their trip from $X_1$ to $X_2$.
+This last operation can be done in $\mathcal{O}(V)$, where $V$ is the number of edges in between the two nodes.
+Then, we use this information to sample one neighbor at each node when iterating from $X_1$.
+This implies not doing a DFS but starting each path at $X_1$.
+
+    The second question that arises is how to take biased samples of paths in a graph, by giving more probability to paths that are similar to the paths we are looking for (i.e., have nodes that are more useful for the solution).
+This implies having information about the paths we are looking for, which we do from the Master (e.g., the same information we use to determine where to repair: the number of errors in the solution).
+Doing this avoids having to generate too many samples to get a good solution.
 

### 2020-05-13 latex/notes/graph_heur/heuristic.tex
last modifications before passing it to another

@@
 
-\subsection{Pattern}
+\section{State of the art}
 
-  We call pattern to a feasible combination of missions assignments and checks for aircraft $i$ between the beginning of period $t$ and the end of period $t^\prime$.
+  The Flight and Maintenance Planning problem was first presented by \ref{} for the civil aviation, where it has received much attention.
+It is also usually called Tail Assignment Problem or Aircraft Assignment Problem and it consists on assigning already scheduled flight legs to a fleet of aircraft (also called tails) while taking into account a number of constraints such as maintenance operations (checks), leg-aircraft compatibility.
+Methods used to solve this problem are grouped by \ref{Khaled2017} in three categories: column generation techniques, multi-comodity flow formulations and time-space network models.
+Several techniques have been used to solve this problem, including compact MIP models, Column Generation, Constraint Programming and Metaheuristics.
 
+  In its military variant, this problem has been studied less.
+It was first presented in \ref{} and the main differences imply the notion of mission instead of using the flight legs.
+Aircraft are assigned to missions for an interval of time periods.
+Since all aircraft return to base after each use, the location of the aircraft (airport) is not taken into account.
+Regarding maintenances, there is only one maintenance workshop.
+This problem usually is concerned with medium and long term checks (B, C and D).
+
+  % talk about rolling horizon
+  % used to partition the planning horizon into pieces, solve those pieces and then join them into a single solution 
 \section{Input data}
@@
 
-        \begin{tabular}{p{15mm}p{140mm}}
+        \begin{tabular}{p{20mm}p{135mm}}
           $t \in \mathcal{TJ}_j$     &  time periods $t \in \mathcal{T}$ in which mission $j$ is active. \\
@@
 
-\section{General logic and master problem}
+\section{General strategy}
 
-  The master problem is used to do the following tasks at each iteration.
+  We will use an approach inspired on the Rolling Horizon methodology in most large planning problems and the Column Generation algorithms used in the Tail Assignment literature.
 
-  \begin{enumerate}
-    \item Find good candidates of time window ($\mathcal{T}_{c}$) and aircrafts ($\mathcal{I}_{c}$) to repair.
-    \item Call the subproblem to obtain a new solution.
-    \item Decide whether to accept or not the new solution.
-  \end{enumerate}
+  The Rolling Horizon concept will be used to slice any current solution and extract a piece, solve (or repair) that piece independently and then incorporate the solution to the whole solution.
+The slicing and the joining will be guaranteed to be compatible with the original solution.
+More formally, a piece sliced in iteration $c$ is defined as a time window ($\mathcal{T}_{c} \in \mathcal{T}$) and a subset of aircraft ($\mathcal{I}_{c} \in \mathcal{I}$).
+The choice of the piece will be determined partly randomly and partly based on the number of soft constraints violations.
+This will constitute the master problem.
 
-  The initial solution ($c=0$) consists of calling the subproblem with a time window the size of the planning horizon ($\mathcal{T}_{0} = \mathcal{T}$) and the complete set of aircraft ($\mathcal{I}_{0} = \mathcal{I}$).
+  The Column Generation concept will be applied when solving each piece.
+A number of variables (patterns $p \in \mathcal{P}_i$) will be created, each one corresponding to an ordered list of flights and checks that can be assigned to each aircraft $i$.
+These variables will then be used to obtain a solution to that piece by assigning each aircraft $i \in \mathcal{I}_{c}$ a pattern $p$.
+In practice, we will not be generating all possible patterns, because the number would be prohibitive.
+Instead, we will be doing sampling, as explained in Section \ref{sec:sampling}.
+This will constitute the subproblem.
 
+  The initial solution ($c=0$) consists of calling the subproblem with a time window the size of the planning horizon ($\mathcal{T}_{c=0} = \mathcal{T}$) and the complete set of aircraft ($\mathcal{I}_{c=0} = \mathcal{I}$).
+
+  A number of alternative subproblems will be tested, as explained in Section \ref{sec:subproblems}.
+
+  \subsection{Pattern representation}
+
+    We define a pattern as any feasible combination of missions assignments and checks for aircraft $i$ inside the planning horizon.
+For a pattern to be feasible, it should comply with maintenance rules regarding flight hours and calendar periods between visits, the initial state of the aircraft, the periods in which each mission is available, the minimum and maximum durations of assignment to missions and the capacity of the aircraft to do a particular mission.
+
+  \subsection{Decomposition strategy}
+
+  The Algorithm \ref{fig:general-logic} presents the general decomposition logic.
+
+    \begin{figure}
+      \centering
       \begin{tikzpicture}[node distance=1.5cm,
@@
       \node (start)             [activityStarts]              {Start};
-    \node (initSol)     [startstop, below of=start]          {Initial solution};
-    \node (maybeStop)      [master, below of=initSol]   {Reached time limit \\or max iterations?};
+      \node (initSol)     [startstop, below of=start]          {$x^* = x = get\_initial\_sol()$\\$OBJ^* = Inf$};
+      \node (maybeStop)      [master, below of=initSol]   {$time > limit$};
       \node (Stop)      [activityStarts, right of=maybeStop, xshift=4cm]   {End};
-    \node (objFunc)      [master, below of=maybeStop]   {Get objective};
-    \node (bestSol)      [master, below of=objFunc]   {Update best solution};
-    \node (prevSol)      [master, below of=bestSol]   {Update previous solution};
-    \node (getErrors)      [master, below of=prevSol]   {Get errors $e$};
-    \node (getCandidate)     [master, below of=getErrors]   {Find candidate $\mathcal{T}_{c}$, $\mathcal{I}_{c}$};
+      \node (objFunc)      [master, below of=maybeStop, yshift=-0.5cm]   {$e= getErrors(x)$\\$OBJ=get\_objective(x,e)$};
+      % \node (updateSol)      [master, below of=objFunc]   {Update best / previous solution};
+      \node (maybeBest)      [master, below of=objFunc]   {$OBJ < OBJ^*$};
+      \node (updateBest)      [master, right of=maybeBest, xshift=3cm]   {$x^* = x$\\$OBJ^* = OBJ$};
+      \node (getCandidate)     [master, below of=maybeBest, yshift=-0.5cm]   {$\mathcal{T}_{c}, \mathcal{I}_{c} = get\_candidate(e)$};
       \node (samplePatterns)      [subproblem, below of=getCandidate]
-                                                        {Sample patterns};
+                                                          {$\mathcal{P}_i = sample\_patterns(\mathcal{T}_{c}) \,\, \forall i \in \mathcal{I}_{c}$};
       \node (decidePatterns)      [subproblem, below of=samplePatterns]
-                                                                  {Decide patterns};
+                                                                    {$p_i^* = decide\_patterns(\mathcal{P}_i)$};
       \node (applyPatterns)       [subproblem, below of=decidePatterns]
-                                                                   {Apply patterns};
-    % \node (ActivityDestroyed) [startstop, below of=onDestroyBlock]
-    %                                                   {Activity is shut down};     
-    % Specification of lines between nodes specified above
-    % with aditional nodes for description 
+                                                                     {$x = apply\_patterns(p_i^*)$};
+
       \draw[->]             (start) -- (initSol);
       \draw[->]     (initSol) -- (maybeStop);
-    \draw[->]     (maybeStop) -- (Stop);
-    \draw[->]     (maybeStop) -- (objFunc);
-    \draw[->]     (objFunc) -- (bestSol);
-    \draw[->]      (bestSol) -- (prevSol);
-    \draw[->]      (prevSol) -- (getErrors);
-    \draw[->]      (getErrors) -- (getCandidate);
+      \draw[->]     (maybeStop) -- node {Yes}  (Stop);
+      \draw[->]     (maybeStop) -- node {No} (objFunc);
+      \draw[->]     (objFunc) -- (maybeBest);
+      \draw[->]      (maybeBest) -- node {Yes} (updateBest);
+      \draw[->]      (updateBest.south) -- ++(0,-1.5) -- ++(-2,0)  -- (getCandidate.east);
+      \draw[->]      (maybeBest) -- node {No} (getCandidate);
       \draw[->]     (getCandidate) -- (samplePatterns);
@@
       \draw[->]      (decidePatterns) -- (applyPatterns);
-    % \draw[->]       (applyPatterns) -- node {The activity is shut down by
-    %                                  user or system} (onDestroyBlock);
-    % \draw[->]    (onDestroyBlock) -- (ActivityDestroyed);
-    % \draw[->]      (decidePatterns) -| node(priorityXMemory)
-    %                                  {higher priority $\rightarrow$ more memory}
-    %                                  (ActivityEnds);
-    % \draw           (applyPatterns) -| (priorityXMemory);
-    % \draw[->]     (ActivityEnds)  |- node [yshift=-2cm, text width=3.1cm]
-                                      % {User navigates back to the activity}
-                                      % (initSol);
-    % \draw[->]      (applyPatterns.east) -- (objFunc.east);
-  \draw[->] (applyPatterns.west) -- ++(-2.6,0) -- ++(0,10) -- ++(0,2) --                
+      \draw[->] (applyPatterns.west) -- ++(-2.6,0) -- ++(0,8) -- ++(0,2) --                
        (maybeStop.west);
       \end{tikzpicture}
+      \caption{General algorithm. In yellow is the master, in green is the subproblem} \label{fig:general-logic}
+    \end{figure}
 
@@
     Algorithm \ref{algo:graph-building} shows the procedure executed to create the graph for every aircraft. 
-The logic of $get\_neighbors$ is described in Function \ref{algo:node-neighbors}.
-TODO: finish describing the $get\_neighbors$ function.
+    % The logic of $get\_neighbors$ is described in Function \ref{algo:node-neighbors}. TODO: finish describing the $get\_neighbors$ function.
 
@@
 
-    \begin{algorithm}
-    \DontPrintSemicolon
-    \SetKwData{node}{$n$}
-    \KwData{\\$n$: current node.}
-    \BlankLine
-    \Begin{
-      \If{$\node.end ==last\_period$}{
-        $return [get\_sink\_node()]$\;
-      }
-      $\node.Get\_maint\_neighbors()$\;
-      $\node.Get\_mission\_neighbors()$\;
-      $\node.Get\_nothing\_neighbors()$\;
-    }
-    \caption{Get neighbors from node}\label{algo:node-neighbors}
-    \end{algorithm}
+    % \begin{algorithm}
+    % \DontPrintSemicolon
+    % \SetKwData{node}{$n$}
+    % \KwData{\\$n$: current node.}
+    % \BlankLine
+    % \Begin{
+    %   \If{$\node.end ==last\_period$}{
+    %     $return [get\_sink\_node()]$\;
+    %   }
+    %   $\node.Get\_maint\_neighbors()$\;
+    %   $\node.Get\_mission\_neighbors()$\;
+    %   $\node.Get\_nothing\_neighbors()$\;
+    % }
+    % \caption{Get neighbors from node}\label{algo:node-neighbors}
+    % \end{algorithm}
 
@@
   \subsection{Sampling paths}
+    \label{sec:sampling}
 
-    The previous procedure can be trivially modified so it renders a maximum number of paths at each time, by stopping the DFS process after a number of paths is returned.
+    The previous procedure can be easily modified so it renders a maximum number of paths at each time, by stopping the DFS process after a number of paths is returned.
 This, nevertheless, results in a extremely biased sampling of the paths since, by construction, the DFS guarantees that consecutive paths share most of the path.
@@
 \section{Subproblem}
-
-  The subproblem is a pattern-aircraft assignment problem. 
-  In each iteration $c$, it takes as input the following:
-
-  \begin{enumerate}
-    \item A range of consecutive dates $\mathcal{T}_{c} \subset \mathcal{T}$ (time window).
-    \item A set of aircraft $\mathcal{I}_{c} \subset \mathcal{I}$.
-  \end{enumerate}
+  \label{sec:subproblems}
 
@@
 
-      Auxiliary variables control the status of each aircraft or group of aircraft.
+      Auxiliary elastic variables control the status of each aircraft or group of aircraft in order to always obtain a feasible solution from the subproblem.
 

### 2020-06-22 R/graph_reports/report.Rmd
updated and cleaned R and python reports code.

---
title: "Results from graph heuristic"
author: "Franco Peschiera"
date: "6/18/2020"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
setwd('../')
source('graph_reports/export_results.R')

solver <- list(prise_srv3_20200605_good='HEUR', 
               serv_cluster1_20200609_good='HEUR', 
               serv_cluster1_20200609_4='HEUR',
               serv_cluster1_20200616='HEUR',
               serv_cluster1_20200617='HEUR',
               serv_cluster1_20200617_2='HEUR', 
               serv_cluster1_20200617_3='HEUR')

```

This is a summary of the tests done so far. I've only used 5 instances per scenario (too few, maybe?) and a limit of 20 minutes.

## On the instances

They have been generated by the same method as the others. I have changed the possibilities of frequency of maintenances so there are less possibilities of maintenance than before (the window to make the maintenance is smaller, the max frequency is the same). The main reason is that since the original mathematical model assumes only two maintenances and the generated graphs are agnostic to this (1, 2, 3, or more maintenances), I can not permit a frequency of less than half the planning horizon without allowing the possibility of having compatibility issues between the two methods. Example: in a horizon of 90 periods, having a maintenance distance of 40 permits having 3 maintenances, one at the beginning, one in the middle, one at the end. This is impossible in the model but possible as a graph solution and causes problems when moving solutions from one to the other.

## The current heuristic

"short" = shortest path algorithm to generate a new pattern for a single aircraft using the generated graph. Usually weights depend on the current solution and the time window is the whole planning period.   
"classic-mip" = solve the whole mip problem by fixing everything that happens outside a "time window" + "set of aircraft" combination. This uses the OR Spectrum MIP without the pseudo-cuts.

The last version of the algorithm, tested below does the following:

1. Create graphs.
2. Finds an initial solution by adding one aircraft at a time using "short".
3. Iterate until time limit or iteration limit:
    1. Randomly picking one of two: "short" or "classic-mip".
    2. Generate a random candidate:
        3. For "short", this is a random aircraft and the whole planning period.
        4. For "classic-mip" this is 30-60 aircraft and 20-66 periods, stopping at 30s or 2% gap.
    3. Keep the solution or not depending on some simulated annealing logic (which is almost "keep if better").

## Quality of the initial solution

Both solvers and the graph method find a feasible solution relatively fast (fast meaning as soon as the searching starts).
But the initial solution obtained by "short" is quite good compared with the MIPs.

In the Table below, for each instance of each scenario, the percentage difference between the initial solution and the best known one for each instance $(initial- best)/best$. In case of the hybrid methods, they sometimes obtain a solution around 2-5 times (200%- 500%) away from the best known one before solving any model. CPLEX does a bad initial solution.

```{r initial}
exps <- c('serv_cluster1_20200617_3', 'serv_cluster1_20200617', 'serv_cluster1_20200610')
exp_names <- c('graph_50', 'graph_10', 'cplex')
initial <- compare_initial(exps, exp_names, solver=solver)
initial %>% kable(format.args = list(big.mark = "."))
```

## Analyze on pre-processing time impact

For the "pure" models we count the time inside the solver, which comes after the pre-processing and model building. This usually takes very little (several seconds, less than 20).

For the hybrid models, we count from the start: we count the mip model pre-processing (which is the same: we do it only once), we count the mip model building (which we do several times but does not take too long). More importantly, we count the time to build the graphs. This depends on the size of the graphs and can take between 60 and 300 seconds.

The size of the graph depends on the initial data: aircraft that are in bad shape are going to have two maintenances and so will need a bigger graph (with more possibilities). The size can be reduced by rounding down the nodes with respect to the remaining flight time (`rft`) of the aircraft. This is, in effect, reducing the valid space of the solution. Of course, this will help both during the generation (less nodes= less time to build) and the later exploitation of the graph (the shortest path will be faster, the assignment of weights will be faster too). On the other hand, we will rely more on the MIP to improve solutions, because it will obtain some patterns that the graph is unable to reach.

In the Figure, `graph_10` means we rounded all nodes to the nearest 10 in terms of `rft`. `graph_50` is more aggressive in reducing the size of the graph. Also, I took the first minute to improve the scale.

```{r pre_processing, out.width='100%'}
exps <- c('serv_cluster1_20200617_3', 'serv_cluster1_20200617', 'serv_cluster1_20200610')
exp_names <- c('graph_50', 'graph_10', 'cplex')
progress <- get_progress(exps, exp_names, solver=solver)
progress %<>% filter(Time>60)
draw_progress(progress, num_aircraft_filt=60, log_scale_y=FALSE)
```

## Comparison on very large datasets

In large datasets, there is not much difference between the approaches in the solution after 20 minutes.

```{r large}
exps <- c('serv_cluster1_20200616', 'serv_cluster1_20200615')
exp_names <- c('graph', 'cplex')
large <- get_summary_table(exps, exp_names)
large %>% kable(format.args = list(big.mark = "."))
```

And the progress. These results need to be updated with a better version of the graph algorithm that initializes a little faster and obtains better initial solutions.

```{r large_progress, out.width='100%'}
exps <- c('serv_cluster1_20200616', 'serv_cluster1_20200615')
exp_names <- c('graph', 'cplex')
progress <- get_progress(exps, exp_names, solver=solver)
progress %<>% filter(Time>60)
draw_progress(progress, num_aircraft_filt=105, log_scale_y=FALSE)
```

## Using CBC or CPLEX

Apparently there is not a great gain in using CBC or CPLEX for solving the MIP subproblem. This is probably caused by the fact that the MIP problems are relatively small.

```{r cbc}
exps <- c('serv_cluster1_20200617_3', 'serv_cluster1_20200617_2', 'serv_cluster1_20200610')
exp_names <- c('graph_cbc', 'graph_cplex', 'cplex')
get_summary_table(exps, exp_names) %>% kable()
```

And the progress:

```{r cbc_progress, out.width='100%'}
exps <- c('serv_cluster1_20200617_3', 'serv_cluster1_20200617_2', 'serv_cluster1_20200610')
exp_names <- c('graph_cbc', 'graph_cplex', 'cplex')
progress <- get_progress(exps, exp_names, solver=solver)
progress %<>% filter(Time>60)
draw_progress(progress, num_aircraft_filt=75, log_scale_y=FALSE)
```

### 2020-06-22 R/graph_reports/datasets.R

@@ -2,13 +2,14 @@ library(reticulate)
 source('nps_reports/functions.R')
 
 get_python_module <- function(rel_path, name){
-    use_virtualenv('~/Documents/projects/OPTIMA/python/venv/', required = TRUE)
+    path_to_python_dir = '/home/pchtsp/Documents/projects/OPTIMA/python/'
+    use_virtualenv('%svenv/' %>% sprintf(path_to_python_dir), required = TRUE)
     # use_condaenv('cvenv', conda='c:/Anaconda3/Scripts/conda.exe', required=TRUE)
     # c:\Anaconda3\envs\cvenv\Scripts\pip.exe install -r requirements
     py_discover_config()
     sysp = import('sys')
     opt_path = ''
-    python_path <- '%s/%s../python/' %>% sprintf(getwd(), opt_path)
+    python_path <- path_to_python_dir
     sysp$path <- c(python_path, sysp$path)
     scripts_path = paste0(python_path, rel_path)
     compare_sto <- import_from_path(name, path=scripts_path)
@@ -17,7 +18,7 @@ get_python_module <- function(rel_path, name){
 
 get_compare_sto <- function() get_python_module('articles', 'Graphs2020')
 
-get_generic_compare <- function(dataset_list, scenario_filter=NULL, exp_names=NULL, get_progress=FALSE, zip=TRUE){
+get_generic_compare <- function(dataset_list, scenario_filter=NULL, exp_names=NULL, ...){
     compare_sto <- get_compare_sto()
     if (exp_names %>% is.null){
         # This assumes the dataset_list has only two datasets!
@@ -29,7 +30,10 @@ get_generic_compare <- function(dataset_list, scenario_filter=NULL, exp_names=NU
         # so it has to have length 2
         scenario_filter <- c(scenario_filter, 'WORKAROUND')
     }
-    df_original <- compare_sto$compare_experiments(exp_list=dataset_list)
+    # df_original <- compare_sto$compare_experiments(exp_list=dataset_list)
+    df_original <- compare_sto$compare_experiments(exp_list=dataset_list, 
+                                                 scenarios=scenario_filter, 
+                                                 ...)
     dataset_names <- dataset_list %>% lapply(function(x) x)
     df_original %<>% 
         mutate(dataset=dataset_names[experiment+1] %>% unlist,

### 2020-06-22 R/graph_reports/export_results.R

@@ -1,8 +1,124 @@
 source('nps_reports/functions.R')
 source('graph_reports/datasets.R')
 
-stop()
+# stop()
 
-data <- get_generic_compare(c('prise_srv3_20200527_2', 'prise_srv3_20200527'),
-                    exp_names = list('medium', 'small')
-                    )
+get_summary_table <- function(exps, exp_names, wider=TRUE){
+    data <- get_generic_compare(exps, exp_names = exp_names)
+    data_nn <- data %>% 
+        mutate(
+            num_aircraft = scenario %>% str_extract('\\d') %>% as.integer %>% multiply_by(15),
+            objective = objective %>% round
+               ) %>% 
+        select(instance, experiment, num_aircraft, objective) %>% 
+        filter(instance %>% is.na %>% not) %>% 
+        arrange(experiment, num_aircraft, instance)
+    
+    if (!wider){
+        return(data_nn)
+    }
+    data_nn %>% 
+        pivot_wider(id_cols = c(instance, num_aircraft), 
+                    names_from = experiment, 
+                    values_from = objective) %>% 
+        arrange(num_aircraft, instance)
+}
+
+# print(n=40)
+
+
+# PROGRESS:
+# stop()
+
+get_progress <- function(exps, exp_names, solver){
+    progress <- get_generic_compare(exps, exp_names = exp_names, get_progress=TRUE, solver=solver)
+    
+    progress %>% 
+        mutate(num_aircraft = scenario %>% str_extract('\\d') %>% as.integer %>% multiply_by(15)) %>% 
+        select(experiment, num_aircraft, instance, Time, BestInteger) %>% 
+        mutate(BestInteger=as.numeric(BestInteger),
+               Time = as.numeric(Time)) %>% 
+        filter(!is.null(BestInteger))
+}
+
+draw_progress <- function(progress, num_aircraft_filt, count_preprocess=TRUE, log_scale_y=TRUE){
+
+    data_graph <- 
+        progress %>% 
+        filter(num_aircraft==num_aircraft_filt) %>% 
+        mutate(row = sprintf("f=%s i=%s", num_aircraft, instance))
+    
+    # "cheating":
+    if (!count_preprocess){
+        data_graph <- data_graph %>% group_by(experiment, instance) %>% mutate(Time=Time - min(Time))    
+    }
+    
[…]

### 2020-06-22 python/articles/COR2019.py

@@ -1,10 +1,8 @@
-import reports.reports as rep
+from articles import MOSIM2018 as rep
 import os
-import package.experiment as exp
 import pandas as pd
 import numpy as np
 import package.params as pm
-import pytups.superdict as sd
 import dfply as dp
 from dfply import X
 import reports.rpy_graphs as rg

### 2020-06-22 python/articles/Graphs2020.py

@@ -169,7 +169,7 @@ def compare_num_paths():
     #  '9': 6204}
 
 
-def compare_experiments(**kwargs):
+def compare_experiments(exp_list, get_progress=False, **kwargs):
 
     def get_solstats(batch):
         objFunc = batch.\
@@ -179,9 +179,26 @@ def compare_experiments(**kwargs):
             vapply(lambda v: dict(objective=v))
         return batch.format_df(objFunc).drop('name', axis=1)
 
-    return comp.get_df_comparison(**kwargs, get_log=False, solstats_func=get_solstats, zip=True)
+    result = comp.get_df_comparison(exp_list, get_log=True, solstats_func=get_solstats, zip=True,
+                                    get_progress=get_progress, **kwargs)
+    if not get_progress:
+        return result
+
+    def expand_df(index):
+        row = result.iloc[[index]]
+        progress = row.iloc[0].progress.reset_index()[['Time', 'BestInteger']]
+        rows = row.loc[row.index.repeat(len(progress))]
+        rows = rows.drop('progress', axis=1).reset_index()
+        return pd.concat([rows, progress], axis=1)
+
+    tables = [expand_df(i) for i in range(len(result)) if len(result.iloc[i].progress)]
+    return pd.concat(tables, axis=0, join='inner')
 
 
 if __name__ == '__main__':
     # table()
-    compare_experiments(exp_list=['prise_srv3_20200527_2'])
+    t2 = compare_experiments(exp_list=['prise_srv3_20200605_good', 'prise_srv3_20200603_2'],
+                             solver=dict(prise_srv3_20200605_good='HEUR'), get_progress=True)
+    # t = compare_experiments(exp_list=['prise_srv3_20200603_2'], get_progress=True)
+    # [t.iloc[x].progress.tail(10) for x in range(len(t))]
+    # [t2.iloc[x].progress.tail(10) for x in range(len(t2))]

### 2020-06-22 python/package/batch.py

@@ -11,6 +11,7 @@ import os
 import zipfile
 import pandas as pd
 import shutil
+import re
 
 
 class Batch(object):
@@ -77,16 +78,24 @@ class Batch(object):
 
     def get_solver(self):
         opt_info = self.get_options()
+        available_solvers = ['CPLEX', 'GUROBI', 'CBC']
+        default = 'CPLEX'
         try:
             el = list(opt_info.keys())[0]
             engine = opt_info[el]['solver']
-            if '.' in engine:
-                engine, solver = engine.split('.')
-            else:
-                solver = engine
         except:
-            solver = 'CPLEX'
-        return solver
+            return default
+        if '.' in engine:
+            engine, solver = engine.split('.')
+        else:
+            solver = engine
+        if solver in available_solvers:
+            return solver
+        # one last try
+        for s in available_solvers:
+            if re.search(s, solver):
+                return s
+        return default
 
     def get_logs(self, get_progress=False):
         if self.logs is not None:
@@ -153,9 +162,12 @@ class Batch(object):
         table = self.format_df(log_info)
 
         for name in ['matrix', 'presolve', 'matrix_post']:
-            aux_table = table[name].apply(pd.Series)
-            aux_table.columns = [name + '_' + c for c in aux_table.columns]
-            table = pd.concat([table, aux_table], axis=1)
+            try:
+                aux_table = table[name].apply(pd.Series)
+                aux_table.columns = [name + '_' + c for c in aux_table.columns]
+                table = pd.concat([table, aux_table], axis=1)
+            except:
+                pass
 
         return table
 
@@ -300,12 +312,13 @@ class ZipBatch(Batch):
         self.cases = self.get_instances_paths().vapply(load_data)
         return self.cases
 
-    def get_logs(self, get_progress=False):
+    def get_logs(self, get_progress=False, solver=None):
         if self.logs is not None:
             return self.logs
 
         zipobj = zipfile.ZipFile(self.path)
-        solver = self.get_solver()
+        if not solver:
+            solver = self.get_solver()
 
         def _read_zip(x):
             try:
@@ -313,13 +326,20 @@ class ZipBatch(Batch):
             except:
                 return 0
 
+        if solver == 'HEUR':
+            func_to_get_log = get_info_heur
+            filename = '/output.log'
+        else:
+            func_to_get_log = ol.get_info_solver
[…]

### 2020-06-22 python/reports/compare_experiments.py

@@ -30,7 +30,7 @@ def empty_logs(batch):
     return batch.format_df(col)
 
 
-def get_df_comparison(exp_list, scenarios=None, get_progress=False, zip=True, get_log=True, solstats_func=None):
+def get_df_comparison(exp_list, scenarios=None, get_progress=False, zip=True, get_log=True, solstats_func=None, solver=None):
     """
     :param list exp_list: list of names of experiments
     :param list scenarios: optional list of scenarios to filter batches
@@ -46,7 +46,10 @@ def get_df_comparison(exp_list, scenarios=None, get_progress=False, zip=True, ge
         else:
             batch1 = ba.Batch(path=params.PATHS['results'] + experiment, scenarios=scenarios)
         if get_log:
-            table = batch1.get_log_df(get_progress=get_progress)
+            _solver = None
+            if solver:
+                _solver = solver.get(experiment)
+            table = batch1.get_log_df(get_progress=get_progress, solver=_solver)
         else:
             table = empty_logs(batch1)
         table_errors = batch1.get_errors_df().drop('name', axis=1)

### 2020-06-22 python/reports/ggpy_graphs.py

@@ -7,8 +7,7 @@ import orloge as log
 import numpy as np
 import data.dates as aux
 import package.experiment as exp
-import reports.reports as rep
-
+from articles import MOSIM2018 as rep
 
 path_root = PATHS['root']
 path_abs = PATHS['experiments']
@@ -76,10 +75,9 @@ def progress_graph():
     ################################################
     # Progress
     ################################################
-    # TODO: correct this to new logging library
     path = path_abs + '201801131817/results.log'
-    result_log = log.LogFile(path)
-    table = result_log.get_progress_cplex()
+    result_log = log.get_info_solver('CPLEX', path)
+    table = result_log['progress']
     table.rename(columns={'ItCnt': 'it', 'Objective': 'relax', 'BestInteger': 'obj'},
                  inplace=True)
     table = table[table.it.str.match(r'\s*\d')][['it', 'relax', 'obj']]

### 2020-09-04 R/presentations/EDAA202003/presentation.Rmd
changes to EDAA202003 ppt.

--- 2020-02-28 R/presentations/roadef202002/presentation.Rmd
@@
 ---
-title: "An alternative MIP formulation for the Military Flight and Maintenance Planning problem"
+title: "Predicting solutions for the Military Flight and Maintenance Planning problem"
 author: |
@@
     | \tiny (franco.peschiera@isae-supaero.fr)
-date: "February, 2020"
+date: "September, 2020"
 output:
@@
 &emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
-&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
-ROADEF2020</span></div>
+&emsp;&emsp;&emsp;&emsp;&emsp;&emsp;
+EDAA Journée doctorants 2020</span></div>
 
@@
 1. Problem.
-1. State of the art.
-1. MIP formulations.
+2. State of the art.
+1. Solving techniques.
 2. Machine Learning.
@@
 
+### Military aircraft maintenance
+
+.pull-left[
+
+* Mirage 2000 fleet.
+* A single workshop to do the maintenance operations.
+* Check type D (overhaul maintenances): 
+  * every 5 years or 1000 flight hours and 
+  * takes 4-6 months.
+* Fleets of 15-60 aircraft.
+* Planning horizons of 90 periods (= 7.5 years).
+
+]
+
+.pull-right[
+```{r, echo=FALSE, out.width='100%', warning=FALSE}
+knitr::include_graphics('Mirage_2000C.jpg')
+```
+
+*French Mirage 2000C. Public Domain*
+]
+
+---
+
 ### Problem
@@
 * A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
-* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the end of the horizon.
+* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the each period of the horizon.
 
@@
 
-### Previous formulation
+### "Classical" solving techniques
 
-* $a_{jti}$   :  =1 if mission $j \in J$ in period $t \in \mathcal{T}_j$ is realized with aircraft $i \in \mathcal{I}_j$, 0 otherwise.  
-* $m_{it}$    :  =1 if aircraft $i \in I$ starts a check in period $t \in \mathcal{T}$, 0 otherwise.  
-* $u_{it}$    :  flown time (continuous) by aircraft $i \in I$ during period $t \in \mathcal{T}$.
-
-\begin{align}
- & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
-    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight1}\\
- & u_{it} \geq U^{min} (1 - \sum_{t' \in \mathcal{T}^s_t} m_{it'})
-    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:flight2}\\
- & rft_{it} \leq rft_{i(t-1)} - u_{it} + H^M m_{it}
-    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:first_rft_upper}\\
-& rft_{it} \in [0,H^M]
-      & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:first-mu}
-\end{align}
-
----
-
-### New formulation
-
-* $a_{ijtt'}$ : =1 if aircraft $i$ starts an assignment to mission $j$ at the beginning of period $t$ and finishes at the end of period $t'$, zero otherwise.  
-* $m_{ip}$ : =1 if aircraft $i$ uses check pattern $p$, zero otherwise.
-  * each pattern $p$ has a single feasible combination of check starts for an aircraft during the whole planning (usually only 1-2 checks per aircraft).
-
-\begin{align}
-  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^{M} + M (1 - m_{ip}) & \notag \\
-  & \hspace{200px}  i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \label{eq:cycle_hours2}\\
-\end{align}
-
-```{r, echo=FALSE, out.width='100%', warning=FALSE}
-knitr::include_graphics('calendar3.png')
-
-```
-
----
-
-### Formulation
+$a_{ijt_1t_2}$: mission $j$ is assigned to aircraft $i$ between $t_1$ and $t_2$.  
+$m_{it_1t_2}$: aircraft $i$ starts a check in $t_1$ and another check in $t_2$.
 
@@
 
-### Is this better?
-
-1. It uses 3 times the number of constraints and 3 times the number of variables.
-  * variables: 11000 => 28000.
-  * constraints: 13000 => 48000.
-1. It is still better. Better lineal relaxation, better performance.
-
----
-
-### Distance between maintenances
-
-.pull-left[
-* The distance between maintenance has a maximum of $E^{M}$ periods.
-* Depending on the instance, the optimal distance can be shorter.
-* This distance conditions the total number of patterns to create.
-]
-
-.pull-right[
-```{r, echo=FALSE, out.width='100%'}
-imgs = c('hist_mean_dist_complete_IT000125_20190716.png')
-knitr::include_graphics(imgs)
-```
-]
-
----
-
 ### Forecasting + Optimization
@@
 
-1. Train a statistical model to predict the mean distance between maintenances for any given instance.
-2. Use this information to limit all possible combinations of patterns to generate.
+1. Train a statistical model to predict characteristics of good (optimal?) solutions for any given instance.
+2. Use this information to limit the solution space and make it easier to find those potentially good solutions.
 
@@
 
-1. **Performance**: a smaller model is easier to solve.
+1. **Performance**: a smaller problem is easier to solve.
 2. **User feedback**: direct feedback about the solution without needing to solve any model.
@@
 
-**The better we're able to predict the optimal distance between maintenances for the whole fleet, the less optimality we will lose**
+**The better we're able to predict those characteristics for the whole fleet, the less optimality we will lose**
 
@@
 
-.pull-left[
 * **Technique**: *Quantile regressions* to estimate upper and lower bounds.
@@
 * **Output features**: mean distance between maintenances.
-]
 
@@
 
-.pull-right[
-```{r, echo=FALSE, out.width='100%'}
-imgs = c('QuantReg_mean_consum_upper_mean_dist_complete_IT000125_20190716.png')
-knitr::include_graphics(imgs)
+```{r, echo=FALSE, out.width='100%', warning=FALSE}
+knitr::include_graphics('calendar3.png')
 ```
-]
 

### 2020-09-09 R/presentations/EDAA202003/presentation.Rmd

@@
     | \tiny (franco.peschiera@isae-supaero.fr)
-date: "September, 2020"
+date: "March, 2020"
 output:
@@
 1. Problem.
-2. State of the art.
+1. Context.
 1. Solving techniques.
@@
 * A single workshop to do the maintenance operations.
-* Check type D (overhaul maintenances): 
-  * every 5 years or 1000 flight hours and 
-  * takes 4-6 months.
+* Type D (big maintenances): 
+  * every 5 years and 
+  * take 4-6 months.
 * Fleets of 15-60 aircraft.
@@
 
+### Predictive maintenance
+
+.center[
+**Predicting *when* aircraft will need a maintenance**.
+]
+--
+
+* This is not what we do.
+* We make maintenance plans for the whole fleet, taking into account maintenance needs.
+* So, we use this data to make better maintenance plans.
+* Nevertheless, we will predict *something* during this presentation.
+
+---
+
 ### Problem
@@
 * A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
-* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the each period of the horizon.
+* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the end of the horizon.
 
@@
 
-$a_{ijt_1t_2}$: mission $j$ is assigned to aircraft $i$ between $t_1$ and $t_2$.  
-$m_{it_1t_2}$: aircraft $i$ starts a check in $t_1$ and another check in $t_2$.
+good model + good solver = solution.
 
@@
 
+### Distance between maintenances
+
+```{r, echo=FALSE, out.width='100%', warning=FALSE}
+knitr::include_graphics('calendar3.png')
+
+```
+
+.pull-left[
+* The distance between maintenance has a maximum of $E^{M}$ periods.
+* Depending on the instance, the optimal distance can be shorter.
+* This distance conditions the total number of patterns to create.
+
+]
+
+.pull-right[
+```{r, echo=FALSE, out.width='70%'}
+imgs = c('hist_mean_dist_complete_IT000125_20190716.png')
+knitr::include_graphics(imgs)
+```
+]
+
+---
+
 ### Prediction model
 
+.pull-left[
 * **Technique**: *Quantile regressions* to estimate upper and lower bounds.
@@
 * **Output features**: mean distance between maintenances.
+]
 
@@
 
-```{r, echo=FALSE, out.width='100%', warning=FALSE}
-knitr::include_graphics('calendar3.png')
+.pull-right[
+```{r, echo=FALSE, out.width='100%'}
+imgs = c('QuantReg_mean_consum_upper_mean_dist_complete_IT000125_20190716.png')
+knitr::include_graphics(imgs)
 ```
+]
 

### 2020-09-09 R/presentations/EDAA202003/presentation.Rmd
changes to EDAA ppt

--- 2020-09-04 R/presentations/EDAA202003/presentation.Rmd
@@
 2. Machine Learning.
-2. Experiments.
+2. Results.
 3. Conclusions.
@@
 
-### Experiments
-
-* Number of instances: medium (1000), large (1000) and very large (1000).
-* Time limit at 3600 seconds.
-* We seeded instance generation for better comparison.
-* CPLEX running 1 thread.
-
---
-
-Largest instances have 60 aircraft, 90 periods, ~30 missions (4 active missions at any given time).
-
-1. Create forecasting model based in 5000 small instances.
-2. Use forecasting model to predict bounds on distance between maintenances: $\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$.
-3. Implement the pseudo-cut:
-
-\begin{align}
-    & m_{ip} = 0 & p_{t'} - p_t < \hat{\mu}_{t'-t}^{lb} - tol \label{eq:dist_lb} \\
-    & m_{ip} = 0  &  p_{t'} - p_t > \hat{\mu}_{t'-t}^{ub} + tol \label{eq:dist_ub}
-\end{align}
-
-4. Recycling.
-
----
-
 ### How good is it (performance)
@@
 
+## Communications and publications
+
+Conferences:
+
+* **ROADEF2018**: Maintenance planning on French military aircraft operations.
+* **MOSIM2018**: Bi-objective MIP formulation for the optimization of maintenance planning on French military aircraft operations.
+* **CLAIO2018**: A novel MIP formulation for the optimization problem of maintenance planning of military aircraft.
+* **AFIS2018**: Planning aircraft maintenances for the French Airforce.
+* **ROADEF2020**: An alternative MIP formulation for the Military Flight and Maintenance Planning problem.
+
+Journals:
+
+* **OR Spectrum** (published): A novel solution approach with ML-based pseudo-cuts for the Flight and Maintenance Planning problem.
+* **Annals of Operations Research** (submitted): Long term planning of military aircraft flight and maintenance operations
+* **European Journal of Operations Research** (to be submitted): Novel Graph-based matheuristic to solve the Flight and Maintenance Planning problem
+
+Partnerships:
+
+* **Dasault Aviation**: desktop software OPTIMA built, tested, deployed and validated by users.
+
+*Estimated thesis defense: Novembre 2020.*
+
+---
+
 # References

### 2020-09-09 R/presentations/EDAA202003/presentation.Rmd
more changes EDAA.

@@
 author: |
-    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
+    | Franco Peschiera
     | \tiny (franco.peschiera@isae-supaero.fr)
@@
 1. Solving techniques.
-2. Machine Learning.
+2. Machine Learning + Optimization.
 2. Results.
-3. Conclusions.
+3. Future work.
+3. Publications.
 
@@
 
-### Problem (informally)
-
-Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs.
-**Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours.
-**Maintenance operations** have specific rules that govern their frequency and capacity.
-
---
-
-.center[
-```{r, echo=FALSE, out.width='100%', warning=FALSE}
-knitr::include_graphics('calendar3.png')
-
-```
-]
-
----
-
-### Military aircraft maintenance
+### Military aircraft operations
 
@@
 * Mirage 2000 fleet.
+* Fleets of 60-80 aircraft.
 * A single workshop to do the maintenance operations.
@@
   * takes 4-6 months.
-* Fleets of 15-60 aircraft.
 * Planning horizons of 90 periods (= 7.5 years).
+* Missions require:
+  * aircraft.
+  * monthly flight hours.
 
@@
 
-* A series of $j \in \mathcal{J}$ pre-scheduled missions are planned along a horizon divided into $t \in \mathcal{T}$ periods.
-* Each mission requires a certain **number** $R_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $H_j$ **hours** in each period.
-* The heterogeneous fleet requires recurrent preventive maintenance operations (**checks**).
-* A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
-* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the each period of the horizon.
+```{r, echo=FALSE, out.width='100%', warning=FALSE}
+knitr::include_graphics('calendar3.png')
+```
 
---
+A good solution:
 
-**Other**: soft constraints on the state of the fleet at each period, initial conditions, default consumption, reduce variance of frequency of checks.
+* complies with all regulatory rules and mission demands.
+* uses the least possible number of maintenances.
+* keeps the fleet in a "good state" at every period.
 
@@
 * In `r Citep(myBib, "Seif2018")`, a generalization for different types of maintenances and capacities was done.
-* In `r Citep(myBib, "Peschiera2019")`, the problem was proved NP-Complete.
 
@@
 $a_{ijt_1t_2}$: mission $j$ is assigned to aircraft $i$ between $t_1$ and $t_2$.  
-$m_{it_1t_2}$: aircraft $i$ starts a check in $t_1$ and another check in $t_2$.
+$m_{ip}$: aircraft $i$ has a maintenance pattern $p$.
 
 \begin{align}
-  & \text{Max}\;
+  & \text{Min}\;
   \sum_{i \in \mathcal{I}, p \in \mathcal{P}} m_{ip} \times W_p \\
@@
           & j \in \mathcal{J}, t \in \mathcal{TJ}_j  \label{eq:missionres}\\
-  & \sum_{p \in \mathcal{P}_{t}} m_{ip} + \sum_{j \in \mathcal{J}_t \cap \mathcal{J}_i} \sum_{(t_1, t_2) \in \mathcal{T}_{jt}} a_{ijt_1t_2} \leq 1 
-          & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}\\
-  & \sum_{(j, t, t') \in \mathcal{J}\mathcal{T}\mathcal{T}_{ic}} a_{ijtt'} H^\prime_{jtt'} + U^{\prime}_{tc} \leq H^M + M (1 - m_{ip}) & \notag \\
-    & & i \in \mathcal{I}, p \in \mathcal{P}, c \in \mathcal{C}_p \\
 \end{align}
@@
 
-### Forecasting + Optimization
+### Machine Learning + Optimization
 
-**We want to**:
+We:
 
-1. Train a statistical model to predict characteristics of good (optimal?) solutions for any given instance.
-2. Use this information to limit the solution space and make it easier to find those potentially good solutions.
+1. Train a statistical model to predict certain characteristics of good (optimal?) solutions for 5000 simulated instances.
+2. Use this information to *predict* the characteristics of good solutions on a new instance ($\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$).
+3. Apply the prediction to *guide* de solution of said instance.
 
@@
 
-**Benefits**:
-
-1. **Performance**: a smaller problem is easier to solve.
-2. **User feedback**: direct feedback about the solution without needing to solve any model.
-3. **More stable solutions**: Every aircraft flies an amount that is closest to the mean of the fleet.
-
---
-
-**The better we're able to predict those characteristics for the whole fleet, the less optimality we will lose**
+\begin{align}
+    & m_{ip} = 0 & p_{t'} - p_t < \hat{\mu}_{t'-t}^{lb} - tol  \\
+    & m_{ip} = 0  &  p_{t'} - p_t > \hat{\mu}_{t'-t}^{ub} + tol
+\end{align}
 
@@
 
-### Prediction model
+## Results
 
-* **Technique**: *Quantile regressions* to estimate upper and lower bounds.
-* **Training**: 5000 small instances.
-* **Input features**: 
-  * mean flight demand per period,
-  * total remaining flight hours at start (init), 
-  * variance of flight demand, 
-  * demand of special missions, 
-  * number of period where flight demand is cut in two.
-* **Output features**: mean distance between maintenances.
-
---
-
-```{r, echo=FALSE, out.width='100%', warning=FALSE}
-knitr::include_graphics('calendar3.png')
-```
-
----
-
-### How good is it (performance)
-
-Faster solutions, more solutions.
-
---
-
-.center[
-```{r, echo=FALSE, out.width='80%'}
-imgs = c('time_performance_ordered_2tasks.png')
-knitr::include_graphics(imgs)
-```
-]
-
----
-
-### How good is it (optimality)
-
-For instances were an optimal solution was found (optimum degradation):
-* 95% of instances had less than 4% gap with real optimal.
-
-.center[
-```{r, echo=FALSE, out.width='80%'}
-imgs = c('quality_degradation_2tasks.png')
-knitr::include_graphics(imgs)
-```
-]
+* 30% reduction in solution times.
+* 40% more stable solutions.
+* 3-4% optimality loss.
 
@@
 * **Better predictions** with better features, or predicting several characteristics of optimal solutions.
-* **Predict a distribution** and sample patterns from the distribution instead of predicting patterns.
 * **Warm-start Column Generation** with a selected subset of potentially good patterns.
@@
 
-* **Dasault Aviation**: desktop software OPTIMA built, tested, deployed and validated by users.
-
-*Estimated thesis defense: Novembre 2020.*
+* **Dasault Aviation**: desktop software built, tested, deployed and validated by users.
 

### 2020-09-09 R/presentations/EDAA202003/presentation.Rmd
last changes to ppt EDAA.

@@
 
-1. Train a statistical model to predict certain characteristics of good (optimal?) solutions for 5000 simulated instances.
-2. Use this information to *predict* the characteristics of good solutions on a new instance ($\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$).
-3. Apply the prediction to *guide* de solution of said instance.
+1. Train a statistical model to predict certain characteristics of good (optimal?) solutions for 5000 simulated scenario.
+2. Use this information to *predict* the characteristics of good solutions on a new scenario: $\hat{\mu}_{t'-t}^{lb}$,  $\hat{\mu}_{t'-t}^{ub}$.
+3. Apply the prediction to *guide* de solution of said scenario.
 

### 2020-10-07 R/presentations/thesis_defense/presentation.Rmd
thesis defense ppt and dassault documentation.

--- 2020-02-28 R/presentations/roadef202002/presentation.Rmd
@@
 output:
+  beamer_presentation:
+    theme: "Berlin"
+    toc: true
+    slide_level: 2
+    incremental: yes
+    fig_height: 4
+    includes:
+        in_header: header.tex
+  slidy_presentation:
+    css: ./../../../css/baobab_slidy.css
+    incremental: yes
+    footer: "ISAE-SUPAERO, Université de Toulouse, France"
+    toc: true
   xaringan::moon_reader:
@@
     footer: "ISAE-SUPAERO, Université de Toulouse, France"
-  slidy_presentation:
-    css: ./../../../css/baobab_slidy.css
-    incremental: yes
-    footer: "ISAE-SUPAERO, Université de Toulouse, France"
-    toc: true
-  beamer_presentation:
-    theme: "Berlin"
-    toc: true
-    slide_level: 2
-    incremental: yes
-    fig_height: 4
-    includes:
-        in_header: header.tex
 editor_options: 
@@
 
-### Problem (informally)
+### Context
 
-Assign both **missions** and **maintenance operations** to a fleet of aircraft in order to maximize availability and minimize costs.
-**Missions** have fixed start and end times and have particular needs in terms of aircraft and flight hours.
-**Maintenance operations** have specific rules that govern their frequency and capacity.
-
---
-
-.center[
-```{r, echo=FALSE, out.width='100%', warning=FALSE}
-knitr::include_graphics('calendar3.png')
-
-```
-]
+Maintenance, maintenance planning. Flight and maintenance planning. Military. DGA et armée de l'air.
 
@@
 
-* A series of $j \in \mathcal{J}$ pre-scheduled missions are planned along a horizon divided into $t \in \mathcal{T}$ periods.
-* Each mission requires a certain **number** $R_{j}$ of aircraft $i \in \mathcal{I}$ which it employs for $H_j$ **hours** in each period.
-* The heterogeneous fleet requires recurrent preventive maintenance operations (**checks**).
-* A check takes exactly $M^d$ periods and cannot be interrupted. There is a limited **capacity** for maintenances at each period.
-* The objective is to **minimize the total number of checks** while **maximizing the status of the fleet** at the end of the horizon.
-
---
-
-**Other**: soft constraints on the state of the fleet at each period, initial conditions, default consumption, reduce variance of frequency of checks.
+Present the industrial problem for the Mirage 2000.
 
@@
 
-### State of the art
+### Contributions
 
-* FMP: Flight and Maintenance Planning problem.
-* In `r Citep(myBib, "Cho2011")`, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
-* In `r Citep(myBib, "Kozanidis2008")`, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
-* In `r Citep(myBib, "Verhoeff2015")`, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
-* In `r Citep(myBib, "Seif2018")`, a generalization for different types of maintenances and capacities was done.
-* In `r Citep(myBib, "Peschiera2019")`, the problem was proved NP-Complete.
+1. Complexity proof and exact methods.
+2. Pattern-like modeling and machine learning.
+3. Efficient pattern-generation with DP.
 
 ---
-
-### Previous formulation
 

### 2020-10-07 docs/source/heuristic.tex
thesis defense ppt and dassault documentation.

\chapter{Simulated Annealing heuristic}

  The heuristic consists in a Simulated Annealing heuristic detailed in Algorithm \ref{algo:heuristic}. The stop criteria are three: (a) a time limit, (b) getting a solution with 0 errors or (c) an iteration limit. Given the structure of the constraints and the design of the algorithm, the heuristic does not guarantee a feasible solution to the problem.

  The first iteration $c=1$ generates an initial solution by first scheduling random mandatory checks (see Line \ref{algo:maint}) and, then, randomly assigning missions as needed (see Line \ref{algo:miss}). Both functions are explained in detail in Algorithms \ref{algo:AssignChecks} and \ref{algo:AssignMissions}, respectively. At each subsequent iteration $c \geq 2$, the solution is first perturbed by removing or moving maintenances and mission assignments (see Lines \ref{algo:cand} and \ref{algo:rel}), in order to randomly re-schedule checks and then re-assign missions.

  Each candidate selected for releasing in Line \ref{algo:cand} consists on a couple (aircraft, period). These candidates are selected based on the location of errors in the incumbent solution. The perturbations in Line \ref{algo:rel} consist in releasing a slice of the $\mathcal{I} \times \mathcal{T}$ matrix of mission assignments and checks schedules. This slice can be a whole row, i.e., free all assignments and checks for aircraft $i$; a group of columns, i.e., free all assignments and checks for all aircraft between periods $t$ and $t^\prime$; or a combination of the two, i.e., free assignments and checks for subset of aircraft between periods $t$ and $t^\prime$. The type of release is randomly generated.

  After each cycle, the solution is compared with the previous one (see line \ref{algo:accept}) and is accepted depending on the difference in quality, a decreasing temperature and a random factor. The greater the temperature, the greater the probability to accept a new solution that has more errors than the incumbent.

  \begin{algorithm}
    \DontPrintSemicolon
    \SetKwData{bestsol}{$x^*$}\SetKwData{besterr}{$err^*$}
    \SetKwData{newsol}{$x^{\prime}$}\SetKwData{newerr}{$err^{\prime}$}
    \SetKwData{incumbsol}{$x$}\SetKwData{incumberr}{$err$}
    \KwData{\\\incumbsol: incumbent solution, $err$ its errors. \\\newsol: new candidate solution. \\\bestsol: best solution found.\\$T$: temperature.\\$C$: maximum iterations.\\$R$: cooldown rate.}
    \BlankLine
    \Begin{
      $\incumbsol \leftarrow InitializeEmptySolution()$\;
      $\incumberr \leftarrow \besterr \leftarrow GetErrors(\incumbsol)$\;
      $\newsol \leftarrow \incumbsol$\;
      \For{$c \leftarrow 1$ \KwTo $C$}{
        $\newsol \leftarrow AssignChecks(\newsol)$\label{algo:maint}\;
        $\newsol \leftarrow AssignMissions(\newsol)$\label{algo:miss}\;
        $\newerr \leftarrow GetErrors(\newsol)$\;
        \If{$AcceptanceFunc(\incumberr, \newerr, T)$}{\label{algo:accept}
          $\incumbsol, \incumberr \longleftarrow \newsol, \newerr$\;
          \If{$\sum \newerr < \sum \besterr$}{
            $\bestsol, \besterr \longleftarrow \newsol, \newerr$\;
          }
        }
        $T \leftarrow R \times T$\;
        \lIf{$|\besterr| = 0$}{break}
        $C \leftarrow GetCandidatesReassign(\incumberr)$\label{algo:cand}\;
        $\newsol \leftarrow PartialRelease(\incumbsol, C)$\label{algo:rel}\;
      }
    }
    \caption{maintFirst}\label{algo:heuristic}
  \end{algorithm}

  \begin{algorithm}\
    \DontPrintSemicolon
    \SetKwData{sol}{$x$}
    \KwData{\\\sol: the current solution}
    \BlankLine
    \Begin{
      \For{$i \in \mathcal{I}$}{
        $needs \leftarrow GetMaintenanceNeeds(\sol, i)$\;
        $\mathcal{T}^{c} \leftarrow GetMaintenanceCandidates(\sol, i, needs)$\;
        \If{$|\mathcal{T}^{c}| > 0$}{
          $t \leftarrow choice(\mathcal{T}^{c})$\;
          $SetMaintenance(\sol, i, t)$;
        }
      }
    }
    \caption{AssignChecks()}\label{algo:AssignChecks}
  \end{algorithm}

  \begin{algorithm}\
    \DontPrintSemicolon
    \SetKwData{sol}{$x$}
    \KwData{\\\sol: the current solution}
    \BlankLine
    \Begin{
      \For{$j \in shuffle(\mathcal{J})$}{
        $needs \leftarrow CheckMissionNeeds(\sol, j)$\;
        $\mathcal{I}^{c} \leftarrow GetMissionCandidates(j)$\;
        \While{$|\mathcal{I}^{c}| > 0 \land |needs| > 0$}{
          $i \leftarrow choice(\mathcal{I}^{c})$\;
          $\mathcal{T}^{c} \leftarrow GetCandidatePeriods(needs, i)$\;
          \If{$|\mathcal{T}^{c}| = 0$}{
            $\mathcal{I}^{c}.pop(i)$\;
          }
          \For{$t \in shuffle(\mathcal{T}^{c})$}{
            $success \leftarrow SetMissionAssignment(\sol, i, t, j)$\;
            \If{$success$}{
              $needs[t] \leftarrow needs[t] - 1$
            } 
          }
        }
      }
    }
    \caption{AssignMissions()}\label{algo:AssignMissions}
  \end{algorithm}

### 2020-10-07 docs/source/model.tex
thesis defense ppt and dassault documentation.

\chapter{Mission assignment model}

The model assigns discrete missions to aircraft while complying with previously scheduled checks and assigned flight hours to each aircraft.

Missions are already scheduled and consist on a demand for a certain number of aircraft and a number of total flight hours for the whole duration of the mission. Mission assignments cannot be interrupted, i.e., an aircraft is assigned to the whole duration of the mission.

To simplify the notation, overhaul checks (VX) will called M-checks. A M-cycle for an aircraft is defined by the periods between the end of a M-check and the start of the next M-check. Security checks will be called VS-checks. A VS-cycle consists of the periods between the end of an VS-check or M-check and the start of the next VS-check or M-check.

Flight hours from missions need to comply with the M-cycles and VS-cycles, as well as with the previously assigned flight hours. There exist hard limits and soft limits on this compliance.

In order to balance the number of mission assignments among aircraft on the fleet, an increasing cost per unit is applied for each additional mission assigned to the same aircraft.

The maintenance capacity in a given period is reduced proportionally to the number of aircraft that leave for a mission with respect to the size of the fleet. This reduced capacity should not exceed the demand for maintenance from the aircraft that stay on base (i.e., that do not go in mission).

\section{Input data}

The main sets.

  \begin{tabular}{p{15mm}p{120mm}}
    $i \in \mathcal{I}$     &  aircraft.    \\
    $t \in \mathcal{T}$     &  time periods included in the planning horizon. We use $t=0$ for starting conditions and $t=T$ for the last period. \\
    $j \in \mathcal{J}$     &  missions. \\
    $s \in \mathcal{S}$     &  intervals por deviations.
  \end{tabular}

Auxiliary sets.

  \begin{tabular}{p{15mm}p{120mm}}
    $t \in \mathcal{T}_j$        &  time periods $t \in \mathcal{T}$ when mission $j$ is active. \\
    $j \in \mathcal{J}_i$        &  missions $j \in \mathcal{J}$ where aircraft $i$ is suitable. \\
    $j \in \mathcal{J}_t$        &  missions $j \in \mathcal{J}$ that are active in period $t$. \\
    $(t, t') \in \mathcal{L}^{M}_{i}$         & $(t \in \mathcal{T}, t' \in \mathcal{T})$ such that aircraft $i$ starts a M-cycle at the beginning of period $t$ and finishes it at the end of period $t'$. \\
    $j \in \mathcal{J}_{itt'}$         & missions $j \in \mathcal{J}_i$ that start after the beginning of period $t$ and finish before the end of period $t'$.\\
    $(t, t') \in \mathcal{L}^{VS}_{i}$         & $(t \in \mathcal{T}, t' \in \mathcal{T})$ such that aircraft $i$ starts a VS-cycle at the beginning of period $t$ and finishes it at the end of period $t'$.\\
  \end{tabular}

  Parameters.

  \begin{tabular}{p{15mm}p{100mm}p{20mm}}
    $H_j$             & total flight hours required per aircraft for mission $j$. & [hours] \\
    $R_j$             & number of aircraft required per period for mission $j$. & [aircraft] \\
    $C^{max}$         & maximum number of simultaneous aircraft checks. & [aircraft] \\
    $H^{M}$         & maximum number of flight hours between two consecutive M-checks for each aircraft. & [hours] \\
    $H^{VS}$         & maximum number of flight hours between two consecutive VS-checks for each aircraft. & [hours] \\
    $H^{rem}_{itt'}$ & remaining flight hours for aircraft $i$ at the end of the M-cycle that starts in $t$ and ends in $t'$.& [hours] \\
    $C^{rem}_{t}$ & net remaining capacity during period $t$ after discounting aircraft that go in mission.& [hours] \\
    $C^{u}_{it}$ & maintenance capacity used by aircraft $i$ in period $t$. & [days] \\
  \end{tabular}

\section{Variables}

    The following binary decision variables prescribe the assignment of missions and checks to aircraft.

    \begin{tabular}{p{8mm}p{127mm}}
      $a_{ij}$ & has value one if aircraft $i$ is assigned to mission $j$, zero otherwise.
    \end{tabular}

    The following continuous auxiliary variables help balance the different objectives.

    \begin{tabular}{p{8mm}p{127mm}}
      $e^{H}_{itt'}$  & Deviation of cycle flight hours for aircraft $i$ during M-cycle $(t, t') \in \mathcal{L}_i$. \\
      $e^{J}_{is}$ & Deviation in number of missions assigned to aircraft $i$ in interval $s$. \\
      $e^{C}_{t}$ & Deviation in maintenance capacity at end of $t$. \\
    \end{tabular}

\section{Objective function and constraints}

  The objective function (\ref{eq:objective1}) expresses the total deviation from all goals on flight hours per cycle, fleet balance and maintenance capacity.

  \begin{align}
    & \text{Min}\; 
    \sum_{\substack{
            i \in \mathcal{I}, \\ (t, t') \in \mathcal{L}_i
            }
        } e^{H}_{itt'}
    + 100 \sum_{\substack{
            i \in \mathcal{I}, \\ s \in \mathcal{S}
            } 
        } PJ_s \times e^{J}_{is}
    + 1000 \sum_{t \in \mathcal{T}} 
    e^{C}_{t} 
    \label{eq:objective1}
  \end{align}

  \begin{align}
    % only one active mission assignment per aircraft.
    & \sum_{\substack{j \in \\ \mathcal{J}_t \cap \mathcal{J}_i}} 
    a_{ij} \leq 1 
            & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state} \\ 
    % number of resources per mission
    & \sum_{i \in \mathcal{I}_j} a_{ij} = R_j
            & j \in \mathcal{J}  \label{eq:missionres}\\
    % number of missions per resource
    & \sum_{i \in \mathcal{J}_i} a_{ij} = \sum_{s \in \mathcal{S}} e^{J}_{is}
            & i \in \mathcal{I}  \label{eq:num_missions}\\
    % max number of flight hours. hard limit. (M)
    & \sum_{j \in \mathcal{J}_{itt'}} a_{ij} H_j \leq H^{M}
            & i \in \mathcal{I}, (t, t') \in \mathcal{L}^M_{i} \label{eq:maxHoursH}\\
    % max number of flight hours. hard limit. (VS)
    & \sum_{j \in \mathcal{J}_{itt'}} a_{ij} H_j \leq H^{VS}
            & i \in \mathcal{I}, (t, t') \in \mathcal{L}^{VS}_{i} \label{eq:maxHoursVS}\\
    % max number of flight hours. soft limit
    % 
    & e^{H}_{itt'} \geq \sum_{j \in \mathcal{J}_{itt'}} a_{ij} H_j - H^{rem}_{itt'} 
            & i \in \mathcal{I}, (t, t') \in \mathcal{L}^M_{i} \label{eq:extraHours}\\
    % finally, we create the constraint guaranteeing capacity.
    &\sum_{i \in \mathcal{I}} [C^{u}_{it} \times (1 - \sum_{\substack{j \in \\ \mathcal{J}_t \cap \mathcal{J}_i}} a_{ij})] 
    - e^{C}_{t} \leq C^{rem}_{t}
            & t \in \mathcal{T} \label{eq:capacity1}
  \end{align}

  Constraints (\ref{eq:state}) limit the number of simultaneous missions that an aircraft can be assigned to 1. Constraints (\ref{eq:missionres}) guarantee that all missions are covered with the required aircraft. Constraints (\ref{eq:num_missions}) tie the total number of missions for each aircraft to penalize it in the objective function.

  Constraints (\ref{eq:maxHoursH}, \ref{eq:maxHoursVS}) set the absolute maximum number of flight hours in M-cycles and VS-cycles, respectively.

  Constraints (\ref{eq:extraHours}) limit the number of additional hours each aircraft can fly during each M-cycle. This constraint is elastic.

  Constraints (\ref{eq:capacity1}) guarantee that the remaining capacity is enough for the scheduled checks of the aircraft that are not currently assigned to a mission.

  Constraints (\ref{eq:capacity1}) limit the number of unpenalized simultaneous checks. Constraints (\ref{eq:missionres}) enforce aircraft mission requirements. Constraints (\ref{eq:state}) restrict each aircraft to at most one assignment each period.

### 2020-10-07 docs/source/conf.py

@@ -71,3 +71,15 @@ intersphinx_mapping = {'python': ('http://docs.python.org/3', None),
 
                        
                        
+latex_elements = {
+    'preamble': r'''
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
+\usepackage[titles]{tocloft}
+\cftsetpnumwidth {1.25cm}\cftsetrmarg{1.5cm}
+\setlength{\cftchapnumwidth}{0.75cm}
+\setlength{\cftsecindent}{\cftchapnumwidth}
+\setlength{\cftsecnumwidth}{1.25cm}
+''',
+    'fncychap': r'\usepackage[Bjornstrup]{fncychap}',
+    'printindex': r'\footnotesize\raggedright\printindex',
+}

### 2020-10-07 python/solvers/model_dassault.py

@@ -1,4 +1,3 @@
-import math
 import pulp as pl
 import solvers.config as conf
 import package.experiment as exp
@@ -17,6 +16,11 @@ class ModelMissions(exp.Experiment):
         self.model = None
 
     def solve(self, options):
+        """
+        Solves a mip model given a FMP with fixed maintenances
+        :param options:
+        :return:
+        """
 
         resources = self.instance.get_resources()
         candidates = self.instance.get_task_candidates()
@@ -75,7 +79,7 @@ class ModelMissions(exp.Experiment):
             to_dict(None).\
             vapply(lambda v: get_rut(v[0], prev(v[-1])))
 
-
+        # VARIABLES
         self.task = \
             pl.LpVariable.dicts(name="task",
                                 indexs=res_task,
@@ -113,7 +117,7 @@ class ModelMissions(exp.Experiment):
         res_nb_mission = \
             task_period_rest.\
             vfilter(lambda v: len(v) > 1).\
-            apply(lambda k, v: pl.lpSum(self.task[k[1], _v] for _v in v)).\
+            kvapply(lambda k, v: pl.lpSum(self.task[k[1], _v] for _v in v)).\
                 to_tuplist().take([1, 0, 2]).to_dict(2, is_list=False)
         for v in res_nb_mission.values():
             self.model += v <= 1
@@ -126,9 +130,9 @@ class ModelMissions(exp.Experiment):
             self.model += v == task_num_resource[k]
 
         # number of missions per resource
-        task_per_task = res_task.to_dict(1).apply(lambda k, v: pl.lpSum(self.task[k, _v] for _v in v))
+        task_per_resource = res_task.to_dict(1).apply(lambda k, v: pl.lpSum(self.task[k, _v] for _v in v))
         excess = self.num_missions_range.to_tuplist().take([0, 2]).to_dict(1).vapply(pl.lpSum)
-        for r, v in task_per_task.items():
+        for r, v in task_per_resource.items():
             self.model += v == excess[r]
 
         # the sum of flight hours that are assigned need to fall in
@@ -208,6 +212,10 @@ class ModelMissions(exp.Experiment):
         return self.solution
 
     def get_solution(self):
+        """
+        Reads the MIP variables contents after solving and generates a solution
+        :return: formatted solution
+        """
 
         _range_f = self.instance.get_periods_range
         assignments = self.task.vfilter(pl.value).keys_tl()
@@ -231,6 +239,14 @@ class ModelMissions(exp.Experiment):
         return sol.Solution(sol_data)
 
[…]

### 2020-10-26 docs/source/model.tex

@@
 
-  \begin{tabular}{p{15mm}p{120mm}}
+  \begin{tabular}{p{20mm}p{120mm}}
     $i \in \mathcal{I}$     &  aircraft.    \\
@@
 
-  \begin{tabular}{p{15mm}p{120mm}}
+  \begin{tabular}{p{20mm}p{120mm}}
     $t \in \mathcal{T}_j$        &  time periods $t \in \mathcal{T}$ when mission $j$ is active. \\
@@
 
-  \begin{tabular}{p{15mm}p{100mm}p{20mm}}
+  \begin{tabular}{p{20mm}p{100mm}p{20mm}}
     $H_j$             & total flight hours required per aircraft for mission $j$. & [hours] \\
